---
title: "Insulin Bone Marker Manuscript"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(knitr)
library(tidyverse)
library(reshape2)
library(gridExtra)
source("/Users/timvigers/Documents/GitHub/Tim-and-Laura/tim_R_functions.R")
```

```{r echo=FALSE,warning=FALSE}
# Read in dataset (approved by Viral and Vanessa)
full_data <- read.csv("/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/Insulin bone marker/Manuscript/Data_Cleaned/full_data.csv")
# Calculate albumin-corrected calcium
full_data$CA_Corrected.0 <- 0.8*(4.5 - full_data$ALBS.0) + full_data$CA.0
# Remove impossible data per Vanessa
full_data$OPN.0[which(full_data$PtID == "K039-0004")] <- NA
full_data$P1NP.0[which(full_data$PtID == "K059-0015")] <- NA
full_data$P1NP.120[which(full_data$PtID == "K075-0011")] <- NA
```

```{r echo=FALSE,include=FALSE}
# Table 1
vars = c("AgeAtVisit0","Gender","Ethnicity","Race","T1DDuration","BMI",
         "UnitsInsPerKg","HbA1c","InsDeliveryMethod","CurrUseCGM","Vitamin.D.0",
         "CA.0","ALBS.0","CA_Corrected.0","eGFR.0","CREA.S.0","ALP.0")
t1a <- CreateTableOne(vars,data = full_data)
t1a <- print(t1a, exact = c("Ethnicity","Race","CurrUseCGM"),
            printToggle = F)
t1b <- CreateTableOne(vars,strata = "group",data = full_data)
t1b <- print(t1b,nonnormal = c("T1DDuration"), exact = c("Ethnicity","Race","CurrUseCGM"),
            printToggle = F)
t1c <- CreateTableOne(vars = vars[-c(which(vars == "Gender"))],strata = "Gender",data = full_data)
t1c <- print(t1c, exact = c("Ethnicity","Race"),
            printToggle = F)
```

## Table 1a: Descriptive Statistics for Full Cohort

```{r echo=FALSE}
kable(t1a)
```

## Table 1b: Descriptive Statistics by C-Peptide Group

```{r echo=FALSE}
kable(t1b[,1:5])
```

## Table 1c: Descriptive Statistics by Gender

```{r echo=FALSE}
kable(t1c[,1:3])
```

## Figure 1: Change in BTM and related hormones

```{r echo=FALSE}
baseline_btm <- c("P1NP.0","CTX1.0","Osteocalcin.0","ON.0","OPG.0","IGF.0")
min_120_btm <- c("P1NP.120","CTX1.120","Osteocalcin.120","ON.120","OPG.120","IGF.120")
# Calculate percent change for BTMs and insulin
full_data$perc_change_P1NP <- ((full_data$P1NP.120 - full_data$P1NP.0)/full_data$P1NP.0)*100
full_data$perc_change_CTX1 <- ((full_data$CTX1.120 - full_data$CTX1.0)/full_data$CTX1.0)*100
full_data$perc_change_Osteocalcin <- ((full_data$Osteocalcin.120 - full_data$Osteocalcin.0)/full_data$Osteocalcin.0)*100
full_data$perc_change_ON <- ((full_data$ON.120 - full_data$ON.0)/full_data$ON.0)*100
full_data$perc_change_OPG <- ((full_data$OPG.120 - full_data$OPG.0)/full_data$OPG.0)*100
full_data$perc_change_IGF <- ((full_data$IGF.120 - full_data$IGF.0)/full_data$IGF.0)*100
full_data$perc_change_Insulin <- ((full_data$Insulin.120 - full_data$Insulin.0)/full_data$Insulin.0)*100
# Mke plotting dataframe
perc_change_data <- full_data %>%
  select(PtID,group,Gender,perc_change_P1NP:perc_change_Insulin)
perc_change_data <- melt(perc_change_data,id.vars = c("PtID","group","Gender","perc_change_Insulin"))
perc_change_data$variable <- gsub("perc_change_","",perc_change_data$variable)
# Full cohort plot
perc_change_plot_all <- ggplot(perc_change_data,aes(x=value,y=perc_change_Insulin)) + 
  geom_point()+
  facet_wrap(~variable,nrow = 1,scales = "free_x") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank(),
        axis.title.x = element_blank())
# By gender
perc_change_plot_gender <- ggplot(perc_change_data,aes(x=value,y=perc_change_Insulin)) + 
  geom_point(aes(shape=Gender))+
  facet_wrap(~variable,nrow = 1,scales = "free_x") +
  ylab("% Change in Insulin") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90),
        axis.title.x = element_blank())
# By group
perc_change_plot_group <- ggplot(perc_change_data,aes(x=value,y=perc_change_Insulin)) + 
  geom_point(aes(shape=group))+
  facet_wrap(~variable,nrow = 1,scales = "free_x") +
  xlab("% Change in BTM") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank()) +
  labs(shape = "C-Peptide Group")
```

```{r echo=FALSE,warning=FALSE,fig.width=8,fig.height=10,dpi=600}
grid.arrange(perc_change_plot_all,perc_change_plot_gender,perc_change_plot_group, nrow = 3)
```

## Table 2: Comparison of Baseline BTMs by C-Peptide Group

```{r echo=FALSE}
t2 <- CreateTableOne(vars = baseline_btm,strata="group",data=full_data)
nonnormal <- c("ON.0","IGF.0")
t2 <- print(t2,nonnormal = nonnormal,printToggle = F)
kable(t2[,1:5])
```

## Table 3: Comparison of % Change in BTMs by C-Peptide Group

```{r echo=FALSE}
percvars <- c("perc_change_P1NP","perc_change_CTX1","perc_change_Osteocalcin",
              "perc_change_ON","perc_change_OPG","perc_change_IGF")
t3 <- CreateTableOne(vars = percvars,strata="group",data=full_data)
t3 <- print(t3,printToggle = F)
kable(t3[,1:5])
```

## Table 4: Comparison of Baseline BTMs by Gender

```{r echo=FALSE}
# Check residuals
# for (v in baseline_btm) {
#   formula <- as.formula(paste(v,"~Gender"))
#   plot(lm(formula,data = full_data),main = v)
# }
t4 <- CreateTableOne(vars = baseline_btm,strata="Gender",data=full_data)
nonnormal <- c("ON.0","IGF.0")
t4 <- print(t4,nonnormal = nonnormal,printToggle = F)
kable(t4[,1:3])
```

## Table 5: Comparison of % Change in BTMs by Gender

```{r echo=FALSE}
# Check residuals
# for (v in percvars) {
#   formula <- as.formula(paste(v,"~Gender"))
#   plot(lm(formula,data = full_data),main = v)
# }
t5 <- CreateTableOne(vars = percvars,strata="Gender",data=full_data)
t5 <- print(t5,printToggle = F)
kable(t5[,1:3])
```

## Table 6: Percent change in BTM different from 0, full cohort

```{r echo=FALSE}
t6 <- CreateTableOne(vars = percvars,data=full_data,testNormal = t.test)
t6 <- print(t6,printToggle = F)
p <- c("")
for (v in 1:length(percvars)) {
  var <- percvars[v]
  test <- t.test(full_data[,var])
  p[v+1] <- format.pval(test$p.value,digits = 3,eps = 0.001)
}
t6 <- cbind(t6,p)
kable(t6)
```

<!-- ## The effect of percent change in bone markers on percent change in insulin for the full cohort -->

<!-- ### P1NP -->

<!-- ```{r echo=FALSE,results='asis'} -->
<!-- p1np_mod <- lm(perc_change_Insulin ~ perc_change_P1NP,full_data) # Residuals are not great. Looked at without row 52 (high P1NP % change) and log transformed, still not significant. -->
<!-- kable(round(summary(p1np_mod)$coefficients,3)) -->
<!-- cat(paste("Adjusted R Squared:",round(summary(p1np_mod)$adj.r.squared,3))) -->
<!-- ``` -->

<!-- ### CTX1 -->

<!-- ```{r echo=FALSE,results='asis'} -->
<!-- ctx1_mod <- lm(perc_change_Insulin ~ perc_change_CTX1,full_data) -->
<!-- kable(round(summary(ctx1_mod)$coefficients,3)) -->
<!-- cat(paste("Adjusted R Squared:",round(summary(ctx1_mod)$adj.r.squared,3))) -->
<!-- ``` -->

<!-- ### Osteocalcin -->

<!-- ```{r echo=FALSE,results='asis'} -->
<!-- osteocalcin_mod <- lm(perc_change_Insulin ~ perc_change_Osteocalcin,full_data) -->
<!-- kable(round(summary(osteocalcin_mod)$coefficients,3)) -->
<!-- cat(paste("Adjusted R Squared:",round(summary(osteocalcin_mod)$adj.r.squared,3))) -->
<!-- ``` -->

<!-- ### ON -->

<!-- ```{r echo=FALSE,results='asis'} -->
<!-- on_mod <- lm(perc_change_Insulin ~ perc_change_ON,full_data) -->
<!-- kable(round(summary(on_mod)$coefficients,3)) -->
<!-- cat(paste("Adjusted R Squared:",round(summary(on_mod)$adj.r.squared,3))) -->
<!-- ``` -->

<!-- ### IGF -->

<!-- ```{r echo=FALSE,results='asis'} -->
<!-- igf_mod <- lm(perc_change_Insulin ~ perc_change_IGF,full_data) -->
<!-- kable(round(summary(igf_mod)$coefficients,3)) -->
<!-- cat(paste("Adjusted R Squared:",round(summary(igf_mod)$adj.r.squared,3))) -->
<!-- ``` -->