---
title: "AHA Untargeted Metabolomics"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Metabolomics")
library(arsenal)
library(skimr)
library(knitr)
library(tidyverse)
library(limma)
library(mixOmics)
```

```{r data cleaning,echo=FALSE}
# Read in raw data
raw_data <- read.csv("./Data_Cleaned/metabolomics_without_imputation.csv",
                     stringsAsFactors = F,na.strings = c("","NA"))
# Format
data <- raw_data %>% 
  mutate(compound = coalesce(HMDB.No.,LI.Library.No.),
         m.z = Neutral.Mass..Da.+1.0078,
         rt = RT..s.,
         masserror = coalesce(Mass.Error..ppm.,Mass.Error..ppm..1),
         rterror = coalesce(RT.Error..s.,RT.Error..s..1)) %>%
  select(compound,m.z,rt,masserror,rterror,
         IsoMS_results_QC01_190618171010_converted.csv:
           IsoMS_results_S227_Batch07_1400_RA8_converted.csv)
data$compound <- paste(data$compound,data$m.z,data$rt,sep = "_")
# Rename columns
colnames(data) <- gsub("IsoMS_results_","",colnames(data))
colnames(data) <- gsub("*_.*","",colnames(data))
# Import clinical data and sample list
clinical <- read.csv("./Data_Cleaned/targeted.csv",stringsAsFactors = F)
# Three progression groups
cac_groups <- read.table("./Data_Raw/CAC Trajectories 3 groups.txt",sep = "\t",
                         header = T)
# CAC progression - increase in square root calcium volume >= 2.5 between baseline and visit 3
clinical$CACp <- 
  as.factor(ifelse(clinical$c3 - clinical$c1 >= 2.5,"Progression","No Progression"))
# New CAC groups
clinical <- left_join(clinical,cac_groups[,c("StudyID","GROUP")],by = "StudyID")
samples <- read.csv("./Data_Cleaned/sample_list.csv")
# Match sample number to subject ID
clinical$GlobalSampleID <- 
  as.character(samples$Injection[match(clinical$StudyID,samples$SampleID)])
# Relevant columns
clinical <- clinical %>%
  select(GlobalSampleID,age,sex,race,CACp,GROUP) %>%
  filter(!is.na(GlobalSampleID))
# Names
names <- data$compound
# Data matrix
data <- data %>% select(QC01:S227) %>%
  t(.) %>% as.data.frame(.) # Transpose - samples in rows
# Format columns
colnames(data) <- names
data$GlobalSampleID <- as.character(rownames(data))
# Add clinical data
data <- left_join(data,clinical,by = "GlobalSampleID")
data <- data %>% select(GlobalSampleID:GROUP,everything())
data$GROUP[grepl("QC",data$GlobalSampleID)] <- "QC"
data$GROUP <- as.factor(data$GROUP)
# QC batch info
# Injection order
sampledata <- read.csv("./Data_Cleaned/injection_seq.csv",stringsAsFactors = F)
sampledata$batch <- sapply(strsplit(sampledata$SampleName,"_"),`[`,2)
sampledata$batch[nrow(sampledata)] <- sampledata$batch[(nrow(sampledata)-1)]
sampledata$batch <- sub("Batch0","",sampledata$batch)
sampledata <- sampledata %>% fill(batch,.direction = "up") %>%
  mutate(order = 1:n())
# Merge
data <- left_join(data,sampledata[,c("GlobalSampleID","batch")],
                  by = "GlobalSampleID")
data$batch <- as.factor(data$batch)
# Order columns
data <- data %>% select(GlobalSampleID:GROUP,batch,everything())
# Column formats
data[,c("GlobalSampleID","sex","race")] <- 
  lapply(data[,c("GlobalSampleID","sex","race")],as.factor) 
```

# Pre-processing

```{r echo=FALSE}
# Check for samples missing > 80% of compounds, remove them
missing <- which(rowSums(is.na(data[,8:ncol(data)])) / length(8:ncol(data)) > 0.8)
data <- data[-missing,]
# Same for 0 instead of NA - none
missing0 <- which(rowSums(data[,8:ncol(data)] == 0) / length(8:ncol(data)) > 0.8)
# Check for and remove compounds with > 20% missing in research - none
research <- which(colSums(is.na(data[!grepl("QC",data$GlobalSampleID),])) / 
                    (nrow(data))-13 > 0.2)
# or > 80% missing in QC - none
qc <- names(which(colSums(is.na(data[grepl("QC",data$GlobalSampleID),8:ncol(data)])) / 
                    13 > 0.2))
data <- data[,!(names(data) %in% qc)]
# Check for compounds with CV > 0.3 in QC samples
qc_cv <- names(which(lapply(data[grepl("QC",data$GlobalSampleID),8:ncol(data)],function(x){
  sd(x,na.rm = T)/mean(x,na.rm = T)
}) > 0.3))
data <- data[,!(names(data) %in% qc_cv)]
# Remove bad name compound
data[,grep("314_",colnames(data))] <- NULL
```

Removed `r length(missing)` samples missing > 80% of compounds and `r length(missing0)` samples with > 80% of compounds equal to 0. Also removed `r length(research)` compounds with > 20% missing in research samples and `r length(qc)` compounds with > 80% missing in QC samples. Finally, removed `r length(qc_cv)` compounds with CV > 0.3 in QC samples. 

# T tests

```{r echo=FALSE,cache=TRUE}
# Find the effective number of variables for multiple testing adjustment.
# Correlation matrix between all metabolites
mets <- data[,8:ncol(data)]
cor_mat <- cor(mets,use = "complete.obs")
# Get eigenvalues for correlation matrix
values <- eigen(cor_mat)$values
# Calculate effective M
M <- dim(cor_mat)[1]
meff <- 1 + (M - 1)*(1 - (var(values)/M))
# Try Randi's method of the number of PCs needed to explain 95% of variance 
pca <- pca(mets,center = T,scale = T,ncomp = 100)
```

```{r echo=FALSE}
# Non-missing outcome
X <- data[,c(5,6,8:ncol(data))]
X <- X[!is.na(X[,1]),]
X$GROUP <- as.factor(as.character(X$GROUP))
# Log transform
# X[,3:ncol(X)] <- lapply(X[,3:ncol(X)],function(x){log(x,base = 2)})
ts <- lapply(colnames(X)[3:ncol(X)], 
             function(x){
               form <- as.formula(paste0(x,"~CACp"))
               t <- t.test(form,X)
               return(list("metabolite"=x,"p"=t$p.value))
             })
ts <- data.frame(do.call(rbind,ts))
ts$p <- as.numeric(ts$p)
ts <- ts[order(ts$p),]
# Sidak correction
alpha <- 1 - (1 - 0.05)^(1/meff)
# Randi's suggestion
alpha <- 0.05 / 76
# Just using the number of PCs needed to 
ts <- ts[which(ts$p <= alpha),]
```

# ANOVA

## Permutation Tests

Density represents the number of significant metabolites and the red line indicates the number of hits in non-permuted data. P values were not adjusted for multiple comparisons.

```{r echo=FALSE}
# Test all
anova_perms <- lapply(1:nsim,function(y){
  df <- X
  if(y > 1) {
    set.seed(y)
    df$GROUP <- sample(df$GROUP)
  }
  anovas <- lapply(colnames(df)[3:ncol(df)], function(x){
    form <- as.formula(paste0(x,"~GROUP"))
    mod <- lm(form,df)
    f <- anova(mod)
    return(list("metabolite"=x,"p"=f$`Pr(>F)`[1]))
  })
  anovas <- data.frame(do.call(rbind,anovas))
  anovas$p <- as.numeric(anovas$p)
  anovas <- anovas[order(anovas$p),]
  anovas$p.adjusted <- p.adjust(anovas$p,method = "fdr")
  return(length(which(anovas$p < 0.05)))
})
anova_perms <- as.data.frame(unlist(anova_perms))
anova_perms$correct <- anova_perms[1,1]
colnames(anova_perms) <- c("p","correct")
ggplot(anova_perms,aes(x=p)) + theme_bw() + 
  geom_density() +
  geom_vline(mapping = aes(xintercept=correct),color = "red")
```

# Moderated t tests

```{r echo=FALSE,cache=TRUE}
# Design matrix - 0 and 1 for CAC progression
design <- as.data.frame(as.numeric(data$CACp)-1)
colnames(design) <- "CACp"
design <- model.matrix(~CACp,design)
# "Expression" matrix (samples in columns) for each
X <- data[match(rownames(design),rownames(X)),8:ncol(data)]
# Linear model
fit <- lmFit(t(X),design)
fit <- eBayes(fit)
# Results for ChemRICH
results <- topTable(fit,coef = 2,number = nrow(X))
results$hmdb <- sapply(strsplit(rownames(results),"_"),"[[",1)
results$p.value <- results$P.Value
results$t.score <- fit$t[match(rownames(results),rownames(fit$t)),2]
results$Compound <- raw_data$Compound[match(results$hmdb,raw_data$HMDB.No.)]
results$foldchange <- 2^results$logFC
# Get inchikey, pubchem, and smiles
i <- cts_convert(results$hmdb,"Human Metabolome Database","InChIKey")
p <- cts_convert(results$hmdb,"Human Metabolome Database","PubChem CID")
s <- cts_convert(results$hmdb,"Human Metabolome Database","SMID")

# write.csv(results[,c("m.z","p.value","t.score")],row.names = F,
#           file = "./Data_Cleaned/moderated_t_tests_CACp.csv")
```

<!-- # Moderated ANOVA -->

<!-- ```{r echo=FALSE} -->
<!-- # Design matrix - 0 and 1 for CAC progression -->
<!-- design <- as.data.frame(data$GROUP[data$GROUP != "QC"]) -->
<!-- colnames(design) <- "GROUP" -->
<!-- design$GROUP <- as.factor(as.numeric(design$GROUP)) -->
<!-- design <- model.matrix(~GROUP,design) -->
<!-- # "Expression" matrix (samples in columns) for each  -->
<!-- X <- data[match(rownames(design),rownames(X)),8:ncol(data)] -->
<!-- # Linear model -->
<!-- fit <- lmFit(t(X),design) -->
<!-- fit <- eBayes(fit) -->
<!-- # Results -->
<!-- results <- topTable(fit,coef = c(2,3),number = nrow(X)) -->
<!-- results$m.z <- sapply(strsplit(rownames(results),"_"),"[[",2) -->
<!-- results$p.value <- results$P.Value -->
<!-- results$t.score <- fit$t[match(rownames(results),rownames(fit$t)),2] -->
<!-- write.csv(results[,c("m.z","p.value","t.score")],row.names = F, -->
<!--           file = "./Data_Cleaned/moderated_anova_GROUP.csv") -->
<!-- ``` -->

<!-- # Metabolite classes     -->

<!-- ```{r echo=FALSE} -->
<!-- # Import identified compounds. These were found by uploading moderated t test  -->
<!-- # results to the peaks to pathways module in MetaboAnalyst -->
<!-- kegg <- read.csv("./Data_Cleaned/mummichog_matched_compound_all.csv") -->
<!-- # Add to data  -->
<!-- raw_data$Query.Mass <-raw_data$Neutral.Mass..Da.+1.0078 -->
<!-- classes <- left_join(raw_data,kegg,by = "Query.Mass") -->
<!-- # Remove unidentified metabolites -->
<!-- classes <- classes[!is.na(classes$Matched.Compound),] -->
<!-- # Get compound classes -->
<!-- colnames(classes)[which(colnames(classes) == "Matched.Compound")] <-  -->
<!--   "KEGG" -->
<!-- classes <- assign_hierarchy(classes,keep_unknowns = F,identifier = "KEGG") -->
<!-- # Order by class name -->
<!-- classes <- classes[order(classes$Class),] -->
<!-- ``` -->

<!-- # Correlation matrix (clustered by metabolite class) -->

<!-- ```{r echo=FALSE} -->
<!-- # Transpose for correlation matrix -->
<!-- cor_mat <- t(classes[!is.na(classes$Class),24:263]) -->
<!-- colnames(cor_mat) <- classes$Class[!is.na(classes$Class)] -->
<!-- # Clustered heatmap -->
<!-- # Correlation matrix -->
<!-- cor = round(cor(cor_mat,use = "pairwise.complete.obs"),3) -->
<!-- heatmap(cor,Rowv = NA,Colv = NA,main = "No Clustering") -->
<!-- heatmap(cor,main = "Default Clustering") -->
<!-- ``` -->

<!-- # WGCNA -->

<!-- - Samples with > 50% missing are excluded -->

<!-- ```{r echo=FALSE} -->
<!-- # WGCNA dataframe - samples in rows -->
<!-- W <- as.data.frame(data[,8:ncol(data)]) -->
<!-- rownames(W) <- data$GlobalSampleID -->
<!-- # Exclude QC -->
<!-- W <- W[14:nrow(W),] -->
<!-- # Good samples and metabolites only -->
<!-- gsm <- goodSamplesGenes(W) -->
<!-- W <- W[,gsm$goodGenes] -->
<!-- W <- W[gsm$goodSamples,] -->
<!-- # Make sample tree -->
<!-- sampleTree <- hclust(dist(W), method = "average") -->
<!-- plot(sampleTree) -->
<!-- ``` -->