---
title: "TrialNet ECL/HLA analysis for ADA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)


knitr::opts_chunk$set(echo = TRUE)

library(tableone)
library(sjlabelled)
library(ggplot2)
library(dplyr)

# read in data
data <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\Andrea Steck\\ECL ab and genotype\\Data\\191015 final dataset 1287 subjects with genotype.csv",
                 na.strings = c("NA",""," "))
keep <- data[,c("PTID","ECL.IAA.visit.1","ECL.GADA.visit.1","HLA.DR3.4","Genotype","DR.3.present","DR4.Present","Age.at.visit.1",
                "Gender","Number.of.positive.biochemical.antibodies.visit.1")]
keep$numpos <- as.factor(keep$Number.of.positive.biochemical.antibodies.visit.1)

keep$biochem_status[keep$Number.of.positive.biochemical.antibodies.visit.1==0] <- 'Ab negative'
keep$biochem_status[keep$Number.of.positive.biochemical.antibodies.visit.1==1] <- 'Single ab positive'
keep$biochem_status[keep$Number.of.positive.biochemical.antibodies.visit.1>1] <- 'Multiple ab positive'
set_label(keep$biochem_status)="Biochemical ab status"

# remove people who do not have HLA data
keep <- keep[keep$HLA.DR3.4!='Missing',]
keep$HLA.DR3.4 <- droplevels(as.factor(keep$HLA.DR3.4))

# remove participant 136806 because mIAA Z-score is >20, not plausible
keep <- keep[keep$PTID != "136806",]

# calculate Z-score for ECL variables
# IAA mean 0.008839 SD 0.04763116
# GADA 0.8606 SD 2.081496
keep$ecl_iaa_z <- (keep$ECL.IAA.visit.1 - 0.008839) / 0.04763116
keep$ecl_gada_z <- (keep$ECL.GADA.visit.1 - 0.8606) / 2.081496

# categorize ECL variables
keep$ecl_iaa_pos <- as.factor(ifelse(keep$ECL.IAA.visit.1>0.006,"Positive","Negative"))
keep$ecl_gada_pos <- as.factor(ifelse(keep$ECL.GADA.visit.1>0.023,"Positive","Negative"))
keep$ecl_either_pos <- ifelse(keep$ecl_iaa_pos=="Positive" | keep$ecl_gada_pos=="Positive", "Positive", "Negative")
set_label(keep$ecl_either_pos)="ECL ab positive"
keep$ecl_cat <- ifelse(keep$ecl_iaa_pos=="Positive" & keep$ecl_gada_pos=="Positive", "ECL IAA and GADA +", 
                       ifelse(keep$ecl_iaa_pos=="Positive" & keep$ecl_gada_pos=="Negative","ECL IAA +",
                              ifelse(keep$ecl_iaa_pos=="Negative" & keep$ecl_gada_pos=="Positive","ECL GADA +","ECL -")))


# number of subjects is the number with non-missing ECL values
n <- nrow(keep[!is.na(keep$ECL.IAA.visit.1) | !is.na(keep$ECL.GADA.visit.1),])

# make dataframe without DRx/x
keep_nox <- keep[keep$Genotype != "DR x/x",]
keep_nox$Genotype <- droplevels(as.factor(keep_nox$Genotype))

# Table 1
tab1vars <- c("Age.at.visit.1","Gender","HLA.DR3.4","Genotype","DR.3.present","DR4.Present","biochem_status")
tab1 <- CreateTableOne(vars=tab1vars, data=keep[keep$HLA.DR3.4!="Missing",])
tab1 <- print(tab1,varLabels=TRUE,showAllLevels=TRUE)
tab1a <- CreateTableOne(vars=tab1vars, data=keep, strata="ecl_either_pos", test=TRUE)
tab1a <- print(tab1a,varLabels=TRUE,showAllLevels=TRUE)

minage <- min(keep$Age.at.visit.1)
maxage <- max(keep$Age.at.visit.1)

# Z-scores by genotype and haplotype
ecl_cont_vars <- c("ecl_iaa_z","ecl_gada_z","ecl_iaa_pos","ecl_gada_pos")
#nonnorm <- c("ecl_iaa_z","ecl_gada_z")
nonnorm <- NULL
tab2 <- CreateTableOne(vars=ecl_cont_vars, data=keep, strata="Genotype", test=TRUE)
tab2 <- print(tab2,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)
tab2a <- CreateTableOne(vars=ecl_cont_vars, data=keep_nox, strata="Genotype", test=TRUE)
tab2a <- print(tab2a,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)
tab3 <- CreateTableOne(vars=ecl_cont_vars, data=keep, strata="DR.3.present", test=TRUE)
tab3 <- print(tab3,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)
tab4 <- CreateTableOne(vars=ecl_cont_vars, data=keep, strata="DR4.Present", test=TRUE)
tab4 <- print(tab4,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)

# Percentage DR3 and DR4 by ECL category
dr3 <- keep %>% group_by(ecl_cat) %>% count(DR.3.present) %>% mutate(perc=(n/rowsum(n,ecl_cat))*100)
dr3 <- dr3[dr3$DR.3.present=="yes",]
dr4 <- keep %>% group_by(ecl_cat) %>% count(DR4.Present) %>% mutate(perc=(n/rowsum(n,ecl_cat))*100)
dr4 <- dr4[dr4$DR4.Present=="yes",]

f1 <- ggplot(keep, aes(x=Genotype,y=ecl_iaa_z)) + geom_boxplot() + labs(y="ECL IAA Z-score") + annotate("text", x=1, y=4.5, label="p=0.474")

f1a <- ggplot(keep_nox, aes(x=Genotype,y=ecl_iaa_z)) + geom_boxplot() + labs(y="ECL IAA Z-score") + annotate("text", x=1, y=4.5, label="p=0.363")

f2 <- ggplot(keep, aes(x=Genotype,y=ecl_gada_z)) + geom_boxplot() + labs(y="ECL GADA Z-score") + annotate("text", x=1, y=6, label="p=0.012")

f2a <- ggplot(keep_nox, aes(x=Genotype,y=ecl_gada_z)) + geom_boxplot() + labs(y="ECL GADA Z-score") + annotate("text", x=1, y=6, label="p=0.550")

f3 <- ggplot(keep, aes(x=DR.3.present,y=ecl_iaa_z)) + geom_boxplot() + labs(y="ECL IAA Z-score") + annotate("text", x=0.75, y=6, label="p=0.176")

f4 <- ggplot(keep, aes(x=DR.3.present,y=ecl_gada_z)) + geom_boxplot() + labs(y="ECL GADA Z-score") + annotate("text", x=0.75, y=7, label="p<0.001")

f5 <- ggplot(keep, aes(x=DR4.Present,y=ecl_iaa_z)) + geom_boxplot() + labs(y="ECL IAA Z-score") + annotate("text", x=0.75, y=6, label="p=0.042")

f6 <- ggplot(keep, aes(x=DR4.Present,y=ecl_gada_z)) + geom_boxplot() + labs(y="ECL GADA Z-score") + annotate("text", x=0.75, y=7, label="p=0.745")

f7_tbl <- table(keep$DR.3.present,keep$ecl_cat)
f7_p <- round(chisq.test(f7_tbl)$p.value,3)
f7 <- ggplot(dr3, aes(x=ecl_cat,y=perc)) + geom_bar(stat="identity",position = position_dodge(width=2), width=0.5) + labs(x="",y="%") + annotate("text", x=1, y=45, label="p=0.001")

f8_tbl <- table(keep$DR4.Present,keep$ecl_cat)
f8_p <- round(chisq.test(f8_tbl)$p.value,3)
f8 <- ggplot(dr4, aes(x=ecl_cat,y=perc)) + geom_bar(stat="identity",position = position_dodge(width=2), width=0.5) + labs(x="",y="%") + annotate("text", x=1, y=60, label="p=0.001")
```

# Introduction

The purpose of this analysis is to examine the association between ECL antibodies and HLA genotype/haplotype in the TrialNet cohort at visit 1.

Subject 136806 was excluded from analysis because the mIAA Z-score at visit 1 was equal to 21.04.

# Methods

Z-scores for ECL GADA (ecl_gada_z) and ECL IAA (ecl_iaa_z) were calculated as (subject's ab level - mean) / SD.  Subjects were classified as positive for ECL IAA (ecl_iaa_pos) if their value was >0.006, and were classified as positive for ECL GADA (ecl_gada_pos) if their value was >0.023.

ECL Z-scores were not normally distributed, and were compared using the Kruskal-Wallis test.  ECL positivity was compared using the chi-square test.

# Results

There were `r n` subjects with ECL antibody results and HLA genotyping included in this analysis.

Table 1 shows descriptive statistics for the entire cohort, and Table 1a shows descriptive statistics by ECL positivity.  The minimum age was `r minage` and the maximum age was `r maxage`.

# Table 1 - Descriptive statistics for the entire cohort
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1)
```
<br>

# Table 1a - descriptive statistics by ECL positivity
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1a)
```
<br>

The comparison of ECL antibodies by genotype is shown in Table 2.

# Table 2 - ECL antibodies by genotype
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2)
```
<br>

The comparison of ECL antibodies by genotype (excluding DRx/x)is shown in Table 2a.

# Table 2a - ECL antibodies by genotype (excluding DR x/x)
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2a)
```
<br>

The comparison of ECL antibodies by DR3 haplotype is shown in Table 3.

# Table 3 - ECL antibodies by DR3 haplotype
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab3)
```
<br>

The comparison of ECL antibodies by DR4 haplotype is shown in Table 4.

# Table 4 - ECL antibodies by DR4 haplotype
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab4)
```

# Figure 1 - ECL IAA Z-score by genotype
```{r, echo=FALSE}
f1
```

# Figure 1a - ECL IAA Z-score by genotype (excluding DRx/x)
```{r, echo=FALSE}
f1a
```

# Figure 2 - ECL GADA Z-score by genotype
```{r, echo=FALSE}
f2
```

# Figure 2a - ECL GADA Z-score by genotype (excluding DRx/x)
```{r, echo=FALSE}
f2a
```

# Figure 3 - ECL IAA Z-score by DR3 haplotype
```{r, echo=FALSE}
f3
```

# Figure 4 - ECL GADA Z-score by DR3 haplotype
```{r, echo=FALSE}
f4
```

# Figure 5 - ECL IAA Z-score by DR4 haplotype
```{r, echo=FALSE}
f5
```

# Figure 6 - ECL GADA Z-score by DR4 haplotype
```{r, echo=FALSE}
f6
```

# Figure 7 - DR3 haplotype by ECL status
```{r, echo=FALSE}
f7
```

# Figure 8 - DR4 haplotype by ECL status
```{r, echo=FALSE}
f8
```
