---
title: "TrialNet ECL/HLA analysis for ADA - biochemical assays"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)


knitr::opts_chunk$set(echo = TRUE)

library(tableone)
library(sjlabelled)
library(ggplot2)
library(dplyr)

# read in data
data <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\Andrea Steck\\ECL ab and genotype\\Data\\191015 final dataset 1287 subjects with genotype.csv",
                 na.strings = c("NA",""," "))

keep <- data[,c("PTID","mIAA.visit.1","GAD65.visit.1","HLA.DR3.4","Genotype","DR.3.present","DR4.Present","Age.at.visit.1",
                "Gender","Number.of.positive.biochemical.antibodies.visit.1","miaa_cat_2","gad65_cat_2")]
keep$numpos <- as.factor(keep$Number.of.positive.biochemical.antibodies.visit.1)

keep$biochem_status[keep$Number.of.positive.biochemical.antibodies.visit.1==0] <- 'Ab negative'
keep$biochem_status[keep$Number.of.positive.biochemical.antibodies.visit.1==1] <- 'Single ab positive'
keep$biochem_status[keep$Number.of.positive.biochemical.antibodies.visit.1>1] <- 'Multiple ab positive'
set_label(keep$biochem_status)="Biochemical ab status"

# remove people who do not have HLA data
keep <- keep[keep$HLA.DR3.4!='Missing',]
keep$HLA.DR3.4 <- droplevels(keep$HLA.DR3.4)

# remove participant 136806 because mIAA Z-score is >20, not plausible
keep <- keep[keep$PTID != "136806",]

# calculate Z-score for antibodies
# mIAA mean 0.02715 SD 0.1160666
# GAD65 0.2618 SD 0.3162107
keep$miaa_z <- (keep$mIAA.visit.1 - 0.02715) / 0.1160666
keep$gad65_z <- (keep$GAD65.visit.1 - 0.2618) / 0.3162107

# categorize antibodies
keep$miaa_pos <- as.factor(ifelse(keep$mIAA.visit.1>0.01,"Positive","Negative"))
keep$gad65_pos <- as.factor(ifelse(keep$GAD65.visit.1>0.032,"Positive","Negative"))
# for those who are single ab positive, do not count them as positive unless the ab was confirmed
keep[keep$`Number.of.positive.biochemical.antibodies.visit.1`==1 & keep$miaa_pos=="Positive" & 
       keep$miaa_cat_2==0,]$miaa_pos <- "Negative"
keep[keep$`Number.of.positive.biochemical.antibodies.visit.1`==1 & keep$gad65_pos=="Positive" & 
       !is.na(keep$gad65_cat_2) & keep$gad65_cat_2==0,]$gad65_pos <- "Negative"
keep$either_pos <- ifelse(keep$miaa_pos=="Positive" | keep$gad65_pos=="Positive", "Positive", "Negative")
set_label(keep$either_pos)="mIAA or GAD65 ab positive"
keep$ab_cat <- ifelse(keep$miaa_pos=="Positive" & keep$gad65_pos=="Positive", "mIAA and GAD65 +", 
                       ifelse(keep$miaa_pos=="Positive" & keep$gad65_pos=="Negative","mIAA +",
                              ifelse(keep$miaa_pos=="Negative" & keep$gad65_pos=="Positive","GAD65 +","Ab -")))


# number of subjects is the number with non-missing ab values
n <- nrow(keep[!is.na(keep$mIAA.visit.1) | !is.na(keep$GAD65.visit.1),])

# make dataframe without DRx/x
keep_nox <- keep[keep$Genotype != "DR x/x",]
keep_nox$Genotype <- droplevels(keep_nox$Genotype)

# Table 1
tab1vars <- c("Age.at.visit.1","Gender","HLA.DR3.4","Genotype","DR.3.present","DR4.Present","biochem_status")
tab1 <- CreateTableOne(vars=tab1vars, data=keep[keep$HLA.DR3.4!="Missing",])
tab1 <- print(tab1,varLabels=TRUE,showAllLevels=TRUE)
tab1a <- CreateTableOne(vars=tab1vars, data=keep, strata="either_pos", test=TRUE)
tab1a <- print(tab1a,varLabels=TRUE,showAllLevels=TRUE)

minage <- min(keep$Age.at.visit.1)
maxage <- max(keep$Age.at.visit.1)

# Z-scores by genotype and haplotype
cont_vars <- c("miaa_z","gad65_z","miaa_pos","gad65_pos")
#nonnorm <- c("miaa_z","gad65_z")
nonnorm <- NULL
tab2 <- CreateTableOne(vars=cont_vars, data=keep, strata="Genotype", test=TRUE)
tab2 <- print(tab2,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)
tab2a <- CreateTableOne(vars=cont_vars, data=keep_nox, strata="Genotype", test=TRUE)
tab2a <- print(tab2a,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)
tab3 <- CreateTableOne(vars=cont_vars, data=keep, strata="DR.3.present", test=TRUE)
tab3 <- print(tab3,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)
tab4 <- CreateTableOne(vars=cont_vars, data=keep, strata="DR4.Present", test=TRUE)
tab4 <- print(tab4,varLabels=TRUE,showAllLevels=TRUE,nonnormal = nonnorm)

# Percentage DR3 and DR4 by ab category
dr3 <- keep %>% group_by(ab_cat) %>% count(DR.3.present) %>% mutate(perc=(n/rowsum(n,ab_cat))*100)
dr3 <- dr3[dr3$DR.3.present=="yes",]
dr4 <- keep %>% group_by(ab_cat) %>% count(DR4.Present) %>% mutate(perc=(n/rowsum(n,ab_cat))*100)
dr4 <- dr4[dr4$DR4.Present=="yes",]

f1 <- ggplot(keep, aes(x=Genotype,y=miaa_z)) + geom_boxplot() + labs(y="mIAA Z-score") + annotate("text", x=1, y=6, label="p=0.192")

f1a <- ggplot(keep_nox, aes(x=Genotype,y=miaa_z)) + geom_boxplot() + labs(y="mIAA Z-score") + annotate("text", x=1, y=2.25, label="p=0.104")

f2 <- ggplot(keep, aes(x=Genotype,y=gad65_z)) + geom_boxplot() + labs(y="GAD65 Z-score") + annotate("text", x=1, y=3.5, label="p=0.001")

f2a <- ggplot(keep_nox, aes(x=Genotype,y=gad65_z)) + geom_boxplot() + labs(y="GAD65 Z-score") + annotate("text", x=1, y=3.5, label="p=0.504")

f3 <- ggplot(keep, aes(x=DR.3.present,y=miaa_z)) + geom_boxplot() + labs(y="mIAA Z-score") + annotate("text", x=0.75, y=6, label="p=0.310")

f4 <- ggplot(keep, aes(x=DR.3.present,y=gad65_z)) + geom_boxplot() + labs(y="GAD65 Z-score") + annotate("text", x=0.75, y=3.75, label="p<0.001")

f5 <- ggplot(keep, aes(x=DR4.Present,y=miaa_z)) + geom_boxplot() + labs(y="mIAA Z-score") + annotate("text", x=0.75, y=6, label="p=0.985")

f6 <- ggplot(keep, aes(x=DR4.Present,y=gad65_z)) + geom_boxplot() + labs(y="GAD65 Z-score") + annotate("text", x=0.75, y=3.5, label="p=0.353")

f7_tbl <- table(keep$DR.3.present,keep$ab_cat)
f7_p <- round(chisq.test(f7_tbl)$p.value,3)
f7 <- ggplot(dr3, aes(x=ab_cat,y=perc)) + geom_bar(stat="identity",position = position_dodge(width=2), width=0.5) + labs(x="",y="%") + annotate("text", x=1, y=45, label="p=0.016")

f8_tbl <- table(keep$DR4.Present,keep$ab_cat)
f8_p <- round(chisq.test(f8_tbl)$p.value,3)
f8 <- ggplot(dr4, aes(x=ab_cat,y=perc)) + geom_bar(stat="identity",position = position_dodge(width=2), width=0.5) + labs(x="",y="%") + annotate("text", x=1, y=65, label="p=0.022")
```

# Introduction

The purpose of this analysis is to examine the association between mIAA and GAD65 antibodies and HLA genotype/haplotype in the TrialNet cohort at visit 1.

# Methods

In this version of the analysis, subjects who were single antibody positive at the first visit were classified as negative if the same antibody was not positive at the second visit.

Subject 136806 was excluded from analysis because the mIAA Z-score at visit 1 was equal to 21.04.

Z-scores for GAD65 (gad65_z) and mIAA (miaa_z) were calculated as (subject's ab level - mean) / SD.  Subjects were classified as positive for mIAA (miaa_pos) if their value was >0.01, and were classified as positive for GAD65 (gad65_pos) if their value was >0.032.

Antibody Z-scores were not normally distributed, and were compared using the Kruskal-Wallis test.  Antibody positivity was compared using the chi-square test.

# Results

There were `r n` subjects with antibody results and HLA genotyping included in this analysis.

Table 1 shows descriptive statistics for the entire cohort, and Table 1a shows descriptive statistics by antibody positivity.  The minimum age was `r minage` and the maximum age was `r maxage`.

# Table 1 - Descriptive statistics for the entire cohort
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1)
```
<br>

# Table 1a - descriptive statistics by antibody positivity
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1a)
```
<br>

The comparison of antibodies by genotype is shown in Table 2.

# Table 2 - antibodies by genotype
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2)
```
<br>

The comparison of antibodies by genotype (excluding DRx/x) is shown in Table 2a.

# Table 2a - antibodies by genotype (excluding DR x/x)
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2a)
```
<br>

The comparison of antibodies by DR3 haplotype is shown in Table 3.

# Table 3 - antibodies by DR3 haplotype
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab3)
```
<br>

The comparison of antibodies by DR4 haplotype is shown in Table 4.

# Table 4 - antibodies by DR4 haplotype
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab4)
```

# Figure 1 - mIAA Z-score by genotype
```{r, echo=FALSE}
f1
```

# Figure 1a - mIAA Z-score by genotype (excluding DRx/x)
```{r, echo=FALSE}
f1a
```

# Figure 2 - GAD65 Z-score by genotype
```{r, echo=FALSE}
f2
```

# Figure 2a - GAD65 Z-score by genotype (excluding DRx/x)
```{r, echo=FALSE}
f2a
```

# Figure 3 - mIAA Z-score by DR3 haplotype
```{r, echo=FALSE}
f3
```

# Figure 4 - GAD65 Z-score by DR3 haplotype
```{r, echo=FALSE}
f4
```

# Figure 5 - mIAA Z-score by DR4 haplotype
```{r, echo=FALSE}
f5
```

# Figure 6 - GAD65 Z-score by DR4 haplotype
```{r, echo=FALSE}
f6
```

# Figure 7 - DR3 haplotype by antibody status
```{r, echo=FALSE}
f7
```

# Figure 8 - DR4 haplotype by antibody status
```{r, echo=FALSE}
f8
```
