---
title: "Maki - %ZsGreen-1"
author: "Kristen Campbell"
date: "10/7/2019"
output: word_document
---
```{r,echo=FALSE,message=FALSE,results='hide',warning=FALSE}
library(knitr)
#library(labelled)
options(scipen=999)
#code for data prep: 
source('S:/Shared Material/Shared Code/R/temp_table1.R')
source('C:/Users/campbkri/Documents/GitHub/BDC-Code/Maki Nakayama/tcell_analysis.R')

```

# Methods

To compare the %ZsGreen-1 by more than 2 groups, one way anova (ANOVA) was used.  The overall F-test was reported to test for any differences in %ZsGreen-1 by group, and Tukey's Honestly Significant Difference (Tukey's HSD) was used to perform pairwise comparisons while maintaining an alpha of 0.05. 

To compare the %ZsGreen-1 by 2 groups (AM vs. LO), a two sample independent t-test was used.

Analysis was performed using R version 3.6.0 and an alpha level of 0.05 was maintained for all tests.

## Table 1: Overall Comparison (8 Groups):

Overall, there was a significant difference in %ZsGreen-1 Positive by fluorochorme group (p= `r all.p`).  The pairwise groups that were signficantly different from each other are listed in Table 1. 

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tuk, align = c("l", "l"))
```


## Table 2: Comparison (4 Groups):
Overall, there was a significant difference in %ZsGreen-1 Positive by fluorochorme group (p= `r four.p`).  The pairwise groups that were signficantly different from each other are listed in Table 2. 

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tuk.4, align = c("l", "l"))
```

## Table 3: Comparison (2 Groups):

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.2grp, align = c("l", "l"))
```
