---
title: "A1c Analysis"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)

# data upload
analysis_raw <- read_csv("S:/Laura/BDC/Projects/Cari Berget/a1c_hcl_omnipod/DATA/HCLClinicalObservati_DATA_2023-06-14_1345.csv")
```

```{r, include = FALSE}
# VARS of interest
analysis_t1 = analysis_raw %>% select(record_id, redcap_event_name, gly_time_am, gly_time_am_ltd, glytime)
```

Objectives: 
1.	To describe use of the Omnipod 5 AID system and glycemic outcomes in a clinical sample of youth starting the Omnipod 5 system for their clinical care. 
2. Describe psychosocial impacts of the Omnipod 5 system in a clinical sample of youth starting Omnipod 5 for their clinical care


1.	Describe change over time in System Use and glycemic variables in Linear mixed model with the following time points: 
•	Baseline (up to 14 days prior to starting HCL)
•	1 month post HCL start
•	3 month post HCL start
•	6 month post HCL start
•	*would like p-values for difference between baseline and each time point

o	Groups: 
o	Full cohort
o	Low, middle, high baseline A1C 
	Low= ≤7%
	Middle= 7.1-8.9%
	High= 9%+
o	Age groups
	Preschool = 5.9 and younger
	Child= 6-13.9
	Adolescent= 14.0-17.9
	Young adult= 18+

o	 discontinuation defined as <10% Automated Mode use at any time point.  All patients with >10% Automated Mode use included in model at each time point. Report #s of Automated Mode discontinuers over time as well as CGM use (% Sensor Wear) for discontinuers over time
Variables:
System use variables
•	% time in Automated Mode
•	% time in Automated: Limited
•	% time in Automated: Activity
•	% time in Manual Mode
•	% Sensor wear
•	Total Daily Dose (TDD)
•	% TDD as bolus
•	% TDD as basal
•	# meal boluses per day 
•	# boluses/day
•	% overrides
Glycemic Variables:
•	HbA1c (%)
•	GMI (%)
•	Average sensor glucose value
•	Standard Deviation of Sensor Glucose Values
•	% under 54 mg/dL
•	% time sensor values 54-69 mg/dL
•	% time sensor values 70-180 mgdL
•	% time sensor values 181-250 mg/dL
•	% time sensor values > 250 mg/dL
•	Analysis for % time in ranges
o	70-180, >180, >250, <70, <54
•	NOTE: to calculate the time > 180 will need to combine the ’181-250’ data point + ‘>250’ data point
•	NOTE: to calculate the time <70, will need to combine the ‘<54’ and ’54-69’ data point
•	% Meeting ADA goals of
o	HbA1c ≤ 7.0%
o	TIR (70-180) ≥ 70% 
o	TBR (<70) < 4%
o	HbA1c ≤ 7.0%, TIR≥ 70%, TBR< 4% 

2. Psychosocial analysis: 
o	Survey data for parents 
o	INSPIRE
o	Hypoglycemia Fear Survey (HFS)
o	Problem Areas in Diabetes (PAID-PR)
o	Survey for youth
o	Hypoglycemia Fear survey (HFS)
o	Problem Areas in Diabetes (PAID)- Pediatric 
o	INSPIRE
o	SEEDS Survey (only for ages 13+)
