---
title: "A1c Analysis"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(table1)
library(nlme)
library(lme4)
library(emmeans)

# data upload
analysis_raw <- read_csv("S:/Laura/BDC/Projects/Cari Berget/a1c_hcl_omnipod/DATA/HCLClinicalObservati_DATA_2023-06-14_1345.csv")

# analysis_grps = analysis_raw %>% select(record_id, redcap_event_name, demo_c_hcl)
```

```{r functions, include = FALSE}
# fit model fun
fit_mod = function(outcome,df){

    # Fit model
  f = as.formula(paste0(outcome,"~","redcap_event_name"))
  mod = lme(f,
            random = ~1|record_id,
            data = df,
            na.action = na.omit)
 
  print(outcome)
  
    # Anova
  mod_anova = anova.lme(mod, type="marginal")
  print(kable(mod_anova,digits = 3,caption = "Test of Overall Effect"))

 
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ redcap_event_name, adjust="none")
  mod_means2 = emmeans(mod,specs=~ redcap_event_name, adjust="none")

  print(kable(mod_means$contrasts[1:3,],digits = 3,caption = "Timepoint Contrasts"))
  print(kable(mod_means2,digits = 3,caption = "Timepoint Means"))
 #print(kable(mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE)))

   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
}

```

```{r dc and scoring, include = FALSE}
# timepoints
tps = c("child_bl_arm_1", "child_m1_arm_1", "child_m3_arm_1", "child_m6_arm_1")

# SYSTEM USE/GLYCEMIC VARS of interest
analysis_gly_t1 = analysis_raw %>% select(record_id, redcap_event_name,demo_c_age, gly_time_am, gly_time_am_ltd, gly_time_am_act,
                                      gly_time_mm, gly_sensorwear, gly_tdd, gly_tdd_bolus, gly_tdd_basal, gly_dailymealbolus, gly_dailybolus,
                                      gly_overrides, gly_a1c, gly_gmi, gly_sensor_mean, gly_sensor_sd, gly_sensor_u54, gly_sensor_54_69,
                                      gly_sensor_70_180, gly_sensor_181_250, gly_sensor_g250) %>% filter(redcap_event_name %in% tps)
# fill in ages
analysis_gly_t1 = analysis_gly_t1 %>% group_by(record_id) %>% fill(demo_c_age) %>% ungroup()

# CHILD BASELINE
psy_soc_child_bl = analysis_raw %>% select(record_id, demo_c_age,redcap_event_name, c_b_header_timepoint:c_b_inspire17,
                                           ya_b_header_timepoint:ya_b_inspire17, seeds_b1:seeds_b20) %>% filter(redcap_event_name == "child_bl_arm_1")
# score surveys
psy_soc_child_bl = psy_soc_child_bl %>% mutate(cb_inspire_total = select(., c_b_inspire1:c_b_inspire17) %>% rowMeans()*25, 
                                               ya_b_inspire_total = select(., ya_b_inspire1:ya_b_inspire17) %>% rowMeans()*25,
                                               inspire_score = ifelse(demo_c_age < 13, cb_inspire_total, ya_b_inspire_total), # INSPIRE TOTAL
                                               
                                               c_hfs_maintain = select(., c_b_hfs_behave3, c_b_hfs_behave4, c_b_hfs_behave7) %>% rowSums(),
                                               c_hfs_worry_bg = select(., c_b_hfs_worry11:c_b_hfs_worry14, c_b_hfs_worry16, c_b_hfs_worry18,
                                                                       c_b_hfs_worry19, c_b_hfs_worry22:c_b_hfs_worry23) %>% rowSums(),
                                               c_hfs_worry_soc = select(., c_b_hfs_worry15,c_b_hfs_worry17,
                                                                        c_b_hfs_worry20, c_b_hfs_worry21, c_b_hfs_worry25) %>% rowSums(),
                                               
                                               ya_hfs_maintain = select(., ya_b_hfs_behave3, ya_b_hfs_behave4, ya_b_hfs_behave7) %>% rowSums(),
                                               ya_hfs_worry_bg = select(., ya_hfs_worry_b11:ya_hfs_worry_b14, ya_hfs_worry_b16, ya_hfs_worry_b18,
                                                                       ya_hfs_worry_b19, ya_hfs_worry_b22:ya_hfs_worry_b23) %>% rowSums(),
                                               ya_hfs_worry_soc = select(., ya_hfs_worry_b15,ya_hfs_worry_b17,
                                                                        ya_hfs_worry_b20, ya_hfs_worry_b21, ya_hfs_worry_b25) %>% rowSums(),
                                               hfs_maintain = ifelse(demo_c_age < 13, c_hfs_maintain, ya_hfs_maintain),
                                               hfs_worry_bg = ifelse(demo_c_age < 13, c_hfs_worry_bg, ya_hfs_worry_bg),
                                               hfs_worry_soc = ifelse(demo_c_age < 13, c_hfs_worry_soc, ya_hfs_worry_soc), # HFS
                                               
                                               cb_paid =  select(., c_b_paid1:c_b_paid20) %>% rowSums(),
                                               cb_paid_score = (80 - cb_paid)*25/20,
                                               ya_paid =  select(., ya_b_paid1:ya_b_paid20) %>% rowSums(),
                                               ya_paid_score = (80 - ya_paid)*25/20,
                                               paid_score = ifelse(demo_c_age < 13, cb_paid_score, ya_paid_score) # PAID
                                   )  %>% select(record_id, redcap_event_name, demo_c_age, inspire_score, hfs_maintain:hfs_worry_soc,paid_score)

# CHILD FU
psy_soc_child_fu = analysis_raw %>% select(record_id,redcap_event_name, c_f_timepoint:c_f_inspire17,
                                           ya_f_header_timepoint:ya_f_inspire17, seeds_fu1:seeds_fu20) %>% filter(redcap_event_name %in% c("child_m1_arm_1","child_m3_arm_1","child_m6_arm_1"))
psy_soc_child_fu = left_join(psy_soc_child_fu, psy_soc_child_bl %>% select(record_id, demo_c_age) %>% unique())

psy_soc_child_fu = psy_soc_child_fu %>% mutate(cb_inspire_total = select(., c_f_inspire1:c_f_inspire17) %>% rowMeans()*25, 
                                               ya_f_inspire_total = select(., ya_f_inspire1:ya_f_inspire17) %>% rowMeans()*25,
                                               inspire_score = ifelse(demo_c_age < 13, cb_inspire_total, ya_f_inspire_total), # INSPIRE TOTAL
                                               
                                               c_hfs_maintain = select(., c_f_hfs_behave3, c_f_hfs_behave4, c_f_hfs_behave7) %>% rowSums(),
                                               c_hfs_worry_bg = select(., c_f_hfs_worry11:c_f_hfs_worry14, c_f_hfs_worry16, c_f_hfs_worry18,
                                                                       c_f_hfs_worry19, c_f_hfs_worry22:c_f_hfs_worry23) %>% rowSums(),
                                               c_hfs_worry_soc = select(., c_f_hfs_worry15,c_f_hfs_worry17,
                                                                        c_f_hfs_worry20, c_f_hfs_worry21, c_f_hfs_worry25) %>% rowSums(),
                                               
                                               ya_hfs_maintain = select(., ya_f_hfs_behave3, ya_f_hfs_behave4, ya_f_hfs_behave7) %>% rowSums(),
                                               ya_hfs_worry_bg = select(., ya_f_hfs_worry11:ya_f_hfs_worry14, ya_f_hfs_worry16, ya_f_hfs_worry18,
                                                                       ya_f_hfs_worry19, ya_f_hfs_worry22:ya_f_hfs_worry23) %>% rowSums(),
                                               ya_hfs_worry_soc = select(., ya_f_hfs_worry15,ya_f_hfs_worry17,
                                                                        ya_f_hfs_worry20, ya_f_hfs_worry21, ya_f_hfs_worry25) %>% rowSums(),
                                               hfs_maintain = ifelse(demo_c_age < 13, c_hfs_maintain, ya_hfs_maintain),
                                               hfs_worry_bg = ifelse(demo_c_age < 13, c_hfs_worry_bg, ya_hfs_worry_bg),
                                               hfs_worry_soc = ifelse(demo_c_age < 13, c_hfs_worry_soc, ya_hfs_worry_soc), # HFS
                                               
                                               cb_paid =  select(., c_f_paid1:c_f_paid20) %>% rowSums(),
                                               cb_paid_score = (80 - cb_paid)*25/20,
                                               ya_paid =  select(., ya_f_paid1:ya_f_paid20) %>% rowSums(),
                                               ya_paid_score = (80 - ya_paid)*25/20,
                                               paid_score = ifelse(demo_c_age < 13, cb_paid_score, ya_paid_score) # PAID
                                   )  %>% select(record_id, redcap_event_name, demo_c_age, inspire_score, hfs_maintain:hfs_worry_soc,paid_score)


# PARENT BL
psy_soc_p_bl = analysis_raw %>% select(record_id,redcap_event_name, p_b_inspire1:p_b_inspire21, p_b_hfs_behave1:p_b_hfs_worry26,
                                       p_b_paid1:p_b_paid18) %>% filter(redcap_event_name == "parent_bl_arm_1")

psy_soc_p_bl = psy_soc_p_bl %>% mutate(inspire_score_p = select(., p_b_inspire1:p_b_inspire21) %>% rowMeans()*25, # inspire
                                               
                                               hfs_maintain_p = select(., p_b_hfs_behave3, p_b_hfs_behave4, p_b_hfs_behave7) %>% rowSums(),
                                               hfs_worry_bg_p = select(., p_b_hfs_worry12:p_b_hfs_worry15, p_b_hfs_worry16, p_b_hfs_worry17,
                                                                       p_b_hfs_worry20, p_b_hfs_worry23:p_b_hfs_worry26) %>% rowSums(),
                                               hfs_worry_soc_p = select(., p_b_hfs_worry16,p_b_hfs_worry18,
                                                                        p_b_hfs_worry19, p_b_hfs_worry21, p_b_hfs_worry22) %>% rowSums(), # HFS
                                               
                                               paid_p =  select(., p_b_paid1:p_b_paid18) %>% rowSums(),
                                               paid_score_p = (72-paid_p)*25/18 
                                   )  %>% select(record_id, redcap_event_name, inspire_score_p, hfs_maintain_p:hfs_worry_soc_p,paid_score_p)

# PARENT FU
psy_soc_p_fu = analysis_raw %>% select(record_id,redcap_event_name, p_f_inspire1:p_f_inspire21, p_f_hfs_behave1:p_f_hfs_worry26,
                                       p_f_paid1:p_f_paid18) %>% filter(redcap_event_name %in% c("parent_m1_arm_1","parent_m3_arm_1","parent_m6_arm_1"))

psy_soc_p_fu = psy_soc_p_fu %>% mutate(inspire_score_p = select(., p_f_inspire1:p_f_inspire21) %>% rowMeans()*25, 
                                       hfs_maintain_p = select(., p_f_hfs_behave3, p_f_hfs_behave4, p_f_hfs_behave7) %>% rowSums(),
                                               hfs_worry_bg_p = select(., p_f_hfs_worry12:p_f_hfs_worry15, p_f_hfs_worry16, p_f_hfs_worry17,
                                                                       p_f_hfs_worry20, p_f_hfs_worry23:p_f_hfs_worry26) %>% rowSums(),
                                               hfs_worry_soc_p = select(., p_f_hfs_worry16,p_f_hfs_worry18,
                                                                        p_f_hfs_worry19, p_f_hfs_worry21, p_f_hfs_worry22) %>% rowSums(), # HFS
                                               
                                               paid_p =  select(., p_f_paid1:p_f_paid18) %>% rowSums(),
                                               paid_score_p = (72-paid_p)*25/18 
                                   )  %>% select(record_id, redcap_event_name, inspire_score_p, hfs_maintain_p:hfs_worry_soc_p,paid_score_p)

# merge datasets
psy_soc_child = full_join(psy_soc_child_bl, psy_soc_child_fu) %>% unique()
psy_soc_p= full_join(psy_soc_p_bl, psy_soc_p_fu) %>% mutate(redcap_event_name = gsub("parent_","child_",redcap_event_name)) %>% unique()

psy_soc = full_join(psy_soc_child, psy_soc_p)%>% unique()
psy_soc = psy_soc%>% group_by(record_id) %>% fill(demo_c_age) %>% ungroup() %>% filter(record_id != 125)
# child outcomes
analysis = full_join(analysis_gly_t1, psy_soc)

analysis$redcap_event_name = factor(analysis$redcap_event_name, levels = c("child_bl_arm_1","child_m1_arm_1","child_m3_arm_1","child_m6_arm_1"),
                                    labels = c("Baseline", "Month 1", "Month 3", "Month 6"))
# discontinuation
analysis = analysis %>% mutate(discontinuation = ifelse(gly_time_am < 10, 1, 0)) 

```

```{r t1setup, include = FALSE}
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))}

child_t1 = table1(~.|redcap_event_name, data = analysis %>% select(-record_id), render.continuous= my.render.cont)
```
# Background

Objectives: 
1.	To describe use of the Omnipod 5 AID system and glycemic outcomes in a clinical sample of youth starting the Omnipod 5 system for their clinical care. 
2.  Describe psychosocial impacts of the Omnipod 5 system in a clinical sample of youth starting Omnipod 5 for their clinical care

-	Describe change over time in System Use and glycemic variables in Linear mixed model
-	% Meeting ADA goals of:HbA1c ≤ 7.0%,TIR (70-180) ≥ 70% 	TBR (<70) < 4%,	HbA1c ≤ 7.0%, TIR≥ 70%, TBR< 4% 
- Report #s of Automated Mode discontinuers over time as well as CGM use (% Sensor Wear) for discontinuers over time

# Outstanding data questions/remarks

- seems like a lot of missing data for the survey outcomes
- no month 1 visits for parents
- very few month 1 visits for psy/soc outcomes for children
- I don't think I have the seeds survey scoring, I may have misplaced this if you sent this already.
- list of subjects to exclude? , i have subj 136, 64 with no age
- is discontinuation considered at baseline? only 2 non baseline below 10% in AM (id34 m3, id131 m1)
- 15 subj dont have a gly time am for baseline visit

# Methods

Discontinuation defined as <10% Automated Mode use at any time point.  All patients with >10% Automated Mode use included in model at each time point

Change in outcomes over time was modeled with a linear mixed model with random intercept for subject and contrasts comparing each timepoint to baseline.  

# Analysis 

## Discontinuation

- 119/141 subjects had < 10% at baseline, 1 subject each for months 1 and 3
- subject 34 missing the sensorwear variable

```{r,echo = FALSE}
kable(analysis %>% filter(discontinuation==1) %>% group_by(redcap_event_name) %>%  summarise(n(), mean_sensorwear = mean(gly_sensorwear, na.rm = T)),digits = 3 )
```

```{r, include=FALSE}
analysis = analysis %>% filter((discontinuation == 0 | redcap_event_name == "Baseline"))
```

## Table 1

```{r t1, echo = FALSE}
child_t1
```

## Full Cohort

## % Meeting ADA Goals
ex: At Baseline, 35% of subjects had an a1c at or below 7%, and 12% of subjects met ADA goals for A1c, TIR and TBR  together.

```{r,echo = FALSE}
kable(analysis %>% mutate(a1c_lt7 = ifelse(gly_a1c <= 7,1,0),
                    tir_gt70 = ifelse(gly_sensor_70_180 >= 70,1,0),
                    tbr_lt4 = ifelse(gly_sensor_u54 + gly_sensor_54_69 < 4, 1,0),
                    all_ada = ifelse(a1c_lt7 + tir_gt70 + tbr_lt4 < 3, 0, 1)) %>% 
  group_by(redcap_event_name) %>%  summarise(p_a1c = mean(a1c_lt7, na.rm = T),
                                             p_tir = mean(tir_gt70, na.rm = T),
                                             p_tbr=  mean(tbr_lt4, na.rm = T),
                                             p_all_ada= mean(all_ada, na.rm = T)), 
  digits = 3,caption = "% meeting ADA Goals")
```

### SU/Glycemic Outcomes

####	% time in Automated Mode
```{r, echo=FALSE, results='asis'}
fit_mod("gly_time_am", analysis)
```

####	% time in Automated: Limited
```{r, echo=FALSE, results='asis'}
fit_mod("gly_time_am_ltd", analysis)
```

####	% time in Automated: Activity
```{r, echo=FALSE, results='asis'}
fit_mod("gly_time_am_act", analysis)
```

####	% time in Manual Mode
```{r, echo=FALSE, results='asis'}
fit_mod("gly_time_mm", analysis)
```

####	% Sensor wear
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensorwear", analysis)
```

####	Total Daily Dose (TDD)
```{r, echo=FALSE, results='asis'}
fit_mod("gly_tdd", analysis)
```

####	% TDD as bolus
```{r, echo=FALSE, results='asis'}
fit_mod("gly_tdd_bolus", analysis)
```

####	% TDD as basal
```{r, echo=FALSE, results='asis'}
fit_mod("gly_tdd_basal", analysis)
```

#### meal boluses per day 
```{r, echo=FALSE, results='asis'}
fit_mod("gly_dailymealbolus", analysis)
```

#### boluses/day
```{r, echo=FALSE, results='asis'}
fit_mod("gly_dailybolus", analysis)
```

####	% overrides
```{r, echo=FALSE, results='asis'}
fit_mod("gly_overrides", analysis)
```

####	HbA1c (%)
```{r, echo=FALSE, results='asis'}
fit_mod("gly_a1c", analysis)
```

####	GMI (%)
```{r, echo=FALSE, results='asis'}
fit_mod("gly_gmi", analysis)
```

####	Average sensor glucose value
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_mean", analysis)
```

####	Standard Deviation of Sensor Glucose Values
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_sd", analysis)
```

####	% under 54 mg/dL
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_u54", analysis)
```

####	% time sensor values 54-69 mg/dL
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_54_69", analysis)
```

####	% time sensor values <70 mg/dL
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_u70", analysis %>% mutate(gly_sensor_u70 = gly_sensor_u54+ gly_sensor_54_69))
```

####	% time sensor values 70-180 mgdL
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_70_180", analysis)
```

####	% time sensor values >180 mg/dL
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_g180", analysis %>% mutate(gly_sensor_g180 = gly_sensor_181_250+ gly_sensor_g250))
```

####	% time sensor values 181-250 mg/dL
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_181_250", analysis)
```

####	% time sensor values > 250 mg/dL
```{r, echo=FALSE, results='asis'}
fit_mod("gly_sensor_g250", analysis)
```

### Psychosocial Outcomes: 

####	Inspire Parents
```{r, echo=FALSE, results='asis'}
fit_mod("inspire_score_p", analysis)
```

####	hfs maintain Parents
```{r, echo=FALSE, results='asis'}
fit_mod("hfs_maintain_p", analysis)
```

####	hfs worry bg Parents
```{r, echo=FALSE, results='asis'}
fit_mod("hfs_worry_bg_p", analysis)
```

####	hfs worry social Parents
```{r, echo=FALSE, results='asis'}
fit_mod("hfs_worry_soc_p", analysis)
```

####	PAID Parents
```{r, echo=FALSE, results='asis'}
fit_mod("paid_score_p", analysis)
```

####	Inspire child
```{r, echo=FALSE, results='asis'}
fit_mod("inspire_score", analysis)
```

####	hfs maintain child
```{r, echo=FALSE, results='asis'}
fit_mod("hfs_maintain", analysis)
```

####	hfs worry bg child
```{r, echo=FALSE, results='asis'}
fit_mod("hfs_worry_bg", analysis)
```

####	hfs worry social child
```{r, echo=FALSE, results='asis'}
fit_mod("hfs_worry_soc", analysis)
```

####	PAID child
```{r, echo=FALSE, results='asis'}
fit_mod("paid_score", analysis)
```


o	Groups: 
o	Full cohort
o	Low, middle, high baseline A1C 
	Low= ≤7%
	Middle= 7.1-8.9%
	High= 9%+
o	Age groups
	Preschool = 5.9 and younger
	Child= 6-13.9
	Adolescent= 14.0-17.9
	Young adult= 18+