---
title: "HCL Overnight Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/670g obs sleep")
library(arsenal)
library(cgmanalysis)
library(car)
library(skimr)
library(Hmisc)
library(knitr)
library(lme4)
library(nlme)
library(MASS)
library(tidyverse)
```

```{r data import and clean,echo=FALSE,include=FALSE}
# Import Cari's data
source("/Users/timvigers/Documents/GitHub/BDC-Code/Erin Cobry/AdvancedClosedLoopCl_R_2020-01-05_1101.r")
# Demographics
demographics <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(child_ya))
# ID to match Erin's - clean Erin's then demographics
demographics$id <- paste0(demographics$record_id,demographics$demographics_initals)
# remove missing
no_data <- c("27","30LR","38SR","41SW")
demographics <- demographics %>% filter(!(id %in% no_data))
# Manually corrected a few dates in CSV file.
# Read in (manually converted to long format from Erin's Excel spreadsheet)
overnight_alarms <- read.csv("./Data_Cleaned/Overnight Alarm Data.csv",
                             na.strings = c("","No Data"))
# Column names
colnames(overnight_alarms) <- c("id","timepoint","start_date","end_date","num_nights",
                                "num_alarms","num_alarms_in_AM","perc_time_in_AM")
overnight_alarms <- overnight_alarms %>% filter(!(id %in% no_data))
# Format date columns
overnight_alarms[,c("start_date","end_date")] <- 
  lapply(overnight_alarms[,c("start_date","end_date")], 
         function(x) {lubridate::mdy(as.character(x))})
# Calculate number of nights 
overnight_alarms$num_nights <- as.numeric(difftime(overnight_alarms$end_date,
                                                   overnight_alarms$start_date,
                                                   units = "days"))
# Numeric timepoints in months
overnight_alarms$numeric_time <- 
  car::recode(overnight_alarms$timepoint,
              "'Baseline' = 0;'Week 2' = 0.5;'T1' = 3;'T2' = 6;'T3' = 9;'T4' = 12")
# Percent AM time
overnight_alarms$perc_time_in_AM <- overnight_alarms$perc_time_in_AM * 100
# Alarms per day
overnight_alarms$alarms_per_day <- 
  overnight_alarms$num_alarms / overnight_alarms$num_nights
overnight_alarms$alarms_AM_per_day <- 
  overnight_alarms$num_alarms_in_AM / overnight_alarms$num_nights
# Get Erin's participants 
demographics <- demographics[which(demographics$id %in% overnight_alarms$id),]
# Combine CGM and pump hx levels
levels(demographics$demographics_cgmhx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
levels(demographics$demographics_pumphx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
```

# Table 1: Descriptive Characteristics

Excluding participants with no CGM data (27,30,38,41).

```{r table 1,echo=FALSE,results='asis'}
# Labels
labels <- list(demographics_t1d_duration = "Diabetes Dur. (years)", 
               demographics_ethnicity.factor = "Ethnicity",
               demographics_sex.factor = "Sex",
               demographics_cgmhx.factor = "CGM History", 
               demographics_pumphx.factor = "Pump History")
# Table 1
t1 <- tableby(~ hba1c + kwt(demographics_age,"median","q1q3") + 
                kwt(demographics_t1d_duration,"median","q1q3") + 
                demographics_sex.factor + demographics_ethnicity.factor +
                demographics_cgmhx.factor + demographics_pumphx.factor,
              data = demographics)
# Print
summary(t1,labelTranslations = labels)
```

```{r clean CGM files,echo=FALSE,include=FALSE,eval=FALSE}
# Merge files and write CSVs with correct dates
# Files
cgm_files <- list.files("./Data_Raw/Project 2 - HCL Overnight Data Project_Zip/Project 2 - HCL Overnight Data Project",
                        recursive = T,full.names = T)
all_cgm <- data.frame()
# Clean all CGM files - this is extremely slow but only needs to be run once
for (file in cgm_files) {
  # Get ID
  id <- strsplit(file,"/")[[1]][5]
  if (id %in% no_data){
    next()
  }
  # Clean data
  table <- read.csv(file,stringsAsFactors = F,na.strings = "")
  sensor <- which(table$Patient.ID == "Sensor")[1]
  if (is.na(sensor)) {
    next()
  }
  table <- table[sensor:nrow(table),]
  colnames(table) <- as.character(table[2,])
  table <- table[which(!is.na(table$`Sensor Glucose (mg/dL)`) & 
                         table$Index != "Index"),]
  if (nrow(table) == 0) {
    next()
  }
  # Format columns
  table$timestamp <- lubridate::mdy_hms(paste(table$Date,table$Time))
  table$sensorglucose <- as.numeric(as.character(table$`Sensor Glucose (mg/dL)`))
  table$subjectid <- id
  table <- table[,c("subjectid","timestamp","sensorglucose")]
  table$timepoint <- NA
  # Get timepoints
  dates <- overnight_alarms[which(overnight_alarms$id == id),]
  # Baseline 
  start <- dates$start_date[dates$timepoint == "Baseline"]
  end <- dates$end_date[dates$timepoint == "Baseline"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "Baseline"
  # Week 2 
  start <- dates$start_date[dates$timepoint == "week 2"]
  end <- dates$end_date[dates$timepoint == "week 2"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "week 2"
  # T1 
  start <- dates$start_date[dates$timepoint == "T1"]
  end <- dates$end_date[dates$timepoint == "T1"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T1"
  # T2 
  start <- dates$start_date[dates$timepoint == "T2"]
  end <- dates$end_date[dates$timepoint == "T2"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T2"
  # T3 
  start <- dates$start_date[dates$timepoint == "T3"]
  end <- dates$end_date[dates$timepoint == "T3"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T3"
  # T4 
  start <- dates$start_date[dates$timepoint == "T4"]
  end <- dates$end_date[dates$timepoint == "T4"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T4"
  # Remove readings with no time period
  table <- table[!is.na(table$timepoint),]
  # Add to big dataframe
  all_cgm <- rbind(all_cgm,table)
}
# Remove duplicates
all_cgm <- unique(all_cgm)
# Split and write CSVs
all_cgm <- split(all_cgm,list(all_cgm$subjectid,all_cgm$timepoint))
lapply(all_cgm,function(x){
  x <- as.data.frame(x)
  filename <- paste0("./Data_Cleaned/Cleaned CGM/",x$subjectid[1],"_",x$timepoint[1],".csv")
  write.csv(x,file = filename,row.names = F,na = "")
})
# CGM variables
cgmvariables("./Data_Cleaned/Cleaned CGM/","./Data_Cleaned",id_filename = T,outputname = "cgm_variables")
```

```{r cgm data,echo=FALSE,include=FALSE}
# Read in CGM data
cgm <- read.csv("./Data_Cleaned/cgm_variables.csv",stringsAsFactors = F)
# ID and timepoint
cgm$id <- sapply(strsplit(cgm$subject_id,"_"),`[[`,1)
cgm$timepoint <- sapply(strsplit(cgm$subject_id,"_"),`[[`,2)
cgm$timepoint <- factor(cgm$timepoint,levels = c("Baseline","week 2","T1","T2","T3","T4"))
# Select columns
cgm <- cgm %>% select(id,timepoint,nighttime_avg_sens_glucose,
                      percent_time_70_180_night,percent_time_under_70_night)
# Merge
overnight_alarms$id <- as.character(overnight_alarms$id)
overnight_alarms <- left_join(overnight_alarms,cgm,by = c("id","timepoint"))
# Order timepoints
overnight_alarms$timepoint <- factor(overnight_alarms$timepoint,
                                     levels = c("Baseline","week 2",
                                                "T1","T2","T3","T4" ))
levels(overnight_alarms$timepoint) <- c("Baseline","Week 2",
                                        "T1","T2","T3","T4" )
# Write to CSV for SAS
overnight_alarms$log_num_nights <- log(overnight_alarms$num_nights)
write.csv(overnight_alarms,file = "./Data_Cleaned/alarms.csv",row.names = F,na = "")
```

# Table 2: Alarm Frequency

```{r alarm table,echo=FALSE}
alarm_table <- overnight_alarms %>% group_by(timepoint) %>%
  summarise(`Mean Alarms per Night` = round(mean(alarms_per_day,na.rm=T),3),
            `Mean Alarms in AM per Night` = round(mean(alarms_AM_per_day,na.rm=T),3))
kable(alarm_table)
```

# Longitudinal Analysis

## Alarms per Day By Timepoint

```{r alarm plot,echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = alarms_per_day,group = id)) + 
  geom_line() + theme_bw() + xlab("Timepoint") + ylab("Alarms per Day")
alarms_over_time
```

## Number of Alarms per Day Model Results

```{r alarm model,echo=FALSE, message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(num_alarms ~ timepoint + perc_time_in_AM + 
                         offset(log(num_nights)),
                       random = ~1|id,data = overnight_alarms,
                       family = "poisson",
                       correlation = corCAR1(form = ~numeric_time))
results <- broom::tidy(mod_ri_car1,"fixed")
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results[,2:4] <- lapply(results[,2:4],function(x) round(x,3))
results$p.value <- format.pval(results$p.value,eps = 0.0001,digits = 3)
kable(results)
```

For each timepoint, the "estimate" column in the table above represents the average difference from baseline after adjusting for percent time in AM. All four timepoints had a significantly lower number of alarms per day than baseline. Also, for each percentage point increase in time spent spent in AM, the number of alarms per day increased on average by `r results$estimate[results$term == "% Time in AM"]` (p = `r results$p.value[results$term == "% Time in AM"]`).

## Nighttime % TIR by Timepoint

```{r tir plot, echo=FALSE,warning=FALSE,dpi=600}
tir_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = percent_time_70_180_night,group = id)) + 
  geom_line() + theme_bw() + xlab("Timepoint") + ylab("Nighttime % TIR")
tir_over_time
```

## Nighttime % TIR Model Results

```{r tir model,echo=FALSE, message=FALSE}
# The outcome looks pretty normal so we can use a LMM with RI and spatial power
mod_ri_car1 <- lme(percent_time_70_180_night ~ timepoint + perc_time_in_AM,
                   random = ~1|id,data = overnight_alarms,na.action = na.omit,
                   correlation = corCAR1(form = ~numeric_time))
# Residuals look good!
results <- broom::tidy(mod_ri_car1,"fixed")
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results[,2:4] <- lapply(results[,2:4],function(x) round(x,3))
results$p.value <- format.pval(results$p.value,eps = 0.0001,digits = 3)
kable(results)
```

After adjusting for percent time in AM, nighttime %TIR was lower at each timepoint compared to baseline. On average, for a participant with 0% time in AM, nighttime %TIR changed by `r results$estimate[results$term == "Week 2"]` at week 2 compared to baseline (p = `r results$p.value[results$term == "Week 2"]`). The interpretation for other timepoints is similar. Also, for each percentage point increase in time spent spent in AM, nighttime %TIR increased on average by `r results$estimate[results$term == "% Time in AM"]` percentage points (p = `r results$p.value[results$term == "% Time in AM"]`). 

## Nighttime Mean Glucose by Timepoint

```{r mean bg plot, echo=FALSE,warning=FALSE,dpi=600}
mean_bg_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = nighttime_avg_sens_glucose,group = id)) + 
  geom_line() + theme_bw() + xlab("Timepoint") + ylab("Nighttime Mean Glucose")
mean_bg_over_time
```

## Nighttime Mean Glucose Model Results

```{r under70 model,echo=FALSE, message=FALSE}
# The outcome looks pretty normal again so we can use a LMM with RI and spatial power
mod_ri_car1 <- lme(nighttime_avg_sens_glucose ~ timepoint + perc_time_in_AM,
                   random = ~1|id,data = overnight_alarms,na.action = na.omit,
                   correlation = corCAR1(form = ~numeric_time))

results <- broom::tidy(mod_ri_car1,"fixed")
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results[,2:4] <- lapply(results[,2:4],function(x) round(x,3))
results$p.value <- format.pval(results$p.value,eps = 0.0001,digits = 3)
kable(results)
```

There was an overall effect of timepoint on nighttime mean glucose (p = 0.0174). After adjusting for percent time in AM, nighttime mean glucose was higher at each timepoint compared to baseline (although the difference was not statistically significant at T3 or T4). On average, for a participant with 0% time in AM, nighttime mean glucose increased by `r results$estimate[results$term == "Week 2"]` at week 2 compared to baseline (p = `r results$p.value[results$term == "Week 2"]`). The interpretation for other timepoints is similar. Also, for each percentage point increase in time spent spent in AM, nighttime mean glucose changed by `r results$estimate[results$term == "% Time in AM"]` on average (p = `r results$p.value[results$term == "% Time in AM"]`). 
