---
title: "HCL Overnight Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/670g obs sleep")
library(arsenal)
library(cgmanalysis)
library(car)
library(skimr)
library(Hmisc)
library(knitr)
library(lme4)
library(nlme)
library(MASS)
library(tidyverse)
```

```{r data import and clean,echo=FALSE,include=FALSE}
# Import Cari's data
source("/Users/timvigers/Documents/GitHub/BDC-Code/Erin Cobry/AdvancedClosedLoopCl_R_2020-01-05_1101.r")
# Demographics
demographics <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(child_ya))
# ID to match Erin's - clean Erin's then demographics
demographics$id <- paste0(demographics$record_id,demographics$demographics_initals)
# remove missing
no_data <- c("27","30LR","38SR","41SW")
demographics <- demographics %>% filter(!(id %in% no_data))
# Manually corrected a few dates in CSV file.
# Read in (manually converted to long format from Erin's Excel spreadsheet)
overnight_alarms <- read.csv("./Data_Cleaned/Overnight Alarm Data 01-21-2020.csv",
                             na.strings = c("","No Data"))
# Column names
colnames(overnight_alarms) <- c("id","timepoint","start_date","end_date","num_nights",
                                "num_alarms","num_alarms_in_AM","time_in_AM_during_cgm",
                                "actual_time_am","cgm_use","perc_cgm_per_week")
overnight_alarms <- overnight_alarms %>% filter(!(id %in% no_data))
# Format date columns
overnight_alarms[,c("start_date","end_date")] <- 
  lapply(overnight_alarms[,c("start_date","end_date")], 
         function(x) {lubridate::mdy(as.character(x))})
# Calculate number of nights and actual AM time
overnight_alarms$num_nights <- as.numeric(difftime(overnight_alarms$end_date,
                                                   overnight_alarms$start_date,
                                                   units = "days"))
overnight_alarms$actual_time_am <- overnight_alarms$time_in_AM_during_cgm * 
  overnight_alarms$perc_cgm_per_week
# Numeric timepoints in months
overnight_alarms$numeric_time <- 
  car::recode(overnight_alarms$timepoint,
              "'Baseline' = 0;'week 2' = 0.5;'T1' = 3;'T2' = 6;'T3' = 9;'T4' = 12")
overnight_alarms$numeric_time <- 
  as.numeric(as.character(overnight_alarms$numeric_time))
# Percent AM time
overnight_alarms$actual_time_am <- overnight_alarms$actual_time_am * 100
# Alarms per day
overnight_alarms$alarms_per_day <- 
  overnight_alarms$num_alarms / overnight_alarms$num_nights
overnight_alarms$alarms_AM_per_day <- 
  overnight_alarms$num_alarms_in_AM / overnight_alarms$num_nights
# Get Erin's participants 
demographics <- demographics[which(demographics$id %in% overnight_alarms$id),]
# Combine CGM and pump hx levels
levels(demographics$demographics_cgmhx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
levels(demographics$demographics_pumphx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
```

# Table 1: Descriptive Characteristics

Excluding participants with no CGM data (IDs 27, 30, 38, and 41).

```{r table 1,echo=FALSE,results='asis'}
# Labels
labels <- list(demographics_t1d_duration = "Diabetes Dur. (years)", 
               demographics_ethnicity.factor = "Ethnicity",
               demographics_sex.factor = "Sex",
               demographics_cgmhx.factor = "CGM History", 
               demographics_pumphx.factor = "Pump History")
# Table 1
t1 <- tableby(~ hba1c + kwt(demographics_age,"median","q1q3") + 
                kwt(demographics_t1d_duration,"median","q1q3") + 
                demographics_sex.factor + demographics_ethnicity.factor +
                demographics_cgmhx.factor + demographics_pumphx.factor,
              data = demographics)
# Print
summary(t1,labelTranslations = labels)
```

```{r clean CGM files,echo=FALSE,include=FALSE,eval=FALSE}
# Merge files and write CSVs with correct dates
# Files
cgm_files <- list.files("./Data_Raw/Project 2 - HCL Overnight Data Project_Zip/Project 2 - HCL Overnight Data Project",
                        recursive = T,full.names = T)
all_cgm <- data.frame()
# Clean all CGM files - this is extremely slow but only needs to be run once
for (file in cgm_files) {
  # Get ID
  id <- strsplit(file,"/")[[1]][5]
  if (id %in% no_data){
    next()
  }
  # Clean data
  table <- read.csv(file,stringsAsFactors = F,na.strings = "")
  sensor <- which(table$Patient.ID == "Sensor")[1]
  if (is.na(sensor)) {
    next()
  }
  table <- table[sensor:nrow(table),]
  colnames(table) <- as.character(table[2,])
  table <- table[which(!is.na(table$`Sensor Glucose (mg/dL)`) & 
                         table$Index != "Index"),]
  if (nrow(table) == 0) {
    next()
  }
  # Format columns
  table$timestamp <- lubridate::mdy_hms(paste(table$Date,table$Time))
  table$sensorglucose <- as.numeric(as.character(table$`Sensor Glucose (mg/dL)`))
  table$subjectid <- id
  table <- table[,c("subjectid","timestamp","sensorglucose")]
  table$timepoint <- NA
  # Get timepoints
  dates <- overnight_alarms[which(overnight_alarms$id == id),]
  # Baseline 
  start <- dates$start_date[dates$timepoint == "Baseline"]
  end <- dates$end_date[dates$timepoint == "Baseline"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "Baseline"
  # Week 2 
  start <- dates$start_date[dates$timepoint == "week 2"]
  end <- dates$end_date[dates$timepoint == "week 2"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "week 2"
  # T1 
  start <- dates$start_date[dates$timepoint == "T1"]
  end <- dates$end_date[dates$timepoint == "T1"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T1"
  # T2 
  start <- dates$start_date[dates$timepoint == "T2"]
  end <- dates$end_date[dates$timepoint == "T2"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T2"
  # T3 
  start <- dates$start_date[dates$timepoint == "T3"]
  end <- dates$end_date[dates$timepoint == "T3"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T3"
  # T4 
  start <- dates$start_date[dates$timepoint == "T4"]
  end <- dates$end_date[dates$timepoint == "T4"]
  table$timepoint[which(table$timestamp > start & table$timestamp <= end)] <- "T4"
  # Remove readings with no time period
  table <- table[!is.na(table$timepoint),]
  # Add to big dataframe
  all_cgm <- rbind(all_cgm,table)
}
# Remove duplicates
all_cgm <- unique(all_cgm)
# Split and write CSVs
all_cgm <- split(all_cgm,list(all_cgm$subjectid,all_cgm$timepoint))
lapply(all_cgm,function(x){
  x <- as.data.frame(x)
  filename <- paste0("./Data_Cleaned/Cleaned CGM/",x$subjectid[1],"_",x$timepoint[1],".csv")
  write.csv(x,file = filename,row.names = F,na = "")
})
# CGM variables
cgmvariables("./Data_Cleaned/Cleaned CGM/","./Data_Cleaned",id_filename = T,outputname = "cgm_variables")
```

```{r cgm data,echo=FALSE,include=FALSE}
# Read in CGM data
cgm <- read.csv("./Data_Cleaned/cgm_variables.csv",stringsAsFactors = F)
# ID and timepoint
cgm$id <- sapply(strsplit(cgm$subject_id,"_"),`[[`,1)
cgm$timepoint <- sapply(strsplit(cgm$subject_id,"_"),`[[`,2)
cgm$timepoint <- factor(cgm$timepoint,levels = c("Baseline","week 2","T1","T2","T3","T4"))
# Select columns
cgm <- cgm %>% select(id,timepoint,nighttime_avg_sens_glucose,
                      percent_time_70_180_night,percent_time_under_70_night)
# Merge
overnight_alarms <- left_join(overnight_alarms,cgm,by = c("id","timepoint"))
# Order timepoints
overnight_alarms$timepoint <- factor(overnight_alarms$timepoint,
                                     levels = c("Baseline","week 2",
                                                "T1","T2","T3","T4" ))
levels(overnight_alarms$timepoint) <- c("Baseline","Week 2",
                                        "T1","T2","T3","T4" )
# Remove those not using CGM
overnight_alarms_cgm <- overnight_alarms %>% filter(cgm_use == "Y")
```

# Table 2: Alarm Frequency

```{r alarm table,echo=FALSE}
alarm_table <- overnight_alarms_cgm %>% group_by(timepoint) %>%
  summarise(`Median Alarms per Night` = round(median(alarms_per_day,na.rm=T),3),
            `Median Alarms in AM per Night` = round(median(alarms_AM_per_day,na.rm=T),3),
            `Median Actual Time in AM` = round(median(actual_time_am,na.rm=T),3))
kable(alarm_table)
```

# Longitudinal Analysis

## Alarms per Day By Timepoint

### Plot

```{r alarm plot,echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = alarms_per_day,group = id)) + 
  geom_line() + theme_bw() + xlab("Timepoint") + ylab("Alarms per Day")
alarms_over_time
```

### Model Results

```{r alarm model,echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(num_alarms ~ timepoint + actual_time_am + 
                         offset(log(num_nights)),
                       random = ~1|id,data = overnight_alarms_cgm,
                       family = "poisson",
                       correlation = corCAR1(form = ~numeric_time))
results <- broom::tidy(mod_ri_car1,effects = "fixed")
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.0001,digits = 3)
results$estimate <- exp(results$estimate)
conf_int <- as.data.frame(intervals(mod_ri_car1)$fixed)
results$lower_ci <- exp(conf_int$lower)
results$upper_ci <- exp(conf_int$upper)
results$estimate <- exp(conf_int$est.)
kable(results[,c("term","estimate","lower_ci","upper_ci","p.value")])
```

There was no effect of timepoint on the rate of alarms per night after adjusting for actual time in AM. There was also no effect of actual time in AM on the rate of alarms per night.

### Number of Observations at Each Timepoint

```{r alarm obs,echo=FALSE,message=FALSE}
alarm_obs <- overnight_alarms_cgm %>% group_by(timepoint) %>% count(!is.na(num_alarms))
kable(alarm_obs[,c("timepoint","n")])
```

## Nighttime % TIR by Timepoint

### Plot

```{r tir plot, echo=FALSE,warning=FALSE,dpi=600}
tir_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = percent_time_70_180_night,group = id)) + 
  geom_line() + theme_bw() + xlab("Timepoint") + ylab("Nighttime % TIR")
tir_over_time
```

### Model Results

```{r tir model,echo=FALSE, message=FALSE}
# The outcome looks pretty normal so we can use a LMM with RI and spatial power
mod_ri_car1 <- lme(percent_time_70_180_night ~ timepoint + actual_time_am,
                   random = ~1|id,data = overnight_alarms_cgm,na.action = na.omit,
                   correlation = corCAR1(form = ~numeric_time))
# Residuals look good!
results <- broom::tidy(mod_ri_car1,"fixed",confi = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.0001,digits = 3)
conf_int <- as.data.frame(intervals(mod_ri_car1,which = "fixed")$fixed)
results$lower_ci <- conf_int$lower
results$upper_ci <- conf_int$upper
results$estimate <- conf_int$est.
kable(results[,c("term","estimate","lower_ci","upper_ci","p.value")])
```

After adjusting for percent time in AM, nighttime %TIR was lower at each timepoint compared to baseline, although none of these differences were statistically significant and there was no overall effect of timepoint. On average, for each percentage point increase in actual time spent spent in AM, nighttime %TIR increased on average by `r results$estimate[results$term == "% Time in AM"]` percentage points (p = `r results$p.value[results$term == "% Time in AM"]`).

### Number of Observations at Each Timepoint

```{r tir obs,echo=FALSE,message=FALSE}
tir_obs <- overnight_alarms_cgm %>% group_by(timepoint) %>% 
  count(!is.na(percent_time_70_180_night)) %>%
  filter(`!is.na(percent_time_70_180_night)` == "TRUE")
kable(tir_obs[,c("timepoint","n")])
```

## Nighttime Mean Glucose by Timepoint

### Plot

```{r mean bg plot, echo=FALSE,warning=FALSE,dpi=600}
mean_bg_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = nighttime_avg_sens_glucose,group = id)) + 
  geom_line() + theme_bw() + xlab("Timepoint") + ylab("Nighttime Mean Glucose")
mean_bg_over_time
```

### Model Results

```{r under70 model,echo=FALSE, message=FALSE}
# The outcome looks pretty normal again so we can use a LMM with RI and spatial power
mod_ri_car1 <- lme(nighttime_avg_sens_glucose ~ timepoint + actual_time_am,
                   random = ~1|id,data = overnight_alarms_cgm,na.action = na.omit,
                   correlation = corCAR1(form = ~numeric_time))
results <- broom::tidy(mod_ri_car1,"fixed")
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.0001,digits = 3)
conf_int <- as.data.frame(intervals(mod_ri_car1,which = "fixed")$fixed)
results$lower_ci <- conf_int$lower
results$upper_ci <- conf_int$upper
results$estimate <- conf_int$est.
kable(results[,c("term","estimate","lower_ci","upper_ci","p.value")])
```

There was no overall effect of timepoint on nighttime mean glucose (p = 0.8424). For each percentage point increase in time spent spent in AM, nighttime mean glucose changed by `r results$estimate[results$term == "% Time in AM"]` on average (p = `r results$p.value[results$term == "% Time in AM"]`). 

### Number of Observations at Each Timepoint

```{r avg gluc obs,echo=FALSE,message=FALSE}
mean_gluc_obs <- overnight_alarms_cgm %>% group_by(timepoint) %>% 
  count(!is.na(nighttime_avg_sens_glucose)) %>%
  filter(`!is.na(nighttime_avg_sens_glucose)` == "TRUE")
kable(mean_gluc_obs[,c("timepoint","n")])
```
