---
title: "670G Obs. ATTD 2019 - Erin Cobry abstract"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "T:")
library(tableone)
library(redcapAPI)
library(knitr)
library(tidyverse)
library(magrittr)
library(dplyr)
```

```{r define functions,include=FALSE}
fill_NAs <- function(data,threshold,idcols) {
  na_thresh <- ceiling(threshold * ncol(data))-length(idcols) # Determine max NAs
  data <- data[rowSums(is.na(data)) < na_thresh,] # Remove rows with NAs over max
  k <- which(is.na(data), arr.ind=TRUE) # Find remaining NAs
  if(length(k)>0) {
    data[k] <- rowMeans(data[,colnames(data)[-idcols]],na.rm = T)[k[,1]] # Replace with means
  }
  data
}
```

```{r data import and formatting,echo=FALSE,include=FALSE}
# Import
knitr::opts_knit$set(root.dir = "T:")
source("T:/Erin Cobry/ATTD Abstract 2019/Data_Cleaned/AdvancedClosedLoopCl_R_2019-09-30_1049.r")
#source("T:\\Cari Berget\\ATTD Abstract 2019\\Data_Cleaned\\AdvancedClosedLoopCl_R_2019-09-30_1049.r")

#tokens <- read.csv("./api_tokens.csv")
#Parent670g <- as.character(tokens$Token[4])
#rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/", token = Parent670g)
#parent_full <- suppressWarnings(exportRecords(rcon))

# A1c as numeric
data$hba1c <- as.numeric(as.character(data$hba1c))
data$hba1c_baseline <- as.numeric(as.character(data$hba1c_baseline))

# Survey data
# Young adults
# PAID
ya_paid <- data %>% select(record_id,redcap_event_name,ya_paid1:ya_paid20)
ya_paid <- fill_NAs(ya_paid,0.75,1:2)
# Score
ya_paid$paid_score <- apply(ya_paid[,3:ncol(ya_paid)],1,sum)*1.25

# HFS scores (baseline and followup)
ya_hfs_b <- data %>% select(record_id,redcap_event_name,
                            ya_hfs_b_behave1:ya_hfs_b_behave15,
                            ya_hfs_b_worry16:ya_hfs_b_worry33) %>% 
  filter(redcap_event_name == "baseline_arm_2")
ya_hfs_f <- data %>% select(record_id,redcap_event_name,
                            ya_hfs_f_behave1:ya_hfs_f_behave15,
                            ya_hfs_f_worry16:ya_hfs_f_worry33) %>%
  filter(redcap_event_name != "baseline_arm_2")
# Change column names to match
colnames(ya_hfs_b) <- gsub("b_","",colnames(ya_hfs_b))
colnames(ya_hfs_f) <- gsub("f_","",colnames(ya_hfs_f))
# Bind
ya_hfs <- rbind(ya_hfs_b,ya_hfs_f)
# Split by HFS subscale, fill NAs
ya_hfs_behave <- ya_hfs %>% select(record_id,redcap_event_name,
                                   ya_hfs_behave1:ya_hfs_behave15)
ya_hfs_behave <- fill_NAs(ya_hfs_behave,0.75,1:2)
ya_hfs_worry <- ya_hfs %>% select(record_id,redcap_event_name,
                                   ya_hfs_worry16:ya_hfs_worry33)
ya_hfs_worry <- fill_NAs(ya_hfs_worry,0.75,1:2)
# Score
ya_hfs_behave$behave_score <- apply(ya_hfs_behave[,3:ncol(ya_hfs_behave)],1,sum)
ya_hfs_worry$worry_score <- apply(ya_hfs_worry[,3:ncol(ya_hfs_worry)],1,sum)
# calculate Z-score
ya_hfs_behave$behave_score_z <- (ya_hfs_behave$behave_score - mean(ya_hfs_behave$behave_score,na.rm=T)) / sd(ya_hfs_behave$behave_score,na.rm=T)
ya_hfs_worry$worry_score_z <- (ya_hfs_worry$worry_score - mean(ya_hfs_worry$worry_score,na.rm=T)) / sd(ya_hfs_worry$worry_score,na.rm=T)

# Child
# PAID Peds still needs 1 subtracted from each score (coded as 1-5 instead of 0-4)
child_paid <- data %>% select(record_id,redcap_event_name,c_paid1:c_paid20)
child_paid[,3:ncol(child_paid)] <- lapply(child_paid[,3:ncol(child_paid)],function(x) x-1)
# Fill NAs
child_paid <- fill_NAs(child_paid,0.75,1:2)
# Score
child_paid$paid_score <- apply(child_paid[,3:ncol(child_paid)],1,sum)*1.25

# HFS
child_hfs_behave <- data %>% select(record_id,redcap_event_name,c_hfs_behave1:c_hfs_behave10)
child_hfs_worry <- data %>% select(record_id,redcap_event_name,c_hfs_worry11:c_hfs_worry25)
# Fill NAs
child_hfs_behave <- fill_NAs(child_hfs_behave,0.75,1:2)
child_hfs_worry <- fill_NAs(child_hfs_worry,0.75,1:2)
# Score
child_hfs_behave$behave_score <- apply(child_hfs_behave[,3:ncol(child_hfs_behave)],1,sum)
child_hfs_worry$worry_score <- apply(child_hfs_worry[,3:ncol(child_hfs_worry)],1,sum)
# calculate Z-score
child_hfs_behave$behave_score_z <- (child_hfs_behave$behave_score - mean(child_hfs_behave$behave_score,na.rm=T)) / sd(child_hfs_behave$behave_score,na.rm=T)
child_hfs_worry$worry_score_z <- (child_hfs_worry$worry_score - mean(child_hfs_worry$worry_score,na.rm=T)) / sd(child_hfs_worry$worry_score,na.rm=T)

# Combine, put back into full dataframe
paid <- rbind(ya_paid[,c("record_id","redcap_event_name","paid_score")],
              child_paid[,c("record_id","redcap_event_name","paid_score")])
behave <- rbind(ya_hfs_behave[,c("record_id","redcap_event_name","behave_score","behave_score_z")],
                child_hfs_behave[,c("record_id","redcap_event_name","behave_score","behave_score_z")])
worry <- rbind(ya_hfs_worry[,c("record_id","redcap_event_name","worry_score","worry_score_z")],
               child_hfs_worry[,c("record_id","redcap_event_name","worry_score","worry_score_z")])
label(paid$redcap_event_name)="Event Name"
label(behave$redcap_event_name)="Event Name"
label(worry$redcap_event_name)="Event Name"

data <- left_join(data,paid,by = c("record_id","redcap_event_name"))
data <- left_join(data,behave,by = c("record_id","redcap_event_name"))
data <- left_join(data,worry,by = c("record_id","redcap_event_name"))

# Parents
# parent_surveys_b <- parent_full %>%
#   select(record_id,timepoint_baselinesurvey:p_hfs_worry26,p_paid1:p_paid18)
# colnames(parent_surveys_b)[2] <- "Timepoint"
# parent_surveys_1 <- parent_full %>%
#   select(record_id,timepoint_survey1:p_hfs_worry26_t1,p_paid1_t1:p_paid18_t1)
# colnames(parent_surveys_1) <- sub("_t1","",colnames(parent_surveys_1))
# colnames(parent_surveys_1)[2] <- "Timepoint"
# parent_surveys_2 <- parent_full %>%
#   select(record_id,timepoint_survey2:p_hfs_worry26_t2,p_paid1_t2:p_paid18_t2)
# colnames(parent_surveys_2) <- sub("_t2","",colnames(parent_surveys_2))
# colnames(parent_surveys_2)[2] <- "Timepoint"
# parent_surveys_3 <- parent_full %>%
#   select(record_id,timepoint_survey3:p_hfs_worry26_t3,p_paid1_t3:p_paid18_t3)
# colnames(parent_surveys_3) <- sub("_t3","",colnames(parent_surveys_3))
# colnames(parent_surveys_3)[2] <- "Timepoint"
# parent_surveys_4 <- parent_full %>%
#   select(record_id,timepoint_survey4:p_hfs_worry26_t4,p_paid1_t4:p_paid18_t4)
# colnames(parent_surveys_4) <- sub("_t4","",colnames(parent_surveys_4))
# colnames(parent_surveys_4)[2] <- "Timepoint"
# parent_surveys <- 
#   suppressWarnings(bind_rows(parent_surveys_b,parent_surveys_1,parent_surveys_2,
#                              parent_surveys_3,parent_surveys_4))
# colnames(parent_surveys) <- sub("p_","",colnames(parent_surveys))
# colnames(parent_surveys) <- sub("hfs_","",colnames(parent_surveys))
# parent_surveys <- parent_surveys %>%
#   filter(!(record_id %in% c("1a","74a"))) %>% # exclude 1a and 74a per Cari
#   arrange(as.numeric(record_id),Timepoint)
# Score
#parent_surveys[,3:ncol(parent_surveys)] <- lapply(parent_surveys[,3:ncol(parent_surveys)],function(x){as.numeric(x)-1})
# Split into PAID and HFS subscales
# parent_surveys_paid <- parent_surveys %>%
#   select(record_id,Timepoint,paid1:paid18)
# parent_surveys_behave <- parent_surveys %>%
#   select(record_id,Timepoint,behave1:behave11)
# parent_surveys_worry <- parent_surveys %>%
#   select(record_id,Timepoint,worry12:worry26)
# Fill NAs
# parent_surveys_paid$Timepoint <- as.character(parent_surveys_paid$Timepoint)
# parent_surveys_paid <- fill_NAs(parent_surveys_paid,0.75,1:2)
# parent_surveys_behave$Timepoint <- as.character(parent_surveys_behave$Timepoint)
# parent_surveys_behave <- fill_NAs(parent_surveys_behave,0.75,1:2)
# parent_surveys_worry$Timepoint <- as.character(parent_surveys_worry$Timepoint)
# parent_surveys_worry <- fill_NAs(parent_surveys_worry,0.75,1:2)
# Score
# PAID
# parent_surveys_paid$paid_score <- 
#   apply(parent_surveys_paid[3:ncol(parent_surveys_paid)],1,
#         function(x) (mean(4-x))*25)
# Behave
# parent_surveys_behave$behave_score <- 
#   apply(parent_surveys_behave[3:ncol(parent_surveys_behave)],1,sum)
# Worry
# parent_surveys_worry$worry_score <- 
#   apply(parent_surveys_worry[3:ncol(parent_surveys_worry)],1,sum)
# Put into one dataframe
# parent_surveys$Timepoint <- as.character(parent_surveys$Timepoint)
# parent_surveys <- left_join(parent_surveys,parent_surveys_paid,
#                   by = c("record_id","Timepoint"))
# parent_surveys <- left_join(parent_surveys,parent_surveys_behave,
#                   by = c("record_id","Timepoint"))
# parent_surveys <- left_join(parent_surveys,parent_surveys_worry,
#                   by = c("record_id","Timepoint"))
# parent_surveys <- parent_surveys %>% 
#   select(record_id,Timepoint,paid_score,behave_score,worry_score)
# parent_surveys <- parent_surveys[rowSums(is.na(parent_surveys)) < 4,]
# parent_surveys$total_score <- parent_surveys$behave_score + parent_surveys$worry_score
# Clean up workspace
#rm(list=setdiff(ls(),c("data","parent_surveys")))

# Subjects who withdrew (per Cari)
withdrawn <- c("41","50")

# remove those who withdrew
data <- subset(data, !(record_id %in% withdrawn))

# delete some extra visits
data <- data[!(data$record_id==17 & data$redcap_event_name=="baseline_arm_2"),]
data <- data[!(data$record_id==21 & data$redcap_event_name=="baseline_arm_2"),]
data <- data[!(data$record_id==8 & data$redcap_event_name=="baseline_arm_2"),]

# calculate change in scores
keep <- c("record_id","redcap_event_name","paid_score","behave_score","worry_score","behave_score_z","worry_score_z","timepoint")
times <- data[data$redcap_event_name %in% c("baseline_arm_1","3_arm_1","5_arm_1","baseline_arm_2","3_arm_2","5_arm_2"),keep]
times$visit[times$redcap_event_name %in% c("baseline_arm_1","baseline_arm_2")] <- "B"
times$visit[times$redcap_event_name %in% c("3_arm_1","3_arm_2")] <- "M6"
times$visit[times$redcap_event_name %in% c("5_arm_1","5_arm_2")] <- "M12"
# make a wide dataset
wide <- reshape(times,timevar="visit",idvar="record_id",direction="wide")
wide$delta_paid6 <- wide$paid_score.M6 - wide$paid_score.B
wide$delta_paid12 <- wide$paid_score.M12 - wide$paid_score.B
wide$delta_behavez_6 <- wide$behave_score_z.M6 - wide$behave_score_z.B
wide$delta_behavez_12 <- wide$behave_score_z.M12 - wide$behave_score_z.B
wide$delta_worryz_6 <- wide$worry_score_z.M6 - wide$worry_score_z.B
wide$delta_worryz_12 <- wide$worry_score_z.M12 - wide$worry_score_z.B

# create variable for change above/below median
wide$delta_paid6_abovemed <- ifelse(wide$delta_paid6>-3.75,1,0)
wide$delta_paid12_abovemed <- ifelse(wide$delta_paid12>-2.64,1,0)
wide$delta_behave6_abovemed <- ifelse(wide$delta_behavez_6>0,1,0)
wide$delta_behave12_abovemed <- ifelse(wide$delta_behavez_12>0,1,0)
wide$delta_worry6_abovemed <- ifelse(wide$delta_worryz_6>-0.11,1,0)
wide$delta_worry12_abovemed <- ifelse(wide$delta_worryz_12>-0.20,1,0)

# create variables for increase, decrease, or no change
wide$delta_paid6_cat <- ifelse(wide$delta_paid6>0,"Increase",ifelse(wide$delta_paid6<0,"Decrease","No change"))
wide$delta_paid12_cat <- ifelse(wide$delta_paid12>0,"Increase",ifelse(wide$delta_paid12<0,"Decrease","No change"))
wide$delta_behave6_cat <- ifelse(wide$delta_behavez_6>0,"Increase",ifelse(wide$delta_behavez_6<0,"Decrease","No change"))
wide$delta_behave12_cat <- ifelse(wide$delta_behavez_12>0,"Increase",ifelse(wide$delta_behavez_12<0,"Decrease","No change"))
wide$delta_worry6_cat <- ifelse(wide$delta_worryz_6>0,"Increase",ifelse(wide$delta_worryz_6<0,"Decrease","No change"))
wide$delta_worry12_cat <- ifelse(wide$delta_worryz_12>0,"Increase",ifelse(wide$delta_worryz_12<0,"Decrease","No change"))

# Demographics
demographics <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(automode_start),!(record_id %in% withdrawn)) %>%
  select(record_id,demographics_dob,demographics_diabetesdx,automode_start,hba1c_date_b,
         hba1c_baseline,demographics_sex.factor,demographics_race.factor,demographics_ethnicity.factor,
         demographics_insurance.factor,demographics_cgmhx.factor,demographics_pumphx.factor) %>%
  mutate(age_at_am_start = round(as.numeric((difftime(automode_start,demographics_dob,units = "days"))/365.25),2),
         t1d_duration_at_am_start = round(as.numeric((difftime(automode_start,demographics_diabetesdx,units = "days"))/365.25),2))
```

```{r table 1,echo=FALSE}
cont_vars <- c("hba1c_baseline","age_at_am_start","t1d_duration_at_am_start")
cat_vars <- c("demographics_sex.factor","demographics_race.factor","demographics_ethnicity.factor",
              "demographics_insurance.factor","demographics_cgmhx.factor","demographics_pumphx.factor")
nonnormal <- c("t1d_duration_at_am_start") # A1c always reported as mean (SD)
t1 <- CreateTableOne(vars = c(cont_vars,cat_vars),data = demographics)
t1 <- print(t1,nonnormal = nonnormal,printToggle = F)
#kable(t1)
```

## Summary statistics for 6 mo change in PAID score (6 mo - baseline)

```{r echo=FALSE, include=TRUE}
summary(wide$delta_paid6)
hist(wide$delta_paid6)
```

## Summary statistics for 12 mo change in PAID score (12 mo - baseline)

```{r echo=FALSE, include=TRUE}
summary(wide$delta_paid12)
hist(wide$delta_paid12)
```

## Summary statistics for 6 mo change in Behave Z-score (6 mo - baseline)
```{r echo=FALSE, include=TRUE}
summary(wide$delta_behavez_6)
hist(wide$delta_behavez_6)
```

## Summary statistics for 12 mo change in Behave Z-score (12 mo - baseline)
```{r echo=FALSE, include=TRUE}
summary(wide$delta_behavez_12)
hist(wide$delta_behavez_12)
```

## Summary statistics for 6 mo change in Worry Z-score (6 mo - baseline)
```{r echo=FALSE, include=TRUE}
summary(wide$delta_worryz_6)
hist(wide$delta_worryz_6)
```
 
## Summary statistics for 12 mo change in Worry Z-score (12 mo - baseline)
```{r echo=FALSE, include=TRUE}
summary(wide$delta_worryz_12)
hist(wide$delta_worryz_12)
``` 