---
title: "670G Obs. ATTD 2019 - Erin Cobry abstract"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "T:")
library(tableone)
library(redcapAPI)
library(knitr)
library(tidyverse)
library(magrittr)
library(dplyr)
```

```{r define functions,include=FALSE}
fill_NAs <- function(data,threshold,idcols) {
  na_thresh <- ceiling(threshold * ncol(data))-length(idcols) # Determine max NAs
  data <- data[rowSums(is.na(data)) < na_thresh,] # Remove rows with NAs over max
  k <- which(is.na(data), arr.ind=TRUE) # Find remaining NAs
  if(length(k)>0) {
    data[k] <- rowMeans(data[,colnames(data)[-idcols]],na.rm = T)[k[,1]] # Replace with means
  }
  data
}
```

```{r data import and formatting,echo=FALSE,include=FALSE}
# Import
knitr::opts_knit$set(root.dir = "T:")
source("T:/Erin Cobry/ATTD Abstract 2019/Data_Cleaned/AdvancedClosedLoopCl_R_2019-09-30_1049.r")
#source("T:\\Cari Berget\\ATTD Abstract 2019\\Data_Cleaned\\AdvancedClosedLoopCl_R_2019-09-30_1049.r")

#tokens <- read.csv("./api_tokens.csv")
#Parent670g <- as.character(tokens$Token[4])
#rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/", token = Parent670g)
#parent_full <- suppressWarnings(exportRecords(rcon))

# A1c as numeric
data$hba1c <- as.numeric(as.character(data$hba1c))
data$hba1c_baseline <- as.numeric(as.character(data$hba1c_baseline))

# Survey data
# Young adults
# PAID
ya_paid <- data %>% select(record_id,redcap_event_name,ya_paid1:ya_paid20)
ya_paid <- fill_NAs(ya_paid,0.75,1:2)
# Score
ya_paid$paid_score <- apply(ya_paid[,3:ncol(ya_paid)],1,sum)*1.25

# HFS scores (baseline and followup)
ya_hfs_b <- data %>% select(record_id,redcap_event_name,
                            ya_hfs_b_behave1:ya_hfs_b_behave15,
                            ya_hfs_b_worry16:ya_hfs_b_worry33) %>% 
  filter(redcap_event_name == "baseline_arm_2")
ya_hfs_f <- data %>% select(record_id,redcap_event_name,
                            ya_hfs_f_behave1:ya_hfs_f_behave15,
                            ya_hfs_f_worry16:ya_hfs_f_worry33) %>%
  filter(redcap_event_name != "baseline_arm_2")
# Change column names to match
colnames(ya_hfs_b) <- gsub("b_","",colnames(ya_hfs_b))
colnames(ya_hfs_f) <- gsub("f_","",colnames(ya_hfs_f))
# Bind
ya_hfs <- rbind(ya_hfs_b,ya_hfs_f)
# Split by HFS subscale, fill NAs
ya_hfs_behave <- ya_hfs %>% select(record_id,redcap_event_name,
                                   ya_hfs_behave1:ya_hfs_behave15)
ya_hfs_behave <- fill_NAs(ya_hfs_behave,0.75,1:2)
ya_hfs_worry <- ya_hfs %>% select(record_id,redcap_event_name,
                                   ya_hfs_worry16:ya_hfs_worry33)
ya_hfs_worry <- fill_NAs(ya_hfs_worry,0.75,1:2)
# Score
ya_hfs_behave$behave_score <- apply(ya_hfs_behave[,3:ncol(ya_hfs_behave)],1,sum)
ya_hfs_worry$worry_score <- apply(ya_hfs_worry[,3:ncol(ya_hfs_worry)],1,sum)
# calculate Z-score
ya_hfs_behave$behave_score_z <- (ya_hfs_behave$behave_score - mean(ya_hfs_behave$behave_score,na.rm=T)) / sd(ya_hfs_behave$behave_score,na.rm=T)
ya_hfs_worry$worry_score_z <- (ya_hfs_worry$worry_score - mean(ya_hfs_worry$worry_score,na.rm=T)) / sd(ya_hfs_worry$worry_score,na.rm=T)

# Child
# PAID Peds still needs 1 subtracted from each score (coded as 1-5 instead of 0-4)
child_paid <- data %>% select(record_id,redcap_event_name,c_paid1:c_paid20)
child_paid[,3:ncol(child_paid)] <- lapply(child_paid[,3:ncol(child_paid)],function(x) x-1)
# Fill NAs
child_paid <- fill_NAs(child_paid,0.75,1:2)
# Score
child_paid$paid_score <- apply(child_paid[,3:ncol(child_paid)],1,sum)*1.25

# HFS
child_hfs_behave <- data %>% select(record_id,redcap_event_name,c_hfs_behave1:c_hfs_behave10)
child_hfs_worry <- data %>% select(record_id,redcap_event_name,c_hfs_worry11:c_hfs_worry25)
# Fill NAs
child_hfs_behave <- fill_NAs(child_hfs_behave,0.75,1:2)
child_hfs_worry <- fill_NAs(child_hfs_worry,0.75,1:2)
# Score
child_hfs_behave$behave_score <- apply(child_hfs_behave[,3:ncol(child_hfs_behave)],1,sum)
child_hfs_worry$worry_score <- apply(child_hfs_worry[,3:ncol(child_hfs_worry)],1,sum)
# calculate Z-score
child_hfs_behave$behave_score_z <- (child_hfs_behave$behave_score - mean(child_hfs_behave$behave_score,na.rm=T)) / sd(child_hfs_behave$behave_score,na.rm=T)
child_hfs_worry$worry_score_z <- (child_hfs_worry$worry_score - mean(child_hfs_worry$worry_score,na.rm=T)) / sd(child_hfs_worry$worry_score,na.rm=T)

# Combine, put back into full dataframe
paid <- rbind(ya_paid[,c("record_id","redcap_event_name","paid_score")],
              child_paid[,c("record_id","redcap_event_name","paid_score")])
behave <- rbind(ya_hfs_behave[,c("record_id","redcap_event_name","behave_score","behave_score_z")],
                child_hfs_behave[,c("record_id","redcap_event_name","behave_score","behave_score_z")])
worry <- rbind(ya_hfs_worry[,c("record_id","redcap_event_name","worry_score","worry_score_z")],
               child_hfs_worry[,c("record_id","redcap_event_name","worry_score","worry_score_z")])
label(paid$redcap_event_name)="Event Name"
label(behave$redcap_event_name)="Event Name"
label(worry$redcap_event_name)="Event Name"

data <- left_join(data,paid,by = c("record_id","redcap_event_name"))
data <- left_join(data,behave,by = c("record_id","redcap_event_name"))
data <- left_join(data,worry,by = c("record_id","redcap_event_name"))

# Parents
# parent_surveys_b <- parent_full %>%
#   select(record_id,timepoint_baselinesurvey:p_hfs_worry26,p_paid1:p_paid18)
# colnames(parent_surveys_b)[2] <- "Timepoint"
# parent_surveys_1 <- parent_full %>%
#   select(record_id,timepoint_survey1:p_hfs_worry26_t1,p_paid1_t1:p_paid18_t1)
# colnames(parent_surveys_1) <- sub("_t1","",colnames(parent_surveys_1))
# colnames(parent_surveys_1)[2] <- "Timepoint"
# parent_surveys_2 <- parent_full %>%
#   select(record_id,timepoint_survey2:p_hfs_worry26_t2,p_paid1_t2:p_paid18_t2)
# colnames(parent_surveys_2) <- sub("_t2","",colnames(parent_surveys_2))
# colnames(parent_surveys_2)[2] <- "Timepoint"
# parent_surveys_3 <- parent_full %>%
#   select(record_id,timepoint_survey3:p_hfs_worry26_t3,p_paid1_t3:p_paid18_t3)
# colnames(parent_surveys_3) <- sub("_t3","",colnames(parent_surveys_3))
# colnames(parent_surveys_3)[2] <- "Timepoint"
# parent_surveys_4 <- parent_full %>%
#   select(record_id,timepoint_survey4:p_hfs_worry26_t4,p_paid1_t4:p_paid18_t4)
# colnames(parent_surveys_4) <- sub("_t4","",colnames(parent_surveys_4))
# colnames(parent_surveys_4)[2] <- "Timepoint"
# parent_surveys <- 
#   suppressWarnings(bind_rows(parent_surveys_b,parent_surveys_1,parent_surveys_2,
#                              parent_surveys_3,parent_surveys_4))
# colnames(parent_surveys) <- sub("p_","",colnames(parent_surveys))
# colnames(parent_surveys) <- sub("hfs_","",colnames(parent_surveys))
# parent_surveys <- parent_surveys %>%
#   filter(!(record_id %in% c("1a","74a"))) %>% # exclude 1a and 74a per Cari
#   arrange(as.numeric(record_id),Timepoint)
# Score
#parent_surveys[,3:ncol(parent_surveys)] <- lapply(parent_surveys[,3:ncol(parent_surveys)],function(x){as.numeric(x)-1})
# Split into PAID and HFS subscales
# parent_surveys_paid <- parent_surveys %>%
#   select(record_id,Timepoint,paid1:paid18)
# parent_surveys_behave <- parent_surveys %>%
#   select(record_id,Timepoint,behave1:behave11)
# parent_surveys_worry <- parent_surveys %>%
#   select(record_id,Timepoint,worry12:worry26)
# Fill NAs
# parent_surveys_paid$Timepoint <- as.character(parent_surveys_paid$Timepoint)
# parent_surveys_paid <- fill_NAs(parent_surveys_paid,0.75,1:2)
# parent_surveys_behave$Timepoint <- as.character(parent_surveys_behave$Timepoint)
# parent_surveys_behave <- fill_NAs(parent_surveys_behave,0.75,1:2)
# parent_surveys_worry$Timepoint <- as.character(parent_surveys_worry$Timepoint)
# parent_surveys_worry <- fill_NAs(parent_surveys_worry,0.75,1:2)
# Score
# PAID
# parent_surveys_paid$paid_score <- 
#   apply(parent_surveys_paid[3:ncol(parent_surveys_paid)],1,
#         function(x) (mean(4-x))*25)
# Behave
# parent_surveys_behave$behave_score <- 
#   apply(parent_surveys_behave[3:ncol(parent_surveys_behave)],1,sum)
# Worry
# parent_surveys_worry$worry_score <- 
#   apply(parent_surveys_worry[3:ncol(parent_surveys_worry)],1,sum)
# Put into one dataframe
# parent_surveys$Timepoint <- as.character(parent_surveys$Timepoint)
# parent_surveys <- left_join(parent_surveys,parent_surveys_paid,
#                   by = c("record_id","Timepoint"))
# parent_surveys <- left_join(parent_surveys,parent_surveys_behave,
#                   by = c("record_id","Timepoint"))
# parent_surveys <- left_join(parent_surveys,parent_surveys_worry,
#                   by = c("record_id","Timepoint"))
# parent_surveys <- parent_surveys %>% 
#   select(record_id,Timepoint,paid_score,behave_score,worry_score)
# parent_surveys <- parent_surveys[rowSums(is.na(parent_surveys)) < 4,]
# parent_surveys$total_score <- parent_surveys$behave_score + parent_surveys$worry_score
# Clean up workspace
#rm(list=setdiff(ls(),c("data","parent_surveys")))

# Subjects who withdrew (per Cari)
withdrawn <- c("41","50")

# remove those who withdrew
data <- subset(data, !(record_id %in% withdrawn))

# delete some extra visits
data <- data[!(data$record_id==17 & data$redcap_event_name=="baseline_arm_2"),]
data <- data[!(data$record_id==21 & data$redcap_event_name=="baseline_arm_2"),]
data <- data[!(data$record_id==8 & data$redcap_event_name=="baseline_arm_2"),]

# convert to numeric
data$time_am <- as.numeric(data$time_am)
data$sensor_70_180 <- as.numeric(data$sensor_70_180)
data$sensor_u54 <- as.numeric(data$sensor_u54)
data$sensor_54_69 <- as.numeric(data$sensor_54_69)

# calculate change in scores
keep <- c("record_id","redcap_event_name","paid_score","behave_score","worry_score","behave_score_z","worry_score_z","timepoint")
times <- data[data$redcap_event_name %in% c("baseline_arm_1","3_arm_1","5_arm_1","baseline_arm_2","3_arm_2","5_arm_2"),keep]
times$visit[times$redcap_event_name %in% c("baseline_arm_1","baseline_arm_2")] <- "B"
times$visit[times$redcap_event_name %in% c("3_arm_1","3_arm_2")] <- "M6"
times$visit[times$redcap_event_name %in% c("5_arm_1","5_arm_2")] <- "M12"
# make a wide dataset
wide <- reshape(times,timevar="visit",idvar="record_id",direction="wide")
wide$delta_paid6 <- wide$paid_score.M6 - wide$paid_score.B
wide$delta_paid12 <- wide$paid_score.M12 - wide$paid_score.B
wide$delta_behavez_6 <- wide$behave_score_z.M6 - wide$behave_score_z.B
wide$delta_behavez_12 <- wide$behave_score_z.M12 - wide$behave_score_z.B
wide$delta_worryz_6 <- wide$worry_score_z.M6 - wide$worry_score_z.B
wide$delta_worryz_12 <- wide$worry_score_z.M12 - wide$worry_score_z.B

# create variable for change above/below median
wide$delta_paid6_abovemed <- ifelse(wide$delta_paid6>-3.75,"Above median","At or below median")
wide$delta_paid12_abovemed <- ifelse(wide$delta_paid12>-2.64,"Above median","At or below median")
wide$delta_behave6_abovemed <- ifelse(wide$delta_behavez_6>0,"Above median","At or below median")
wide$delta_behave12_abovemed <- ifelse(wide$delta_behavez_12>0,"Above median","At or below median")
wide$delta_worry6_abovemed <- ifelse(wide$delta_worryz_6>-0.11,"Above median","At or below median")
wide$delta_worry12_abovemed <- ifelse(wide$delta_worryz_12>-0.20,"Above median","At or below median")

# create variables for increase, decrease, or no change
wide$delta_paid6_cat <- ifelse(wide$delta_paid6>0,"Increase",ifelse(wide$delta_paid6<0,"Decrease","No change"))
wide$delta_paid12_cat <- ifelse(wide$delta_paid12>0,"Increase",ifelse(wide$delta_paid12<0,"Decrease","No change"))
wide$delta_behave6_cat <- ifelse(wide$delta_behavez_6>0,"Increase",ifelse(wide$delta_behavez_6<0,"Decrease","No change"))
wide$delta_behave12_cat <- ifelse(wide$delta_behavez_12>0,"Increase",ifelse(wide$delta_behavez_12<0,"Decrease","No change"))
wide$delta_worry6_cat <- ifelse(wide$delta_worryz_6>0,"Increase",ifelse(wide$delta_worryz_6<0,"Decrease","No change"))
wide$delta_worry12_cat <- ifelse(wide$delta_worryz_12>0,"Increase",ifelse(wide$delta_worryz_12<0,"Decrease","No change"))

# merge the delta vars with baseline characteristics
# Baseline
baseline <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(automode_start),!(record_id %in% withdrawn)) %>%
  select(record_id,demographics_dob,demographics_diabetesdx,automode_start,hba1c_date_b,
         hba1c_baseline,demographics_sex.factor,demographics_race.factor,demographics_ethnicity.factor,
         demographics_insurance.factor,demographics_cgmhx.factor,demographics_pumphx.factor,time_am,
         sensor_u54,sensor_54_69,sensor_70_180) %>%
  mutate(age_at_am_start = round(as.numeric((difftime(automode_start,demographics_dob,units = "days"))/365.25),2),
         t1d_duration_at_am_start = round(as.numeric((difftime(automode_start,demographics_diabetesdx,units = "days"))/365.25),2))
final <- merge(baseline, wide, by="record_id")

# Table 1
cont_vars <- c("hba1c_baseline","age_at_am_start","t1d_duration_at_am_start")
cat_vars <- c("demographics_sex.factor","demographics_race.factor","demographics_ethnicity.factor",
              "demographics_insurance.factor","demographics_cgmhx.factor","demographics_pumphx.factor")
nonnormal <- c("t1d_duration_at_am_start") # A1c always reported as mean (SD)
exact <- c("demographics_pumphx.factor"," demographics_cgmhx.factor")
t1 <- CreateTableOne(vars = c(cont_vars,cat_vars),data = baseline)
t1 <- print(t1,nonnormal = nonnormal,printToggle = F)

# association between changes in survey scores and % AM
lm1 <- lm(delta_paid6 ~ time_am, data=final)
lm2 <- lm(delta_paid12 ~ time_am, data=final)
lm3 <- lm(delta_behavez_6 ~ time_am, data=final)
lm4 <- lm(delta_behavez_12 ~ time_am, data=final)
lm5 <- lm(delta_worryz_6 ~ time_am, data=final)
lm6 <- lm(delta_worryz_12 ~ time_am, data=final)

# compare those above and below the median
# vars are % AM, a1c, TIR, duration, age, hx pump use, hx cgm use, % time hypo (<54 and 54-69)
compvars <- c("time_am","hba1c_baseline","sensor_70_180","t1d_duration_at_am_start","age_at_am_start",
              "demographics_cgmhx.factor","demographics_pumphx.factor","sensor_u54","sensor_54_69")
t2 <- CreateTableOne(vars = compvars,data = final,strata = "delta_paid6_abovemed")
t2 <- print(t2,nonnormal = nonnormal,exact=exact,printToggle = F)
t3 <- CreateTableOne(vars = compvars,data = final,strata = "delta_paid12_abovemed")
t3 <- print(t3,nonnormal = nonnormal,exact=exact,printToggle = F)
t4 <- CreateTableOne(vars = compvars,data = final,strata = "delta_behave6_abovemed")
t4 <- print(t4,nonnormal = nonnormal,exact=exact,printToggle = F)
t5 <- CreateTableOne(vars = compvars,data = final,strata = "delta_behave12_abovemed")
t5 <- print(t5,nonnormal = nonnormal,exact=exact,printToggle = F)
t6 <- CreateTableOne(vars = compvars,data = final,strata = "delta_worry6_abovemed")
t6 <- print(t6,nonnormal = nonnormal,exact=exact,printToggle = F)
t7 <- CreateTableOne(vars = compvars,data = final,strata = "delta_worry12_abovemed")
t7 <- print(t7,nonnormal = nonnormal,exact=exact,printToggle = F)

# compare those who increased and decreased
t8 <- CreateTableOne(vars = compvars,data = final[final$delta_paid6_cat != "No change",],strata = "delta_paid6_cat")
t8 <- print(t8,nonnormal = nonnormal,exact=exact,printToggle = F)
t9 <- CreateTableOne(vars = compvars,data = final[final$delta_paid12_cat != "No change",],strata = "delta_paid12_cat")
t9 <- print(t9,nonnormal = nonnormal,exact=exact,printToggle = F)
t10 <- CreateTableOne(vars = compvars,data = final[final$delta_behave6_cat != "No change",],strata = "delta_behave6_cat")
t10 <- print(t10,nonnormal = nonnormal,exact=exact,printToggle = F)
t11 <- CreateTableOne(vars = compvars,data = final[final$delta_behave12_cat != "No change",],strata = "delta_behave12_cat")
t11 <- print(t11,nonnormal = nonnormal,exact=exact,printToggle = F)
t12 <- CreateTableOne(vars = compvars,data = final[final$delta_worry6_cat != "No change",],strata = "delta_worry6_cat")
t12 <- print(t12,nonnormal = nonnormal,exact=exact,printToggle = F)
t13 <- CreateTableOne(vars = compvars,data = final[final$delta_worry12_cat != "No change",],strata = "delta_worry12_cat")
t13 <- print(t13,nonnormal = nonnormal,exact=exact,printToggle = F)

# add descriptive statistics for PAID scores, AM time
t14 <- CreateTableOne(vars = c("paid_score.M6","paid_score.M12","time_am"),data = final)
t14 <- print(t14,nonnormal = nonnormal,exact=exact,printToggle = F)

```

## Introduction

The purpose of this analysis is to examine associations between changes in survey scores (PAID, FHS) and other measures in the observational 670g study.  

## Methods

The following data manipulation steps were performed:

1) Participants 41 and 50, who withdrew from the study, were excluded from analysis.
2) Only participants with a non-missing AM start value were included in analysis.
3) After scoring the surveys, the FHS worry and behavior scales were converted to Z-scores at each time point so child and young adult scores could be combined.
4) Participants were stratified as above or at or below the median change score for the PAID, worry, and behavior surveys at both 6 and 12 months, as well as by whether they had an increase or decrease in each of the surveys.

The distributions of all variables were examined prior to analysis.  Continuous variables were compared using t-tests (normally distributed) or the Kruskal-Wallis test.  Categorical variables were compared using chi-square or Fisher's exact test.  The associations between % of time in auto-mode and changes in survey scores were examined using linear models.


## Results

```{r echo=FALSE, include=TRUE}
kable(t1,caption="Table 1.  Demographics and clinical characteristics for entire sample.")
```

The association between the change in PAID score by 6 months and % AM time was not significant.

```{r echo=FALSE, include=TRUE}
summary(lm1)
```

The association between the change in PAID score by 12 months and % AM time was significant.

```{r echo=FALSE, include=TRUE}
summary(lm2)
```

The association between the change in behavior Z-score by 6 months and % AM time was not significant.

```{r echo=FALSE, include=TRUE}
summary(lm3)
```

The association between the change in behavior Z-score by 12 months and % AM time was not significant.

```{r echo=FALSE, include=TRUE}
summary(lm4)
```

The association between the change in worry Z-score by 6 months and % AM time was not significant.

```{r echo=FALSE, include=TRUE}
summary(lm5)
```

The association between the change in worry Z-score by 12 months and % AM time was not significant.

```{r echo=FALSE, include=TRUE}
summary(lm6)
```

```{r echo=FALSE, include=TRUE}
kable(t2,caption="Table 2.  Comparison of participants above the median vs. at or below the median in change in PAID score at 6 months.")
```

```{r echo=FALSE, include=TRUE}
kable(t3,caption="Table 3.  Comparison of participants above the median vs. at or below the median in change in PAID score at 12 months.")
```

```{r echo=FALSE, include=TRUE}
kable(t4,caption="Table 4.  Comparison of participants above the median vs. at or below the median in change in behavior Z-score at 6 months.")
```

```{r echo=FALSE, include=TRUE}
kable(t5,caption="Table 5.  Comparison of participants above the median vs. at or below the median in change in behavior Z-score at 12 months.")
```

```{r echo=FALSE, include=TRUE}
kable(t6,caption="Table 6.  Comparison of participants above the median vs. at or below the median in change in worry Z-score at 6 months.")
```

```{r echo=FALSE, include=TRUE}
kable(t7,caption="Table 7.  Comparison of participants above the median vs. at or below the median in change in worry Z-score at 12 months.")
```

```{r echo=FALSE, include=TRUE}
kable(t8, caption="Table 8. Comparison of participants with an increase vs. decrease in change in PAID score at 6 months. ")
```

```{r echo=FALSE, include=TRUE}
kable(t9, caption="Table 9. Comparison of participants with an increase vs. decrease in change in PAID score at 12 months. ")
```

```{r echo=FALSE, include=TRUE}
kable(t10, caption="Table 10. Comparison of participants with an increase vs. decrease in change in behavior Z-score at 6 months. ")
```

```{r echo=FALSE, include=TRUE}
kable(t11, caption="Table 11. Comparison of participants with an increase vs. decrease in change in behavior Z-score at 12 months. ")
```

```{r echo=FALSE, include=TRUE}
kable(t12, caption="Table 12. Comparison of participants with an increase vs. decrease in change in worry Z-score at 6 months. ")
```

```{r echo=FALSE, include=TRUE}
kable(t13, caption="Table 13. Comparison of participants with an increase vs. decrease in change in worry Z-score at 12 months. ")
```

```{r echo=FALSE, include=TRUE}
kable(t14, caption="Table 14. Descriptive statistics for PAID scores and AM time. ")
```







