---
title: "PICLS DSMB report 120522"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(stringr)
library(dplyr)
library(nlme)
library(emmeans)
library(sjPlot)
library(arsenal)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects/Sarit Polsky/PICLS"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
setwd(home_dir)
```

```{r data, include=FALSE}
# read CGM data
cgmdata <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/Cleaned CGM data/REDCap Upload.csv")
cgmdata$pid <- str_sub(cgmdata$subject_id,1,4)
cgmdata$time <- str_sub(cgmdata$subject_id,6,7)
cgmdata$time <- factor(cgmdata$time, c("T1","T2","T3","PP"))

# merge in randomization groups
rand <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")
rand_keep <- rand %>% select(pid,randomization_group)
cgmdata <- merge(cgmdata, rand_keep, by="pid", all.x = T, all.y = T)

# merge in baseline BMI
basebmi <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-BaselineBMI_DATA_2022-11-28_1039.csv")
cgmdata <- merge(cgmdata, basebmi, by="pid", all.x = T, all.y = F)

# remove 102A for now, because we don't have any CGM data
cgmdata <- cgmdata %>% filter(!pid=="102A")

# get baseline values
# CGM variables to test
# %time <54, %time 63-140, %time <63, % time >140 
base <- cgmdata %>% filter(time=="T1")
base$percent_time_under_54_t1 <- base$percent_time_under_54
base$percent_time_63_140_t1 <- base$percent_time_63_140
base$percent_time_under_63_t1 <- base$percent_time_under_63
base$percent_time_over_140_t1 <- base$percent_time_over_140
base$percent_time_over_180_t1 <- base$percent_time_over_180
base$mean_sensor_t1 <- base$mean_sensor
base$percent_time_under_70_t1 <- base$percent_time_under_70
base$percent_time_70_180_t1 <- base$percent_time_70_180
base$percent_time_over_250_t1 <- base$percent_time_over_250
base <- base %>% select(pid,percent_time_under_54_t1, percent_time_63_140_t1, percent_time_under_63_t1, 
                        percent_time_over_140_t1,percent_time_over_180_t1, mean_sensor_t1,
                        percent_time_under_70_t1,percent_time_70_180_t1,percent_time_over_250_t1)
cgmdata <- merge(cgmdata, base, by="pid", all.x = T, all.y = T)
cgmdata$clinic <- as.factor(ifelse(str_sub(cgmdata$pid,1,1)=="1","BDC","OSU"))
# delete PD data for per protocol analysis
cgmdata <- cgmdata %>% filter(!(pid=="111A" & time=="T2"))
cgmdata <- cgmdata %>% filter(!(pid=="111A" & time=="T3"))
cgmdata <- cgmdata %>% filter(!(pid=="114A" & time=="T2"))
cgmdata <- cgmdata %>% filter(!(pid=="114A" & time=="T3"))
cgmdata <- cgmdata %>% filter(!(pid=="117A" & time=="T2"))
cgmdata <- cgmdata %>% filter(!(pid=="117A" & time=="T3"))

# read in a1c data
a1cdata <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/AllLabs_VenousA1c.csv")
a1cdata$Lab.Collection.Trimester <- as.factor(a1cdata$Lab.Collection.Trimester)
a1cdata$pid <- a1cdata$ParticipantID
a1cdata <- a1cdata %>% select(pid,Lab.Collection.Trimester,Labvalue)
# read in preg visit A1c
a1credcap <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/QPregnancyVisits_DATA_2022-12-04_1234.csv")
a1credcap <- a1credcap %>% select(pid,Trimester,a1c_preg)
a1credcap$Lab.Collection.Trimester <- ifelse(a1credcap$Trimester==1,"1st",
                                             ifelse(a1credcap$Trimester==2,"2nd",
                                                    ifelse(a1credcap$Trimester==3,"3rd","Postpartum")))
a1credcap$Labvalue <- a1credcap$a1c_preg
a1credcap <- a1credcap %>% select(pid,Lab.Collection.Trimester,Labvalue)
a1cdata <- rbind(a1cdata,a1credcap)
# now take the average of any multiple A1c values within the same trimester
a1cdata_sum <- a1cdata %>% group_by(pid,Lab.Collection.Trimester) %>% summarise(mean_a1c = mean(Labvalue, na.rm=TRUE))
a1cdata_sum <- merge(a1cdata_sum, rand_keep, by="pid", all.x = F, all.y = T)
a1cdata_sum <- a1cdata_sum %>% select(pid, randomization_group,  Lab.Collection.Trimester, mean_a1c)
# get baseline values
base_a1c <- a1cdata_sum %>% filter(Lab.Collection.Trimester=="1st")
base_a1c$a1c_t1 <- base_a1c$mean_a1c
base_a1c <- base_a1c %>% select(pid, a1c_t1)
a1cdata_sum <- merge(a1cdata_sum, base_a1c, by="pid", all.x = T, all.y = T)
a1cdata_sum <- merge(a1cdata_sum, basebmi, by="pid", all.x = T, all.y = F)
a1cdata_sum$clinic <- as.factor(ifelse(str_sub(a1cdata_sum$pid,1,1)=="1","BDC","OSU"))

# HFS data
hfs <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-HFS_DATA_2022-11-28_1622.csv",
                na.strings = c("NA",""," "))
hfs <- hfs %>% filter(!is.na(fear_visit_date))
hfs$visit <- ifelse(hfs$redcap_event_name=="visit_1_baseline_arm_1","Baseline",
                    ifelse(hfs$redcap_event_name=="visit_10_3032_week_arm_1","Visit 10",
                           ifelse(hfs$redcap_event_name=="visit_16_46_wks_po_arm_1","Visit 16",
                                  ifelse(hfs$redcap_event_name=="visit_4_68_weeks_arm_1","Visit 4","Visit 7"))))
hfs$visit <- factor(hfs$visit, levels = c("Baseline","Visit 4","Visit 7","Visit 10","Visit 16"))
hfs$behavior <- rowSums(hfs[,c("fear_beh_bedtime_snacks","fear_beh_not_alone","fear_beh_finagle_test",
                               "fear_beh_high_when_alone","fear_beh_eat","fear_beh_reduce_meds",
                               "fear_beh_meeting","fear_beh_fast_sugar","fear_beh_no_exercise",
                               "fear_beh_check_often")])
hfs$worry <- rowSums((hfs[,c("fear_wor_unaware","fear_wor_no_food","fear_wor_dizzy","fear_wor_asleep",
                             "fear_wor_embarrass","fear_wor_reaction_alone","fear_wor_seem_drunk",
                             "fear_wor_losing_control","fear_wor_no_help","fear_wor_driving",
                             "fear_wor_work_mistake","fear_wor_work_eval","fear_wor_seizure",
                             "fear_caring_for_others","fear_long_term","fear_faint","fear_insulin_reaction")]))
hfs$total <- hfs$behavior + hfs$worry
# get baseline values
base_hfs <- hfs %>% filter(redcap_event_name=="visit_1_baseline_arm_1")
base_hfs$total_baseline <- base_hfs$total
base_hfs$worry_baseline <- base_hfs$worry
base_hfs$behavior_baseline <- base_hfs$behavior
base_hfs <- base_hfs %>% select(pid,total_baseline,worry_baseline,behavior_baseline)
hfs <- merge(hfs,base_hfs,by="pid",all.x = T, all.y = T)
# Take out visit 4
hfs <- hfs %>% filter(!visit=="Visit 4")
hfs$visit <- factor(hfs$visit)
hfs$clinic <- as.factor(ifelse(str_sub(hfs$pid,1,1)=="1","BDC","OSU"))

# merge in randomization group
hfs <- merge(hfs, rand_keep, by="pid", all.x = F, all.y = T)
```

# Results

This analysis is of the per protocol data set.

## Histogram of number of days of CGM wear per trimester

### Overall

```{r, echo=FALSE, message=FALSE}
hist(cgmdata$num_days)
```

### Trimester 1

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T1",]$num_days)
```

### Trimester 2

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T2",]$num_days)
```

### Trimester 3

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T3",]$num_days)
```

### Post-partum

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="PP",]$num_days)
```

## CGM descriptive statistics

### By trimester

```{r results='asis'}
t1_form = as.formula("time ~ percent_time_under_54 + percent_time_63_140 + percent_time_under_63 + percent_time_over_140 +
                     percent_time_over_180 + mean_sensor + percent_time_0_69 + percent_time_70_180 + percent_time_over_250")
t1 = tableby(t1_form,cgmdata)
summary(t1)
```

### By clinic

```{r results='asis'}
t2_form = as.formula("clinic ~ percent_time_under_54 + percent_time_63_140 + percent_time_under_63 + percent_time_over_140 +
                     percent_time_over_180 + mean_sensor + percent_time_0_69 + percent_time_70_180 + percent_time_over_250")
t2 = tableby(t2_form,cgmdata)
summary(t2)
```

## A1c descriptive statistics

### By trimester

```{r results='asis'}
t1_form = as.formula("Lab.Collection.Trimester ~ mean_a1c")
t1 = tableby(t1_form,a1cdata_sum)
summary(t1)
```

### By clinic

```{r results='asis'}
t2_form = as.formula("clinic ~ mean_a1c")
t2 = tableby(t2_form,a1cdata_sum)
summary(t2)
```


## Percent time <54 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_54 ~ time*randomization_group + clinic, random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent TIR 63-140 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_63_140 ~ time*randomization_group + clinic,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time <63 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_63 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time >140 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_140 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time >180

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_180 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Mean sensor glucose

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(mean_sensor ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time spent <70

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_70 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time spent 70-180

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_70_180 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time spent >250

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_250 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## HbA1c

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(mean_a1c ~ Lab.Collection.Trimester*randomization_group + clinic ,random=~1|pid,data = a1cdata_sum,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("Lab.Collection.Trimester","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("Lab.Collection.Trimester","randomization_group"))
```

## HFS total score

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(total ~ visit*randomization_group + clinic ,random=~1|pid,data = hfs,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("visit","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("visit","randomization_group"))
```

## HFS worry

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(worry ~ visit*randomization_group + clinic ,random=~1|pid,data = hfs,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("visit","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("visit","randomization_group"))
```

## HFS behavior

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(behavior ~ visit*randomization_group + clinic ,random=~1|pid,data = hfs,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("visit","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("visit","randomization_group"))
```
