---
title: "PICLS DSMB report 120522"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(stringr)
library(dplyr)
library(nlme)
library(emmeans)
library(sjPlot)
library(arsenal)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects/Sarit Polsky/PICLS"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
#knitr::opts_knit$set(root.dir = home_dir)
#setwd(home_dir)
```

```{r data, include=FALSE}
# read CGM data
cgmdata <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/Cleaned CGM data/REDCap Upload.csv")
cgmdata$pid <- str_sub(cgmdata$subject_id,1,4)
cgmdata$time <- str_sub(cgmdata$subject_id,6,7)
cgmdata$time <- factor(cgmdata$time, c("Ru","T1","T2","T3","PP"))

# merge in randomization groups
rand <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")
rand$randomization_group <- ifelse(rand$randomization_group==1,"HCL","SAPT")
rand_keep <- rand %>% select(pid,randomization_group)
cgmdata <- merge(cgmdata, rand_keep, by="pid", all.x = T, all.y = T)

# merge in baseline BMI
basebmi <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-BaselineBMI_DATA_2022-11-28_1039.csv")
cgmdata <- merge(cgmdata, basebmi, by="pid", all.x = T, all.y = F)

# remove 102A for now, because we don't have any CGM data
cgmdata <- cgmdata %>% filter(!pid=="102A")

# get baseline values
# CGM variables to test
# %time <54, %time 63-140, %time <63, % time >140 
base <- cgmdata %>% filter(time=="T1")
base$percent_time_under_54_t1 <- base$percent_time_under_54
base$percent_time_63_140_t1 <- base$percent_time_63_140
base$percent_time_under_63_t1 <- base$percent_time_under_63
base$percent_time_over_140_t1 <- base$percent_time_over_140
base$percent_time_over_180_t1 <- base$percent_time_over_180
base$mean_sensor_t1 <- base$mean_sensor
base$percent_time_under_70_t1 <- base$percent_time_under_70
base$percent_time_70_180_t1 <- base$percent_time_70_180
base$percent_time_over_250_t1 <- base$percent_time_over_250
base <- base %>% select(pid,percent_time_under_54_t1, percent_time_63_140_t1, percent_time_under_63_t1, 
                        percent_time_over_140_t1,percent_time_over_180_t1, mean_sensor_t1,
                        percent_time_under_70_t1,percent_time_70_180_t1,percent_time_over_250_t1)
cgmdata <- merge(cgmdata, base, by="pid", all.x = T, all.y = T)
cgmdata$clinic <- as.factor(ifelse(str_sub(cgmdata$pid,1,1)=="1","BDC","OSU"))

# read in a1c data
a1cdata <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/AllLabs_VenousA1c.csv")
a1cdata$Lab.Collection.Trimester <- as.factor(a1cdata$Lab.Collection.Trimester)
a1cdata$pid <- a1cdata$ParticipantID
a1cdata <- a1cdata %>% select(pid,Lab.Collection.Trimester,Labvalue)
# read in preg visit A1c
a1credcap <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/QPregnancyVisits_DATA_2022-12-04_1234.csv")
a1credcap <- a1credcap %>% select(pid,Trimester,a1c_preg)
a1credcap$Lab.Collection.Trimester <- ifelse(a1credcap$Trimester==1,"1st",
                                             ifelse(a1credcap$Trimester==2,"2nd",
                                                    ifelse(a1credcap$Trimester==3,"3rd","Postpartum")))
a1credcap$Labvalue <- a1credcap$a1c_preg
a1credcap <- a1credcap %>% select(pid,Lab.Collection.Trimester,Labvalue)
a1cdata <- rbind(a1cdata,a1credcap)
# now take the average of any multiple A1c values within the same trimester
a1cdata_sum <- a1cdata %>% group_by(pid,Lab.Collection.Trimester) %>% summarise(mean_a1c = mean(Labvalue, na.rm=TRUE))
a1cdata_sum <- merge(a1cdata_sum, rand_keep, by="pid", all.x = F, all.y = T)
a1cdata_sum <- a1cdata_sum %>% select(pid, randomization_group,  Lab.Collection.Trimester, mean_a1c)
# get baseline values
base_a1c <- a1cdata_sum %>% filter(Lab.Collection.Trimester=="1st")
base_a1c$a1c_t1 <- base_a1c$mean_a1c
base_a1c <- base_a1c %>% select(pid, a1c_t1)
a1cdata_sum <- merge(a1cdata_sum, base_a1c, by="pid", all.x = T, all.y = T)
a1cdata_sum <- merge(a1cdata_sum, basebmi, by="pid", all.x = T, all.y = F)
a1cdata_sum$clinic <- as.factor(ifelse(str_sub(a1cdata_sum$pid,1,1)=="1","BDC","OSU"))

###################
# create second dataset of HbA1c data using corrected POC values
a1ccorrected <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/Venous and POC A1cs_corrected POC.csv")
a1ccorrected$pid <- a1ccorrected$Study.ID
a1ccorrected$Labvalue <- a1ccorrected$corrected_A1C
a1ccorrected$Lab.Collection.Trimester <- ifelse(a1ccorrected$Trimester=="1","1st",
                                             ifelse(a1ccorrected$Trimester=="2","2nd",
                                                    ifelse(a1ccorrected$Trimester=="3","3rd","Postpartum")))
a1ccorrected$Lab.Collection.Trimester <- as.factor(a1ccorrected$Lab.Collection.Trimester)
a1ccorrected <- a1ccorrected %>% select(pid,Lab.Collection.Trimester,Labvalue)
# read in preg visit A1c
a1credcap <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/QPregnancyVisits_DATA_2022-12-04_1234.csv")
a1credcap <- a1credcap %>% select(pid,Trimester,a1c_preg)
a1credcap$Lab.Collection.Trimester <- ifelse(a1credcap$Trimester==1,"1st",
                                             ifelse(a1credcap$Trimester==2,"2nd",
                                                    ifelse(a1credcap$Trimester==3,"3rd","Postpartum")))
a1credcap$Labvalue <- a1credcap$a1c_preg
a1credcap <- a1credcap %>% select(pid,Lab.Collection.Trimester,Labvalue)
a1ccorrected <- rbind(a1ccorrected,a1credcap)
# now take the average of any multiple A1c values within the same trimester
a1ccorrected_sum <- a1ccorrected %>% group_by(pid,Lab.Collection.Trimester) %>% summarise(mean_a1c = mean(Labvalue, na.rm=TRUE))
a1ccorrected_sum <- merge(a1ccorrected_sum, rand_keep, by="pid", all.x = F, all.y = T)
a1ccorrected_sum <- a1ccorrected_sum %>% select(pid, randomization_group,  Lab.Collection.Trimester, mean_a1c)
# get baseline values
base_a1ccorrected <- a1ccorrected_sum %>% filter(Lab.Collection.Trimester=="1st")
base_a1ccorrected$a1c_t1 <- base_a1ccorrected$mean_a1c
base_a1ccorrected <- base_a1ccorrected %>% select(pid, a1c_t1)
a1ccorrected_sum <- merge(a1ccorrected_sum, base_a1ccorrected, by="pid", all.x = T, all.y = T)
a1ccorrected_sum <- merge(a1ccorrected_sum, basebmi, by="pid", all.x = T, all.y = F)
a1ccorrected_sum$clinic <- as.factor(ifelse(str_sub(a1ccorrected_sum$pid,1,1)=="1","BDC","OSU"))

###################


# HFS data
hfs <- read.csv("/Volumes/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-HFS_DATA_2022-11-28_1622.csv",
                na.strings = c("NA",""," "))
hfs <- hfs %>% filter(!is.na(fear_visit_date))
hfs$visit <- ifelse(hfs$redcap_event_name=="visit_1_baseline_arm_1","Baseline",
                    ifelse(hfs$redcap_event_name=="visit_10_3032_week_arm_1","Visit 10",
                           ifelse(hfs$redcap_event_name=="visit_16_46_wks_po_arm_1","Visit 16",
                                  ifelse(hfs$redcap_event_name=="visit_4_68_weeks_arm_1","Visit 4","Visit 7"))))
hfs$visit <- factor(hfs$visit, levels = c("Baseline","Visit 4","Visit 7","Visit 10","Visit 16"))
hfs$behavior <- rowSums(hfs[,c("fear_beh_bedtime_snacks","fear_beh_not_alone","fear_beh_finagle_test",
                               "fear_beh_high_when_alone","fear_beh_eat","fear_beh_reduce_meds",
                               "fear_beh_meeting","fear_beh_fast_sugar","fear_beh_no_exercise",
                               "fear_beh_check_often")])
hfs$worry <- rowSums((hfs[,c("fear_wor_unaware","fear_wor_no_food","fear_wor_dizzy","fear_wor_asleep",
                             "fear_wor_embarrass","fear_wor_reaction_alone","fear_wor_seem_drunk",
                             "fear_wor_losing_control","fear_wor_no_help","fear_wor_driving",
                             "fear_wor_work_mistake","fear_wor_work_eval","fear_wor_seizure",
                             "fear_caring_for_others","fear_long_term","fear_faint","fear_insulin_reaction")]))
hfs$total <- hfs$behavior + hfs$worry
# get baseline values
base_hfs <- hfs %>% filter(redcap_event_name=="visit_1_baseline_arm_1")
base_hfs$total_baseline <- base_hfs$total
base_hfs$worry_baseline <- base_hfs$worry
base_hfs$behavior_baseline <- base_hfs$behavior
base_hfs <- base_hfs %>% select(pid,total_baseline,worry_baseline,behavior_baseline)
hfs <- merge(hfs,base_hfs,by="pid",all.x = T, all.y = T)
# Take out visit 4
hfs <- hfs %>% filter(!visit=="Visit 4")
hfs$visit <- factor(hfs$visit)
hfs$clinic <- as.factor(ifelse(str_sub(hfs$pid,1,1)=="1","BDC","OSU"))

# merge in randomization group
hfs <- merge(hfs, rand_keep, by="pid", all.x = F, all.y = T)
```

# Results

This analysis is of the ITT data set. Changes to this version from the previous version of the report:

1) Added number of days of CGM data (variable num_days) to the table of CGM descriptive statistics (section 1.2). See also the histograms in section 1.1.

2) Added method description for the two versions of the HbA1c models (sections 1.13 and 1.14).

3) Added correlation between HbA1c and mean sensor glucose by trimester (section 1.15)

## Histogram of number of days of CGM wear per trimester

### Overall

```{r, echo=FALSE, message=FALSE}
hist(cgmdata$num_days)
```

### Trimester 1

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T1",]$num_days)
```

### Trimester 2

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T2",]$num_days)
```

### Trimester 3

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T3",]$num_days)
```

### Post-partum

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="PP",]$num_days)
```

## CGM descriptive statistics

### By trimester

```{r results='asis'}
t1_form = as.formula("time ~ percent_time_under_54 + percent_time_63_140 + percent_time_under_63 + percent_time_over_140 +
                     percent_time_over_180 + mean_sensor + percent_time_0_69 + percent_time_70_180 + percent_time_over_250 + num_days")
t1 = tableby(t1_form,cgmdata)
summary(t1)
```

### By clinic

```{r results='asis'}
t2_form = as.formula("clinic ~ percent_time_under_54 + percent_time_63_140 + percent_time_under_63 + percent_time_over_140 +
                     percent_time_over_180 + mean_sensor + percent_time_0_69 + percent_time_70_180 + percent_time_over_250 + num_days")
t2 = tableby(t2_form,cgmdata)
summary(t2)
```

## A1c descriptive statistics

### By trimester

```{r results='asis'}
t1_form = as.formula("Lab.Collection.Trimester ~ mean_a1c")
t1 = tableby(t1_form,a1cdata_sum)
summary(t1)
```

### By clinic

```{r results='asis'}
t2_form = as.formula("clinic ~ mean_a1c")
t2 = tableby(t2_form,a1cdata_sum)
summary(t2)
```


## Percent time <54 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_54 ~ time*randomization_group + clinic, random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent TIR 63-140 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_63_140 ~ time*randomization_group + clinic,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time <63 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_63 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time >140 mg/dL

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_140 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time >180

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_180 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Mean sensor glucose

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(mean_sensor ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time spent <70

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_70 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time spent 70-180

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_70_180 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time spent >250

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_250 ~ time*randomization_group + clinic ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## HbA1c - no correction

In this version of the model, POC and venous A1c values were used without making any type of adjustment for the method of measurement.

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(mean_a1c ~ Lab.Collection.Trimester*randomization_group + clinic ,random=~1|pid,data = a1cdata_sum,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("Lab.Collection.Trimester","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("Lab.Collection.Trimester","randomization_group"))
```

## HbA1c - using correction factor for POC HbA1c

There were 38 venous and POC A1cs from 15 participants that were within a week of each other. These values were used to develop a "correction factor" for POC A1c by regressing venous A1c on POC A1c. The parameters from this linear regression model were used to estimate the venous A1c that would have been observed based on the POC A1c. These corrected values were used in the analysis model instead of the original POC A1c.

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(mean_a1c ~ Lab.Collection.Trimester*randomization_group + clinic ,random=~1|pid,data = a1ccorrected_sum,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("Lab.Collection.Trimester","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("Lab.Collection.Trimester","randomization_group"))
```

## Correlation between HbA1c and mean glucose by trimester

### T1

```{r, echo=FALSE, message=FALSE, comment=""}
cgmdata$Lab.Collection.Trimester <- cgmdata$time
x <- cgmdata %>% select(pid, Lab.Collection.Trimester, mean_sensor)
y <- a1ccorrected_sum %>% select(pid, Lab.Collection.Trimester, mean_a1c)
y$Lab.Collection.Trimester <- ifelse(y$Lab.Collection.Trimester=="1st","T1",
                                     ifelse(y$Lab.Collection.Trimester=="2nd","T2",
                                            ifelse(y$Lab.Collection.Trimester=="3rd","T3","PP")))
combined_cgm_a1c <- merge(x, y, by=c("pid","Lab.Collection.Trimester"), all.x = TRUE, all.y = TRUE)
combined_cgm_a1c <- combined_cgm_a1c %>% select(mean_sensor, Lab.Collection.Trimester, mean_a1c)
combined_cgm_a1c_t1 <- combined_cgm_a1c %>% filter(Lab.Collection.Trimester=="T1")
combined_cgm_a1c_t2 <- combined_cgm_a1c %>% filter(Lab.Collection.Trimester=="T2")
combined_cgm_a1c_t3 <- combined_cgm_a1c %>% filter(Lab.Collection.Trimester=="T3")
combined_cgm_a1c_PP <- combined_cgm_a1c %>% filter(Lab.Collection.Trimester=="PP")

cor(combined_cgm_a1c_t1$mean_sensor,combined_cgm_a1c_t1$mean_a1c, use="pairwise.complete.obs")

plot(combined_cgm_a1c_t1$mean_sensor,combined_cgm_a1c_t1$mean_a1c, xlab="Mean Sensor Glucose", ylab="HbA1c")
```

### T2

```{r, echo=FALSE, message=FALSE, comment=""}
cor(combined_cgm_a1c_t2$mean_sensor,combined_cgm_a1c_t2$mean_a1c, use="pairwise.complete.obs")

plot(combined_cgm_a1c_t2$mean_sensor,combined_cgm_a1c_t2$mean_a1c, xlab="Mean Sensor Glucose", ylab="HbA1c")
```

### T3

```{r, echo=FALSE, message=FALSE, comment=""}
cor(combined_cgm_a1c_t3$mean_sensor,combined_cgm_a1c_t3$mean_a1c, use="pairwise.complete.obs")

plot(combined_cgm_a1c_t3$mean_sensor,combined_cgm_a1c_t3$mean_a1c, xlab="Mean Sensor Glucose", ylab="HbA1c")
```

### PP

```{r, echo=FALSE, message=FALSE, comment=""}
cor(combined_cgm_a1c_PP$mean_sensor,combined_cgm_a1c_PP$mean_a1c, use="pairwise.complete.obs")

plot(combined_cgm_a1c_PP$mean_sensor,combined_cgm_a1c_PP$mean_a1c, xlab="Mean Sensor Glucose", ylab="HbA1c")
```

## HFS total score

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(total ~ visit*randomization_group + clinic ,random=~1|pid,data = hfs,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("visit","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("visit","randomization_group"))
```

## HFS worry

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(worry ~ visit*randomization_group + clinic ,random=~1|pid,data = hfs,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("visit","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("visit","randomization_group"))
```

## HFS behavior

### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(behavior ~ visit*randomization_group + clinic ,random=~1|pid,data = hfs,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("visit","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("visit","randomization_group"))
```
