---
title: "PICLS DSMB report 120522"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(stringr)
library(dplyr)
library(nlme)
library(emmeans)
library(sjPlot)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects/Sarit Polsky/PICLS"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
setwd(home_dir)
```

```{r data, include=FALSE}
# read CGM data
cgmdata <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/Cleaned CGM data/REDCap Upload.csv")
cgmdata$pid <- str_sub(cgmdata$subject_id,1,4)
cgmdata$time <- str_sub(cgmdata$subject_id,6,7)
cgmdata$time <- factor(cgmdata$time, c("T1","T2","T3","PP"))

# merge in randomization groups
rand <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")
rand_keep <- rand %>% select(pid,randomization_group)
cgmdata <- merge(cgmdata, rand_keep, by="pid", all.x = T, all.y = T)

# merge in baseline BMI
basebmi <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-BaselineBMI_DATA_2022-11-28_1039.csv")
cgmdata <- merge(cgmdata, basebmi, by="pid", all.x = T, all.y = F)

# remove 102A for now, because we don't have any CGM data
cgmdata <- cgmdata %>% filter(!pid=="102A")

# get baseline values
# CGM variables to test
# %time <54, %time 63-140, %time <63, % time >140 
base <- cgmdata %>% filter(time=="T1")
base$percent_time_under_54_t1 <- base$percent_time_under_54
base$percent_time_63_140_t1 <- base$percent_time_63_140
base$percent_time_under_63_t1 <- base$percent_time_under_63
base$percent_time_over_140_t1 <- base$percent_time_over_140
base <- base %>% select(pid,percent_time_under_54_t1, percent_time_63_140_t1, percent_time_under_63_t1, percent_time_over_140_t1)
cgmdata <- merge(cgmdata, base, by="pid", all.x = T, all.y = T)

# read in a1c data
a1cdata <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/AllLabs_VenousA1c.csv")
a1cdata$Lab.Collection.Trimester <- as.factor(a1cdata$Lab.Collection.Trimester)
a1cdata$pid <- a1cdata$ParticipantID
# now take the average of any multiple A1c values within the same trimester
a1cdata_sum <- a1cdata %>% group_by(pid,Lab.Collection.Trimester) %>% summarise(mean_a1c = mean(Labvalue, na.rm=TRUE))
a1cdata_sum <- merge(a1cdata_sum, rand_keep, by="pid", all.x = F, all.y = T)
a1cdata_sum <- a1cdata_sum %>% select(pid, randomization_group,  Lab.Collection.Trimester, mean_a1c)
# get baseline values
base_a1c <- a1cdata_sum %>% filter(Lab.Collection.Trimester=="1st")
base_a1c$a1c_t1 <- base_a1c$mean_a1c
base_a1c <- base_a1c %>% select(pid, a1c_t1)
a1cdata_sum <- merge(a1cdata_sum, base_a1c, by="pid", all.x = T, all.y = T)
a1cdata_sum <- merge(a1cdata_sum, basebmi, by="pid", all.x = T, all.y = F)

# HFS data
hfs <- read.csv("B:/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-HFS_DATA_2022-11-28_1622.csv")
hfs$time <- ifelse(hfs$redcap_event_name=="visit_1_baseline_arm_1","T1",
                        ifelse(hfs$redcap_event_name %in% c("visit_8_2nd_trimes_arm_1","visit_9_2nd_trimes_arm_1"),"T2",
                               ifelse(hfs$redcap_event_name %in% c("visit_11_3rd_trime_arm_1","visit_12_3rd_trime_arm_1","visit_13_3rd_trime_arm_1"),"T3",
                                      ifelse(hfs$redcap_event_name=="visit_16_46_wks_po_arm_1","PP",NA))))
x <- hfs[,c("pid","redcap_event_name")]
write.csv(x,"C:/temp/visits_trimesters.csv", row.names = F)
```

# Results

## Histogram of number of days of CGM wear per trimester

### Overall

```{r, echo=FALSE, message=FALSE}
hist(cgmdata$num_days)
```

### Trimester 1

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T1",]$num_days)
```

### Trimester 2

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T2",]$num_days)
```

### Trimester 3

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="T3",]$num_days)
```

### Post-partum

```{r, echo=FALSE, message=FALSE}
hist(cgmdata[cgmdata$time=="PP",]$num_days)
```

## Percent time <54 mg/dL

### Without adjustment for baseline

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_54 ~ time*randomization_group,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

### With adjustment for baseline value and baseline BMI

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_54 ~ time*randomization_group + percent_time_under_54_t1 + bmi_baseline,
           random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent TIR 63-140 mg/dL

### Without adjustment for baseline

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_63_140 ~ time*randomization_group ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

### With adjustment for baseline value and baseline BMI

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_63_140 ~ time*randomization_group + percent_time_63_140_t1 + bmi_baseline,
           random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time <63 mg/dL

### Without adjustment for baseline

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_63 ~ time*randomization_group ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

### With adjustment for baseline value and baseline BMI

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_under_63 ~ time*randomization_group + percent_time_under_63_t1 + bmi_baseline,
           random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## Percent time >140 mg/dL

### Without adjustment for baseline 

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_140 ~ time*randomization_group ,random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

### With adjustment for baseline value and baseline BMI

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(percent_time_over_140 ~ time*randomization_group + percent_time_over_140_t1 + bmi_baseline,
           random=~1|pid,data = cgmdata,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("time","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("time","randomization_group"))
```

## HbA1c

### Without adjustment for baseline

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(mean_a1c ~ Lab.Collection.Trimester*randomization_group ,random=~1|pid,data = a1cdata_sum,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("Lab.Collection.Trimester","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("Lab.Collection.Trimester","randomization_group"))
```

### With adjustment for baseline value and baseline BMI

#### ANOVA table

```{r, echo=FALSE, message=FALSE}
mod <- lme(mean_a1c ~ Lab.Collection.Trimester*randomization_group + a1c_t1 + bmi_baseline,
           random=~1|pid,data = a1cdata_sum,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,c("Lab.Collection.Trimester","randomization_group"))
mod_pairs <-  pairs(mod_means,adjust="tukey", simple="each")

kable(mod_anova)
```

#### Model means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_means
```

#### Pairwise comparison of means

```{r, echo=FALSE, message=FALSE, comment=""}
mod_pairs
```

#### Plot

```{r, echo=FALSE, message=FALSE}
plot_model(mod, type="pred", terms=c("Lab.Collection.Trimester","randomization_group"))
```

