---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(readxl)
library(stringr)
library(lubridate)
library(nlme)
library(lme4)
library(lmerTest)
library(table1)


#data upload
Final_Baseline <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final Baseline.xlsx")
Final_2Tri <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final 2nd trimester.xlsx")
Final_3Tri <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final 3rd Trimester.xlsx")
Final_PP <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final Postpartum.xlsx")

mp_meal <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Missing Participants Export.xlsx", 
    sheet = "Meal Totals")
mp_daily <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Missing Participants Export.xlsx", 
    sheet = "Daily Totals")

PICLS <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/PICLSStudyHCLVsSAPTI_DATA_2024-05-24_1006(1).csv")
# groups and gest dates and centiles
gestdates <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data cleaning/CGM Data Check_6.8.23.csv") 
groups <- read.csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")%>% select(pid, randomization_group)
groups$randomization_group = factor(groups$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )

centiles = read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/BulkCentileCalc_Global_v8.0.6.1_PICLS_v5_CJ.xlsm", 
    sheet = "Import") %>% filter(!is.na(`Data ID`)) %>% mutate(pid = `Data ID`)


```


```{r dc, include = FALSE}
######################### ctrc data clean ######################################
baseline_ctrc = Final_Baseline %>% mutate(tp = "Baseline", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

tri2_ctrc = Final_2Tri %>% mutate(tp = "Trimester 2", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

tri3_ctrc = Final_3Tri %>% mutate(tp = "Trimester 3", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

pp_ctrc = Final_PP %>% mutate(tp = "Post Partum", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

mp_ctrc = mp_meal %>% mutate(pid_string = `Participant/Menu ID`) %>% select(pid_string,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

mp_ctrc = mp_ctrc %>% mutate(pid = stringr::word(pid_string, 1), tri = word(pid_string,2)) 
mp_ctrc = mp_ctrc %>% mutate(tp = case_when(tri == "3rd" ~ "Trimester 3",
                                            tri == "Base" ~ "Baseline",
                                            tri == "Post" ~ "Post Partum"))
mp_ctrc = mp_ctrc %>% select(-tri, -pid_string)


# merge
ctrc = full_join(baseline_ctrc, tri2_ctrc)
ctrc = full_join(ctrc, tri3_ctrc)
ctrc = full_join(ctrc, pp_ctrc)
ctrc = full_join(ctrc, mp_ctrc)
ctrc = right_join(groups, ctrc)

ctrc= ctrc %>% mutate(mufapufa = `% Calories from MUFA` + `% Calories from MUFA`)
gestdates = gestdates %>% select(pid, EDD_Final:`Delivery Date`)
ctrc2 = left_join(ctrc, gestdates)
ctrc2$date = ctrc2$`Date of Intake`
ctrc2 = ctrc2 %>% mutate(tp = case_when(date < mdy(SecondTri) ~ "Trimester 1",
                                        date >= mdy(SecondTri) & date < mdy(ThirdTri) ~ "Trimester 2",
                                        date >= mdy(ThirdTri) & date < mdy(`Delivery Date`) ~ "Trimester 3",
                                        date >= mdy(`Delivery Date`) ~ "Post Partum" ))
ctrc2$tp = factor(ctrc2$tp, levels = c("Trimester 1", "Trimester 2", "Trimester 3", "Post Partum"))

ctrc2 = ctrc2 %>% unique()

########################### MDIET   ############################################
maternal_diet_subj_day = ctrc2 %>% group_by(randomization_group, tp,pid, `Date of Intake`) %>% summarise(CarbTotal = sum(`Total Carbohydrate (g)`),
                                                                 Carb_Perc_Caltotal = mean(`% Calories from Carbohydrate`), # carbs
                                                                 FatTotal = sum(`Total Fat (g)`),
                                                                 Fat_Perc_Caltotal = mean(`% Calories from Fat`),
                                                                 SFA_Perc_Caltotal = mean(`% Calories from SFA`),
                                                                 MUFA_plus_PUFA_Perc_Caltotal = mean(mufapufa), #fats
                                                                 ProteinTotal = sum(`Total Protein (g)`),
                                                                 Protein_Perc_Caltotal = mean(`% Calories from Protein`),
                                                                 Animal_Protein_Perc = mean((`Animal Protein (g)`/`Total Protein (g)`)*100, na.rm=T),
                                                                 Plant_Protein_Perc = mean((`Vegetable Protein (g)`/`Total Protein (g)`)*100, na.rm=T), # protein
                                                                 FiberTotal = sum(`Total Dietary Fiber (g)`),
                                                                 GlyIndex_b = mean(`Glycemic Index (bread reference)`),
                                                                 GlyIndex_g = mean(`Glycemic Index (glucose reference)`),
                                                                 Cal_Total = sum(`Energy (kcal)`),
                                                                 Sfa_total = sum(`Total Saturated Fatty Acids (SFA) (g)`),
                                                                 Animal_total = sum(`Animal Protein (g)`),
                                                                 Plant_total = sum(`Vegetable Protein (g)`)
                                                                 )

maternal_diet_group = maternal_diet_subj_day %>% group_by(randomization_group, tp) %>% summarise_all(.funs = mean) %>% select(-pid) %>% ungroup() %>% select(-`Date of Intake`)
maternal_diet_group2 = maternal_diet_subj_day %>% group_by( tp) %>% summarise_all(.funs = mean) %>% select(-pid, -randomization_group) %>% ungroup() %>% select(-`Date of Intake`)


################################################################################

####################### Standardized Diets ##########################################

####### EDIT HERE

# standardized fiber intake averaged across the three trimesters, carb, fat and saturated fat, and total protein, animal and plant protein intake as a percentage of total calories
standardized_groups = maternal_diet_subj_day %>% group_by(pid, tp) %>% select(pid, tp, randomization_group, Cal_Total, FiberTotal,
                                                                              CarbTotal, FatTotal, Sfa_total, ProteinTotal, Animal_Protein_Perc, Plant_Protein_Perc, Animal_total, Plant_total) %>%
  mutate(s_fiber = FiberTotal*2000/Cal_Total,
         s_carb = CarbTotal*2000/Cal_Total,
         s_fat = FatTotal*2000/Cal_Total,
         s_sfat = Sfa_total*2000/Cal_Total,
         s_protein = ProteinTotal*2000/Cal_Total,
         s_animal_perc = (Animal_total*2000/Cal_Total)/s_protein*100,
         s_plant_perc = (Plant_total*2000/Cal_Total)/s_protein*100)

# ###################### Fiber Diet totals ############################################
# 
# fiber_groups = maternal_diet_subj_day %>% group_by(pid, tp) %>% select(pid, tp, randomization_group, FiberTotal) %>% summarise(avg_daily_fiber = mean(FiberTotal))
# fiber_groups = fiber_groups %>% mutate(Diet_Quality = case_when(avg_daily_fiber < 15 ~ "Poor/Low Quality",
#                                                                 avg_daily_fiber >= 15 & avg_daily_fiber < 28 ~ "Moderate Quality",
#                                                                 avg_daily_fiber>= 28 ~"High Quality"))
# 
# fiber_groups = left_join(fiber_groups, groups)
# fiber_groups$Diet_Quality = factor(fiber_groups$Diet_Quality, levels = c("Poor/Low Quality", "Moderate Quality","High Quality"))

################################################################################

###################### STdized Grouped into trimester avg #####################################

standardized_tri_avg = standardized_groups %>% group_by(pid, tp)  %>% summarise(avg_daily_fiber_s = mean(s_fiber),
                                                                           avg_daily_carb_s = mean(s_carb),
                                                                           avg_daily_fat_s = mean(s_fat),
                                                                           avg_daily_satfat_s = mean(s_sfat),
                                                                           avg_daily_protein_s = mean(s_protein),
                                                                           avg_daily_animal_perc_s = mean(s_animal_perc),
                                                                           avg_daily_plant_perc_s = mean(s_plant_perc))

# diet_groups = diet_groups %>% mutate(Fiber = ifelse(avg_daily_fiber >= 28, "High Quality","Not High Quality"),
#                                      Carb = ifelse(avg_daily_carb >= 175 & avg_daily_carb < 220, "High Quality","Not High Quality"),
#                                      Fat = ifelse(avg_daily_fat_perc <= 30, "High Quality","Not High Quality"),
#                                      SatFat = ifelse(avg_daily_satfat_perc < 10, "High Quality","Not High Quality"),
#                                      Overall = ifelse(Fiber=="High Quality"&Carb=="High Quality"&Fat=="High Quality"&SatFat=="High Quality", "High Quality", "Not High Quality"))
# diet_groups = left_join(diet_groups, groups)
################################################################################

###################### MATERNAL WEIGHT GAIN ############################################

cleanedcgm_noacet <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Raw/cleanedcgm_noacet.csv")
gmi_data = cleanedcgm_noacet %>% mutate(pid = word(subject_id,1,sep = "_"),tri = word(subject_id,2,sep = "_")) %>% select(pid, tri, gmi)
gmi_data = gmi_data %>% mutate(tp = case_when(tri == "T1" ~ "Trimester 1",
                                              tri == "T2" ~ "Trimester 2",
                                              tri == "T3" ~ "Trimester 3",
                                              tri == "PP" ~ "Post Partum"
                                              )) %>% filter(!is.na(tp)) %>% select(-tri)



Preg_visits <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/QPregnancyVisits_DATA_2022-12-04_1234.csv")

Preg_visits = left_join(Preg_visits, gestdates)

Preg_visits = Preg_visits %>% mutate(tp = case_when(mdy(date_preg) < mdy(SecondTri) ~ "Trimester 1",
                                        mdy(date_preg) >= mdy(SecondTri) & mdy(date_preg) < mdy(ThirdTri) ~ "Trimester 2",
                                        mdy(date_preg) >= mdy(ThirdTri) & mdy(date_preg) < mdy(`Delivery Date`) ~ "Trimester 3",
                                        mdy(date_preg) >= mdy(`Delivery Date`) ~ "Post Partum" ))
Preg_visits$tp = factor(Preg_visits$tp, levels = c("Trimester 1", "Trimester 2", "Trimester 3", "Post Partum"))
matweights = Preg_visits %>% select(pid, tp, weight_preg, date_preg,FirstTri,SecondTri,ThirdTri,`Delivery Date`, bmi_preg)

# CHANGING TYPO FROM 68KG to 99 KG
matweights[141, 3] = 99.0

precons = PICLS %>% filter(redcap_event_name =="preconception_arm_1")%>% select(pid, weight_precon)

all_weights = left_join(matweights, precons) %>% filter(!is.na(tp))
last_weights = all_weights %>% group_by(pid, tp) %>% filter(!is.na(weight_preg)) %>% mutate(last_weight = ifelse(date_preg == max(date_preg), 1, 0)) %>% filter(last_weight == 1) %>% ungroup()

# avg weight gain by trimester, then converted to pounds to classify as excessive weight gain
last_weights = last_weights %>% mutate(weight_gain_total = weight_preg - weight_precon,
                                       week_length = as.numeric(difftime(mdy(date_preg), mdy(FirstTri), units="weeks")),
                                       avg_weekly_weight_gain = weight_gain_total/week_length,
                                       avg_weekly_weight_gain_lbs = avg_weekly_weight_gain*2.20462,
                                       weight_gain_total_lbs = weight_gain_total*2.20462)

last_weights = last_weights %>% mutate(excessive_weight_gain_wk = case_when(bmi_preg < 25 ~ ifelse(avg_weekly_weight_gain_lbs > 0.92, 1, 0),
                                                                         bmi_preg < 30 & bmi_preg >=25 ~ ifelse(avg_weekly_weight_gain_lbs > 0.66, 1, 0),
                                                                         bmi_preg >= 30  ~ ifelse(avg_weekly_weight_gain_lbs > 0.52, 1, 0)),
                                       excessive_weight_gain_total_tri = case_when(tp == "Trimester 3" & bmi_preg < 25~ ifelse(weight_gain_total_lbs > 35, 1,0),
                                                                               tp == "Trimester 3" & bmi_preg < 30 & bmi_preg >=25~ ifelse(weight_gain_total_lbs > 25, 1,0),
                                                                               tp == "Trimester 3" & bmi_preg >= 30~ ifelse(weight_gain_total_lbs > 20, 1,0)))

last_weights_tri = last_weights %>% filter(tp != "Post Partum") %>% group_by(pid) %>%fill(excessive_weight_gain_total_tri, .direction="updown")
last_weights_tri = last_weights_tri %>% mutate(excessive_weight_gain = ifelse(excessive_weight_gain_wk == 1 |tp == "Trimester 3" & excessive_weight_gain_total_tri == 1, 1, 0))
# post partum
last_weights_pp = last_weights %>% filter(tp == "Post Partum") %>% mutate(excessive_weight_gain_total_pp = case_when(bmi_preg < 25~ ifelse(weight_gain_total_lbs > 35, 1,0),
                                                                                bmi_preg < 30 & bmi_preg >=25~ ifelse(weight_gain_total_lbs > 25, 1,0),
                                                                                bmi_preg >= 30~ ifelse(weight_gain_total_lbs > 20, 1,0)))
last_weights_pp = last_weights_pp %>% mutate(excessive_weight_gain = ifelse(excessive_weight_gain_wk == 1 |excessive_weight_gain_total_pp == 1, 1, 0))

# tables
excessive_weight_tbl = last_weights_tri %>% group_by(tp) %>% summarise(count = sum(excessive_weight_gain, na.rm=T), percent = mean(excessive_weight_gain, na.rm=T)*100)
excessive_weight_tbl_pp = last_weights_pp %>% group_by(tp) %>% summarise(count = sum(excessive_weight_gain, na.rm=T), percent = mean(excessive_weight_gain, na.rm=T)*100)

# analysis for weekly mwg 
maternal_weight_gain_tri = last_weights_tri %>% select(pid, tp, excessive_weight_gain, avg_weekly_weight_gain_lbs, weight_gain_total_lbs)
avg_mwg_tri = maternal_weight_gain_tri %>% group_by(pid) %>% summarise(weekly_weight_gain_avg = mean(avg_weekly_weight_gain_lbs, na.rm =T))
###################### ANALYSIS SET FOR AIM 1 #####################################
mwg_total = full_join(maternal_weight_gain_tri, last_weights_pp %>% select(pid, tp, excessive_weight_gain, avg_weekly_weight_gain_lbs, weight_gain_total_lbs)) %>% unique()

analysis_aim1_tri = left_join(standardized_tri_avg, mwg_total) %>% unique()

#### Diet quality

analysis_aim1_tri = analysis_aim1_tri %>% mutate(s_fiber_group = case_when(avg_daily_fiber_s < 15 ~ "Low Quality",
                                                                         avg_daily_fiber_s >= 15 & avg_daily_fiber_s < 28 ~ "Medium Quality",
                                                                         avg_daily_fiber_s >= 28 ~ "High Quality"),
                                                 s_Fiber = case_when(avg_daily_fiber_s < 28 ~ "Not High Quality",
                                                                     avg_daily_fiber_s >= 28 ~ "High Quality"),
                                                 s_Carb = case_when(avg_daily_carb_s >= 175 & avg_daily_carb_s < 220 ~ "High Quality",
                                                                    avg_daily_carb_s < 175 | avg_daily_carb_s >= 220 ~ "Not High Quality"),
                                                 s_Fat = case_when(avg_daily_fat_s <= 30 ~ "High Quality",
                                                                   avg_daily_fat_s > 30 ~ "Not High Quality"),
                                                 s_SatFat = case_when(avg_daily_satfat_s < 10 ~ "High Quality",
                                                                      avg_daily_satfat_s >= 10 ~ "Not High Quality"),
                                                 s_Overall = ifelse(s_Fiber=="High Quality"& s_Carb=="High Quality"&s_Fat=="High Quality"&s_SatFat=="High Quality", "High Quality", "Not High Quality"))

analysis_aim1_tri$s_fiber_group = factor(analysis_aim1_tri$s_fiber_group, levels = c("Low Quality", "Medium Quality", "High Quality"))

analysis_m1_tri = analysis_aim1_tri %>% filter(tp != "Post Partum")
analysis_m1 = analysis_m1_tri %>% ungroup() %>% group_by(pid) %>% select(avg_daily_fiber_s:avg_daily_plant_perc_s, avg_weekly_weight_gain_lbs) %>% 
  summarise(fiber_s = mean(avg_daily_fiber_s),
            carb_s = mean(avg_daily_carb_s),
            fat_s = mean(avg_daily_fat_s),
            sfat_s = mean(avg_daily_satfat_s),
            protein_s = mean(avg_daily_protein_s),
            animal_perc_s = mean(avg_daily_animal_perc_s),
            plant_perc_s = mean(avg_daily_plant_perc_s),
            avg_wk_weight_gain_lbs = mean(avg_weekly_weight_gain_lbs, na.rm=T))

adjust = last_weights_tri %>% filter(tp == "Trimester 3") %>% select(pid, week_length)
analysis_m1 = left_join(analysis_m1, adjust)
analysis_m1[analysis_m1$pid == "110A", "week_length"] = 14.428571

##############################
base_bmis = PICLS %>% select(pid, bmi_baseline) %>% filter(!is.na(bmi_baseline))
analysis_m1_tri = left_join(analysis_m1_tri, base_bmis)

#analysis_m1_pp = analysis_aim1_tri %>% filter(tp == "Post Partum")
sg_pp = standardized_groups %>% filter(tp == "Post Partum") %>% group_by(pid, tp)  %>% summarise(avg_daily_fiber = mean(FiberTotal),
                                                                           avg_daily_carb = mean(CarbTotal),
                                                                           avg_daily_fat = mean(FatTotal),
                                                                           avg_daily_satfat = mean(Sfa_total),
                                                                           avg_daily_protein = mean(ProteinTotal),
                                                                           avg_daily_animal_perc = mean(Animal_Protein_Perc),
                                                                           avg_daily_plant_perc = mean(Plant_Protein_Perc))
analysis_m1_pp = left_join(mwg_total %>% filter(tp == "Post Partum"), sg_pp ) %>% filter(!is.na(avg_daily_fiber))




# Fiber > 28 
# 175 <= Carb < 220
# Fat <=30 
# SatFat <10  (n = 6 for both satfat/fat joint)
```

```{r function, include = F}
# fit model fun
fit_mod = function(assoc,df){
  
  # Fit model
  print(assoc)  
  f = as.formula(paste0("avg_wk_weight_gain_lbs","~",assoc, "+ week_length"))
  mod = lm(f, data = df, na.action = na.omit)
  print(summary(mod))
  ggplot(data = df, aes_string(assoc, "avg_wk_weight_gain_lbs")) + geom_point() + theme_classic()
 
}

fit_lr_mod = function(assoc, df, interaction=FALSE){
  if(interaction==FALSE){
  print(assoc)  
  f = as.formula(paste0("excessive_weight_gain","~",assoc, "+ tp + bmi_baseline + (1|pid)"))
  mod = glmer(f, df, family = binomial)
  summary(mod)
  }
  else{
  print(assoc)  
  f = as.formula(paste0("excessive_weight_gain","~",assoc, "*bmi_baseline + tp  + (1|pid)"))
  mod = glmer(f, df, family = binomial)
  summary(mod)
  }
}

fit_mod_pp = function(assoc,df){
  
  # Fit model
  print(assoc)  
  f = as.formula(paste0("weight_gain_total_lbs","~",assoc))
  mod = lm(f, data = df, na.action = na.omit)
  print(summary(mod))
  ggplot(data = df, aes_string(assoc, "weight_gain_total_lbs")) + geom_point() + theme_classic()
 
}

```

# Outstanding Data questions/remarks

# Methods

Subjects maternal weight gain was measured using the latest date in a trimester compared with the preconception weight. This was then averaged into a weekly weight gain over all trimesters.

Excessive weight gain(either by weekly weight gain or total weight gain) was derived, and was stratified by timepoint.

Linear Models for Maternal Weight Gain were fit with standardized fiber intake, carb, fat and saturated fat, and total protein, animal and plant protein intake as a percentage of total calories averaged across the three trimesters. Standardization was based on a 2000 calorie diet and averaged first over trimesters, then overall. Avg Weekly Maternal Weight gain was measured as the average weekly weight gain across trimesters then averaged all together. Linear Models were adjusted for weeks of gestation between the first and last weight measurement used.

Generalized Linear Mixed Models with a binomial link function (Mixed Effects Logistic Regression) for Excessive Weight Gain were fit adjusting for enrollment maternal BMI. Similar Models were fit including an interaction term between the dietary independent variable and Maternal BMI at enrollment

Linear Models for Post Partum Maternal Weight Retention were fit with Average Daily fiber intake, carb, fat and saturated fat, and total protein, animal and plant protein intake as a percentage of total calories. Post Partum Weight retention is defined  as the difference between post partum weight and first pregnancy weight.

# Appended 8/27/2024

## Standardized Diet Quality (Overall and Fiber)

### Overall Criteria

n = 15 for high quality standardized fiber (4 more than non standardized), n = 19 for high quality standardized carbs (3 more than non standardized)

note: This might be driven by extrapolation/higher fat lower cal meals that met criteria when averaged, but when standardized caused to switch groups (in previous analysis we had ~ 20 and 30 percent meeting high quality for fat and satfat)

```{r}
table1(~s_Fiber + s_Carb + s_Fat + s_SatFat + s_Overall|tp, data = analysis_aim1_tri)
```

### Fiber Criteria

```{r}
table1(~s_fiber_group|tp, data = analysis_aim1_tri)
```

# Analysis

## Excessive Weight Gain

### Excessive Weight Gain table

Excessive weight gain is defines as avg weight gain of 0.92lbs/wk, 0.66lbs/wk, or 0.52lbs/wk for Normal, Overweight, and Obese Maternal BMI,
Or a total weight gain of 35, 25, or 20 lbs respectively.

Below is the excessive weight gain distribution by trimester.

```{r, results='asis'}
# trimester excessive weight gain counts
kable(excessive_weight_tbl)
# post-partum
kable(excessive_weight_tbl_pp)
```

### Mixed Logistic Regression Models

Standardized Fiber, Carb, Fat, SatFat, Protein, and Percentage of Animal/Plant Protein were not found to be associated with excessive weight gain.

#### Fiber

fiber group not associated with EWG

```{r}
fit_lr_mod("avg_daily_fiber_s", analysis_m1_tri, F)

fit_lr_mod("s_fiber_group", analysis_m1_tri, F)
```

#### Carb

```{r}
fit_lr_mod("avg_daily_carb_s", analysis_m1_tri, F)
```

#### Fat

```{r}
fit_lr_mod("avg_daily_fat_s", analysis_m1_tri, F)
```

#### Sat Fat

```{r}
fit_lr_mod("avg_daily_satfat_s", analysis_m1_tri, F)
```

#### Protein

```{r}
fit_lr_mod("avg_daily_protein_s", analysis_m1_tri, F)
```

#### Animal Protein

```{r}
fit_lr_mod("avg_daily_animal_perc_s", analysis_m1_tri, F)
```

#### Plant Protein

```{r}
fit_lr_mod("avg_daily_plant_perc_s", analysis_m1_tri, F)
```

### Mixed Logistic Regression Models with interaction term

Most models had trouble converging including the interaction term. I would interpret terms here with a grain of salt.

#### Fiber

```{r}
fit_lr_mod("avg_daily_fiber_s", analysis_m1_tri, T)
```

#### Carb

```{r}
fit_lr_mod("avg_daily_carb_s", analysis_m1_tri, T)
```

#### Fat

```{r}
fit_lr_mod("avg_daily_fat_s", analysis_m1_tri, T)
```

#### Sat Fat

```{r}
fit_lr_mod("avg_daily_satfat_s", analysis_m1_tri, T)
```

#### Protein

```{r}
fit_lr_mod("avg_daily_protein_s", analysis_m1_tri, T)
```

#### Animal Protein

```{r}
fit_lr_mod("avg_daily_animal_perc_s", analysis_m1_tri, T)
```

#### Plant Protein

```{r}
fit_lr_mod("avg_daily_plant_perc_s", analysis_m1_tri, T)
```

## Maternal Weight Gain

Standardized Fiber, Carb, Fat, SatFat, Protein, and Percentage of Animal/Plant Protein were not found to be associated with maternal weight gain.

### Linear Models

#### Fiber

```{r}
fit_mod("fiber_s", analysis_m1)
```

#### Carb

```{r}
fit_mod("carb_s", analysis_m1)
```

#### Fat

```{r}
fit_mod("fat_s", analysis_m1)
```

#### Sat Fat

```{r}
fit_mod("sfat_s", analysis_m1)
```

#### Protein

```{r}
fit_mod("protein_s", analysis_m1)
```

#### Animal Protein

```{r}
fit_mod("animal_perc_s", analysis_m1)
```

#### Plant Protein

```{r}
fit_mod("plant_perc_s", analysis_m1)
```

## Post Partum Weight Retention

### Linear Models

Standardized Fiber, Carb, Fat, SatFat, Protein, and Percentage of Animal/Plant Protein were not found to be associated with Post Partum Maternal Weight retention.

#### Fiber

```{r}
fit_mod_pp("avg_daily_fiber", analysis_m1_pp)
```

#### Carb

```{r}
fit_mod_pp("avg_daily_carb", analysis_m1_pp)
```

#### Fat

```{r}
fit_mod_pp("avg_daily_fat", analysis_m1_pp)
```

#### Sat Fat

```{r}
fit_mod_pp("avg_daily_satfat", analysis_m1_pp)
```

#### Protein

```{r}
fit_mod_pp("avg_daily_protein", analysis_m1_pp)
```

#### Animal Protein

```{r}
fit_mod_pp("avg_daily_animal_perc", analysis_m1_pp)
```

#### Plant Protein

```{r}
fit_mod_pp("avg_daily_plant_perc", analysis_m1_pp)
```