---
title: "PICLS_cs"
author: "Casey Sakamoto"
date: "2023-11-13"
output: html_document
---

```{r setup, include = FALSE}
library(knitr)
library(stringr)
library(tidyverse)
library(nlme)
library(emmeans)
library(readxl)
library(openxlsx)
library(lubridate)
library(data.table)

#data upl  S:/Laura/BDC\Projects\Sarit Polsky\PICLS
######## FIRST SHEET. TDD STUFF
tdd_files_list <- list.files(path = 'S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_tdd_dfs <- lapply(as.list(tdd_files_list), function(x) read_xlsx(x, sheet = 1))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
tdd_df_names <- paste0(str_extract(basename(tdd_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_tdd_dfs) <- tdd_df_names

########## SECIND SHEET PUMP ADJ
pump_files_list <- list.files(path = 'S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_pumpdfs <- lapply(as.list(pump_files_list), function(x) read_xlsx(x, sheet = 2, col_types = c("text", "date", "text", "numeric", "numeric", "numeric", "numeric", "numeric")))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
pump_df_names <- paste0(str_extract(basename(pump_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_pumpdfs) <- pump_df_names
```

```{r data clean, include=FALSE}
# tdd
varnames = c("date", "gweek", "gday", "tdd_u", "td_bolus", "td_basal", "bolus_perc", "basal_perc", "sampleid")

column_namer = function(x){
  df = x
  colnames(df) = varnames
  df
}

# want to add id and combine datasets
list_of_tdd_dfs = mapply(cbind, list_of_tdd_dfs, "SampleID"=tdd_df_names, SIMPLIFY=F)
#list_of_tdd_dfs[[2]]$`Gestational Week` = as.Date(list_of_tdd_dfs[[2]]$`Gestational Week`)
list_of_tdd_dfs[[22]] = list_of_tdd_dfs[[22]] %>% select(-Index)

list_of_tdd_dfs = lapply(list_of_tdd_dfs, column_namer)


tdd_df = do.call(rbind, list_of_tdd_dfs)
tdd_df = tdd_df %>% filter(!is.na(gweek))

# pump settings   ,not sure if we need"Date of pump adjustment"
pumpvarnames = c("Gestational age","Date of pump adjustment","Start Time","Basal units/hour","Carb ratio (g/U)","Correction bolus","Total basal (U)","Active insulin time (hours)" )

column_namer_p = function(x){
  df = x
  df = df %>% select(pumpvarnames)

  return(df)
}


list_of_pumpdfs = lapply(list_of_pumpdfs, column_namer_p)
list_of_pumpdfs = mapply(cbind, list_of_pumpdfs, "sampleid"=pump_df_names, SIMPLIFY=F)

pump_df = do.call(rbind, list_of_pumpdfs)


## get into analysis/table form
pump_adj_df = pump_df %>% mutate(adj_count = ifelse(!is.na(`Date of pump adjustment`), 1, 0))
pump_adj_df$gweek = factor(tstrsplit(pump_adj_df$`Gestational age`, "w")[[1]], levels = c("Preconception", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
                                                                                             "11", "12", "13", "14", "15", "16", "17", "18", "19", "20",
                                                                                             "21", "22", "23", "24", "25", "26", "27", "28", "29", "30",
                                                                                             "31", "32", "33", "34", "35", "36", "37", "38", "Postpartum"))
pump_adj_df = pump_adj_df %>% filter(!is.na(gweek))
test = pump_adj_df %>% group_by(gweek) %>% summarise(sum = sum(adj_count))

## add in wks order for the desc tables, and months for the model

## model function will be outcome = int + group + time + groupTime

```

We will plan to analyze the number of pump adjustments made per gestational week and the average basal rates, carb ratios, and correction factors for each meal period (breakfast, lunch, evening, overnight) for each time period in pregnancy (4-8 weeks gestation, etc.).

these are the pump settings we're looking at: total daily dose, total daily bolus, total daily basal, and percent basal/bolus

```{r pump data}


```
