---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(readxl)
library(stringr)
library(lubridate)
library(nlme)
library(lme4)
library(lmerTest)
library(table1)

rename <- function(df) {
  df$"Omega-6 Fatty Acids (g)"  <- NA
  colnames(df[,c("PUFA 18:2 (linoleic acid) (g)")]) <- 
    c("PUFA 18:2 (linoleic acid, undifferentiated) (g)")
  names(df)[names(df) == 'PUFA 18:2 (linoleic acid) (g)'] <- 'PUFA 18:2 (linoleic acid, undifferentiated) (g)'
  names(df)[names(df) == 'PUFA 18:3 (linolenic acid) (g)'] <- 'PUFA 18:3 (linolenic acid, undifferentiated) (g)'
  names(df)[names(df) == 'PUFA 20:4 (arachidonic acid) (g)'] <- 'PUFA 20:4 (arachidonic acid, undifferentiated) (g)'
  names(df)[names(df) == 'PUFA 20:5 (eicosapentaenoic acid [EPA]) (g)'] <- 'PUFA 20:5 n-3 (eicosapentaenoic acid [EPA]) (g)'
  names(df)[names(df) == 'PUFA 22:5 (docosapentaenoic acid [DPA]) (g)'] <- 'PUFA 22:5 n-3 (docosapentaenoic acid [DPA]) (g)'
  names(df)[names(df) == 'PUFA 22:6 (docosahexaenoic acid [DHA]) (g)'] <- 'PUFA 22:6 n-3 (docosahexaenoic acid [DHA]) (g)'
  df$"PUFA 18:2 n-6 (linoleic acid [LA]) (g)" <- NA
  df$"PUFA 18:3 n-6 (gamma-linolenic acid [GLA]) (g)" <- NA
  df$"PUFA 20:4 n-6 (arachidonic acid [AA]) (g)" <- NA
  return(df)
}

#data upload
Final_Baseline <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final Baseline.xlsx")
Final_2Tri <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final 2nd trimester.xlsx")
Final_3Tri <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final 3rd Trimester.xlsx")
Final_PP <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final Postpartum.xlsx")

mp_meal <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Missing Participants Export.xlsx", 
    sheet = "Meal Totals")
mp_daily <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Missing Participants Export.xlsx", 
    sheet = "Daily Totals")
# divide mp_meal into trimesters
mp_meal_Baseline <- mp_meal %>% filter(str_detect(`Participant/Menu ID`, "Base"))
mp_meal_Baseline$`Participant/Menu ID` <- substr(mp_meal_Baseline$`Participant/Menu ID`, 1, 4)
mp_meal_3rd_Tri <- mp_meal %>% filter(str_detect(`Participant/Menu ID`, "3rd"))
mp_meal_3rd_Tri$`Participant/Menu ID` <- substr(mp_meal_3rd_Tri$`Participant/Menu ID`, 1, 4)
mp_meal_Post <- mp_meal %>% filter(str_detect(`Participant/Menu ID`, "Post"))
mp_meal_Post$`Participant/Menu ID` <- substr(mp_meal_Post$`Participant/Menu ID`, 1, 4)
# combine
Final_Baseline <- rbind(Final_Baseline, mp_meal_Baseline)
Final_3Tri <- rbind(Final_3Tri, mp_meal_3rd_Tri)
Final_PP <- rbind(Final_PP, mp_meal_Post)

# per Scott, remove data for 119A, 7/14/21
Final_3Tri <- Final_3Tri %>% filter(!(`Participant/Menu ID` == "119A" & `Date of Intake` == as.Date("2021-07-14")))
# per Scott, remove data for 203B, 6/17/20
Final_Baseline <- Final_Baseline %>% filter(!(`Participant/Menu ID` == "203B" & `Date of Intake` == as.Date("2020-06-17")))
# per Scott, remove data for 202B, 12/23/19
Final_Baseline <- Final_Baseline %>% filter(!(`Participant/Menu ID` == "202B" & `Date of Intake` == as.Date("2019-12-23")))
# per Scott, remove data for 202B, 12/9/20
Final_2Tri <- Final_2Tri %>% filter(!(`Participant/Menu ID` == "202B" & `Date of Intake` == as.Date("2020-12-09")))

# year for 211B was entered as 2021 but really 2022
temp <- Final_PP %>% filter(`Participant/Menu ID` == "211B")
temp$`Date of Intake` <- temp$`Date of Intake` %m+% years(1)
Final_PP <- Final_PP %>% filter(!`Participant/Menu ID` == "211B")
Final_PP <- rbind(Final_PP, temp)

# read in archive files and check if they already exist
# probably baseline for 111A
# per Scott, do not include this file
# extra1 <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/NDS export archive/18-1798 Nutrient Analysis Export 1st Trimester.xlsx")
# a <- colnames(extra1)
# a <- data.frame(cbind(a, rep(1, ncol(extra1))))
# names(a) <- c("varname", "in_extra")
# b <- colnames(Final_Baseline)
# b <- data.frame(cbind(b, rep(1, ncol(Final_Baseline))))
# names(b) <- c("varname", "in_final_baseline")
# ab <- full_join(a, b, by = "varname")
# # combine
# extra1 <- rename(extra1)
# Final_Baseline <- rbind(Final_Baseline, extra1)

# this file is 2nd trimester for 108A, already in dataset
extra2 <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/NDS export archive/18-1798 Nutrient Analysis Export 2nd Trimester 9.10.2020.xlsx")

# 2nd trimester 100A
extra3 <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/NDS export archive/18-1798 Nutrient Analysis Export 2nd Trimester.xlsx")
a <- colnames(extra3)
a <- data.frame(cbind(a, rep(1, ncol(extra3))))
names(a) <- c("varname", "in_extra")
b <- colnames(Final_2Tri)
b <- data.frame(cbind(b, rep(1, ncol(Final_2Tri))))
names(b) <- c("varname", "in_final_2Tri")
ab <- full_join(a, b, by = "varname")
extra3 <- rename(extra3)
Final_2Tri <- rbind(Final_2Tri, extra3)

# file 18-1798 Nutrient Analysis Export 3rd Trimester 9.10.2020 has mix of data from 100A and 112A, but already in dataset

# 3rd trimester 108A
# per Scott, do not include this file
# extra4 <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/NDS export archive/18-1798 Nutrient Analysis Export 3rd Trimester.xlsx")
# a <- colnames(extra4)
# a <- data.frame(cbind(a, rep(1, ncol(extra4))))
# names(a) <- c("varname", "in_extra")
# b <- colnames(Final_3Tri)
# b <- data.frame(cbind(b, rep(1, ncol(Final_3Tri))))
# names(b) <- c("varname", "in_final_3Tri")
# ab <- full_join(a, b, by = "varname")
# extra4 <- rename(extra4)
# Final_3Tri <- rbind(Final_3Tri, extra4)

# file 18-1798 Nutrient Analysis Export 110A 2nd Tri is 2nd trimester for 110A, already in dataset
# file 18-1798 Nutrient Analysis Export Baseline 9.10.2020 is a mix of baseline data, all of it already in dataset

# file 18-1798 Nutrient Analysis Export Baseline has data from 102A, 106A which are not on master list of food logs
# file also has baseline for 108A which was missing
# also baseline for 112A which was missing
# also baseline for 113A which was missing
extra5 <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/NDS export archive/18-1798 Nutrient Analysis Export Baseline.xlsx")
extra5 <- extra5 %>% filter(`Participant/Menu ID` %in% c("108A", "112A", "113A"))
a <- colnames(extra5)
a <- data.frame(cbind(a, rep(1, ncol(extra5))))
names(a) <- c("varname", "in_extra")
b <- colnames(Final_Baseline)
b <- data.frame(cbind(b, rep(1, ncol(Final_Baseline))))
names(b) <- c("varname", "in_final_baseline")
ab <- full_join(a, b, by = "varname")
write_csv(ab,"/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/mismatch_colnames.csv")
extra5 <- rename(extra5)
Final_Baseline <- rbind(Final_Baseline, extra5)

# file 18-1798 Nutrient Analysis Export Post-Partum 9.10.2020 has PP for 100A, already in dataset

# file 18-1798 Nutrient Analysis Export Post-Partum has PP for 108A which does not appear to be in dataset, but is also not marked missing
extra6 <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/NDS export archive/18-1798 Nutrient Analysis Export Post-Partum.xlsx")
extra6 <- extra6 %>% filter(`Participant/Menu ID` %in% c("108A"))
a <- colnames(extra6)
a <- data.frame(cbind(a, rep(1, ncol(extra6))))
names(a) <- c("varname", "in_extra")
b <- colnames(Final_PP)
b <- data.frame(cbind(b, rep(1, ncol(Final_PP))))
names(b) <- c("varname", "in_final_PP")
ab <- full_join(a, b, by = "varname")
extra6 <- rename(extra6)
Final_PP <- rbind(Final_PP, extra6)

# file 18-1798 Nutrient Analysis Export Post-Partum has 3rd trimester for 107A (mislabeled as PP) which does not appear to be in dataset, but is also not marked missing
extra6a <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/NDS export archive/18-1798 Nutrient Analysis Export Post-Partum.xlsx")
extra6a <- extra6a %>% filter(`Participant/Menu ID` %in% c("107A"))
a <- colnames(extra6a)
a <- data.frame(cbind(a, rep(1, ncol(extra6a))))
names(a) <- c("varname", "in_extra")
b <- colnames(Final_3Tri)
b <- data.frame(cbind(b, rep(1, ncol(Final_3Tri))))
names(b) <- c("varname", "in_final_3Tri")
ab <- full_join(a, b, by = "varname")
extra6a <- rename(extra6a)
Final_3Tri <- rbind(Final_3Tri, extra6a)

# read in corrected file sent by CTRC
corrected <- read_excel('/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 11.22.2024 Export.xlsx')
a <- colnames(corrected)
a <- data.frame(cbind(a, rep(1, ncol(corrected))))
names(a) <- c("varname", "in_extra")
b <- colnames(Final_3Tri)
b <- data.frame(cbind(b, rep(1, ncol(Final_3Tri))))
names(b) <- c("varname", "in_final_3Tri")
ab <- full_join(a, b, by = "varname")
corrected$`Meal Data Field 1 Descriptor` <- NULL
corrected$`Meal Data Field 1 Response` <- NULL
corrected$`Meal Data Field 2 Descriptor` <- NULL
corrected$`Meal Data Field 2 Response` <- NULL
corrected$`Meal Data Field 3 Descriptor` <- NULL
corrected$`Meal Data Field 3 Response` <- NULL
corrected_base <- corrected %>% filter(str_detect(`Participant/Menu ID`, "Base") | str_detect(`Participant/Menu ID`, "1st"))
corrected_base$`Participant/Menu ID` <- substr(corrected_base$`Participant/Menu ID`, 1, 4)
corrected_2nd <- corrected %>% filter(str_detect(`Participant/Menu ID`, "2nd"))
corrected_2nd$`Participant/Menu ID` <- substr(corrected_2nd$`Participant/Menu ID`, 1, 4)
corrected_3rd <- corrected %>% filter(str_detect(`Participant/Menu ID`, "3rd"))
corrected_3rd$`Participant/Menu ID` <- substr(corrected_3rd$`Participant/Menu ID`, 1, 4)
# combine corrected data into existing datasets
Final_Baseline <- rbind(Final_Baseline, corrected_base)
Final_2Tri <- rbind(Final_2Tri, corrected_2nd)
Final_3Tri <- rbind(Final_3Tri, corrected_3rd)

PICLS <- read_csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/PICLSStudyHCLVsSAPTI_DATA_2024-05-24_1006(1).csv")
# groups and gest dates and centiles
gestdates <- read_csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data cleaning/CGM Data Check_6.8.23.csv") 
groups <- read.csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")%>% select(pid, randomization_group)
groups$randomization_group = factor(groups$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )

centiles = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/BulkCentileCalc_Global_v8.0.6.1_PICLS_v5_CJ.xlsm", 
    sheet = "Import") %>% filter(!is.na(`Data ID`)) %>% mutate(pid = `Data ID`)


```


```{r dc, include = FALSE}
######################### ctrc data clean ######################################
baseline_ctrc = Final_Baseline %>% mutate(tp = "Baseline", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

tri2_ctrc = Final_2Tri %>% mutate(tp = "Trimester 2", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

tri3_ctrc = Final_3Tri %>% mutate(tp = "Trimester 3", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

pp_ctrc = Final_PP %>% mutate(tp = "Post Partum", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

mp_ctrc = mp_meal %>% mutate(pid_string = `Participant/Menu ID`) %>% select(pid_string,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`,`Energy (kcal)` )

mp_ctrc = mp_ctrc %>% mutate(pid = stringr::word(pid_string, 1), tri = word(pid_string,2)) 
mp_ctrc = mp_ctrc %>% mutate(tp = case_when(tri == "3rd" ~ "Trimester 3",
                                            tri == "Base" ~ "Baseline",
                                            tri == "Post" ~ "Post Partum"))
mp_ctrc = mp_ctrc %>% select(-tri, -pid_string)


# merge
ctrc = full_join(baseline_ctrc, tri2_ctrc)
ctrc = full_join(ctrc, tri3_ctrc)
ctrc = full_join(ctrc, pp_ctrc)
ctrc = full_join(ctrc, mp_ctrc)
ctrc = right_join(groups, ctrc)
# 107A is in the above file

ctrc= ctrc %>% mutate(mufapufa = `% Calories from MUFA` + `% Calories from MUFA`)
gestdates = gestdates %>% select(pid, EDD_Final:`Delivery Date`)
ctrc2 = left_join(ctrc, gestdates)
ctrc2$date = ctrc2$`Date of Intake`
ctrc2 = ctrc2 %>% mutate(tp = case_when(date < mdy(SecondTri) ~ "Trimester 1",
                                        date >= mdy(SecondTri) & date < mdy(ThirdTri) ~ "Trimester 2",
                                        date >= mdy(ThirdTri) & date < mdy(`Delivery Date`) ~ "Trimester 3",
                                        date >= mdy(`Delivery Date`) ~ "Post Partum" ))
ctrc2$tp = factor(ctrc2$tp, levels = c("Trimester 1", "Trimester 2", "Trimester 3", "Post Partum"))

ctrc2 = ctrc2 %>% unique()

########################### MDIET   ############################################
maternal_diet_subj_day = ctrc2 %>% group_by(randomization_group, tp,pid, `Date of Intake`) %>% summarise(CarbTotal = sum(`Total Carbohydrate (g)`),
                                                                 Carb_Perc_Caltotal = mean(`% Calories from Carbohydrate`), # carbs
                                                                 FatTotal = sum(`Total Fat (g)`),
                                                                 Fat_Perc_Caltotal = mean(`% Calories from Fat`),
                                                                 SFA_Perc_Caltotal = mean(`% Calories from SFA`),
                                                                 MUFA_plus_PUFA_Perc_Caltotal = mean(mufapufa), #fats
                                                                 ProteinTotal = sum(`Total Protein (g)`),
                                                                 Protein_Perc_Caltotal = mean(`% Calories from Protein`),
                                                                 Animal_Protein_Perc = mean((`Animal Protein (g)`/`Total Protein (g)`)*100, na.rm=T),
                                                                 Plant_Protein_Perc = mean((`Vegetable Protein (g)`/`Total Protein (g)`)*100, na.rm=T), # protein
                                                                 FiberTotal = sum(`Total Dietary Fiber (g)`),
                                                                 GlyIndex_b = mean(`Glycemic Index (bread reference)`),
                                                                 GlyIndex_g = mean(`Glycemic Index (glucose reference)`),
                                                                 Cal_Total = sum(`Energy (kcal)`),
                                                                 Sfa_total = sum(`Total Saturated Fatty Acids (SFA) (g)`),
                                                                 Animal_total = sum(`Animal Protein (g)`),
                                                                 Plant_total = sum(`Vegetable Protein (g)`)
                                                                 )

# write file for Matthew to review
write.csv(maternal_diet_subj_day, "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/Data_clean/maternal_diet_subj_day.csv")

maternal_diet_group = maternal_diet_subj_day %>% group_by(randomization_group, tp) %>% summarise_all(.funs = mean) %>% select(-pid) %>% ungroup() %>% select(-`Date of Intake`)
maternal_diet_group2 = maternal_diet_subj_day %>% group_by( tp) %>% summarise_all(.funs = mean) %>% select(-pid, -randomization_group) %>% ungroup() %>% select(-`Date of Intake`)


################################################################################

####################### Standardized Diets ##########################################

####### EDIT HERE

#### add in extra columns for grams, then a column for standardized percentages of calorie totals below

#### add in a table for these standardized percentage values 

## converting grams of fat/prot/carb to calories: 9 fat, 4 carb, prot, 2 fiber

# standardized fiber intake averaged across the three trimesters, carb, fat and saturated fat, and total protein, animal and plant protein intake as a percentage of total calories
standardized_groups = maternal_diet_subj_day %>% group_by(pid, tp) %>% select(pid, tp, randomization_group, Cal_Total, FiberTotal,
                                                                              CarbTotal, FatTotal, Sfa_total, ProteinTotal, Animal_Protein_Perc, Plant_Protein_Perc, Animal_total, Plant_total) %>%
  mutate(# standardized percentage of calorie total
         s_fiber_calperc = (2*FiberTotal)/Cal_Total,
         s_carb_calperc = (4*CarbTotal)/Cal_Total,
         s_fat_calperc = (9*FatTotal)/Cal_Total,
         s_sfat_calperc = (9*Sfa_total)/Cal_Total,
         s_protein_calperc = (4*ProteinTotal)/Cal_Total,

         # standardized grams
         s_fiber_g = (FiberTotal)*2000/Cal_Total,
         s_carb_g = (CarbTotal)*2000/Cal_Total,
         s_fat_g = (FatTotal)*2000/Cal_Total,
         s_sfat_g = (Sfa_total)*2000/Cal_Total,
         s_protein_g = (ProteinTotal)*2000/Cal_Total,
         s_animal_perc = (Animal_total)/ProteinTotal*100,
         s_plant_perc = (Plant_total)/ProteinTotal*100,
         
         # Caltotals
         caltotal = Cal_Total)

# ###################### Fiber Diet totals ############################################
# 
# fiber_groups = maternal_diet_subj_day %>% group_by(pid, tp) %>% select(pid, tp, randomization_group, FiberTotal) %>% summarise(avg_daily_fiber = mean(FiberTotal))
# fiber_groups = fiber_groups %>% mutate(Diet_Quality = case_when(avg_daily_fiber < 15 ~ "Poor/Low Quality",
#                                                                 avg_daily_fiber >= 15 & avg_daily_fiber < 28 ~ "Moderate Quality",
#                                                                 avg_daily_fiber>= 28 ~"High Quality"))
# 
# fiber_groups = left_join(fiber_groups, groups)
# fiber_groups$Diet_Quality = factor(fiber_groups$Diet_Quality, levels = c("Poor/Low Quality", "Moderate Quality","High Quality"))

################################################################################

###################### STdized Grouped into trimester avg #####################################

standardized_tri_avg = standardized_groups %>% group_by(pid, tp)  %>% summarise(avg_daily_fiber_g_s = mean(s_fiber_g),
                                                                           avg_daily_carb_g_s = mean(s_carb_g),
                                                                           avg_daily_fat_g_s = mean(s_fat_g),
                                                                           avg_daily_satfat_g_s = mean(s_sfat_g),
                                                                           avg_daily_protein_g_s = mean(s_protein_g),

                                                                           avg_daily_fiber_calperc = mean(s_fiber_calperc)*100,
                                                                           avg_daily_carb__calperc = mean(s_carb_calperc)*100,
                                                                           avg_daily_fat__calperc = mean(s_fat_calperc)*100,
                                                                           avg_daily_satfat__calperc = mean(s_sfat_calperc)*100,
                                                                           avg_daily_protein__calperc = mean(s_protein_calperc)*100,
                                                                           avg_daily_animal_perc_s = mean(s_animal_perc),
                                                                           avg_daily_plant_perc_s = mean(s_plant_perc),
                                                                           avg_caltotal = mean(Cal_Total))

# diet_groups = diet_groups %>% mutate(Fiber = ifelse(avg_daily_fiber >= 28, "High Quality","Not High Quality"),
#                                      Carb = ifelse(avg_daily_carb >= 175 & avg_daily_carb < 220, "High Quality","Not High Quality"),
#                                      Fat = ifelse(avg_daily_fat_perc <= 30, "High Quality","Not High Quality"),
#                                      SatFat = ifelse(avg_daily_satfat_perc < 10, "High Quality","Not High Quality"),
#                                      Overall = ifelse(Fiber=="High Quality"&Carb=="High Quality"&Fat=="High Quality"&SatFat=="High Quality", "High Quality", "Not High Quality"))
# diet_groups = left_join(diet_groups, groups)
################################################################################

###################### MATERNAL WEIGHT GAIN ############################################

cleanedcgm_noacet <- read_csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Raw/cleanedcgm_noacet.csv")
gmi_data = cleanedcgm_noacet %>% mutate(pid = word(subject_id,1,sep = "_"),tri = word(subject_id,2,sep = "_")) %>% select(pid, tri, gmi)
gmi_data = gmi_data %>% mutate(tp = case_when(tri == "T1" ~ "Trimester 1",
                                              tri == "T2" ~ "Trimester 2",
                                              tri == "T3" ~ "Trimester 3",
                                              tri == "PP" ~ "Post Partum"
                                              )) %>% filter(!is.na(tp)) %>% select(-tri)



Preg_visits <- read_csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/QPregnancyVisits_DATA_2022-12-04_1234.csv")

Preg_visits = left_join(Preg_visits, gestdates)

Preg_visits = Preg_visits %>% mutate(tp = case_when(mdy(date_preg) < mdy(SecondTri) ~ "Trimester 1",
                                        mdy(date_preg) >= mdy(SecondTri) & mdy(date_preg) < mdy(ThirdTri) ~ "Trimester 2",
                                        mdy(date_preg) >= mdy(ThirdTri) & mdy(date_preg) < mdy(`Delivery Date`) ~ "Trimester 3",
                                        mdy(date_preg) >= mdy(`Delivery Date`) ~ "Post Partum" ))
Preg_visits$tp = factor(Preg_visits$tp, levels = c("Trimester 1", "Trimester 2", "Trimester 3", "Post Partum"))
matweights = Preg_visits %>% select(pid, tp, weight_preg, date_preg,FirstTri,SecondTri,ThirdTri,`Delivery Date`, bmi_preg)

# CHANGING TYPO FROM 68KG to 99 KG
matweights[141, 3] = 99.0

precons = PICLS %>% filter(redcap_event_name =="preconception_arm_1")%>% select(pid, weight_precon)

all_weights = left_join(matweights, precons) %>% filter(!is.na(tp))
last_weights = all_weights %>% group_by(pid, tp) %>% filter(!is.na(weight_preg)) %>% mutate(last_weight = ifelse(date_preg == max(date_preg), 1, 0)) %>% filter(last_weight == 1) %>% ungroup()

# avg weight gain by trimester, then converted to pounds to classify as excessive weight gain
last_weights = last_weights %>% mutate(weight_gain_total = weight_preg - weight_precon,
                                       week_length = as.numeric(difftime(mdy(date_preg), mdy(FirstTri), units="weeks")),
                                       avg_weekly_weight_gain = weight_gain_total/week_length,
                                       avg_weekly_weight_gain_lbs = avg_weekly_weight_gain*2.20462,
                                       weight_gain_total_lbs = weight_gain_total*2.20462)

last_weights = last_weights %>% mutate(excessive_weight_gain_wk = case_when(bmi_preg < 25 ~ ifelse(avg_weekly_weight_gain_lbs > 0.92, 1, 0),
                                                                         bmi_preg < 30 & bmi_preg >=25 ~ ifelse(avg_weekly_weight_gain_lbs > 0.66, 1, 0),
                                                                         bmi_preg >= 30  ~ ifelse(avg_weekly_weight_gain_lbs > 0.52, 1, 0)),
                                       excessive_weight_gain_total_tri = case_when(tp == "Trimester 3" & bmi_preg < 25~ ifelse(weight_gain_total_lbs > 35, 1,0),
                                                                               tp == "Trimester 3" & bmi_preg < 30 & bmi_preg >=25~ ifelse(weight_gain_total_lbs > 25, 1,0),
                                                                               tp == "Trimester 3" & bmi_preg >= 30~ ifelse(weight_gain_total_lbs > 20, 1,0)))

last_weights_tri = last_weights %>% filter(tp != "Post Partum") %>% group_by(pid) %>%fill(excessive_weight_gain_total_tri, .direction="updown")
last_weights_tri = last_weights_tri %>% mutate(excessive_weight_gain = ifelse(excessive_weight_gain_wk == 1 |tp == "Trimester 3" & excessive_weight_gain_total_tri == 1, 1, 0))
# post partum
last_weights_pp = last_weights %>% filter(tp == "Post Partum") %>% mutate(excessive_weight_gain_total_pp = case_when(bmi_preg < 25~ ifelse(weight_gain_total_lbs > 35, 1,0),
                                                                                bmi_preg < 30 & bmi_preg >=25~ ifelse(weight_gain_total_lbs > 25, 1,0),
                                                                                bmi_preg >= 30~ ifelse(weight_gain_total_lbs > 20, 1,0)))
last_weights_pp = last_weights_pp %>% mutate(excessive_weight_gain = ifelse(excessive_weight_gain_wk == 1 |excessive_weight_gain_total_pp == 1, 1, 0))

# tables
excessive_weight_tbl = last_weights_tri %>% group_by(tp) %>% summarise(count = sum(excessive_weight_gain, na.rm=T), percent = mean(excessive_weight_gain, na.rm=T)*100)
excessive_weight_tbl_pp = last_weights_pp %>% group_by(tp) %>% summarise(count = sum(excessive_weight_gain, na.rm=T), percent = mean(excessive_weight_gain, na.rm=T)*100)

# analysis for weekly mwg 
maternal_weight_gain_tri = last_weights_tri %>% select(pid, tp, excessive_weight_gain, avg_weekly_weight_gain_lbs, weight_gain_total_lbs)
avg_mwg_tri = maternal_weight_gain_tri %>% group_by(pid) %>% summarise(weekly_weight_gain_avg = mean(avg_weekly_weight_gain_lbs, na.rm =T))
###################### ANALYSIS SET FOR AIM 1 #####################################
mwg_total = full_join(maternal_weight_gain_tri, last_weights_pp %>% select(pid, tp, excessive_weight_gain, avg_weekly_weight_gain_lbs, weight_gain_total_lbs)) %>% unique()

analysis_aim1_tri = left_join(standardized_tri_avg, mwg_total) %>% unique()

#### Diet quality

analysis_aim1_tri = analysis_aim1_tri %>% mutate(s_fiber_group = case_when(avg_daily_fiber_g_s < 15 ~ "Low Quality",
                                                                         avg_daily_fiber_g_s >= 15 & avg_daily_fiber_g_s < 28 ~ "Medium Quality",
                                                                         avg_daily_fiber_g_s >= 28 ~ "High Quality"),
                                                 s_Fiber = case_when(avg_daily_fiber_g_s < 28 ~ "Not High Quality",
                                                                     avg_daily_fiber_g_s >= 28 ~ "High Quality"),
                                                 s_Carb = case_when(avg_daily_carb_g_s >= 175 & avg_daily_carb_g_s < 220 ~ "High Quality",
                                                                    avg_daily_carb_g_s < 175  ~ "Not High Quality (< 175)", 
                                                                    avg_daily_carb_g_s >= 220 ~ "Not High Quality (>220)"),
                                                 s_Fat = case_when(avg_daily_fat__calperc <= 30 ~ "High Quality",
                                                                   avg_daily_fat__calperc > 30 ~ "Not High Quality"),
                                                 s_SatFat = case_when(avg_daily_satfat__calperc < 10 ~ "High Quality",
                                                                      avg_daily_satfat__calperc >= 10 ~ "Not High Quality"),
                                                 s_Overall = ifelse(s_Fiber=="High Quality"& s_Carb=="High Quality"&s_Fat=="High Quality"&s_SatFat=="High Quality", "High Quality", "Not High Quality"))

analysis_aim1_tri$s_fiber_group = factor(analysis_aim1_tri$s_fiber_group, levels = c("Low Quality", "Medium Quality", "High Quality"))

analysis_m1_tri = analysis_aim1_tri %>% filter(tp != "Post Partum")
analysis_m1 = analysis_m1_tri %>% ungroup() %>% group_by(pid) %>% select(avg_daily_fiber_g_s:avg_daily_plant_perc_s, avg_weekly_weight_gain_lbs) %>% 
  summarise(fiber_s = mean(avg_daily_fiber_g_s),
            carb_s = mean(avg_daily_carb_g_s),
            fat_s = mean(avg_daily_fat_g_s),
            sfat_s = mean(avg_daily_satfat_g_s),
            protein_s = mean(avg_daily_protein_g_s),
            animal_perc_s = mean(avg_daily_animal_perc_s),
            plant_perc_s = mean(avg_daily_plant_perc_s),
            avg_wk_weight_gain_lbs = mean(avg_weekly_weight_gain_lbs, na.rm=T))

adjust = last_weights_tri %>% filter(tp == "Trimester 3") %>% select(pid, week_length)
analysis_m1 = left_join(analysis_m1, adjust)
analysis_m1[analysis_m1$pid == "110A", "week_length"] = 14.428571

##############################
base_bmis = PICLS %>% select(pid, bmi_baseline) %>% filter(!is.na(bmi_baseline))
analysis_m1_tri = left_join(analysis_m1_tri, base_bmis)

#analysis_m1_pp = analysis_aim1_tri %>% filter(tp == "Post Partum")
sg_pp = standardized_groups %>% filter(tp == "Post Partum") %>% group_by(pid, tp)  %>% summarise(avg_daily_fiber = mean(FiberTotal),
                                                                           avg_daily_carb = mean(CarbTotal),
                                                                           avg_daily_fat = mean(FatTotal),
                                                                           avg_daily_satfat = mean(Sfa_total),
                                                                           avg_daily_protein = mean(ProteinTotal),
                                                                           avg_daily_animal_perc = mean(Animal_Protein_Perc),
                                                                           avg_daily_plant_perc = mean(Plant_Protein_Perc))
analysis_m1_pp = left_join(mwg_total %>% filter(tp == "Post Partum"), sg_pp ) %>% filter(!is.na(avg_daily_fiber))



analysis_aim1_tri$standard_effect = 2000/analysis_aim1_tri$avg_caltotal
analysis_aim1_tri = left_join(analysis_aim1_tri, base_bmis)
# Fiber > 28 
# 175 <= Carb < 220
# Fat <=30 
# SatFat <10  (n = 6 for both satfat/fat joint)
```

```{r function, include = F}
# fit model fun
fit_mod = function(assoc,df){
  
  # Fit model
  print(assoc)  
  f = as.formula(paste0("avg_wk_weight_gain_lbs","~",assoc, "+ week_length"))
  mod = lm(f, data = df, na.action = na.omit)
  print(summary(mod))
  ggplot(data = df, aes_string(assoc, "avg_wk_weight_gain_lbs")) + geom_point() + theme_classic()
 
}

fit_lr_mod = function(assoc, df, interaction=FALSE){
  if(interaction==FALSE){
  print(assoc)  
  f = as.formula(paste0("excessive_weight_gain","~",assoc, "+ tp + bmi_baseline + (1|pid)"))
  mod = glmer(f, df, family = binomial)
  summary(mod)
  }
  else{
  print(assoc)  
  f = as.formula(paste0("excessive_weight_gain","~",assoc, "*bmi_baseline + tp  + (1|pid)"))
  mod = glmer(f, df, family = binomial)
  summary(mod)
  }
}

fit_mod_pp = function(assoc,df){
  
  # Fit model
  print(assoc)  
  f = as.formula(paste0("weight_gain_total_lbs","~",assoc))
  mod = lm(f, data = df, na.action = na.omit)
  print(summary(mod))
  ggplot(data = df, aes_string(assoc, "weight_gain_total_lbs")) + geom_point() + theme_classic()
 
}

```

# Outstanding Data questions/remarks

# Methods

Subjects maternal weight gain was measured using the latest date in a trimester compared with the preconception weight. This was then averaged into a weekly weight gain over all trimesters.

Excessive weight gain(either by weekly weight gain or total weight gain) was derived, and was stratified by timepoint.

Linear Models for Maternal Weight Gain were fit with standardized fiber intake, carb, fat and saturated fat, and total protein, animal and plant protein intake as a percentage of total calories averaged across the three trimesters. Standardization was based on a 2000 calorie diet and averaged first over trimesters, then overall. Avg Weekly Maternal Weight gain was measured as the average weekly weight gain across trimesters then averaged all together. Linear Models were adjusted for weeks of gestation between the first and last weight measurement used.

Generalized Linear Mixed Models with a binomial link function (Mixed Effects Logistic Regression) for Excessive Weight Gain were fit adjusting for enrollment maternal BMI. Similar Models were fit including an interaction term between the dietary independent variable and Maternal BMI at enrollment

Linear Models for Post Partum Maternal Weight Retention were fit with Average Daily fiber intake, carb, fat and saturated fat, and total protein, animal and plant protein intake as a percentage of total calories. Post Partum Weight retention is defined  as the difference between post partum weight and first pregnancy weight.

# Appended 8/27/2024

## visit count/ctrc log count


```{r}
# ctrc files aggregated from maternal_diet_subj_day dataframe into standardized_tri_Avg datafram in order to get a visit for tri 1 ,2 ,3, pp

table(standardized_tri_avg$tp)


table(standardized_tri_avg$pid, standardized_tri_avg$tp)

```

## Standardized Diet Quality (Overall and Fiber)

Standardized diet quality was based on taking the average daily total grams consumed for each nutrition type, converting to calories (9 per gram for fats, etc), then standardizing to a 2000 calorie diet total.

### Overall Criteria

n = 15 for high quality standardized fiber (4 more than non standardized), n = 19 for high quality standardized carbs (3 more than non standardized)


#### Overall Criteria

```{r}
table1(~s_Fiber + s_Carb + s_Fat + s_SatFat + s_Overall|tp, data = analysis_aim1_tri)
```

#### Dietary distributions by timepoint

standardized daily values (g)

```{r}
table1(~avg_daily_fiber_g_s + avg_daily_carb_g_s  + avg_daily_fat_g_s + avg_daily_satfat_g_s + avg_daily_protein_g_s |tp, data = analysis_aim1_tri)
```

standardizing effect sizes

```{r}
table1(~ standard_effect |tp, data = analysis_aim1_tri)
```

#### Calorie Percentage distributions by timepoint

```{r}
table1(~avg_caltotal+avg_daily_fiber_calperc+ avg_daily_carb__calperc  + avg_daily_fat__calperc +
         avg_daily_satfat__calperc +  avg_daily_protein__calperc+avg_daily_animal_perc_s + avg_daily_plant_perc_s |tp, data = analysis_aim1_tri)

```


### Fiber Criteria

```{r}
table1(~s_fiber_group|tp, data = analysis_aim1_tri)
```

### Excessive weight gain 3rd tri only

note: poor distn of EWG

```{r}
analysis_aim1_tri3 = analysis_aim1_tri %>% filter(tp == "Trimester 3")
table(analysis_aim1_tri3$excessive_weight_gain, analysis_aim1_tri3$s_Fiber)

ew_lr_tri3 = glm(excessive_weight_gain ~ s_Fiber+ bmi_baseline, data = analysis_aim1_tri3, family = "binomial")
summary(ew_lr_tri3)
```

# Analysis

## Excessive Weight Gain

### Excessive Weight Gain table

Excessive weight gain is defines as avg weight gain of 0.92lbs/wk, 0.66lbs/wk, or 0.52lbs/wk for Normal, Overweight, and Obese Maternal BMI,
Or a total weight gain of 35, 25, or 20 lbs respectively.

Below is the excessive weight gain distribution by trimester.

```{r, results='asis'}
# trimester excessive weight gain counts
kable(excessive_weight_tbl)
# post-partum
kable(excessive_weight_tbl_pp)
```

### Mixed Logistic Regression Models

Standardized Fiber, Carb, Fat, SatFat, Protein, and Percentage of Animal/Plant Protein were not found to be associated with excessive weight gain.

#### Fiber

fiber group not associated with EWG

```{r}
fit_lr_mod("avg_daily_fiber_g_s", analysis_m1_tri, F)

fit_lr_mod("s_fiber_group", analysis_m1_tri, F)
```

#### Carb

```{r}
fit_lr_mod("avg_daily_carb_g_s", analysis_m1_tri, F)
```

#### Fat

```{r}
fit_lr_mod("avg_daily_fat_g_s", analysis_m1_tri, F)
```

#### Sat Fat

```{r}
fit_lr_mod("avg_daily_satfat_g_s", analysis_m1_tri, F)
```

#### Protein

```{r}
fit_lr_mod("avg_daily_protein_g_s", analysis_m1_tri, F)
```

#### Animal Protein

```{r}
fit_lr_mod("avg_daily_animal_perc_s", analysis_m1_tri, F)
```

#### Plant Protein

```{r}
fit_lr_mod("avg_daily_plant_perc_s", analysis_m1_tri, F)
```

### Mixed Logistic Regression Models with interaction term

Most models had trouble converging including the interaction term. I would interpret terms here with a grain of salt.

#### Fiber

```{r}
fit_lr_mod("avg_daily_fiber_g_s", analysis_m1_tri, T)
```

#### Carb

```{r}
fit_lr_mod("avg_daily_carb_g_s", analysis_m1_tri, T)
```

#### Fat

```{r}
fit_lr_mod("avg_daily_fat_g_s", analysis_m1_tri, T)
```

#### Sat Fat

```{r}
fit_lr_mod("avg_daily_satfat_g_s", analysis_m1_tri, T)
```

#### Protein

```{r}
fit_lr_mod("avg_daily_protein_g_s", analysis_m1_tri, T)
```

#### Animal Protein

```{r}
fit_lr_mod("avg_daily_animal_perc_s", analysis_m1_tri, T)
```

#### Plant Protein

```{r}
fit_lr_mod("avg_daily_plant_perc_s", analysis_m1_tri, T)
```

## Maternal Weight Gain

Standardized Fiber, Carb, Fat, SatFat, Protein, and Percentage of Animal/Plant Protein were not found to be associated with maternal weight gain.

### Linear Models

#### Fiber

```{r}
fit_mod("fiber_s", analysis_m1)
```

#### Carb

```{r}
fit_mod("carb_s", analysis_m1)
```

#### Fat

```{r}
fit_mod("fat_s", analysis_m1)
```

#### Sat Fat

```{r}
fit_mod("sfat_s", analysis_m1)
```

#### Protein

```{r}
fit_mod("protein_s", analysis_m1)
```

#### Animal Protein

```{r}
fit_mod("animal_perc_s", analysis_m1)
```

#### Plant Protein

```{r}
fit_mod("plant_perc_s", analysis_m1)
```

## Post Partum Weight Retention

### Linear Models

Standardized Fiber, Carb, Fat, SatFat, Protein, and Percentage of Animal/Plant Protein were not found to be associated with Post Partum Maternal Weight retention.

#### Fiber

```{r}
fit_mod_pp("avg_daily_fiber", analysis_m1_pp)
```

#### Carb

```{r}
fit_mod_pp("avg_daily_carb", analysis_m1_pp)
```

#### Fat

```{r}
fit_mod_pp("avg_daily_fat", analysis_m1_pp)
```

#### Sat Fat

```{r}
fit_mod_pp("avg_daily_satfat", analysis_m1_pp)
```

#### Protein

```{r}
fit_mod_pp("avg_daily_protein", analysis_m1_pp)
```

#### Animal Protein

```{r}
fit_mod_pp("avg_daily_animal_perc", analysis_m1_pp)
```

#### Plant Protein

```{r}
fit_mod_pp("avg_daily_plant_perc", analysis_m1_pp)
```