---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(readxl)
library(stringr)
library(lubridate)
library(nlme)
library(table1)


#data upload
Final_Baseline <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final Baseline.xlsx")
Final_2Tri <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final 2nd trimester.xlsx")
Final_3Tri <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final 3rd Trimester.xlsx")
Final_PP <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Final Postpartum.xlsx")

mp_meal <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Missing Participants Export.xlsx", 
    sheet = "Meal Totals")
mp_daily <- read_excel("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/CTRC files/18-1798 Nutrient Analysis Export/18-1798 Missing Participants Export.xlsx", 
    sheet = "Daily Totals")

PICLS <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS 3 NUTRITION LOGS/PICLSStudyHCLVsSAPTI_DATA_2024-05-24_1006(1).csv")
# groups and gest dates
gestdates <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data cleaning/CGM Data Check_6.8.23.csv") 
groups <- read.csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")%>% select(pid, randomization_group)
groups$randomization_group = factor(groups$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )
```


```{r dc, include = FALSE}
######################### ctrc data clean ######################################
baseline_ctrc = Final_Baseline %>% mutate(tp = "Baseline", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`)

tri2_ctrc = Final_2Tri %>% mutate(tp = "Trimester 2", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`)

tri3_ctrc = Final_3Tri %>% mutate(tp = "Trimester 3", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`)

pp_ctrc = Final_3Tri %>% mutate(tp = "Post Partum", pid = `Participant/Menu ID`) %>% select(pid, tp,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`)

mp_ctrc = mp_meal %>% mutate(pid_string = `Participant/Menu ID`) %>% select(pid_string,`Date of Intake`, `Meal Name`,`Meal Time`,
                                                                                          `Total Carbohydrate (g)`,`% Calories from Carbohydrate`, `Total Grains (ounce equivalents)`,
                                                                                          `Total Protein (g)`, `Animal Protein (g)`, `Vegetable Protein (g)`,`% Calories from Protein`,
                                                                                          `Total Fat (g)`,`% Calories from Fat`, `Total Saturated Fatty Acids (SFA) (g)`,
                                                                                          `Total Monounsaturated Fatty Acids (MUFA) (g)`,`% Calories from MUFA`,
                                                                                          `Total Polyunsaturated Fatty Acids (PUFA) (g)`,`% Calories from PUFA`,
                                                                                          `Total Dietary Fiber (g)`, `Soluble Dietary Fiber (g)`, `Insoluble Dietary Fiber (g)`,
                                                                                          `Water (g)`, 108:110,`Glycemic Index (bread reference)`,`% Calories from SFA`,
                                                                                          `Glycemic Load (bread reference)`,
                                                                                          `Glycemic Load (glucose reference)`, `Glycemic Index (glucose reference)`)

mp_ctrc = mp_ctrc %>% mutate(pid = stringr::word(pid_string, 1), tri = word(pid_string,2)) 
mp_ctrc = mp_ctrc %>% mutate(tp = case_when(tri == "3rd" ~ "Trimester 3",
                                            tri == "Base" ~ "Baseline",
                                            tri == "Post" ~ "Post Partum"))
mp_ctrc = mp_ctrc %>% select(-tri, -pid_string)


# merge
ctrc = full_join(baseline_ctrc, tri2_ctrc)
ctrc = full_join(ctrc, tri3_ctrc)
ctrc = full_join(ctrc, pp_ctrc)
ctrc = full_join(ctrc, mp_ctrc)
ctrc = right_join(groups, ctrc)

ctrc= ctrc %>% mutate(mufapufa = `% Calories from MUFA` + `% Calories from MUFA`)
gestdates = gestdates %>% select(pid, EDD_Final:`Delivery Date`)
ctrc2 = left_join(ctrc, gestdates)
ctrc2$date = ctrc2$`Date of Intake`
ctrc2 = ctrc2 %>% mutate(tp = case_when(date < mdy(SecondTri) ~ "Trimester 1",
                                        date >= mdy(SecondTri) & date < mdy(ThirdTri) ~ "Trimester 2",
                                        date >= mdy(ThirdTri) & date < mdy(`Delivery Date`) ~ "Trimester 3",
                                        date >= mdy(`Delivery Date`) ~ "Post Partum" ))
ctrc2$tp = factor(ctrc2$tp, levels = c("Trimester 1", "Trimester 2", "Trimester 3", "Post Partum"))


########################### MDIET   ############################################
maternal_diet_subj_day = ctrc2 %>% group_by(randomization_group, tp,pid, `Date of Intake`) %>% summarise(CarbTotal = sum(`Total Carbohydrate (g)`),
                                                                 Carb_Perc_Caltotal = mean(`% Calories from Carbohydrate`), # carbs
                                                                 FatTotal = sum(`Total Fat (g)`),
                                                                 Fat_Perc_Caltotal = mean(`% Calories from Fat`),
                                                                 SFA_Perc_Caltotal = mean(`% Calories from SFA`),
                                                                 MUFA_plus_PUFA_Perc_Caltotal = mean(mufapufa), #fats
                                                                 ProteinTotal = sum(`Total Protein (g)`),
                                                                 Protein_Perc_Caltotal = mean(`% Calories from Protein`),
                                                                 Animal_Protein_Perc = mean((`Animal Protein (g)`/`Total Protein (g)`)*100, na.rm=T),
                                                                 Plant_Protein_Perc = mean((`Vegetable Protein (g)`/`Total Protein (g)`)*100, na.rm=T), # protein
                                                                 FiberTotal = sum(`Total Dietary Fiber (g)`),
                                                                 GlyIndex_b = mean(`Glycemic Index (bread reference)`),
                                                                 GlyIndex_g = mean(`Glycemic Index (glucose reference)`)
                                                                 )

maternal_diet_group = maternal_diet_subj_day %>% group_by(randomization_group, tp) %>% summarise_all(.funs = mean) %>% select(-pid) %>% ungroup() %>% select(-`Date of Intake`)
maternal_diet_group2 = maternal_diet_subj_day %>% group_by( tp) %>% summarise_all(.funs = mean) %>% select(-pid, -randomization_group) %>% ungroup() %>% select(-`Date of Intake`)


################################################################################
###################### Fiber Diet totals ############################################

fiber_groups = maternal_diet_subj_day %>% group_by(pid, tp) %>% select(pid, tp, randomization_group, FiberTotal) %>% summarise(avg_daily_fiber = mean(FiberTotal))
fiber_groups = fiber_groups %>% mutate(Diet_Quality = case_when(avg_daily_fiber < 15 ~ "Poor/Low Quality",
                                                                avg_daily_fiber >= 15 & avg_daily_fiber < 28 ~ "Moderate Quality",
                                                                avg_daily_fiber>= 28 ~"High Quality"))

fiber_groups = left_join(fiber_groups, groups)
fiber_groups$Diet_Quality = factor(fiber_groups$Diet_Quality, levels = c("Poor/Low Quality", "Moderate Quality","High Quality"))

################################################################################

###################### Overall Diet totals #####################################

diet_groups = maternal_diet_subj_day %>% group_by(pid, tp)  %>% summarise(avg_daily_fiber = mean(FiberTotal),
                                                                           avg_daily_carb = mean(CarbTotal),
                                                                           avg_daily_fat_perc = mean(Fat_Perc_Caltotal),
                                                                           avg_daily_satfat_perc = mean(SFA_Perc_Caltotal))

diet_groups = diet_groups %>% mutate(Fiber = ifelse(avg_daily_fiber >= 28, "High Quality","Not High Quality"),
                                     Carb = ifelse(avg_daily_carb >= 175 & avg_daily_carb < 220, "High Quality","Not High Quality"),
                                     Fat = ifelse(avg_daily_fat_perc <= 30, "High Quality","Not High Quality"),
                                     SatFat = ifelse(avg_daily_satfat_perc < 10, "High Quality","Not High Quality"),
                                     Overall = ifelse(Fiber=="High Quality"&Carb=="High Quality"&Fat=="High Quality"&SatFat=="High Quality", "High Quality", "Not High Quality"))
diet_groups = left_join(diet_groups, groups)
################################################################################

###################### MIXED MODELS ############################################

cleanedcgm_noacet <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Raw/cleanedcgm_noacet.csv")
gmi_data = cleanedcgm_noacet %>% mutate(pid = word(subject_id,1,sep = "_"),tri = word(subject_id,2,sep = "_")) %>% select(pid, tri, gmi)
gmi_data = gmi_data %>% mutate(tp = case_when(tri == "T1" ~ "Trimester 1",
                                              tri == "T2" ~ "Trimester 2",
                                              tri == "T3" ~ "Trimester 3",
                                              tri == "PP" ~ "Post Partum"
                                              )) %>% filter(!is.na(tp)) %>% select(-tri)



Preg_visits <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/QPregnancyVisits_DATA_2022-12-04_1234.csv")

Preg_visits = left_join(Preg_visits, gestdates)

Preg_visits = Preg_visits %>% mutate(tp = case_when(mdy(date_preg) < mdy(SecondTri) ~ "Trimester 1",
                                        mdy(date_preg) >= mdy(SecondTri) & mdy(date_preg) < mdy(ThirdTri) ~ "Trimester 2",
                                        mdy(date_preg) >= mdy(ThirdTri) & mdy(date_preg) < mdy(`Delivery Date`) ~ "Trimester 3",
                                        mdy(date_preg) >= mdy(`Delivery Date`) ~ "Post Partum" ))
Preg_visits$tp = factor(Preg_visits$tp, levels = c("Trimester 1", "Trimester 2", "Trimester 3", "Post Partum"))
matweights = Preg_visits %>% select(pid, tp, weight_preg, date_preg,FirstTri,SecondTri,ThirdTri,`Delivery Date`)

# maternal weight gain
matweights_tri = matweights %>% group_by(pid, tp) %>% summarise(avg_tri_weight = mean(weight_preg, na.rm=T)) %>% filter(!is.na(tp))
analysis_mwg = left_join(matweights_tri,fiber_groups) %>% group_by(pid) %>%fill(randomization_group, .direction = "updown") %>% filter(!is.na(randomization_group))
analysis_mwg = left_join(analysis_mwg, gmi_data)
analysis_mwg$tp = factor(analysis_mwg$tp, levels = c("Trimester 1", "Trimester 2", "Trimester 3", "Post Partum"))

MWG_model = lme(avg_tri_weight ~ avg_daily_fiber + tp + gmi, random = ~1|pid,
                data = analysis_mwg,
                na.action = na.omit)
summary(MWG_model)
# outcome: mwg ~ fiber + GMI + tp |pid

# nns


################################################################################
```
# Outstanding Data questions/remarks

Double check aggregating goals for aims 2 & 3 -- Daily aggregation then average across subj?

Double Check variables to use for outcomes: maternal weight, neonatal size/ how to measure them by trimester: last weight, net weight gain etc.

# Methods

Nutritional Dates set by gestational dating/trimester.

Maternal Diet assessed by group-timepoint daily means across subjects.

Diet Quality assessed by totalling Fiber intake by day then averaging across subject-trimesters and stratifying into Poor/Low, Moderate, and High Quality Diet groups. Similarly for other Diet outcomes for Overall High Quality Diet aim.

Mixed Models For Diet Quality (thru Fiber outcome) were fit with random intercept for subject, and adjusting for trimester and GMI for Maternal Weight gain and Neonatal size outcomes. Associations were assessed.

# Analysis

## Maternal Diet

```{r aim1}
print("OVERALL")
kable(maternal_diet_group2[by=.(tp)])
# print("Rand Groups")
# kable(maternal_diet_group[by=.(tp, randomization_group)])
```
## Diet Quality via Fiber Intake 

```{r aim2}
table1(~Diet_Quality|tp, data = fiber_groups)
#table1(~Diet_Quality|tp*randomization_group, data = fiber_groups)
```

## Diet Quality via Overall Criteria

Fiber > 28 - n = 13
175 <= Carb < 220 - n = 9
Fat <=30 - n = 8
SatFat <10 - n = 12 (n = 6 for both satfat/fat joint)

3 categories - n = 3

-- double check how to aggregate the diet totals

```{r aim3}
table1(~Fiber+Carb+Fat+SatFat+Overall|tp, data = diet_groups)
#table1(~Fiber+Carb+Fat+SatFat+Overall|tp*randomization_group, data = diet_groups)

# check distn before aggregating by subj-tri
print("DAILY TOTALS")
hist(maternal_diet_subj_day$FiberTotal, main = "Total Dietary Fiber (g)")
hist(maternal_diet_subj_day$CarbTotal, main = "Total Carb")
hist(maternal_diet_subj_day$Fat_Perc_Caltotal, main = "Total Fat %")
hist(maternal_diet_subj_day$SFA_Perc_Caltotal, main = "Total SFA %")
```

## Association of Diet thru Fiber Intake independent of GMI

### Maternal Weight Gain

Diet Quality using fiber as a surrogate marker is significantly associated with maternal weight gain independent of GMI (p = 0.03).

On average, we would expect every gram of daily fiber consumed to associate with a 0.13kg (SE = 0.05) decrease in Maternal weight

```{r mwg model}
summary(MWG_model)
```

### Neonatal Size

```{r nns model}

```
