---
title: "PICLS_cs"
author: "Casey Sakamoto"
date: "2023-11-13"
output: html_document
---

```{r setup, include = FALSE}
library(knitr)
library(stringr)
library(tidyverse)
library(nlme)
library(emmeans)
library(readxl)
library(openxlsx)
library(lubridate)
library(data.table)

#data upl  S:/Laura/BDC\Projects\Sarit Polsky\PICLS
######## FIRST SHEET. TDD STUFF
tdd_files_list <- list.files(path = 'S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_tdd_dfs <- lapply(as.list(tdd_files_list), function(x) read_xlsx(x, sheet = 1))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
tdd_df_names <- paste0(str_extract(basename(tdd_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_tdd_dfs) <- tdd_df_names

########## SECIND SHEET PUMP ADJ
pump_files_list <- list.files(path = 'S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_pumpdfs <- lapply(as.list(pump_files_list), function(x) read_xlsx(x, sheet = 2, col_types = c("text", "date", "text", "numeric", "numeric", "numeric", "numeric", "numeric")))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
pump_df_names <- paste0(str_extract(basename(pump_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_pumpdfs) <- pump_df_names

#### groups

```

```{r data clean, include=FALSE}
# tdd
varnames = c("date", "gweek", "gday", "tdd_u", "td_bolus", "td_basal", "bolus_perc", "basal_perc", "sampleid")

column_namer = function(x){
  df = x
  colnames(df) = varnames
  df
}

# want to add id and combine datasets
list_of_tdd_dfs = mapply(cbind, list_of_tdd_dfs, "SampleID"=tdd_df_names, SIMPLIFY=F)
#list_of_tdd_dfs[[2]]$`Gestational Week` = as.Date(list_of_tdd_dfs[[2]]$`Gestational Week`)
list_of_tdd_dfs[[22]] = list_of_tdd_dfs[[22]] %>% select(-Index)

list_of_tdd_dfs = lapply(list_of_tdd_dfs, column_namer)


tdd_df = do.call(rbind, list_of_tdd_dfs)
tdd_df = tdd_df %>% filter(!is.na(gweek))

# add in groups, months and pid
tdd_df$pid = substr(tdd_df$sampleid,1,4) 
tdd_df = left_join(tdd_df, groups)
tdd_df$randomization_group = factor(tdd_df$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )
tdd_df_wk_avg = tdd_df %>% group_by(pid, gweek) %>% summarise(tdd_u_wk = mean(tdd_u, na.rm = T),
                                                              td_bolus_wk = mean(td_bolus, na.rm = T),
                                                              td_basal_wk = mean(td_basal, na.rm = T),
                                                              bolus_perc_wk = mean(bolus_perc, na.rm = T),
                                                              basal_perc_wk = mean(basal_perc, na.rm = T))

tdd_df = tdd_df %>% mutate(gmonth = case_when(gweek %in% c("1","2","3","4") ~ "1",
                                              gweek %in% c("5","6","7","8") ~ "2",
                                              gweek %in% c("9","10","11","12") ~ "3",
                                              gweek %in% c("13","14","15","16") ~ "4",
                                              gweek %in% c("17","18","19","20") ~ "5",
                                              gweek %in% c("21","22","23","24") ~ "6",
                                              gweek %in% c("25","26","27","28") ~ "7",
                                              gweek %in% c("29","30","31","32") ~ "8",
                                              gweek %in% c("33","34","35","36") ~ "9",
                                              gweek %in% c("37","38","39") ~ "10",
                                              TRUE ~ gweek))
tdd_df$gmonth = factor(tdd_df$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))
tdd_df_month = tdd_df %>% group_by(pid, gmonth) %>% summarise(tdd_u_m = mean(tdd_u, na.rm = T),
                                                              td_bolus_m = mean(td_bolus, na.rm = T),
                                                              td_basal_m = mean(td_basal, na.rm = T),
                                                              bolus_perc_m = mean(bolus_perc, na.rm = T),
                                                              basal_perc_m = mean(basal_perc, na.rm = T))
tdd_df_month = left_join(tdd_df_month, groups %>% select(pid, randomization_group))

# pump settings   ,not sure if we need"Date of pump adjustment"
pumpvarnames = c("Gestational age","Date of pump adjustment","Start Time","Basal units/hour","Carb ratio (g/U)","Correction bolus","Total basal (U)","Active insulin time (hours)" )

column_namer_p = function(x){
  df = x
  df = df %>% select(pumpvarnames)

  return(df)
}


list_of_pumpdfs = lapply(list_of_pumpdfs, column_namer_p)
list_of_pumpdfs = mapply(cbind, list_of_pumpdfs, "sampleid"=pump_df_names, SIMPLIFY=F)

pump_df = do.call(rbind, list_of_pumpdfs)


## get into analysis/table form
pump_adj_df = pump_df %>% mutate(adj_count = ifelse(!is.na(`Date of pump adjustment`), 1, 0))
pump_adj_df$gweek = factor(tstrsplit(pump_adj_df$`Gestational age`, "w")[[1]], levels = c("Preconception", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
                                                                                             "11", "12", "13", "14", "15", "16", "17", "18", "19", "20",
                                                                                             "21", "22", "23", "24", "25", "26", "27", "28", "29", "30",
                                                                                             "31", "32", "33", "34", "35", "36", "37", "38", "Postpartum"))
pump_adj_df = pump_adj_df %>% filter(!is.na(gweek))
pump_adj_df$pid = substr(pump_adj_df$sampleid,1,4) 
pump_adj_df = left_join(pump_adj_df, groups)
pump_adj_df$randomization_group = factor(pump_adj_df$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )


n_pump_adj = pump_adj_df %>% group_by(gweek, randomization_group) %>% summarise(sum = sum(adj_count))

## add in wks order for the desc tables, and months for the model

## model function will be outcome = int + group + time + groupTime 1 = hcl 2 = sapt
```

```{r models fun, include = F}
contrasts = c("1 randomization_group1 - 1 randomization_group2",
"2 randomization_group1 - 2 randomization_group2",
"3 randomization_group1 - 3 randomization_group2",
"4 randomization_group1 - 4 randomization_group2",
"5 randomization_group1 - 5 randomization_group2",
"6 randomization_group1 - 6 randomization_group2",
"7 randomization_group1 - 7 randomization_group2",
"8 randomization_group1 - 8 randomization_group2",
"9 randomization_group1 - 9 randomization_group2",
"10 randomization_group1 - 10 randomization_group2")
# "1 randomization_group1 - 1 randomization_group2" in case we need pre/post
# "1 randomization_group1 - 1 randomization_group2"
fit_mod = function(outcome, df){

    # Fit model
  f = as.formula(paste0(outcome,"~","gmonth + randomization_group + gmonth*randomization_group"))
  mod = lme(f,
            random = ~1|pid,
            data = df,
            na.action = na.omit)
    cat("\n")
  print(outcome)
  cat("\n")
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ gmonth*randomization_group, adjust="none")

  mod_means2 = emmeans(mod,specs=~ gmonth*randomization_group, adjust="none")
  
  print(kable(mod_means$contrasts,digits = 3,caption = "Timepoint Contrasts"))[c(39,61,82, 102, 121, 139, 156, 172, 187, 201)]
  print(kable(mod_means2,digits = 3,caption = "Timepoint Means"))

}
```

We will plan to analyze the number of pump adjustments made per gestational week and the average basal rates, carb ratios, and correction factors for each meal period (breakfast, lunch, evening, overnight) for each time period in pregnancy (4-8 weeks gestation, etc.).

# Descriptive statistics per gestational week 

## number pump adjustments by group

```{r}
pump_adj_df %>% group_by(gweek, randomization_group) %>% summarise(sum = sum(adj_count))
```

#

# Models

# Non-PostPartum Months

# TDD data

## total daily dose 
```{r}
fit_mod("tdd_u_m", tdd_df_month)
```
total daily bolus 
```{r}
fit_mod("td_bolus_m", tdd_df_month)
```
total daily basal 
```{r}
fit_mod("td_basal_m", tdd_df_month)
```
percent basal  
```{r}
fit_mod("basal_perc_m", tdd_df_month)
```
percent bolus 
```{r}
fit_mod("bolus_perc_m", tdd_df_month)
```

# PostPartum Weeks

# Number of changes to pump settings per week/month 

## Basal units/hour 

## Carb ratio 

## Correction bolus (breakfast, lunch, afternoon snack, dinner – which hours correspond to each meal? Does this matter if we are just analyzing the number of changes per week/month?) 
