---
title: "PICLS_cs"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include = FALSE}
library(knitr)
library(stringr)
library(tidyverse)
library(nlme)
library(emmeans)
library(readxl)
library(openxlsx)
library(lubridate)
library(data.table)
library(table1)
library(chron)
library(htmlTable)
library(sjPlot)

#data upl  S:/Laura/BDC\Projects\Sarit Polsky\PICLS
######## FIRST SHEET. TDD STUFF
tdd_files_list <- list.files(path = '/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_tdd_dfs <- lapply(as.list(tdd_files_list), function(x) read_xlsx(x, sheet = 1))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
tdd_df_names <- paste0(str_extract(basename(tdd_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_tdd_dfs) <- tdd_df_names

########## SECIND SHEET PUMP ADJ
pump_files_list <- list.files(path = '/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_pumpdfs <- lapply(as.list(pump_files_list), function(x) read_xlsx(x, sheet = 2, col_types = c("text", "date", "text", "numeric", "numeric", "numeric", "numeric", "numeric")))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
pump_df_names <- paste0(str_extract(basename(pump_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_pumpdfs) <- pump_df_names

#### groups
groups <- read.csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")%>% select(pid, randomization_group)

```

```{r data clean, include=FALSE}
# tdd
varnames = c("date", "gweek", "gday", "tdd_u", "td_bolus", "td_basal", "bolus_perc", "basal_perc", "sampleid")

column_namer = function(x){
  df = x
  colnames(df) = varnames
  df
}

# want to add id and combine datasets
list_of_tdd_dfs = mapply(cbind, list_of_tdd_dfs, "SampleID"=tdd_df_names, SIMPLIFY=F)
#list_of_tdd_dfs[[2]]$`Gestational Week` = as.Date(list_of_tdd_dfs[[2]]$`Gestational Week`)
list_of_tdd_dfs[[22]] = list_of_tdd_dfs[[22]] %>% select(-Index)

list_of_tdd_dfs = lapply(list_of_tdd_dfs, column_namer)


tdd_df = do.call(rbind, list_of_tdd_dfs)
tdd_df = tdd_df %>% filter(!is.na(gweek))

# add in groups, months and pid
tdd_df$pid = substr(tdd_df$sampleid,1,4) 
tdd_df = left_join(tdd_df, groups)
tdd_df$randomization_group = factor(tdd_df$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )
tdd_df_wk_avg = tdd_df %>% group_by(pid, gweek) %>% summarise(tdd_u_wk = mean(tdd_u, na.rm = T),
                                                              td_bolus_wk = mean(td_bolus, na.rm = T),
                                                              td_basal_wk = mean(td_basal, na.rm = T),
                                                              bolus_perc_wk = mean(bolus_perc, na.rm = T),
                                                              basal_perc_wk = mean(basal_perc, na.rm = T))
#write.csv2(tdd_df, file = "ref_dates_picls.csv")

tdd_df = tdd_df %>% mutate(gmonth = case_when(gweek %in% c("1","2","3","4") ~ "1",
                                              gweek %in% c("5","6","7","8") ~ "2",
                                              gweek %in% c("9","10","11","12") ~ "3",
                                              gweek %in% c("13","14","15","16") ~ "4",
                                              gweek %in% c("17","18","19","20") ~ "5",
                                              gweek %in% c("21","22","23","24") ~ "6",
                                              gweek %in% c("25","26","27","28") ~ "7",
                                              gweek %in% c("29","30","31","32") ~ "8",
                                              gweek %in% c("33","34","35","36") ~ "9",
                                              gweek %in% c("37","38","39") ~ "10",
                                              TRUE ~ gweek))
tdd_df$gmonth = factor(tdd_df$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))
tdd_df_month = tdd_df %>% group_by(pid, gmonth) %>% summarise(tdd_u_m = mean(tdd_u, na.rm = T),
                                                              td_bolus_m = mean(td_bolus, na.rm = T),
                                                              td_basal_m = mean(td_basal, na.rm = T),
                                                              bolus_perc_m = mean(bolus_perc, na.rm = T),
                                                              basal_perc_m = mean(basal_perc, na.rm = T))
tdd_df_month = left_join(tdd_df_month, groups %>% select(pid, randomization_group))
tdd_df_month$randomization_group = factor(tdd_df_month$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )

tdd_df_month_a = tdd_df_month %>% filter(gmonth %in% c("1","2","3","4","5","6","7","8","9","10"))

# pump settings   ,not sure if we need"Date of pump adjustment"
pumpvarnames = c("Gestational age","Date of pump adjustment","Start Time","Basal units/hour","Carb ratio (g/U)","Correction bolus","Total basal (U)","Active insulin time (hours)" )


# lets just case when this
# 0.166666666666667 = 4
# 0.145833333333333 = 3.5
# 0.135416666666667 = 3.25
# 0.125 = 3
# 0.114583333333333  = 2.75
# 0.104166666666667 = 2.5
# 0.09375000 = 2.25
# 0.0833333333333333  = 2.00

column_namer_p = function(x){
  df = x
  df = df %>% select(pumpvarnames) %>% mutate(activetime = case_when(`Active insulin time (hours)` == 4 ~ 4,
                                                                     `Active insulin time (hours)` == 3 ~ 3,
                                                                     `Active insulin time (hours)` == 1/6 ~ 4,
                                                                     `Active insulin time (hours)` == 7/48 ~ 3.5,
                                                                     `Active insulin time (hours)` == 13/96  ~ 3.25,
                                                                     `Active insulin time (hours)` == 0.125  ~ 3,
                                                                     `Active insulin time (hours)` == 11/96  ~ 2.75,
                                                                     `Active insulin time (hours)` == 5/48 ~ 2.5,
                                                                     `Active insulin time (hours)` == 0.09375000 ~ 2.25,
                                                                     `Active insulin time (hours)` == 1/12 ~ 2.0,))

  return(df)
}




list_of_pumpdfs = lapply(list_of_pumpdfs, column_namer_p)
list_of_pumpdfs = mapply(cbind, list_of_pumpdfs, "sampleid"=pump_df_names, SIMPLIFY=F)

pump_df = do.call(rbind, list_of_pumpdfs)

####### APPENDED SEP 2024 ################################
pump_df = pump_df %>% mutate(time_conv = case_when(grepl(":", `Start Time`) ~ `Start Time`,
                                                     grepl(".", `Start Time`) ~ as.character(times(as.numeric(`Start Time`)))))


pump_df = pump_df %>% mutate(time_conv_l = str_length(time_conv))
pump_df = pump_df %>% mutate(time_conv2 = ifelse(time_conv_l == 8, gsub('.{3}$', '', time_conv), time_conv))
pump_df = pump_df %>% mutate(time_conv3 = ifelse(time_conv_l == 4, paste0("0",time_conv2), time_conv2),
                             time_conv_final = hm(time_conv3)) 
pump_df = pump_df %>% select(-c(time_conv:time_conv_l))



pump_df2 = pump_df %>% group_by(sampleid) %>% fill(`Gestational age`, .direction = "down")
pump_df2 = pump_df2 %>% group_by(sampleid) %>% fill(`Date of pump adjustment`, .direction = "down") %>% filter(!is.na(`Start Time`))


pump_df2 = pump_df2%>% group_by(sampleid, `Date of pump adjustment`) %>% fill(c("Basal units/hour","Carb ratio (g/U)","Correction bolus"), .direction = "down")
#table(pump_df$time_conv3)

pump_set_df = pump_df2

pump_set_df$gweek = factor(tstrsplit(pump_set_df$`Gestational age`, "w")[[1]], levels = c("Preconception", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
                                                                                             "11", "12", "13", "14", "15", "16", "17", "18", "19", "20",
                                                                                             "21", "22", "23", "24", "25", "26", "27", "28", "29", "30",
                                                                                             "31", "32", "33", "34", "35", "36", "37", "38", "Postpartum"))
#pump_set_df = pump_set_df %>% filter(!is.na(gweek))
pump_set_df$pid = substr(pump_set_df$sampleid,1,4) 
pump_set_df = left_join(pump_set_df, groups)
pump_set_df$randomization_group = factor(pump_set_df$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )

pump_set_df = pump_set_df  %>% mutate(gmonth = case_when(gweek %in% c("1","2","3","4") ~ "1",
                                              gweek %in% c("5","6","7","8") ~ "2",
                                              gweek %in% c("9","10","11","12") ~ "3",
                                              gweek %in% c("13","14","15","16") ~ "4",
                                              gweek %in% c("17","18","19","20") ~ "5",
                                              gweek %in% c("21","22","23","24") ~ "6",
                                              gweek %in% c("25","26","27","28") ~ "7",
                                              gweek %in% c("29","30","31","32") ~ "8",
                                              gweek %in% c("33","34","35","36") ~ "9",
                                              gweek %in% c("37","38","39") ~ "10",
                                              TRUE ~ gweek)
                                      )

# table(pump_set_df$time_conv3)
# meal times
breakfast = c("05:00","06:00","06:30","07:00", "08:00", "09:00", "09:30", "10:00")
lunch = c("11:00", "11:30", "12:00", "13:00", "13:30", "14:00", "14:30", "15:00", "15:30")
dinner = c("16:00", "16:30", "17:00", "17:30", "18:00", "18:30", "19:00", "19:30", "20:00", "20:30", "21:00")
overnight = c("22:00", "22:30", "23:00", "00:00", "01:30", "02:00", "03:00", "04:00", "04:30")
pump_set_df = pump_set_df %>% mutate(meal_period = case_when(time_conv3 %in% breakfast ~ "Breakfast",
                                                             time_conv3 %in% lunch ~ "Lunch",
                                                             time_conv3 %in% dinner ~ "Dinner",
                                                             time_conv3 %in% overnight ~ "Overnight"))
day = c("06:00", "06:30", "07:00", "08:00", "09:00", "09:30", "10:00", "11:00", "11:30", "12:00", "13:00", "13:30", "14:00", "14:30", "15:00", "15:30", "16:00", "16:30", "17:00", "17:30", "18:00", "18:30", "19:00", "19:30", "20:00", "20:30")
night = c("21:00", "22:00", "22:30", "23:00", "00:00", "01:30", "02:00", "03:00", "04:00", "04:30", "05:00")
pump_set_df = pump_set_df %>% mutate(day_period = case_when(time_conv3 %in% day ~ "Day",
                                                            time_conv3 %in% night ~ "Night"))

# "Carb ratio (g/U)"
avg_pumpset_cr_week = pump_set_df %>% group_by(pid, gweek,gmonth, meal_period, randomization_group) %>% summarise(carb_ratio_avg = mean(`Carb ratio (g/U)`, na.rm = T)) %>% ungroup()
avg_pumpset_cr_week$gmonth = factor(avg_pumpset_cr_week$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))

avg_pumpset_cr_month = avg_pumpset_cr_week %>% group_by(pid, gmonth, meal_period, randomization_group) %>% summarise(carb_ratio_avg_m = mean(carb_ratio_avg, na.rm = T)) %>% ungroup()
avg_pumpset_cr_month_breakfast = avg_pumpset_cr_month %>% filter(meal_period == "Breakfast")
avg_pumpset_cr_month_lunch = avg_pumpset_cr_month %>% filter(meal_period == "Lunch")
avg_pumpset_cr_month_dinner = avg_pumpset_cr_month %>% filter(meal_period == "Dinner")
avg_pumpset_cr_month_overnight = avg_pumpset_cr_month %>% filter(meal_period == "Overnight")

pumpset_cr_df = avg_pumpset_cr_month %>% group_by(randomization_group, gmonth, meal_period) %>% summarise(Avg_Carb_Ratio_gU = mean(carb_ratio_avg_m, na.rm = T)) %>% ungroup()

pumpset_cr_df = pumpset_cr_df[order(pumpset_cr_df$gmonth, pumpset_cr_df$meal_period),]

# "Correction bolus"
avg_pumpset_cb_week = pump_set_df %>% group_by(pid, gweek,gmonth, day_period, randomization_group) %>% summarise(correction_bolus_avg = mean(`Correction bolus`, na.rm = T)) %>% ungroup()
avg_pumpset_cb_week$gmonth = factor(avg_pumpset_cb_week$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))

avg_pumpset_cb_month = avg_pumpset_cb_week %>% group_by(pid, gmonth, day_period, randomization_group) %>% summarise(correction_bolus_avg_m = mean(correction_bolus_avg, na.rm = T)) %>% ungroup()
avg_pumpset_cb_month_day = avg_pumpset_cb_month %>% filter(day_period == "Day") 
avg_pumpset_cb_month_night = avg_pumpset_cb_month %>% filter(day_period == "Night") 

pumpset_cb_df = avg_pumpset_cb_month %>% group_by(randomization_group, gmonth, day_period) %>% summarise(Avg_Correction_Bolus = mean(correction_bolus_avg_m, na.rm = T)) %>% ungroup()

pumpset_cb_df = pumpset_cb_df[order(pumpset_cb_df$gmonth, pumpset_cb_df$day_period),]

##########################################################

## get into analysis/table form
pump_adj_df = pump_df %>% mutate(adj_count = ifelse(!is.na(`Date of pump adjustment`), 1, 0))
pump_adj_df$gweek = factor(tstrsplit(pump_adj_df$`Gestational age`, "w")[[1]], levels = c("Preconception", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
                                                                                             "11", "12", "13", "14", "15", "16", "17", "18", "19", "20",
                                                                                             "21", "22", "23", "24", "25", "26", "27", "28", "29", "30",
                                                                                             "31", "32", "33", "34", "35", "36", "37", "38", "Postpartum"))
pump_adj_df = pump_adj_df %>% filter(!is.na(gweek))
pump_adj_df$pid = substr(pump_adj_df$sampleid,1,4) 
pump_adj_df = left_join(pump_adj_df, groups)
pump_adj_df$randomization_group = factor(pump_adj_df$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )

# PUMP ADJUSTS
n_pump_adj = pump_adj_df %>% group_by(gweek, pid,randomization_group) %>% summarise(sum = sum(adj_count))


n_pump_adj = n_pump_adj %>% mutate(gmonth = case_when(gweek %in% c("1","2","3","4") ~ "1",
                                              gweek %in% c("5","6","7","8") ~ "2",
                                              gweek %in% c("9","10","11","12") ~ "3",
                                              gweek %in% c("13","14","15","16") ~ "4",
                                              gweek %in% c("17","18","19","20") ~ "5",
                                              gweek %in% c("21","22","23","24") ~ "6",
                                              gweek %in% c("25","26","27","28") ~ "7",
                                              gweek %in% c("29","30","31","32") ~ "8",
                                              gweek %in% c("33","34","35","36") ~ "9",
                                              gweek %in% c("37","38","39") ~ "10",
                                              TRUE ~ gweek))

n_pump_adj$gmonth = factor(n_pump_adj$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))

n_pump_adj_month = n_pump_adj %>% group_by(pid, gmonth, randomization_group) %>% summarise(pump_adj_m = sum(sum, na.rm = T))

# AVERAGE ACTIVE INSULIN TIME 

n_aai = pump_adj_df %>% group_by(gweek, pid,randomization_group) %>% summarise(Average_Active_Insulin = mean(activetime, na.rm = T))


n_aai = n_aai %>% mutate(gmonth = case_when(gweek %in% c("1","2","3","4") ~ "1",
                                              gweek %in% c("5","6","7","8") ~ "2",
                                              gweek %in% c("9","10","11","12") ~ "3",
                                              gweek %in% c("13","14","15","16") ~ "4",
                                              gweek %in% c("17","18","19","20") ~ "5",
                                              gweek %in% c("21","22","23","24") ~ "6",
                                              gweek %in% c("25","26","27","28") ~ "7",
                                              gweek %in% c("29","30","31","32") ~ "8",
                                              gweek %in% c("33","34","35","36") ~ "9",
                                              gweek %in% c("37","38","39") ~ "10",
                                              TRUE ~ gweek))

n_aai$gmonth = factor(n_aai$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))

n_aai_month = n_aai %>% group_by(pid, gmonth, randomization_group) %>% summarise(Average_Active_Insulin_m = mean(Average_Active_Insulin, na.rm = T))
# aai_df = n_aai_month %>% group_by( gmonth, randomization_group) %>% summarise(mean_Average_Active_Insulin_Time = mean(Average_Active_Insulin_m, na.rm = T),
#                                                                               sd_Average_Active_Insulin_Time = sd(Average_Active_Insulin_m, na.rm = T))
aai_df = n_aai_month %>% group_by( gmonth, randomization_group) %>% summarise(mean_Average_Active_Insulin_Time = mean(Average_Active_Insulin_m, na.rm = T))

#######################################
test = n_pump_adj %>% group_by(randomization_group, gweek) %>% summarise(mean_adj_by_subj = mean(sum), n_subj = n_distinct(pid))
## model function will be outcome = int + group + time + groupTime 1 = hcl 2 = sapt


# for postpartum take the subset of pp dates and create var that subtracts from earliest date to get week
# rinse repeat models

```

```{r models fun, include = F}

fit_mod = function(outcome, df){

    # Fit model
  f = as.formula(paste0(outcome,"~","gmonth + randomization_group + gmonth*randomization_group"))
  mod = lme(f,
            random = ~1|pid,
            data = df,
            correlation = corAR1(),
            na.action = na.omit)
    cat("\n")
  print(outcome)
  cat("\n")
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ gmonth*randomization_group, adjust="none")

  mod_means2 = emmeans(mod,specs=~ gmonth*randomization_group, adjust="none")
  
  mod_anova <- anova.lme(mod, type="marginal")
  
  print(mod_anova)
  print(kable(mod_means$contrasts[c(35,57,78,98, 117,135, 152, 168, 183,197)],digits = 3,caption = "Timepoint Contrasts", row.names = T))
  print(kable(mod_means2,digits = 3,caption = "Timepoint Means"))

}

fit_mod_cr = function(outcome, df){

    # Fit model
  f = as.formula(paste0(outcome,"~","gmonth + randomization_group + gmonth*randomization_group"))
  mod = lme(f,
            random = ~1|pid,
            data = df,
            correlation = corAR1(),
            na.action = na.omit)
    cat("\n")
  print(outcome)
  cat("\n")
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ gmonth*randomization_group, adjust="none")

  mod_means2 = emmeans(mod,specs=~ gmonth*randomization_group, adjust="none")
  

  print(kable(mod_means$contrasts[c(35,57,78,98, 117,135, 152, 168, 183,197)],digits = 3,caption = "Timepoint Contrasts", row.names = T))
  print(kable(mod_means2,digits = 3,caption = "Timepoint Means"))

}

```

# Methods

Linear mixed models were fit with random intercept for patient including group, gest month, and group time interaction. Contrasts were performed to compare groups at each gest month.

# Descriptive statistics 

## number pump adjustments by group

Mean Subject pump adjustments by group and gest week. Noted in PICLS meeting during early weeks only a few subjects had data

```{r}
#kable(pump_adj_df %>% group_by(gweek, randomization_group) %>% summarise(sum = sum(adj_count)))
kable(test%>% arrange(.,gweek))
```

## Carb Ratios

```{r}
kable(pumpset_cr_df)
```

## Correction Bolus

```{r}
kable(pumpset_cb_df)
```

## Average Active Insulin Time

```{r}
kable(aai_df%>% arrange(.,gmonth))
```

# Models

The first table in each set of model output is the overall ANOVA table. The p-value for the term labeled "gmonth:randomization_group" is the test of whether the trends in the outcome differed by group. If that p-value is significant, refer to the table of timepoint contrast for the comparison of the two groups at each timepoint.

## TDD data

### total daily dose 

```{r}
fit_mod("tdd_u_m", tdd_df_month)

```

### total daily bolus 

```{r}
fit_mod("td_bolus_m", tdd_df_month)
```

### total daily basal 

```{r}
fit_mod("td_basal_m", tdd_df_month)
```

### percent basal

```{r}
fit_mod("basal_perc_m", tdd_df_month)
```

### percent bolus 

```{r}
fit_mod("bolus_perc_m", tdd_df_month)
```

## Pump settings

### Number of changes to pump settings per month

```{r}
fit_mod("pump_adj_m", n_pump_adj_month)
```

### Carb ratios

#### Breakfast

```{r}
fit_mod("carb_ratio_avg_m", avg_pumpset_cr_month_breakfast)
```

#### Lunch

```{r}
fit_mod("carb_ratio_avg_m", avg_pumpset_cr_month_lunch)
```

#### Dinner

```{r}
fit_mod("carb_ratio_avg_m", avg_pumpset_cr_month_dinner)
```

#### Overnight

```{r}
fit_mod("carb_ratio_avg_m", avg_pumpset_cr_month_overnight)
```

### Correction bolus

#### Day

```{r}
fit_mod("correction_bolus_avg_m", avg_pumpset_cb_month_day)
```

#### Night

```{r}
fit_mod("correction_bolus_avg_m", avg_pumpset_cb_month_night)
```

### Average active insulin time

```{r}
fit_mod("Average_Active_Insulin_m", n_aai_month)
```