---
title: "PICLS_cs"
author: "Casey Sakamoto"
date: "2023-11-13"
output: html_document
---

```{r setup, include = FALSE}
library(knitr)
library(stringr)
library(tidyverse)
library(nlme)
library(emmeans)
library(readxl)
library(openxlsx)
library(lubridate)
library(data.table)

#data upl  S:/Laura/BDC\Projects\Sarit Polsky\PICLS
######## FIRST SHEET. TDD STUFF
tdd_files_list <- list.files(path = 'S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_tdd_dfs <- lapply(as.list(tdd_files_list), function(x) read_xlsx(x, sheet = 1))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
tdd_df_names <- paste0(str_extract(basename(tdd_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_tdd_dfs) <- tdd_df_names

########## SECIND SHEET PUMP ADJ
pump_files_list <- list.files(path = 'S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/CGM_Pump_ins',full.names = TRUE)
list_of_pumpdfs <- lapply(as.list(pump_files_list), function(x) read_xlsx(x, sheet = 2, col_types = c("text", "date", "text", "numeric", "numeric", "numeric", "numeric", "numeric")))

# Create a vector of names based on the first word of the filename + "Balance"
# Note that we can't use empty space in object names, hence the underscore
pump_df_names <- paste0(str_extract(basename(pump_files_list), "[^ ]+"), "_TDD_df")

# Assign the names to our list of dfs
names(list_of_pumpdfs) <- pump_df_names

#### groups
groups <- read.csv("S:/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Clean/PICLSStudyHCLVsSAPTI-RandomizedPatients_DATA_2022-11-14_1616.csv")%>% select(pid, randomization_group)

```

```{r data clean, include=FALSE}
# tdd
varnames = c("date", "gweek", "gday", "tdd_u", "td_bolus", "td_basal", "bolus_perc", "basal_perc", "sampleid")

column_namer = function(x){
  df = x
  colnames(df) = varnames
  df
}

# want to add id and combine datasets
list_of_tdd_dfs = mapply(cbind, list_of_tdd_dfs, "SampleID"=tdd_df_names, SIMPLIFY=F)
#list_of_tdd_dfs[[2]]$`Gestational Week` = as.Date(list_of_tdd_dfs[[2]]$`Gestational Week`)
list_of_tdd_dfs[[22]] = list_of_tdd_dfs[[22]] %>% select(-Index)

list_of_tdd_dfs = lapply(list_of_tdd_dfs, column_namer)


tdd_df = do.call(rbind, list_of_tdd_dfs)
tdd_df = tdd_df %>% filter(!is.na(gweek))

# add in groups, months and pid
tdd_df$pid = substr(tdd_df$sampleid,1,4) 
tdd_df = left_join(tdd_df, groups)
tdd_df$randomization_group = factor(tdd_df$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )
tdd_df_wk_avg = tdd_df %>% group_by(pid, gweek) %>% summarise(tdd_u_wk = mean(tdd_u, na.rm = T),
                                                              td_bolus_wk = mean(td_bolus, na.rm = T),
                                                              td_basal_wk = mean(td_basal, na.rm = T),
                                                              bolus_perc_wk = mean(bolus_perc, na.rm = T),
                                                              basal_perc_wk = mean(basal_perc, na.rm = T))

tdd_df = tdd_df %>% mutate(gmonth = case_when(gweek %in% c("1","2","3","4") ~ "1",
                                              gweek %in% c("5","6","7","8") ~ "2",
                                              gweek %in% c("9","10","11","12") ~ "3",
                                              gweek %in% c("13","14","15","16") ~ "4",
                                              gweek %in% c("17","18","19","20") ~ "5",
                                              gweek %in% c("21","22","23","24") ~ "6",
                                              gweek %in% c("25","26","27","28") ~ "7",
                                              gweek %in% c("29","30","31","32") ~ "8",
                                              gweek %in% c("33","34","35","36") ~ "9",
                                              gweek %in% c("37","38","39") ~ "10",
                                              TRUE ~ gweek))
tdd_df$gmonth = factor(tdd_df$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))
tdd_df_month = tdd_df %>% group_by(pid, gmonth) %>% summarise(tdd_u_m = mean(tdd_u, na.rm = T),
                                                              td_bolus_m = mean(td_bolus, na.rm = T),
                                                              td_basal_m = mean(td_basal, na.rm = T),
                                                              bolus_perc_m = mean(bolus_perc, na.rm = T),
                                                              basal_perc_m = mean(basal_perc, na.rm = T))
tdd_df_month = left_join(tdd_df_month, groups %>% select(pid, randomization_group))
tdd_df_month$randomization_group = factor(tdd_df_month$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )

tdd_df_month_a = tdd_df_month %>% filter(gmonth %in% c("1","2","3","4","5","6","7","8","9","10"))

# pump settings   ,not sure if we need"Date of pump adjustment"
pumpvarnames = c("Gestational age","Date of pump adjustment","Start Time","Basal units/hour","Carb ratio (g/U)","Correction bolus","Total basal (U)","Active insulin time (hours)" )

column_namer_p = function(x){
  df = x
  df = df %>% select(pumpvarnames)

  return(df)
}


list_of_pumpdfs = lapply(list_of_pumpdfs, column_namer_p)
list_of_pumpdfs = mapply(cbind, list_of_pumpdfs, "sampleid"=pump_df_names, SIMPLIFY=F)

pump_df = do.call(rbind, list_of_pumpdfs)


## get into analysis/table form
pump_adj_df = pump_df %>% mutate(adj_count = ifelse(!is.na(`Date of pump adjustment`), 1, 0))
pump_adj_df$gweek = factor(tstrsplit(pump_adj_df$`Gestational age`, "w")[[1]], levels = c("Preconception", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
                                                                                             "11", "12", "13", "14", "15", "16", "17", "18", "19", "20",
                                                                                             "21", "22", "23", "24", "25", "26", "27", "28", "29", "30",
                                                                                             "31", "32", "33", "34", "35", "36", "37", "38", "Postpartum"))
pump_adj_df = pump_adj_df %>% filter(!is.na(gweek))
pump_adj_df$pid = substr(pump_adj_df$sampleid,1,4) 
pump_adj_df = left_join(pump_adj_df, groups)
pump_adj_df$randomization_group = factor(pump_adj_df$randomization_group, levels = c(1,2), labels = c("HCL","SAPT") )


n_pump_adj = pump_adj_df %>% group_by(gweek, pid,randomization_group) %>% summarise(sum = sum(adj_count))


n_pump_adj = n_pump_adj %>% mutate(gmonth = case_when(gweek %in% c("1","2","3","4") ~ "1",
                                              gweek %in% c("5","6","7","8") ~ "2",
                                              gweek %in% c("9","10","11","12") ~ "3",
                                              gweek %in% c("13","14","15","16") ~ "4",
                                              gweek %in% c("17","18","19","20") ~ "5",
                                              gweek %in% c("21","22","23","24") ~ "6",
                                              gweek %in% c("25","26","27","28") ~ "7",
                                              gweek %in% c("29","30","31","32") ~ "8",
                                              gweek %in% c("33","34","35","36") ~ "9",
                                              gweek %in% c("37","38","39") ~ "10",
                                              TRUE ~ gweek))

n_pump_adj$gmonth = factor(n_pump_adj$gmonth, levels = c("Preconception","1","2","3","4","5","6","7","8","9","10","Postpartum"))

n_pump_adj_month = n_pump_adj %>% group_by(pid, gmonth, randomization_group) %>% summarise(pump_adj_m = sum(sum, na.rm = T))




## model function will be outcome = int + group + time + groupTime 1 = hcl 2 = sapt


# for postpartum take the subset of pp dates and create var that subtracts from earliest date to get week
# rinse repeat models

```

```{r models fun, include = F}

fit_mod = function(outcome, df){

    # Fit model
  f = as.formula(paste0(outcome,"~","gmonth + randomization_group + gmonth*randomization_group"))
  mod = lme(f,
            random = ~1|pid,
            data = df,
            na.action = na.omit)
    cat("\n")
  print(outcome)
  cat("\n")
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ gmonth*randomization_group, adjust="none")

  mod_means2 = emmeans(mod,specs=~ gmonth*randomization_group, adjust="none")
  

  print(kable(mod_means$contrasts[c(35,57,78,98, 117,135, 152, 168, 183,197)],digits = 3,caption = "Timepoint Contrasts", row.names = T))
  print(kable(mod_means2,digits = 3,caption = "Timepoint Means"))

}


```

# Methods

Pump adjustments by gest week were provided. Linear mixed models were fit with random intercept for patient including group, gest month, and group time interaction. Contrasts were performed to compare groups at each gest month.

# Descriptive statistics per gestational week 

## number pump adjustments by group

```{r}
kable(pump_adj_df %>% group_by(gweek, randomization_group) %>% summarise(sum = sum(adj_count)))
```

# Models

# TDD data

## total daily dose 

There were no significant monthly differences between HCL and SAPT group for TDD by month.

```{r}
fit_mod("tdd_u_m", tdd_df_month)
```
## total daily bolus 

There was a significant difference between tdb at month 10 between groups (p = 0.03). The HCL on average had 18.6 (se 7.9) higher tdb than the sapt group.

```{r}
fit_mod("td_bolus_m", tdd_df_month)
```

## total daily basal 

There was a significant difference between tdb at month 9 between groups (p = 0.03). The HCL on average had 9.8 (se 4.2) lower td basal than the sapt group.


```{r}
fit_mod("td_basal_m", tdd_df_month)
```
## percent basal

There was a significant difference in percent bolus between groups in months 2 and 9 (p = 0.03, 0.01)
HCL's percent bolus is on average 12.0 (se 5.3) points lower at 2 months and 13.7 (se = 4.9) points lower than the SAPT group at 9 months.

```{r}
fit_mod("basal_perc_m", tdd_df_month)
```

## percent bolus 

There was a significant difference in percent bolus between groups in months 2 and 9 (p = 0.03, 0.01)
HCL's percent bolus is on average 12.0 (se 5.3) points higher at 2 months and 13.7 (se = 4.9) points higher than the SAPT group at 9 months.

```{r}
fit_mod("bolus_perc_m", tdd_df_month)
```

# Number of changes to pump settings per month

There were no signifcant differences in the amount of changes to pump settings per month between groups.

```{r}
fit_mod("pump_adj_m", n_pump_adj_month)
```
