---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readxl)
library(lubridate)
library(table1)
library(nlme) # mixed models
library(emmeans)
library(ggrepel)

#data
visit_level <- read_excel("S:/Laura/BDC/Projects/Lauren Waterman/AID_inYA/AID A1c master with notes.xlsx", 
    sheet = "VisitLevel")
patient_level <- read_excel("S:/Laura/BDC/Projects/Lauren Waterman/AID_inYA/AID A1c master with notes.xlsx", 
    sheet = "PatientLevel")
```

```{r dataclean, include=FALSE}
# demographics -- patient level
# note that there are sums and a small table at the bottom of the spreadsheet, so need to remove the rows where no MRN
demog_df = patient_level %>% filter(!is.na(MRN))

# create retention rate var
disco_df = demog_df %>% mutate(disc_perm = ifelse(`currently on`=="n",1,0),
                               disc_temp = ifelse(`temp d/c`== "y" |`temp d/c`== "Y" ,1,0),
                               temp_reason = `...21`) %>% select(MRN, disc_perm, disc_temp, temp_reason, AID_StartDate)
disco_df = disco_df %>% mutate(temp_dc_13 = ifelse(temp_reason == 1 | temp_reason == 3, 1,0),
                               any_dc = ifelse(disc_perm == 1 | disc_temp == 1, 1, 0))

# remove pregnant visits
visitlevel = visit_level %>% filter(Pregnant == 'n')
visitlevel = visitlevel %>% filter(DaysFromIndexVisitDate<721)

# categorize into 90 day increments
# create days group
visitlevel = visitlevel %>% mutate(time_cat = case_when(DaysFromIndexVisitDate < 0  ~ "Baseline",
                                                    DaysFromIndexVisitDate >=0 & DaysFromIndexVisitDate < 90 ~ "0-89",
                                                    DaysFromIndexVisitDate >=90 & DaysFromIndexVisitDate < 180 ~ "90-179",
                                                    DaysFromIndexVisitDate >=180 & DaysFromIndexVisitDate < 270 ~ "180-269",
                                                    DaysFromIndexVisitDate >= 270 & DaysFromIndexVisitDate < 360 ~ "270-359",
                                                    DaysFromIndexVisitDate >= 360 & DaysFromIndexVisitDate < 450 ~ "360-449",
                                                    DaysFromIndexVisitDate >= 450 & DaysFromIndexVisitDate < 540 ~ "450-539",
                                                    DaysFromIndexVisitDate >= 540 & DaysFromIndexVisitDate < 630 ~ "540-629",
                                                    DaysFromIndexVisitDate >= 630 & DaysFromIndexVisitDate < 721 ~ "630-719"))
visitlevel$time_cat = factor(visitlevel$time_cat, levels = c("Baseline", "0-89", "90-179", "180-269", "270-359","360-449","450-539","540-629","630-719"))

############################
analysis = left_join(visitlevel, demog_df)
analysis = left_join(analysis, disco_df)

# within each subj time group take the closest days to the end of the category (least neg baseline, closest to 90, 179, etc)
analysis = analysis %>% group_by(MRN, time_cat) %>% filter(DaysFromIndexVisitDate == max(DaysFromIndexVisitDate)) %>% ungroup()

############################

# create age vars
analysis = analysis %>% mutate(Age = as.numeric(difftime(VisitDate, DOB))/365.25,
                               Age_cat = case_when(Age < 19 ~ "14-18",
                                                   Age < 23 & Age >= 19 ~ "19-22",
                                                   Age < 26 & Age >= 23 ~ "23-25",
                                                   Age >= 26  ~ "26-30"),
                               
                               a1c = `A1cValue_+/-7daysFromVisit`,
                               a1c_l7 = ifelse(a1c < 7, 1, 0))


# gather baseline and 2nd visit for insulin regimen changed

analysis$any_dc = ifelse(is.na(analysis$any_dc), 0, analysis$any_dc)
analysis$any_dc = factor(analysis$any_dc, levels = c(0,1), labels = c("Never DC", "Any DC"))

baseline = analysis %>% filter(time_cat == "Baseline")
baseline_a1c = baseline %>% mutate(baseline_a1c = case_when(`A1cValue_+/-7daysFromVisit` < 7 ~ "< 7",
                                                            `A1cValue_+/-7daysFromVisit` < 9 & `A1cValue_+/-7daysFromVisit` >= 7 ~ "7-9",
                                                            `A1cValue_+/-7daysFromVisit` > 9 ~ "> 9"),
                                   baseline_reg = InsulinRegimen) %>% select(MRN, baseline_a1c, AID_StartDate, baseline_reg)

analysis = left_join(analysis, baseline_a1c)
```

```{r t1setup, include = FALSE}
# create table 1
# table 1 includes: diabetes duration at time of aid initiation (days = 0), age at day = 0, prior insulin use (non aid pump or mdi, system switched to, race/ethn, insurance cat)
table1_df = analysis %>% filter(time_cat=="Baseline") %>%select(Age_cat, Race_Ethnicity, Gender, DiabetesDuration, InsulinRegimen, CGM, InsuranceCategory_IndexVisitDate, any_dc)

t1 = table1(~Age_cat+ Race_Ethnicity+ Gender+ DiabetesDuration+ InsulinRegimen+ CGM+ InsuranceCategory_IndexVisitDate|any_dc, data = table1_df, render.continuous = c(. = "Mean (SD)"))

##### find missing MRNS #####
table1_nobl = analysis%>% filter(time_cat=="Baseline") %>%select(MRN) %>% unique()
anal_mrn = analysis%>% select(MRN) %>% unique()

setdiff(anal_mrn$MRN,table1_nobl$MRN)
```

```{r models, include = FALSE}
####### ALL SUBJ ###############
# Fit random intercept model
mod_aa = lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc +time_cat, 
            random = ~1|MRN,
            data = analysis,
            na.action = na.omit)
# Anova
mod_aa_anova = anova.lme(mod_aa, type="marginal")
# Means
# mod_aa_means = emmeans(mod_aa,specs=pairwise ~ time_cat, adjust="none")
emm_options(rg.limit=12000)
mod_aa_means = emmeans(mod_aa, trt.vs.ctrl ~ time_cat, adjust="none")

#################################

######### A1c GRPS ##############

# Low a1c
analysis_lowa1c = analysis %>% filter(baseline_a1c == "< 7")
# Fit random intercept model
mod_lowa1c= lme(a1c ~ InsulinRegimen  + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc +time_cat, 
            random = ~1|MRN,
            data = analysis_lowa1c,
            na.action = na.omit)

# Anova
mod_lowa1c_anova = anova.lme(mod_lowa1c, type="marginal")
# Means
mod_lowa1c_means = emmeans(mod_lowa1c, trt.vs.ctrl ~ time_cat, adjust="none")

# MID a1c
analysis_mida1c = analysis %>% filter(baseline_a1c == "7-9")
# Fit random intercept model
mod_mida1c= lme(a1c ~ InsulinRegimen  + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc +time_cat, 
            random = ~1|MRN,
            data = analysis_mida1c,
            na.action = na.omit)

# Anova
mod_mida1c_anova = anova.lme(mod_mida1c, type="marginal")
# Means
mod_mida1c_means = emmeans(mod_mida1c, trt.vs.ctrl ~ time_cat, adjust="none")

# high a1c
analysis_higha1c = analysis %>% filter(baseline_a1c == "> 9")
# Fit random intercept model
mod_higha1c= lme(a1c ~ InsulinRegimen + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc +time_cat, 
            random = ~1|MRN,
            data = analysis_higha1c,
            na.action = na.omit)

# Anova
mod_higha1c_anova = anova.lme(mod_higha1c, type="marginal")
# Means
mod_higha1c_means = emmeans(mod_higha1c, trt.vs.ctrl ~ time_cat, adjust="none")
#################################



######### Age GRPS ##############

# Age 14-18
analysis18 = analysis %>% filter(Age_cat == "14-18")
# Fit random intercept model
mod_18 = lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc + time_cat, 
            random = ~1|MRN,
            data = analysis18,
            na.action = na.omit)

# Anova
mod_18_anova = anova.lme(mod_18, type="marginal")

# Means
mod_18_means = emmeans(mod_18, trt.vs.ctrl ~ time_cat, adjust="none")

# Age 19-22
analysis22 = analysis %>% filter(Age_cat == "19-22")
# Fit random intercept model
mod_22= lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc + time_cat, 
            random = ~1|MRN,
            data = analysis22,
            na.action = na.omit)

# Anova
mod_22_anova = anova.lme(mod_22, type="marginal")

# Means
mod_22_means = emmeans(mod_22, trt.vs.ctrl ~ time_cat, adjust="none")

# Age 23-25
analysis25 = analysis %>% filter(Age_cat == "23-25")
# Fit random intercept model
mod_25 = lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc + time_cat, 
            random = ~1|MRN,
            data = analysis25,
            na.action = na.omit)

# Anova
mod_25_anova = anova.lme(mod_25, type="marginal")

# Means
mod_25_means = emmeans(mod_25, trt.vs.ctrl ~ time_cat, adjust="none")

# Age 26-30
analysis30 = analysis %>% filter(Age_cat == "26-30")
# Fit random intercept model
mod_30 = lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc + time_cat, 
            random = ~1|MRN,
            data = analysis30,
            na.action = na.omit)

# Anova
mod_30_anova = anova.lme(mod_30, type="marginal")

# Means
mod_30_means = emmeans(mod_30, trt.vs.ctrl ~ time_cat, adjust="none")

# # Age 30+
# analysis30p = analysis %>% filter(Age_cat == "> 30")
# # Fit random intercept model
# mod_30p = lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + any_dc + time_cat, 
#             random = ~1|MRN,
#             data = analysis30p,
#             na.action = na.omit)
# 
# # Anova
# mod_30p_anova = anova.lme(mod_30p, type="marginal")
# 
# # Means
# mod_30p_means = emmeans(mod_30p, trt.vs.ctrl ~ time_cat, adjust="none")

#################################


######### insulin GRPS ##############MDI/Ratio Insulin Pump
# mdi
analysis_mdi = analysis %>% filter(baseline_reg == "MDI/Ratio")
# Fit random intercept model
mod_mdi= lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + time_cat, 
            random = ~1|MRN,
            data = analysis_mdi,
            na.action = na.omit)

# Anova
mod_mdi_anova = anova.lme(mod_mdi, type="marginal")
# Means
mod_mdi_means = emmeans(mod_mdi, trt.vs.ctrl ~ time_cat, adjust="none")

# pump
analysis_pump = analysis %>% filter(baseline_reg == "Insulin Pump")
# Fit random intercept model
mod_pump= lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + time_cat, 
            random = ~1|MRN,
            data = analysis_pump,
            na.action = na.omit)

# Anova
mod_pump_anova = anova.lme(mod_pump, type="marginal")
# Means
mod_pump_means = emmeans(mod_pump, trt.vs.ctrl ~ time_cat, adjust="none")
#################################


######### discontinued GRPS ##############
# nodc
analysis_nodc = analysis %>% filter(any_dc == "Never DC")
# Fit random intercept model
mod_nodc= lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + time_cat, 
            random = ~1|MRN,
            data = analysis_nodc,
            na.action = na.omit)

# Anova
mod_nodc_anova = anova.lme(mod_nodc, type="marginal")
# Means
mod_nodc_means = emmeans(mod_nodc, trt.vs.ctrl ~ time_cat, adjust="none")

# anydc
analysis_anydc = analysis %>% filter(any_dc == "Any DC")
# Fit random intercept model
mod_anydc= lme(a1c ~ InsulinRegimen + baseline_a1c + Age + DiabetesDuration + Race_Ethnicity + Gender + InsuranceCategory_IndexVisitDate + time_cat, 
            random = ~1|MRN,
            data = analysis_anydc,
            na.action = na.omit)

# Anova
mod_anydc_anova = anova.lme(mod_anydc, type="marginal")
# Means
mod_anydc_means = emmeans(mod_anydc, trt.vs.ctrl ~ time_cat, adjust="none")
#################################
```

```{r plots, include = FALSE}

############# A1C #################
# plots 
## means df
plotaa = data.frame(mod_aa_means$emmeans) %>% mutate(cohort = "Full Cohort")
plotlowa1c = data.frame(mod_lowa1c_means$emmeans)%>% mutate(cohort = "< 7")
plotmida1c = data.frame(mod_mida1c_means$emmeans)%>% mutate(cohort = "7-9")
plothigha1c = data.frame(mod_higha1c_means$emmeans)%>% mutate(cohort = "> 9")


plot_a1c_df = full_join(plotaa, plotlowa1c)
plot_a1c_df = full_join(plot_a1c_df, plotmida1c)
plot_a1c_df = full_join(plot_a1c_df, plothigha1c)


plot_a1c_df$cohort = factor(plot_a1c_df$cohort, levels = c("Full Cohort", "< 7", "7-9", "> 9"))
plot_a1c_df$time_cat = factor(plot_a1c_df$time_cat, labels = c("Baseline", "0-3 Months", "3-6 Months", "6-9 Months", "9-12 Months", "12-15 Months", "15-18 Months", "18-21 Months", "21-24 Months"))

lineplot_a1c = ggplot(plot_a1c_df, aes(x = time_cat, y = emmean, linetype = cohort, group = cohort)) +
  geom_line()  + geom_point(aes(shape = cohort)) + 
  geom_errorbar(aes(ymin = emmean - SE,ymax = emmean + SE), width = 0.2) + 
  ylab("HbA1c %") + xlab("Time") + theme_classic() 
#################################


########### AGE GROUPs ###############
plot18 = data.frame(mod_18_means$emmeans)%>% mutate(cohort = "14-18")
plot22 = data.frame(mod_22_means$emmeans)%>% mutate(cohort = "19-22")
plot25 = data.frame(mod_25_means$emmeans)%>% mutate(cohort = "23-25")
plot30 = data.frame(mod_30_means$emmeans)%>% mutate(cohort = "26-30")
# plot30p = data.frame(mod_30p_means$emmeans)%>% mutate(cohort = "> 30")


plot_age_df = full_join(plotaa, plot18)
plot_age_df = full_join(plot_age_df, plot22)
plot_age_df = full_join(plot_age_df, plot25)
plot_age_df = full_join(plot_age_df, plot30)
# plot_age_df = full_join(plot_age_df, plot30p)

plot_age_df$cohort = factor(plot_age_df$cohort, levels = c("Full Cohort", "14-18", "19-22", "23-25", "26-30"))
plot_age_df$time_cat = factor(plot_age_df$time_cat, labels = c("Baseline", "0-3 Months", "3-6 Months", "6-9 Months", "9-12 Months", "12-15 Months", "15-18 Months", "18-21 Months", "21-24 Months"))

lineplot_age = ggplot(plot_age_df, aes(x = time_cat, y = emmean, linetype = cohort, group = cohort)) +
  geom_line()  + geom_point(aes(shape = cohort)) + 
  geom_errorbar(aes(ymin = emmean - SE,ymax = emmean + SE), width = 0.2) + 
  ylab("HbA1c %") + xlab("Time") + theme_classic() 

#################################

############# Baseline Ins reg #################
plotmdi = data.frame(mod_mdi_means$emmeans)%>% mutate(cohort = "MDI/Ratio")
plotpump = data.frame(mod_pump_means$emmeans)%>% mutate(cohort = "Insulin Pump")


plot_ir_df = full_join(plotaa, plotmdi)
plot_ir_df = full_join(plot_ir_df, plotpump)

plot_ir_df$cohort = factor(plot_ir_df$cohort, levels = c("Full Cohort", "MDI/Ratio", "Insulin Pump"))
plot_ir_df$time_cat = factor(plot_ir_df$time_cat, labels = c("Baseline", "0-3 Months", "3-6 Months", "6-9 Months", "9-12 Months", "12-15 Months", "15-18 Months", "18-21 Months", "21-24 Months"))

lineplot_ir = ggplot(plot_ir_df, aes(x = time_cat, y = emmean, linetype = cohort, group = cohort)) +
  geom_line()  + geom_point(aes(shape = cohort)) + 
  geom_errorbar(aes(ymin = emmean - SE,ymax = emmean + SE), width = 0.2) + 
  ylab("HbA1c %") + xlab("Time") + theme_classic() 
#################################


############# Any DC #################
# plots 
## means df
plotnodc = data.frame(mod_nodc_means$emmeans)%>% mutate(cohort = "No DC")
plotanydc = data.frame(mod_anydc_means$emmeans)%>% mutate(cohort = "Any DC")


plot_dc_df = full_join(plotaa, plotnodc)
plot_dc_df = full_join(plot_dc_df, plotanydc)

plot_dc_df$cohort = factor(plot_dc_df$cohort, levels = c("Full Cohort", "No DC", "Any DC"))
plot_dc_df$time_cat = factor(plot_dc_df$time_cat, labels = c("Baseline", "0-3 Months", "3-6 Months", "6-9 Months", "9-12 Months", "12-15 Months", "15-18 Months", "18-21 Months", "21-24 Months"))

lineplot_dc = ggplot(plot_dc_df, aes(x = time_cat, y = emmean, linetype = cohort, group = cohort)) +
  geom_line()  + geom_point(aes(shape = cohort)) + 
  geom_errorbar(aes(ymin = emmean - SE,ymax = emmean + SE), width = 0.2) + 
  ylab("HbA1c %") + xlab("Time") + theme_classic() 
#################################

```

# Outstanding Data Remarks/Questions

- 2 subjects each with MRN 1864674 and 1936618 in patient level -- doesn't seem to be a problem at visit level; looks like only one set of visits per MRN

- 156/3340 visits pregnant = y were excluded
- 47/3184 visits were "multiple" visits in a window; value closest to the window border was used similar to the abstract 
- = 3137 visits used

- several subjects don't have a visit for Baseline ( days < 0 ): not included in desc stats table for now
(this n for desc table will be 613 (624-9))

# Methods

- A descriptive table for baseline measurements was produced

- A mixed model with random intercept was fit for A1c values and time, adjusting for age, sex, diabetes duration, race, insurance, baseline a1c, retention rate,and insulin regimen. Contrasts were generated to compare A1cs at each time and produce marginal estimates.

- plots of these marginal means and SE were produced

# Analysis

## Visits and Discontinuations

- Number of visits by timepoint really only dropped off after the 2y mark
- 98 of 624 participants had some sort of discontinuation, 24 of which was permanent. Of the discontinuations, 39 had either supply issues (24) or cost/insurance issues (15) as the specific reason for discontinuation

```{r demog/t1, echo=FALSE}
print("Number of Visits by Time (days from index), Age, and Time and Age")

kable(analysis %>% group_by(time_cat, Age_cat) %>% summarise(n()))
kable(analysis%>% group_by(Age_cat) %>% summarise(n()))
kable(analysis%>% group_by(time_cat) %>% summarise(n()))

print("Overall Retention Rates")
disco_df %>% summarise(sum(disc_perm), sum(disc_temp), sum(any_dc), sum(temp_dc_13, na.rm=T))
kable(analysis %>% group_by(time_cat) %>% count(`Discontinued?`))


print("A1c Rates")
kable(analysis %>% group_by(time_cat) %>% summarise(mean(a1c),perc_under7= mean(a1c_l7)))
```


also hoping to see % < 7 a1c at each timepoint & discontinuations (each visit by above)


## Demographics

Overall and split by retention rates

```{r t1, echo = FALSE}
t1
```

## Baseline A1c vs post-AID A1c 

### Entire Cohort

- Discontinuations (p <0.0001) and Baseline A1c (p < 0.0001) are both associated with A1c.

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod aa}
kable(mod_aa_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_aa_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_aa_means$emmeans, digits = 2, caption = "Timepoint Means")
```

### A1c Groups

```{r}
lineplot_a1c
```

#### < 7 A1c Baseline

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod lowa1c}
kable(mod_lowa1c_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_lowa1c_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_lowa1c_means$emmeans, digits = 2, caption = "Timepoint Means")
```

#### 7-9 A1c Baseline

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod mida1c}
kable(mod_mida1c_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_mida1c_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_mida1c_means$emmeans, digits = 2, caption = "Timepoint Means")
```

#### > 9 A1c Baseline

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod higha1c}
kable(mod_higha1c_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_higha1c_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_higha1c_means$emmeans, digits = 2, caption = "Timepoint Means")
```

### Age Groups

```{r}
lineplot_age
```

#### 14-18

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod 18}
kable(mod_18_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_18_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_18_means$emmeans, digits = 2, caption = "Timepoint Means")
```


#### 19-22

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod 22}
kable(mod_22_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_22_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_22_means$emmeans, digits = 2, caption = "Timepoint Means")
```

#### 23-25

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod 25}
kable(mod_25_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_25_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_25_means$emmeans, digits = 2, caption = "Timepoint Means")
```

#### 26-30

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod 30}
kable(mod_30_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_30_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_30_means$emmeans, digits = 2, caption = "Timepoint Means")
```

### Baseline Insulin Regimen

```{r}
lineplot_ir
```

#### mdi
- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod mdi}
kable(mod_mdi_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_mdi_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_mdi_means$emmeans, digits = 2, caption = "Timepoint Means")
```

#### pump

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod pump}
kable(mod_pump_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_pump_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_pump_means$emmeans, digits = 2, caption = "Timepoint Means")
```

### Retention Rate

```{r}
lineplot_dc
```


#### no dc

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod nodc}
kable(mod_nodc_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_nodc_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_nodc_means$emmeans, digits = 2, caption = "Timepoint Means")
```


#### anydc

- There was a significant difference between Post-AID and Baseline A1cs (see contrasts below)

- Marginal Means provided in timepoint means below

```{r mod anydc}
kable(mod_anydc_anova,digits = 3,caption = "Test of Overall Effect")
kable(mod_anydc_means$contrasts,digits = 3,caption = "Timepoint Contrasts")
kable(mod_anydc_means$emmeans, digits = 2, caption = "Timepoint Means")
```