---
title: "citeseq mofa analysis"
author: "Casey Sakamoto"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Install and load necessary packages
library(MOFA2)
library(ggplot2)
library(gridExtra)
library(tidyverse)
```

```{r}
#load model
outfile = file.path(getwd(),"model_nc_fast")
model <- load_model(outfile)

# overview of data
# significant amount of missing data (gray)
# makes sense as spread out like in howard's plot
plot_data_overview(model)

# Variance explained for every factor in per view
df = as.data.frame(model@cache$variance_explained$r2_per_factor[[1]])
top_2_max_values <- function(df) {
  for (col in colnames(df)) {
    if (is.numeric(df[[col]])) {  # Only process numeric columns
      sorted_indices <- order(df[[col]], decreasing = TRUE)
      top_2_values <- df[[col]][sorted_indices[1:2]]
      top_2_rows <- rownames(df)[sorted_indices[1:2]]
      
      cat("RNA cluster", col, "\n")
      for (i in 1:2) {
        cat("% Var Explained", top_2_values[i], "Component:", top_2_rows[i], "\n")
      }
      cat("\n")
    }
  }
}
top_2_max_values(df)
# using each group (rna cluster) as a view it looks like a single factor accounts for ~ 20-40% of the variation within.
# for clusters 0-11, the most important factors are factors 1-12 (not 1-1 for which explains which)
plot_variance_explained(model, x="view", y="factor")

#visualisation of factors 1-12 by idda1c as a start
# maybe not so interesting -- change to above/below median?
# model@samples_metadata
plot_factor(model, 
  factor = 1:12,
  color_by = "Baseline_IDDA1C",
  dodge = T
)
```

# important features cluster 0

```{r}
# visualization of feature weights
plot_weights(model,
  view = "0",
  factor = 11,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "0",
  factor = 11,
  nfeatures = 10
)
```

# important features cluster 1

```{r}
# visualization of feature weights
plot_weights(model,
  view = "1",
  factor = 9,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "1",
  factor = 9,
  nfeatures = 10
)
```

# important features cluster 2

```{r}
# visualization of feature weights
plot_weights(model,
  view = "2",
  factor = 12,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "2",
  factor = 12,
  nfeatures = 10
)
```

# important features cluster 3

```{r}
# visualization of feature weights
plot_weights(model,
  view = "3",
  factor = 6,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "3",
  factor = 6,
  nfeatures = 10
)
```

# important features cluster 4

```{r}
# visualization of feature weights
plot_weights(model,
  view = "4",
  factor = 10,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "4",
  factor = 10,
  nfeatures = 10
)
```

# important features cluster 5

```{r}
# visualization of feature weights
plot_weights(model,
  view = "5",
  factor = 5,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "5",
  factor = 5,
  nfeatures = 10
)
```

# important features cluster 6

```{r}
# visualization of feature weights
plot_weights(model,
  view = "6",
  factor = 8,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "6",
  factor = 8,
  nfeatures = 10
)
```

# important features cluster 7

```{r}
# visualization of feature weights
plot_weights(model,
  view = "7",
  factor = 3,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "7",
  factor = 3,
  nfeatures = 10
)
```

# important features cluster 8

note that two components explained a high % of variance for cluster 8: 7 and 4

```{r}
# visualization of feature weights
plot_weights(model,
  view = "8",
  factor = 7,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "8",
  factor = 7,
  nfeatures = 10
)

# visualization of feature weights
plot_weights(model,
  view = "8",
  factor = 4,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "8",
  factor = 4,
  nfeatures = 10
)
```

# important features cluster 9

```{r}
# visualization of feature weights
plot_weights(model,
  view = "9",
  factor = 1,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "9",
  factor = 1,
  nfeatures = 10
)
```

# important features cluster 10

```{r}
# visualization of feature weights
plot_weights(model,
  view = "10",
  factor = 2,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "10",
  factor = 2,
  nfeatures = 10
)
```

# important features cluster 11

```{r}
# visualization of feature weights
plot_weights(model,
  view = "11",
  factor = 1,
  nfeatures = 5,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
plot_top_weights(model,
  view = "11",
  factor = 1,
  nfeatures = 10
)
```

