---
title: "ASK/Fr1da SARS-COV-2 ab"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(tableone)
library(knitr)
library(car)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

################
# ASK analyses #
################

ask_data <- read.csv("B:\\Projects\\Marian Rewers\\ASK\\ASK Fr1da\\Data raw\\ASK dataset for Laura Feb 13 limited.csv",na.strings = c("NA",""," "))
ask_data$DM_pos <- ifelse(ask_data$MAB=="Yes" | ask_data$SBB=="Yes","YES","NO")
ask_data$DM_pos_01 <- ifelse(ask_data$DM_pos=="YES",1,
                                ifelse(!is.na(ask_data$DM_pos),0,NA))
ask_data$COVID_Infection2_01 <- ifelse(ask_data$COVID_Infection2=="Yes",1,
                                   ifelse(!is.na(ask_data$COVID_Infection2),0,NA))
ask_data$RaceEthnicity_2group <- as.factor(ifelse(ask_data$RaceEthnicity %in% c("NHW","Unknown"),"NHW","Not NHW"))

t1 <- CreateTableOne(vars=c("AgeAtScreen","DM_pos","MAB","SBB"),data=ask_data)
t1 <- as.data.frame(print(t1,exact=c("DM_pos","MAB","SBB"),nonnormal = "AgeAtScreen",printToggle = F))

t1_stratified <- CreateTableOne(vars=c("AgeAtScreen","DM_pos","MAB","SBB"),strata = "COVID_Infection2", data=ask_data)
t1_stratified <- as.data.frame(print(t1_stratified,exact=c("DM_pos","MAB","SBB"),nonnormal = "AgeAtScreen",printToggle = F))

mod1 <- glm(data=ask_data, DM_pos_01 ~ COVID_Infection2_01, family="binomial")
or1 <- exp(cbind(OR = coef(mod1), confint(mod1)))

mod3 <- glm(data=ask_data, DM_pos_01 ~ COVID_Infection2_01 + AgeAtScreen + Sex + T1D_Fam + RaceEthnicity_2group, family="binomial")
or3 <- exp(cbind(OR = coef(mod3), confint(mod3)))

##################
# Fr1da analyses #
##################

fr1da_data <- read.csv("B:\\Projects\\Marian Rewers\\ASK\\ASK Fr1da\\Data raw\\Fr1da dataset for Laura Feb 13 limited.csv",na.strings = c("NA",""," "))
fr1da_data$DM_pos <- ifelse(fr1da_data$MAB=="Yes" | fr1da_data$SBB=="Yes","YES","NO")
fr1da_data$DM_pos_01 <- ifelse(fr1da_data$DM_pos=="YES",1,
                                ifelse(!is.na(fr1da_data$DM_pos),0,NA))
fr1da_data$COVID_Infection_01 <- ifelse(fr1da_data$COVID_Infection =="TRUE",1,
                                   ifelse(!is.na(fr1da_data$COVID_Infection ),0,NA))

t1_fr1da <- CreateTableOne(vars=c("AgeAtScreen","DM_pos","MAB","SBB"),data=fr1da_data)
t1_fr1da <- as.data.frame(print(t1_fr1da,exact=c("DM_pos","MAB","SBB"),nonnormal = "AgeAtScreen",printToggle = F))

t1_fr1da_stratified <- CreateTableOne(vars=c("AgeAtScreen","DM_pos","MAB","SBB"),strata = "COVID_Infection", data=fr1da_data)
t1_fr1da_stratified <- as.data.frame(print(t1_fr1da_stratified,exact=c("DM_pos","MAB","SBB"),nonnormal = "AgeAtScreen",printToggle = F))

mod1_fr1da <- glm(data=fr1da_data, DM_pos_01 ~ COVID_Infection , family="binomial")
or1_fr1da <- exp(cbind(OR = coef(mod1_fr1da), confint(mod1_fr1da)))

mod3_fr1da <- glm(data=fr1da_data, DM_pos_01 ~ COVID_Infection  + AgeAtScreen + Sex + T1D_Fam, family="binomial")
or3_fr1da <- exp(cbind(OR = coef(mod3_fr1da), confint(mod3_fr1da)))


```

# ASK analyses

## Descriptive statistics, overall

```{r echo=FALSE}
kable(t1)
```

## Descriptive statistics, by COVID status

```{r echo=FALSE}
kable(t1_stratified)
```

## Unadjusted logistic regression model

The p-value (labeled "Pr(>Chisq)") tests whether there is a significant association between DM Ab positivity and SARS-Cov-2 positivity.  The association was not statistically significant.  The odds ratio was `r round(or1[2,1],2)`, which means that people who tested positive for COVID-19 antibodies had `r round(or1[2,1],2)` times the odds of having positive DM antibodies, although this was not significant.

```{r echo=FALSE,comment=""}
Anova(mod1,type = "III")
or1
```

## Adjusted logistic regression model - continuous age, race/ethnicity 2 categories 

The p-value (labeled "Pr(>Chisq)") for each term in the model tests whether there is a significant association between each variable and DM antibody positivity.  

The odds ratio for SARS-Cov-2 positivity was `r round(or3[2,1],2)`, which means that people with SARS-Cov-2 positivity had `r round(or3[2,1],2)` times the odds of having positive DM antibodies antibodies, and this was not significant.

```{r echo=FALSE,comment=""}
Anova(mod3,type = "III")
or3
```

# Fr1da analyses

## Descriptive statistics, overall

```{r echo=FALSE}
kable(t1_fr1da)
```

## Descriptive statistics, by COVID status

```{r echo=FALSE}
kable(t1_fr1da_stratified)
```

## Unadjusted logistic regression model

The p-value (labeled "Pr(>Chisq)") tests whether there is a significant association between DM Ab positivity and SARS-Cov-2 positivity.  The association was not statistically significant.  The odds ratio was `r round(or1_fr1da[2,1],2)`, which means that people who tested positive for COVID-19 antibodies had `r round(or1_fr1da[2,1],2)` times the odds of having positive DM antibodies, although this was not significant.

```{r echo=FALSE,comment=""}
Anova(mod1_fr1da,type = "III")
or1_fr1da
```

## Adjusted logistic regression model - continuous age

The p-value (labeled "Pr(>Chisq)") for each term in the model tests whether there is a significant association between each variable and DM antibody positivity.  

The odds ratio for SARS-Cov-2 positivity was `r round(or3_fr1da[2,1],2)`, which means that people with SARS-Cov-2 positivity had `r round(or3_fr1da[2,1],2)` times the odds of having positive DM antibodies antibodies, and this was not significant.

```{r echo=FALSE,comment=""}
Anova(mod3_fr1da,type = "III")
or3_fr1da
```