---
title: "High Risk Program"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(tidyverse)
library(nlme)
library(redcapAPI)
library(knitr)
# Check OS and alter file path accordingly.
if (.Platform$OS.type == "windows") {pathstart <- "//ucdenver.pvt/"} else if (.Platform$OS.type == "unix"){pathstart <- "/Volumes/"}
source("/Users/timvigers/Documents/GitHub/Tim-and-Laura/tim_R_functions.R")
```

```{r echo=FALSE,cache=TRUE}
# REDCap API data import
source("/Users/timvigers/Documents/GitHub/BDC-Code/api_tokens.R")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = majidi_high_risk)
data_full <- exportRecords(rcon)
```

```{r echo=FALSE,include=FALSE}
dat <- data_full
dat$record_id <- as.character(dat$record_id)
# import start and end dates, merge with dat, fill down id columns
dates <- read.csv(paste0(pathstart,"som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Shideh Majidi/High Risk Program/data_Cleaned/StartEndDates.csv"))
dates <- dates %>% select(record_id,program_start_date,program_end_date,Gender)
dates$record_id <- as.character(dates$record_id)
dat <- left_join(dat,dates,by = "record_id") %>%
  fill(record_id,name,mrn,control,case_id,dob)
# Fix incorrect visit date.
dat$visitdate[which(dat$visitdate == "2006-04-01")] <- "2016-04-01"
# Re-format columns
dat$record_id <- as.factor(dat$record_id)
dat$case_id <- as.factor(dat$case_id)
dat$Gender <- as.factor(dat$Gender)
dat$a1c <- as.numeric(dat$a1c)
dat$visitdate <- lubridate::ymd(as.character(dat$visitdate))
dat$dob <- lubridate::ymd(as.character(dat$dob))
dat$program_start_date <- lubridate::mdy(as.character(dat$program_start_date))
dat$program_end_date <- lubridate::mdy(as.character(dat$program_end_date))
dat$case_id <- as.factor(dat$case_id)
dat$control <- as.factor(dat$control)
dat$lowbg <- as.numeric(dat$lowbg)
dat$targetbg <- as.numeric(dat$targetbg)
dat$highbg <- as.numeric(dat$highbg)
dat$hospitalbefore <- as.numeric(dat$hospitalbefore)
dat$hosp_dur <- as.numeric(dat$hosp_dur)
dat$hospitalpost <- as.numeric(dat$hospitalpost)
dat$nsbefore <- as.numeric(dat$nsbefore)
dat$ns_dur <- as.numeric(dat$ns_dur)
dat$nspost <- as.numeric(dat$nspost)
dat$cancelbefore <- as.numeric(dat$cancelbefore)
dat$cancel_dur <- as.numeric(dat$cancel_dur)
dat$cancelpost <- as.numeric(dat$cancelpost)
# Remove participants with unverifiable BGs
incorrect.bgs <- read.csv(paste0(pathstart,"som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Shideh Majidi/High Risk Program/data_Cleaned/NotVerifiedBG.csv"))
incorrect.bgs$Record.ID <- as.character(incorrect.bgs$Record.ID)
dat <- dat %>%
  filter(!(record_id %in% incorrect.bgs$Record.ID))
# Calculate days between visits, age, t1d duration. Find those with complete dat.
dat <- dat %>% 
  arrange(record_id) %>%
  group_by(record_id) %>%
  mutate(days = visitdate - lag(visitdate)) %>%
  mutate(Age = round(as.numeric(difftime(visitdate,dob,units = "days")/365.2425),3)) %>%
  mutate(T1Dyears = round(as.numeric(difftime(dx,dob,units = "days")/365.2425),3)) %>%
  mutate(base_a1c = a1c[2]) %>%
  mutate(complete = length(which(!is.na(visitdate))))
```

```{r echo=FALSE}
# Match each case to 1 control
match <- dat %>%
  group_by(record_id) %>%
  filter(row_number()==1)
match$treat <- ifelse(match$control == "No",1,0)
# find those with most dat
match <- match %>%
  group_by(case_id,treat) %>%
  filter(complete == max(complete)) %>%
  arrange(case_id,desc(treat))
# Remove case ID groups without only controls. If two controls are tied for 
# amount of dat, use baseline A1c as a tiebreaker.
match <- match %>%
  group_by(case_id) %>%
  mutate(a1c_diff = abs(base_a1c - base_a1c[1])) %>% 
  filter(n() > 1) %>%
  slice(c(1,which(a1c_diff == min(a1c_diff[2:n()])))) %>%
  arrange(case_id,treat)
# 4 cases were excluded based on unverifiable BG values, 
# so should be 31 cases and 31 controls.

# Remove all unnecessary controls and blank rows
dat <- dat %>%
  filter(record_id %in% match$record_id) %>%
  filter(!is.na(visitdate) | redcap_event_name == "event_1_arm_1")
# Add timepoint based on visit date, remove unnecessary rows.
dat$timepoint <- ifelse(dat$visitdate <= dat$program_start_date,"Before",
                         ifelse(dat$visitdate > dat$program_start_date & 
                                  dat$visitdate <= dat$program_end_date,"During","After"))
dat$timepoint <- factor(dat$timepoint,levels = c("Before","During","After"))
```

```{r echo=FALSE,include=FALSE}
# Table 1
t1vars <- c("Age","T1Dyears","a1c","hospitalbefore","hosp_dur","hospitalpost",
            "nsbefore","ns_dur","nspost","cancelbefore","cancel_dur",
            "cancelpost","Gender","race_ethnicity","insurance","treatment___1",
            "treatment___2","treatment___3")
demos1 <- filter(dat, redcap_event_name == "event_1_arm_1") %>% 
  select("record_id","treatment___1","treatment___2","treatment___3",t1vars[-c(1,3)])
demos2 <- filter(dat, redcap_event_name == "event_2_arm_1") %>%
  select("record_id","control","case_id","a1c","Age")
demos <- merge(demos1,demos2, by = "record_id") %>%
  arrange(case_id,control)
# Make table 1
t1 <- CreateTableOne(t1vars,strata = "control",dat = demos,
                     testNonNormal = wilcox.test,
                     argsNonNormal = list(paired = TRUE),
                     testNormal = t.test,
                     argsNormal = list(paired = TRUE))

t1 <- print(t1,exact = c("insurance","race_ethnicity","treatment___3"),
            nonnormal = norm.check(demos,t1vars[1:12]))
```

### Table 1: Descriptive Statistics at Baseline Stratified by Group

```{r echo=FALSE}
kable(t1[,1:3])
```

```{r echo=FALSE}
# Filter out rows with just demographic information
mixed <- dat %>% filter(!is.na(visitdate))
# Mixed models of change in average before vs. during and during vs. after
# Average at each timepoint
avgs <- mixed %>%
  group_by(record_id,timepoint) %>%
  summarise(case_id = case_id[1],
            control = control[1],
            mean_a1c = mean(a1c,na.rm=T),
            mean_bgchecks = mean(bgchecks,na.rm=T),
            mean_low_bg = mean(lowbg,na.rm=T),
            mean_target_bg = mean(targetbg,na.rm=T),
            mean_high_bg = mean(highbg,na.rm=T),
            mean_days = mean(days,na.rm=T),
            sw = sum(sw == "Yes"))
# Differences before vs. during
bef_dur <- avgs %>%
  filter(timepoint %in% c("Before","During")) %>%
  arrange(record_id,timepoint) %>%
  group_by(record_id) %>%
  summarise(case_id = case_id[1],
            control = control[1],
            a1c = mean_a1c[2] - mean_a1c[1],
            bgchecks = mean_bgchecks[2]-mean_bgchecks[1],
            low_bg = mean_low_bg[2]-mean_low_bg[1],
            targetbg = mean_target_bg[2] - mean_target_bg[1],
            highbg = mean_high_bg[2]-mean_high_bg[1],
            days = mean_days[2] - mean_days[1],
            sw = sw[2] - sw[1])
bef_dur[bef_dur == "NaN"] <- NA
# Differences during vs. after
dur_aft <- avgs %>%
  filter(timepoint %in% c("During","After")) %>%
  arrange(record_id,timepoint) %>%
  group_by(record_id) %>%
  summarise(case_id = case_id[1],
            control = control[1],
            a1c = mean_a1c[2] - mean_a1c[1],
            bgchecks = mean_bgchecks[2]-mean_bgchecks[1],
            low_bg = mean_low_bg[2]-mean_low_bg[1],
            targetbg = mean_target_bg[2] - mean_target_bg[1],
            highbg = mean_high_bg[2]-mean_high_bg[1],
            days = mean_days[2] - mean_days[1],
            sw = sw[2] - sw[1])
dur_aft[dur_aft == "NaN"] <- NA
```

### Mixed Models

```{r echo=FALSE}
# A1c
# plot by timepoint
plot <- ggplot(avgs, aes(x=timepoint,y = mean_a1c,group = record_id))+
  geom_line(aes(color = control)) +
  ylab("Average HbA1c") +
  xlab("Timepoint") + 
  ggtitle("Figure 1: Average HbA1c by Timepoint")
plot
# Models
# Before vs. during
a1c <- lme(a1c ~ control,random = ~1|case_id,data = bef_dur,na.action = na.omit)
fixed <- summary(a1c)$tTable
kable(fixed, caption = "Average HbA1c: Change from Before to During")
# During vs. after
a1c <- lme(a1c ~ control,random = ~1|case_id,data = dur_aft,na.action = na.omit)
fixed <- summary(a1c)$tTable
kable(fixed, caption = "Average HbA1c: Change from During to After")
```

Each participant's HbA1c values were averaged for each time point (before, during, and after the intervention). The tables above represent the average change between time points for cases and controls. So on average HbA1c for cases decreased by 0.741 (although this wasn't statistically significant at the 0.05 level) and increased by 0.544 (1.285 + (-0.741) = 0.544) for controls (p = 0.019) when comparing during the intervention to before. 

Comparing after the intervention to during the intervention, on average HbA1c increased by 0.087 for cases, and increased by 0.041 (0.087 + (-0.046)) for controls, although neither change was statistically significant (p = 0.793,p = 0.920 respectively).

The interpretation for the following tables is the same. 

```{r echo=FALSE,warning=FALSE}
# bg checks
# plot by timepoint
plot <- ggplot(avgs, aes(x=timepoint,y = mean_bgchecks,group = record_id))+
  geom_line(aes(color = control)) +
  ylab("Average BG Checks") +
  xlab("Timepoint") + 
  ggtitle("Figure 2: Average BG Checks by Timepoint")
plot
# Models
# Before vs. during
bgchecks <- lme(bgchecks ~ control,random = ~1|case_id,data = bef_dur,na.action = na.omit)
fixed <- summary(bgchecks)$tTable
kable(fixed, caption = "Average BG Checks: Change from Before to During")
# During vs. after
bgchecks <- lme(bgchecks ~ control,random = ~1|case_id,data = dur_aft,na.action = na.omit)
fixed <- summary(bgchecks)$tTable
kable(fixed, caption = "Average BG Checks: Change from During to After")
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
# Days
# plot by timepoint
plot <- ggplot(avgs, aes(x=timepoint,y = mean_days,group = record_id))+
  geom_line(aes(color = control)) +
  ylab("Average Days Between Visits") +
  xlab("Timepoint") + 
  ggtitle("Figure 3: Average Days Between Visits")
plot
# Models
# Before vs. during
days <- lme(days ~ control,random = ~1|case_id,data = bef_dur,na.action = na.omit)
fixed <- summary(days)$tTable
kable(fixed, caption = "Average Days Between Visits: Change from Before to During")
# During vs. after
days <- lme(days ~ control,random = ~1|case_id,data = dur_aft,na.action = na.omit)
fixed <- summary(days)$tTable
kable(fixed, caption = "Average Days Between Visits: Change from During to After")
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
# Target BG
# plot by timepoint
plot <- ggplot(avgs, aes(x=timepoint,y = mean_target_bg,group = record_id))+
  geom_line(aes(color = control)) +
  ylab("Target BG") +
  xlab("Timepoint") + 
  ggtitle("Figure 4: Target BG")
plot
# Models
# Before vs. during
targetbg <- lme(targetbg ~ control,random = ~1|case_id,data = bef_dur,na.action = na.omit)
fixed <- summary(targetbg)$tTable
kable(fixed, caption = "Target BG: Change from Before to During")
# During vs. after
targetbg <- lme(targetbg ~ control,random = ~1|case_id,data = dur_aft,na.action = na.omit)
fixed <- summary(targetbg)$tTable
kable(fixed, caption = "Target BG: Change from During to After")
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
# High BG
# plot by timepoint
plot <- ggplot(avgs, aes(x=timepoint,y = mean_high_bg,group = record_id))+
  geom_line(aes(color = control)) +
  ylab("High BG") +
  xlab("Timepoint") + 
  ggtitle("Figure 5: High BG")
plot
# Models
# Before vs. during
highbg <- lme(highbg ~ control,random = ~1|case_id,data = bef_dur,na.action = na.omit)
fixed <- summary(highbg)$tTable
kable(fixed, caption = "High BG: Change from Before to During")
# During vs. after
highbg <- lme(highbg ~ control,random = ~1|case_id,data = dur_aft,na.action = na.omit)
fixed <- summary(highbg)$tTable
kable(fixed, caption = "High BG: Change from During to After")
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
# Low BG
# plot by timepoint
plot <- ggplot(avgs, aes(x=timepoint,y = mean_low_bg,group = record_id))+
  geom_line(aes(color = control)) +
  ylab("Low BG") +
  xlab("Timepoint") + 
  ggtitle("Figure 6: Low BG")
plot
# Models
# Before vs. during
lowbg <- lme(low_bg ~ control,random = ~1|case_id,data = bef_dur,na.action = na.omit)
fixed <- summary(lowbg)$tTable
kable(fixed, caption = "Low BG: Change from Before to During")
# During vs. after
lowbg <- lme(low_bg ~ control,random = ~1|case_id,data = dur_aft,na.action = na.omit)
fixed <- summary(lowbg)$tTable
kable(fixed, caption = "Low BG: Change from During to After")
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
# SW Visits
# plot by timepoint
plot <- ggplot(avgs, aes(x=timepoint,y = sw,group = record_id))+
  geom_line(aes(color = control)) +
  ylab("Total SW Visits") +
  xlab("Timepoint") + 
  ggtitle("Figure 7: Total SW Visits")
plot
# Models
# Before vs. during
sw <- lme(sw ~ control,random = ~1|case_id,data = bef_dur,na.action = na.omit)
fixed <- summary(sw)$tTable
kable(fixed, caption = "Total SW Visits: Change from Before to During")
# During vs. after
sw <- lme(sw ~ control,random = ~1|case_id,data = dur_aft,na.action = na.omit)
fixed <- summary(sw)$tTable
kable(fixed, caption = "Total SW Visits: Change from During to After")
```