---
title: "BBGD! Mixed Models"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(skimr)
library(knitr)
library(tidyverse)
library(caret)
library(glmnet)
library(Hmisc)
library(UpSetR)
library(lme4)
library(lmerTest)
library(car)
knitr::opts_chunk$set(echo = FALSE)
home_dir = ifelse(.Platform$OS.type != "unix",
                  "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kim Driscoll/Bring BG Down/",
                  "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kim Driscoll/Bring BG Down/")
knitr::opts_knit$set(root.dir = home_dir)
set.seed(1017)
```

```{r data}
df = read.csv("./Data_Cleaned/cleaned_df.csv",na.strings = "")
# Variable formatting
df$X = NULL
df$c_gender = factor(df$c_gender)
# Dichotomize variables
df$Total.Household.Income = factor(df$Total.Household.Income)
levels(df$Total.Household.Income) = 
  c("$90K+","<$90K","<$90K","<$90K","<$90K","<$90K","<$90K","<$90K","<$90K","$90K+","<$90K")
df$Highest.Degree.Parent.Completed = factor(df$Highest.Degree.Parent.Completed)
levels(df$Highest.Degree.Parent.Completed) = 
  c("Associate's or High School","Bachelor's or above","Bachelor's or above",
    "Associate's or High School","Bachelor's or above")
df$Parent.s.Marital.Status = factor(df$Parent.s.Marital.Status)
levels(df$Parent.s.Marital.Status) = 
  c("Single","Living with Partner","Living with Partner","Single")
df = df %>% unite(Parent.Race.Ethnicity,Parent.Ethnicity,Parent.Race,na.rm = T,remove = F)
df$Parent.Race.Ethnicity[df$Parent.Race.Ethnicity == ""] = NA
df$Parent.Race.Ethnicity = factor(df$Parent.Race.Ethnicity)
levels(df$Parent.Race.Ethnicity) = c("Ethnic Minority","Ethnic Minority","White")
# Outcome cutoffs
df$Parent.Social.Consequences.Subscale.of.HFS = 
  cut(df$Parent.Social.Consequences.Subscale.of.HFS,right = F,
      breaks = c(-Inf,9,Inf),labels = c("Not Elevated","Elevated"))
df$Parent.GAD.7.Score = 
  cut(df$Parent.GAD.7.Score,right = F,
      breaks = c(-Inf,10,Inf),labels = c("Not Elevated","Elevated"))
# Fix birthdays and age
df$c_dob[df$participant_id == 420] = "2007-05-15"
df$c_dob[df$participant_id == 429] = "2008-01-08"
df$age = as.numeric(difftime(lubridate::ymd(df$t1_date),lubridate::ymd(df$c_dob),
                             units = "days"))/365.25
```

# Table 1: Participant Characteristics

```{r results='asis'}
# Table 1
t1 = tableby(treatment_group~age+c_gender+Parent.Maintain.High.BG.of.HFS+Parent.Worry.Helplessness.Subscale.of.HFS+Parent.Social.Consequences.Subscale.of.HFS+Parent.GAD.7.Score+Private.Insurance+fe(Total.Household.Income)+fe(Parent.s.Marital.Status)+fe(Highest.Degree.Parent.Completed)+fe(Parent.Race.Ethnicity)+fe(Primary.Male.Caretaker),
             data = df[df$studyvisit == 1,])
summary(t1,pfootnote = T)
```

# Methods

All variables from Table 1 were included as covariates in model selection despite not being significantly different in univariate tests. Model selection for covariates was performed at study visit 1, because elasticnet regularization has not been implemented for longitudinal models. Time was treated as a continuous variable, and final models were re-fit using all timepoints. Elasticnet was used in place of the lasso because it is more effective when selecting between correlated variables. Optimal tuning parameters were selected to minimize root mean square error based on leave one out cross validation.

Household income and education were dichotomized into >= \$90K vs. < $90K and associate's degree or high school vs. bachelor's degree and above. Gender, income, and race were selected for all outcomes, and were thus included in the final model results (along with age, which is of clinical interest).

# Model selection

```{r}
model_select = function(outcome,df = vis_1,fam = "gaussian",pred = predictors){
  df = df[,c(outcome,pred)]
  # CV
  cv = trainControl(method = "LOOCV")
  # Fit
  form = as.formula(paste0(outcome,"~",paste0(pred,collapse = "+")))
  hit_elnet = train(form, data = df,method = "glmnet",trControl = cv,
                    na.action = "na.omit",family = fam)
  # Best
  best = hit_elnet$bestTune
  # Final model 
  lasso_df = df[complete.cases(df),]
  mod = glmnet(y = data.matrix(lasso_df[,outcome]),x = data.matrix(lasso_df[,pred]),
               alpha = as.numeric(best["alpha"]),lambda = as.numeric(best["lambda"]),
               family = fam)
  return(rownames(coef(mod))[which(round(coef(mod),3)!=0 & 
                                     rownames(coef(mod)) != "(Intercept)")])
}
# Just baseline obs. for model selection
vis_1 = df[df$studyvisit == 1,]
# Predictors
predictors = c("treatment_group","age","c_gender","Private.Insurance","Total.Household.Income",
               "Parent.s.Marital.Status","Highest.Degree.Parent.Completed",
               "Parent.Race.Ethnicity")
```

## Parent Maintain High BG of HFS

```{r}
maintain_coefs = model_select(outcome = "Parent.Maintain.High.BG.of.HFS")
kable(maintain_coefs,col.names = "Selected Variables")
```

## Parent Worry/Helplessness Subscale of HFS

```{r}
worry_coefs = model_select(outcome = "Parent.Worry.Helplessness.Subscale.of.HFS")
kable(worry_coefs,col.names = "Selected Variables")
```

## Parent Social Consequences Subscale of HFS

This outcome was dichotomized as >= 9 vs. < 9. However, the models failed to converge due to lack of variability in the outcome.

```{r eval=FALSE}
social_coefs = model_select(outcome = "Parent.Social.Consequences.Subscale.of.HFS",fam = "binomial")
```

## Parent GAD-7 Score

This outcome was dichotomized as >= 10 vs. < 10.

```{r}
gad_coefs = model_select(outcome = "Parent.GAD.7.Score",fam = "binomial")
kable(gad_coefs,col.names = "Selected Variables")
```

## Upset plot

Social consequences were not included in model selection due to model instability (very view people had elevated scores).

```{r}
# Format output
selected = do.call(rbind,list(
  predictors %in% maintain_coefs,
  predictors %in% worry_coefs,
  predictors %in% gad_coefs
))
colnames(selected) = predictors
selected = data.frame(apply(selected,2,as.numeric))
# Plot
upset(selected,nsets = ncol(selected),order.by = "freq")
```

# Model results

```{r}
fit_mod = function(outcome,pred = predictors,dat = df,fam = "gaussian"){
  form = as.formula(paste0(outcome,"~","studyvisit*treatment_group+",
                           paste0(pred,collapse = "+"),"+(1|participant_id)"))
  if(fam == "gaussian"){
    mod = lmer(form,data = dat)
  } else {
    mod = glmer(form,data = dat,family = fam)
  }
  return(mod)
}
# Selected predictors
predictors = 
  c("age","c_gender","Total.Household.Income","Parent.Race.Ethnicity")
# Fill down
df = df %>% group_by(participant_id) %>% 
  fill(all_of(predictors))
```

## Parent Maintain High BG of HFS

```{r}
mod = fit_mod(outcome = "Parent.Maintain.High.BG.of.HFS")
kable(Anova(mod,type = "III"),caption = "Type III Tests of Fixed Effects")
kable(summary(mod)$coefficients,caption = "Mixed Model Coefficients")
```

## Parent Worry/Helplessness Subscale of HFS

```{r}
mod = fit_mod(outcome = "Parent.Worry.Helplessness.Subscale.of.HFS")
kable(Anova(mod,type = "III"),caption = "Type III Tests of Fixed Effects")
kable(summary(mod)$coefficients,caption = "Mixed Model Coefficients")
```

## Parent Social Consequences Subscale of HFS

Would only converge when including age and gender.

```{r}
df$Parent.Social.Consequences.Subscale.of.HFS = 
  as.numeric(df$Parent.Social.Consequences.Subscale.of.HFS) - 1
mod = fit_mod(outcome = "Parent.Social.Consequences.Subscale.of.HFS",
              fam = "binomial",pred = predictors[c(1,2)])
kable(Anova(mod,type = "III"),caption = "Type III Tests of Fixed Effects")
kable(summary(mod)$coefficients,caption = "Mixed Model Coefficients")
```

## Parent GAD-7 Score

Would only converge when including age and gender.

```{r}
df$Parent.GAD.7.Score = as.numeric(df$Parent.GAD.7.Score)-1
mod = fit_mod(outcome = "Parent.GAD.7.Score",fam = "binomial",
              pred = predictors[c(1,2)])
kable(Anova(mod,type = "III"),caption = "Type III Tests of Fixed Effects")
kable(summary(mod)$coefficients,caption = "Mixed Model Coefficients")
```
