---
title: "BBGD! Mixed Models Sub-Sample Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(arsenal)
library(skimr)
library(knitr)
library(tidyverse)
library(caret)
library(glmnet)
library(Hmisc)
library(UpSetR)
library(GLMMadaptive)
library(lme4)
library(lmerTest)
library(emmeans)
library(grid)
library(gridExtra)
knitr::opts_chunk$set(echo = FALSE, dpi = 1200, fig.height = 12, fig.width = 12)
if (Sys.info()["sysname"] == "Windows") {
  home_dir <- "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kim Driscoll/Bring BG Down"
} else if (Sys.info()["sysname"] == "Linux") {
  home_dir <- "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kim Driscoll/Bring BG Down"
} else if (Sys.info()["sysname"] == "Darwin") {
  home_dir <- "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kim Driscoll/Bring BG Down"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
```

```{r data}
df <- read.csv("./Data_Cleaned/cleaned_df.csv", na.strings = "")
# Variable formatting
df$X <- NULL
df$c_gender <- factor(df$c_gender)
# Dichotomize variables
df$Total.Household.Income <- factor(df$Total.Household.Income)
levels(df$Total.Household.Income) <-
  c("$90K+", "<$90K", "<$90K", "<$90K", "<$90K", "<$90K", "<$90K", "<$90K", "<$90K", "$90K+", "<$90K")
df$Highest.Degree.Parent.Completed <- factor(df$Highest.Degree.Parent.Completed)
levels(df$Highest.Degree.Parent.Completed) <-
  c(
    "Associate's or High School", "Bachelor's or above", "Bachelor's or above",
    "Associate's or High School", "Bachelor's or above"
  )
df$Parent.s.Marital.Status <- factor(df$Parent.s.Marital.Status)
levels(df$Parent.s.Marital.Status) <-
  c("Single", "Living with Partner", "Living with Partner", "Single")
df <- df %>% unite(Parent.Race.Ethnicity, Parent.Ethnicity, Parent.Race, na.rm = T, remove = F)
df$Parent.Race.Ethnicity[df$Parent.Race.Ethnicity == ""] <- NA
df$Parent.Race.Ethnicity <- factor(df$Parent.Race.Ethnicity)
levels(df$Parent.Race.Ethnicity) <- c("Ethnic Minority", "Ethnic Minority", "White")
# Fix birthdays and age
df$c_dob[df$participant_id == 420] <- "2007-05-15"
df$c_dob[df$participant_id == 429] <- "2008-01-08"
df$age <- as.numeric(difftime(lubridate::ymd(df$t1_date), lubridate::ymd(df$c_dob),
  units = "days"
)) / 365.25
# Fill down
df <- df %>%
  filter(participant_id != 430) %>%
  group_by(participant_id) %>%
  fill(Total.Household.Income)
# Write for checking SAS models
write.csv(df, file = "./Data_Cleaned/cleaned_data_for_sas.csv", row.names = F)
```

# Model results

```{r}
fit_mod <- function(outcome, pred = predictors, dat = df, fam = "gaussian") {
  form <- as.formula(paste0(
    outcome, "~", "factor(studyvisit)*treatment_group+",
    paste0(pred, collapse = "+"), "+(1|participant_id)"
  ))
  zi_form <- as.formula(paste0(
    outcome, "~", "studyvisit*treatment_group+",
    paste0(pred, collapse = "+")
  ))
  fixed <- as.formula(paste0("~", paste0(pred, collapse = "+")))
  if (fam == "gaussian") {
    mod <- lmer(form, data = dat)
  } else if (fam == "zi.poisson") {
    # Complete cases
    dat <- dat[, c("participant_id", "studyvisit", "treatment_group", outcome, pred)]
    dat <- dat[complete.cases(dat), ]
    mod <-
      mixed_model(update(form, . ~ . - (1 | participant_id)),
        random = ~ 1 | participant_id, zi_fixed = ~treatment_group,
        data = dat, family = zi.poisson()
      )
  } else {
    mod <- glmer(form, data = dat, family = fam)
  }
  return(mod)
}
# Selected predictors
predictors <-
  c("age", "c_gender", "Total.Household.Income")
form <- as.formula(paste0(
  "~", "studyvisit*treatment_group+",
  paste0(predictors, collapse = "+")
))
# Fill down
df <- df %>%
  group_by(participant_id) %>%
  fill(all_of(predictors))
# Get those with high values
high_maintain = df %>% group_by(participant_id) %>%
  mutate(high_maintain = max(Parent.Maintain.High.BG.of.HFS,na.rm = T)) %>% 
  filter(high_maintain >= 7) %>% select(-high_maintain)

high_worry = df %>% group_by(participant_id) %>%
  mutate(high_worry = max(Parent.Worry.Helplessness.Subscale.of.HFS,na.rm = T)) %>% 
  filter(high_worry >= 24) %>% select(-high_worry)

high_social = df %>% group_by(participant_id) %>%
  mutate(high_social = max(Parent.Social.Consequences.Subscale.of.HFS,na.rm = T)) %>% 
  filter(high_social >= 9) %>% select(-high_social)
```

## Parent Maintain High BG of HFS

```{r}
mod <- lmer(update(form, Parent.Maintain.High.BG.of.HFS ~ . + (1 | participant_id)), data = high_maintain)
kable(summary(mod)$coefficients, caption = "Mixed Model Coefficients",digits = 3)
```

Detailed model information:

```{r}
summary(mod)
```

```{r}
# Holly! Here below is how to get model means
# Model visit as categorical for easier plotting
plot_form <- as.formula(paste0(
  "~", "factor(studyvisit)*treatment_group+",
  paste0(predictors, collapse = "+")
))
mod <- lmer(update(plot_form, Parent.Maintain.High.BG.of.HFS ~ . + (1 | participant_id)), data = df)
means <- data.frame(emmeans(mod, ~ studyvisit + factor(treatment_group)))
# Plot
fig1a <- ggplot(means, aes(x = studyvisit, y = emmean, linetype = factor(treatment_group))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(
    x = studyvisit, y = emmean,
    ymin = emmean - SE, ymax = emmean + SE
  ), width = 0.1, inherit.aes = F) +
  scale_linetype_manual(
    values = c("dashed", "solid"), name = "Group",
    labels = c("Control", "Treatment")
  ) +
  ggtitle("Caregiver Maintain High BG") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(), axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
fig1a
```

## Parent Worry/Helplessness Subscale of HFS

```{r}
mod <- lmer(update(form, Parent.Worry.Helplessness.Subscale.of.HFS ~ . + (1 | participant_id)),
  data = df
)
kable(summary(mod)$coefficients, caption = "Mixed Model Coefficients",digits = 3)
```

Detailed model information:

```{r}
summary(mod)
```

```{r}
# Holly! Here below is how to get model means
# Model visit as categorical for easier plotting
plot_form <- as.formula(paste0(
  "~", "factor(studyvisit)*treatment_group+",
  paste0(predictors, collapse = "+")
))
mod <- lmer(update(plot_form, Parent.Worry.Helplessness.Subscale.of.HFS ~ . + (1 | participant_id)),
  data = df
)
means <- data.frame(emmeans(mod, ~ studyvisit + factor(treatment_group)))
# Plot
fig1b <- ggplot(means, aes(x = studyvisit, y = emmean, linetype = factor(treatment_group))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(
    x = studyvisit, y = emmean,
    ymin = emmean - SE, ymax = emmean + SE
  ), width = 0.1, inherit.aes = F) +
  scale_linetype_manual(
    values = c("dashed", "solid"), name = "Group",
    labels = c("Control", "Treatment")
  ) +
  ggtitle("Caregiver Worry/Helplessness About Low BG") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(), axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
fig1b
```

## Parent Social Consequences Subscale of HFS

This outcome was analyzed using a zero-inflated Poisson model with random intercept for participant. Coefficients are interpreted on the log count scale. 

```{r}
mod <- mixed_model(update(form, Parent.Social.Consequences.Subscale.of.HFS ~ .),
  random = ~ 1 | participant_id, zi_fixed = ~1,
  data = df, family = zi.poisson()
)
kable(summary(mod)$coef_table, caption = "Mixed Model Coefficients",digits = 3)
```

Detailed model information:

```{r}
summary(mod)
```

```{r}
# Holly! Here below is how to get model means
# Model visit as categorical for easier plotting
plot_form <- as.formula(paste0(
  "Parent.Social.Consequences.Subscale.of.HFS~",
  "factor(studyvisit)*treatment_group+",
  paste0(predictors, collapse = "+")
))
mod <- mixed_model(plot_form,
  random = ~ 1 | participant_id, zi_fixed = ~1,
  data = df, family = zi.poisson()
)
means <- data.frame(emmeans(mod, ~ studyvisit + factor(treatment_group)))
fig1c <- ggplot(means, aes(x = studyvisit, y = emmean, linetype = factor(treatment_group))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(
    x = studyvisit, y = emmean,
    ymin = emmean - SE, ymax = emmean + SE
  ), width = 0.1, inherit.aes = F) +
  scale_linetype_manual(
    values = c("dashed", "solid"), name = "Group",
    labels = c("Control", "Treatment")
  ) +
  ggtitle("Caregiver Worry About Social Consequences") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(), axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
fig1c
```

# Combined Figure

```{r}
grid.arrange(
  arrangeGrob(fig1a + theme(legend.position = "none"),
    fig1b + theme(legend.position = "none"),
    fig1c + theme(legend.position = "none"),
    nrow = 2,
    left = textGrob("Mean Score",
      rot = 90, vjust = 1,
      gp = gpar(fontface = "bold", cex = 1.5)
    )
  ),
  bottom = textGrob("Visit",
    hjust = 1,
    gp = gpar(fontface = "bold", cex = 1.5)
  )
)
```
