---
title: "BBGD! Mixed Models Sub-Sample Analysis"
author: "Tim Vigers, Casey Sakamoto & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(Hmisc)
library(lme4)
library(lmerTest)
library(emmeans)
library(readr)
library(table1)

knitr::opts_chunk$set(echo = FALSE, dpi = 1200, fig.height = 12, fig.width = 12)
if (Sys.info()["sysname"] == "Windows") {
  home_dir <- "S:/Laura/BDC/Projects/Kim Driscoll/Bring BG Down"
} else if (Sys.info()["sysname"] == "Linux") {
  home_dir <- "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kim Driscoll/Bring BG Down"
} else if (Sys.info()["sysname"] == "Darwin") {
  home_dir <- "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kim Driscoll/Bring BG Down"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
```

```{r data, include= FALSE}
# get data and covariates
df <- read_csv("S:/Laura/BDC/Projects/Kim Driscoll/Bring BG Down/Data_Cleaned/bbgd_master_Sakamoto.csv")
analysis = df %>% select(participant_id, studyvisit, treatment_group,`Total Household Income`,c_gender,
                         `Parent Maintain High BG of HFS`, `Parent Worry/Helplessness Subscale of HFS`,
                         `Parent Social Consequences Subscale of HFS`)

# Fill down
analysis <- analysis %>%
  group_by(participant_id) %>%
  fill(`Total Household Income`)

# predictors for 1-6
# assess time, trt,, sex, income, time*trt
predictors16 = c("factor(studyvisit)","factor(treatment_group)", "factor(c_gender)", "`Total Household Income`","factor(studyvisit)*factor(treatment_group)")

# predictors 89
predictors89 = c("factor(studyvisit)", "factor(treatment_group)", "factor(c_gender)", "factor(studyvisit)*factor(treatment_group)")

#predictos10
predictors10 = c("factor(studyvisit)", "factor(treatment_group)", "factor(studyvisit)*factor(treatment_group)")
# table 1 
# mean(sd), range for times 1,678 for each score by sex
table_df = analysis %>% select(studyvisit, c_gender, 
                         `Parent Maintain High BG of HFS`, `Parent Worry/Helplessness Subscale of HFS`,
                         `Parent Social Consequences Subscale of HFS`) %>%
                        filter(studyvisit %in% c(1,6,7,8))
table_df$participant_id = factor(table_df$participant_id)
# wide to long
table_df = gather(table_df, subscale, score,`Parent Maintain High BG of HFS`:`Parent Social Consequences Subscale of HFS`)
#table1

testt1 = table_df %>% group_by(studyvisit, subscale,c_gender) %>% summarise(mean = mean(score, na.rm = T), sd = sd(score, na.rm = T), min = min(score, na.rm = T), max = max(score, na.rm = T))
```

```{r, include = FALSE}
fit_mod <- function(outcome, pred = predictors, dat = df) {
  
  form <- as.formula(paste0(
    outcome, "~", paste0(pred, collapse = "+"), "+(1|participant_id)"
  ))

    mod <- lmer(form, data = dat)
   
  return(mod)
}
```
# Methods

# Model results

```{r, include = F}
# 1
mod1_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors16, dat = analysis %>% filter(studyvisit<7))
mod1_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors16, dat = analysis %>% filter(studyvisit<7))
mod1_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors16, dat = analysis %>% filter(studyvisit<7))

a1 = kable(summary(mod1_mh)$coefficients, digits = 3)
b1 = kable(summary(mod1_wl)$coefficients, digits = 3)
c1 = kable(summary(mod1_ws)$coefficients, digits = 3)

# 2
mod2_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors16, dat = analysis)
mod2_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors16, dat = analysis)
mod2_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors16, dat = analysis)

a2 = kable(summary(mod2_mh)$coefficients, digits = 3)
b2 = kable(summary(mod2_wl)$coefficients, digits = 3)
c2 = kable(summary(mod2_ws)$coefficients, digits = 3)

# 3
mod3_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors16, dat = analysis %>% filter(studyvisit<7 & `Parent Maintain High BG of HFS` >= 7))
mod3_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(studyvisit<7 & `Parent Worry/Helplessness Subscale of HFS` >= 24))
mod3_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(studyvisit<7 & `Parent Social Consequences Subscale of HFS` >= 9))

a3 = kable(summary(mod3_mh)$coefficients, digits = 3)
b3 = kable(summary(mod3_wl)$coefficients, digits = 3)
c3 = kable(summary(mod3_ws)$coefficients, digits = 3)

# 4
mod4_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors16, dat = analysis %>% filter(`Parent Maintain High BG of HFS` >= 7))
mod4_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(`Parent Worry/Helplessness Subscale of HFS` >= 24))
mod4_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(`Parent Social Consequences Subscale of HFS` >= 9))

a4 = kable(summary(mod4_mh)$coefficients, digits = 3)
b4 = kable(summary(mod4_wl)$coefficients, digits = 3)
c4 = kable(summary(mod4_ws)$coefficients, digits = 3)

# 5
mod5_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors16, dat = analysis %>% filter(studyvisit<7 &`Parent Maintain High BG of HFS` < 7))
mod5_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(studyvisit<7 &`Parent Worry/Helplessness Subscale of HFS` < 24))
mod5_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(studyvisit<7 &`Parent Social Consequences Subscale of HFS` < 9))

a5 = kable(summary(mod5_mh)$coefficients, digits = 3)
b5 = kable(summary(mod5_wl)$coefficients, digits = 3)
c5 = kable(summary(mod5_ws)$coefficients, digits = 3)

# 6
mod6_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors16, dat = analysis %>% filter(`Parent Maintain High BG of HFS` < 7))
mod6_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(`Parent Worry/Helplessness Subscale of HFS` < 24))
mod6_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors16, dat = analysis%>% filter(`Parent Social Consequences Subscale of HFS` < 9))

a6 = kable(summary(mod6_mh)$coefficients, digits = 3)
b6 = kable(summary(mod6_wl)$coefficients, digits = 3)
c6 = kable(summary(mod6_ws)$coefficients, digits = 3)

# 8
mod8_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors89, dat = analysis %>% filter(studyvisit<7))
mod8_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors89, dat = analysis %>% filter(studyvisit<7))
mod8_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors89, dat = analysis %>% filter(studyvisit<7))

a8 = kable(summary(mod8_mh)$coefficients, digits = 3)
b8 = kable(summary(mod8_wl)$coefficients, digits = 3)
c8 = kable(summary(mod8_ws)$coefficients, digits = 3)

# 9
mod9_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors89, dat = analysis)
mod9_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors89, dat = analysis)
mod9_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors89, dat = analysis)

a9 = kable(summary(mod9_mh)$coefficients, digits = 3)
b9 = kable(summary(mod9_wl)$coefficients, digits = 3)
c9 = kable(summary(mod9_ws)$coefficients, digits = 3)

# 10 1
mod10_mh = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors10, dat = analysis%>% filter(studyvisit<7 & c_gender == 1))
mod10_wl = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors10, dat = analysis%>% filter(studyvisit<7  & c_gender == 1))
mod10_ws = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors10, dat = analysis%>% filter(studyvisit<7  & c_gender == 1))

a10 = kable(summary(mod10_mh)$coefficients, digits = 3)
b10 = kable(summary(mod10_wl)$coefficients, digits = 3)
c10 = kable(summary(mod10_ws)$coefficients, digits = 3)

# 10 2
mod10_mh2 = fit_mod(outcome = "`Parent Maintain High BG of HFS`", pred = predictors10, dat = analysis%>% filter(studyvisit<7 & c_gender == 2))
mod10_wl2 = fit_mod(outcome = "`Parent Worry/Helplessness Subscale of HFS`", pred = predictors10, dat = analysis%>% filter(studyvisit<7  & c_gender == 2))
mod10_ws2 = fit_mod(outcome = "`Parent Social Consequences Subscale of HFS`", pred = predictors10, dat = analysis%>% filter(studyvisit<7  & c_gender == 2))

a102 = kable(summary(mod10_mh2)$coefficients, digits = 3)
b102 = kable(summary(mod10_wl2)$coefficients, digits = 3)
c102 = kable(summary(mod10_ws2)$coefficients, digits = 3)
```

## 1. LMM with times 1-6

### Maintain High

```{r}
a1
```

### Worry Low

```{r}
b1
```

### Worry Social

```{r}
c1
```

## 2. LMM with times 1-8

### Maintain High

```{r}
a2
```

### Worry Low

```{r}
b2
```

### Worry Social

```{r}
c2
```

## 3. LMM with times 1-6 and exceeding clinical cutpoints

### Maintain High

```{r}
a3
```

### Worry Low

```{r}
b3
```

### Worry Social

```{r}
c3
```

## 4. LMM with times 1-8 and exceeding clinical cutpoints

### Maintain High

```{r}
a4
```

### Worry Low

```{r}
b4
```

### Worry Social

```{r}
c4
```

## 5. LMM with times 1-6 and NOT exceeding clinical cutpoints

### Maintain High

```{r}
a5
```

### Worry Low

```{r}
b5
```

### Worry Social

```{r}
c5
```

## 6. LMM with times 1-8 and NOT exceeding clinical cutpoints

### Maintain High

```{r}
a6
```

### Worry Low

```{r}
b6
```

### Worry Social

```{r}
c6
```

## 7.

```{r}
testt1
```

## 8. LMM with times 1-6 no income adjust

### Maintain High

```{r}
a8
```

### Worry Low

```{r}
b8
```

### Worry Social

```{r}
c8
```

## 9. LMM with times 1-8 no income adjust

### Maintain High

```{r}
a9
```

### Worry Low

```{r}
b9
```

### Worry Social

```{r}
c9
```

## 10. LMM with times 1-6 no income adjust split up by sex

NOTE: couldnt find a data dictionary anywhere so I left the treatment and sex variables in original format

### sex == 1

#### Maintain High

```{r}
a10
```

#### Worry Low

```{r}
b10
```

#### Worry Social

```{r}
c10
```

### sex == 2
#### Maintain High

```{r}
a102
```

#### Worry Low

```{r}
b102
```

#### Worry Social

```{r}
c102
```