---
title: "Functional Data Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: beamer_presentation
header-includes:
  \usepackage{bm}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(patchwork)
library(splines)
library(gam)
data(Wage,package = "ISLR")
data(cancer, package="survival")
```

```{r}
dat = read.csv("/Users/timvigers/Dropbox/Work/Presentations/BDC/for_jm_v3_kaci.csv")
par(mfrow=c(2,4),oma=c(0,0,2,0),mar=c(3,2,2,1))

##plot 1
patient<-subset(dat,dat$idn==320)
patient<-patient[order(patient$visit_num),] 
plot(patient$tac_months,patient$TAC,xlim=c(1,48),ylim=c(0,20),xlab="Months Post-Tx",ylab="TAC Trough Level",xaxt="n")
axis(1,at=c(1,3,6,12,24,36,48))
lines(patient$tac_months,patient$TAC)
abline(v=patient$left_months[1],lty=2)
abline(v=patient$right_months[1])
##plot 2
patient<-subset(dat,dat$idn==14)
patient<-patient[order(patient$visit_num),] 
plot(patient$tac_months,patient$TAC,xlim=c(1,48),ylim=c(0,20),ylab="",xlab="",xaxt="n")
axis(1,at=c(1,3,6,12,24,36,48))
lines(patient$tac_months,patient$TAC)
abline(v=patient$left_months[1],lty=2)
abline(v=patient$right_months[1])
##plot 3
patient<-subset(dat,dat$idn==189)
patient<-patient[order(patient$visit_num),] 
plot(patient$tac_months,patient$TAC,xlim=c(1,48),ylim=c(0,20),ylab="",xlab="",xaxt="n")
axis(1,at=c(1,3,6,12,24,36,48))
lines(patient$tac_months,patient$TAC)
abline(v=patient$left_months[1],lty=2)
abline(v=patient$right_months[1])
##plot 4
patient<-subset(dat,dat$idn==465)
patient<-patient[order(patient$visit_num),] 
plot(patient$tac_months,patient$TAC,xlim=c(1,48),ylim=c(0,20),ylab="",xlab="",xaxt="n")
axis(1,at=c(1,3,6,12,24,36,48))
lines(patient$tac_months,patient$TAC)
abline(v=patient$left_months[1],lty=2)
abline(v=patient$right_months[1])
mtext('4 Random Individuals with dnDSA', side=3, line=0, outer=TRUE,cex=0.8,font=2)
```

## What Is Functional Data Analysis (FDA)?

- Most statistics we are familiar with use some sort of parametric model $p(x|\bm{\theta})$, and we try to estimate the parameters of *p*.
- For example, we might assume the data are normally distributed $N(\mu,\sigma^2)$ and use the sample mean $\bar{x}$ as an estimate of the location parameter $\mu$.
- Sometimes, however, we do not want to assume a density function in advance, and are interested in estimating $p$ itself (the smooth function that we assume generates our data).

## Goals and Uses

- The goals of FDA are essentially the same as other areas of statistics:
1. Represent data to aid in analysis and to highlight characteristics
2. Study patterns and sources of variation
3. Explain variation in an outcome using independent variables

- Common uses of FDA:
- Growth data
- Imaging data 
- Wearable device data
- Accelerometry (physical activity) 
- Heart rate
- **Continuous glucose monitoring** (CGM)

## Semi- and Non-Parametric Regression (Scatterplot Smoothing)

- The basic linear model is: $y_i= \beta_0+\beta_1 x_i+\epsilon_i$ for $1\leq i \leq n$
- Each predictor in our model has a simple form, and the overall shape is determined by the model rather than the data (e.g. a straight line in a linear model, one bend in a quadratic model, etc.)
- The non-parametric version of this model is: $y_i = f(x_i)+\epsilon_i$
- $f(x_i)$ is an arbitrary function.
- The only constraint is that $f(x_i)$ must be a smooth function, so this model is much more flexible.
- A semi-parametric model essentially uses parametric models for some predictors, and non-parametric models for others.

## Penalized Splines

- A natural question at this point is: "What do you mean $f(x_i)$ is an arbitrary smooth function? Can you give us an example?"
- A penalized spline model is one relatively simple option for $f(x_i)$: $$f(x)=\beta_0+\beta_1 x + \sum_{k=1}^K{u_k(x-\kappa_k)_+}$$ where $x_+\equiv max(x,0)$ and the $\kappa_ks$ are pre-set values that we choose.
- In other words, we are fitting multiple short lines that meet at our $\kappa_ks$ (often called "knots").

## Penalized Splines (for Stats Nerds)

- Estimation of the model coefficients is a fairly straightforward constrained optimization problem where we want to minimize $$\sum_{i=1}^n{(y_i - f(x_i))^2}\text{ with }\sum_{k=1}^K{u_k^2}\leq C$$

- This is equivalent to penalized least squares with smoothing parameter $\lambda\geq0$, where we want to minimize: $$\sum_{i=1}^n{(y_i - f(x_i))^2}+\lambda\sum_{k=1}^K{u_k^2}$$

## Polynomial Examples

```{r}
l = ggplot(Wage,aes(x=age,y=wage)) +
  geom_point(alpha=0.1) + 
  geom_smooth(method='lm',formula=y~x,se=F,color="blue",size=2) +
  theme_bw() + ggtitle("Linear") +
  xlab("Age") + ylab("Wage")

q = ggplot(Wage,aes(x=age,y=wage)) +
  geom_point(alpha=0.1) + 
  geom_smooth(method='lm',formula=y~x+I(x^2),se=F,color="blue",size=2) +
  theme_bw() + ggtitle("Quadratic") +
  xlab("Age") + ylab("Wage")

(l + q)
```

## Splines

```{r}
age_grid = seq(min(Wage$age),max(Wage$age))
fit = lm(wage~bs(age,knots=c(25,40,60)),data = Wage)
pred = predict(fit,newdata = list(age = age_grid))
pred = data.frame(cbind(age_grid,pred))
ggplot(pred,aes(x=age_grid,y=pred)) + 
  geom_line(color="blue",size=2) +
  geom_point(data = Wage,aes(x=age,y=wage),inherit.aes = F,alpha=0.1) +
  theme_bw() + ggtitle("Cubic Splines, 3 Knots") +
  xlab("Age") + ylab("Wage")
```

## Generalized Additive Models (GAMs)

- GAMs are a non-parametric extension of generalized linear models.
- Generalized linear models relax some of the assumptions of a linear model.
- We replace $E(\bm{Y}|\bm{X})=\sum_1^p\beta_j X_j$ with the additive model $E(\bm{Y}|\bm{X})=\sum_1^p s_j(X_j)$
- The conditional mean of the response is modeled using the sum of smooth functions of the predictors. 
- We can think of splines are an extension of simple linear regression (1 predictor) and GAMs as an extension of multiple linear regression.
- For the wage dataset we can adjust the model for year and education: $wage = \beta_0+f_1(year)+f_2(age)+f_3(education)+\epsilon$

## Wage GAM

```{r}
gam = gam(wage~s(year,4)+s(age,5)+education,data = Wage)
par(mfrow=c(1,3))
plot(gam)
```

## GAM Summary

```{r}
summary(gam)
```

## Application to Survival Analysis

```{r}
surv_gam = gam(tac_months~factor(hlamis)+factor(Ethnic_group)+s(age),
           family=cox.ph(),data=dat,weights=dsa_yn)

summary(surv_gam)


np <- 300
j <- 320
newd <- data.frame(time=seq(0,3000,length=np))
dname <- names(dat)
for (n in dname) newd[[n]] <- rep(dat[[n]][j],np)
newd$time <- seq(0,3000,length=np)
fv <- predict(surv_gam,newdata=newd,type="response",se=TRUE)
plot(newd$time,fv$fit,type="l",ylim=c(0,1),xlab="time",ylab="survival")
lines(newd$time,fv$fit+2*fv$se.fit,col=2)
lines(newd$time,fv$fit-2*fv$se.fit,col=2)


patient<-subset(dat,dat$idn==320)
patient<-patient[order(patient$visit_num),] 
plot(patient$tac_months,patient$TAC,xlim=c(1,48),ylim=c(0,20),xlab="Months Post-Tx",ylab="TAC Trough Level",xaxt="n")
axis(1,at=c(1,3,6,12,24,36,48))
lines(patient$tac_months,patient$TAC)
abline(v=patient$left_months[1],lty=2)
abline(v=patient$right_months[1])
```

## References

1. Ramsay, J. O., and B. W. Silverman. Functional Data Analysis. 2nd ed. Springer Series in Statistics. New York: Springer, 2005.

2. Harezlak J, Ruppert D, Wand MP. Semiparametric Regression with R. 1st ed. 2018. Springer New York: Imprint: Springer; 2018. doi:10.1007/978-1-4939-8853-2

3. James G, Witten D, Hastie T, Tibshirani R, eds. An Introduction to Statistical Learning: With Applications in R. Springer; 2013.

4. {pammtools}: Survival Analysis Using Generalized Additive Mixed Models (Andreas Bender). www.youtube.com, https://www.youtube.com/watch?v=ZvQH0lBDwWc. Accessed 1 June 2021.