---
title: "SEEDS Bot Hunting"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lubridate)
library(DT)
library(stringdist)
library(SimplyAgree)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
```

```{r}
df = read.csv("./Laurel Messer/SEEDS/Data_Cleaned/Complete_data_1-12-2022.csv",na.strings = "")
```

# Time to complete

```{r results='asis',warning=FALSE}
# Time at end of consent to time at end of survey
df$time = difftime(mdy_hm(df$Survey.Timestamp.1),
                   mdy_hm(df$Survey.Timestamp),units = "mins")
# Table
t1 = tableby(~notest(time,"Nmiss","median","q1q3"),df)
summary(t1)
```

## Time < 5 minutes

```{r}
df %>% filter(time < 5) %>%
  select(First.name,Last.name,Email.address,Organization.,Practice.name.,
         Practice.setting.,time) %>% arrange(time) %>% datatable(.)
df$TIM_verification_Real...1..quest...2..fake...3[df$time < 5] = 3
```

# Matching names to email

Names and emails (converted to lower case with numbers and punctuation removed) were converted to q-grams and the Euclidean distance between the two was calculated. For full names, 2-grams were used, and for initials single letters were used. 

A name distance > 4 was chosen somewhat arbitrarily as exclusion criterion, but it appeared to capture obviously fake emails without accidentally excluding those with initials in their email address. 

```{r}
# Emails, names, and initials 
emails = tolower(sub("@.*","",df$Email.address))
emails = gsub("[^a-zA-Z]", "", emails)
names = tolower(paste0(df$First.name,df$Last.name))
names = gsub("[^a-zA-Z]", "", names)
initials = tolower(paste0(substr(df$First.name,1,1),substr(df$Last.name,1,1)))
names_emails = cbind(names,initials,emails)
# Split into 2-grams and calculate distance
max_length = max(c(sapply(names,nchar),sapply(emails,nchar)),na.rm = T)
df$dist_name_2gram = apply(names_emails,1,function(r){
  q = data.frame(qgrams(r["names"],r["emails"],q = 2))
  # Normalize to length
  q[,(ncol(q)+1):max_length] = 0
  as.numeric(dist(q))
})
df$dist_initials_2gram = apply(names_emails,1,function(r){
  q = data.frame(qgrams(r["initials"],r["emails"],q = 1))
  # Normalize to length
  q[,(ncol(q)+1):max_length] = 0
  as.numeric(dist(q))
})
# Print
df %>% select(First.name,Last.name,Email.address,dist_name_2gram) %>% 
  filter(dist_name_2gram > 4) %>%
  arrange(desc(dist_name_2gram)) %>% datatable(.)
# Mark as fake
df$TIM_verification_Real...1..quest...2..fake...3[df$dist_name_2gram > 4] = 3
```

# Identical responses

There were no identical responses across all questions except for incomplete records. 

## Free text

```{r}
df %>% count(X1..Do.you.have.any.other.comments.insights.about.the.Tools.listed.above.that.would.help.us.create.meaningful.Tools..Please.be.as.descriptive.as.possible.) %>% arrange(desc(n)) %>% filter(n > 1) %>% kable(.)

df %>% count(X2..In.addition.to.the.Tools.described.above..what.other.support.or.resources.would.help.you.utilize.diabetes.devices.more.in.your.clinical.practice..Please.be.as.descriptive.as.possible.) %>% arrange(desc(n)) %>% filter(n > 1) %>% kable(.)

df %>% count(X3..Imagine.if.you.your.practice.could.obtain.personalized.support.for.diabetes.devices.from.a.specialty.diabetes.center..Please.describe.in.as.much.detail.as.possible.what.that.would.ideally.entail.Example..Would.you.prefer.one.time.set.up.assistance..periodic.check.ins..or.more.intensive.communication..What.would.be.the.most.important.topics.for.support..workflow..how.to.adjust.insulin.doses..device.training..patient.education..etc....) %>% arrange(desc(n)) %>% filter(n > 1) %>% kable(.)

df %>% count(X4..Time.management.is.an.enormous.issue.for.clinical.care.appointments..In.what.ways.could.you.envision.an.Online.Diabetes.Device.Pathway.helping.with.time.management.during.or.after.clinical.care.visits.) %>% arrange(desc(n)) %>% filter(n > 1) %>% kable(.)
```

# Laurel, Greg, and Tim (with Casey's help) combined

```{r}
# Combine Tim and Casey
tim = read.csv("./Laurel Messer/SEEDS/Data_Cleaned/bots_flagged_TV2022-01-14.csv",
               na.strings = "") %>% arrange(Record.ID)
casey = read.csv("./Laurel Messer/SEEDS/Data_Cleaned/bots_flagged_CS2022-01-14.csv",
                 na.strings = c("","NA")) %>% arrange(Record.ID)
# Merge
laurel = read.csv("./Laurel Messer/SEEDS/Data_Cleaned/Complete_data_1-12-2022_Laurel.csv")
greg = read.csv("./Laurel Messer/SEEDS/Data_Cleaned/Complete_data_1-12-2022_Greg.csv")
final = full_join(laurel %>% select(Record.ID,LAUREL_Real...1..quest...2..fake...3),
                  greg %>% select(Record.ID,GREG_verification_Real...1..quest...2..fake...3),
                  by = "Record.ID") %>%
  full_join(.,tim %>% select(Record.ID,TIM_verification_Real...1..quest...2..fake...3),
            by = "Record.ID")
colnames(final) = c("ID","Laurel","Greg","Tim")
# Compare
tim_greg = agree_test(final$Tim,final$Greg)
greg_laurel = agree_test(final$Greg,final$Laurel)
tim_laurel = agree_test(final$Tim,final$Laurel)
# Add up
final$Sum = rowSums(final[,2:ncol(final)])
final$Final = cut(final$Sum,c(-Inf,3,8,Inf),labels = c("Real","Questionable","Fake"))
# Add email back
final$Email = laurel$Email.address[match(final$ID,laurel$Record.ID)]
final = final %>% select(ID,Email,everything())
# Write
write.csv(final,file = "./Laurel Messer/SEEDS/Data_Cleaned/SEEDS data flagged 2022-01-18.csv",
          row.names = F,na = "")
```

Laurel and Greg agreed approximately `r round(as.numeric(greg_laurel$ccc.xy[1])*100,1)`% of the time (`r paste0(round(as.numeric(greg_laurel$ccc.xy[2])*100,1),"% - ",round(as.numeric(greg_laurel$ccc.xy[3])*100,1),"%")`), Laurel and Tim agreed approximately `r round(as.numeric(tim_laurel$ccc.xy[1])*100,1)`% of the time (`r paste0(round(as.numeric(tim_laurel$ccc.xy[2])*100,1),"% - ",round(as.numeric(tim_laurel$ccc.xy[3])*100,1),"%")`), and Tim and Greg agreed approximately `r round(as.numeric(tim_greg$ccc.xy[1])*100,1)`% of the time (`r paste0(round(as.numeric(tim_greg$ccc.xy[2])*100,1),"% - ",round(as.numeric(tim_greg$ccc.xy[3])*100,1),"%")`).