---
title: "A1C by Tech Trends 2016 vs 2020"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(readxl)
library(lubridate)
library(table1)
library(nlme) # mixed models
library(emmeans)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Todd Alonso/A1c trends 2017 to 2021 by tech"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)

#import data
a1c_tech = read_excel("S:/Laura/BDC/Projects/Todd Alonso/A1c trends 2017 to 2021 by tech/Data raw/Change in A1c 2017 to 2021 by tech_only encounters to be used for analysis.xlsx")
```

```{r data clean, include=FALSE}
# stratify into oct-march 2016-2017 and oct-march 2020-2021
analysis = a1c_tech %>% 
  mutate(time_group = case_when(year(VisitDate) == 2016 | year(VisitDate) == 2017 ~ "2016-2017",
                                year(VisitDate) == 2020 | year(VisitDate) == 2021 ~ "2020-2021"),
         tech_group = case_when(tolower(InsulinRegimen) == "insulin pump" & CGM == "Yes" ~ "CGM + Pump",
                                tolower(InsulinRegimen) == "insulin pump" & CGM != "Yes" ~ "Pump Alone",
                                tolower(InsulinRegimen) != "insulin pump" & CGM == "Yes" ~ "CGM Alone",
                                TRUE ~ "None")) %>% 
  select(EPICMRN, VisitDate, Age_A1cResultDate, A1c_Value, tech_group, InsulinRegimen, CGM, time_group, DiabetesDuration_A1cResultDate) 

# factor tech type for order
analysis$tech_group = factor(analysis$tech_group, levels = c("CGM + Pump", "CGM Alone", "Pump Alone", "None"))

# check for multicollinearity between age and diabetes duration
cor(analysis$Age_A1cResultDate, a1c_tech$DiabetesDuration_A1cResultDate) # 0.58 correlation, pretty strong


# vif seems ok though
# test_mod = lm(A1c_Value ~ Age_A1cResultDate + DiabetesDuration_A1cResultDate + time_group + tech_group, data = analysis)
# test_mod2 = lm(A1c_Value ~ Age_A1cResultDate  + time_group + tech_group, data = analysis)
# 
# vif(test_mod)
# vif(test_mod2)
```

```{r analysis, include=FALSE}
# table 1
a1c_table1 = table1(~ A1c_Value + Age_A1cResultDate + DiabetesDuration_A1cResultDate | tech_group*time_group,
                    render.continuous = c(.="Mean (SD)"),
                     data = analysis)

# Fit random intercept model
a1c_mod = lme(A1c_Value ~ Age_A1cResultDate + DiabetesDuration_A1cResultDate + tech_group + time_group + tech_group*time_group,
            random = ~1|EPICMRN,
            data = analysis,
            na.action = na.omit)

a1c_mod2 = lme(A1c_Value ~ Age_A1cResultDate +  tech_group + time_group + tech_group*time_group,
            random = ~1|EPICMRN,
            data = analysis,
            na.action = na.omit)
# Anova
mod_anova = anova.lme(a1c_mod, type="marginal")
mod2_anova =  anova.lme(a1c_mod2, type="marginal")
# Means
mod_means = emmeans(a1c_mod,specs=pairwise ~ tech_group:time_group, adjust="tukey")
  
# Plot
a1c_predict = cbind(analysis, a1cpred = predict(a1c_mod))
```

# Outstanding Data Questions/ Remarks

* Not much demographics provided in dataset; let me know if you'd like anything added

# Methods

A1c values were modeled with a Linear Mixed Model with time groups(2016-2017, 2020-2021) and tech groups(CGM+Pump, CGM only, Pump Only, None) and an interaction term of time and tech, adjusting for age and diabetes duration. A random intercept for subject was added to account for subjects with repeated measurements. Contrasts of estimated marginal means (least squares means) were used to determine the difference within tech groups at each time point.

# Analysis

## Demographics Table
```{r Table 1}
a1c_table1
```

## Model Fit and Contrasts

Adjusting for age and diabetes duration, there is a significant difference in A1c within the different tech groups across the 2016-2017 and 2020-2021 time periods (p < 0.0001).

The average A1c decreased by 0.44 (95% CI : 0.29, 0.58) in the CGM + Pump group (p < 0.0001), increased by 0.31 (95% CI: 0.09, 0.53) in the Pump Only groups (p = 0.022), and increased by 0.45 (95% CI: 0.26, 0.64) in the No Tech group (p < 0.0001). There was no significant difference in the time periods within the CGM only group (p = 0.84) 


```{r model}
kable(mod_anova,digits = 3,caption = "Test of Overall Effect")

kable(mod_means$contrasts[c(4,11,17,22),],digits = 3,caption = "Timepoint Means")

ggplot(data = a1c_predict,aes(x = time_group, y = a1cpred, group = tech_group, color = tech_group))+
  geom_smooth(method = lm, se = F,na.rm = T) + 
  xlab("Time Groups") + ylab("Predicted Mean A1c") + theme_classic() 
```


```{r plot, include=FALSE}
a1ctech_plot_loess = ggplot(data = analysis, aes(x = Age_VisitDate, y = A1c_Value, color = tech_group, linetype = time_group)) + geom_smooth(method = "loess", se = FALSE) + theme_classic() + labs(title = "Loess")

a1ctech_plot_gam = ggplot(data = analysis, aes(x = Age_VisitDate, y = A1c_Value, color = tech_group, linetype = time_group)) + geom_smooth(method = "gam", se = FALSE) + theme_classic() + labs(title = "gam")

a1ctech_plot_lm = ggplot(data = analysis, aes(x = Age_VisitDate, y = A1c_Value, color = tech_group, linetype = time_group)) + geom_smooth(method = "lm", se = FALSE) + theme_classic() + labs(title = "lm")

# a1ctech_plot_loess
# a1ctech_plot_gam
# a1ctech_plot_lm
```
