---
title: "DCCT/EDIC Proteomics Gene Set Enrichment Analysis"
author: "Laura Pyle & Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
library(enrichplot)
knitr::opts_knit$set(root.dir = "~/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/BDC/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01")
```

# Protein-to-gene linkage

We used the `biomaRt` R package to link UniProt IDs to Entrez Gene IDs, based on Ensembl's most recent version of the Human Genes BioMart. 

See `~/GitHub/BDC-Code/Janet Snell-Bergeon/DCCT and EDIC/create_omics_linkage_file.R` for additional documentation.

# GSEA

```{r data cleaning}
# Import results
res <- read.csv("Proteomics/Reports/DCCT EDIC CARV.csv", na.strings = c("", "NA"))
res <- res[!is.na(res$logFC), ]
# Import annotations
load("./Data_Raw/annotations.RData")
# Add to dataframe
res$entrez <- protein_anotation$Entrez[match(res$protid, protein_anotation$UniProt)]
```

A total of `r sum(is.na(res$entrez))` proteins were not linked to an Entrez ID (out of `r nrow(res)`).

## Unadjusted

```{r}
unadjusted <- res$logFC[!is.na(res$entrez)]
names(unadjusted) <- res$entrez[!is.na(res$entrez)]
unadjusted <- sort(unadjusted, decreasing = T)
dups <- duplicated(names(unadjusted))
y <- gsePathway(unadjusted[!dups], pvalueCutoff = 1, pAdjustMethod = "BH", verbose = FALSE)
x <- setReadable(y, "org.Hs.eg.db", "ENTREZID")
```

A total of `r sum(dups)` gene IDs were duplicated, so those with lower effect sizes were removed for GSEA analysis.

### Dotplot

```{r}
dotplot(y)
```

### Ridgeplot

```{r fig.height=10,fig.width=8}
ridgeplot(y)
```

### Network plot

```{r}
cnetplot(x)
```

### UpSet plot

```{r}
upsetplot(y) + ylab("logFC")
```

### GSEA plot

```{r}
gseaplot2(y, geneSetID = 1:5)
```

## Adjusted for GROUP, VisitAge, SEX, Visit, Duration

```{r}
res = read.csv("Proteomics/Reports/DCCT EDIC CARV adjusted GROUP VisitAge SEX Visit Duration.csv",
               na.strings = c("","NA"))
res <- res[!is.na(res$logFC), ]
res$entrez <- protein_anotation$Entrez[match(res$protid, protein_anotation$UniProt)]

adjusted <- res$logFC[!is.na(res$entrez)]
names(adjusted) <- res$entrez[!is.na(res$entrez)]
adjusted <- sort(adjusted, decreasing = T)
dups <- duplicated(names(adjusted))
y <- gsePathway(adjusted[!dups], pvalueCutoff = 1, pAdjustMethod = "BH", verbose = FALSE)
x <- setReadable(y, "org.Hs.eg.db", "ENTREZID")
```

A total of `r sum(dups)` gene IDs were duplicated, so those with lower effect sizes were removed for GSEA analysis.

### Dotplot

```{r}
dotplot(y)
```

### Ridgeplot

```{r fig.height=10,fig.width=8}
ridgeplot(y)
```

### Network plot

```{r}
cnetplot(x)
```

### UpSet plot

```{r}
upsetplot(y) + ylab("logFC")
```

### GSEA plot

```{r}
gseaplot2(y, geneSetID = 1:5)
```
