---
title: "EDIC Lipidomics"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_knit$set(root.dir = "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/EDIC/Lipidomics")
knitr::opts_knit$set(root.dir = "~/Dropbox/Work/Janet Snell-Bergeon/EDIC/Lipidomics")
library(PCAtools)
library(readxl)
library(tidyverse)
library(limma)
library(knitr)
```

```{r include=FALSE}
# Import
intensities = read.csv("./Data_Raw/EDIC lipidomics data matrix_Skyline_2021.02.25.csv",stringsAsFactors = F)
# Get compound info in separate frame
compound_info = intensities[,"LipidSpecies"]
intensities$LipidSpecies = NULL
# Transpose and format
intensities = as.data.frame(t(intensities))
names(compound_info) = colnames(intensities)
intensities$sample_id = gsub("\\.","-",rownames(intensities))
intensities = intensities %>% select(sample_id,everything())
intensities[,2:ncol(intensities)] = lapply(intensities[,2:ncol(intensities)],function(c){log(as.numeric(c))})
# Import sample injection order
injection_order = read_excel("./Data_Raw/EDIC lipidomics run order.xlsx")
colnames(injection_order) = "sample_id"
injection_order$sample_id = sub("QC-EDIC","EDIC-QC",injection_order$sample_id)
injection_order$sample_id[is.na(injection_order$sample_id)] = paste0("Failed-QC-",1:sum(is.na(injection_order$sample_id))) # For now assume that the missing are QCs
injection_order$injection_order = 1:nrow(injection_order)
# Batch information
batches = diff(c(0,grep("QC",injection_order$sample_id),nrow(injection_order)))
injection_order$batch = rep(1:length(batches),times = batches)
# Remove duplicates and QC for now
injection_order = injection_order[-c(which(duplicated(injection_order$sample_id)),grep("Failed-QC",injection_order$sample_id)),]
# Get sample info
sample.info = read.csv("./Data_Raw/QEDIC_WP_DDA_60SPD_StudyInformat.csv")
# Get sample IDs, add leading zeroes and "EDIC" to number
sample.info$sample_id = sample.info$ProteomicsSampleID
sample.info$sample_id = paste0("EDIC-",str_pad(sample.info$sample_id,3,"left",0))
sample.info = sample.info %>% select(sample_id,everything())
# Combine
sample.info = merge(injection_order,sample.info,all = T,by = "sample_id")
sample.info = sample.info[order(sample.info$injection_order),]
# Try two batches
sample.info$batch2 = ifelse(sample.info$batch < 6,1,2)
# Plot by HARD status
sample.info$class = sample.info$HARD
sample.info$class[which(is.na(sample.info$class))] = "QC"
sample.info$class = factor(sample.info$class)
# Column names
colnames(intensities)[2:ncol(intensities)] = compound_info
```

# Moderated T Tests

## CARV

### All

Moderated t tests were performed to test for differences in lipids between those who experienced CARV and those who did not. The results immediately below were not stratified by sex or storage time.

```{r results='asis'}
# limma stratified by storage time
sample.info$storage_time = sample.info$VisitAge - sample.info$AGE00
sample.info$storage_time = cut(sample.info$storage_time,2,labels = c("< 14",">= 14"))
low = sample.info[sample.info$storage_time=="< 14",]
high = sample.info[sample.info$storage_time==">= 14",]
# Make design matrix
carv <- model.matrix(~CARV,sample.info[!is.na(sample.info$storage_time),])
# Unadjusted models
y <- t(intensities[intensities$sample_id %in% sample.info[!is.na(sample.info$storage_time),"sample_id"],2:ncol(intensities)])
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
kable(topTable(fit,coef = 2))
```

#### Adjusted for sex, age, diabetes duration, treatment group

```{r results='asis'}
# Make design matrix
carv <- model.matrix(~CARV+SEX+AGE+DURATION+GROUP,sample.info)
# Unadjusted models
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y))
kable(head(res,10),caption = "Not adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05)` significant lipids.

```{r results='asis'}
# Make design matrix
carv <- model.matrix(~CARV+SEX+AGE+DURATION+GROUP+storage_time,sample.info)
# Unadjusted models
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y))
kable(head(res,10),caption = "Adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05)` significant lipids.

### Stratified by Above or Below Median Storage Time

```{r results='asis'}
# Make design matrices
carv_low <- model.matrix(~CARV,low)
carv_high <- model.matrix(~CARV,high)
# Unadjusted models
y_low <- t(intensities[intensities$sample_id %in% low$sample_id,2:ncol(intensities)])
y_high <- t(intensities[intensities$sample_id %in% high$sample_id,2:ncol(intensities)])
fit_low <- lmFit(y_low,carv_low)
fit_low <- eBayes(fit_low)
fit_high <- lmFit(y_high,carv_high)
fit_high <- eBayes(fit_high)
# Results
low_res = topTable(fit_low,coef = 2,number = nrow(y_low))
kable(head(low_res,10),caption = "Low Storage Time")
high_res = topTable(fit_high,coef = 2,number = nrow(y_high))
kable(head(high_res,10),caption = "High Storage Time")
```

There were `r sum(low_res$P.Value < 0.05)` significant lipids in the low group and `r sum(high_res$P.Value < 0.05)` in the high group.

### Stratified by Sex

```{r results='asis'}
# Split sample info
male = sample.info[sample.info$SEX == "M",]
female = sample.info[sample.info$SEX == "F",]
# Make design matrices
carv_male <- model.matrix(~CARV,male)
carv_female <- model.matrix(~CARV,female)
# Unadjusted models
y_male <- t(intensities[intensities$sample_id %in% male$sample_id,2:ncol(intensities)])
y_female <- t(intensities[intensities$sample_id %in% female$sample_id,2:ncol(intensities)])
fit_male <- lmFit(y_male,carv_male)
fit_male <- eBayes(fit_male)
fit_female <- lmFit(y_female,carv_female)
fit_female <- eBayes(fit_female)
# Results
male_res = topTable(fit_male,coef = 2,number = nrow(y_low))
kable(head(male_res,10),caption = "Males")
female_res = topTable(fit_female,coef = 2,number = nrow(y_high))
kable(head(female_res,10),caption = "Females")
```

There were `r sum(male_res$P.Value < 0.05)` significant lipids in the male group and `r sum(female_res$P.Value < 0.05)` in the female group.

## HARD

### All

Moderated t tests were performed to test for differences in lipids between those who experienced HARD and those who did not. The results immediately below were not stratified by sex or storage time.

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~HARD,sample.info[!is.na(sample.info$storage_time),])
# Unadjusted models
y <- t(intensities[intensities$sample_id %in% sample.info[!is.na(sample.info$storage_time),"sample_id"],2:ncol(intensities)])
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
kable(topTable(fit_low,coef = 2))
```

### Adjusted for sex, age, diabetes duration, treatment group

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~HARD+SEX+AGE+DURATION+GROUP,sample.info)
# Unadjusted models
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y))
kable(head(res,10),caption = "Not adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05)` significant lipids.

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~HARD+SEX+AGE+DURATION+GROUP+storage_time,sample.info)
# Unadjusted models
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y))
kable(head(res,10),caption = "Adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05)` significant lipids.

### Stratified by Above or Below Median Storage Time

```{r results='asis'}
# Make design matrices
hard_low <- model.matrix(~HARD,low)
hard_high <- model.matrix(~HARD,high)
# Unadjusted models
y_low <- t(intensities[intensities$sample_id %in% low$sample_id,2:ncol(intensities)])
y_high <- t(intensities[intensities$sample_id %in% high$sample_id,2:ncol(intensities)])
fit_low <- lmFit(y_low,hard_low)
fit_low <- eBayes(fit_low)
fit_high <- lmFit(y_high,hard_high)
fit_high <- eBayes(fit_high)
# Results
low_res = topTable(fit_low,coef = 2,number = nrow(y_low))
kable(head(low_res,10),caption = "Low Storage Time")
high_res = topTable(fit_high,coef = 2,number = nrow(y_high))
kable(head(high_res,10),caption = "High Storage Time")
```

There were `r sum(low_res$P.Value < 0.05)` significant lipids in the low group and `r sum(high_res$P.Value < 0.05)` in the high group.

### Stratified by Sex

```{r results='asis'}
# Make design matrices
hard_male <- model.matrix(~HARD,male)
hard_female <- model.matrix(~HARD,female)
# Unadjusted models
y_male <- t(intensities[intensities$sample_id %in% male$sample_id,2:ncol(intensities)])
y_female <- t(intensities[intensities$sample_id %in% female$sample_id,2:ncol(intensities)])
fit_male <- lmFit(y_male,hard_male)
fit_male <- eBayes(fit_male)
fit_female <- lmFit(y_female,hard_female)
fit_female <- eBayes(fit_female)
# Results
male_res = topTable(fit_male,coef = 2,number = nrow(y_low))
kable(head(male_res,10),caption = "Males")
female_res = topTable(fit_female,coef = 2,number = nrow(y_high))
kable(head(female_res,10),caption = "Females")
```

There were `r sum(male_res$P.Value < 0.05)` significant lipids in the male group and `r sum(female_res$P.Value < 0.05)` in the female group.
