---
title: "EDIC Lipidomics"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_knit$set(root.dir = "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
knitr::opts_knit$set(root.dir = "~/Documents/Work/Janet Snell-Bergeon/EDIC/Lipidomics")
library(PCAtools)
library(readxl)
library(tidyverse)
```

```{r include=FALSE}
# Import
intensities = read.csv("./Data_Raw/EDIC lipidomics data matrix_Skyline_2021.02.25.csv",stringsAsFactors = F)
# Get compound info in separate frame
compound_info = intensities[,"LipidSpecies"]
intensities$LipidSpecies = NULL
# Transpose and format
intensities = as.data.frame(t(intensities))
names(compound_info) = colnames(intensities)
intensities$sample_id = gsub("\\.","-",rownames(intensities))
intensities = intensities %>% select(sample_id,everything())
intensities[,2:ncol(intensities)] = lapply(intensities[,2:ncol(intensities)],as.numeric)
# Import sample injection order
injection_order = read_excel("./Data_Raw/EDIC lipidomics run order.xlsx")
colnames(injection_order) = "sample_id"
injection_order$sample_id = sub("QC-EDIC","EDIC-QC",injection_order$sample_id)
injection_order$sample_id[is.na(injection_order$sample_id)] = paste0("Failed-QC-",1:sum(is.na(injection_order$sample_id))) # For now assume that the missing are QCs
injection_order$injection_order = 1:nrow(injection_order)
# Batch information
batches = diff(c(0,grep("QC",injection_order$sample_id),nrow(injection_order)))
injection_order$batch = rep(1:length(batches),times = batches)
# Remove duplicates and QC for now
injection_order = injection_order[-c(which(duplicated(injection_order$sample_id)),grep("Failed-QC",injection_order$sample_id)),]
# Get sample info
sample.info = read.csv("./Data_Raw/QEDIC_WP_DDA_60SPD_StudyInformat.csv")
# Get sample IDs, add leading zeroes and "EDIC" to number
sample.info$sample_id = sample.info$ProteomicsSampleID
sample.info$sample_id = paste0("EDIC-",str_pad(sample.info$sample_id,3,"left",0))
sample.info = sample.info %>% select(sample_id,everything())
# Combine
sample.info = merge(injection_order,sample.info,all = T,by = "sample_id")
sample.info = sample.info[order(sample.info$injection_order),]
# Try two batches
sample.info$batch2 = ifelse(sample.info$batch < 6,1,2)
# Plot by HARD status
sample.info$class = sample.info$HARD
sample.info$class[which(is.na(sample.info$class))] = "QC"
sample.info$class = factor(sample.info$class)
```

### Sample QC

#### PCA Scree plot

```{r scree}
ordered.intensities <- merge(sample.info[,c("sample_id", "injection_order")], intensities)
#Order by injection order - then the order of the samples will be the same as in sample.info
ordered.intensities <- ordered.intensities[order(ordered.intensities$injection_order),]
sample.ids <- ordered.intensities$sample_id
pca.mat <- as.matrix(t(ordered.intensities[,-c(1,2)]))
#Remove compounds with no variability so PCA does not fall over
no.var.compounds <- which(apply(pca.mat, 1, var) == 0)
if (length(no.var.compounds) > 0) {
  pca.mat <- pca.mat[-no.var.compounds,]  
}
# Complete cases
pca.mat = pca.mat[complete.cases(pca.mat),]
#The colnames of the PCA input matrix and the rownames of the PCA metadata must be set 
#and must contain the same IDs and in the same order
colnames(pca.mat) <- sample.ids
rownames(sample.info) <- sample.info$sample_id
pca = pca(pca.mat, metadata = sample.info, center=TRUE, scale=TRUE)
screeplot(pca, components=getComponents(pca, 1:10), hline = 50)
```

#### PCA scatterplots

The dashed horizontal and vertical lines denote 2 SD away from the mean, to be used for identifying outliers

```{r pca}
nr.sd <- 2
pca.frame <- pca$rotated
pca.frame$class <- sample.info$class
pca.frame$batch <- factor(sample.info$batch)
pca.frame$batch2 <- factor(sample.info$batch2)
pca.frame$CARV <- factor(sample.info$CARV)
pca.frame$HARD <- factor(sample.info$HARD)
pca.frame$SEX <- factor(sample.info$SEX)
pca.frame$Study = factor(sub("_.*","",sample.info$Visit_Code))
# Look at difference of sample storage visitduration - visit 00
pca.frame$SAMPLE_AGE = sample.info$VisitAge - sample.info$AGE00
pca.frame$SAMPLE_AGE = cut(pca.frame$SAMPLE_AGE,c(-Inf,median(pca.frame$SAMPLE_AGE),Inf))
#PC1 and PC2
x.min <- mean(pca.frame$PC1) - nr.sd*sd(pca.frame$PC1)
x.max <- mean(pca.frame$PC1) + nr.sd*sd(pca.frame$PC1)
y.min <- mean(pca.frame$PC2) - nr.sd*sd(pca.frame$PC2)
y.max <- mean(pca.frame$PC2) + nr.sd*sd(pca.frame$PC2)
# Color by class
ggplot(pca.frame, aes(x=PC1, y=PC2, color=CARV)) +
  geom_point() +
  scale_shape_manual(values=1:8) + theme_bw() +
  geom_hline(yintercept=y.min, linetype="dashed") +
  geom_hline(yintercept=y.max, linetype="dashed") +
  geom_vline(xintercept=x.min, linetype="dashed") +
  geom_vline(xintercept=x.max, linetype="dashed") 
# Color by class
ggplot(pca.frame, aes(x=PC1, y=PC2, color=class)) +
  geom_point() +
  scale_shape_manual(values=1:8) + theme_bw() +
  geom_hline(yintercept=y.min, linetype="dashed") +
  geom_hline(yintercept=y.max, linetype="dashed") +
  geom_vline(xintercept=x.min, linetype="dashed") +
  geom_vline(xintercept=x.max, linetype="dashed") 
# Color by sex
ggplot(pca.frame, aes(x=PC1, y=PC2, color=SEX)) +
  geom_point() +
  scale_shape_manual(values=1:8) + theme_bw() +
  geom_hline(yintercept=y.min, linetype="dashed") +
  geom_hline(yintercept=y.max, linetype="dashed") +
  geom_vline(xintercept=x.min, linetype="dashed") +
  geom_vline(xintercept=x.max, linetype="dashed") 
# Color by age
ggplot(pca.frame, aes(x=PC1, y=PC2, color=SAMPLE_AGE)) +
  geom_point() +
  scale_shape_manual(values=1:8) + theme_bw() +
  geom_hline(yintercept=y.min, linetype="dashed") +
  geom_hline(yintercept=y.max, linetype="dashed") +
  geom_vline(xintercept=x.min, linetype="dashed") +
  geom_vline(xintercept=x.max, linetype="dashed")
# Color by study
ggplot(pca.frame, aes(x=PC1, y=PC2, color=Study)) +
  geom_point() +
  scale_shape_manual(values=1:8) + theme_bw() +
  geom_hline(yintercept=y.min, linetype="dashed") +
  geom_hline(yintercept=y.max, linetype="dashed") +
  geom_vline(xintercept=x.min, linetype="dashed") +
  geom_vline(xintercept=x.max, linetype="dashed")
# Color by batch
ggplot(pca.frame, aes(x=PC1, y=PC2, color=batch)) +
  geom_point() +
  scale_shape_manual(values=1:8) + theme_bw() +
  geom_hline(yintercept=y.min, linetype="dashed") +
  geom_hline(yintercept=y.max, linetype="dashed") +
  geom_vline(xintercept=x.min, linetype="dashed") +
  geom_vline(xintercept=x.max, linetype="dashed") 
# Color by fewer batches
ggplot(pca.frame, aes(x=PC1, y=PC2, color=batch2)) +
  geom_point() +
  scale_shape_manual(values=1:8) + theme_bw() +
  geom_hline(yintercept=y.min, linetype="dashed") +
  geom_hline(yintercept=y.max, linetype="dashed") +
  geom_vline(xintercept=x.min, linetype="dashed") +
  geom_vline(xintercept=x.max, linetype="dashed") 
```

#### Remove PCA outliers

Remove samples that are PCA outliers.

```{r pca_outliers,echo=TRUE}
#PC1 
PC1.min <- mean(pca.frame$PC1) - nr.sd*sd(pca.frame$PC1)
PC1.max <- mean(pca.frame$PC1) + nr.sd*sd(pca.frame$PC1)
PC1.outliers <- sample.info$sample_id[(pca.frame$PC1 < PC1.min) | (pca.frame$PC1 > PC1.max)]

#PC2
PC2.min <- mean(pca.frame$PC2) - nr.sd*sd(pca.frame$PC2)
PC2.max <- mean(pca.frame$PC2) + nr.sd*sd(pca.frame$PC2)
PC2.outliers <- sample.info$sample_id[(pca.frame$PC2 < PC2.min) | (pca.frame$PC2 > PC2.max)]

#Create output data frame with outliers
outlier.frame <- data.frame(
  sample_id=c(PC1.outliers, PC2.outliers),
  reason=c(rep("PC1 outlier", length(PC1.outliers)), rep("PC2 outlier", length(PC2.outliers)))
)
dim(outlier.frame)
outlier.ids <- unique(outlier.frame$sample_id)
length(outlier.ids)
outlier.ids
#Remove the outliers from the intensities data frame, so they are no longer present for the 
#subsequent QC of the comounds
dim(intensities)
intensities <- intensities[!(intensities$sample_id %in% outlier.ids),]
dim(intensities)
dim(sample.info)
sample.info <- sample.info[!(sample.info$sample_id %in% outlier.ids),]
dim(sample.info)
```


#### Number of peaks per sample ordered by injection order

```{r nr_peaks}
nrPeaks <- function(i) {
  return (sum(i != 1 & !is.na(i)))
}
ordered.intensities <- merge(sample.info[,c("sample_id", "injection_order")], intensities)
#Order by injection order - then the order of the samples will be the same as in sample.info
ordered.intensities <- ordered.intensities[order(ordered.intensities$injection_order),-c(1,2)]
nr.peaks <- apply(ordered.intensities, 1, nrPeaks)
nr.peaks.frame <- sample.info[order(sample.info$injection_order),]
nr.peaks.frame$nr_peaks <- nr.peaks
# Plot
ggplot(nr.peaks.frame, aes(x=injection_order, y=nr_peaks, shape=factor(batch), color=class)) +
  geom_point() +
  scale_shape_manual(values=1:15) 
```

#### Mean intensity per sample ordered by injection order

```{r mean_intensity}
meanIntensity <- function(i) {
  return (mean(i,na.rm = T))
}
mean.intensity <- apply(ordered.intensities, 1, meanIntensity)
avg.intensity.frame <- sample.info[order(sample.info$injection_order),]
avg.intensity.frame$mean_intensity <- mean.intensity
ggplot(avg.intensity.frame , aes(x=injection_order, y=mean_intensity, shape=factor(batch), color=class)) +
  geom_point() +
  scale_shape_manual(values=1:15)
```

### Compound QC

#### Intensity boxplots

```{r compound intensity}
wide.frame <- merge(intensities, sample.info)
keycol <- "compound_id"
valuecol <- "intensity"
gathercols <- names(wide.frame)[grep("^V", names(wide.frame))] # Assume column names start with "V"
long.frame <- gather_(wide.frame, keycol, valuecol, gathercols)
long.frame <- long.frame[order(long.frame$injection_order),]
long.frame$log2_intensity <- log2(as.numeric(long.frame$intensity))
long.frame <- long.frame[long.frame$log2_intensity > 0,]
boxplot(log2_intensity ~ batch, data=long.frame)
```

#### Histogram of percent missing values per compound

##### To change

```{r distro_missing_compounds}
compound.frame <- merge(sample.info, intensities)
compound.cols <- names(compound.frame)[startsWith(names(compound.frame), "V")]
percMissCompounds <- function(i) {
  return ((sum(i==1)/length(i))*100)
}
compound.miss.frame <- compound.frame %>% 
  group_by(class) %>%
  summarize_at(vars(compound.cols), list(percMissCompounds))
head(compound.miss.frame[,1:10])
keycol <- "compound_id"
valuecol <- "percent_missing"
compound.miss.long.frame <- gather_(compound.miss.frame, keycol, valuecol, compound.cols)
ggplot(compound.miss.long.frame, aes(x=percent_missing, fill=class)) +
    geom_histogram(position = 'dodge', breaks=seq(0, 100, 10)) +
    scale_x_continuous(breaks = seq(0, 100, 10))
```

#### Identify compounds with large missingness in research samples

##### To change

* Set the qc.classes variable to the possible class values of QC samples
* Set the max.perc.research.missingness variable to the maximum percentage allowed missingness in research samples; defaults to 20\%

```{r remove_high_missing_compounds}
qc.classes <- c("QC")
max.perc.research.missingness <- 20
compound.research.frame <- merge(sample.info[!(sample.info$class %in% qc.classes),], intensities)
compound.cols <- names(compound.research.frame)[startsWith(names(compound.research.frame), "C")]
compound.research.miss.frame <- compound.research.frame %>%
  summarize_at(vars(compound.cols), list(percMissCompounds))
head(compound.research.miss.frame[,1:10])
head(t(compound.research.miss.frame[,1:10]))
compound.research.miss.long.frame <- t(compound.research.miss.frame)
sum(compound.research.miss.long.frame > max.perc.research.missingness,na.rm = T)
miss.remove.list <- rownames(compound.research.miss.long.frame)[compound.research.miss.long.frame > max.perc.research.missingness]
```

#### Identify compounds present only in research samples

These compounds must be removed as they cannot be normalized by methods that assume QC samples.
Note that if a compound has a large % missingness in QC samples (but is detected in a small number of QC samples]),
it is indicative of perhaps not really being detected in the QC samples, and hence these compounds should be removed too
(to much missingness in QC samples may also affect normalization).

##### To change

* Set the max.perc.qc.missingness variable to the maximum percentage allowed missingness in QC samples; defaults to 80\%

```{r remove_research_only_compounds}
max.perc.qc.missingness <- 80
compound.qc.miss.long.frame <- compound.miss.long.frame[compound.miss.long.frame$class %in% qc.classes,]
compound.qc.miss.long.frame <- compound.qc.miss.long.frame[!(compound.qc.miss.long.frame$compound_id %in% miss.remove.list),]
researchonly.remove.list <- unique(compound.qc.miss.long.frame$compound_id[compound.qc.miss.long.frame$percent_missing > max.perc.qc.missingness])
length(researchonly.remove.list)
```

#### Identify compounds with large CV in QC samples

##### To change

* Set the cv.threshold variable to the maximum CV allowed in QC samples; defaults to 0.3

```{r remove_CV_compounds}
cv.threshold <- 0.3
compound.qc.frame <- merge(sample.info[(sample.info$class %in% qc.classes),], intensities[,!(colnames(intensities) %in% c(miss.remove.list, researchonly.remove.list))])
compound.cols <- names(compound.qc.frame)[startsWith(names(compound.qc.frame), "C")]
cV <- function(x) {
  x <- x[x != 1]
  return(sd(x)/mean(x))
}
cv.frame <- compound.qc.frame %>%
  group_by(class) %>%
  summarize_at(vars(compound.cols), list(cV))
head(cv.frame[,1:10])
keycol <- "compound_id"
valuecol <- "cv"
cv.long.frame <- gather_(cv.frame, keycol, valuecol, compound.cols)
ggplot(cv.long.frame, aes(x=cv, fill=class)) +
    geom_histogram(position = 'dodge', breaks=seq(0, max(cv.long.frame$cv)+0.1, 0.1)) +
    scale_x_continuous(breaks = seq(0, max(cv.long.frame$cv+0.1), 0.1))
cv.remove.list <- unique(cv.long.frame$compound_id[cv.long.frame$cv > cv.threshold])
length(cv.remove.list)

```

#### Write sample and compound failures and create a clean output file

```{r write_output}
#Write files with information on sample and compound failures
write.table(outlier.frame,
            paste0(data.dir, "sample_QC_failures.txt"),
             sep="\t", quote=F, row.names=F, col.names=T)
compound.failure.frame <- data.frame(compound_id=c(miss.remove.list, researchonly.remove.list, cv.remove.list),
                       reason=c(rep("High % missing values", length(miss.remove.list)),
                                rep("Not in QC samples", length(researchonly.remove.list)),
                                rep("High CV in QC samples", length(cv.remove.list))))
write.table(compound.failure.frame,
             paste0(data.dir, "compound_QC_failures.txt"),
             sep="\t", quote=F, row.names=F, col.names=T)

#Write clean files with sample and compound QC failures removed
#(note that sample failures were removed from intensities prior to compound QC)
keep.cols <- names(intensities)[!(names(intensities) %in% compound.failure.frame$compound_id)]
dim(intensities)
intensities <- intensities[,keep.cols]
dim(intensities)
write.table(intensities,
             paste0(data.dir, "compound_intensities_clean.txt"),
             sep="\t", quote=F, row.names=F, col.names=T)

```


# Questions for Randi

1. Why does the number of peaks code count the number of compounds > 1 in each sample? Am I supposed to change that to a more meaningful threshold?


