---
title: "EDIC Glycated Proteomics"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_knit$set(root.dir = "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/EDIC/Lipidomics")
knitr::opts_knit$set(root.dir = "~/Documents/Work/Janet Snell-Bergeon/EDIC/Proteomics")
library(tidyverse)
library(limma)
library(knitr)
```

```{r include=FALSE}
# Import
intensities = read.csv("./Data_Clean/peptide_groups.csv",stringsAsFactors = F)
# Get compound info in separate frame
compound_info = intensities[,"Sequence"]
intensities = intensities[,grep("Abundance_Sample",colnames(intensities))]
# Transpose and format
intensities = as.data.frame(t(intensities))
intensities$sample_id = gsub("Abundance_","",rownames(intensities))
intensities = intensities %>% select(sample_id,everything())
intensities[,2:ncol(intensities)] = lapply(intensities[,2:ncol(intensities)],function(c){log(as.numeric(c))})
# Get sample info and IDs
sample.info = read.csv("./Data_Clean/study_info.csv")
id_list = read.csv("./Data_Clean/id_list.csv")
id_list = id_list[,c("Subject.ID","Raw.File.Name")]
sample.info = left_join(sample.info,id_list,by = "Subject.ID")
sample.info = sample.info[!(duplicated(sample.info$Raw.File.Name)),]
# Get sample IDs, add leading zeroes and "EDIC" to number
sample.info = sample.info %>% arrange(Raw.File.Name)
sample.info$sample_id = sample.info$Raw.File.Name
```

# Moderated T Tests

## CARV

### All

Moderated t tests were performed to test for differences in proteins between those who experienced CARV and those who did not. The results immediately below were not stratified by sex or storage time.

```{r results='asis'}
# limma stratified by storage time
sample.info$storage_time = sample.info$VisitAge - sample.info$AGE00
sample.info$storage_time = cut(sample.info$storage_time,2,labels = c("< 14",">= 14"))
low = sample.info[sample.info$storage_time=="< 14",]
high = sample.info[sample.info$storage_time==">= 14",]
# Make design matrix
carv <- model.matrix(~CARV,sample.info[!is.na(sample.info$storage_time),])
# Unadjusted models
y <- t(intensities[,2:ncol(intensities)])
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
kable(topTable(fit,coef = 2,sort.by = "p"))
```

#### Adjusted for sex, age, diabetes duration, treatment group

```{r results='asis'}
# Make design matrix
carv <- model.matrix(~CARV+SEX+AGE+DURATION+GROUP,sample.info)
# Unadjusted models
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y),sort.by = "p")
kable(head(res,10),caption = "Not adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05,na.rm=T)` nominally significant proteins.

```{r results='asis'}
# Make design matrix
carv <- model.matrix(~CARV+SEX+AGE+DURATION+GROUP+storage_time,sample.info)
# Unadjusted models
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y),sort.by = "p")
kable(head(res,10),caption = "Adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05,na.rm=T)` nominally significant proteins.

### Stratified by Above or Below Median Storage Time

```{r results='asis'}
# Make design matrices
carv_low <- model.matrix(~CARV,low)
carv_high <- model.matrix(~CARV,high)
# Unadjusted models
y_low <- t(intensities[intensities$sample_id %in% low$sample_id,2:ncol(intensities)])
y_high <- t(intensities[intensities$sample_id %in% high$sample_id,2:ncol(intensities)])
fit_low <- lmFit(y_low,carv_low)
fit_low <- eBayes(fit_low)
fit_high <- lmFit(y_high,carv_high)
fit_high <- eBayes(fit_high)
# Results
low_res = topTable(fit_low,coef = 2,number = nrow(y_low),sort.by = "p")
kable(head(low_res,10),caption = "Low Storage Time")
high_res = topTable(fit_high,coef = 2,number = nrow(y_high),sort.by = "p")
kable(head(high_res,10),caption = "High Storage Time")
```

There were `r sum(low_res$P.Value < 0.05,na.rm=T)` nominally significant proteins in the low group and `r sum(high_res$P.Value < 0.05,na.rm=T)` in the high group.

### Stratified by Sex

```{r results='asis'}
# Split sample info
male = sample.info[sample.info$SEX == "M",]
female = sample.info[sample.info$SEX == "F",]
# Make design matrices
carv_male <- model.matrix(~CARV,male)
carv_female <- model.matrix(~CARV,female)
# Unadjusted models
y_male <- t(intensities[intensities$sample_id %in% male$sample_id,2:ncol(intensities)])
y_female <- t(intensities[intensities$sample_id %in% female$sample_id,2:ncol(intensities)])
fit_male <- lmFit(y_male,carv_male)
fit_male <- eBayes(fit_male)
fit_female <- lmFit(y_female,carv_female)
fit_female <- eBayes(fit_female)
# Results
male_res = topTable(fit_male,coef = 2,number = nrow(y_low),sort.by = "p")
kable(head(male_res,10),caption = "Males")
female_res = topTable(fit_female,coef = 2,number = nrow(y_high),sort.by = "p")
kable(head(female_res,10),caption = "Females")
```

There were `r sum(male_res$P.Value < 0.05,na.rm=T)` nominally significant proteins in the male group and `r sum(female_res$P.Value < 0.05,na.rm=T)` in the female group.

## HARD

### All

Moderated t tests were performed to test for differences in proteins between those who experienced HARD and those who did not. The results immediately below were not stratified by sex or storage time.

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~HARD,sample.info[!is.na(sample.info$storage_time),])
# Unadjusted models
y <- t(intensities[intensities$sample_id %in% sample.info[!is.na(sample.info$storage_time),"sample_id"],2:ncol(intensities)])
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
kable(topTable(fit_low,coef = 2,sort.by = "p"))
```

### Adjusted for sex, age, diabetes duration, treatment group

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~HARD+SEX+AGE+DURATION+GROUP,sample.info)
# Unadjusted models
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y),sort.by = "p")
kable(head(res,10),caption = "Not adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05,na.rm=T)` nominally significant proteins.

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~HARD+SEX+AGE+DURATION+GROUP+storage_time,sample.info)
# Unadjusted models
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y),sort.by = "p")
kable(head(res,10),caption = "Adjusted for storage time")
```

There were `r sum(res$P.Value < 0.05,na.rm=T)` nominally significant proteins.

### Stratified by Above or Below Median Storage Time

```{r results='asis'}
# Make design matrices
hard_low <- model.matrix(~HARD,low)
hard_high <- model.matrix(~HARD,high)
# Unadjusted models
y_low <- t(intensities[intensities$sample_id %in% low$sample_id,2:ncol(intensities)])
y_high <- t(intensities[intensities$sample_id %in% high$sample_id,2:ncol(intensities)])
fit_low <- lmFit(y_low,hard_low)
fit_low <- eBayes(fit_low)
fit_high <- lmFit(y_high,hard_high)
fit_high <- eBayes(fit_high)
# Results
low_res = topTable(fit_low,coef = 2,number = nrow(y_low),sort.by = "p")
kable(head(low_res,10),caption = "Low Storage Time")
high_res = topTable(fit_high,coef = 2,number = nrow(y_high),sort.by = "p")
kable(head(high_res,10),caption = "High Storage Time")
```

There were `r sum(low_res$P.Value < 0.05,na.rm=T)` nominally significant proteins in the low group and `r sum(high_res$P.Value < 0.05,na.rm=T)` in the high group.

### Stratified by Sex

```{r results='asis'}
# Make design matrices
hard_male <- model.matrix(~HARD,male)
hard_female <- model.matrix(~HARD,female)
# Unadjusted models
y_male <- t(intensities[intensities$sample_id %in% male$sample_id,2:ncol(intensities)])
y_female <- t(intensities[intensities$sample_id %in% female$sample_id,2:ncol(intensities)])
fit_male <- lmFit(y_male,hard_male)
fit_male <- eBayes(fit_male)
fit_female <- lmFit(y_female,hard_female)
fit_female <- eBayes(fit_female)
# Results
male_res = topTable(fit_male,coef = 2,number = nrow(y_low),sort.by = "p")
kable(head(male_res,10),caption = "Males")
female_res = topTable(fit_female,coef = 2,number = nrow(y_high),sort.by = "p")
kable(head(female_res,10),caption = "Females")
```

There were `r sum(male_res$P.Value < 0.05,na.rm=T)` nominally significant proteins in the male group and `r sum(female_res$P.Value < 0.05,na.rm=T)` in the female group.

# Storage Time in Controls

Moderated t tests were performed to test for differences in proteins between controls (GROUP variable is "STANDARD") whose samples were stored above the median storage time and those whose samples were below the median.

```{r results='asis'}
# controls stratified by storage time
controls = sample.info[sample.info$GROUP == "STANDARD",]
# Make design matrix
stor <- model.matrix(~storage_time,controls)
# Unadjusted models
y <- t(intensities[intensities$sample_id %in% controls$sample_id,2:ncol(intensities)])
fit <- lmFit(y,stor)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,sort.by = "p",number = nrow(y))
kable(head(res,10))
```

There were `r sum(res$adj.P.Val < 0.05,na.rm=T)` proteins that were significantly different by storage time after FDR p value adjustment.
