---
title: "DCCT Untargeted Metabolomics"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arsenal)
library(skimr)
library(knitr)
library(mixOmics)
library(tidyverse)
library(limma)
library(webchem)
library(omu)
library(PCAtools)
library(dplyr)
library(stringr)
home_dir = ifelse(.Platform$OS.type != "unix","T:/Janet Snell-Bergeon/AHA collaborative grant/Metabolomics",
                  "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Metabolomics")
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning,echo=FALSE}
#setwd(home_dir)
# Read in raw data
raw_data <- 
  read.csv("/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Metabolomics/EDIC metabolomics final data matrix.csv", stringsAsFactors = F,na.strings = c("","NA"))
compound.info <- raw_data %>% select(NovaMT.Library.No.,External.Identifier,Compound, m.z_light, m.z_heavy)
# Format
data <- raw_data %>% 
  mutate(compound = coalesce(NovaMT.Library.No.,External.Identifier),
         m.z = Neutral.Mass..Da.+1.0078,
         rt = RT..s.,
         masserror = Mass.Error..ppm.,
         rterror = RT.Error..s.) %>%  select(compound,m.z,rt,masserror,rterror,CVD:Healthy.124)
data$compound <- paste(data$compound,data$m.z,data$rt,sep = "_")
# Rename columns
#colnames(data) <- gsub("IsoMS_results_","",colnames(data))
#colnames(data) <- gsub("*_.*","",colnames(data))
# Names
names <- data$compound
# Data matrix
data <- data %>% select(CVD:Healthy.124) %>% t(.) %>% as.data.frame(.) # Transpose - samples in rows
# Format columns
colnames(data) <- names
data$GlobalSampleID <- as.character(rownames(data))

# get actual study IDs
id_labels <- read.csv("/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Metabolomics/metabolomics_without_imputation.csv", nrows = 1)
id_labels <- id_labels[,17:ncol(id_labels)]
id_labels <- t(id_labels)
a <- row.names(id_labels)
id_labels <- cbind(id_labels, a)
colnames(id_labels) <- c("group_id_labels","CURRENT_LABEL")
id_labels <- as.data.frame(id_labels)
id_labels$CURRENT_LABEL <- word(id_labels$CURRENT_LABEL,2,sep = "\\.")
id_labels <- id_labels %>% filter(group_id_labels %in% c("CVD","HD","Healthy"))
data <- cbind(data, id_labels)
data_group <- row.names(data)
data_group <- word(data_group,1,sep = "\\.")
data <- cbind(data,data_group)
# read in linking file
linking <- read.csv("/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Data_Raw/snell_bergeon_21550_Linking_File.csv")
linking$MASK_PAT <- linking$mask_pat
linking <- linking %>% select(MASK_PAT, CURRENT_LABEL)
linking$CURRENT_LABEL <- str_replace(linking$CURRENT_LABEL ," ", "_")
data <- merge(data, linking, by="CURRENT_LABEL", all.x = T, all.y = F)
# read in study data
study_dat <- read.csv("/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Proteomics/Raw data/QEDIC_WP_DDA_60SPD_StudyInformat.csv")
# need to link study data to results file
data <- merge(data, study_dat, by="MASK_PAT", all.x = T, all.y = F)
# reorder columns
data <- cbind(data[,1:2],data[,412:ncol(data)],data[,3:411])
```

# Pre-processing

```{r echo=FALSE}
# Check for samples missing > 80% of compounds, remove them
#missing <- (rowSums(is.na(data[,131:ncol(data)])) / length(131:ncol(data))) > 0.8
#data <- data[-missing,]
# Same for 0 instead of NA - none
#missing0 <- which(rowSums(data[,10:ncol(data)] == 0) / length(10:ncol(data)) > 0.8)
# Check for and remove compounds with > 20% missing in research - none
#research <- which(colSums(is.na(data[!grepl("QC",data$GlobalSampleID),])) / 
#                    (nrow(data))-13 > 0.2)
# or > 80% missing in QC - none
#qc <- names(which(colSums(is.na(data[grepl("QC",data$GlobalSampleID),10:ncol(data)])) / 
#                    13 > 0.2))
#data <- data[,!(names(data) %in% qc)]
# Check for compounds with CV > 0.3 in QC samples
#qc_cv <- names(which(lapply(data[grepl("QC",data$GlobalSampleID),131:ncol(data)],function(x){
#  sd(x,na.rm = T)/mean(x,na.rm = T)
#}) > 0.3))
#data <- data[,!(names(data) %in% qc_cv)]
# Remove bad name compound
#data[,grep("314_",colnames(data))] <- NULL
```


# Moderated T Tests

## CARV

Moderated t tests were performed to test for differences in proteins between those who experienced CARV and those who did not. 

```{r results='asis'}
# Make design matrix
carv <- model.matrix(~data$CARV)
# Unadjusted models
y <- t(data[,131:ncol(data)])
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number=nrow(y),sort.by = "p")
#res$NovaMT.Library.No. <- word(row.names(res),1,sep = "\\_")
#res <- merge(res, compound.info, by=c("NovaMT.Library.No."), all.x = T, all.y = T)
kable(res,row.names = F)
write.csv(res,"/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Metabolomics/Reports/metabolomics_by_CARV.csv")
```

#### Adjusted for sex, age, diabetes duration, treatment group

```{r results='asis'}
# Make design matrix
carv <- model.matrix(~CARV+SEX+AGE+DURATION+GROUP,data)
# Unadjusted models
fit <- lmFit(y,carv)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y),sort.by = "p")
kable(head(res,10),caption = "Not adjusted for storage time",row.names = F)
write.csv(res,"/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Metabolomics/Reports/metabolomics_by_CARV_adj_sex_age_duration_group.csv")
```

## HARD

Moderated t tests were performed to test for differences in proteins between those who experienced HARD and those who did not. 

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~data$HARD)
# Unadjusted models
y <- t(data[,131:ncol(data)])
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number=nrow(y),sort.by = "p")
#res$NovaMT.Library.No. <- word(row.names(res),1,sep = "\\_")
#res <- merge(res, compound.info, by=c("NovaMT.Library.No."), all.x = T, all.y = T)
kable(res,row.names = F)
write.csv(res,"/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Metabolomics/Reports/metabolomics_by_HARD.csv")
```

#### Adjusted for sex, age, diabetes duration, treatment group

```{r results='asis'}
# Make design matrix
hard <- model.matrix(~HARD+SEX+AGE+DURATION+GROUP,data)
# Unadjusted models
fit <- lmFit(y,hard)
fit <- eBayes(fit)
# Results
res = topTable(fit,coef = 2,number = nrow(y),sort.by = "p")
kable(head(res,10),caption = "Not adjusted for storage time",row.names = F)
write.csv(res,"/Volumes/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/CACTI DCCT EDIC omics R01/Metabolomics/Reports/metabolomics_by_HARD_adj_sex_age_duration_group.csv")
```