---
title: "670G Obs. ATTD 2019"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(tableone)
library(knitr)
library(nlme)
library(pROC)
library(tidyverse)
```

```{r define functions,include=FALSE}
fill_NAs <- function(data,threshold,idcols) {
  na_thresh <- ceiling(threshold * ncol(data))-length(idcols) # Determine max NAs
  data <- data[rowSums(is.na(data)) < na_thresh,] # Remove rows with NAs over max
  k <- which(is.na(data), arr.ind=TRUE) # Find remaining NAs
  if(length(k)>0) {
    data[k] <- rowMeans(data[,colnames(data)[-idcols]],na.rm = T)[k[,1]] # Replace with means
  }
  data
}
```

```{r data import and formatting,echo=FALSE,include=FALSE}
# Import
source("./Cari Berget/ATTD Abstract 2019/Data_Cleaned/AdvancedClosedLoopCl_R_2019-09-30_1049.r")
# A1c as numeric
data$hba1c <- as.numeric(as.character(data$hba1c))
data$hba1c_baseline <- as.numeric(as.character(data$hba1c_baseline))
data$hba1c <- ifelse(data$gyl_timepoint.factor == "Baseline",data$hba1c_baseline,data$hba1c)
# Missing glycemic timepoints to baseline
data$gyl_timepoint.factor[is.na(data$gyl_timepoint.factor)] <- "Baseline"
# Subjects who withdrew (per Cari)
withdrawn <- c("41","50")
# Demographics
demographics <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(automode_start),!(record_id %in% withdrawn)) %>%
  select(record_id,demographics_dob,demographics_diabetesdx,automode_start,
         hba1c_date_b,hba1c_baseline,demographics_sex.factor,
         demographics_race.factor,demographics_ethnicity.factor,
         demographics_insurance.factor,demographics_cgmhx.factor,demographics_pumphx.factor) %>%
  mutate(age_at_am_start = round(as.numeric((difftime(automode_start,demographics_dob,units = "days"))/365.25),2),
         t1d_duration_at_am_start = round(as.numeric((difftime(automode_start,demographics_diabetesdx,units = "days"))/365.25),2))
# Days from AM start
days <- data %>% group_by(record_id) %>%
  mutate(date = coalesce(as.character(hba1c_date_b),as.character(date_visit))) %>%
  fill(automode_start) %>%
  mutate(days_to_am = abs(as.numeric((difftime(automode_start,date,
                                               units = "days"))))) %>%
  select(record_id,gyl_timepoint.factor,days_to_am) %>% 
  filter(!is.na(gyl_timepoint.factor) & !is.na(days_to_am))
days$gyl_timepoint.factor <- as.character(days$gyl_timepoint.factor)
# Long to wide for demographics frame
days <- days %>% spread(gyl_timepoint.factor,days_to_am)
# Column names
colnames(days) <- c("record_id","baseline_to_am","am_to_m1","am_to_t1","am_to_t2",
                    "am_to_t3","am_to_t4")
# add to demographics
demographics <- left_join(demographics,days,by = "record_id")
# 12 month data
year <- data %>% filter(record_id %in% demographics$record_id) %>% 
  group_by(record_id) %>%
  summarise(av_am_use = mean(time_am,na.rm = T))
year <- left_join(data[grep("5",data$redcap_event_name),c("record_id","hba1c")],
                  year, by = "record_id")
# Split HbA1c into above vs. below or = 7.5
year$hba1c <- cut(year$hba1c,breaks = c(-Inf,7.5,Inf),labels = c("<=7.5%",">7.5%"))
```

# Table 1: Demographics

```{r table 1,echo=FALSE}
cont_vars <- c("hba1c_baseline","age_at_am_start","t1d_duration_at_am_start",
               "baseline_to_am","am_to_m1","am_to_t1","am_to_t2","am_to_t3",
               "am_to_t4")
cat_vars <- c("demographics_sex.factor","demographics_race.factor",
              "demographics_ethnicity.factor","demographics_insurance.factor",
              "demographics_cgmhx.factor","demographics_pumphx.factor")
nonnormal <- c("t1d_duration_at_am_start","baseline_to_am","am_to_m1","am_to_t1",
               "am_to_t2","am_to_t3","am_to_t4") # A1c always reported as mean (SD)
t1 <- CreateTableOne(vars = c(cont_vars,cat_vars),data = demographics)
t1 <- print(t1,nonnormal = nonnormal,printToggle = F)
kable(t1)
```

```{r discontinuers,echo=FALSE}
# Find discontinuers
t1_discont <- data %>%
  filter(gyl_timepoint.factor == "Time 1", time_am < 10) %>%
  .$record_id
t2_discont <- data %>%
  filter(gyl_timepoint.factor == "Time 2", time_am < 10) %>%
  .$record_id
t3_discont <- data %>%
  filter(gyl_timepoint.factor == "Time 3", time_am < 10) %>%
  .$record_id
t4_discont <- data %>%
  filter(gyl_timepoint.factor == "Time 4", time_am < 10) %>%
  .$record_id
all_discont <- unique(c(t1_discont,t2_discont,t3_discont,t4_discont))
first_six <- unique(c(t1_discont,t2_discont))
# Intermittent discontinuers
intermittent <- data %>% filter(!is.na(time_am),gyl_timepoint.factor != "Baseline") %>% 
  select(record_id,redcap_event_name,time_am)
intermittent$time_am <- ifelse(intermittent$time_am < 10,0,1)
intermittent <- intermittent %>% group_by(record_id) %>%
  summarise(rle = length(rle(time_am)$lengths))
intermittent <- intermittent$record_id[which(intermittent$rle > 2)]
```

```{r descriptive models,echo=FALSE}
# Model data with AM < 10 excluded
mod_data <- data %>% filter(gyl_timepoint.factor == "Baseline" | time_am >=10)
# Models and model means
# AM use
am_use_mod <- lme(time_am ~ gyl_timepoint.factor,random=~1|record_id,
                  data = mod_data[mod_data$gyl_timepoint.factor != "Baseline",],
                  na.action = na.omit)
am_use_mod_means <- lme(time_am ~ gyl_timepoint.factor-1,random=~1|record_id,
                        mod_data[mod_data$gyl_timepoint.factor != "Baseline",],
                        na.action = na.omit)
# TIR
tir_70_180_mod <- lme(sensor_70_180 ~ gyl_timepoint.factor,random=~1|record_id,
                      data = mod_data[mod_data$gyl_timepoint.factor != "Baseline",],na.action = na.omit)
tir_70_180_mod_means <- lme(sensor_70_180 ~ gyl_timepoint.factor-1,random=~1|record_id,
                            data = mod_data[mod_data$gyl_timepoint.factor != "Baseline",],na.action = na.omit)
# HbA1c
hba1c_mod <- lme(hba1c ~ gyl_timepoint.factor,random=~1|record_id,
                 data = mod_data,na.action = na.omit)
hba1c_mod_means <- lme(hba1c ~ gyl_timepoint.factor-1,random=~1|record_id,
                       data = mod_data,na.action = na.omit)
```

# Outcomes over time

## AM Use (%)

```{r am use diff, echo=FALSE}
diff <- summary(am_use_mod)$tTable
rownames(diff) <- sub("gyl_timepoint.factor","",rownames(diff))
kable(diff,caption = "Difference From Baseline")
```

This table shows the difference between each timepoint and month 1 (the intercept). So for example, on average AM use was 16.1 points lower at month 6 compared to baseline (p = 0.0003).

```{r am use means, echo=FALSE}
means <- summary(am_use_mod_means)$tTable
rownames(means) <- sub("gyl_timepoint.factor","",rownames(means))
kable(means,caption = "Model Means")
```

This table shows the average AM use at each time point. The p-values here indicate whether or not the average is close to 0, so I would ignore them when reporting these results. 

The interpretation of the following tables is the same, but the differences are from baseline visit (all time points were compared to month 1 for AM use, as it is the first visit with AM use data). All models only included baseline observations and visits with AM use >= 10%.

## TIR (%)

```{r tir diff, echo=FALSE}
diff <- summary(tir_70_180_mod)$tTable
rownames(diff) <- sub("gyl_timepoint.factor","",rownames(diff))
kable(diff,caption = "Difference From Baseline")
```

```{r tir means, echo=FALSE}
means <- summary(tir_70_180_mod_means)$tTable
rownames(means) <- sub("gyl_timepoint.factor","",rownames(means))
kable(means,caption = "Model Means")
```

## HbA1c (%)

```{r hba1c diff, echo=FALSE}
diff <- summary(hba1c_mod)$tTable
rownames(diff) <- sub("gyl_timepoint.factor","",rownames(diff))
kable(diff,caption = "Difference From Baseline")
```

```{r hba1c means, echo=FALSE}
means <- summary(hba1c_mod_means)$tTable
rownames(means) <- sub("gyl_timepoint.factor","",rownames(means))
kable(means,caption = "Model Means")
```

# Discontinuers

```{r echo=FALSE}
discont_table <- as.data.frame(matrix(nrow = 1,ncol = 6))
colnames(discont_table) <- c("T1","T2","T3","T4","Total","Intermittent")
rownames(discont_table) <- "n"
discont_table$T1 <- length(t1_discont)
discont_table$T2 <- length(t2_discont)
discont_table$T3 <- length(t3_discont)
discont_table$T4 <- length(t4_discont)
discont_table$Total <- length(all_discont)
discont_table$Intermittent <- length(intermittent)
kable(t(discont_table),caption = "Discontinuers")
```

The table above includes the number of participants with AM use < 10% at each time point, the total number of participants who discontinued at some point in the study ("Total"), and the number of participants who discontinued at one visit, then were counted as a continuer at a later visit, then discontinued again ("Intermittent"). 

# ROC Analysis

```{r ROC analysis,echo=FALSE,message=FALSE,dpi=600}
roc <- roc(year,hba1c,av_am_use)
plot.roc(roc)
results <- as.data.frame(coords(roc,x="best",input = "threshold",best.method = "youden",
                  transpose = FALSE))
results$AUC <- auc(roc)
kable(results,caption = "ROC Analysis Results")
```

The optimal average AM use (across all visits) threshold based on the Youden index is 68.9%. This has a specificity of 62.5% and a sensitivity of 84.1% to detect whether or not HbA1c was <=7.5 % or >7.5 % after 12 months. AUC is 0.76.