---
title: "670G Obs. ATTD 2019"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(tableone)
library(knitr)
library(nlme)
library(pROC)
library(tidyverse)
```

```{r define functions,include=FALSE}
fill_NAs <- function(data,threshold,idcols) {
  na_thresh <- ceiling(threshold * ncol(data))-length(idcols) # Determine max NAs
  data <- data[rowSums(is.na(data)) < na_thresh,] # Remove rows with NAs over max
  k <- which(is.na(data), arr.ind=TRUE) # Find remaining NAs
  if(length(k)>0) {
    data[k] <- rowMeans(data[,colnames(data)[-idcols]],na.rm = T)[k[,1]] # Replace with means
  }
  data
}
```

```{r data import and formatting,echo=FALSE,include=FALSE}
# Import
source("./Cari Berget/ATTD Abstract 2019/Data_Cleaned/AdvancedClosedLoopCl_R_2019-09-30_1049.r")
# A1c as numeric
data$hba1c <- as.numeric(as.character(data$hba1c))
data$hba1c_baseline <- as.numeric(as.character(data$hba1c_baseline))
# Subjects who withdrew (per Cari)
withdrawn <- c("41","50")
# Demographics
demographics <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(automode_start),!(record_id %in% withdrawn)) %>%
  select(record_id,demographics_dob,demographics_diabetesdx,automode_start,hba1c_date_b,
         hba1c_baseline,demographics_sex.factor,demographics_race.factor,demographics_ethnicity.factor,
         demographics_insurance.factor,demographics_cgmhx.factor,demographics_pumphx.factor) %>%
  mutate(age_at_am_start = round(as.numeric((difftime(automode_start,demographics_dob,units = "days"))/365.25),2),
         t1d_duration_at_am_start = round(as.numeric((difftime(automode_start,demographics_diabetesdx,units = "days"))/365.25),2))
# 12 month data
year <- data %>% filter(record_id %in% demographics$record_id) %>% 
  group_by(record_id) %>%
  summarise(av_am_use = mean(time_am,na.rm = T))
year <- left_join(data[grep("5",data$redcap_event_name),c("record_id","hba1c")],
                  year, by = "record_id")
# Split HbA1c into above vs. below or = 7.5
year$hba1c <- cut(year$hba1c,breaks = c(-Inf,7.5,Inf),labels = c("<=7.5%",">7.5%"))
```

```{r table 1,echo=FALSE}
cont_vars <- c("hba1c_baseline","age_at_am_start","t1d_duration_at_am_start")
cat_vars <- c("demographics_sex.factor","demographics_race.factor","demographics_ethnicity.factor",
              "demographics_insurance.factor","demographics_cgmhx.factor","demographics_pumphx.factor")
nonnormal <- c("t1d_duration_at_am_start") # A1c always reported as mean (SD)
t1 <- CreateTableOne(vars = c(cont_vars,cat_vars),data = demographics)
t1 <- print(t1,nonnormal = nonnormal,printToggle = F)
kable(t1)
```

```{r descriptive models,echo=FALSE}
# Rename redcap events to combine child and YA
mod_data <- data
mod_data$redcap_event_name <- factor(sub("_arm.*","",mod_data$redcap_event_name),
                                     levels = c("baseline","1","2","3","4","5"),
                                     labels = c("Baseline","M1","M3","M6","M9","M12"))
# Models and model means
# AM use
am_use_mod <- lme(time_am ~ redcap_event_name,random=~1|record_id,
                  data = mod_data[mod_data$redcap_event_name != "Baseline",],na.action = na.omit)
am_use_mod_means <- lme(time_am ~ redcap_event_name-1,random=~1|record_id,
                        mod_data[mod_data$redcap_event_name != "Baseline",],na.action = na.omit)
# TIR
tir_70_180_mod <- lme(sensor_70_180 ~ redcap_event_name,random=~1|record_id,data = mod_data,na.action = na.omit)
tir_70_180_mod_means <- lme(sensor_70_180 ~ redcap_event_name-1,random=~1|record_id,data = mod_data,na.action = na.omit)
# HbA1c
hba1c_mod <- lme(hba1c ~ redcap_event_name,random=~1|record_id,data = mod_data,na.action = na.omit)
hba1c_mod_means <- lme(hba1c ~ redcap_event_name-1,random=~1|record_id,data = mod_data,na.action = na.omit)
```

# Outcomes over time

## AM Use (%)

```{r am use diff, echo=FALSE}
diff <- summary(am_use_mod)$tTable
rownames(diff) <- sub("redcap_event_name","",rownames(diff))
kable(diff,caption = "Difference From Baseline")
```

This table shows the difference between each timepoint and baseline (the intercept). So for example, on average AM use was 14.7 points lower at month 6 compared to baseline (p = 0.0003).

```{r am use means, echo=FALSE}
means <- summary(am_use_mod_means)$tTable
rownames(means) <- sub("redcap_event_name","",rownames(means))
kable(means,caption = "Model Means")
```

This table shows the average AM use at each time point. The p-values here indicate whether or not the average is close to 0, so I would ignore them when reporting these results. 

The interpretation of the following tables is the same.

## TIR (%)

```{r tir diff, echo=FALSE}
diff <- summary(tir_70_180_mod)$tTable
rownames(diff) <- sub("redcap_event_name","",rownames(diff))
kable(diff,caption = "Difference From Baseline")
```

```{r tir means, echo=FALSE}
means <- summary(tir_70_180_mod_means)$tTable
rownames(means) <- sub("redcap_event_name","",rownames(means))
kable(means,caption = "Model Means")
```

## HbA1c (%)

```{r hba1c diff, echo=FALSE}
diff <- summary(hba1c_mod)$tTable
rownames(diff) <- sub("redcap_event_name","",rownames(diff))
kable(diff,caption = "Difference From Baseline")
```

```{r hba1c means, echo=FALSE}
means <- summary(hba1c_mod_means)$tTable
rownames(means) <- sub("redcap_event_name","",rownames(means))
kable(means,caption = "Model Means")
```

# ROC Analysis

```{r ROC analysis,echo=FALSE,message=FALSE}
roc <- roc(year,hba1c,av_am_use)
plot.roc(roc)
results <- as.data.frame(coords(roc,x="best",input = "threshold",best.method = "youden",
                  transpose = FALSE))
results$AUC <- auc(roc)
kable(results,caption = "ROC Analysis Results")
```

The optimal average AM use threshold based on the Youden index is 68.9%. This has a specificity of 62.5% and a sensitivity of 84.1% to detect whether or not HbA1c was <=7.5 % or >7.5 % after 12 months. AUC is 0.76.