---
title: "Clinic Team Clinic - Shideh Majidi"
author: "Kristen Campbell and Laura Pyle"
date: "`r Sys.Date()`"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(lsmeans)

#setwd("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/")
source("S:/Shared Material/Shared Code/R/temp_table1.R")

##functions specifically for this project
#source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/00_data.R")
source("C:/Users/pylell/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/00_data.R")
#source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/04_noshows.R")
source("C:/Users/pylell/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/04_noshows.R")

#source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/01_tables.R")
source("C:/Users/pylell/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/01_tables.R")
#source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/03_analysis_FULL.R")
source("C:/Users/pylell/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/03_analysis_FULL.R")



```

# Statistical Methods 

Data was summarized using frequency and percentage (categorical variables), mean and standard deviation (sd) (normally distributed continuous variables), or median and interquartile range (non-normally distributed continuous variables).  Comparisons of baseline data by group were performed using two sample independent t-tests, Chi-square tests, or Fisher's exact tests.  

Aim 1a: In Control patients only, the A1c will be different post-RTC depending on whether the patient participated in <4 versus 4+ RTC visits.  A linear mixed model was fit to control subjects using A1c as the outcome.  The main predictors were number of RTC visits (<4 versus 4+), time period (RTC versus post-RTC), and an interaction between number of RTC visits and time period.  The model was also adjusted for A1c at baseline, days from first visit, gender, insurance (public vs private), and Race/Ethnicity (Non-hispanic white versus other).  If the interaction term was significant, this would indicate that A1c trajectory in the post-RTC period differed depending on the number of RTC visits. 

Aim 1b: In CTC patients only, the A1c will be different post-RTC depending on whether the patient participated in <4 versus 4+ RTC visits.  The same modeling process was followed as in Aim 1a, except the model was only adjusted for baseline A1c, and days from first visit due to the small sample size.  

Aim 2: In the full cohort (Control+CTC), the trajectory of A1c Post-RTC will differ depending on whether the participant was in the control group or CTC group. A linear mixed model was fit to all subjects' post-RTC visit data with the outcome of A1c.  The main predictor was group (control versus CTC), and the model was adjusted for A1c at baseline, days from first visit, and number of RTC visits (<4 versus 4+).  If the group term was significant, this would indicate that A1c trajectory differed depending on study group.

# Summary Statistics 
**Table 1: Patient Demographics**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table1, align = c("l", "l"))
```

**Table 2: Diabetes Characteristics at Baseline**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table2, align = c("l", "l"))
```

\newpage

**P-values for change in A1c by Group**

These p-values correspond to a test of whether the change in A1c in each group (see mean change in Table 2) is significantly different from zero.

```{r,echo=FALSE,message=FALSE,warning=FALSE,comment=""}
t.test(dat.one[dat.one$group=="Control",]$avg_a1c_change,mu=0)
t.test(dat.one[dat.one$group=="CTC",]$avg_a1c_change,mu=0)
```

**Table 3: Visit Information**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table3, align = c("l", "l"))
```

**Figure 1: A1c Trajectories, CTC Group**
Figure from 02_plots.R: A1c_CTC.jpeg

Each plot represents a separate subject in the CTC group (N=14 total).  The solid red line is the subject's A1c over time, and the dotted black line represents the first CTC visit.

**Figure 2: A1c Trajectories, Control Group**
Figure from 02_plots.R: A1c_Control.jpeg

Each plot represents a separate subject in the Control group (a random sample of 14, out of the total group of `r nrow(subset(dat.one,dat.one$group=="Control"))`).  The solid red line is the subject's A1c over time, and the dotted black line represents the first Routine visit.

# Aim 1a: Control Group Only

<!-- # linear model -->
<!-- ```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE} -->
<!-- kable(lm_1a_tab, align = c("l", "l")) -->
<!-- ``` -->

# linear mixed model
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lmm_1a_sum, align = c("l", "l"))
```

After adjusting for all other variables in the model, baseline A1c was associated with average A1c (higher baseline = higher average) and the average A1c in the Post-RTC period was higher (`r lmm_1a_sum[7,3]`). There was no difference in A1c between RTC and Post-RTC time periods depending on whether the patient participated in 4+ RTC visits.

# Aim 1b: CTC Group Only

<!-- # linear model -->
<!-- ```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE} -->
<!-- kable(lm_1b_tab, align = c("l", "l")) -->
<!-- ``` -->

<!-- After adjusting for the average A1c during the RTC period, those with 4+ RTC visits had, on average a change in A1c (from RTC period to CTC period) of 0.08 lower than those with <4 RTC visits (non-significant).   In general, the direction is not intuitive (more RTC visits means A1c increased more) -->

# linear mixed model
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lmm_1b_sum, align = c("l", "l"))
```

After adjusting for all other variables in the model, baseline A1c was associated with average A1c (higher baseline = higher average) and the average A1c in the CTC period was higher (`r lmm_1b_sum[5,3]`). There was no difference in A1c between RTC and CTC time periods depending on whether the patient participated in 4+ RTC visits.

# Aim 2: All Patients

<!-- # linear model -->
<!-- ```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE} -->
<!-- kable(lm_2_tab, align = c("l", "l")) -->
<!-- ``` -->

# linear mixed model (post time period only)
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lmm_2_sum, align = c("l", "l"))
```

A1c did not differ significantly post-RTC for those in CTC group compared to the control group (p=`r lmm_2_sum[5,3]`).




