---
title: "Clinic Team Clinic - Shideh Majidi"
author: "Kristen Campbell"
date: "`r Sys.Date()`"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
#setwd("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/")
source("S:/Shared Material/Shared Code/R/temp_table1.R")

##functions specifically for this project
source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/00_data.R")
source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/04_noshows.R")

source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/01_tables.R")
source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Shideh Majidi/Clinic Team Clinic/03_analysis_FULL.R")



```

# Statistical Methods 

# Summary Statistics 
**Table 1: Patient Demographics**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table1, align = c("l", "l"))
```

**Table 2: Diabetes Characteristics at Baseline**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table2, align = c("l", "l"))
```

**Table 3: Visit Information**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table3, align = c("l", "l"))
```

**Figure 1: A1c Trajectories, CTC Group**
Figure from 02_plots.R: A1c_CTC.jpeg

Each plot represents a separate subject in the CTC group (N=14 total).  The solid red line is the subject's A1c over time, and the dotted black line represents the first CTC visit.

**Figure 2: A1c Trajectories, Control Group**
Figure from 02_plots.R: A1c_Control.jpeg

Each plot represents a separate subject in the Control group (a random sample of 14, out of the total group of `r nrow(subset(dat.one,dat.one$group=="Control"))`).  The solid red line is the subject's A1c over time, and the dotted black line represents the first Routine visit.

# Aim 1a: Control Group Only

# linear model
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lm_1a_tab, align = c("l", "l"))
```

# linear mixed model
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lmm_1a_sum, align = c("l", "l"))
```

# Aim 1b: CTC Group Only

# linear model
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lm_1b_tab, align = c("l", "l"))
```

# linear mixed model
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lmm_1b_sum, align = c("l", "l"))
```

# Aim 2: All Patients

# linear model
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lm_2_tab, align = c("l", "l"))
```

# linear mixed model (post time period only)
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(lmm_2_sum, align = c("l", "l"))
```






