---
title: "sleep_qa_ada2024"
author: "Casey Sakamoto"
date: "2023-11-28"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(redcapAPI)
library(tidyverse)
library(emmeans)
library(performance)
library(knitr)
library(table1)
library(readr)
library(readxl)
library(lme4)

# data upload
sd_df = read_csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2024/SleepDisruptors QA/SleepDisruptorsInAdo_DATA_2023-11-28_1536.csv")

# promis t tables
Depression_T <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/ADA 2024/SleepDisruptors QA/Depression T-tables.xlsx", 
    sheet = "Pedaitric Depressive Symptoms 8")
Depression_T$raw = Depression_T$`Raw Score`
Depression_T$depression_t = Depression_T$`T-score`


Anxiety_T <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/ADA 2024/SleepDisruptors QA/Anxiety T-tables.xlsx", 
    sheet = "Pediatric Anxiety 8a")
Anxiety_T$raw = Anxiety_T$`Raw Summed Score`
Anxiety_T$anxiety_t = Anxiety_T$`T-Score`


Psychological_Stress_T <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/ADA 2024/SleepDisruptors QA/Psychological Stress Experience T-tables.xlsx", 
    sheet = "Pediatric Psych Stress Exp 4a")
Psychological_Stress_T$raw= Psychological_Stress_T$`Raw Score`
Psychological_Stress_T$psych_stress_t = Psychological_Stress_T$`T-score`

Sleep_Disturbance_T <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/ADA 2024/SleepDisruptors QA/Sleep Disturbance and Sleep Related Impairment T Tables.xlsx", 
    sheet = "Pediatric Sleep Disturbance")
Sleep_Disturbance_T$raw = Sleep_Disturbance_T$`Raw Summed Score`
Sleep_Disturbance_T$sleep_disturbance_t = Sleep_Disturbance_T$`T-Score`


Impairment_T <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/ADA 2024/SleepDisruptors QA/Sleep Disturbance and Sleep Related Impairment T Tables.xlsx", 
    sheet = "Pediatric Sleep-Related Impairm")
Impairment_T$raw = Impairment_T$`Raw Summed Score`
Impairment_T$sleep_impairment_t = Impairment_T$`T-Score`
```


```{r data clean, include = FALSE}
# Data cleaning and transformation
sd_df = sd_df %>%
  mutate(
    # Recategorizing parent gender
    gender_m_fo = case_when(
      parent_gender == 1 ~ "Male",
      parent_gender == 2 | parent_gender == 3 ~ "Female/Other"
    ),
    # Categorizing age into groups
    age_cat = case_when(
      age < 12 & age >= 6 ~ "6-12",
      age >= 12 ~ "13-18"
    ),
    # Recoding race variables
    race = case_when(
      race___1 == 1 ~ "White, Non-Hispanic",
      race___2 == 1 ~ "Black/African American",
      race___3 == 1 ~ "Hispanic/Latino",
      race___4 == 1 ~ "Native Hawaiian/Pacific Islander",
      race___5 == 1 ~ "Asian",
      race___6 == 1 ~ "American Indian/Alaskan Native",
      race___8 == 1 ~ "Don't Wish to Answer",
      TRUE ~ "Unknown"
    )
  )

# Recoding categorical variables into factors with labels
sd_df$insulinmethod = factor(sd_df$insulinmethod, levels = c("1", "2"), labels = c("Injections", "Insulin Pump"))
sd_df$currenthcltech = factor(sd_df$currenthcltech, levels = c("1", "0"), labels = c("Yes", "No"))
sd_df$parenteducation = factor(sd_df$parenteducation, levels = c("2", "3", "4", "5", "6", "7", "8"),
                               labels = c("Less than Highschool", "GED", "Highschool Graduate",
                                          "Trade/Technical School", "Some College", "College Graduate", "Graduate School"))
sd_df$current_cgm_use = factor(sd_df$current_cgm_use, levels = c("1", "2"), labels = c("Yes", "No"))
sd_df$rurality = factor(sd_df$rurality, levels = c("1", "2"), labels = c("Rural", "Urban"))

# Extracting demographic covariates for analysis
demog_df = sd_df %>%
  select(record_id, redcap_event_name, gender_m_fo, age, age_cat, race, parenteducation, diabetesduration, 
         insulinmethod, currenthcltech, current_cgm_use, rurality) %>%
  filter(redcap_event_name == "enrollment_arm_1")

# Extracting HbA1c data from a specific survey event
a1c_col = sd_df %>%
  filter(redcap_event_name == "surveys_arm_1") %>%
  select(record_id, hba1c)

# Merging demographic and HbA1c data
demog_df = left_join(demog_df, a1c_col) %>%
  select(-record_id, -redcap_event_name) %>%
  filter(!is.na(gender_m_fo)) 
rm(a1c_col) # Removing temporary variable to free memory

# Selecting key demographic variables for analysis
demogs = sd_df %>%
  select(record_id, redcap_event_name, gender_m_fo, age_cat, currenthcltech, current_cgm_use) %>%
  filter(redcap_event_name == "enrollment_arm_1") %>%
  select(-redcap_event_name)

# Selecting key sleep and diabetes-related variables for analysis
analysis = sd_df %>%
  select(record_id, redcap_event_name,
         totalsleep_avg, onsetlatency_avg, sleepefficency_avg, waso_avg,   # Sleep measures
         sensor_70_180, sensor_181_250, sensor_g251, sensor_u54, sensor_55_69, sensor_mean # Diabetes measures
  ) %>%
  filter(redcap_event_name == "surveys_arm_1") %>%
  select(-redcap_event_name)

# Merging demographic data with analysis dataset
analysis = left_join(analysis, demogs) 
rm(demogs) # Removing temporary variable

# Creating new computed variables for hypoglycemia and hyperglycemia
analysis = analysis %>%
  mutate(sensor_hypoglycemic = sensor_u54 + sensor_55_69,
         sensor_hyperglycemic = sensor_181_250 + sensor_g251)

# Processing survey responses
survey_df = sd_df %>%
  mutate(
    # Calculating Morningness/Eveningness Questionnaire (MESC) total score
    mesc_total = (6 - me_oob) + me_getup + (5 - me_gym) + (5 - me_testtime) +
                 (5 - me_besttime) + (6 - me_p_bedtime) + me_malert + (6 - me_b_bedtime) +
                 me_6am + (5 - me_time2alert),
    # Adjusting survey responses to start from 0 instead of 1
    across(snore_ht:interrupts, ~ .x - 1)
  )

# Extracting MESC total scores
mesc = survey_df %>%
  select(record_id, mesc_total) %>%
  filter(!is.na(mesc_total))

# Calculating Sleep-Related Breathing Disorder (SRDB) score
srdb = survey_df %>%
  filter(redcap_event_name == "surveys_arm_2") %>%
  mutate(SRDB_score = rowMeans(across(snore_ht:interrupts), na.rm = T)) %>%
  select(record_id, SRDB_score)

# Standardizing record_id values
srdb$record_id = sub("^.", "1", srdb$record_id)

# Calculating Hypoglycemia Fear Survey (HFS) total score
hfs = survey_df %>%
  filter(redcap_event_name == "surveys_arm_1") %>%
  mutate(across(bedtime_snacks:having_low_sugar, ~ .x - 1),
         hfs_total = rowSums(across(bedtime_snacks:having_low_sugar), na.rm = FALSE)) %>%
  select(record_id, hfs_total)

# Processing Child PROMIS survey scores
promis = survey_df %>%
  filter(redcap_event_name == "surveys_arm_1") %>%
  mutate(
    sleep_dist_raw = rowSums(across(sd_falling:sd_tossed)),  # Sleep Disturbance
    sleep_imp_raw = rowSums(across(sri_sleepy:sri_mood)),   # Sleep Impairment
    psy_stress_raw = rowSums(across(ps_stressed:ps_manage)), # Psychological Stress
    anxiety_raw = rowSums(across(a_awful:a_bed)),           # Anxiety
    depress_raw = rowSums(across(ds_sad:ds_hard))           # Depression
  ) %>%
  select(record_id, sleep_dist_raw:depress_raw)

# Mapping raw scores to standardized T-scores
promis = left_join(promis, Anxiety_T, by = c('anxiety_raw' = 'raw')) %>%
  select(record_id, sleep_dist_raw:depress_raw, anxiety_t)

promis = left_join(promis, Depression_T, by = c('depress_raw' = 'raw')) %>%
  select(record_id, sleep_dist_raw:depress_raw, anxiety_t, depression_t)

promis = left_join(promis, Sleep_Disturbance_T, by = c('sleep_dist_raw' = 'raw')) %>%
  select(record_id, sleep_dist_raw:depress_raw, anxiety_t, depression_t, sleep_disturbance_t)

promis = left_join(promis, Impairment_T, by = c('sleep_imp_raw' = 'raw')) %>%
  select(record_id, sleep_dist_raw:depress_raw, anxiety_t, depression_t, sleep_disturbance_t, sleep_impairment_t)

promis = left_join(promis, Psychological_Stress_T, by = c('psy_stress_raw' = 'raw')) %>%
  select(record_id, sleep_dist_raw:depress_raw, anxiety_t, depression_t, sleep_disturbance_t, sleep_impairment_t, psych_stress_t)

# Merging survey data
surveys = full_join(promis, hfs) %>% unique()
surveys = full_join(surveys, srdb) %>% unique()
surveys = full_join(surveys, mesc) %>% unique()

# Merging all survey and analysis datasets
analysis = full_join(analysis, surveys) %>% unique()

# Converting total sleep time from HH:MM format to minutes
analysis$hoursc = vapply(strsplit(as.character(analysis$totalsleep_avg), ":"), `[`, 1, FUN.VALUE = character(1))
analysis$minsc = vapply(strsplit(as.character(analysis$totalsleep_avg), ":"), `[`, 2, FUN.VALUE = character(1))
analysis$sleeptotal_avg = as.numeric(analysis$hoursc) * 60 + as.numeric(analysis$minsc)
```



```{r functions, include = F}
fit_mod = function(outcome,df){

  # Fit random intercept model
  f = as.formula(paste0(outcome,"~ gender_m_fo"))
  mod = lm(f,data = df)
  
  f2 = as.formula(paste0(outcome,"~ gender_m_fo + factor(age_cat)"))
  mod2 = lm(f2,data = df)

  print("Base Model")
  cat("\n")
  print(summary(mod))
  cat("\n")
  print("Age Adj Model")
  cat("\n")
  print(summary(mod2))
}

pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
```

# Data remarks
For the age categories of 6-12/13-18 there was a pretty uneven split; I included the model results in case they were of interest, but especially with the categorical outcomes where we essentially stratify age and then gender group theres very few subjects in each category.

# Methods

Demographics stratified by gender group were produced. 

For numeric outcomes t-tests (adjusting for age and not adjusting) were performed. 

For categorical outcomes a chi square test of independence performed.

# Demographics

```{r demog table}
t1 = table1(~ . |gender_m_fo, data = demog_df, overall=F, extra.col=list(`P-value`=pvalue) )
t2 = table1(~ . |age_cat, data = demog_df )
t3 = table1(~ . |age_cat*gender_m_fo, data = demog_df )

t1;t2;t3
```


# Model Output

## Sleep Measures

### Total sleep

Gender was not significantly associated with avg total sleep (min) in either the base or age adjusted models (p = .86, 0.42)

```{r tsm}
fit_mod('sleeptotal_avg', analysis)
```

### Sleep latency

Gender was not significantly associated with sleep latency in either the base or age adjusted models (p = .74, .99)

```{r slm}
fit_mod('onsetlatency_avg', analysis)
```

### Sleep efficiency

Gender was not significantly associated with sleep efficiency in either the base or age adjusted models (p = .65, .70)

```{r sem}
fit_mod('sleepefficency_avg', analysis)
```

### waso

Gender was not significantly associated with WASO avgs in either the base or age adjusted models (p = .32, .57)

```{r wasom}
fit_mod('waso_avg', analysis)
```


## Diabetes measures 

### Time in range (70-180)

Gender was not significantly associated with TIR in either the base or age adjusted models (p = .83, .89)

```{r tirmm}
fit_mod('sensor_70_180', analysis)
```

### Time hypoglycemic (<70 – combine <54 and 55-69)

Gender was not significantly associated with TBR in either the base or age adjusted models (p = .06, .06)

```{r tbrmm}
fit_mod('sensor_hypoglycemic', analysis)
```

### Time hyperglycemic (>180 – combine 181-250 and >250)

Gender was not significantly associated with TBR in either the base or age adjusted models (p = .10, .10)

```{r tarmm}
fit_mod('sensor_hyperglycemic', analysis)
```

### HCL use (for those on HCL, if RedCap value blank do not include) 

Gender was not significantly associated with HCL use p = 1

```{r hclmm}
#table(analysis$gender_m_fo, analysis$currenthcltech)
chisq.test(analysis$gender_m_fo, analysis$currenthcltech)$p.value
```


### Sensor use (for those on sensor, if RedCap value blank do not include) -- log reg?

Gender was not significantly associated with sensor use p = 1

```{r summ}
#table(analysis$gender_m_fo, analysis$current_cgm_use)
chisq.test(analysis$gender_m_fo, analysis$current_cgm_use)$p.value
```

### Sensor Glucose Mean

Gender was not significantly associated with mean sensor glucose (p= 0.6), or in the age adjusted model (p = 0.65)

```{r msgs}
fit_mod("sensor_mean", analysis)
```

## Survey measures 

### Morningness/Eveningness Questionnaire 

Gender is not associated with mesc scores (p = 0.61) in the unadjusted model, and the age adjusted (p = 0.85).

```{r mescn}
fit_mod("mesc_total", analysis)
```

### SRDB Scale

Gender is associated with SRDB scores (p = 0.03) in the unadjusted model, but not the age adjusted (p = 0.07). Male subjects on average had 0.07 (SE = 0.033)lower score than female/other group.

```{r srdbn}
fit_mod("SRDB_score", analysis)
```

### Hypoglycemia Fear Survey child/Teen

Gender was not significantly associated with HFS total scores in both unadjusted and age adjusted models (p = 0.6, 0.5)

```{r hfsm}
fit_mod("hfs_total", analysis)
```

### Child_PROMIS_Sleep Disturbance

Gender was not significantly associated with promis sleep disturbance t scores scores in both unadjusted and age adjusted models (p = 0.18, 0.28)

```{r pdismsd}
fit_mod("sleep_disturbance_t", analysis)
``` 

### Child_PROMIS_Sleep Related Impairment 

Gender is associated with Sleep related impairment promis t scores  (p = 0.03) in the unadjusted model, but not the age adjusted (p = 0.10). Male subjects on average had 4.7 (SE = 2.1)lower score than the female/other group.

```{r promsd}
fit_mod("sleep_impairment_t", analysis)
``` 


### Child_PROMIS_Psychological stress

Gender is associated with psychological stress promis t scores  (p = 0.01) in the unadjusted model, and the age adjusted (p = 0.02). Male subjects on average had 5.1 (SE = 1.8)lower scores than the female/other group in the unadjusted model and 4.8 (se = 1.9) in the age adjusted model.

```{r pssmsd}
fit_mod("psych_stress_t", analysis)
``` 

### Child_PROMIS_Anxiety

Gender is associated with anxiety promis t scores  (p = 0.03) in the unadjusted model, but not the age adjusted (p = 0.07). Male subjects on average had 5.4 (SE = 2.4)lower scores than the female/other group.

```{r pand}
fit_mod("anxiety_t", analysis)
``` 


### Child_PROMIS_Depressive Symptoms 

Gender is associated with depresion promis t scores  (p = 0.005) in the unadjusted model, and the age adjusted (p = 0.01). Male subjects on average had 6.8 (SE = 2.3)lower scores than the female/other group in the unadjusted model and 6.5(se 2.5) in the adjusted.

```{r desd}
fit_mod("depression_t", analysis)
``` 