---
title: "sleep_qa_ada2024"
author: "Casey Sakamoto"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
library(redcapAPI)
library(tidyverse)
library(emmeans)
library(performance)
library(knitr)
library(table1)
library(readr)
# data upload
sd_df = read_csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2024/SleepDisruptors QA/SleepDisruptorsInAdo_DATA_2023-11-28_1536.csv")
```


```{r data clean, include = FALSE}
sd_df = sd_df %>% mutate(gender_m_fo = case_when(parent_gender == 1 ~ "Male",
                              parent_gender == 2 | parent_gender == 3 ~"Female/Other"
                              ),
                         age_cat = case_when(age < 12 & age >= 6 ~ "6-12",
                                             age >= 12 ~ "13-18"),
                         race = case_when(race___1 == 1 ~ "White, Non-Hispanic",
                                          race___2 == 1 ~ "Black/African American",
                                          race___3 == 1 ~ "Hispanic/Latino",
                                          race___4 == 1 ~ "Native Hawaiian/Pacific Islander",
                                          race___5 == 1 ~ "Asian",
                                          race___6 == 1 ~ "American Indian/Alaskan Native",
                                          race___8 == 1 ~ "Don't Wish to Answer",
                                          TRUE ~ "Unknown"
                                          ))
sd_df$insulinmethod = factor(sd_df$insulinmethod, levels = c("1","2"), labels = c("Injections", "Insulin Pump"))
sd_df$currenthcltech = factor(sd_df$currenthcltech, levels = c("1","0"), labels = c("Yes", "No"))
sd_df$current_cgm_use= factor(sd_df$current_cgm_use, levels = c("1","2"), labels = c("Yes", "No"))
sd_df$rurality= factor(sd_df$rurality, levels = c("1","2"), labels = c("Rural", "Urban"))

# table(sd_df$gender_m_fo)


# demog covariates
demog_df = sd_df %>% select(record_id, redcap_event_name,gender_m_fo ,age, age_cat,race , diabetesduration, insulinmethod, currenthcltech, current_cgm_use, rurality) %>% 
  filter(redcap_event_name == "enrollment_arm_1")

a1c_col = sd_df %>% filter(redcap_event_name == "surveys_arm_1") %>% select(record_id, hba1c) 

demog_df = left_join(demog_df, a1c_col) %>% select( -record_id, -redcap_event_name) %>% filter(!is.na(gender_m_fo)) ; rm(a1c_col)
#demog_df = left_join(demog_df, a1c_col) %>% filter(!is.na(gender_m_fo))
# table(demog_df$gender_m_fo)

#  run analyses w both age and agecat
#### TO DO ADD SURVEYS

demogs = sd_df %>% select(record_id,redcap_event_name,gender_m_fo ,age_cat, currenthcltech, current_cgm_use) %>% 
  filter(redcap_event_name == "enrollment_arm_1") %>% select(-redcap_event_name)
analysis = sd_df %>% select(record_id,redcap_event_name,
                            totalsleep_avg, onsetlatency_avg, sleepefficency_avg, waso_avg,   # sleep meas
                            sensor_70_180, sensor_181_250, sensor_g251, sensor_u54, sensor_55_69 # diab meas
) %>% filter(redcap_event_name == "surveys_arm_1") %>% select(-redcap_event_name)

analysis = left_join(analysis, demogs) ; rm(demogs)
```

Variables:
Sleep measures (RedCap wrist actigraphy variable)
Total sleep
Sleep latency
Sleep efficiency
WASO
Diabetes measures (RedCap Pump & CGM Glycemic Data OR Injection & BG Meter Glycemic Data Instruments [depending on the insulin method for each patient])
Time in range (70-180)
Time hypoglycemic (<70 – combine <54 and 55-69)
Time hyperglycemic (>180 – combine 181-250 and >250)
HCL use (for those on HCL, if RedCap value blank do not include) -- log reg?
Sensor use (for those on sensor, if RedCap value blank do not include) -- log reg

Also look at some of the survey measures – you should have scoring t-tables from Erin to do so
Morningness/Eveningness Questionnaire 
SRDB Scale
Hypoglycemia Fear Survey child/Teen
Child_PROMIS_Sleep Disturbance
Child_PROMIS_Sleep Related Impairment 
Child_PROMIS_Psychological stress
Child_PROMIS_Anxiety
Child_PROMIS_Depressive Symptoms 

```{r functions, include = F}
fit_mod = function(outcome,df,plot = T,diagnostics = F){

  # Fit random intercept model
  f = as.formula(paste0(outcome,"~ gender_m_fo"))
  mod = lmer(f,data = df)
  
  print(summary(mod))
}


```

Demographics:
Age (RedCap variable in baseline demographics instrument) --- for analyses interested in both cat and cts
Race/ethnicity (RedCap variable in baseline demographics instrument)
HbA1c (RedCap variable in Pump and CGM Glycemic Data form)
T1D duration (RedCap T1D duration equation in baseline demographics instrument) 
Number on pump (RedCap variable [insulinmethod] injection vs pump in baseline demographics instrument)
Number on HCL (RedCap variable [currenthcltech] closed loop system in baseline demographics instrument)
Number on sensor (RedCap variable [current_cgm_use] in baseline demographics instrument)
Percent rural (Redcap variable [rurality] in baseline demographics instrument)

```{r demog table, include = F}
t1 = table1(~ . |gender_m_fo, data = demog_df )
```

to do : score surveys, finish function (lin and log), fit, interpret models
```{r models, include=FALSE}


```

