---
title: "NOVO Transition"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,dpi = 600)
knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/NOVO Transition")
library(arsenal)
library(skimr)
library(knitr)
library(cgmanalysis)
library(tidyverse)
library(XML)
library(Hmisc)
library(emmeans)
```

```{r import}
# Import
source("/Users/timvigers/GitHub/BDC-Code/Viral Shah/Transition/NOVOTransitionStudy_R_2020-10-29_1804.R")
```

```{r cgm,eval=FALSE}
# Get cgm data function
cgm_by_timepoint = function(indir,outdir,regex,timepoint){
  # List patient folders
  l = list.files(indir)
  l_full = list.files(indir,full.names = T)
  l = l_full[!is.na(suppressWarnings(as.numeric(l)))]
  # Find timepoint files in patient folders
  f = lapply(l, function(f){
    files = list.files(f,full.names = T,pattern = ".*.xml")
    w = which(grepl(regex,tolower(files)))
    files[w]
  })
  f = unlist(f)
  # Import randomization files and write as csv to clean data folder
  dir.create(outdir,showWarnings = FALSE)
  lapply(f, function(file){
    ext = tools::file_ext(file)
    if(ext == "xml"){
      doc <- xmlParse(file)
      l <- xmlToList(doc)
      id = as.numeric(strsplit(file,"/")[[1]][4])
      l <- l[["GlucoseReadings"]]
      times <- lapply(l, function(x) {x[["DisplayTime"]]})
      times <- do.call(rbind,times)
      sensor <- lapply(l, function(x) {x[["Value"]]})
      sensor <- do.call(rbind,sensor)
      table <- as.data.frame(cbind(times,sensor))
      colnames(table) <- c("timestamp","sensorglucose")
      table$subjectid <- NA
      table$subjectid[1] <- id
      table <- table[,c("subjectid","timestamp","sensorglucose")]
      table$timestamp = lubridate::ymd_hms(table$timestamp)
      write.csv(table,file = paste0(outdir,id,"_",timepoint,".csv"),
                row.names = F,na="")
    }
  })
}
# Randomization
cgm_by_timepoint(indir="./Data_Raw/Patient data",
                 outdir = "./Data_Cleaned/Randomization/",
                 regex = "random",
                 timepoint = "randomization")
# Baseline
cgm_by_timepoint(indir="./Data_Raw/Patient data",
                 outdir = "./Data_Cleaned/Screening/",
                 regex = "screen",
                 timepoint = "screening")
# CGM variables
cgmvariables("./Data_Cleaned/Randomization/",
             "./Data_Cleaned","cgm_randomization")
cgmvariables("./Data_Cleaned/Screening/",
             "./Data_Cleaned","cgm_screening")
```

```{r data clean}
# Subject id from record id (record_id = 1 dropped out)
data$subject_id = data$record_id + 1
# Import CGM data
cgm_randomization = read.csv("./Data_Cleaned/cgm_randomization.csv")
colnames(cgm_randomization)[2:ncol(cgm_randomization)] = 
  paste0(colnames(cgm_randomization)[2:ncol(cgm_randomization)],"_r")
cgm_screening = read.csv("./Data_Cleaned/cgm_screening.csv")
colnames(cgm_screening)[2:ncol(cgm_screening)] = 
  paste0(colnames(cgm_screening)[2:ncol(cgm_screening)],"_s")
# Import insulin data
insulin = read.csv("./Data_Cleaned/insulin.csv",na.strings = "")
insulin$subject_id = as.numeric(sapply(strsplit(insulin$Id,"-"),"[[",2))
# Add to REDCap data
data = full_join(data,insulin,by = "subject_id")
data = full_join(data,cgm_randomization,by = "subject_id")
data = full_join(data,cgm_screening,by = "subject_id")
# Check if they have CGM data
data$cgm_randomization = factor(!is.na(data$date_cgm_placement_r),
                                labels = c("No","Yes"))
# Surveys
# IDSS score: mean of items 1-14 (reverse code items: 1, 3-4, 7-13)
reverse = c("idss_complicated","idss_embarrassing","idss_hassle_to_use",
            "idss_expensive","idss_spontaneous","idss_many_pieces",
            "idss_benefit","idss_inconvinient","idss_time",
            "idss_hassle_to_carry")
reverse_v2 = paste0(reverse,"_randomization_v2")
data[,reverse_v2] = lapply(data[,reverse_v2], function(x){6-x})
all_idss = c(reverse,"idss_good_control","idss_control",
             "idss_works_well","idss_future")
all_idss_v2 = paste0(all_idss,"_randomization_v2")
data$idss_randomization_v2 = 
  apply(data,1,function(r){mean(as.numeric(r[all_idss_v2]))})
# Change in % time > 180
data$change_time_over_180 = 
  data$percent_time_over_180_r - data$percent_time_over_180_s
# Insulin per kg
data$basal_dose = data$V1.Basal.dose/data$weight
data$bolus_dose = data$V1.Bolus.dose/data$weight
```

# Table 1: Descriptive Characteristics

```{r results='asis',warning=FALSE}
# Table 1
t1_vars = c("age","gender.factor","bmi","gold_score.factor",
            "diabetes_duration","a1c","insulin_pump.factor",
            "basal_dose","bolus_dose","education.factor",
            "insurance_private.factor","insurance_medicaid.factor",
            "insurance_medicare.factor","insurance_medical.factor",
            "insurance_no_answer.factor","household_income.factor",
            "diab_conditions___1.factor","diab_conditions___2.factor",
            "diab_conditions___3.factor","diab_conditions___4.factor",
            "diab_conditions___5.factor","cgm_randomization")
t1_form = as.formula(paste0("randomization_group.factor~",
                            paste0(t1_vars,collapse = "+")))
table_1 = 
  tableby(t1_form,data = 
            droplevels(data[!is.na(data$randomization_group.factor),]))
# Print
summary(table_1,labelTranslations = 
          list(age="Age",gender.factor="Gender",
               gold_score.factor = "Gold Score",
               insulin_pump.factor = "Insulin Pump",
               basal_dose = "Basal Dose",bolus_dose="Bolus Dose",
               education.factor = "Education",
               insurance_medicaid.factor = "Medicaid",
               insurance_medical.factor="No Medical Insurance",
               insurance_medicare.factor="Medicare",
               insurance_medicaid.factor="Medicaid",
               insurance_private.factor="Private",
               insurance_no_answer.factor="No Answer",
               household_income.factor="Household Income",
               diab_conditions___1.factor="Diabetic Retinopathy",
               diab_conditions___2.factor="Diabetic Neuropathy",
               diab_conditions___3.factor="Diabetic Nephropathy",
               diab_conditions___4.factor="High Cholesterol",
               diab_conditions___5.factor="High Blood Pressure",
               cgm_randomization = "CGM Data at Randomization"))
```

# Figure 1: Percent time spent > 180 mg/dL

```{r fig 1,warning=FALSE}
ggplot(data[!is.na(data$randomization_group.factor),],
       aes(x=randomization_group.factor,y=percent_time_over_180_r))+
  geom_boxplot() + 
  theme_bw() + xlab("") + ylab("% Time Above 180 mg/dL")
```

# Figure 2: Change in percent time spent > 180 mg/dL

```{r fig 2,warning=FALSE}
ggplot(data[!is.na(data$randomization_group.factor),],
       aes(x=randomization_group.factor,y=change_time_over_180))+
  geom_boxplot() + 
  theme_bw() + xlab("") + ylab("Change in % Time Above 180 mg/dL")
```

# Table 2: Differences between treatment group adjusted for BMI and Medicaid status

Means and standard errors (SE) are averaged over the levels of Medicaid status and reported for the average BMI in each treatment group. All outcomes were compared using linear regression except for the number of correction boluses, which was modeled using Poisson regression.

```{r table 2}
data$no_of_corrections = 
  suppressWarnings(
    as.numeric(
      data$No.of.correction.boluses.in.first.72.hours.of.randomization))
vars = list("% TIR"="percent_time_70_180_r",
            "% > 180 mg/dL"="percent_time_over_180_r",
            "Change in % > 180 mg/dL"="change_time_over_180",
            "% Time < 70 mg/dL"="percent_time_under_70_r",
            "% Time < 54 mg/dL"="percent_time_under_54_r",
            "Number of correction boluses in first 72 hours"=
              "no_of_corrections","IDSS"="idss_randomization_v2")
group = "randomization_group.factor"
table_2 = lapply(vars,function(var){
  form = 
    as.formula(paste0(var,"~",group,"+bmi+insurance_medicaid.factor"))
  mod = lm(form,data = data)
  if(grepl("no",var)){
    mod = glm(form,data = data,family = "poisson")
  }
  res = as.data.frame(summary(mod)$coefficients)
  p = res[2,4]
  means = as.data.frame(emmeans(mod,as.formula(paste0("~",group))))
  return(c(means[1,2],means[1,3],means[2,2],means[2,3],p))
})
table_2 = as.data.frame(do.call(rbind,table_2))
rownames(table_2) = names(vars)
kable(table_2,col.names = c("SOC Mean","SOC SE","OLP Mean","OLP SE","p"),
      digits = 3)
```
