---
title: ""
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(broom)
library(survival)
library(performance)
library(knitr)
library(pROC)
library(naniar)
library(readxl)
library(psych)
library(pheatmap)
library(tableone)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggplot2)
library(reshape2)
library(lubridate)


# load in metab data
QAHA_Biomarkers_IDs = read_excel("S:/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Combined predictive model/QAHA_Biomarkers_IDs.xls")
```

```{r data clean, include = FALSE}
# the goal is to transform each of these metrics into a slope (visit max - visit1)/(datemax - date 1) - dates in years
test = QAHA_Biomarkers_IDs %>% select(StudyID, Date_V1:Date_V4, ACV1, acv2, acv3, acv4, CKDepiV1:CKDEpiV4, age, gender, durcatv1)
test$acv2 = as.numeric(test$acv2)
test$acv3 = as.numeric(test$acv3)

test$CKDepiV1 = as.numeric(test$CKDepiV1)
test$ckdepiv2 = as.numeric(test$ckdepiv2)
test$ckdepiv3 = as.numeric(test$ckdepiv3)

# All subjects have a visit 1 date but 14 have no acv1 -- drop these or use a 2 if they have it?
# All subjects with no visit 2 or 3 dates also dont have a acv value


# 13 subjects have an acv4 value but no date -- ask janet

# 8 subj with only acv1
options(scipen = 20)
test2 = test %>% mutate(slope_acv = case_when(!is.na(Date_V4) & !is.na(acv4) ~ (acv4 - ACV1)/(as.double(difftime(Date_V4, Date_V1, units = "days"))/365),
                                              !is.na(Date_V3) & !is.na(acv3) ~ (acv3 - ACV1)/(as.double(difftime(Date_V3, Date_V1, units = "days"))/365),
                                              !is.na(Date_V2) & !is.na(acv2) ~ (acv2 - ACV1)/(as.double(difftime(Date_V2, Date_V1, units = "days"))/365)
                                              ),
                        slope_egfr = case_when(!is.na(Date_V4) & !is.na(CKDEpiV4) ~ (CKDEpiV4 - CKDepiV1)/(as.double(difftime(Date_V4, Date_V1, units = "days"))/365),
                                              !is.na(Date_V3) & !is.na(ckdepiv3) ~ (ckdepiv3 - CKDepiV1)/(as.double(difftime(Date_V3, Date_V1, units = "days"))/365),
                                              !is.na(Date_V2) & !is.na(ckdepiv2) ~ (ckdepiv2 - CKDepiV1)/(as.double(difftime(Date_V2, Date_V1, units = "days"))/365)))

# 22 subj missing; 8 with only acv1, 14 without, of the 14, 9 have at least 2 dates -- exclude

# Function and master data
#source("D:/Repositories/shared-resources/Machine Learning/Tim - ElasticNet CV/easy_elasticnet.R")
source("C:/Repositories/shared-resources/Machine Learning/Tim - ElasticNet CV/easy_elasticnet.R")
load("S:/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/aha_master_data_no_snps.Rdata")
analysis = test2; rm(test, test2)

# get omics stuff from each sheet
omics_umm_fdr <- read_excel("S:/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Reports/AHA_omics_ttests_baseline_renal.xlsx", 
    sheet = "untargeted_metab_moderated_FDR") %>% filter(adj.P.Val < 0.05) %>% mutate(name = `...1`) %>% select(name)
omics_tmm_fdr <- read_excel("S:/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Reports/AHA_omics_ttests_baseline_renal.xlsx", 
    sheet = "targeted_metab_moderated_FDR")%>% filter(adj.P.Val < 0.05)%>% mutate(name = `...1`) %>% select(name)
omics_gpr_fdr <- read_excel("S:/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Reports/AHA_omics_ttests_baseline_renal.xlsx", 
    sheet = "global_prot_moderated_FDR")%>% filter(adj.P.Val < 0.05)%>% mutate(name = `...1`) %>% select(name)
omics_glpr_fdr <- read_excel("S:/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Reports/AHA_omics_ttests_baseline_renal.xlsx", 
    sheet = "glycated_prot_moderated_FDR")%>% filter(adj.P.Val < 0.05)%>% mutate(name = `...1`) %>% select(name)
omics_lm_fdr <- read_excel("S:/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Reports/AHA_omics_ttests_baseline_renal.xlsx", 
    sheet = "lipidomics_moderated_FDR")%>% filter(adj.P.Val < 0.05)%>% mutate(name = `...1`) %>% select(name)

omics = full_join(omics_umm_fdr, omics_tmm_fdr)
omics = full_join(omics, omics_glpr_fdr)
omics = full_join(omics, omics_lm_fdr)
omics = full_join(omics, omics_gpr_fdr)

omics = unique(c(omics$name))
limited_omics = omics

# vars not in the original dataset
names = colnames(df)
nothere = setdiff(omics, names)
omics = setdiff(omics, nothere)
limited_omics = omics
# omics selected by moderated t-test and PLSDA

# this was from prior code
# omics = unique(c("gly_P02671.7","gly_P02671.5","gly_P02671.4","gly_P02671.3",
#           "AAA01201000179.0807263.4","gly_P02679","HMDB00510161.0696402.9",
#           "AAA01201000179.0809235.7","gly_P02675.4","gly_P02675","gly_P02647",
#           "P02655","gly_P02647.1","HMDB00510161.0694436","gly_P02652",
#           "gly_P01009","gly_P01834.11","P01817","AcCa 10:3",
#           "HMDB0028822260.1378628.6","AAA01201000179.0807263.4",
#           "gly_P02671.3","gly_P02671.7"))
# limited_omics = c("AAA01201000179.0807263.4","gly.P02671.7","gly.P02671.3",
#           "gly.P02671.5","gly.P02671.4")


# Take out people without diabetes, remove diabetes status from predictors
df = df[df$diabetic == 1,]
clinical_predictors = clinical_predictors[-which(clinical_predictors == "dia")]

# Log transform 
df[,omics] = log(df[,omics])
# Fix names
colnames(df)[which(colnames(df) %in% omics)] = make.names(omics,unique = T,allow_ = F)
omics = make.names(omics,unique = T,allow_ = F)

limited_omics = omics
# Variables for null model
null_vars <- c("age","gender","durcatv1")

analysis$StudyID = as.numeric(analysis$StudyID)

test = analysis %>% select(StudyID, slope_acv, slope_egfr)
analysis2 = left_join( df, test)
```

# appended 6/5/2023

Some significant omics variables from the spreadsheets were not in the original dataset. These were:

```{r}
print(nothere)
```



# Change in ACV

## Null Model (Age, Sex, DiaDur)
```{r}
n_acv_change  = as.formula(paste0("slope_acv~",paste0(null_vars,collapse = "+")))
null_acv_mod = lm(formula = n_acv_change,data = analysis)
kable(tidy(null_acv_mod),digits = 3)
```

## Change in acv per year

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
acv_change_select = easy_elasticnet(data = analysis2,outcome = "slope_acv",
                           predictors = clinical_predictors,
                           model_type = "gaussian")
```

## Clinical model results

```{r}
f_acv_change = as.formula(paste0("slope_acv ~",
                                 paste0(acv_change_select,collapse = "+")))
clin_acv_mod = lm(formula = f_acv_change,data = analysis2)
kable(tidy(clin_acv_mod),digits = 3)
```

## Omics model results (no imputation)

```{r}
omics_acv_mod = lm(formula = update(f_acv_change,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
         data = analysis2)
kable(tidy(omics_acv_mod),digits = 3)
```

## Model comparison

```{r}
mc_acv = as.data.frame(compare_performance(clin_acv_mod,omics_acv_mod,null_acv_mod,
                                       metrics = c("R2","R2_adj"),verbose = F))
mc_acv$Name = c("Clinical Model","Omics Model","Null model")
kable(mc_acv,digits = 3)
```

# Change in EGFR

## Null Model (Age, Sex, DiaDur)
```{r}
n_egfr_change  = as.formula(paste0("slope_egfr~",paste0(null_vars,collapse = "+")))
null_egfr_mod = lm(formula = n_egfr_change,data = analysis)
kable(tidy(null_egfr_mod),digits = 3)
```

## Change in egfr per year

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
egfr_change_select = easy_elasticnet(data = analysis2,outcome = "slope_egfr",
                           predictors = clinical_predictors,
                           model_type = "gaussian")
```

## Clinical model results

```{r}
f_egfr_change = as.formula(paste0("slope_egfr ~",
                                 paste0(egfr_change_select,collapse = "+")))
clin_egfr_mod = lm(formula = f_egfr_change,data = analysis2)
kable(tidy(clin_egfr_mod),digits = 3)
```

## Omics model results (no imputation)

```{r}
omics_egfr_mod = lm(formula = update(f_egfr_change,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
         data = analysis2)
kable(tidy(omics_egfr_mod),digits = 3)
```

## Model comparison

```{r}
mc_egfr = as.data.frame(compare_performance(clin_egfr_mod,omics_egfr_mod,null_egfr_mod,
                                       metrics = c("R2","R2_adj"),verbose = F))
mc_egfr$Name = c("Clinical Model","Omics Model","Null model")
kable(mc_egfr,digits = 3)
```
