---
title: "Insulin Cost Power Calculation"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(pwr)
library(ggplot2)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Todd Alonso/Waterman Alonso insulin pricing"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
```

# Hypothesis

Patients with private insurance, and are based in Colorado, will observe a more significant reduction in co-payments compared to those on public insurance or those with out of state insurance.

# Methods

The detectable effect sizes were calculated for significance level 0.05, power 0.8, and a sample size range of 100-1000 in increments of 100. Effect size is measured in Cohen's D, where $D = \frac{|\mu_1 - \mu_2|}{\sigma}$. Cohen's D represents the detectable effect size in terms of the standard deviation. Cohen suggests that D values of 0.2, 0.5, and 0.8 represent small, medium, and large effect sizes.

# Analysis

```{r pwr calc}
# mentioned the sample size could be in the range of the 100s
# change the sample size
n = seq(50, 500, 50)
nn = length(n)


# obtain sample sizes
effsize = rep(NA, nn)
for (i in 1:nn){
    result = pwr.t.test(n = n[i],
    sig.level = .05, power = 0.8,
    alternative = "two.sided")
    effsize[i] = result$d
}
df = as.data.frame(cbind(n, effsize))
# set up graph
plot(y = effsize, x = n,
 type="b",
  ylab="Cohen's D",
  xlab="n in each group" )

# add annotation (grid lines, title, legend)
title("Sample Size Estimation for Cohen's D Effect Size")
```

* With a sample size of 100 (50 per group), the detectable effect size is 0.57
* With a sample size of 300, the detectable effect size is 0.32
* With a sample size of 500, the detectable effect size is 0.25
* With a sample size of 1000, the detectable effect size is 0.18


