---
title: "Cobry: JDRF Phase 1 Manuscript"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
---

```{r setup, include=FALSE}
library(redcapAPI)
library(tidyverse)
library(emmeans)
library(performance)
library(knitr)
#library(table1)
library(arsenal)
library(readr)
library(readxl)
library(lme4)

knitr::opts_chunk$set(echo = FALSE, comment = "")

# promis tables
Depression_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Depression T-tables.xlsx", 
    sheet = "Pedaitric Depressive Symptoms 8")
Depression_T$raw = Depression_T$`Raw Score`
Depression_T$child_depression_t = Depression_T$`T-score`
# LP added: parent version of depression
pp_dep_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Depression T-tables.xlsx", 
                          sheet = "Parent Proxy Depressive Sx 6a")
pp_dep_short$parent_depression_t <- pp_dep_short$`T-score`

Anxiety_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Anxiety T-tables.xlsx", 
    sheet = "Pediatric Anxiety 8a")
Anxiety_T$raw = Anxiety_T$`Raw Summed Score`
Anxiety_T$child_anxiety_t = Anxiety_T$`T-Score`
# LP added: parent version of anxiety
pp_anx_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Anxiety T-tables.xlsx", 
                          sheet = "Parent Proxy Anxiety 8a")
pp_anx_short$parent_anxiety_t <- pp_anx_short$`T-Score`

Psychological_Stress_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Psychological Stress Experience T-tables.xlsx", 
    sheet = "Pediatric Psych Stress Exp 4a")
Psychological_Stress_T$raw= Psychological_Stress_T$`Raw Score`
Psychological_Stress_T$child_psych_stress_t = Psychological_Stress_T$`T-score`
# LP added: parent version of psychological stress
pp_pse_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Psychological Stress Experience T-tables.xlsx", 
                             sheet = "Parent Proxy Psy Stress Exp 4a")
pp_pse_short$parent_psych_stress_t <- pp_pse_short$`T-score`

Sleep_Disturbance_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Sleep Disturbance and Sleep Related Impairment T Tables.xlsx", 
    sheet = "Pediatric Sleep Disturbance")
Sleep_Disturbance_T$raw = Sleep_Disturbance_T$`Raw Summed Score`
Sleep_Disturbance_T$child_sleep_disturbance_t = Sleep_Disturbance_T$`T-Score`
# do not see parent version of sleep disturbance

Impairment_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Sleep Disturbance and Sleep Related Impairment T Tables.xlsx", 
    sheet = "Pediatric Sleep-Related Impairm")
Impairment_T$raw = Impairment_T$`Raw Summed Score`
Impairment_T$child_sleep_impairment_t = Impairment_T$`T-Score`
# do not see parent version of sleep impairment

# LP added
# do not see parent version of life satisfaction
peds_lifesatisfaction_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Life Satisfaction T-tables.xlsx", 
                                         sheet = "Pedaitric Life Satisfaction 4a")
peds_lifesatisfaction_short$child_lifesastisfaction_t <- peds_lifesatisfaction_short$`T-Score`
# child and parent version of physical activity
peds_physact_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Physical Activity T-tables.xlsx", 
                                          sheet = "Pediatric Physical Activity 4a")
peds_physact_short$child_physact_t <- peds_physact_short$`T-score`
pp_physact_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Physical Activity T-tables.xlsx", 
                                 sheet = "Parent Proxy Phys Activity 4a")
pp_physact_short$parent_physact_t <- pp_physact_short$`T-score`
# child and parent version of positive affect
peds_posaff_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Positive Affect T-tables.xlsx", 
                                 sheet = "Pediatric Positive Affect 4a")
peds_posaff_short$child_posaff_t <- peds_posaff_short$`T-score`
pp_posaff_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Positive Affect T-tables.xlsx", 
                              sheet = "Parent Proxy Positive Affect 4a")
pp_posaff_short$parent_posaff_t <- pp_posaff_short$`T-score`
# child and parent version of anger
peds_anger_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Anger T-tables.xlsx", 
                            sheet = "Pediatric Self Report 5a")
peds_anger_short$child_anger_t <- peds_anger_short$`T-Score`
pp_anger_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Anger T-tables.xlsx", 
                         sheet = "Parent Proxy 5a")
pp_anger_short$parent_anger_t <- pp_anger_short$`T-Score`
# child and family versions of family relationships
peds_fr_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Relationships T-tables.xlsx", 
                            sheet = "Pediatric Family Relation 8a")
peds_fr_short$child_fr_t <- peds_fr_short$`T-score`
pp_fr_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Relationships T-tables.xlsx", 
                          sheet = "Parent Prox Family Relat 8a")
pp_fr_short$parent_fr_t <- pp_fr_short$`T-score`
# child version of peer relationships - do not see parent
peds_pr_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Relationships T-tables.xlsx", 
                             sheet = "Pediatric Peer Relationships 8a")
peds_pr_short$child_pr_t <- peds_pr_short$`T-score`
```


```{r data clean, include = FALSE}
# data upload
sd_df = read_csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/SleepDisruptorsInAdo_DATA_2024-10-30_0826.csv")

sd_df = sd_df %>% mutate(gender_m_fo = case_when(parent_gender == 1 ~ "Male",
                              parent_gender == 2 | parent_gender == 3 ~"Female/Other"
                              ),
                         age_cat = case_when(age < 12 & age >= 6 ~ "6-12",
                                             age >= 12 ~ "13-18"),
                         race = case_when(race___1 == 1 ~ "White, Non-Hispanic",
                                          race___2 == 1 ~ "Black/African American",
                                          race___3 == 1 ~ "Hispanic/Latino",
                                          race___4 == 1 ~ "Native Hawaiian/Pacific Islander",
                                          race___5 == 1 ~ "Asian",
                                          race___6 == 1 ~ "American Indian/Alaskan Native",
                                          race___8 == 1 ~ "Don't Wish to Answer",
                                          TRUE ~ "Unknown"
                                          ))
sd_df$insulinmethod = factor(sd_df$insulinmethod, levels = c("1","2"), labels = c("Injections", "Insulin Pump"))
sd_df$currenthcltech = factor(sd_df$currenthcltech, levels = c("1","0"), labels = c("Yes", "No"))
sd_df$parenteducation = factor(sd_df$parenteducation, levels = c("2","3","4","5","6","7","8"),
                               labels = c("Less than Highschool", "GED", "Highschool Graduate",
                                          "Trade/Technical School", "Some College", "College Graduate", "Graduate School"))
sd_df$current_cgm_use= factor(sd_df$current_cgm_use, levels = c("1","2"), labels = c("Yes", "No"))
sd_df$rurality= factor(sd_df$rurality, levels = c("1","2"), labels = c("Rural", "Urban"))

# Score PROMIS
source("/Users/pylell/Documents/GitHub/BDC-Code/Erin Cobry/JDRF Phase 1 manuscript/promis-scoring-r.r")
child_scores <- score_all_promis_scales(sd_df, respondent = "child")
child_scores <- dplyr::select_all(child_scores, .funs = funs(paste0("c_", .)))
sd_df <- cbind(sd_df, child_scores)
parent_scores <- score_all_promis_scales(sd_df, respondent = "parent")
parent_scores <- dplyr::select_all(parent_scores, .funs = funs(paste0("p_", .)))
sd_df <- cbind(sd_df, parent_scores)
# add scaled scores
source("/Users/pylell/Documents/GitHub/BDC-Code/Erin Cobry/JDRF Phase 1 manuscript/promis scaled scores v3.r")
# Create lookup tables
lookup_tables <- create_promis_lookup_tables()
all_scores <- convert_all_to_tscores(scores = sd_df, lookup_tables = lookup_tables)

# now read in the parent version of family relationships
p_fam <- read.csv('/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/SleepDisruptorsInAdo-PROMISParentFamRelat_DATA_2025-03-18_0745.csv')
p_fam <- p_fam %>% select(record_id, redcap_event_name, promis_parent_proxy_sf_v10_fam_relationships_8a_tscore)
all_scores <- left_join(all_scores, p_fam, by = c("record_id", "redcap_event_name"))

# overwrite sd_df with df that contains t-scores
sd_df <- all_scores

# demog covariates
demog_df = sd_df %>% select(record_id, redcap_event_name, season, gender_m_fo, age, age_cat, stateresidence, rurality, race, 
                            childsgradeinschool, school_format, diabetesduration, insulinmethod, pump_brand, current_cgm_use,
                            cgm_type, currenthcltech, dka, recent_dka, severe_hypo_episode, recent_severe_hypo,
                            parenteducation, child_sleep_condition, child_nightime_awakening, nightime_awakening_2,
                            parent_sleep_concerns, concern_sleep_disorder, intvent_benefit) %>% 
  filter(redcap_event_name == "enrollment_arm_1")
a1c_col = sd_df %>% filter(redcap_event_name == "surveys_arm_1") %>% select(record_id, hba1c) 
demog_df = left_join(demog_df, a1c_col, by = "record_id") %>% select(-redcap_event_name) %>% filter(!is.na(gender_m_fo)) ; rm(a1c_col)
tanner_df <- sd_df %>% select(record_id, redcap_event_name, mstage2, fstage2, drstage) %>% 
  filter(redcap_event_name == "enrollment_arm_1")
tanner_df <- tanner_df %>% rowwise() %>% mutate(self_stage = max(mstage2, fstage2, na.rm = T))
tanner_df$self_stage <- ifelse(is.infinite(tanner_df$self_stage), NA, tanner_df$self_stage)
demog_df <- left_join(demog_df, tanner_df, by = "record_id")

# parent scores are in survey arm 2
pscores <- sd_df %>% filter(redcap_event_name == "surveys_arm_2") %>%
  select(record_id, p_physical_tscore, p_positive_affect_tscore, p_psychological_stress_tscore,
                            p_anger_tscore, p_anxiety_tscore, p_depressive_tscore, p_family_tscore, p_peer_tscore,
                            promis_parent_proxy_sf_v10_fam_relationships_8a_tscore)
# need to make IDs match
pscores$record_id <- str_replace(pscores$record_id, "2", "1")

analysis = sd_df %>% select(record_id,redcap_event_name,
                            time_am, time_mm, sensor_wear, eve_sensor_low, eve_sensor_below,
                            eve_sensor_target, eve_sensor_above, eve_sensor_high, eve_sensor_mean, eve_sensor_sd,
                            night_sensor_low, night_sensor_below, night_sensor_target, night_sensor_above,
                            night_sensor_high, night_sensor_mean, night_sensor_sd, sensor_g251,
                            sensor_181_250, sensor_70_180, sensor_55_69, sensor_u54, sensor_mean, sensor_sd,
                            gmi, tdd, tdd_basal, tdd_bolus, dailybolus, dailymealbolus, sleepmode, exercisemode,
                            avg_bedtime, getup_avg, timeinbed_avg, onsetlatency_avg, sleepefficency_avg, 
                            waso_avg, awake_avg, c_life_satisfaction_tscore, c_physical_tscore, 
                            c_positive_affect_tscore,  c_psychological_stress_tscore, 
                             c_anger_tscore,  c_anxiety_tscore,
                            c_depressive_tscore,  c_family_tscore,
                             c_peer_tscore,  c_sleep_disturbance_tscore,
                            c_sleep_impairment_tscore) %>% filter(redcap_event_name == "surveys_arm_1") %>% select(-redcap_event_name)
analysis = left_join(analysis, pscores, by = "record_id")

analysis = left_join(demog_df, analysis) 
analysis = analysis %>% mutate(sensor_hypoglycemic = sensor_u54 + sensor_55_69,
                               sensor_hyperglycemic = sensor_181_250 + sensor_g251)


survey_df = sd_df %>% mutate(mesc_total = (6 - me_oob) + me_getup + (5 - me_gym) + (5 - me_testtime)
                             + (5 - me_besttime) + (6 - me_p_bedtime) + me_malert + (6-me_b_bedtime)
                             + me_6am + (5 - me_time2alert),# Morningness/Eveningness Questionnaire
                             across(snore_ht:interrupts, ~ .x - 1)) 
mesc = survey_df %>% select(record_id, mesc_total)
mesc = mesc %>% filter(!is.na(mesc_total))
  
srdb = survey_df %>% filter(redcap_event_name %in% c("surveys_arm_1", "surveys_arm_2")) %>% mutate(SRDB_score = rowMeans(across(snore_ht:interrupts), na.rm = T)) %>%
  select(record_id, SRDB_score) # SRDB Scale
srdb$record_id = sub("^.", "1", srdb$record_id)
srdb <- srdb %>% filter(!is.na(SRDB_score))

hfs = survey_df %>% filter(redcap_event_name == "surveys_arm_1") %>% mutate(across(bedtime_snacks:having_low_sugar, ~ .x - 1),
                           hfs_total = rowSums(across(bedtime_snacks:having_low_sugar), na.rm=F)) %>% select(record_id, hfs_total)# Hypoglycemia Fear Survey child/Teen

surveys = full_join(srdb, hfs, by = c("record_id")) %>% unique()
surveys = full_join(surveys, mesc) %>% unique()
surveys = surveys %>% filter(!is.na(mesc_total) | !is.na(SRDB_score) | !is.na(hfs_total))
analysis = left_join(analysis,surveys, by = c("record_id")) %>% unique()

analysis$hoursc = vapply(strsplit(as.character(analysis$sleeptotal_avg),":"), `[`, 1, FUN.VALUE=character(1))
analysis$minsc = vapply(strsplit(as.character(analysis$sleeptotal_avg),":"), `[`, 2, FUN.VALUE=character(1))
analysis$sleeptotal_avg = as.numeric(analysis$hoursc)*60 + as.numeric(analysis$minsc)
analysis$sleeptotal_avg <- as.numeric(analysis$sleeptotal_avg) / 60

# create variable for sufficient sleep, using 7 hours for 13 and older, 8 hours for younger than 13
analysis$sufficient_sleep <- ifelse(analysis$age < 13 & analysis$sleeptotal_avg >= 8, 1, 
                                    ifelse(analysis$age < 13 & analysis$sleeptotal_avg < 8, 0,
                                           ifelse(analysis$age >= 13 & analysis$sleeptotal_avg >= 7, 1, 0)))
```

```{r functions, include = F}
fit_mod = function(pred,df){

  # Fit random intercept model
  #f = as.formula(paste0(outcome,"~ gender_m_fo"))
  f = as.formula(paste0("sleeptotal_avg ~", pred))
  
  
  mod = lm(f,data = df)
  
  f2 = as.formula(paste0("sleeptotal_avg ~", pred, "+ factor(age_cat)"))
  mod2 = lm(f2,data = df)

  print("Base Model")
  cat("\n")
  print(summary(mod))
  cat("\n")
  print("Age Adj Model")
  cat("\n")
  print(summary(mod2))
}

fit_mod_log = function(pred,df){

  # Fit random intercept model
  #f = as.formula(paste0(outcome,"~ gender_m_fo"))
  f = as.formula(paste0("sufficient_sleep ~", pred))
  
  
  mod = glm(f,data = df, family = "binomial")
  
  f2 = as.formula(paste0("sufficient_sleep ~", pred, "+ factor(age_cat)"))
  mod2 = glm(f2,data = df, family = "binomial")

  print("Base Model")
  cat("\n")
  print(summary(mod))
  cat("\n")
  print("Age Adj Model")
  cat("\n")
  print(summary(mod2))
}

fit_mod_gender = function(outcome,df){
  f = as.formula(paste0(outcome,"~ gender_m_fo"))

  mod = lm(f,data = df)
  
  f2 = as.formula(paste0(outcome , "~ gender_m_fo + factor(age_cat)"))
  mod2 = lm(f2,data = df)

  print("Base Model")
  cat("\n")
  print(summary(mod))
  cat("\n")
  print("Age Adj Model")
  cat("\n")
  print(summary(mod2))
}


pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
```

# LP data comments

- ID 1032AH-2 removed

# Methods

Demographics stratified by gender group were produced. 

Linear regressions with total sleep time as the outcome were performed, both unadjusted and adjusted for age category.

# Demographics

## By gender group

```{r create_t1, include=FALSE}
t1 <- tableby(gender_m_fo ~ as.factor(season) + gender_m_fo + age + age_cat + as.factor(stateresidence) + rurality + race + as.factor(childsgradeinschool) + as.factor(school_format) + 
diabetesduration + insulinmethod + as.factor(pump_brand) + current_cgm_use + as.factor(cgm_type) + 
currenthcltech + as.factor(dka) + as.factor(severe_hypo_episode) + 
parenteducation + as.factor(child_sleep_condition) + as.factor(child_nightime_awakening) + nightime_awakening_2 + parent_sleep_concerns + concern_sleep_disorder + intvent_benefit + as.factor(hba1c) + as.factor(mstage2) + 
as.factor(fstage2) + as.factor(drstage) + as.factor(self_stage), data = demog_df)
```

```{r t1, results='asis'}
summary(t1)
```

## By age group

```{r, include=FALSE}
t2 <- tableby(age_cat ~ as.factor(season) + gender_m_fo + age + age_cat + as.factor(stateresidence) + rurality + race + as.factor(childsgradeinschool) + as.factor(school_format) + 
diabetesduration + insulinmethod + as.factor(pump_brand) + current_cgm_use + as.factor(cgm_type) + 
currenthcltech + as.factor(dka) + as.factor(severe_hypo_episode) + 
parenteducation + as.factor(child_sleep_condition) + as.factor(child_nightime_awakening) + nightime_awakening_2 + parent_sleep_concerns + concern_sleep_disorder + intvent_benefit + hba1c + as.factor(mstage2) + 
as.factor(fstage2) + as.factor(drstage) + as.factor(self_stage), data = demog_df)
```

```{r , results='asis'}
summary(t2)
```

# Model Output - Continuous sleep duration

## Sleep Measures

### Sleep latency

```{r slm}
fit_mod('onsetlatency_avg', analysis)
```

### Sleep efficiency

```{r sem}
fit_mod('sleepefficency_avg', analysis)
```

### waso

```{r wasom}
fit_mod('waso_avg', analysis)
```

## Diabetes measures 

### Time in range (70-180)

```{r tirmm}
fit_mod('sensor_70_180', analysis)
```

### Time hypoglycemic (<70 – combine <54 and 55-69)

```{r tbrmm}
fit_mod('sensor_hypoglycemic', analysis)
```

### Time hyperglycemic (>180 – combine 181-250 and >250)

```{r tarmm}
fit_mod('sensor_hyperglycemic', analysis)
```

### Sensor Glucose Mean

```{r msgs}
fit_mod("sensor_mean", analysis)
```

## Survey measures 

### Morningness/Eveningness Questionnaire 

```{r mescn}
fit_mod("mesc_total", analysis)
```

### SRDB Scale

```{r srdbn}
fit_mod("SRDB_score", analysis)
```

### Hypoglycemia Fear Survey child/Teen

```{r hfsm}
fit_mod("hfs_total", analysis)
```

### Child Life Satisfaction

```{r }
fit_mod("c_life_satisfaction_tscore", analysis)
``` 

### Child Physical Activity

```{r }
fit_mod("c_physical_tscore", analysis)
``` 

### Parent Physical Activity

```{r }
fit_mod("p_physical_tscore", analysis)
``` 

### Child Positve Affect

```{r }
fit_mod("c_positive_affect_tscore", analysis)
``` 

### Parent Positive Affect

```{r }
fit_mod("p_positive_affect_tscore", analysis)
``` 

### Child Psychological Stress

```{r }
fit_mod("c_psychological_stress_tscore", analysis)
``` 

### Parent Psychological Stress

```{r }
fit_mod("p_psychological_stress_tscore", analysis)
``` 

### Child Anger

```{r }
fit_mod("c_anger_tscore", analysis)
``` 

### Parent Anger

```{r }
fit_mod("p_anger_tscore", analysis)
``` 

### Child Anxiety

```{r }
fit_mod("c_anxiety_tscore", analysis)
``` 

### Parent Anxiety

```{r }
fit_mod("p_anxiety_tscore", analysis)
``` 

### Child Depression

```{r }
fit_mod("c_depressive_tscore", analysis)
``` 

### Parent Depression

```{r}
fit_mod("p_depressive_tscore", analysis)
``` 

### Child Family Relationships

```{r}
fit_mod("c_family_tscore", analysis)
``` 

### Parent Family Relationships

```{r}
fit_mod("promis_parent_proxy_sf_v10_fam_relationships_8a_tscore", analysis)
``` 

### Child Peer Relationships

```{r}
fit_mod("c_peer_tscore", analysis)
``` 

### Parent Peer Relationships

```{r}
#fit_mod("p_peer_tscore", analysis)
``` 

### Child Sleep Disturbance

```{r}
fit_mod("c_sleep_disturbance_tscore", analysis)
``` 

### Child Sleep Impairment

```{r}
fit_mod("c_sleep_impairment_tscore", analysis)
``` 

#######

# Model Output - Sufficient sleep

## Sleep Measures

### Sleep latency

```{r }
fit_mod_log('onsetlatency_avg', analysis)
```

### Sleep efficiency

```{r }
fit_mod_log('sleepefficency_avg', analysis)
```

### waso

```{r }
fit_mod_log('waso_avg', analysis)
```

## Diabetes measures 

### Time in range (70-180)

```{r }
fit_mod_log('sensor_70_180', analysis)
```

### Time hypoglycemic (<70 – combine <54 and 55-69)

```{r }
fit_mod_log('sensor_hypoglycemic', analysis)
```

### Time hyperglycemic (>180 – combine 181-250 and >250)

```{r }
fit_mod_log('sensor_hyperglycemic', analysis)
```

### Sensor Glucose Mean

```{r }
fit_mod_log("sensor_mean", analysis)
```

## Survey measures 

### Morningness/Eveningness Questionnaire 

```{r }
fit_mod_log("mesc_total", analysis)
```

### SRDB Scale

```{r }
fit_mod_log("SRDB_score", analysis)
```

### Hypoglycemia Fear Survey child/Teen

```{r }
fit_mod_log("hfs_total", analysis)
```

### Child Life Satisfaction

```{r }
fit_mod_log("c_life_satisfaction_tscore", analysis)
``` 

### Child Physical Activity

```{r }
fit_mod_log("c_physical_tscore", analysis)
``` 

### Parent Physical Activity

```{r }
fit_mod_log("p_physical_tscore", analysis)
``` 

### Child Positve Affect

```{r }
fit_mod_log("c_positive_affect_tscore", analysis)
``` 

### Parent Positive Affect

```{r }
fit_mod_log("p_positive_affect_tscore", analysis)
``` 

### Child Psychological Stress

```{r }
fit_mod_log("c_psychological_stress_tscore", analysis)
``` 

### Parent Psychological Stress

```{r }
fit_mod_log("p_psychological_stress_tscore", analysis)
``` 

### Child Anger

```{r }
fit_mod_log("c_anger_tscore", analysis)
``` 

### Parent Anger

```{r }
fit_mod_log("p_anger_tscore", analysis)
``` 

### Child Anxiety

```{r }
fit_mod_log("c_anxiety_tscore", analysis)
``` 

### Parent Anxiety

```{r }
fit_mod_log("p_anxiety_tscore", analysis)
``` 

### Child Depression

```{r }
fit_mod_log("c_depressive_tscore", analysis)
``` 

### Parent Depression

```{r}
fit_mod_log("p_depressive_tscore", analysis)
``` 

### Child Family Relationships

```{r}
fit_mod_log("c_family_tscore", analysis)
``` 

### Parent Family Relationships

```{r}
fit_mod_log("promis_parent_proxy_sf_v10_fam_relationships_8a_tscore", analysis)
``` 

### Child Peer Relationships

```{r}
fit_mod_log("c_peer_tscore", analysis)
``` 

### Parent Peer Relationships

```{r}
#fit_mod_log("p_peer_tscore", analysis)
``` 

### Child Sleep Disturbance

```{r}
fit_mod_log("c_sleep_disturbance_tscore", analysis)
``` 

### Child Sleep Impairment

```{r}
fit_mod_log("c_sleep_impairment_tscore", analysis)
``` 

# Model Output - Gender

## Sleep Measures

### Sleep latency

```{r }
fit_mod_gender('onsetlatency_avg', analysis)
```

### Sleep efficiency

```{r }
fit_mod_gender('sleepefficency_avg', analysis)
```

### waso

```{r }
fit_mod_gender('waso_avg', analysis)
```

## Diabetes measures 

### Time in range (70-180)

```{r }
fit_mod_gender('sensor_70_180', analysis)
```

### Time hypoglycemic (<70 – combine <54 and 55-69)

```{r }
fit_mod_gender('sensor_hypoglycemic', analysis)
```

### Time hyperglycemic (>180 – combine 181-250 and >250)

```{r }
fit_mod_gender('sensor_hyperglycemic', analysis)
```

### Sensor Glucose Mean

```{r }
fit_mod_gender("sensor_mean", analysis)
```

## Survey measures 

### Morningness/Eveningness Questionnaire 

```{r }
fit_mod_gender("mesc_total", analysis)
```

### SRDB Scale

```{r }
fit_mod_gender("SRDB_score", analysis)
```

### Hypoglycemia Fear Survey child/Teen

```{r }
fit_mod_gender("hfs_total", analysis)
```

### Child Life Satisfaction

```{r }
fit_mod_gender("c_life_satisfaction_tscore", analysis)
``` 

### Child Physical Activity

```{r }
fit_mod_gender("c_physical_tscore", analysis)
``` 

### Parent Physical Activity

```{r }
fit_mod_gender("p_physical_tscore", analysis)
``` 

### Child Positve Affect

```{r }
fit_mod_gender("c_positive_affect_tscore", analysis)
``` 

### Parent Positive Affect

```{r }
fit_mod_gender("p_positive_affect_tscore", analysis)
``` 

### Child Psychological Stress

```{r }
fit_mod_gender("c_psychological_stress_tscore", analysis)
``` 

### Parent Psychological Stress

```{r }
fit_mod_gender("p_psychological_stress_tscore", analysis)
``` 

### Child Anger

```{r }
fit_mod_gender("c_anger_tscore", analysis)
``` 

### Parent Anger

```{r }
fit_mod_gender("p_anger_tscore", analysis)
``` 

### Child Anxiety

```{r }
fit_mod_gender("c_anxiety_tscore", analysis)
``` 

### Parent Anxiety

```{r }
fit_mod_gender("p_anxiety_tscore", analysis)
``` 

### Child Depression

```{r }
fit_mod_gender("c_depressive_tscore", analysis)
``` 

### Parent Depression

```{r}
fit_mod_gender("p_depressive_tscore", analysis)
``` 

### Child Family Relationships

```{r}
fit_mod_gender("c_family_tscore", analysis)
``` 

### Parent Family Relationships

```{r}
fit_mod_gender("promis_parent_proxy_sf_v10_fam_relationships_8a_tscore", analysis)
``` 

### Child Peer Relationships

```{r}
fit_mod_gender("c_peer_tscore", analysis)
``` 

### Parent Peer Relationships

```{r}
#fit_mod_gender("p_peer_tscore", analysis)
``` 

### Child Sleep Disturbance

```{r}
fit_mod_gender("c_sleep_disturbance_tscore", analysis)
``` 

### Child Sleep Impairment

```{r}
fit_mod_gender("c_sleep_impairment_tscore", analysis)
``` 
