---
title: "Cobry: JDRF Phase 1 Manuscript"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
---

```{r setup, include=FALSE}
library(redcapAPI)
library(tidyverse)
library(emmeans)
library(performance)
library(knitr)
library(table1)
library(readr)
library(readxl)
library(lme4)

# data upload
sd_df = read_csv("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/SleepDisruptorsInAdo_DATA_2024-10-30_0826.csv")

# promis tables
Depression_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Depression T-tables.xlsx", 
    sheet = "Pedaitric Depressive Symptoms 8")
Depression_T$raw = Depression_T$`Raw Score`
Depression_T$depression_t = Depression_T$`T-score`
# LP added: parent version of depression
pp_dep_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Depression T-tables.xlsx", 
                          sheet = "Parent Proxy Depressive Sx 6a")

Anxiety_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Anxiety T-tables.xlsx", 
    sheet = "Pediatric Anxiety 8a")
Anxiety_T$raw = Anxiety_T$`Raw Summed Score`
Anxiety_T$anxiety_t = Anxiety_T$`T-Score`
# LP added: parent version of anxiety
pp_anx_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Anxiety T-tables.xlsx", 
                          sheet = "Parent Proxy Anxiety 8a")

Psychological_Stress_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Psychological Stress Experience T-tables.xlsx", 
    sheet = "Pediatric Psych Stress Exp 4a")
Psychological_Stress_T$raw= Psychological_Stress_T$`Raw Score`
Psychological_Stress_T$psych_stress_t = Psychological_Stress_T$`T-score`
# LP added: parent version of psychological stress
pp_pse_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Psychological Stress Experience T-tables.xlsx", 
                             sheet = "Parent Proxy Psy Stress Exp 4a")

Sleep_Disturbance_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Sleep Disturbance and Sleep Related Impairment T Tables.xlsx", 
    sheet = "Pediatric Sleep Disturbance")
Sleep_Disturbance_T$raw = Sleep_Disturbance_T$`Raw Summed Score`
Sleep_Disturbance_T$sleep_disturbance_t = Sleep_Disturbance_T$`T-Score`
# do not see parent version of sleep disturbance

Impairment_T <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/JDRF Career Development Grant/Phase 1 Manuscript/Data_raw/Sleep Disturbance and Sleep Related Impairment T Tables.xlsx", 
    sheet = "Pediatric Sleep-Related Impairm")
Impairment_T$raw = Impairment_T$`Raw Summed Score`
Impairment_T$sleep_impairment_t = Impairment_T$`T-Score`
# do not see parent version of sleep impairment

# LP added
# do not see parent version of life satisfaction
peds_lifesatisfaction_short <- read_excel('/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/Survey scoring/PROMIS_scoring_and_T-tables_-_Non_sleep_measures/Life Satisfaction T-tables.xlsx')
# child and parent version of physical activity
peds_physact_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Physical Activity T-tables.xlsx", 
                                          sheet = "Pediatric Physical Activity 4a")
pp_physact_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Physical Activity T-tables.xlsx", 
                                 sheet = "Parent Proxy Phys Activity 4a")
# child and parent version of positive affect
peds_posaff_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Positive Affect T-tables.xlsx", 
                                 sheet = "Pediatric Positive Affect 4a")
pp_posaff_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Positive Affect T-tables.xlsx", 
                              sheet = "Parent Proxy Positive Affect 4a")
# child and parent version of anger
peds_anger_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Anger T-tables.xlsx", 
                            sheet = "Pediatric Self Report 5a")
pp_anger_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Anger T-tables.xlsx", 
                         sheet = "Parent Proxy 5a")
# child and family versions of family relationships
peds_fr_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Relationships T-tables.xlsx", 
                            sheet = "Pediatric Family Relation 8a")
pp_fr_short = read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Relationships T-tables.xlsx", 
                          sheet = "Parent Prox Family Relat 8a")
# child version of peer relationships - do not see parent
peds_pr_short <- read_excel("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Relationships T-tables.xlsx", 
                             sheet = "Pediatric Peer Relationships 8a")
```


```{r data clean, include = FALSE}
sd_df = sd_df %>% mutate(gender_m_fo = case_when(parent_gender == 1 ~ "Male",
                              parent_gender == 2 | parent_gender == 3 ~"Female/Other"
                              ),
                         age_cat = case_when(age < 12 & age >= 6 ~ "6-12",
                                             age >= 12 ~ "13-18"),
                         race = case_when(race___1 == 1 ~ "White, Non-Hispanic",
                                          race___2 == 1 ~ "Black/African American",
                                          race___3 == 1 ~ "Hispanic/Latino",
                                          race___4 == 1 ~ "Native Hawaiian/Pacific Islander",
                                          race___5 == 1 ~ "Asian",
                                          race___6 == 1 ~ "American Indian/Alaskan Native",
                                          race___8 == 1 ~ "Don't Wish to Answer",
                                          TRUE ~ "Unknown"
                                          ))
sd_df$insulinmethod = factor(sd_df$insulinmethod, levels = c("1","2"), labels = c("Injections", "Insulin Pump"))
sd_df$currenthcltech = factor(sd_df$currenthcltech, levels = c("1","0"), labels = c("Yes", "No"))
sd_df$parenteducation = factor(sd_df$parenteducation, levels = c("2","3","4","5","6","7","8"),
                               labels = c("Less than Highschool", "GED", "Highschool Graduate",
                                          "Trade/Technical School", "Some College", "College Graduate", "Graduate School"))
sd_df$current_cgm_use= factor(sd_df$current_cgm_use, levels = c("1","2"), labels = c("Yes", "No"))
sd_df$rurality= factor(sd_df$rurality, levels = c("1","2"), labels = c("Rural", "Urban"))

# demog covariates
demog_df = sd_df %>% select(record_id, redcap_event_name, season, gender_m_fo, age, age_cat, stateresidence, rurality, race, 
                            childsgradeinschool, school_format, diabetesduration, insulinmethod, pump_brand, current_cgm_use,
                            cgm_type, currenthcltech, dka, recent_dka, severe_hypo_episode, recent_severe_hypo,
                            parenteducation, child_sleep_condition, child_nightime_awakening, nightime_awakening_2,
                            parent_sleep_concerns, concern_sleep_disorder, intvent_benefit) %>% 
  filter(redcap_event_name == "enrollment_arm_1")
a1c_col = sd_df %>% filter(redcap_event_name == "surveys_arm_1") %>% select(record_id, hba1c) 
demog_df = left_join(demog_df, a1c_col, by = "record_id") %>% select(-redcap_event_name) %>% filter(!is.na(gender_m_fo)) ; rm(a1c_col)
tanner_df <- sd_df %>% select(record_id, redcap_event_name, mstage2, fstage2, drstage) %>% 
  filter(redcap_event_name == "enrollment_arm_1")
tanner_df <- tanner_df %>% rowwise() %>% mutate(self_stage = max(mstage2, fstage2, na.rm = T))
demog_df <- left_join(demog_df, tanner_df, by = "record_id")

analysis = sd_df %>% select(record_id,redcap_event_name,
                            time_am, time_mm, sensor_wear, eve_sensor_low, eve_sensor_below,
                            eve_sensor_target, eve_sensor_above, eve_sensor_high, eve_sensor_mean, eve_sensor_sd,
                            night_sensor_low, night_sensor_below, night_sensor_target, night_sensor_above,
                            night_sensor_high, night_sensor_mean, night_sensor_sd, sensor_g251,
                            sensor_181_250, sensor_70_180, sensor_55_69, sensor_u54, sensor_mean, sensor_sd,
                            gmi, tdd, tdd_basal, tdd_bolus, dailybolus, dailymealbolus, sleepmode, exercisemode,
                            avg_bedtime, getup_avg, timeinbed_avg, totalsleep_avg, onsetlatency_avg, sleepefficency_avg, 
                            waso_avg, awake_avg  
) %>% filter(redcap_event_name == "surveys_arm_1") %>% select(-redcap_event_name)

analysis = left_join(demog_df, analysis) 
analysis = analysis %>% mutate(sensor_hypoglycemic = sensor_u54 + sensor_55_69,
                               sensor_hyperglycemic = sensor_181_250 + sensor_g251)

surveys <- sd_df %>% filter(redcap_event_name %in% c("surveys_arm_1"))

# PROMIS Life satisfaction
ls_df <- surveys %>% select(record_id, redcap_event_name, ls_satisfied, ls_happy, 
                            ls_good, ls_wanted)
# Score 4-item version
results_4item <- score_promis_satisfaction(
   data = ls_df,
   item1 = "ls_satisfied",
   item2 = "ls_happy",
   item3 = "ls_good",
   item4 = "ls_wanted", t_scores_file = '/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/Survey scoring/PROMIS_scoring_and_T-tables_-_Non_sleep_measures/Life Satisfaction T-tables.xlsx'
 )
results_4item$ls_t_score <- results_4item$t_score
results_4item <- results_4item %>% select(record_id, ls_t_score)
analysis <- left_join(analysis, results_4item, by = "record_id")

# PROMIS Physical activity - child
results_pa_child <- score_promis_activity_child(
   data = ls_df,
  item1,  # "I could do sports and exercise that other kids my age could do"
                                        item2,  # "I had trouble doing sports or exercise"
                                        item3,  # "I could do physical activities like running"
                                        item4,  # "I could do physical activities like riding a bike"
                                        item5,  # "I could keep up when I played with other kids"
                                        item6,  # "I could move my body quickly"
                                        item7,  # "I could run a mile without stopping"
                                        item8,  # "I could do physical activities or exercise without getting tired quickly"
  t_scores_file = '/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ADA 2022/COVID/Data clean/Physical Activity T-tables.xlsx'
 )

# PROMIS Physical activity - parent

# PROMIS Positve affect - child

# PROMIS Positive affect - parent

# PROMIS psychological stress - child

# PROMIS psychological stress - parent

# PROMIS anger - child

# PROMIS anger - parent

# PROMIS anxiety - child

# PROMIS anxiety - parent

# PROMIS depressie sx - child

# PROMIS depressive sc - parent

# PROMIS family relationships - Child

# PROMIS family relationships - parent

# PROMIS peer relationships - child

# PROMIS peer relationships - parent

# PROMIS sleep disurbance - child

# PROMIS sleep disturbance - parent

# PROMIS sleep related impairment - child

# PROMIS sleep related impairment - parent

# Morningness/eveningness - child

# Morningness/eveningness - parent

# Hypoglycemia fear - child

# Hypoglyemia fear - parent

# SRDB





# old survey scoring below


# Also look at some of the survey measures – you should have scoring t-tables from Erin to do so
# survey scoring

survey_df = sd_df %>% mutate(mesc_total = (6 - me_oob) + me_getup + (5 - me_gym) + (5 - me_testtime)
                             + (5 - me_besttime) + (6 - me_p_bedtime) + me_malert + (6-me_b_bedtime)
                             + me_6am + (5 - me_time2alert),# Morningness/Eveningness Questionnaire
                             across(snore_ht:interrupts, ~ .x - 1)) 

mesc = survey_df %>% select(record_id, mesc_total) %>% filter(!is.na(mesc_total))
  
srdb = survey_df %>% filter(redcap_event_name == "surveys_arm_2") %>% mutate(SRDB_score = rowMeans(across(snore_ht:interrupts), na.rm = T)) %>%
  select(record_id, SRDB_score) # SRDB Scale

srdb$record_id = sub("^.", "1", srdb$record_id)

hfs = survey_df %>% filter(redcap_event_name == "surveys_arm_1") %>% mutate(across(bedtime_snacks:having_low_sugar, ~ .x - 1),
                           hfs_total = rowSums(across(bedtime_snacks:having_low_sugar), na.rm=F)) %>% select(record_id, hfs_total)# Hypoglycemia Fear Survey child/Teen

# Child_PROMIS_Sleep Disturbance
# Child_PROMIS_Sleep Related Impairment 
# Child_PROMIS_Psychological stress
# Child_PROMIS_Anxiety
# Child_PROMIS_Depressive Symptoms 
promis = survey_df %>% filter(redcap_event_name == "surveys_arm_1") %>% mutate(
  sleep_dist_raw = rowSums(across(sd_falling:sd_tossed)),
  sleep_imp_raw =  rowSums(across(sri_sleepy:sri_mood)),
  psy_stress_raw =  rowSums(across(ps_stressed:ps_manage)),
  anxiety_raw =  rowSums(across(a_awful:a_bed)),
  depress_raw =  rowSums(across(ds_sad:ds_hard))) %>% select(record_id, sleep_dist_raw:depress_raw)


promis = left_join(promis, Anxiety_T, by = c('anxiety_raw' = 'raw')) %>% select(record_id, sleep_dist_raw:depress_raw,anxiety_t)

promis = left_join(promis, Depression_T, by = c('depress_raw' = 'raw')) %>% select(record_id, sleep_dist_raw:depress_raw,anxiety_t,depression_t)

promis = left_join(promis, Sleep_Disturbance_T, by = c('sleep_dist_raw' = 'raw')) %>% select(record_id, sleep_dist_raw:depress_raw,anxiety_t,depression_t,sleep_disturbance_t)

promis = left_join(promis, Impairment_T, by = c('sleep_imp_raw' = 'raw')) %>% select(record_id, sleep_dist_raw:depress_raw,anxiety_t,depression_t,sleep_disturbance_t,sleep_impairment_t)

promis = left_join(promis, Psychological_Stress_T, by = c('psy_stress_raw' = 'raw')) %>% select(record_id, sleep_dist_raw:depress_raw,anxiety_t,depression_t,sleep_disturbance_t,sleep_impairment_t,psych_stress_t)

surveys = full_join(promis, hfs) %>% unique()
surveys = full_join(surveys, srdb) %>% unique()
surveys = full_join(surveys, mesc) %>% unique()

analysis = left_join(analysis,surveys) %>% unique()

analysis$hoursc = vapply(strsplit(as.character(analysis$totalsleep_avg),":"), `[`, 1, FUN.VALUE=character(1))
analysis$minsc = vapply(strsplit(as.character(analysis$totalsleep_avg),":"), `[`, 2, FUN.VALUE=character(1))
analysis$sleeptotal_avg = as.numeric(analysis$hoursc)*60 + as.numeric(analysis$minsc)
```



```{r functions, include = F}
fit_mod = function(outcome,df){

  # Fit random intercept model
  f = as.formula(paste0(outcome,"~ gender_m_fo"))
  mod = lm(f,data = df)
  
  f2 = as.formula(paste0(outcome,"~ gender_m_fo + factor(age_cat)"))
  mod2 = lm(f2,data = df)

  print("Base Model")
  cat("\n")
  print(summary(mod))
  cat("\n")
  print("Age Adj Model")
  cat("\n")
  print(summary(mod2))
}

pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
```

# LP data comments

- ID 1032AH-2 removed


# Data remarks
For the age categories of 6-12/13-18 there was a pretty uneven split; I included the model results in case they were of interest, but especially with the categorical outcomes where we essentially stratify age and then gender group theres very few subjects in each category.

# Methods

Demographics stratified by gender group were produced. 

For numeric outcomes t-tests (adjusting for age and not adjusting) were performed. 

For categorical outcomes a chi square test of independence performed.

# Demographics

```{r demog table}
t1 = table1(~ . |gender_m_fo, data = demog_df, overall=F, extra.col=list(`P-value`=pvalue) )
t2 = table1(~ . |age_cat, data = demog_df )
t3 = table1(~ . |age_cat*gender_m_fo, data = demog_df )

t1;t2;t3
```


# Model Output

## Sleep Measures

### Total sleep

Gender was not significantly associated with avg total sleep (min) in either the base or age adjusted models (p = .86, 0.42)

```{r tsm}
fit_mod('sleeptotal_avg', analysis)
```

### Sleep latency

Gender was not significantly associated with sleep latency in either the base or age adjusted models (p = .74, .99)

```{r slm}
fit_mod('onsetlatency_avg', analysis)
```

### Sleep efficiency

Gender was not significantly associated with sleep efficiency in either the base or age adjusted models (p = .65, .70)

```{r sem}
fit_mod('sleepefficency_avg', analysis)
```

### waso

Gender was not significantly associated with WASO avgs in either the base or age adjusted models (p = .32, .57)

```{r wasom}
fit_mod('waso_avg', analysis)
```


## Diabetes measures 

### Time in range (70-180)

Gender was not significantly associated with TIR in either the base or age adjusted models (p = .83, .89)

```{r tirmm}
fit_mod('sensor_70_180', analysis)
```

### Time hypoglycemic (<70 – combine <54 and 55-69)

Gender was not significantly associated with TBR in either the base or age adjusted models (p = .06, .06)

```{r tbrmm}
fit_mod('sensor_hypoglycemic', analysis)
```

### Time hyperglycemic (>180 – combine 181-250 and >250)

Gender was not significantly associated with TBR in either the base or age adjusted models (p = .10, .10)

```{r tarmm}
fit_mod('sensor_hyperglycemic', analysis)
```

### HCL use (for those on HCL, if RedCap value blank do not include) 

Gender was not significantly associated with HCL use p = 1

```{r hclmm}
#table(analysis$gender_m_fo, analysis$currenthcltech)
chisq.test(analysis$gender_m_fo, analysis$currenthcltech)$p.value
```


### Sensor use (for those on sensor, if RedCap value blank do not include) -- log reg?

Gender was not significantly associated with sensor use p = 1

```{r summ}
#table(analysis$gender_m_fo, analysis$current_cgm_use)
chisq.test(analysis$gender_m_fo, analysis$current_cgm_use)$p.value
```

### Sensor Glucose Mean

Gender was not significantly associated with mean sensor glucose (p= 0.6), or in the age adjusted model (p = 0.65)

```{r msgs}
fit_mod("sensor_mean", analysis)
```

## Survey measures 

### Morningness/Eveningness Questionnaire 

Gender is not associated with mesc scores (p = 0.61) in the unadjusted model, and the age adjusted (p = 0.85).

```{r mescn}
fit_mod("mesc_total", analysis)
```

### SRDB Scale

Gender is associated with SRDB scores (p = 0.03) in the unadjusted model, but not the age adjusted (p = 0.07). Male subjects on average had 0.07 (SE = 0.033)lower score than female/other group.

```{r srdbn}
fit_mod("SRDB_score", analysis)
```

### Hypoglycemia Fear Survey child/Teen

Gender was not significantly associated with HFS total scores in both unadjusted and age adjusted models (p = 0.6, 0.5)

```{r hfsm}
fit_mod("hfs_total", analysis)
```

### Child_PROMIS_Sleep Disturbance

Gender was not significantly associated with promis sleep disturbance t scores scores in both unadjusted and age adjusted models (p = 0.18, 0.28)

```{r pdismsd}
fit_mod("sleep_disturbance_t", analysis)
``` 

### Child_PROMIS_Sleep Related Impairment 

Gender is associated with Sleep related impairment promis t scores  (p = 0.03) in the unadjusted model, but not the age adjusted (p = 0.10). Male subjects on average had 4.7 (SE = 2.1)lower score than the female/other group.

```{r promsd}
fit_mod("sleep_impairment_t", analysis)
``` 


### Child_PROMIS_Psychological stress

Gender is associated with psychological stress promis t scores  (p = 0.01) in the unadjusted model, and the age adjusted (p = 0.02). Male subjects on average had 5.1 (SE = 1.8)lower scores than the female/other group in the unadjusted model and 4.8 (se = 1.9) in the age adjusted model.

```{r pssmsd}
fit_mod("psych_stress_t", analysis)
``` 

### Child_PROMIS_Anxiety

Gender is associated with anxiety promis t scores  (p = 0.03) in the unadjusted model, but not the age adjusted (p = 0.07). Male subjects on average had 5.4 (SE = 2.4)lower scores than the female/other group.

```{r pand}
fit_mod("anxiety_t", analysis)
``` 


### Child_PROMIS_Depressive Symptoms 

Gender is associated with depresion promis t scores  (p = 0.005) in the unadjusted model, and the age adjusted (p = 0.01). Male subjects on average had 6.8 (SE = 2.3)lower scores than the female/other group in the unadjusted model and 6.5(se 2.5) in the adjusted.

```{r desd}
fit_mod("depression_t", analysis)
``` 