---
title: "CGM Sampling Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Desktop")
library(arsenal)
library(skimr)
library(knitr)
library(cgmanalysis)
library(tidyverse)
```

```{r import A1c data}
# Import
subjects <- read.csv("./Data_Cleaned/subject_info.csv",na.strings = c("","(-)"))
# Names to lower case and no middle or suffixes for matching
subjects[,c("FirstName","LastName")] <- 
  lapply(subjects[,c("FirstName","LastName")],function(x){
    x <- tolower(x)
    x <- sub(" .*","",x)
  })
subjects$name <- paste0(subjects$FirstName,subjects$LastName)
subjects$name <- gsub(" ","",subjects$name)
```

```{r organize CGM data,eval=FALSE}
# Set output directory
outdir <- "./Data_Cleaned/cgm"
# Iterate through CGM data, format, and name
files <- list.files("./Data_Raw/Patient 90 days",full.names = T)
for (f in files) {
  ext <- tools::file_ext(f)
  if(grepl("xls",ext)){
    dat <- readxl::read_excel(f,col_types = "text")
    dat <- as.data.frame(dat)
  } else if (ext == "csv"){
    dat <- read.csv(f,na.strings = "")
  }
  name <- sub(" *\\ ","",tolower(paste0(dat[1,5],dat[2,5])))
  name <- gsub("[[:digit:]]","",name)
  name <- gsub(" ","",name)
  date <- lubridate::mdy(subjects$MostRecentVisitDate[match(name,subjects$name)])
  if(is.na(date)){stop(paste("No date:",f))}
  # Format data
  calibration <- which(dat[,3]=="Calibration")
  if (length(calibration)>0){dat <- dat[-calibration,]}
  colnames(dat)[grep("glucose",tolower(colnames(dat)))] <- "sensorglucose"
  colnames(dat)[grep("timestamp",tolower(colnames(dat)))] <- "timestamp"
  dat$timestamp <- lubridate::ymd_hms(sub("T"," ",dat$timestamp))
  dat <- dat[-c(1:min(which(!is.na(dat$timestamp)))),]
  dat$subjectid <- name
  dat <- dat[,c("subjectid","timestamp","sensorglucose")]
  dat <- dat[-which(dat$sensorglucose=="Low"),]
  dat$sensorglucose <- suppressWarnings(as.numeric(dat$sensorglucose))
  
  day14 <- dat[dat$timestamp <= date & dat$timestamp >= (date-14),]
  day30 <- dat[dat$timestamp <= date & dat$timestamp >= (date-30),]
  day90 <- dat[dat$timestamp <= date & dat$timestamp >= (date-90),]
  # Write CSV files
  if (nrow(day14 > 0)){write.csv(day14,paste0(outdir,"/day14/",name,"day14.csv"),na = "",row.names = F)}
  if (nrow(day30 > 0)){write.csv(day30,paste0(outdir,"/day30/",name,"day30.csv"),na = "",row.names = F)}
  if (nrow(day90 > 0)){write.csv(day90,paste0(outdir,"/day90/",name,"day90.csv"),na = "",row.names = F)}
}
```

```{r cgmvariables,eval=FALSE}
# CGM variables on each folder
cgmvariables("./Data_Cleaned/cgm/day14","./Data_Cleaned/",
             outputname = "day14variables")
cgmvariables("./Data_Cleaned/cgm/day30","./Data_Cleaned",
             outputname = "day30variables")
cgmvariables("./Data_Cleaned/cgm/day90","./Data_Cleaned",
             outputname = "day90variables")
```

```{r combine data}
vars <- c("percent_time_70_180","percent_time_over_180","percent_time_over_250",
          "percent_time_under_70","percent_time_under_54","cv","average_sensor")
var_names <- c("% TIR 70-180","% Time > 180","% Time > 250",
               "% Time < 70","% Time < 54","CV","Mean Glucose")
# Import cgm variables - limit to those with > 10 days of data
day14 <- read.csv("./Data_Cleaned/day14variables.csv") %>%
  filter(percent_cgm_wear <=100) %>% 
  select(subject_id,percent_cgm_wear,all_of(vars)) %>% 
  left_join(subjects[,c("name","MostRecentA1C")],by = c("subject_id" = "name"))
day30 <- read.csv("./Data_Cleaned/day30variables.csv") %>%
  filter(percent_cgm_wear <=100) %>% 
  select(subject_id,percent_cgm_wear,all_of(vars)) %>% 
  left_join(subjects[,c("name","MostRecentA1C")],by = c("subject_id" = "name"))
day90 <- read.csv("./Data_Cleaned/day90variables.csv") %>%
  filter(percent_cgm_wear <=100) %>% 
  select(subject_id,percent_cgm_wear,all_of(vars)) %>% 
  left_join(subjects[,c("name","MostRecentA1C")],by = c("subject_id" = "name"))
```

```{r analysis}
corr_analysis <- function(df14,df30,df90,cgm_perc = 75){
  # Remove low percentage wear
  df14 <- df14[which(df14$percent_cgm_wear>=cgm_perc),]
  df30 <- df30[which(df30$percent_cgm_wear>=cgm_perc),]
  df90 <- df90[which(df90$percent_cgm_wear>=cgm_perc),]
  # Correlations
  df1 <- cor(df14[,3:ncol(df14)])
  df2 <- cor(df30[,3:ncol(df30)])
  df3 <- cor(df90[,3:ncol(df90)])
  df <- bind_rows(df1[1:7,8],df2[1:7,8],df3[1:7,8])
  df <- df^2
  df$time <- c(14,30,90)
  long <- df %>% pivot_longer(-time)
  p <- ggplot(long,aes(x = time,y = value,color = name)) + geom_line() + geom_point() +
    ylim(0,1) + theme_bw() + 
    xlab("Days in Sampling Period") + ylab("R^2")
  print(p)
  df$n <- c(nrow(df14),nrow(df30),nrow(df90))
  res <- df %>% select(time,n,all_of(vars))
  kable(res,col.names = c("Days of CGM Wear","n",var_names),caption = "R^2 for Correlation with HbA1c",digits = 3)
}
```

# Results

Each participant's CGM data was cleaned and split into time periods of 14, 30, and 90 days priors to most recent HbA1c reading. CGM summary statistics were calculated for each time period using the cgmanalysis R package. Here we report the $R^2$ for the correlation between CGM metrics and HbA1c, given 14, 30, or 90 days of CGM data prior to HbA1c reading. These $R^2$ values are also reported at various levels of percent CGM wear.

## 60% CGM Wear

```{r results 60}
corr_analysis(day14,day30,day90,60)
```

## 65% CGM Wear

```{r results 65}
corr_analysis(day14,day30,day90,65)
```

## 70% CGM Wear

```{r results 70}
corr_analysis(day14,day30,day90,70)
```

## 75% CGM Wear

```{r results 75}
corr_analysis(day14,day30,day90,75)
```

## 80% CGM Wear

```{r results 80}
corr_analysis(day14,day30,day90,80)
```

## 85% CGM Wear

```{r results 85}
corr_analysis(day14,day30,day90,85)
```

## 90% CGM Wear

```{r results 90}
corr_analysis(day14,day30,day90,90)
```

## 95% CGM Wear

```{r results 95}
corr_analysis(day14,day30,day90,95)
```

