---
title: "CGM Sampling Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,dpi = 600)
#knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kaan Akturk/CGM Sampling")
 setwd("S:/Laura/BDC/Projects/Kaan Akturk/CGM Sampling/")
library(arsenal)
library(knitr)
library(cgmanalysis)
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r import A1c data}
# Import
 setwd("S:/Laura/BDC/Projects/Kaan Akturk/CGM Sampling/")
subjects <- read.csv("./Data_Cleaned/subject_info.csv",na.strings = c("","(-)"))
# Names to lower case and no middle or suffixes for matching
subjects[,c("FirstName","LastName")] <- 
  lapply(subjects[,c("FirstName","LastName")],function(x){
    x <- tolower(x)
    x <- sub(" .*","",x)
  })
subjects$name <- paste0(subjects$FirstName,subjects$LastName)
subjects$name <- gsub(" ","",subjects$name)
# Set days and sampling percentages
days <- seq(7,91,by = 7)
sampling <- seq(45,95,by = 5)
```

```{r organize CGM data,eval=FALSE}
# Set output directory
# setwd("S:/Laura/BDC/Projects/Kaan Akturk/CGM Sampling/")
outdir <- "./Data_Cleaned/cgm"
#outdir <- "./Data_Cleaned/test"
# Iterate through CGM data, format, and name
files <- list.files("./Data_Raw/Patient 90 days",full.names = T)
for (f in files) {
  ext <- tools::file_ext(f)
  if(grepl("xls",ext)){
    dat <- readxl::read_excel(f,col_types = "text")
    dat <- as.data.frame(dat)
  } else if (ext == "csv"){
    dat <- read.csv(f,na.strings = "")
  }
  name <- sub(" *\\ ","",tolower(paste0(dat[1,5],dat[2,5])))
  name <- gsub("[[:digit:]]","",name)
  name <- gsub(" ","",name)
  date <- 
    lubridate::mdy(subjects$MostRecentVisitDate[
      match(name,subjects$name)])
  if(is.na(date)){stop(paste("No date:",f))}
  # Format data
  calibration <- which(dat[,3]=="Calibration")
  if (length(calibration)>0){dat <- dat[-calibration,]}
  colnames(dat)[grep("glucose",tolower(colnames(dat)))] <- "sensorglucose"
  colnames(dat)[grep("timestamp",tolower(colnames(dat)))] <- "timestamp"
  dat$timestamp <- lubridate::ymd_hms(sub("T"," ",dat$timestamp))
  dat <- dat[-c(1:min(which(!is.na(dat$timestamp)))),]
  dat$subjectid <- name
  dat <- dat[,c("subjectid","timestamp","sensorglucose")]
  dat <- dat[-which(dat$sensorglucose=="Low"),]
  dat$sensorglucose <- suppressWarnings(as.numeric(dat$sensorglucose))
  
  for(d in days){
    dat_name <- paste0("day",d)
    # Create new dataframe
    dir <- paste0(outdir,"/day",d)
    assign(dat_name,dat[dat$timestamp <= date & 
                          dat$timestamp >= (date-d),])
    # Make new directory
    dir.create(dir,showWarnings = F)
    # Write
    if (nrow(get(dat_name)) > 0){
      write.csv(get(dat_name),paste0(dir,"/",name,dat_name,".csv"),
                na = "",row.names = F)
    }
  }
}
```

```{r cgmvariables,eval=FALSE}
# CGM variables on each folder
for (d in days) {
  cgmvariables(paste0("./Data_Cleaned/cgm","/",paste0("day",d)),
               "./Data_Cleaned/",
               outputname = paste0("day",d,"variables"))
}
```

# Table 1: Descriptive Characteristics

```{r t1,results='asis'}
demographics <- subjects[1:336,]
table1 <- tableby(~ Age + Gender + MostRecentA1C,data = demographics)
summary(table1)
```

```{r combine data}
setwd("S:/Laura/BDC/Projects/Kaan Akturk/CGM Sampling/")

vars <- c("percent_time_70_180","percent_time_over_180",
          "percent_time_over_250","percent_time_under_70",
          "percent_time_under_54","cv","mean_sensor")
var_names <- c("% TIR 70-180","% Time > 180","% Time > 250",
               "% Time < 70","% Time < 54","CV","Mean Glucose")
# Import cgm variables by days from HbA1c
for (d in days) {
  filename <- paste0("day",d,"variables",".csv")
  dat <- read.csv(paste0("./Data_Cleaned/",filename))
  dat <- dat[,c("subject_id","percent_cgm_wear",vars)]
  dat$sample_days <- d
  if(d == days[1]){cgm <- dat} else {cgm <- rbind(cgm,dat)}
}
rm(dat)
# Add Hba1c data
cgm <- left_join(cgm,subjects[,c("name","MostRecentA1C")],by = c("subject_id" = "name"))
# Iterate through CGM sampling percentages
r <- tail(split.data.frame(cgm,cgm$sample_days),1)
r <- do.call(rbind,r)
for (s in sampling){
  # Calculate correlations
  lim <- cgm[cgm$percent_cgm_wear >= s,]
  c <- split.data.frame(lim,lim$sample_days)
  c <- head(c,-1)
  c <- lapply(c, function(x){
    x <- as.data.frame(x)
    d <- x$sample_days[1]
    ids <- intersect(x$subject_id,r$subject_id)
    x <- x[x$subject_id %in% ids,]
    r <- r[r$subject_id %in% ids,]
    x <- x[order(x$subject_id),]
    r <- r[order(r$subject_id),]
    co <- cor(x[,vars],r[,vars])
    n = nrow(x)
    m_a1c <- cor(x$percent_time_70_180,r$MostRecentA1C)
    all <- as.numeric(c(d,s,n,c(diag(co),m_a1c)^2))
    names(all) <- c("sample_days","sample_percentage","n",vars)
    return(all)
  }) 
  c <- as.data.frame(do.call(rbind,c))
  if (s == sampling[1]){plot <- c} else {plot <- rbind(plot,c)}
}
plot = plot[1:length(plot)-1] # remove mean hba1c and tir per kaan
plot$sample_percentage <- paste0(plot$sample_percentage,"% CGM Sampling")
t <- split.data.frame(plot,plot$sample_percentage)
```

## $R^2$ for CGM Metrics With the 3-Month Sampling Period Values

### All metrics faceted by CGM sampling percentage

```{r plot,fig.width=8,fig.height=8}
# Plot
plotvars <- c(vars)
plotvarnames <- c("% 70-180 mg/dL","% >180 mg/dL","% >250 mg/dL",
                  "% <70 mg/dL","% <54 mg/dL","CV","Mean glucose")
plot <- plot %>% pivot_longer(all_of(plotvars))

p <- ggplot(plot,aes(x=sample_days,y=value,color=name)) + geom_line() + 
  facet_wrap("sample_percentage") + 
  theme_classic() + ylab("R^2") + xlab("Days in sampling period") + 
  scale_color_discrete(name="",breaks = plotvars,labels = plotvarnames) +
  scale_x_continuous(breaks = days[seq(1,length(days),2)]) + 
  scale_y_continuous(breaks = seq(0,1,by = 0.1),sec.axis = sec_axis( trans=~sqrt(.), name="R"))
p
```

### All CGM sampling percentages faceted by metric

```{r plot 2,fig.width=8,fig.height=8}
# Format labels
plot$name <- as.factor(plot$name)
levels(plot$name) <- c("Mean glucose","CV",
                       "% 70-180 mg/dL","% >180 mg/dL","% >250 mg/dL",
                       "% <54 mg/dL","% <70 mg/dL")

p2 <- ggplot(plot,aes(x=sample_days,y=value,color=sample_percentage)) + 
  geom_line() + 
  facet_wrap("name") + 
  theme_classic() + ylab("R^2") + xlab("Days in sampling period") +
  scale_color_discrete(name="") +
  scale_x_continuous(breaks = days[seq(1,length(days),2)]) + 
  scale_y_continuous(breaks = seq(0,1,by = 0.1))

p2
```

## $R^2$ Tables

### By Sampling Percentage

```{r tables fine percentages,results='asis'}
invisible(lapply(t, function(x){
  x <- as.data.frame(x)
  title <- x$sample_percentage[1] 
  x$sample_percentage <- NULL
  print(kable(x,col.names = c("Days of Sampling","n",plotvarnames),
              row.names = F,caption = title,digits = 3))
}))
```

### By CGM Metric

```{r tables fine by metric,results='asis'}
t2 <- split.data.frame(plot,plot$name)
invisible(lapply(t2, function(x){
  x <- as.data.frame(x)
  title <- as.character(x$name[1])
  x$n = NULL
  x <- x %>% pivot_wider(names_from = sample_percentage) %>%
    select(-name)
  colnames(x)[1] <- "Days of Sampling"
  print(kable(x,row.names = F,caption = title,digits = 3))
}))
```

# Three CGM sampling groups

```{r three group data}
cgm$percent_wear_group <- 
  cut(cgm$percent_cgm_wear,breaks = c(-Inf,50,70,Inf),
      labels = c("<50%","50-70%",">70%"),right = F)
# Iterate through CGM sampling percentages
r <- tail(split.data.frame(cgm,cgm$sample_days),1)
r <- do.call(rbind,r)
for (s in levels(cgm$percent_wear_group)){
  # Calculate correlations
  lim <- cgm[cgm$percent_wear_group == s,]
  c <- split.data.frame(lim,lim$sample_days)
  c <- head(c,-1)
  c <- lapply(c, function(x){
    x <- as.data.frame(x)
    d <- x$sample_days[1]
    ids <- intersect(x$subject_id,cgm$subject_id)
    x <- x[x$subject_id %in% ids,]
    r2 <- r[r$subject_id %in% ids,]
    x <- x[order(x$subject_id),]
    r2 <- r2[order(r2$subject_id),]
    n <- nrow(x)
    co <- suppressWarnings(cor(x[,vars],r2[,vars]))
    m_a1c <- cor(x$percent_time_70_180,r2$MostRecentA1C)
    all <- c(d,s,n,c(diag(co),m_a1c)^2)
    names(all) <- c("sample_days","percent_wear_group","n",
                    vars)
    return(all)
  }) 
  c <- as.data.frame(do.call(rbind,c))
  if (s == levels(cgm$percent_wear_group)[1]){
    plot1 <- c
  } else {plot1 <- rbind(plot1,c)}
}


plot1[,c(1,3:ncol(plot1))] <- 
  lapply(plot1[,c(1,3:ncol(plot1))],function(x){as.numeric(as.character(x))})

plot1 = plot1[1:length(plot1)-1]
t3 <- split.data.frame(plot1,plot1$percent_wear_group)


```

## $R^2$ for CGM Metrics With the 3-Month Sampling Period Values

### All metrics faceted by CGM sampling percentage

```{r group plot 1,warning=FALSE}
plot1 <- plot1 %>% pivot_longer(all_of(plotvars))

p <- ggplot(plot1[plot1$name == "percent_time_70_180",],
            aes(x=sample_days,y=value)) + geom_line(aes(linetype=percent_wear_group)) +
  theme_classic() + ylab("R^2") + xlab("Days in sampling period") +
  scale_color_discrete(name="CGM Wear") +
  scale_x_continuous(breaks = days[seq(1,length(days),1)]) + 
  scale_y_continuous(sec.axis = sec_axis( trans=~sqrt(.), name="R"),
                     limits = c(0.6,1)) 
p

# 
```

### All CGM sampling percentages faceted by metric

```{r group plot 2,fig.width=8,fig.height=8}
# Format labels
plot1$name <- as.factor(plot1$name)
levels(plot1$name) <- c("Mean glucose","CV",
                        "% 70-180 mg/dL","% >180 mg/dL","% >250 mg/dL",
                        "% <54 mg/dL","% <70 mg/dL")

p2 <- 
  ggplot(plot1,aes(x=sample_days,y=value)) + 
  geom_line(aes(linetype=percent_wear_group)) + 
  facet_wrap("name") + 
  theme_classic() + ylab("R^2") + xlab("Days in sampling period") +
  scale_color_discrete(name="") +
  scale_x_continuous(breaks = days[seq(1,length(days),2)]) + 
  scale_y_continuous(breaks = seq(0,1,by = 0.1), sec.axis = sec_axis( trans=~sqrt(.), name="R"))

p2
```

## $R^2$ Tables

### By Sampling Percentage

```{r 3 group tables,results='asis'}
invisible(lapply(t3, function(x){
  x <- as.data.frame(x)
  title <- x$percent_wear_group[1]
  x$sample_percentage <- NULL
  x$percent_wear_group <- NULL
  print(kable(x,col.names = c("Days of Sampling","n",plotvarnames),
              row.names = F,caption = title,digits = 3))
}))
```

### By CGM Metric

```{r 3 group tables by metric,results='asis'}
t4 <- split.data.frame(plot1,plot1$name)
invisible(lapply(t4, function(x){
  x <- as.data.frame(x)
  title <- as.character(x$name[1])
  x$sample_percentage <- NULL
  x$n <- NULL
  x <- x %>% pivot_wider(names_from = percent_wear_group) %>%
    select(-name)
  colnames(x)[1] <- "Days of Sampling"
  print(kable(x,
              row.names = F,caption = title,digits = 3))
}))
```

# No CGM Sampling Cutoff

```{r}
# Iterate through CGM sampling days
c <- split.data.frame(cgm,cgm$sample_days)
r <- tail(c,1)
r <- do.call(rbind,r)
c <- head(c,-1)
c <- lapply(c, function(x){
  x <- as.data.frame(x)
  d <- x$sample_days[1]
  ids <- intersect(x$subject_id,r$subject_id)
  x <- x[x$subject_id %in% ids,]
  r2 <- r[r$subject_id %in% ids,]
  x <- x[order(x$subject_id),]
  r2 <- r2[order(r2$subject_id),]
  n <- nrow(x)
  co <- suppressWarnings(cor(x[,vars],r2[,vars]))
  m_a1c <- cor(x$percent_time_70_180,r2$MostRecentA1C)
  all <- c(d,n,c(diag(co),m_a1c)^2)
  names(all) <- c("sample_days","n",vars)
  return(all)
}) 
c <- as.data.frame(do.call(rbind,c))
c = c[1:length(c)-1]
# Plot
plot2 = c %>% pivot_longer(all_of(plotvars))
plot2$name <- as.factor(plot2$name)
levels(plot2$name) <- c("Mean glucose","CV",
                        "% 70-180 mg/dL","% >180 mg/dL","% >250 mg/dL",
                        "% <54 mg/dL","% <70 mg/dL")
p3 <- 
  ggplot(plot2,aes(x=sample_days,y=value)) + 
  geom_line(aes( color = name)) +
  theme_classic() + ylab("R^2") + xlab("Days in sampling period") +
  scale_color_discrete(name="") +
  scale_x_continuous(breaks = days[seq(1,length(days),2)]) + 
  scale_y_continuous(breaks = seq(0,1,by = 0.1), sec.axis = sec_axis( trans=~sqrt(.), name="R"))

p3
# in aes: linetype = name,
# scale_linetype_manual(values = c(rep("solid",4), rep("dashed",4))) + scale_color_manual(values=c("black", "blue", "red", "green","black", "blue", "red", "green"))+
```

## $R^2$ Tables

```{r,results='asis'}
kable(c,col.names = c("Days of Sampling","n",plotvarnames),
      row.names = F,digits = 3)
```


