---
title: "HCL and Sleep (ADA 2022)"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(redcapAPI)
library(arsenal)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r api import,include=FALSE}
token = read.table("./Erin Cobry/ADA 2022/HCL and Sleep/api_token.txt")[1,1]
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",token = token)
data = exportRecords(rcon,dates = F)
ids = data$record_id
times = data$redcap_event_name
seasons = read.csv("./Erin Cobry/Summer vs. School Sleep/Data_Clean/summer_vs_school.csv",na.strings = "")
seasons = seasons[complete.cases(seasons),]
adult_visits = seasons
adult_visits$ID = seasons$ID - 100
seasons = rbind(adult_visits,seasons)
seasons$ID = as.character(seasons$ID)
seasons = seasons %>% 
  pivot_longer(cols = c("Summer","School"),names_to = "Period",values_to = "Visit")
seasons$Visit[seasons$ID < 200 & seasons$Visit == "M3"] = "hcl_month_3_day_1_arm_1"
seasons$Visit[seasons$ID < 200 & seasons$Visit == "M6"] = "hcl_month_6_day_1_arm_1"
seasons$Visit[seasons$ID >= 200 & seasons$Visit == "M3"] = "hcl_month_3_day_1_arm_2"
seasons$Visit[seasons$ID >= 200 & seasons$Visit == "M6"] = "hcl_month_6_day_1_arm_2"
seasons = seasons %>% rename(record_id = "ID",timepoint = "Visit")
```

# Table 1a: Parent Characteristics

```{r results='asis'}
# Basic demographics
parent_demographics = exportRecords(rcon,forms = "parent_demographics")
parent_demographics$record_id = as.character(ids)
parent_demographics$age = 
  as.numeric(difftime(parent_demographics$parent_demographics_timestamp,
                      parent_demographics$parent_bday,units = "days"))/365.25
parent_demographics = parent_demographics %>% 
  select(record_id,age,parent_gender,race,parent_check_night,child_check_glucose,
         insulinmethod,previous_cgm_use,currenthcltech) %>% 
  filter(record_id %in% seasons$record_id)
parent_demographics = 
  parent_demographics[rowSums(is.na(parent_demographics)) < (ncol(parent_demographics)-1),]
# Table
f = as.formula(paste("~",paste0(colnames(parent_demographics)[-1],collapse = "+")))
t1a = tableby(f,data = parent_demographics)
summary(t1a,digits = 2,labelTranslations = list(age = "Age (years)",
                                                currenthcltech = "Currently using HCL?"))
```

# Table 1b: Child Characteristics

```{r results='asis'}
# Basic demographics
child_demographics = exportRecords(rcon,forms = "child_demographics_chart_review")
child_demographics$record_id = as.character(ids)
# Get Hba1c
hba1c = exportRecords(rcon,forms = "glycemic_data")
hba1c$record_id = as.character(ids)
hba1c = hba1c %>% filter(record_id >= 200,gyl_timepoint == "Baseline") %>%
  select(record_id,hba1c)
# Format
child_demographics$t1d_duration = 
  as.numeric(difftime(child_demographics$consent_date,
                      child_demographics$t1d_diagnosis,units = "days"))/365.25
child_demographics = child_demographics %>% 
  select(record_id,cons_age,childgender,t1d_duration) %>%
  filter(record_id %in% seasons$record_id)
child_demographics = 
  child_demographics[rowSums(is.na(child_demographics)) < (ncol(child_demographics)-1),]
# Add Hba1c
child_demographics = left_join(child_demographics,hba1c,by = "record_id")
# Table
f = as.formula(paste("~",paste0(colnames(child_demographics)[-1],collapse = "+")))
t1a = tableby(f,data = child_demographics)
summary(t1a,digits = 3,labelTranslations = list(t1d_duration = "T1D Duration (years)"))
```

# Methods

Because some participants were included in both the summer and school year groups, a paired t test was used to compare summertime and school year for participants with both. 

# Actigraphy

Average sleep time was converted to minutes for all analyses.

```{r}
# Get total sleep time, sleep efficiency, and WASO
activity = exportRecords(rcon,forms = "wrist_actigraphy")
activity$record_id = as.character(ids)
activity$timepoint = times
child_activity = activity %>% 
  select(record_id,timepoint,totalsleep_avg,sleepefficency_avg,waso_avg,timeinbed_avg) %>% 
  filter(record_id %in% seasons$record_id,record_id >= 200,
         grepl("month_3|month_6",timepoint)) %>%
  mutate(totalsleep_avg = as.numeric(lubridate::hms(as.character(totalsleep_avg)))/60,
         timeinbed_avg = as.numeric(lubridate::hms(as.character(timeinbed_avg)))/60)
child_activity = left_join(child_activity,child_demographics,by = "record_id")
child_activity = left_join(child_activity,seasons,by = c("record_id", "timepoint"))
child_activity$sufficient = apply(child_activity,1,function(r){
  age = as.numeric(r["cons_age"])
  time = as.numeric(r["totalsleep_avg"])/60
  if (is.na(time) | is.na(age)){
    suff = NA
  } else if(age <= 2){
    if(time >= 11 & time <= 14){suff = "Yes"} else {suff = "No"}
  } else if (age <= 5){
    if(time >= 10 & time <= 13){suff = "Yes"} else {suff = "No"}
  } else if (age <= 12){
    if(time >= 9 & time <= 12){suff = "Yes"} else {suff = "No"}
  } else if (age <= 17){
    if(time >= 8 & time <= 10){suff = "Yes"} else {suff = "No"}
  } else if (age > 17) {
    if(time >= 7){suff = "Yes"} else {suff = "No"}
  }
})
child_activity$sufficient = as.factor(child_activity$sufficient)
# Same for parents
parent_activity = activity %>% 
  select(record_id,timepoint,totalsleep_avg,sleepefficency_avg,waso_avg,timeinbed_avg) %>% 
  filter(record_id %in% seasons$record_id,record_id < 200,
         grepl("month_3|month_6",timepoint)) %>%
  mutate(avg_sleep = as.character(totalsleep_avg),
         totalsleep_avg = as.numeric(lubridate::hms(totalsleep_avg))/60,
         timeinbed_avg = as.numeric(lubridate::hms(timeinbed_avg))/60,
         sufficient = ifelse(totalsleep_avg >= 7*60,"Yes","No"))
parent_activity$sufficient = as.factor(parent_activity$sufficient)
parent_activity = left_join(parent_activity,seasons,by = c("record_id", "timepoint"))
```

## Children

```{r results='asis'}
p <- paired(Period ~ totalsleep_avg + sleepefficency_avg + waso_avg + timeinbed_avg + sufficient, 
            data = child_activity,id = record_id, signed.rank.exact = FALSE)
summary(p,labelTranslations = list(totalsleep_avg = "Average Total Sleep Time (minutes)",
                                   timeinbed_avg = "Average Time  in Bed (minutes)",
                                   sufficient = "Sufficient sleep?"))
```

## Parents

```{r results='asis'}
p <- paired(Period ~ totalsleep_avg + sleepefficency_avg + waso_avg + timeinbed_avg + sufficient, 
            data = parent_activity,id = record_id, signed.rank.exact = FALSE)
summary(p,labelTranslations = list(totalsleep_avg = "Average Total Sleep Time (minutes)",
                                   timeinbed_avg = "Average Time  in Bed (minutes)",
                                   sufficient = "Sufficient sleep?"))
```

# Glycemic outcomes

GMI is a linear transformation of mean glucose, so was not included in this report. 

```{r}
# Get total sleep time, sleep efficiency, and WASO
glycemic = exportRecords(rcon,forms = "glycemic_data")
glycemic$record_id = as.character(ids)
glycemic$timepoint = times
glycemic = glycemic %>% 
  mutate(sensor_u70 = sensor_55_69 + sensor_u54,
         sensor_g180 = sensor_181_250 + sensor_g251) %>% 
  select(record_id,timepoint,sensor_u70,sensor_70_180,sensor_g180,sensor_mean) %>% 
  filter(record_id %in% seasons$record_id,record_id >= 200,
         grepl("month_3|month_6",timepoint))
glycemic = left_join(glycemic,seasons,by = c("record_id", "timepoint"))
```

```{r results='asis'}
p <- paired(Period ~ sensor_u70+sensor_70_180+sensor_g180+sensor_mean, 
            data = glycemic,id = record_id, signed.rank.exact = FALSE)
summary(p,labelTranslations = list(sensor_u70 = "% time sensor values < 70 mg/dL",
                                   sensor_g180 = "% time sensor values > 180 mg/dL"))
```

# Surveys

## PSQI 

```{r}
psqi = exportRecords(rcon,forms = "psqi",dates = F)
psqi$record_id = as.character(ids)
psqi$timepoint = times
psqi = psqi %>% select(record_id,timepoint,psqi_total) %>% 
  filter(record_id %in% seasons$record_id,record_id < 200,
         grepl("month_3|month_6",timepoint))
psqi = left_join(psqi,seasons,by = c("record_id", "timepoint"))
```

```{r results='asis'}
p <- paired(Period ~ psqi_total,data = psqi,id = record_id, signed.rank.exact = FALSE)
summary(p)
```

## PROMIS: Sleep Disturbance

### Children

```{r}
# Import select, filter, etc.
promis_peds = exportRecords(rcon,forms = "promis_pediatric_817")
promis_peds$record_id = as.character(ids)
promis_peds$timepoint = times
promis_peds = promis_peds %>% select(record_id,timepoint,falling_asleep:bad_mood_sleepy) %>%
  filter(record_id %in% seasons$record_id,grepl("enrollment|month_3|month_6",timepoint),
         record_id >= 200) %>%
  filter(record_id %in% seasons$record_id,grepl("month_3|month_6",timepoint),record_id >= 200) %>%
  mutate(across(falling_asleep:bad_mood_sleepy,.fns = as.numeric),
         through_night = 6 - through_night) %>% 
  mutate(disturbance = rowSums(select(.,falling_asleep:tossed_and_turned)),
         impairment = rowSums(select(.,daytime:bad_mood_sleepy)))
# Convert to T scores
peds_disturbance = read.csv("./Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/peds_sleep_disturbance.csv")
peds_impairment = read.csv("./Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/peds_sleep_impairment.csv")
promis_peds$disturbance = 
  peds_disturbance$T.Score[match(promis_peds$disturbance,
                                 peds_disturbance$Raw.Summed.Score)]
promis_peds$impairment = 
  peds_impairment$T.Score[match(promis_peds$impairment,
                                peds_impairment$Raw.Summed.Score)]
promis_peds = left_join(promis_peds,seasons,by = c("record_id", "timepoint"))
```

```{r results='asis'}
p <- paired(Period ~ disturbance + impairment,
            data = promis_peds,id = record_id,signed.rank.exact = FALSE)
p <- paired(Period ~ disturbance,data = promis_peds,id = record_id, signed.rank.exact = FALSE)
summary(p)
```

### Parents

```{r}
# Import select, filter, etc.
promis_parent = exportRecords(rcon,forms = "promis_parent_proxy_817")
promis_parent$record_id = as.character(ids)
promis_parent$timepoint = times
promis_parent = promis_parent %>% select(record_id,timepoint,asleep:bad_mood) %>%
  filter(record_id %in% seasons$record_id,grepl("enrollment|month_3|month_6",timepoint),
         record_id < 200) %>%
  mutate(across(asleep:bad_mood,.fns = as.numeric),
         night = 6 - night) %>% 
  mutate(disturbance = rowSums(select(.,asleep:tossed)),
         impairment = rowSums(select(.,sleepy:bad_mood)))
promis_parent$timepoint = sub("enrollment_","baseline_",promis_parent$timepoint)
# Convert to T scores
parent_disturbance = read.csv("./Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/parent_proxy_sleep_disturbance.csv")
parent_impairment = read.csv("./Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/parent_proxy_sleep_impairment.csv")
promis_parent$disturbance = 
  parent_disturbance$T.Score[match(promis_parent$disturbance,
                                   parent_disturbance$Raw.Summed.Score)]
promis_parent$impairment = 
  parent_impairment$T.Score[match(promis_parent$impairment,
                                  parent_impairment$Raw.Summed.Score)]
promis_parent = left_join(promis_parent,seasons,by = c("record_id", "timepoint"))
```

```{r results='asis'}
p <- paired(Period ~ disturbance + impairment,
            data = promis_parent,id = record_id,signed.rank.exact = FALSE)
summary(p)
```

# Awakenings

Awakenings were analyzed using a generalized linear mixed effects model for the negative binomial family, with a random effect for participant. The "Summer" term indicates the difference between school and summer, with school time as the reference.

```{r}
# Data cleaning
data$record_id = as.character(data$record_id)
data$timepoint = times
child_awakenings = data %>% 
  select(record_id,timepoint,diabetes_awakenings,all_awakenings) %>% 
  filter(record_id %in% seasons$record_id,grepl("month_3|month_6",timepoint),
         record_id >= 200)
child_awakenings = left_join(child_awakenings,seasons,by = c("record_id", "timepoint"))
# Parents
parent_awakenings = data %>% 
  select(record_id,timepoint,diabetes_awakenings,all_awakenings) %>% 
  filter(record_id %in% seasons$record_id,grepl("month_3|month_6",timepoint),
         record_id < 200)
parent_awakenings = left_join(parent_awakenings,seasons,by = c("record_id", "timepoint"))
```

### Children

```{r results='asis'}
mod = glmer.nb(diabetes_awakenings ~ Period + (1|record_id),data = child_awakenings)
res = tidy(mod,"fixed")
res$term = sub("Period","",res$term)
kable(res[,2:ncol(res)],digits = 3)
```

### Parents

```{r results='asis'}
mod = glmer.nb(diabetes_awakenings ~ Period + (1|record_id),data = parent_awakenings)
res = tidy(mod,"fixed")
res$term = sub("Period","",res$term)
kable(res[,2:ncol(res)],digits = 3)
```

## All Awakenings

### Children

```{r results='asis'}
mod = glmer.nb(all_awakenings ~ Period + (1|record_id),data = child_awakenings)
res = tidy(mod,"fixed")
res$term = sub("Period","",res$term)
kable(res[,2:ncol(res)],digits = 3)
```

### Parents

```{r results='asis'}
mod = glmer.nb(all_awakenings ~ Period + (1|record_id),data = parent_awakenings)
res = tidy(mod,"fixed")
res$term = sub("Period","",res$term)
kable(res[,2:ncol(res)],digits = 3)
```
