---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readxl)
library(redcapAPI)
library(nlme)
library(lme4)
library(emmeans)
library(broom.mixed)
library(lubridate)
library(table1)

# data and subj list upload
HCL_Study_Subjects = read_excel("S:/Laura/BDC/Projects/Erin Cobry/Summer vs. School Sleep/Data_Clean/HCL_Study_Subjects1.xlsx") %>% filter(!is.na(DOB))
include = HCL_Study_Subjects$Subject
# hcl and sleep dataset
token = read.table("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/api_token.txt")[1,1]
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",token = token)
data = exportRecords(rcon,dates = F)
ids = data$record_id
times = data$redcap_event_name
exclude = c(101,201,102,202,103,203,110,210,119,219,121,221,131,231,141,241)
```





```{r data clean, include=FALSE}

#no m6:200,202,205,206,207,208,243,244,245,246 --- for paired stuff

# exclude everyone in exclude from data 
# kids only
data = data %>% filter((record_id %in% include),record_id > 199) %>% 
  select(record_id, redcap_event_name,all_awakenings, diabetes_awakenings,totalsleep_avg, sleepefficency_avg,
         waso_avg, timeinbed_avg, sensor_mean, sensor_70_180,  sensor_g251, sensor_181_250,
         sensor_55_69, sensor_u54) %>%
  mutate(timepoint_cs = case_when(redcap_event_name == "baseline_day_1_arm_2" ~ "Baseline (before HCL)",
                                  redcap_event_name == "hcl_month_3_day_1_arm_2" ~ "3 Months",
                                  redcap_event_name == "hcl_month_6_day_1_arm_2" ~ "6 Months")) %>%
  filter(!is.na(timepoint_cs))

# get the excel sheet in a position i can work with
baseline = HCL_Study_Subjects %>%  mutate(visit = "Baseline (before HCL)",visit_date = Baseline, visit_type = `School/Summer...5`) %>% select(Subject, DOB, Age, visit, visit_type, visit_date) 
m3 = HCL_Study_Subjects %>%  mutate(visit = "3 Months",visit_date = M3, visit_type = `School/Summer...7`) %>% select(Subject, DOB, Age, visit, visit_type, visit_date) 
m6 = HCL_Study_Subjects %>%  mutate(visit = "6 Months",visit_date = M6, visit_type = `School/Summer...9`) %>% select(Subject, DOB, Age, visit, visit_type, visit_date) 

# join
study_subjects = full_join(baseline, m3)
study_subjects = full_join(study_subjects, m6) 
study_subjects$Subject = as.character(study_subjects$Subject)

data$record_id = as.character(data$record_id)
data$timepoint_cs = as.character(data$timepoint_cs)

analysis = left_join(data, study_subjects, 
                     by = c("record_id" = "Subject",
                            "timepoint_cs" = "visit"
                            ))

analysis$totalsleep_avg = hm(analysis$totalsleep_avg)
analysis$timeinbed_avg = hm(analysis$timeinbed_avg)
analysis = analysis %>%  mutate(totalsleep_avg_min = hour(totalsleep_avg)* 60 + minute(totalsleep_avg),
                                                                         timeinbed_avg_min = hour(timeinbed_avg)* 60 + minute(timeinbed_avg),
                                                                         sensor_u70 = sensor_55_69 + sensor_u54,
                                                                         sensor_g180 = sensor_181_250 + sensor_g251
                                                                         )

analysis = analysis %>% filter(!is.na(visit_type))
# number of visits per subject; need to try and get to 2?
#table(analysis$record_id)

# fit model fun
fit_mod = function(outcome,df){

    # Fit model
  f = as.formula(paste0(outcome,"~","timepoint_cs"))
  mod = lme(f,
            random = ~1|record_id,
            data = df,
            na.action = na.omit)
 
  #print(summary(mod))
  
    # Anova
  mod_anova = anova.lme(mod, type="marginal")
  print(kable(mod_anova,digits = 3,caption = "Test of Overall Effect"))

 
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ timepoint_cs, adjust="none")
  mod_means2 = emmeans(mod,specs=~ timepoint_cs, adjust="none")

  print(kable(mod_means$contrasts[c(2,3)],digits = 3,caption = "Timepoint Contrasts"))
  print(kable(mod_means2,digits = 3,caption = "Timepoint Means"))
 # print(kable(mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE)))

   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
}

```

# 1: Table for sleep duration by age group and summer/school
```{r table, include = FALSE}
m6 = c(200,202,205,206,207,208,243,244,245,246)
analysis_table = analysis %>% mutate(age_grp = case_when(Age > 6 & Age < 13 ~ "6-12", Age > 13 & Age < 18 ~ "13-18")) %>% 
  filter(!(record_id %in% m6 & timepoint_cs == "6 Months"))

sleep_table = table1( ~ totalsleep_avg_min|visit_type*age_grp,
                      data = analysis_table,
                      topclass ="Rtable1-zebra",
                      render.continuous=c( "Mean (SD)"="Mean (SD)"))
```

```{r t1}
sleep_table
```

# 2: Differences in Sleep measures for summer/school 

```{r section 2, include=FALSE}
# All Awakenings
mod = glmer.nb(all_awakenings ~ visit_type + (1|record_id),data = analysis)
res = tidy(mod,"fixed")
res$term = sub("Period","",res$term)

# diabetes related awakenings
mod_dra = glmer.nb(diabetes_awakenings ~ visit_type + (1|record_id),data = analysis)
res_dra = tidy(mod_dra,"fixed")
res_dra$term = sub("Period","",res_dra$term)

# Total Sleep Time
mod_ts = lme(totalsleep_avg_min ~ visit_type, random = ~1|record_id, data = analysis, na.action = na.omit)
res_ts = tidy(mod_ts, "fixed")
res_ts$term = sub("Period","",res_ts$term)

# Sleep Efficiency
mod_se = lme(sleepefficency_avg ~ visit_type, random = ~1|record_id, data = analysis, na.action = na.omit)
res_se = tidy(mod_se, "fixed")
res_se$term = sub("Period","",res_se$term)

# WASO
mod_w = lme(waso_avg ~ visit_type, random = ~1|record_id, data = analysis, na.action = na.omit)
res_w = tidy(mod_w, "fixed")
res_w$term = sub("Period","",res_w$term)
```
Awakenings were analyzed using a generalized linear mixed effects model for the negative binomial family, with a random effect for participant. Sleep time, sleep efficiency, average WASO were analyzed using a linear mixed effects model with a random effect for participant.

The "Summer" term indicates the difference between school and summer, with school time as the reference.

## All Awakenings

There is not a significant difference in awakenings between the school and summer groups (p = 0.831)

```{r summer/school awa}
kable(res[,2:ncol(res)],digits = 3)
```

## Diabetes Related Awakenings

There is not a significant difference in awakenings between the school and summer groups (p = 0.829)

```{r summer/school d awa}
kable(res_dra[,2:ncol(res_dra)],digits = 3)
```

## Total Sleep Time

There is an association between total sleep time and school/summer groups (p = 0.031). On average, the summer group gets 40.32 (95%ci: 4.46, 76.18) more total sleep minutes than the School group.



```{r ss tst}
kable(res_ts[,2:ncol(res_ts)],digits = 3)
```

## Sleep Efficiency

There is not a significant difference in sleep efficiency between the school and summer groups (p = 0.439)


```{r ss se}
kable(res_se[,2:ncol(res_se)],digits = 3)
```


## WASO

There is not a significant difference in average WASO scores between the school and summer groups (p = 0.224)


```{r ss waso}
kable(res_w[,2:ncol(res_w)],digits = 3)
```


# 3: Just Summer visits baseline vs hcl

note: only 3 Baseline visits (subj 200, 202, 205) , 11 3mo, 10 6mo

```{r summerdata, include= FALSE}
analysis_summer = analysis %>% filter(visit_type == "Summer")
```

## avg total sleep time (min)

There is not a significant difference in avg total sleep time from Baseline to Month 3 (p = 0.942), or from Baseline to Month 6 (p= 0.522)
```{r total sleep sum}
fit_mod("totalsleep_avg_min", df = analysis_summer)
```

## avg sleep efficiency

There is not a significant difference in avg sleep efficiency from Baseline to Month 3 (p = 0.332), or from Baseline to Month 6 (p= 0.509)

```{r sleep eff sum}
fit_mod("sleepefficency_avg", df = analysis_summer)
```

## waso avg

There is not a significant difference in avg waso from Baseline to Month 3 (p = 0.692), or from Baseline to Month 6 (p= 0.359)


```{r waso sum}
fit_mod("waso_avg", df = analysis_summer)
```

## time in bed (min) avg

There is not a significant difference in avg minutes spent in bed from Baseline to Month 3 (p = 0.321), or from Baseline to Month 6 (p= 0.321)

```{r timeinbed sum}
fit_mod("timeinbed_avg_min", df = analysis_summer)
```

## tir 70-180

There is a significant difference in TIR between
Baseline and Month 3: Increase by 44.7 (21.759, 67.681) (p = 0.001)
and Baseline and Month 6:increase by 36.8 (14.64288, 59.02512) (p = 0.003)
```{r tir70180 sum}
fit_mod("sensor_70_180", df = analysis_summer)
```

## tbr 70
There is not a significant difference in TBR 70 from Baseline to Month 3 (p = 0.64), or from Baseline to Month 6 (p= 0.45)

```{r tbr70 sum}
fit_mod("sensor_u70", df = analysis_summer)
```

## tar 180

There is a significant difference in TAR between
Baseline and Month 3: Decrease by 44.97 (21.89, 68.05) (p = 0.001)
and Baseline and Month 6: Decrease by 37.60 (15.30, 59.90) (p = 0.003)

```{r tar180 sum}
fit_mod("sensor_g180", df = analysis_summer)
```

## sensor mean glucose

There is a significant difference in smg between
Baseline and Month 3: Decrease by 73.42 (38.03, 108.77) (p < 0.0001)
and Baseline and Month 6: Decrease by 37.60 (26.38, 94.74) (p = 0.002)
```{r smg sum}
fit_mod("sensor_mean", df = analysis_summer)
```

# 4: Just School visits baseline vs hcl

note:n =  36 bl, 24 m3, 19 m6

```{r schooldata, include= FALSE}
analysis_school = analysis %>% filter(visit_type == "School")
```


## avg total sleep time (min)

There is not a significant difference in avg total sleep time (min) from BL to M3 (p = 0.172)
There is a significant difference in avg total sleep time (min) between Baseline and Month 6: Decrease by 23.5 (5.90 41.16) (p = 0.013)


```{r total sleepschool}
fit_mod("totalsleep_avg_min", df = analysis_school)
```

## avg sleep efficiency

There is not a significant difference in avg sleep efficiency from Baseline to Month 3 (p = 0.828), or from Baseline to Month 6 (p= 0.881)

```{r total sleep _school}
fit_mod("sleepefficency_avg", df = analysis_school)
```

## waso avg

There is not a significant difference in avg waso from Baseline to Month 3 (p = 0.697), or from Baseline to Month 6 (p= 0.109)


```{r waso _school}
fit_mod("waso_avg", df = analysis_school)
```

## time in bed avg

There is not a significant difference in avg time in bed from Baseline to Month 3 (p = 0.070), or from Baseline to Month 6 (p= 0.073)

```{r timeinbed _school}
fit_mod("timeinbed_avg_min", df = analysis_school)
```

## tir 70-180

There is not a significant difference in TIR from Baseline to Month 3 (p = 0.403), or from Baseline to Month 6 (p= 0.827)


```{r tir70180 _school}
fit_mod("sensor_70_180", df = analysis_school)
```

## tbr 70

There is not a significant difference in tbr70 from Baseline to Month 3 (p = 0.383), or from Baseline to Month 6 (p= 0.724)


```{r tbr70 _school}
fit_mod("sensor_u70", df = analysis_school)
```

## tar 180

There is not a significant difference in TAR180 from Baseline to Month 3 (p = 0.552), or from Baseline to Month 6 (p= 0.654)


```{r tar180 _school}
fit_mod("sensor_g180", df = analysis_school)
```

## sensor mean glucose

There is not a significant difference in smg from Baseline to Month 3 (p = 0.656), or from Baseline to Month 6 (p= 0.897)

```{r smg _school}
fit_mod("sensor_mean", df = analysis_school)
```
 

