---
title: "HCL and sleep: changes in sleep after starting Control IQ"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tableone)
library(dplyr)
library(stringr)
library(nlme)
```

```{r echo=FALSE,include=FALSE}
source("T:\\Erin Cobry\\Prospective HCL and sleep study\\Data raw\\AdvancedHybridClosed-Demographics_R_2021-11-08_1921.r")
demo <- data
demo <- demo[!(demo$record_id %in% c(121,221,101,201,102,202,103,203,121,221,111,211,131,231,141,241)),]
demo_child <- subset(demo, record_id>=200)
demo_child$consent_date <- as.Date(as.character(demo_child$consent_date))
demo_adult <- subset(demo, record_id<200)
demo_adult$parent_bday <- as.Date(as.character(demo_adult$parent_bday))
# to calculate age, need to get consent date from child demo
consent <- demo_child %>% select(record_id, consent_date)
consent$record_id <- consent$record_id-100
demo_adult$consent_date <- NULL
demo_adult <- merge(demo_adult,consent,by="record_id",all.x=T,all.y = T)
demo_adult$age <- floor(((demo_adult$consent_date - demo_adult$parent_bday)/365.25))
demo_adult$age <- ifelse(demo_adult$age==0,NA,demo_adult$age)
demo_adult$parent_check_night <- as.numeric(demo_adult$parent_check_night)
# calculate child duration
demo_child$t1d_diagnosis <- as.Date(as.character(demo_child$t1d_diagnosis))
demo_child$t1d_duration <- as.numeric(((demo_child$consent_date - demo_child$t1d_diagnosis)/365.25))
# merge in insulin use, CGM use, HCL use from parent questionnaire
ins <- demo_adult[c("record_id","insulinmethod.factor","previous_cgm_use.factor","currenthcltech.factor")]
ins$record_id <- ins$record_id+100
demo_child$insulinmethod.factor <- NULL
demo_child$previous_cgm_use.factor <- NULL
demo_child$currenthcltech.factor <- NULL
demo_child <- merge(demo_child,ins,by="record_id",all.x=T,all.y=T)
demo_child$cons_age <- as.numeric(demo_child$cons_age)

source("T:\\Erin Cobry\\Prospective HCL and sleep study\\Data raw\\AdvancedHybridClosed-ATTDAbstracts2022_R_2021-11-08_2000.r")
data <- data[!(data$record_id %in% c(121,221,101,201,102,202,103,203,121,221,111,211,131,231,141,241)),]
# fix total sleep variable
data$totalsleep_avg <- as.character(data$totalsleep_avg)
data$hours <-  as.numeric(word(data$totalsleep_avg,1,sep=":"))
data$minutes <- as.numeric(word(data$totalsleep_avg,2,sep=":"))
data$totalsleep_avg_num <- data$hours + (data$minutes/60)
data$redcap_event_name.factor <- droplevels(data$redcap_event_name.factor)
# keep needed actigraphy variables
actigraphy <- data[,c("record_id","redcap_event_name","totalsleep_avg_num","sleepefficency_avg","waso_avg")]
# create new actigraphy visit variable
actigraphy$visit <- ifelse(actigraphy$redcap_event_name %in% c("baseline_day_1_arm_1","baseline_day_1_arm_2"),"B","WK1")
actigraphy_wide <- reshape(actigraphy, idvar = "record_id", timevar = "visit", direction = "wide")
actigraphy_wide$d_tot <- as.numeric(actigraphy_wide$totalsleep_avg_num.WK1 - actigraphy_wide$totalsleep_avg_num.B)
actigraphy_wide$d_eff <- as.numeric(actigraphy_wide$sleepefficency_avg.WK1 - actigraphy_wide$sleepefficency_avg.B)
actigraphy_wide$d_waso <- as.numeric(actigraphy_wide$waso_avg.WK1 - actigraphy_wide$waso_avg.B)
#label(actigraphy_wide$d_tot) <- "Change in total sleep time"
#label(actigraphy_wide$d_eff) <- "Change in sleep efficiency"
#label(actigraphy_wide$d_waso) <- "Change in WASO"
actigraphy_wide_child <- actigraphy_wide[actigraphy_wide$record_id>=200,]
actigraphy_wide_adult <- actigraphy_wide[actigraphy_wide$record_id<200,]

# create new variables indicating whether the child got <9, <8, <7 hours at baseline
actigraphy_wide_child$b_9hours <- ifelse(is.na(actigraphy_wide_child$totalsleep_avg_num.B),NA,
                                         ifelse(actigraphy_wide_child$totalsleep_avg_num.B<9,0,1))
actigraphy_wide_child$b_8hours <- ifelse(is.na(actigraphy_wide_child$totalsleep_avg_num.B),NA,
                                         ifelse(actigraphy_wide_child$totalsleep_avg_num.B<8,0,1))
actigraphy_wide_child$b_7hours <- ifelse(is.na(actigraphy_wide_child$totalsleep_avg_num.B),NA,
                                         ifelse(actigraphy_wide_child$totalsleep_avg_num.B<7,0,1))







# adult demo table
# need to add child BG checks when those data are clean
adult_demo_vars <- c("age","parent_gender.factor","race.factor","parent_check_night")
t1_adult <- CreateTableOne(data=demo_adult,vars=adult_demo_vars)
t1_adult <- print(t1_adult,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T,nonnorm=c("age"),test = F)

# child demo table
child_demo_vars <- c("cons_age","childgender.factor","t1d_duration","insulinmethod.factor","previous_cgm_use.factor",
                     "currenthcltech.factor")
t1_child <- CreateTableOne(data=demo_child,vars=child_demo_vars)
t1_child <- print(t1_child,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T,nonnorm=c("t1d_duration"),test = F)

# adult actigraphy table
actvars <- c("d_tot","d_eff","d_waso","totalsleep_avg_num.B","totalsleep_avg_num.WK1","sleepefficency_avg.B",
             "sleepefficency_avg.WK1","waso_avg.B","waso_avg.WK1")
mean <- apply(actigraphy_wide_adult[,actvars],2,mean,na.rm=T)
sd <- apply(actigraphy_wide_adult[,actvars],2,sd,na.rm=T)
mean <- round(mean,4)
sd <- round(sd,4)
a <- actigraphy_wide_adult[,c("d_tot","d_eff","d_waso","totalsleep_avg_num.B","totalsleep_avg_num.WK1","sleepefficency_avg.B",
             "sleepefficency_avg.WK1","waso_avg.B","waso_avg.WK1")]
p.value <- NULL
for (i in 1:length(actvars)) {
  p.value[i] <- t.test(a[,i],mu=0)$p.value
}
act_table_adult <- cbind(mean,sd,p.value)
row.names(act_table_adult) <- c("Change in total sleep time (hrs)","Change in sleep efficiency (%)","Change in WASO (min)",
                                "Baseline total sleep time (hrs)","Wk1 total sleep time (hrs)","Baseline sleep efficiency (%)",
                                "Wk1 sleep efficiency (%)","Baseline WASO (min)","Wk1 WASO (min)")

# child actigraphy table
actvars <- c("d_tot","d_eff","d_waso","totalsleep_avg_num.B","totalsleep_avg_num.WK1","sleepefficency_avg.B",
             "sleepefficency_avg.WK1","waso_avg.B","waso_avg.WK1")
mean <- apply(actigraphy_wide_child[,actvars],2,mean,na.rm=T)
mean <- round(mean,4)
sd <- apply(actigraphy_wide_child[,actvars],2,sd,na.rm=T)
sd <- round(sd,4)
a <- actigraphy_wide_child[,c("d_tot","d_eff","d_waso","totalsleep_avg_num.B","totalsleep_avg_num.WK1","sleepefficency_avg.B",
             "sleepefficency_avg.WK1","waso_avg.B","waso_avg.WK1")]
p.value <- NULL
for (i in 1:length(actvars)) {
  p.value[i] <- t.test(a[,i],mu=0)$p.value
}
act_table_child <- cbind(mean,sd,p.value)
row.names(act_table_child) <- c("Change in total sleep time (hrs)","Change in sleep efficiency (%)","Change in WASO (min)",
                                "Baseline total sleep time (hrs)","Wk1 total sleep time (hrs)","Baseline sleep efficiency (%)",
                                "Wk1 sleep efficiency (%)","Baseline WASO (min)","Wk1 WASO (min)")

# adult diary table
diaryvars <- c("d_tot_awakenings","d_t1d_awakenings","tot_awakenings.B","tot_awakenings.WK1","tot_t1d_awakenings.B","tot_t1d_awakenings.WK1")
mean <- apply(diary_wide_adult[,diaryvars],2,mean,na.rm=T)
mean <- round(mean,4)
sd <- apply(diary_wide_adult[,diaryvars],2,sd,na.rm=T)
sd <- round(sd,4)
a <- diary_wide_adult[,c("d_tot_awakenings","d_t1d_awakenings","tot_awakenings.B","tot_awakenings.WK1","tot_t1d_awakenings.B","tot_t1d_awakenings.WK1")]
p.value <- NULL
for (i in 1:length(diaryvars)) {
  p.value[i] <- round(t.test(a[,i],mu=0)$p.value,4)
}
diary_table_adult <- cbind(mean,sd,p.value)
row.names(diary_table_adult) <- c("Change in number of total awakenings","Change in number of diabetes-related awakenings",
                                  "Baseline total awakenings","Wk1 total awakenings","Baseline diabetes-related awakenings",
                                  "Wk1 diabetes-related awakenings")

# child diary table
diaryvars <- c("d_tot_awakenings","d_t1d_awakenings","tot_awakenings.B","tot_awakenings.WK1","tot_t1d_awakenings.B","tot_t1d_awakenings.WK1")
mean <- apply(diary_wide_child[,diaryvars],2,mean,na.rm=T)
mean <- round(mean,4)
sd <- apply(diary_wide_child[,diaryvars],2,sd,na.rm=T)
sd <- round(sd,4)
a <- diary_wide_child[,c("d_tot_awakenings","d_t1d_awakenings","tot_awakenings.B","tot_awakenings.WK1","tot_t1d_awakenings.B","tot_t1d_awakenings.WK1")]
p.value <- NULL
for (i in 1:length(diaryvars)) {
  p.value[i] <- round(t.test(a[,i],mu=0)$p.value,4)
}
diary_table_child <- cbind(mean,sd,p.value)
row.names(diary_table_child) <- c("Change in number of total awakenings","Change in number of diabetes-related awakenings",
                                  "Baseline total awakenings","Wk1 total awakenings","Baseline diabetes-related awakenings",
                                  "Wk1 diabetes-related awakenings")

# now create tables for child variables by age group
age_child <- demo_child[,c("record_id","cons_age")]
age_child$age_cat <- ifelse(age_child$cons_age>=2 & age_child$cons_age<=5,"Ages 2-5",
                            ifelse(age_child$cons_age<=12,"Ages 6-12","Ages 13-17"))
actigraphy_wide_child <- merge(actigraphy_wide_child,age_child,by="record_id",all.x = T,all.y = T)
diary_wide_child$record_id <- diary_wide_child$Record.ID
diary_wide_child <- merge(diary_wide_child,age_child,by="record_id",all.x = T,all.y = T)

t_act_age <- CreateTableOne(data=actigraphy_wide_child,vars=actvars,strata = "age_cat")
t_act_age <- print(t_act_age,printToggle = F,varLabels=F,missing=T,showAllLevels = T,test = F)

t_diary_age <- CreateTableOne(data=diary_wide_child,vars=diaryvars,strata = "age_cat")
t_diary_age <- print(t_diary_age,printToggle = F,varLabels=F,missing=T,showAllLevels = T,test = F)


```

# Background

The purpose of this analysis is to test whether sleep measures, by actigraphy and self-report, change 1 week after starting Control IQ, in both adults and children.

# Methods

The actigraphy sleep variables were all average values (average total sleep time, average sleep efficiency, average WASO).  The total number of awakenings and number of diabetes-related awakenings were calculated during the week of sleep diary data for each visit.  Paired t-tests were used to compare sleep measures at baseline and at 1 week.

Note that the length of awakening variables had a large amount of missing data; therefore, those measures are not included in this analysis.  Of the first awakenings reported, `r permiss_a1`% had the length missing.  Similarly, the percentage of missing length data for the 2nd, 3rd, and 4th awakenings were `r permiss_a2`%,`r permiss_a3`%, and `r permiss_a4`%.

# Results

```{r, echo=FALSE, message=FALSE}
kable(t1_adult,caption="Table 1. Demographics and clinical characteristics in parents.")
```

<br>

```{r, echo=FALSE, message=FALSE}
kable(t1_child,caption="Table 2. Demographics and clinical characteristics in children.")
```

\newpage

```{r, echo=FALSE, message=FALSE}
kable(act_table_adult,caption="Table 3.  Change in sleep variables at 1 week after starting control IQ, in adults.")
```

<br>

```{r, echo=FALSE, message=FALSE}
kable(act_table_child,caption="Table 4.  Change in sleep variables at 1 week after starting control IQ, in children.")
```

<br>

```{r, echo=FALSE, message=FALSE}
kable(diary_table_adult,caption="Table 5.  Change in awakenings at 1 week after starting control IQ, in adults.")
```

<br>

```{r, echo=FALSE, message=FALSE}
kable(diary_table_child,caption="Table 6.  Change in awakenings at 1 week after starting control IQ, in children.")
```

\newpage

<br>

```{r, echo=FALSE, message=FALSE}
kable(t_act_age,caption="Table 7.  Sleep variables in children, by age group")
```

<br>

```{r, echo=FALSE, message=FALSE}
kable(t_diary_age,caption="Table 8.  Awakening variables in children, by age group")
```
