---
title: "ASBMR_QCT"
author: "Casey Sakamoto"
date: "6/3/2022"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(Hmisc)
library(stringr)
library(table1)
library(knitr)
library(lubridate)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Viral Shah/K23"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/K23"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/K23"
}
knitr::opts_knit$set(root.dir = home_dir)

setwd("S:/Laura/BDC/Projects/Viral Shah/K23")
rm(home_dir)
```

```{r dataup, include = FALSE}
BSTM_t1d = read_csv("BSTM_t1d_raw.csv")

# grab vars of interest
analysis = BSTM_t1d %>% filter(qct == 1) %>% mutate(age_onset = as.numeric(difftime(date_of_diabetes_diagnosis, date_of_birth)/365.25),
                                                    age_onset_gt20 = case_when(age_onset > 20~1,
                                                                               age_onset<= 20 ~ 0))%>%
  select(record_id, age,age_onset_gt20,t1d_diagnosis, demo_gender1, postmenopause_pill_42695d, are_you_currently_smoking_e66c9b, do_you_drink_3_or_more_uni_ca666a, stairs_flight, sports_past_week,did_you_participate_in_any, sport_team_text, regular_activity_past_week, last_week_regular_number_of_times, fall_diab_duration, hba1c_value_k, egfr,albumin_serum, creatinine_serum, pth_baseline, calcium_serum, p1np_baseline, ctx_baseline, age_menopause_7904ab, fe_hip_stance:qct_complete)

# some hba1c are listed with a % sign (LITERALLY WHY); need to remove that
analysis = analysis %>% mutate(hba1c = 
                                 case_when(str_detect(hba1c_value_k, "%") ~ as.numeric(str_sub(hba1c_value_k, 1,str_length(hba1c_value_k)-1)),
                                           TRUE ~ as.numeric(hba1c_value_k))) %>% select(-hba1c_value_k)

# need to fix ctx with the truncated values by dividing by 2
analysis = analysis %>% mutate(ctx_bl = 
                                 case_when(str_detect(ctx_baseline, "<") ~ (as.numeric(str_sub(ctx_baseline, 2))/2),
                                           TRUE ~ as.numeric(ctx_baseline)),
                              ) %>% select(-c( ctx_baseline))
analysis[analysis == "not done" ] <- NA

# add in menopausal duration
analysis = analysis %>% mutate(menopause_dur = age - age_menopause_7904ab)
label(analysis$menopause_dur) = "Menopausal Duration"

label(analysis$t1d_diagnosis) = "Clinical T1D Diagnosis"
analysis$t1d_diagnosis = factor(analysis$t1d_diagnosis, labels = c("Control", "T1D"))

analysis$demo_gender1 = factor(analysis$demo_gender1, labels = c("Female"))
analysis$postmenopause_pill_42695d = factor(analysis$postmenopause_pill_42695d, labels = c("No HRT", "HRT"))
analysis$are_you_currently_smoking_e66c9b = factor(analysis$are_you_currently_smoking_e66c9b, labels = c("No", "Yes"))

label(analysis$fall_diab_duration) = "T1D Duration (Years)"
label(analysis$do_you_drink_3_or_more_uni_ca666a) = "Alcohol > 3 Drinks"
label(analysis$demo_gender1) = "Sex"
label(analysis$age) = "Age"
label(analysis$postmenopause_pill_42695d) = "HRT"
label(analysis$are_you_currently_smoking_e66c9b) = "Current Smoker"
label(analysis$hba1c) = "HBA1C"
label(analysis$egfr) = "eGFR"
label(analysis$albumin_serum) = "Serum Albumin"
label(analysis$creatinine_serum) = "Serum Creatinine"
label(analysis$pth_baseline) = "PTH"
label(analysis$calcium_serum) = "Calcium"
label(analysis$p1np_baseline) = "P1NP"
label(analysis$ctx_bl) = "CTX"

# factor the phys act variables
label(analysis$sports_past_week) = "Sports/Recreation Past Week"
analysis$sports_past_week = factor(analysis$sports_past_week, labels = c("No", "Yes"))

label(analysis$did_you_participate_in_any) = "Sports/Recreation Past Year"
analysis$did_you_participate_in_any = factor(analysis$did_you_participate_in_any, labels = c("No", "Yes"))

label(analysis$regular_activity_past_week) = "Regular Activity Past Week"
analysis$regular_activity_past_week = factor(analysis$regular_activity_past_week, labels = c("No", "Yes"))

# function for pvalue in table1
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

# function to fit interaction linear mod 

# add in?
# "fall_diab_duration","+","age_onset_gt20"

fit_mod = function(outcome,df){

    # Fit p1np model
  f = as.formula(paste0("p1np_baseline~",outcome, "*", "t1d_diagnosis","+","age","+","hba1c"))
  mod = lm(f,
            data = df,
            na.action = na.omit)
   cat("\n")
  cat(paste("P1NP Model"))
  cat("\n")
 
  print(summary(mod))
   
 
  # Fit ctx model
  f2 = as.formula(paste0("ctx_bl~",outcome,"*","t1d_diagnosis","+","age","+","hba1c"))
  mod2 = lm(f2,
            data = df,
            na.action = na.omit)
   cat("\n")
  cat(paste("CTX Model"))
  cat("\n")
 
  print(summary(mod2))
   
  }

```


```{r plots}
# predict the p1np/ctx for hba1c interaction
analysis_plot = analysis %>% select(record_id,t1d_diagnosis, hba1c, p1np_baseline, ctx_bl) %>% filter(!is.na(p1np_baseline) | !is.na(ctx_bl))

# splitup data set to make lines & plot
analysis_control = analysis_plot %>% filter(t1d_diagnosis == "Control")
analysis_t1d = analysis_plot %>% filter(t1d_diagnosis == "T1D")
#p1np
hba1c_p = ggplot(data = analysis_plot , aes(x = hba1c, y = p1np_baseline,  color = t1d_diagnosis)) + 
      geom_point(na.rm = T) + geom_smooth(method = lm, se = F,na.rm = T, data = analysis_control)  + geom_smooth(method = lm, se = F,na.rm = T, data = analysis_t1d) +
      theme_classic() 

hba1c_c = ggplot(data = analysis_plot , aes(x = hba1c, y = ctx_bl,  color = t1d_diagnosis)) + 
      geom_point(na.rm = T) + geom_smooth(method = lm, se = F,na.rm = T, data = analysis_control)  + geom_smooth(method = lm, se = F,na.rm = T, data = analysis_t1d) +
      theme_classic() 
hba1c_c = hba1c_c + theme(text = element_text(size = 20)) + labs(y = "CTX Baseline")
hba1c_p = hba1c_p + theme(text = element_text(size = 20)) + labs(y = "P1NP Baseline")
```
# Baseline Characteristics
```{r t1}
bl_t1 = table1(~  age+demo_gender1+postmenopause_pill_42695d+menopause_dur+are_you_currently_smoking_e66c9b+do_you_drink_3_or_more_uni_ca666a+sports_past_week+did_you_participate_in_any+regular_activity_past_week+fall_diab_duration+hba1c+ egfr+albumin_serum+creatinine_serum+pth_baseline+calcium_serum|t1d_diagnosis,data = analysis, render.continuous=c(.="Mean (SD)") )
bl_t1_2 = table1(~p1np_baseline+ctx_bl|t1d_diagnosis,data = analysis, render.continuous=c(.="Mean (SD)"),overall = FALSE, extra.col=list(`P-value`=pvalue) )

bl_t1

bl_t1_2
```

# QCT Hip Differences 
```{r t2}
analysis_2 = analysis %>% select(t1d_diagnosis,fe_hip_stance:l2_bmd_mg_cm_3 )
bl_t2 = table1(~.  |t1d_diagnosis,data = analysis_2, render.continuous=c(.="Mean (SD)"), extra.col=list(`P-value`=pvalue),overall = FALSE )

bl_t2
```

# Bone Quality Predictors
Models adjusted for Age, diabetes duration, age of onset (<20 years vs >20 years), and HbA1c  

## tot_bon_tot_hip_mass
### P1NP: Not associated with P1np (p = 0.42 ), or different between groups (p = 0.72)
### CTX: Not associated with CTX (p = 0.38 ), or different between groups (p = 0.57)


## cort_bone_comp_tot_hip_vbmd
### P1NP: Not associated with P1np (p = 0.58 ), or different between groups (p = 0.97)
### CTX: Not associated with CTX (p = 0.68 ), or different between groups (p = 0.83)

## l1_bmd_mg_cm_3
### P1NP: Not associated with P1np (p = 0.74), or different between groups (p = 0.94)
### CTX: Not associated with CTX (p = 0.45 ), or different between groups (p = 0.85)


## l2_bmd_mg_cm_3
### P1NP: Not associated with P1np (p = 0.42 ), or different between groups (p = 0.72)
### CTX: Not associated with CTX (p = 0.74 ), or different between groups (p = 0.92)


```{r plotts}
# hba1c_p
# hba1c_c
```

# Code Appendix
```{r qct interactions}
# any interaction between QCT variables

# tot_bon_tot_hip_mass
fit_mod("tot_bon_tot_hip_mass", analysis)

# cort_bone_comp_tot_hip_vbmd
fit_mod("cort_bone_comp_tot_hip_vbmd", analysis)

# l1_bmd_mg_cm_3
fit_mod("l1_bmd_mg_cm_3", analysis)

# l2_bmd_mg_cm_3
fit_mod("l2_bmd_mg_cm_3", analysis)
```