---
title: "ASBMR Abstract"
author: "Casey Sakamoto"
date: "6/6/2022"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(Hmisc)
library(stringr)
library(table1)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Viral Shah/K23"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/K23"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/K23"
}
knitr::opts_knit$set(root.dir = home_dir)

setwd("S:/Laura/BDC/Projects/Viral Shah/K23")
rm(home_dir)
```

```{r data import and clean, include=FALSE}
# import data
BSTM_t1d = read_csv("BSTM_t1d.csv")
analysis_t = read_csv("bstm_t1d_raw.csv")
cgm_summary = read_csv("S:/Laura/BDC/Projects/Viral Shah/K23/Data_Clean/cgm_r1_summary_stats.csv")
cgm_summary = cgm_summary %>% mutate(record_id = gsub("(\\w$)", "\\U\\1", ID, perl = TRUE)) %>% filter(`Sensor Readings` > 100)
# only want cgm data with >70% (>940.8 for 15 min, and >2822.4 for 5 min intervals)

# only interested in patients with t1d
# want fingerstick glucose, 14 day cgm metrics, a1c, p1np and ctx
analysis = BSTM_t1d %>% filter(`2. Clinical Diagnosis of T1D` == "Yes") %>% 
  select(`Record ID`, `Finger Stick Blood Glucose mg/dl`, `HbA1C value:`, `P1NP Baseline ng/ml`, `CTX Baseline ng/ml`)

# fingerstick bg to numeric and fix "not done" to NA
analysis[ analysis == "not done" ] <- NA
analysis = analysis %>% mutate(fsbg = as.numeric(`Finger Stick Blood Glucose mg/dl`)) %>% select(- `Finger Stick Blood Glucose mg/dl`)

# some hba1c are listed with a % sign (LITERALLY WHY); need to remove that
analysis = analysis %>% mutate(hba1c = 
                                 case_when(str_detect(`HbA1C value:`, "%") ~ as.numeric(str_sub(`HbA1C value:`, 1,str_length(`HbA1C value:`)-1)),
                                           TRUE ~ as.numeric(`HbA1C value:`))) %>% select(-`HbA1C value:`)

# need to fix ctx with the truncated values by dividing by 2
analysis = analysis %>% mutate(ctx_bl = 
                                 case_when(str_detect(`CTX Baseline ng/ml`, "<") ~ (as.numeric(str_sub(`CTX Baseline ng/ml`, 2))/2),
                                           TRUE ~ as.numeric(`CTX Baseline ng/ml`)),
                               p1np_bl = `P1NP Baseline ng/ml`) %>% select(-c(`P1NP Baseline ng/ml`, `CTX Baseline ng/ml`))
id_list = analysis$`Record ID`
cgm_summary = cgm_summary %>% filter(record_id %in% id_list) %>% distinct()
analysis = left_join(analysis, cgm_summary, by = c("Record ID" = "record_id"))
# list of subj id's and date for tim
# subjdates = BSTM_t1d %>% filter(`2. Clinical Diagnosis of T1D` == "Yes") %>% 
#   select(`Record ID`, `Date of visit`, `Date of the study visit:`)
# write.csv(subjdates, file = "asbmr_subjdates_cgm.csv")
```

```{r table data import, include=FALSE}
# variable select
demog_df = analysis_t %>% filter(t1d_diagnosis == 1) %>% select(record_id, age, demo_gender1, fall_diab_duration, postmenopause_pill_42695d, are_you_currently_smoking_e66c9b, do_you_drink_3_or_more_uni_ca666a, hba1c_value_k, egfr,albumin_serum, creatinine_serum, pth_baseline, calcium_serum, p1np_baseline, ctx_baseline, finger_stick_blood_glucose)

# some hba1c are listed with a % sign (LITERALLY WHY); need to remove that
demog_df = demog_df %>% mutate(hba1c = 
                                 case_when(str_detect(hba1c_value_k, "%") ~ as.numeric(str_sub(hba1c_value_k, 1,str_length(hba1c_value_k)-1)),
                                           TRUE ~ as.numeric(hba1c_value_k))) %>% select(-hba1c_value_k)

# need to fix ctx with the truncated values by dividing by 2
demog_df = demog_df %>% mutate(ctx_bl = 
                                 case_when(str_detect(ctx_baseline, "<") ~ (as.numeric(str_sub(ctx_baseline, 2))/2),
                                           TRUE ~ as.numeric(ctx_baseline)),
                              ) %>% select(-c( ctx_baseline))
demog_df[demog_df == "not done" ] <- NA
demog_df = demog_df %>% mutate(fsbg = as.numeric(finger_stick_blood_glucose)) %>% select(- `finger_stick_blood_glucose`)

demog_df$demo_gender1 = factor(demog_df$demo_gender1, labels = c("Female"))
demog_df$postmenopause_pill_42695d = factor(demog_df$postmenopause_pill_42695d, labels = c("No HRT", "HRT"))
demog_df$are_you_currently_smoking_e66c9b = factor(demog_df$are_you_currently_smoking_e66c9b, labels = c("No", "Yes"))

label(demog_df$fall_diab_duration) = "T1D Duration (Years)"
label(demog_df$demo_gender1) = "Sex"
label(demog_df$age) = "Age"
label(demog_df$postmenopause_pill_42695d) = "HRT"
label(demog_df$are_you_currently_smoking_e66c9b) = "Current Smoker"
label(demog_df$hba1c) = "HBA1C"
label(demog_df$egfr) = "eGFR"
label(demog_df$albumin_serum) = "Serum Albumin"
label(demog_df$creatinine_serum) = "Serum Creatinine"
label(demog_df$pth_baseline) = "PTH"
label(demog_df$calcium_serum) = "Calcium"
label(demog_df$p1np_baseline) = "P1NP"
label(demog_df$ctx_bl) = "CTX"
label(demog_df$fsbg) = "Finger Stick Blood Glucose"

demog_df = left_join(demog_df, cgm_summary, by = "record_id") 
demog_df = demog_df[-14,] # duplicate id
# table 1 create
demogs = table1(~age+demo_gender1+fall_diab_duration+postmenopause_pill_42695d+are_you_currently_smoking_e66c9b+ `14 Day Mean` + `14 Day TIR < 70` + 
                   `14 Day TIR 70 - 140` + `14 Day TIR 70 - 180` + `14 Day TIR Over 180`+ CV + hba1c+egfr+
                  albumin_serum+ creatinine_serum+pth_baseline+calcium_serum+p1np_baseline+ctx_bl + fsbg,
       data=demog_df,  render.continuous=c(.="Mean (SD)"))
```

```{r functions}
# function to output scatter, corr and linear mod
fit_mod = function(outcome,df,plot = T,diagnostics = F){
  
    # Fit p1np model
  f = as.formula(paste0("p1np_bl~",outcome))
  mod = lm(f,
            data = df,
            na.action = na.omit)
   cat("\n")
  cat(paste("P1NP Model"))
  cat("\n")
 
  print(summary(mod))
   

  
   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
 
  # Fit ctx model
  f2 = as.formula(paste0("ctx_bl~",outcome))
  mod2 = lm(f2,
            data = df,
            na.action = na.omit)
   cat("\n")
  cat(paste("CTX Model"))
  cat("\n")
 
  print(summary(mod2))
   

  
   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod2),"observations."))
  cat("\n") 
  
  # Plots
  if(plot){
    p = 
      ggplot(data = df,aes_string(x = outcome, y = "p1np_bl")) + 
      geom_point(na.rm = T) + geom_smooth(method = lm, se= F, na.rm = T)  +
      theme_classic() 
    
    
        p2 = 
      ggplot(data = df,aes_string(x = outcome, y = "ctx_bl")) + 
      geom_point(na.rm = T) + geom_smooth(method = lm, se = F,na.rm = T)  + 
      theme_classic() 
    
      p = p + theme(text = element_text(size = 20))
      p2 = p2 + theme(text = element_text(size = 20))
      
        if(outcome == "`14 Day TIR 70 - 140`" | outcome == "`14 Day TIR 70 - 180`" ){
          p = p + scale_x_continuous(limits = c(),breaks = c(25,50,75,100))
          p2 = p2 + scale_x_continuous(breaks = c(25,50,75,100))
        }
        
         if(outcome == "`14 Day TIR Over 180`"){
          p = p + scale_x_continuous(breaks = c(10,20,30,40,50))
          p2 = p2 + scale_x_continuous(breaks = c(10,20,30,40,50))
         }
        
                 if(outcome == "`14 Day TIR < 70`"){
          p = p + scale_x_continuous(breaks = c(0,2,4,6,8,10))
          p2 = p2 + scale_x_continuous(breaks = c(0,2,4,6,8,10))
        }
        print(p)
        print(p2)
  }
 
   # Check model
  if(diagnostics){
    print(check_model(mod))
  }
}

```

# Outstanding Data Questions/ Remarks
Remarks as of 6/6/2022

* truncated values of ctx (some data listed as "<0.076") were inputted as 0.5 * truncated value
* Subjects with less than 70% CGM readings were not included for cgm analysis

# Methods

Simple linear models  were fit with Fingerstick Glucose, 14 Days CGM metrics, and A1C respectively to assess the linear relationship with with P1NP and CTX. Only subjects with T1D diagnosis (n = 31) were included in analysis.

# Analysis

## Added in P1NP Model With Covariates 6/23/2022
```{r p2np}

```

## Demographics Table

```{r t1}
demogs
```

## Fingerstick Glucose 

* very weak linear relationships between P1NP and CTX and Fingerstick Glucose (r = -0.18 and -0.05 respectively)
```{r fingerstickglucose}
# correlations
f_c1 = cor(analysis$fsbg, analysis$p1np_bl, use = "pairwise.complete.obs")
f_c2 = cor(analysis$fsbg, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "fsbg", df = analysis)

cat("\n")
cat(paste("P1NP Pearson Corr Coef = ",f_c1))
cat(paste("CTX Pearson Corr Coef = ",f_c2))
cat("\n")
```

## 14 Day CGM Metrics

### 14 Day Mean

* weak linear relationship between 14 day mean and p1np and ctx (r = 0.18, 0.03 respectively)

```{r 14dm}
# correlations
m14d_c1 = cor(analysis$`14 Day Mean`, analysis$p1np_bl, use = "pairwise.complete.obs")
m14d_c2 = cor(analysis$`14 Day Mean`, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "`14 Day Mean`", df = analysis)

cat("\n")
cat(paste("P1NP Pearson Corr Coef = ",m14d_c1))
cat(paste("CTX Pearson Corr Coef = ",m14d_c2))
cat("\n")
```

### 14 Day TIR < 70

* very weak linear relationship between 14 day mean and p1np and ctx (r = 0.12, -0.10 respectively)

```{r tir lt70}
# correlations
b70_c1 = cor(analysis$`14 Day TIR < 70`, analysis$p1np_bl, use = "pairwise.complete.obs")
b70_c2 = cor(analysis$`14 Day TIR < 70`, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "`14 Day TIR < 70`", df = analysis)

cat("\n")
(cat(paste("P1NP Pearson Corr Coef = ",b70_c1)))
(cat(paste("CTX Pearson Corr Coef = ",b70_c2)))
cat("\n")
```

### 14 Day TIR 70-140

* weak linear relationship between 14 day mean and p1np and ctx (r = -0.09, -0.01 respectively)

```{r tir70140}
# correlations
b70_140_c1 = cor(analysis$`14 Day TIR 70 - 140`, analysis$p1np_bl, use = "pairwise.complete.obs")
b70_140_c2 = cor(analysis$`14 Day TIR 70 - 140`, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "`14 Day TIR 70 - 140`", df = analysis)

cat("\n")
(cat(paste("P1NP Pearson Corr Coef = ",b70_140_c1)))
(cat(paste("CTX Pearson Corr Coef = ",b70_140_c2)))
cat("\n")
```

### 14 Day TIR 70-180

* weak linear relationship between 14 day mean and p1np and ctx (r = -0.24, -0.07 respectively)

```{r tir70180}
# correlations
b70_180_c1 = cor(analysis$`14 Day TIR 70 - 180`, analysis$p1np_bl, use = "pairwise.complete.obs")
b70_180_c2 = cor(analysis$`14 Day TIR 70 - 180`, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "`14 Day TIR 70 - 180`", df = analysis)

cat("\n")
(cat(paste("P1NP Pearson Corr Coef = ",b70_180_c1)))
(cat(paste("CTX Pearson Corr Coef = ",b70_180_c2)))
cat("\n")
```

### 14 Day TIR >180

* weak linear relationship between 14 day mean and p1np and ctx (r = 0.22, 0.08 respectively)

```{r tirgt180}
# correlations
g180_c1 = cor(analysis$`14 Day TIR Over 180`, analysis$p1np_bl, use = "pairwise.complete.obs")
g180_c2 = cor(analysis$`14 Day TIR Over 180`, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "`14 Day TIR Over 180`", df = analysis)

cat("\n")
(cat(paste("P1NP Pearson Corr Coef = ",g180_c1)))
(cat(paste("CTX Pearson Corr Coef = ",g180_c2)))
cat("\n")
```


## A1C

 * very weak linear relationships between P1NP and CTX and AIC (r = -0.18 and -0.18 respectively)
```{r hba1c}
# correlations
h_c1 = cor(analysis$hba1c, analysis$p1np_bl, use = "pairwise.complete.obs")
h_c2 = cor(analysis$hba1c, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "hba1c", df = analysis)

cat("\n")
cat(paste("P1NP Pearson Corr Coef = ",h_c1))
cat(paste("CTX Pearson Corr Coef = ",h_c2))
cat("\n")
```
