---
title: "ASBMR Abstract"
author: "Casey Sakamoto"
date: "4/27/2022"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(Hmisc)
library(stringr)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Viral Shah/K23"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/K23"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/K23"
}
knitr::opts_knit$set(root.dir = home_dir)

setwd("S:/Laura/BDC/Projects/Viral Shah/K23")
rm(home_dir)
```

```{r data import and clean, include=FALSE}
# import data
BSTM_t1d <- read_csv("BSTM_t1d.csv")

# only interested in patients with t1d
# want fingerstick glucose, 14 day cgm metrics, a1c, p1np and ctx
analysis = BSTM_t1d %>% filter(`2. Clinical Diagnosis of T1D` == "Yes") %>% 
  select(`Record ID`, `Finger Stick Blood Glucose mg/dl`, `HbA1C value:`, `P1NP Baseline ng/ml`, `CTX Baseline ng/ml`)

# fingerstick bg to numeric and fix "not done" to NA
analysis[ analysis == "not done" ] <- NA
analysis = analysis %>% mutate(fsbg = as.numeric(`Finger Stick Blood Glucose mg/dl`)) %>% select(- `Finger Stick Blood Glucose mg/dl`)

# some hba1c are listed with a % sign (LITERALLY WHY); need to remove that
analysis = analysis %>% mutate(hba1c = 
                                 case_when(str_detect(`HbA1C value:`, "%") ~ as.numeric(str_sub(`HbA1C value:`, 1,str_length(`HbA1C value:`)-1)),
                                           TRUE ~ as.numeric(`HbA1C value:`))) %>% select(-`HbA1C value:`)

# need to fix ctx with the truncated values by dividing by 2
analysis = analysis %>% mutate(ctx_bl = 
                                 case_when(str_detect(`CTX Baseline ng/ml`, "<") ~ (as.numeric(str_sub(`CTX Baseline ng/ml`, 2))/2),
                                           TRUE ~ as.numeric(`CTX Baseline ng/ml`)),
                               p1np_bl = `P1NP Baseline ng/ml`) %>% select(-c(`P1NP Baseline ng/ml`, `CTX Baseline ng/ml`))

# list of subj id's and date for tim
# subjdates = BSTM_t1d %>% filter(`2. Clinical Diagnosis of T1D` == "Yes") %>% 
#   select(`Record ID`, `Date of visit`, `Date of the study visit:`)
# write.csv(subjdates, file = "asbmr_subjdates_cgm.csv")
```

```{r functions}
# function to output scatter, corr and linear mod
fit_mod = function(outcome,df,plot = T,diagnostics = F){
  
    # Fit p1np model
  f = as.formula(paste0("p1np_bl~",outcome))
  mod = lm(f,
            data = df,
            na.action = na.omit)
   cat("\n")
  cat(paste("P1NP Model"))
  cat("\n")
 
  print(summary(mod))
   

  
   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
 
  # Fit ctx model
  f2 = as.formula(paste0("ctx_bl~",outcome))
  mod2 = lm(f,
            data = df,
            na.action = na.omit)
   cat("\n")
  cat(paste("CTX Model"))
  cat("\n")
 
  print(summary(mod2))
   

  
   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod2),"observations."))
  cat("\n") 
  
  # Plots
  if(plot){
    p = 
      ggplot(data = df,aes_string(x = outcome, y = "p1np_bl")) + 
      geom_point(na.rm = T) + geom_smooth(method = lm, se= F, na.rm = T) + 
      theme_bw() 
    print(p)
    
        p2 = 
      ggplot(data = df,aes_string(x = outcome, y = "ctx_bl")) + 
      geom_point(na.rm = T) + geom_smooth(method = lm, se = F,na.rm = T) + 
      theme_bw() 
    print(p2)
  }
 
   # Check model
  if(diagnostics){
    print(check_model(mod))
  }
}

```

# Outstanding Data Questions/ Remarks
Remarks as of 4/27/2022
* Waiting on clean CGM data as of 4/27/2022
* truncated values of ctx (some data listed as "<0.076") were inputted as 0.5 * truncated value
* get a list of subj ids and visit dates for tim for cgm data 


# Methods

Simple linear models  were fit with Fingerstick Glucose, 14 Days CGM metrics, and A1C respectively to assess the linear relationship with with P1NP and CTX. Only subjects with T1D diagnosis (n = 31) were included in analysis.

# Analysis

## Fingerstick Glucose 

```{r fingerstickglucose}
# correlations
f_c1 = cor(analysis$fsbg, analysis$p1np_bl, use = "pairwise.complete.obs")
f_c2 = cor(analysis$fsbg, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "fsbg", df = analysis)

cat("\n")
cat(paste("P1NP Pearson Corr Coef = ",f_c1))
cat(paste("CTX Pearson Corr Coef = ",f_c2))
cat("\n")
```

## 14 Day CGM Metrics

```{r cgm data}
```

## A1C
 
```{r hba1c}
# correlations
h_c1 = cor(analysis$hba1c, analysis$p1np_bl, use = "pairwise.complete.obs")
h_c2 = cor(analysis$hba1c, analysis$ctx_bl, use = "pairwise.complete.obs")

# model and outputs
fit_mod(outcome = "hba1c", df = analysis)

cat("\n")
cat(paste("P1NP Pearson Corr Coef = ",h_c1))
cat(paste("CTX Pearson Corr Coef = ",h_c2))
cat("\n")
```
