---
title: "Gasteroparesis Study"
author: "Kristen Campbell"
date: "March 8, 2019"
output: word_document
editor_options: 
  chunk_output_type: console
---


```{r,echo=FALSE,message=FALSE,results='hide',warning=FALSE}
library(knitr)
library(nlme)
setwd("C:/Users/campbkri/Documents/GitHub/BDC-Code/Viral Shah/Gastroparesis/Code")
source('01_data.R')

##Summary Tables:
table1<-final_table(dat,c('Age','Gender','Years.w.Diabetes','Weight_3mo_pre',
                          'Avg.daily.Basal_3mo_pre','Avg.daily.Bolus_3mo_pre','active.time_3mo_use','Carb.Ratio_3mo_pre'),
                  dat$Group,margin=2,single=F,2, col.names=T, summary.stat='median')
table1

dat.g<-subset(dat,dat$Group=="Gastroparesis")
label(dat.g$duration_of_gastro)<-"Years from Dx of Gastroparesis to 1st CGM placement"

table_g<-final_table(dat.g,c('duration_of_gastro','gci_pre','gci_post'),
                  dat.g$ID,margin=2,single=F,2, col.names=T, summary.stat='median')
table_g

gci.p<-round(wilcox.test(dat.g$gci_pre,dat.g$gci_post,paired=T)$p.value,2)


tab.wear<-final_table(dat,c('total_sensor_readings_1','total_sensor_readings_2','total_sensor_readings_3','total_sensor_readings_overall',
                          'total_days_1','total_days_2','total_days_3','total_days_overall',
                          'avg_total_sensor_per_day_1','avg_total_sensor_per_day_2','avg_total_sensor_per_day_3','avg_total_sensor_per_day_overall'
                            ),
                  dat$Group,margin=2,single=F,2, col.names=T, summary.stat='median')
tab.wear

##Outcome Tables:
#outcome 1 (A1C)
table_a1c<-final_table(dat,c('A1C_3mo_pre','A1C_3mo_after', 'A1C_6mo_post','a1c_mean','a1c_change'),
                  dat$Group,margin=2,single=F,2, col.names=T, summary.stat='median')
table_a1c

a1c.p<-round(wilcox.test(dat$A1C_3mo_pre,dat$A1C_6mo_post,paired=T)$p.value,2)
a1c.p1<-round(wilcox.test(dat$A1C_3mo_pre,dat$A1C_3mo_after,paired=T)$p.value,2)
a1c.p2<-round(wilcox.test(dat$A1C_3mo_after,dat$A1C_6mo_post,paired=T)$p.value,2)

a1c.p.control<-round(wilcox.test(dat$A1C_3mo_pre[dat$Group=="Control"],dat$A1C_6mo_post[dat$Group=="Control"],paired=T)$p.value,2)
a1c.p.control1<-round(wilcox.test(dat$A1C_3mo_pre[dat$Group=="Control"],dat$A1C_3mo_after[dat$Group=="Control"],paired=T)$p.value,2)
a1c.p.control2<-round(wilcox.test(dat$A1C_3mo_after[dat$Group=="Control"],dat$A1C_6mo_post[dat$Group=="Control"],paired=T)$p.value,2)


a1c.p.gastro<-round(wilcox.test(dat$A1C_3mo_pre[dat$Group=="Gastroparesis"],dat$A1C_6mo_post[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
a1c.p.gastro1<-round(wilcox.test(dat$A1C_3mo_pre[dat$Group=="Gastroparesis"],dat$A1C_3mo_after[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
a1c.p.gastro2<-round(wilcox.test(dat$A1C_3mo_after[dat$Group=="Gastroparesis"],dat$A1C_6mo_post[dat$Group=="Gastroparesis"],paired=T)$p.value,2)


#outcome 2 (CGM)
table_cgm<-final_table(dat,c('average_sensor_1','average_sensor_2','average_sensor_3',
                             'daytime_avg_sensor_glucose_1','daytime_avg_sensor_glucose_2',
                             'daytime_avg_sensor_glucose_3','nighttime_avg_sens_glucose_1',
                             'nighttime_avg_sens_glucose_2','nighttime_avg_sens_glucose_3',
                             'cv_1','cv_2','cv_3'),
                  dat$Group,margin=2,single=F,1, col.names=T, summary.stat='median')
table_cgm

table_cgm_strat<-final_table(dat,c('percent_time_under_70_1','percent_time_under_70_2','percent_time_under_70_3',
                             'percent_time_70_180_1','percent_time_70_180_2','percent_time_70_180_3',
                             'percent_time_over_180_1','percent_time_over_180_2','percent_time_over_180_3'),
                  dat$Group,margin=2,single=F,1, col.names=T, summary.stat='median')
table_cgm_strat

table_cgm_changes<-final_table(dat,c('average_sensor_change_12','average_sensor_change_23','average_sensor_change_13',
                             'daytime_avg_sensor_change_12','daytime_avg_sensor_change_23','daytime_avg_sensor_change_13',
                             'nighttime_avg_sensor_change_12','nighttime_avg_sensor_change_23','nighttime_avg_sensor_change_13',
                             'cv_change_12','cv_change_23','cv_change_13'),
                  dat$Group,margin=2,single=F,1, col.names=T, summary.stat='median')
table_cgm_changes


table_cgm_changes_strat<-final_table(dat,
                                     c('percent_time_under_70_change_12','percent_time_under_70_change_23','percent_time_under_70_change_13',
                             'percent_time_70_180_change_12','percent_time_70_180_change_23','percent_time_70_180_change_13',
                             'percent_time_over_180_change_12','percent_time_over_180_change_23','percent_time_over_180_change_13'),
                  dat$Group,margin=2,single=F,1, col.names=T, summary.stat='median')
table_cgm_changes_strat


table_other_changes<-final_table(dat,c('carb_ratio_change', 'active_mean_change','weight_change',
                                       'basal_change','bolus_change'),
                  dat$Group,margin=2,single=F,1, col.names=T, summary.stat='median')
table_other_changes

#####AVERAGE:
#overall
hypo_12.p<-round(wilcox.test(dat$percent_time_under_70_1,dat$percent_time_under_70_2,paired=T)$p.value,4)
hypo_23.p<-round(wilcox.test(dat$percent_time_under_70_2,dat$percent_time_under_70_3,paired=T)$p.value,4)
hypo_13.p<-round(wilcox.test(dat$percent_time_under_70_1,dat$percent_time_under_70_3,paired=T)$p.value,4)
#controls
hypo_12.c.p<-round(wilcox.test(dat$percent_time_under_70_1[dat$Group=="Control"],
                              dat$percent_time_under_70_2[dat$Group=="Control"],paired=T)$p.value,3)
hypo_23.c.p<-round(wilcox.test(dat$percent_time_under_70_2[dat$Group=="Control"],
                              dat$percent_time_under_70_3[dat$Group=="Control"],paired=T)$p.value,3)
hypo_13.c.p<-round(wilcox.test(dat$percent_time_under_70_1[dat$Group=="Control"],
                              dat$percent_time_under_70_3[dat$Group=="Control"],paired=T)$p.value,3)
#gastro
hypo_12.g.p<-round(wilcox.test(dat$percent_time_under_70_1[dat$Group=="Gastroparesis"],
                              dat$percent_time_under_70_2[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
hypo_23.g.p<-round(wilcox.test(dat$percent_time_under_70_2[dat$Group=="Gastroparesis"],
                              dat$percent_time_under_70_3[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
hypo_13.g.p<-round(wilcox.test(dat$percent_time_under_70_1[dat$Group=="Gastroparesis"],
                              dat$percent_time_under_70_3[dat$Group=="Gastroparesis"],paired=T)$p.value,3)

norm_12.p<-round(wilcox.test(dat$percent_time_70_180_1,dat$percent_time_70_180_2,paired=T)$p.value,4)
norm_23.p<-round(wilcox.test(dat$percent_time_70_180_2,dat$percent_time_70_180_3,paired=T)$p.value,4)
norm_13.p<-round(wilcox.test(dat$percent_time_70_180_1,dat$percent_time_70_180_3,paired=T)$p.value,4)
#controls
norm_12.c.p<-round(wilcox.test(dat$percent_time_70_180_1[dat$Group=="Control"],
                              dat$percent_time_70_180_2[dat$Group=="Control"],paired=T)$p.value,3)
norm_23.c.p<-round(wilcox.test(dat$percent_time_70_180_2[dat$Group=="Control"],
                              dat$percent_time_70_180_3[dat$Group=="Control"],paired=T)$p.value,3)
norm_13.c.p<-round(wilcox.test(dat$percent_time_70_180_1[dat$Group=="Control"],
                              dat$percent_time_70_180_3[dat$Group=="Control"],paired=T)$p.value,3)
#gastro
norm_12.g.p<-round(wilcox.test(dat$percent_time_70_180_1[dat$Group=="Gastroparesis"],
                              dat$percent_time_70_180_2[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
norm_23.g.p<-round(wilcox.test(dat$percent_time_70_180_2[dat$Group=="Gastroparesis"],
                              dat$percent_time_70_180_3[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
norm_13.g.p<-round(wilcox.test(dat$percent_time_70_180_1[dat$Group=="Gastroparesis"],
                              dat$percent_time_70_180_3[dat$Group=="Gastroparesis"],paired=T)$p.value,3)

hyper_12.p<-round(wilcox.test(dat$percent_time_over_180_1,dat$percent_time_over_180_2,paired=T)$p.value,4)
hyper_23.p<-round(wilcox.test(dat$percent_time_over_180_2,dat$percent_time_over_180_3,paired=T)$p.value,4)
hyper_13.p<-round(wilcox.test(dat$percent_time_over_180_1,dat$percent_time_over_180_3,paired=T)$p.value,4)
#controls
hyper_12.c.p<-round(wilcox.test(dat$percent_time_over_180_1[dat$Group=="Control"],
                              dat$percent_time_over_180_2[dat$Group=="Control"],paired=T)$p.value,3)
hyper_23.c.p<-round(wilcox.test(dat$percent_time_over_180_2[dat$Group=="Control"],
                              dat$percent_time_over_180_3[dat$Group=="Control"],paired=T)$p.value,3)
hyper_13.c.p<-round(wilcox.test(dat$percent_time_over_180_1[dat$Group=="Control"],
                              dat$percent_time_over_180_3[dat$Group=="Control"],paired=T)$p.value,3)
#gastro
hyper_12.g.p<-round(wilcox.test(dat$percent_time_over_180_1[dat$Group=="Gastroparesis"],
                              dat$percent_time_over_180_2[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
hyper_23.g.p<-round(wilcox.test(dat$percent_time_over_180_2[dat$Group=="Gastroparesis"],
                              dat$percent_time_over_180_3[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
hyper_13.g.p<-round(wilcox.test(dat$percent_time_over_180_1[dat$Group=="Gastroparesis"],
                              dat$percent_time_over_180_3[dat$Group=="Gastroparesis"],paired=T)$p.value,3)
#####SD:
#overall
cv_12.p<-round(wilcox.test(dat$cv_1,dat$cv_2,paired=T)$p.value,4)
cv_23.p<-round(wilcox.test(dat$cv_2,dat$cv_3,paired=T)$p.value,4)
cv_13.p<-round(wilcox.test(dat$cv_1,dat$cv_3,paired=T)$p.value,4)
#controls
cv_12.c.p<-round(wilcox.test(dat$cv_1[dat$Group=="Control"],
                              dat$cv_2[dat$Group=="Control"],paired=T)$p.value,3)
cv_23.c.p<-round(wilcox.test(dat$cv_2[dat$Group=="Control"],
                              dat$cv_3[dat$Group=="Control"],paired=T)$p.value,3)
cv_13.c.p<-round(wilcox.test(dat$cv_1[dat$Group=="Control"],
                              dat$cv_3[dat$Group=="Control"],paired=T)$p.value,3)
#gastro
cv_12.g.p<-round(wilcox.test(dat$cv_1[dat$Group=="Gastroparesis"],
                              dat$cv_2[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
cv_23.g.p<-round(wilcox.test(dat$cv_2[dat$Group=="Gastroparesis"],
                              dat$cv_3[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
cv_13.g.p<-round(wilcox.test(dat$cv_1[dat$Group=="Gastroparesis"],
                              dat$cv_3[dat$Group=="Gastroparesis"],paired=T)$p.value,3)

#####DAYTIME:
#overall
day_12.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_1,dat$daytime_avg_sensor_glucose_2,paired=T)$p.value,4)
day_23.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_2,dat$daytime_avg_sensor_glucose_3,paired=T)$p.value,4)
day_13.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_1,dat$daytime_avg_sensor_glucose_3,paired=T)$p.value,4)
#controls
day_12.c.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_1[dat$Group=="Control"],
                              dat$daytime_avg_sensor_glucose_2[dat$Group=="Control"],paired=T)$p.value,3)
day_23.c.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_2[dat$Group=="Control"],
                              dat$daytime_avg_sensor_glucose_3[dat$Group=="Control"],paired=T)$p.value,3)
day_13.c.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_1[dat$Group=="Control"],
                              dat$daytime_avg_sensor_glucose_3[dat$Group=="Control"],paired=T)$p.value,3)
#gastro
day_12.g.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_1[dat$Group=="Gastroparesis"],
                              dat$daytime_avg_sensor_glucose_2[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
day_23.g.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_2[dat$Group=="Gastroparesis"],
                              dat$daytime_avg_sensor_glucose_3[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
day_13.g.p<-round(wilcox.test(dat$daytime_avg_sensor_glucose_1[dat$Group=="Gastroparesis"],
                              dat$daytime_avg_sensor_glucose_3[dat$Group=="Gastroparesis"],paired=T)$p.value,3)

#####NIGHTTIME:
#overall
night_12.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_1,dat$nighttime_avg_sens_glucose_2,paired=T)$p.value,4)
night_23.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_2,dat$nighttime_avg_sens_glucose_3,paired=T)$p.value,4)
night_13.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_1,dat$nighttime_avg_sens_glucose_3,paired=T)$p.value,4)
#controls
night_12.c.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_1[dat$Group=="Control"],
                              dat$nighttime_avg_sens_glucose_2[dat$Group=="Control"],paired=T)$p.value,3)
night_23.c.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_2[dat$Group=="Control"],
                              dat$nighttime_avg_sens_glucose_3[dat$Group=="Control"],paired=T)$p.value,3)
night_13.c.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_1[dat$Group=="Control"],
                              dat$nighttime_avg_sens_glucose_3[dat$Group=="Control"],paired=T)$p.value,3)
#gastro
night_12.g.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_1[dat$Group=="Gastroparesis"],
                              dat$nighttime_avg_sens_glucose_2[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
night_23.g.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_2[dat$Group=="Gastroparesis"],
                              dat$nighttime_avg_sens_glucose_3[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
night_13.g.p<-round(wilcox.test(dat$nighttime_avg_sens_glucose_1[dat$Group=="Gastroparesis"],
                              dat$nighttime_avg_sens_glucose_3[dat$Group=="Gastroparesis"],paired=T)$p.value,3)


table_cgm_change<-final_table(dat,c('average_sensor_change_12','average_sensor_change_23','average_sensor_change_13'),
                  dat$Group,margin=2,single=F,2, col.names=T, summary.stat='median')
table_cgm_change


##percent time:
avg_12.p<-round(wilcox.test(dat$average_sensor_1,dat$average_sensor_2,paired=T)$p.value,4)
avg_23.p<-round(wilcox.test(dat$average_sensor_2,dat$average_sensor_3,paired=T)$p.value,4)
avg_13.p<-round(wilcox.test(dat$average_sensor_1,dat$average_sensor_3,paired=T)$p.value,4)
#controls
avg_12.c.p<-round(wilcox.test(dat$average_sensor_1[dat$Group=="Control"],
                              dat$average_sensor_2[dat$Group=="Control"],paired=T)$p.value,3)
avg_23.c.p<-round(wilcox.test(dat$average_sensor_2[dat$Group=="Control"],
                              dat$average_sensor_3[dat$Group=="Control"],paired=T)$p.value,3)
avg_13.c.p<-round(wilcox.test(dat$average_sensor_1[dat$Group=="Control"],
                              dat$average_sensor_3[dat$Group=="Control"],paired=T)$p.value,3)
#gastro
avg_12.g.p<-round(wilcox.test(dat$average_sensor_1[dat$Group=="Gastroparesis"],
                              dat$average_sensor_2[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
avg_23.g.p<-round(wilcox.test(dat$average_sensor_2[dat$Group=="Gastroparesis"],
                              dat$average_sensor_3[dat$Group=="Gastroparesis"],paired=T)$p.value,2)
avg_13.g.p<-round(wilcox.test(dat$average_sensor_1[dat$Group=="Gastroparesis"],
                              dat$average_sensor_3[dat$Group=="Gastroparesis"],paired=T)$p.value,3)



####models:
a1c_change<-lm(dat$A1C_6mo_post~dat$Group+dat$Age+dat$A1C_3mo_pre)
summary(a1c_change)
#plot(a1c_change)

a1c_mod<-fx_model_linear('a1c_change',c('Group','Age'),data=dat)

##repeated measures:
results<-lme(average_sensor~factor(Period),random=~1|subject_id,data=dat.long)
summary(results)

results<-lme(average_sensor~factor(Period)*Group,random=~1|subject_id,data=dat.long)
summary(results)

results<-lme(average_sensor~factor(Period)*Group,random=~1|subject_id/Period,data=dat.long)
summary(results)

####actual analysis:
results_avg<-lme(average_sensor~factor(Period)*Group+total_sensor_readings,random=~1|subject_id,data=dat.long)
summary(results_avg)

results_sd<-lme(standard_deviation~factor(Period)*Group+total_sensor_readings,random=~1|subject_id,data=dat.long)
summary(results_sd)

results_day<-lme(daytime_avg_sensor_glucose~factor(Period)*Group+total_sensor_readings,random=~1|subject_id,data=dat.long)
summary(results_day)

results_night<-lme(nighttime_avg_sens_glucose~factor(Period)*Group+total_sensor_readings,random=~1|subject_id,data=dat.long)
summary(results_night)

```

#Statistical Methods

Throughout the results, the three periods will be abbreviated as: 
+ Pre: Months -3 to 0 pre-670G 
+ Post-1: Months 0 to 3 post-670G 
+ Post-2: Months 3 to 6 post-670G 

Demographics and clinical characteristics were summarized using median and interquartile range (IQR) or count and proportion (%).  Changes in variables of interest between any two periods were calculated as differences.  Comparisons between groups were performed using Wilcoxon rank sum tests or Fisher's exact tests, due to the small sample size. Wilcoxon signed rank tests were used to compare changes between two periods within a group.  Linear regression was used to test for an association between the change (pre-670G minus post-670G) in A1C (outcome) with Group (gastroparesis versus control), adjusting for age.
       


#Summary Statistics
**Table 1: Clinical Characteristics, by Group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table1, align = c("l", "l"))
```

Table 1 compares basic demographics and clinical characteristics between the Gastroparesis and Control groups.  The two groups were similar in all variables. 

**Table 2: Clinical Characteristics, Gastroparesis Group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table_g, align = c("l", "l"))
```

The median time from diagnosis of gastroparesis to the 670G placement was `r substr(table_g[1,2],1,3)` years.  There was not a significant difference between GCI scores pre versus post (`r substr(table_g[2,2],1,3)` vs. `r substr(table_g[3,2],1,1)`; p=`r gci.p`).

**Table 3: CGM Amount of Wear**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.wear, align = c("l", "l"))
```

Table 3 describes the amount of CGM data we have for each period and each group.

#Primary Outcome: A1C
**Table 3a: A1C Summary Statistics**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table_a1c, align = c("l", "l"))
```

A1C levels in the control group were slightly higher compared to the gastroparesis group in the Pre period.  There was not a significant difference between groups in A1c levels during Post-2, or in the change between Post-2 versus Pre.

There was a significant difference overall between Pre and Post-2 A1C values (`r substr(table_a1c[1,2],1,4)` vs. `r substr(table_a1c[2,2],1,4)`p=`r a1c.p`).  Within subjects with Gastroparesis, there was not a significant difference between Pre and Post-2 A1C (`r substr(table_a1c[1,3],1,3)` vs. `r substr(table_a1c[2,3],1,3)`; p=`r a1c.p.gastro`).  Within control subjects, there also was not a significant difference between Pre and Post-2 A1C (`r substr(table_a1c[1,4],1,3)` vs. `r substr(table_a1c[2,4],1,3)`;p=`r a1c.p.control`).

**Table 3b: A1C Model: Outcome Pre A1C minus Post-2 A1C**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(a1c_mod, align = c("l", "l"))
```

After adjusting for age, Group was not significantly associated with the change in A1C from Pre to Post-2 (p=`r a1c_mod[1,3]`)

#Secondary Outcome: CGM
**Table 4a: CGM Summary Statistics**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table_cgm, align = c("l", "l"))
```

Table 4a reports the median (IQR) of various CGM summary measurements.  None of the CGM measures showed significant differences between patients with gastroparesis and controls. 

Overall 24-hour mean glucose was significantly different between Pre and Post-1 (p=`r avg_12.p`), as well as between Post-1 and Post-2 (p=`r avg_23.p`).  Within the control group, 24-hour mean glucose was also different between Pre and Post-1 (p=`r avg_12.c.p`), and between Post-1 and Post-2 (p=`r avg_23.c.p`).  Within the gastroparesis group, 24-hour mean glucose measurements did not differ between any period.

Daytime mean glucose was significantly different between Pre and Post-1 (p=`r day_12.p`), and between Post-1 and Post-2 (p=`r day_23.p`). Within the control group, daytime mean glucose was also different between Pre and Post-1 (p=`r day_12.c.p`), and between Post-1 and Post-2 (p=`r day_23.c.p`). Within the gastroparesis group, none of the SD glucose measurements differed between period.

Nighttime mean glucose was significantly different between Pre and Post-1 (p=`r night_12.p`), and between Post-1 and Post-2 (p=`r night_23.p`). Within the control group, daytime mean glucose was also different between Pre and Post-1 (p=`r night_12.c.p`), and between Post-1 and Post-2 (p=`r night_23.c.p`). Within the gastroparesis group, none of the SD glucose measurements differed between period.

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table_cgm_strat, align = c("l", "l"))
```

**Table 4b: CGM Changes**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table_cgm_changes, align = c("l", "l"))
```

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table_cgm_changes_strat, align = c("l", "l"))
```

**Table 5: Other Changes**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(table_other_changes, align = c("l", "l"))
```


