---
title: "Introduction to Joint Modeling"
subtitle: "For Pyle ADA Talk"
author: "Kristen Miller"
date: "3-24-2026"
date-format: "MMMM D, YYYY"
format:
  revealjs:
    theme: white
    slide-number: c/t
    progress: true
    chalkboard: false
    transition: fade
    background-transition: fade
    embed-resources: true
    highlight-style: github
    code-fold: false
    code-line-numbers: true
    fig-width: 7
    fig-height: 4
    fig-align: center
    footer: "Pyle ADA Talk"
    logo: "uwmdi_logo.jpg"
    toc: false
    smaller: false
    scrollable: false
    incremental: false
    css: intro-biostatistics.css
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
set.seed(42)
```

# Part I: Foundations {background-color="#028090"}

## What is joint modeling?

::: {.fragment}

<div class="stat-box">
A statistical model of two (or more) related outcomes, most often a longitudinal and a time-to-event outcome 
</div>

:::

::: {.columns}
::: {.column width="50%"}

::: {.fragment}
<div class="stat-box" style="margin-top:0.7em;">
<strong>Longitudinal outcome</strong><br>
is tracked over time in the same subjects, also called a <em>repeated measure</em>
</div>

:::

::: {.fragment}
```{r pop-sample-diagram, echo=FALSE, fig.height=4}
# Install and load necessary packages if you haven't already
# install.packages("ggplot2")
# install.packages("lme4")
library(ggplot2)
library(lme4) # Contains the sleepstudy dataset

# Load the example longitudinal data
data(sleepstudy)

# Create a spaghetti plot with an overall trend line
ggplot(data = sleepstudy, aes(x = Days, y = Reaction, group = Subject)) +
  geom_line(color = "gray", alpha = 0.5) + # Individual lines (lightly colored)
  geom_point(color = "gray", alpha = 0.5) + # Individual points (optional)
  geom_smooth(
    aes(group = 1),
    method = "lm",
    color = "blue",
    linewidth = 1.5,
    se = TRUE
  ) + # Overall linear trend
  labs(
    title = "Longitudinal Plot of Reaction Time over Days",
    x = "Days of Sleep Deprivation",
    y = "Reaction Time (ms)"
  ) +
  theme_minimal()
```

:::
::: 

::: {.column width="50%"}

::: {.fragment}
<div class="stat-box" style="margin-top:0.7em;">
<strong>Time-to-event outcome</strong><br>
measures the length of time until an occurrence of an event, also called a <em>survival outcome</em>
</div>

:::

::: {.fragment}

```{r pop-sample-diagram2, echo=FALSE, fig.height=4}
library(survival)
library(survminer)
surv_object_all <- Surv(time = lung$time, event = lung$status == 2)
fit_all <- survfit(surv_object_all ~ 1, data = lung)
# Plot the curve using ggsurvplot for publication-ready plots
ggsurvplot(
  fit_all,
  data = lung,
  title = "Survival Probability of Lung Cancer Patients",
  conf.int = TRUE, # Add confidence intervals
  risk.table = F, # Add the number-at-risk table
  ggtheme = theme_light(), # Use a clean theme,
  legend = "none"
)
```
:::
:::
:::

## Why joint modeling?

::: {.fragment}
<div class="stat-box" style="margin-top:0.7em;">
<strong>Analyzing these outcomes separately introduces bias:</strong><br>

### Problem 1: Informative Dropout
Patients who experience the event early tend to have worse longitudinal trajectories. Ignoring this creates **selection bias**.

### Problem 2: Measurement Error
Using observed (error-prone) biomarker values as time-varying covariates in a Cox model leads to **attenuation bias** — underestimating the true association.

### Problem 3: Efficiency
Separate models discard information. A joint model **borrows strength** across both submodels, improving parameter estimation efficiency.
</div>

:::
:::
:::

## Model Specification

A joint model has two linked submodels:

### Longitudinal Submodel
$$
Y_i(t) = m_i(t) + \varepsilon_i(t), \quad \varepsilon_i(t) \sim \mathcal{N}(0, \sigma^2)
$$
$$
m_i(t) = \mathbf{x}_i^\top(t)\boldsymbol{\beta} + \mathbf{z}_i^\top(t)\mathbf{b}_i, \quad \mathbf{b}_i \sim \mathcal{N}(\mathbf{0}, \mathbf{D})
$$

### Time-to-Event Submodel
$$
h_i(t) = h_0(t) \exp\!\left[\boldsymbol{\gamma}^\top \mathbf{w}_i + \alpha \, m_i(t)\right]
$$

### The Association Parameter
The parameter $\alpha$ **links** the two submodels — it captures the effect of the true (latent) longitudinal trajectory on the hazard of the event.
:::
:::
:::

## Joint modeling in diabetes research

::: {.fragment}

### Example 1:
Do longitudinal trajectories of HbA1c associate with diabetic retinopathy? (https://doi.org/10.1038/s41598-022-06240-5)

### Example 2:
Does drug response predict the risk of side effects in patients starting T2D therapy? (https://doi.org/10.2147/CLEP.S179555)

### Example 3:
Can daily glucose measures predict preterm birth in women with T1D? (https://doi.org/10.1155/2020/3074532)

</div>
