---
title: "ZnT8 Trasnporter Haploinsufficiency"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(performance)
library(car)
library(emmeans)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

home_dir = ifelse(.Platform$OS.type != "unix","B:/Projects/Howard Davidson/Zinc transporter/Data raw",
                  "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
knitr::opts_knit$set(root.dir = home_dir)

```

```{r, include=FALSE}
# read in data
fig3 <- read.csv("./Fig 3 Calcuim Oscillation.csv")
fig3 <- data.frame(t(fig3))
colnames(fig3) <- "Calcium.oscillation"
fig3$genotype <- ifelse(grepl("WT",rownames(fig3)),"WT",
                        ifelse(grepl("KO",rownames(fig3)),"KO","HE"))
fig3$treatment <- ifelse(grepl("Non",rownames(fig3)),"Non-treatment","Treatment")

fig4_atp <- read.csv("./Fig 4 ATP production.csv")
fig4_atp <- data.frame(t(fig4_atp))
colnames(fig4_atp) <- "ATP.production"
fig4_atp$genotype <- ifelse(grepl("WT",rownames(fig4_atp)),"WT",
                        ifelse(grepl("KO",rownames(fig4_atp)),"KO","HE"))
fig4_atp$treatment <- ifelse(grepl("Non",rownames(fig4_atp)),"Non-treatment","Treatment")

fig4_basalresp <- read.csv("./Fig 4 Basal respiration.csv")
fig4_basalresp <- data.frame(t(fig4_basalresp))
colnames(fig4_basalresp) <- "Basal.respiration"
fig4_basalresp$genotype <- ifelse(grepl("WT",rownames(fig4_basalresp)),"WT",
                        ifelse(grepl("KO",rownames(fig4_basalresp)),"KO","HE"))
fig4_basalresp$treatment <- ifelse(grepl("Non",rownames(fig4_basalresp)),"Non-treatment","Treatment")

fig4_maxresp <- read.csv("./Fig 4 Maximal respiration.csv")
fig4_maxresp <- data.frame(t(fig4_maxresp))
colnames(fig4_maxresp) <- "Max.respiration"
fig4_maxresp$genotype <- ifelse(grepl("WT",rownames(fig4_maxresp)),"WT",
                        ifelse(grepl("KO",rownames(fig4_maxresp)),"KO","HE"))
fig4_maxresp$treatment <- ifelse(grepl("Non",rownames(fig4_maxresp)),"Non-treatment","Treatment")

fig4_protonleak <- read.csv("./Fig 4 Proton leak.csv")
fig4_protonleak <- data.frame(t(fig4_protonleak))
colnames(fig4_protonleak) <- "Proton.leak"
fig4_protonleak$genotype <- ifelse(grepl("WT",rownames(fig4_protonleak)),"WT",
                        ifelse(grepl("KO",rownames(fig4_protonleak)),"KO","HE"))
fig4_protonleak$treatment <- ifelse(grepl("Non",rownames(fig4_protonleak)),"Non-treatment","Treatment")

fig2d2f <- read.csv("./Figure 2D~2F merge all genotypes.csv")
fig2d2f <- data.frame(t(fig2d2f))
colnames(fig2d2f) <- "GSIS"
fig2d2f$genotype <- ifelse(grepl("WT",rownames(fig2d2f)),"WT",
                        ifelse(grepl("KO",rownames(fig2d2f)),"KO","HE"))
fig2d2f$treatment <- ifelse(grepl("Non",rownames(fig2d2f)),"Non-treatment","Treatment")
fig2d2f$log_gsis <- log(fig2d2f$GSIS)

supfig3_atp <- read.csv("./Sup fig 3 ATP production.csv")
supfig3_atp <- data.frame(t(supfig3_atp))
colnames(supfig3_atp) <- "ATP.production"
supfig3_atp$genotype <- ifelse(grepl("WT",rownames(supfig3_atp)),"WT",
                        ifelse(grepl("KO",rownames(supfig3_atp)),"KO","HE"))
supfig3_atp$treatment <- ifelse(grepl("Non",rownames(supfig3_atp)),"Non-treatment","Treatment")

supfig3_basalresp <- read.csv("./Sup fig 3 Basal respiration.csv")
supfig3_basalresp <- data.frame(t(supfig3_basalresp))
colnames(supfig3_basalresp) <- "Basal.respiration"
supfig3_basalresp$genotype <- ifelse(grepl("WT",rownames(supfig3_basalresp)),"WT",
                        ifelse(grepl("KO",rownames(supfig3_basalresp)),"KO","HE"))
supfig3_basalresp$treatment <- ifelse(grepl("Non",rownames(supfig3_basalresp)),"Non-treatment","Treatment")

supfig3_maxresp <- read.csv("./Sup fig 3 Maximal respiration.csv")
supfig3_maxresp <- data.frame(t(supfig3_maxresp))
colnames(supfig3_maxresp) <- "Max.respiration"
supfig3_maxresp$genotype <- ifelse(grepl("WT",rownames(supfig3_maxresp)),"WT",
                        ifelse(grepl("KO",rownames(supfig3_maxresp)),"KO","HE"))
supfig3_maxresp$treatment <- ifelse(grepl("Non",rownames(supfig3_maxresp)),"Non-treatment","Treatment")

supfig3_proton <- read.csv("./Sup fig 3 Proton leak.csv")
supfig3_proton <- data.frame(t(supfig3_proton))
colnames(supfig3_proton) <- "Proton.leak"
supfig3_proton$genotype <- ifelse(grepl("WT",rownames(supfig3_proton)),"WT",
                        ifelse(grepl("KO",rownames(supfig3_proton)),"KO","HE"))
supfig3_proton$treatment <- ifelse(grepl("Non",rownames(supfig3_proton)),"Non-treatment","Treatment")

supfig_inscontent <- read.csv("./Sup fig Insulin Content.csv")
supfig_inscontent <- data.frame(t(supfig_inscontent))
colnames(supfig_inscontent) <- "Insulin.content"
supfig_inscontent$genotype <- ifelse(grepl("WT",rownames(supfig_inscontent)),"WT",
                        ifelse(grepl("KO",rownames(supfig_inscontent)),"KO","HE"))
supfig_inscontent$treatment <- ifelse(grepl("Non",rownames(supfig_inscontent)),"Non-treatment","Treatment")

supfig_insratio <- read.csv("./Sup fig Insulin ratio.csv")
supfig_insratio <- data.frame(t(supfig_insratio))
colnames(supfig_insratio) <- "Insulin.ratio"
supfig_insratio$genotype <- ifelse(grepl("WT",rownames(supfig_insratio)),"WT",
                        ifelse(grepl("KO",rownames(supfig_insratio)),"KO","HE"))
supfig_insratio$treatment <- ifelse(grepl("Non",rownames(supfig_insratio)),"Non-treatment","Treatment")

supfig_insulin <- read.csv("./Sup fig Insulin.csv")
supfig_insulin <- data.frame(t(supfig_insulin))
colnames(supfig_insulin) <- "Insulin"
supfig_insulin$genotype <- ifelse(grepl("WT",rownames(supfig_insulin)),"WT",
                        ifelse(grepl("KO",rownames(supfig_insulin)),"KO","HE"))
supfig_insulin$treatment <- ifelse(grepl("Non",rownames(supfig_insulin)),"Non-treatment","Treatment")

supfig_proinsulin <- read.csv("./Sup fig Proinsulin.csv")
supfig_proinsulin <- data.frame(t(supfig_proinsulin))
colnames(supfig_proinsulin) <- "Proinsulin"
supfig_proinsulin$genotype <- ifelse(grepl("WT",rownames(supfig_proinsulin)),"WT",
                        ifelse(grepl("KO",rownames(supfig_proinsulin)),"KO","HE"))
supfig_proinsulin$treatment <- ifelse(grepl("Non",rownames(supfig_proinsulin)),"Non-treatment","Treatment")

# function for model with interaction
mod_int <- function(data, outcome) {
  df = data
  form = as.formula(paste0(outcome,"~ genotype*treatment"))
  mod <- lm(form, data=df)
  print(check_model(mod))
  print(Anova(mod,type = "III"))
  means <- emmeans(mod, pairwise ~ treatment|genotype)
  print(means)
  p <- emmip(mod, treatment ~ genotype)
  print(p)
}

mod_noint <- function(data, outcome) {
  df = data
  form = as.formula(paste0(outcome,"~ genotype + treatment"))
  mod <- lm(form, data=df)
  print(check_model(mod))
  print(Anova(mod,type = "III"))
  means <- emmeans(mod, pairwise ~ treatment)
  print(means)
  means_g <- emmeans(mod, pairwise ~ genotype)
  print(means_g)
}
```

# How to interpret these results

- For each outcome, there are two sets of models.  The first contains an interaction term for genotype and treatment.  The second model does not contain the interaction term.  If the interaction term is not significant, it's better to use the model without the interaction term to interpret the effects of treatment and genotype.

- For each model, the first set of output is a set of diagnostic plots to evaluate the model fit.  All the models had good fit, with the exception of the GSIS outcome in Figure 2. There appear to be a lot of values close to zero in that variable, so it is not normally distributed.  There are models that could be a better fit to this type of data, but they are harder to interpret.  A simple log transformation did not improve the fit.

- The next part of the output is the ANOVA table.  The column labeled "Pr(>F)" gives the p-value for each term in the model.  The term "genotype:treatment" is the interaction between genotype and treatment.

- The next part of the output provides the means for each group, and the p-values for the pairwise comparison of groups.  For models with an interaction term, groups are formed by crossing treatment and genotype.  

# Figure 2d-f

## GSIS

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=fig2d2f, outcome="GSIS")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=fig2d2f, outcome="GSIS")
```

# Figure 3

## Calcium oscillation

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=fig3, outcome="Calcium.oscillation")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=fig3, outcome="Calcium.oscillation")
```

# Figure 4

## ATP production

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=fig4_atp, outcome="ATP.production")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=fig4_atp, outcome="ATP.production")
```

## Basal respiration

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=fig4_basalresp, outcome="Basal.respiration")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=fig4_basalresp, outcome="Basal.respiration")
```

## Maximal respiration

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=fig4_maxresp, outcome="Max.respiration")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=fig4_maxresp, outcome="Max.respiration")
```

## Proton leak

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=fig4_protonleak, outcome="Proton.leak")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=fig4_protonleak, outcome="Proton.leak")
```

# Supplemental figure 3

## ATP production

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig3_atp, outcome="ATP.production")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig3_atp, outcome="ATP.production")
```

## Basal respiration

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig3_basalresp, outcome="Basal.respiration")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig3_basalresp, outcome="Basal.respiration")
```

## Maximal respiration

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig3_maxresp, outcome="Max.respiration")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig3_maxresp, outcome="Max.respiration")
```

## Proton leak

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig3_proton, outcome="Proton.leak")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig3_proton, outcome="Proton.leak")
```

# Supplemental figure proinsulin:insulin ratio

## Insulin content

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig_inscontent, outcome="Insulin.content")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig_inscontent, outcome="Insulin.content")
```

## Insulin ratio

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig_insratio, outcome="Insulin.ratio")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig_insratio, outcome="Insulin.ratio")
```

## Insulin

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig_insulin, outcome="Insulin")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig_insulin, outcome="Insulin")
```

## Proinsulin

### Model with interaction

```{r echo=FALSE, comment=""}
mod_int(data=supfig_proinsulin, outcome="Proinsulin")
```

### Model without interaction

```{r echo=FALSE, comment=""}
mod_noint(data=supfig_proinsulin, outcome="Proinsulin")
```
