---
title: "Advanced CGM and OGTT: Joint Models"
author: "Tim Vigers"
date: "now"
date-format: "MMMM D, YYYY [at] hh:mm a"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    embed-resources: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(plotly)
library(nlme)
library(JMbayes2)
library(ggsurvfit)
library(survminer)
library(knitr)
library(gtsummary)
library(parallel)
base_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/home/tvigers/Documents/Data",
  "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Tims-Mac-Studio.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Mac-Studio" = "/Users/lpyle/Library/CloudStorage/OneDrive-SharedLibraries-UW",
  "MacBook-Pro-51.local" = "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW(2)",
  "twigs" = "/home/tim/Documents/Data"
)
data_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/BDC/Andrea Steck/Advanced CGM and OGTT",
  "Tims-MacBook-Air.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Tims-Mac-mini.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Tims-Mac-Studio.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Mac" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Mac-Studio" = "/Tim Vigers - UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "MacBook-Pro-51.local" = "/Tim Vigers - UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "twigs" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT"
)
home_dir <- paste0(base_dir, data_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
setwd(home_dir)
# Load data
# See BDC-Code/Andrea Steck/Advanced CGM and OGTT/create_analysis_dataset.R
load("./Data_Clean/analysis_dataset.RData")
# Filter people >= age 30
cgm_lmm = cgm_lmm |> filter(Age < 30)
# Scale and center the variables of interest
# cgm_lmm$mean_glucose <- scale(cgm_lmm$mean_glucose)
# cgm_lmm$sd_glucose <- scale(cgm_lmm$sd_glucose)
# cgm_lmm$cv_glucose <- scale(cgm_lmm$cv_glucose)
# cgm_lmm$perc_time_over_140 <- scale(cgm_lmm$perc_time_over_140)
# cgm_lmm$hba1c <- scale(cgm_lmm$hba1c)
# Limit survival group to those included in mixed model DF
cgm_surv = cgm_surv |> filter(ID %in% cgm_lmm$ID)
# JM options
jm_control <- list(
  n_iter = 1e5,
  n_burnin = 1e3,
  n_thin = 2,
  n_chains = 4,
  cores = 4
)
# Make a simplified dataset for Andrea to pick participants for examples
newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |>
  group_by(ID) |>
  select(ID, Group, Age, AgeEndpoint) |>
  summarise(
    BaselineAge = round(first(Age), 1),
    AgeEndpoint = round(unique(AgeEndpoint), 1),
    Group = unique(Group),
    VisitsWithCGM = n()
  )
# Save analysis datasets
write.csv(
  newdat,
  file = "./Data_Clean/simplified_data.csv",
  row.names = FALSE,
  na = ""
)
write.csv(
  cgm_lmm,
  file = "./Data_Clean/cgm_data.csv",
  row.names = FALSE,
  na = ""
)
write.csv(
  cgm_surv,
  file = "./Data_Clean/survival_data.csv",
  row.names = FALSE,
  na = ""
)
```

# Participant characteristics

```{r}
cgm_surv |>
  select(-ID, -event) |>
  tbl_summary(
    by = Group,
    statistic = list(
      AgeEndpoint ~ "{median} ({min}, {max})"
    )
  ) |>
  add_p() |>
  add_overall() |>
  separate_p_footnotes()
```

# CGM metrics summarized across the whole day

```{r}
#| fig-cap: Mean glucose over time
mean_plot <- ggplot(
  cgm_lmm,
  aes(x = Age, y = mean_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Mean Glucose (mg/dL)") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(mean_plot)
```

```{r}
#| fig-cap: Glucose SD over time
sd_plot <- ggplot(
  cgm_lmm,
  aes(x = Age, y = sd_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Glucose SD (mg/dL)") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(sd_plot)
```

```{r}
survfit2(
  Surv(AgeEndpoint, event) ~ 1,
  data = cgm_surv
) |>
  ggsurvfit() +
  labs(
    x = "Age (years)",
    y = "Overall survival probability"
  ) +
  add_confidence_interval()
```

## Cox model

For the survival submodel, AIC was compared between a Cox proportional hazards model and Weibull, exponential, gaussian, logistic,lognormal and log logistic models. The Cox PH model had the lowest AIC and visual inspection of model residuals suggested good model fit. AIC and concordance was also better for the Cox model with age as the time variable compared to the Cox model with time from first CGM as the time variable.

We also compared the Cox model adjusted for sex, race/ethnicity, FDR status, and maximum AB group to a simplified model adjusted for sex only. Model performance was very similar, but the simplified model was better by BIC (479.2 vs. 488.0).

When debating whether to use age or time from first CGM as the time variable, AIC for the Cox model was better when using age but AIC was better for the linear model using time from first CGM. We decided that clinical usefulness should be the tiebreaker and therefore continued with age as the time variable in all models.

```{r}
# Cox model
cox_fit <- coxph(
  Surv(AgeEndpoint, event) ~ sex +
    Race_Ethn2 +
    screen_FDR_GP +
    maxAB_group,
  x = T,
  model = T,
  data = cgm_surv
)
cox_fit_simple <- coxph(
  Surv(AgeEndpoint, event) ~ sex,
  x = T,
  model = T,
  data = cgm_surv
)
BIC(cox_fit, cox_fit_simple)
cox_fit <- coxph(
  Surv(AgeEndpoint, event) ~ sex,
  x = T,
  model = T,
  data = cgm_surv
)
```

### Results

```{r}
summary(cox_fit)
```

### Diagnostics

```{r}
ggcoxdiagnostics(cox_fit, type = "martingale", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "deviance", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "score", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "schoenfeld", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "partial", linear.predictions = F)
```

## Linear model results

Models with random slope and random intercept for participant were better than those with just a random intercept by AIC. The linear model using years from first CGM as the time variable was better than the model using age as the time variable by AIC.

```{r}
lme_fit_mean_ri <- lme(
  mean_glucose ~ Age,
  random = ~ 1 | ID,
  data = cgm_lmm
)
lme_fit_mean_rs <- lme(
  mean_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
AIC(lme_fit_mean_ri, lme_fit_mean_rs)
plot_df <- lme_fit_mean_rs$data
plot_df$pred <- predict(lme_fit_mean_rs)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = mean_glucose), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
# Final model
lme_fit_mean <- lme(
  mean_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
```

# Joint model results

## Mean glucose

```{r}
lme_fit_mean <- lme(
  mean_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
performance::check_model(lme_fit_mean)
plot_df <- lme_fit_mean$data
plot_df$pred <- predict(lme_fit_mean)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = mean_glucose), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
joint_fit_mean <- jm(
  cox_fit,
  lme_fit_mean,
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_mean)
ggtraceplot(joint_fit_mean, parm = "betas")
```

## Glucose SD

```{r}
lme_fit_sd <- lme(
  sd_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
performance::check_model(lme_fit_sd)
plot_df <- lme_fit_sd$data
plot_df$pred <- predict(lme_fit_sd)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = sd_glucose), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
joint_fit_sd <- jm(
  cox_fit,
  lme_fit_sd,
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_sd)
ggtraceplot(joint_fit_sd, parm = "betas")
```

## Glucose CV

```{r}
lme_fit_cv <- lme(
  cv_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
performance::check_model(lme_fit_cv)
plot_df <- lme_fit_cv$data
plot_df$pred <- predict(lme_fit_cv)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = cv_glucose), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
joint_fit_cv <- jm(
  cox_fit,
  lme_fit_cv,
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_cv)
ggtraceplot(joint_fit_cv, parm = "betas")
```

## TAR140

```{r}
lme_fit_tar <- lme(
  perc_time_over_140 ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
performance::check_model(lme_fit_tar)
plot_df <- lme_fit_tar$data
plot_df$pred <- predict(lme_fit_tar)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = perc_time_over_140), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
joint_fit_tar <- jm(
  cox_fit,
  lme_fit_tar,
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_tar)
ggtraceplot(joint_fit_tar, parm = "betas")
```

## HbA1c

```{r}
lme_fit_hba1c <- lme(
  hba1c ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
performance::check_model(lme_fit_hba1c)
plot_df <- lme_fit_hba1c$data
plot_df$pred <- predict(lme_fit_hba1c)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = hba1c), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
joint_fit_hba1c <- jm(
  cox_fit,
  lme_fit_hba1c,
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_hba1c)
ggtraceplot(joint_fit_hba1c, parm = "betas")
```

## HbA1c and Mean

```{r}
joint_fit_hba1c_mean <- jm(
  cox_fit,
  list(lme_fit_hba1c, lme_fit_mean),
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_hba1c_mean)
ggtraceplot(joint_fit_hba1c_mean, parm = "betas")
```

## HbA1c and SD

```{r}
joint_fit_hba1c_sd <- jm(
  cox_fit,
  list(lme_fit_hba1c, lme_fit_sd),
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_hba1c_sd)
ggtraceplot(joint_fit_hba1c_sd, parm = "betas")
```

## HbA1c and CV

```{r}
joint_fit_hba1c_cv <- jm(
  cox_fit,
  list(lme_fit_hba1c, lme_fit_cv),
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_hba1c_cv)
ggtraceplot(joint_fit_hba1c_cv, parm = "betas")
```

## HbA1c and TAR140

```{r}
joint_fit_hba1c_tar <- jm(
  cox_fit,
  list(lme_fit_hba1c, lme_fit_tar),
  time_var = "Age",
  id_var = "ID",
  control = jm_control
)
summary(joint_fit_hba1c_tar)
ggtraceplot(joint_fit_hba1c_tar, parm = "betas")
```

# Dynamic predictions

## HbA1c and CV

```{r}
# Progressors
newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "01-00375")
predLongHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  process = "event",
  times = seq(15, 18, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized CV"),
  xlab = "Age",
  main = "01-00375 (Progressed at Age 15.7)"
)

newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "04-02273")
predLongHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  process = "event",
  times = seq(10, 13, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized CV"),
  xlab = "Age",
  main = "04-02273 (Progressed at Age 10.4)"
)

newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "39390-3")
predLongHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  process = "event",
  times = seq(21.5, 24.5, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized CV"),
  xlab = "Age",
  main = "39390-3 (Progressed at Age 21.7)"
)
# Non-Progressors
newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "01-00149")
predLongHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  process = "event",
  times = seq(19, 22, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized CV"),
  xlab = "Age",
  main = "01-00149 (Did Not Progress)"
)

newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "03-00456")
predLongHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  process = "event",
  times = seq(21, 24, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized CV"),
  xlab = "Age",
  main = "03-00456 (Did Not Progress)"
)


newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "98-00003")
predLongHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_cv,
  newdata = newdat,
  process = "event",
  times = seq(19.5, 22.5, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized CV"),
  xlab = "Age",
  main = "98-00003 (Did Not Progress)"
)
```

## HbA1c and TA140

```{r}
newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "01-00375")
predLongHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  process = "event",
  times = seq(15, 18, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized TA140"),
  xlab = "Age",
  main = "01-00375 (Progressed at Age 15.7)"
)


newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "04-02273")
predLongHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  process = "event",
  times = seq(10, 13, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized TA140"),
  xlab = "Age",
  main = "04-02273 (Progressed at Age 10.4)"
)


newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "39390-3")
predLongHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  process = "event",
  times = seq(21.5, 24.5, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized TA140"),
  xlab = "Age",
  main = "39390-3 (Progressed at Age 21.7)"
)
# Non-Progressors
newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "01-00149")
predLongHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  process = "event",
  times = seq(19, 22, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized TA140"),
  xlab = "Age",
  main = "01-00149 (Did Not Progress)"
)


newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "03-00456")
predLongHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  process = "event",
  times = seq(21, 24, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized TA140"),
  xlab = "Age",
  main = "03-00456 (Did Not Progress)"
)


newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |> filter(ID == "98-00003")
predLongHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  return_newdata = TRUE,
  process = "longitudinal",
  type = "mean_subject"
)
predSurvHbA1c <- predict(
  joint_fit_hba1c_tar,
  newdata = newdat,
  process = "event",
  times = seq(19.5, 22.5, length.out = 100),
  return_newdata = TRUE
)
plot(
  predLongHbA1c,
  predSurvHbA1c,
  outcomes = 1:2,
  ylab_event = "Progression Probability",
  ylab_long = c("Normalized HbA1c", "Normalized TA140"),
  xlab = "Age",
  main = "98-00003 (Did Not Progress)"
)
```

### AUC table

```{r}
# Make a list of all the models we've looked at
models <- list(
  joint_fit_mean,
  joint_fit_sd,
  joint_fit_cv,
  joint_fit_tar,
  joint_fit_hba1c,
  joint_fit_hba1c_mean,
  joint_fit_hba1c_sd,
  joint_fit_hba1c_cv,
  joint_fit_hba1c_tar
)
names(models) <- c(
  "Mean",
  "SD",
  "CV",
  "TA140",
  "HbA1c",
  "HbA1c + Mean",
  "HbA1c + SD",
  "HbA1c + CV",
  "HbA1c + TA140"
)
# Make a grid for AUC testing
auc_grid <- expand.grid(
  start = seq(6, 20, by = 1),
  dt = 1:3
)
# Get AUC at various time ranges for each model
newdat <- full_join(cgm_lmm, cgm_surv)
auc_grid_results <- apply(auc_grid, 1, function(r) {
  start <- as.numeric(r["start"])
  dt <- as.numeric(r["dt"])
  auc_results <- sapply(names(models), function(n) {
    tv <- tryCatch(
      tvROC(
        models[[n]],
        newdata = newdat,
        Tstart = start,
        Dt = dt
      ),
      error = function(e) {
        return(NA)
      }
    )
    if (length(tv) == 1) {
      return(NA)
    } else {
      return(tvAUC(tv)$auc)
    }
  })
  res <- data.frame(auc_results) |>
    rownames_to_column(var = "Model") |>
    rename(AUC = auc_results) |>
    mutate(up_to = start, window = dt)
  return(res)
})

t <- do.call(rbind, auc_grid_results)


auc_grid_results <- auc_grid_results |>
  pivot_wider(
    values_from = AUC,
    names_from = window
  ) |>
  group_split(up_to, .keep = F)
names(auc_grid_results) <- unique(auc_grid$start)
invisible(lapply(names(auc_grid_results), function(n) {
  cat("\n")
  cat("\n")
  cat("#### Using data up to", n, "years")
  cat("\n")
  cat("\n")
  print(kable(auc_grid_results[[n]]))
  cat("\n")
  cat("\n")
}))
```

# Combined predictions via super-learning

```{r}
#| eval: false
cv_dats <- create_folds(cgm_lmm, V = 5, id_var = "ID")
# A function that will fit all the above models on the CV data.
fit_models <- function(fold) {
  # Cox model
  cox_fit <- coxph(
    Surv(AgeEndpoint, event) ~ sex +
      Race_Ethn2 +
      screen_FDR_GP +
      maxAB_group,
    x = T,
    model = T,
    data = cgm_surv[cgm_surv$ID %in% fold$ID, ]
  )
  # Mixed models
  print("Mean")
  lme_fit_mean <- lme(
    mean_glucose ~ Age,
    random = ~ Age | ID,
    data = fold,
    control = list(opt = "optim", msMaxIter = 1000)
  )
  print("SD")
  lme_fit_sd <- lme(
    sd_glucose ~ Age,
    random = ~ Age | ID,
    data = fold,
    control = list(opt = "optim", msMaxIter = 1000)
  )
  print("CV")
  lme_fit_cv <- lme(
    cv_glucose ~ Age,
    random = ~ Age | ID,
    data = fold,
    control = list(opt = "optim", msMaxIter = 1000)
  )
  print("TAR")
  lme_fit_tar <- lme(
    perc_time_over_140 ~ Age,
    random = ~ Age | ID,
    data = fold,
    control = list(opt = "optim", msMaxIter = 1000)
  )
  print("HbA1c")
  lme_fit_hba1c <- lme(
    hba1c ~ Age,
    random = ~ Age | ID,
    data = fold,
    control = list(opt = "optim", msMaxIter = 1000)
  )
  # Joint models
  joint_fit_mean <- jm(
    cox_fit,
    lme_fit_mean,
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_sd <- jm(
    cox_fit,
    lme_fit_sd,
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_cv <- jm(
    cox_fit,
    lme_fit_cv,
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_tar <- jm(
    cox_fit,
    lme_fit_tar,
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_hba1c <- jm(
    cox_fit,
    lme_fit_hba1c,
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_hba1c_mean <- jm(
    cox_fit,
    list(lme_fit_hba1c, lme_fit_mean),
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_hba1c_sd <- jm(
    cox_fit,
    list(lme_fit_hba1c, lme_fit_sd),
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_hba1c_cv <- jm(
    cox_fit,
    list(lme_fit_hba1c, lme_fit_cv),
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  joint_fit_hba1c_tar <- jm(
    cox_fit,
    list(lme_fit_hba1c, lme_fit_tar),
    time_var = "Age",
    id_var = "ID",
    control = jm_control
  )
  # Format output
  out <- list(
    Mean = joint_fit_mean,
    SD = joint_fit_sd,
    CV = joint_fit_cv,
    TAR = joint_fit_tar,
    HbA1c = joint_fit_hba1c,
    HbA1c_Mean = joint_fit_hba1c_mean,
    HbA1c_SD = joint_fit_hba1c_sd,
    HbA1c_CV = joint_fit_hba1c_cv,
    HbA1c_TAR = joint_fit_hba1c_tar
  )
  class(out) <- "jmList"
  return(out)
}
models_folds <- lapply(cv_dats$training, fit_models)
```
