---
title: "Advanced CGM and OGTT: Joint Models"
author: "Tim Vigers"
date: "now"
date-format: "MMMM D, YYYY [at] hh:mm a"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    embed-resources: true
    fig-cap-location: top
    page-layout: full
editor: source
engine: knitr
---

```{r setup}
#| include: false
library(tidyverse)
library(plotly)
library(nlme)
library(JMbayes2)
library(ggsurvfit)
library(survminer)
library(knitr)
library(gtsummary)
library(parallel)
library(reticulate)
base_dir <- switch(Sys.info()["nodename"],
  "togo" = "/home/tvigers/Documents/Data",
  "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Tims-Mac-Studio.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Mac-Studio" = "/Users/lpyle/Library/CloudStorage/OneDrive-SharedLibraries-UW",
  "MacBook-Pro-51.local" = "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW(2)",
  "Kristens-MacBook-Pro.local" = "/Users/kristenmiller/Library/CloudStorage/OneDrive-UW",
  "twigs" = "/home/tim/Documents/Data"
)
data_dir <- switch(Sys.info()["nodename"],
  "togo" = "/BDC/Andrea Steck/Advanced CGM and OGTT",
  "Tims-MacBook-Air.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Tims-Mac-mini.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Tims-Mac-Studio.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Mac" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Mac-Studio" = "/Tim Vigers - UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "MacBook-Pro-51.local" = "/Tim Vigers - UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Kristens-MacBook-Pro.local" = "/Tim Vigers's files - UWMDI/Andrea Steck",
  "twigs" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT"
)
home_dir <- paste0(base_dir, data_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
setwd(home_dir)
set.seed(1017)
# Load data
# See BDC-Code/Andrea Steck/Advanced CGM and OGTT/create_analysis_dataset.R
load("./Data_Clean/analysis_dataset.RData")
# Filter people >= age 30 and those who progressed at their first CGM visit.
# We don't know for sure that the CGM wear happened prior to progression, so
# these people may not be in the risk set.
no_risk <- cgm_surv$ID[which(cgm_surv$EndTime == 0 & cgm_surv$event == 1)]
cgm_lmm <- cgm_lmm |> filter(Age < 30, !ID %in% no_risk)
# Remove TRIALNET
# cgm_lmm <- cgm_lmm |> filter(study != "TRIALNET")
# Scale and center the variables of interest
# cgm_lmm$mean_glucose <- scale(cgm_lmm$mean_glucose)
# cgm_lmm$sd_glucose <- scale(cgm_lmm$sd_glucose)
# cgm_lmm$cv_glucose <- scale(cgm_lmm$cv_glucose)
# cgm_lmm$perc_time_over_140 <- scale(cgm_lmm$perc_time_over_140)
# cgm_lmm$hba1c <- scale(cgm_lmm$hba1c)
# Limit survival group to those included in mixed model DF
cgm_surv <- cgm_surv |> filter(ID %in% cgm_lmm$ID)
# Make a simplified dataset for Andrea to pick participants for examples
newdat <- full_join(cgm_lmm, cgm_surv)
newdat <- newdat |>
  group_by(ID) |>
  select(ID, Group, Age, AgeEndpoint) |>
  summarise(
    BaselineAge = round(first(Age), 1),
    AgeEndpoint = round(unique(AgeEndpoint), 1),
    Group = unique(Group),
    VisitsWithCGM = n()
  )
# Save analysis datasets
write.csv(
  newdat,
  file = "./Data_Clean/simplified_data.csv",
  row.names = FALSE,
  na = ""
)
write.csv(
  cgm_lmm,
  file = "./Data_Clean/cgm_data.csv",
  row.names = FALSE,
  na = ""
)
write.csv(
  cgm_surv,
  file = "./Data_Clean/survival_data.csv",
  row.names = FALSE,
  na = ""
)
# Test peak finding on first CGM
cgm_test <- split.data.frame(cgm, f = ~ ID + dov_CGM, drop = TRUE)
cgm_test <- cgm_test[[1]]
# Calculate minutes from first glucose
cgm_test$mins <- ymd_hms(paste(cgm_test$Date, cgm_test$Time))
cgm_test$mins <- as.numeric(cgm_test$mins - cgm_test$mins[1]) / 60
```

```{python}
import numpy as np
import pandas as pd
from scipy.signal import find_peaks
from scipy.signal import peak_widths

# Import cgm data from R
df = r.cgm_test
# Get glucose and time columns
glucose = df["SensorValue"].values
time = df["mins"].values
# Sampling interval (assumes uniform spacing)
dt = np.median(np.diff(time))  # minutes per sample
samples_per_2h = int(2 * 60 / dt)  # minimum distance between peaks
# Find peaks
peaks, props = find_peaks(
    glucose,
    height=60,  # peak height <= 60 mg/dL
    distance=samples_per_2h,  # ≥ 2 h between peaks
)
peaks, props
# Filter for duration >= 30 min
# Measure width at the level of peak 50 (i.e., where glucose is 50 mg/dL below peak)
# rel_height=1.0 means at the base (prominence reference level)
# Adjust rel_height to control where width is measured
widths, width_heights, left_ips, right_ips = peak_widths(glucose, peaks, rel_height=1)
left_ips, right_ips
# Convert width to hours
widths_mins = widths * dt
# Keep only peaks where elevated duration >= 30
duration_mask = widths_mins >= 30
peaks_filtered = peaks[duration_mask]
widths_filtered = widths_mins[duration_mask]
props_filtered = {k: v[duration_mask] for k, v in props.items()}
# Assemble results
peak_df = pd.DataFrame(
    {
        "peak_time_min": time[peaks_filtered],
        "peak_glucose": glucose[peaks_filtered],
        "duration_min": widths_filtered,
        "peak_idx": peaks_filtered,
    }
)
```

```{r}
peak_df <- py$peak_df
```

# Participant characteristics

```{r}
cgm_surv |>
  select(-ID, -event) |>
  tbl_summary(
    by = Group,
    statistic = list(
      AgeEndpoint ~ "{median} ({min}, {max})"
    )
  ) |>
  add_p() |>
  add_overall() |>
  separate_p_footnotes()
```

# CGM metrics summarized across the whole day

```{r}
#| fig-cap: Mean glucose over time
mean_plot <- ggplot(
  cgm_lmm,
  aes(x = Age, y = mean_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Mean Glucose (mg/dL)") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(mean_plot)
```

```{r}
#| fig-cap: Glucose SD over time
sd_plot <- ggplot(
  cgm_lmm,
  aes(x = Age, y = sd_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Glucose SD (mg/dL)") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(sd_plot)
```

```{r}
survfit2(
  Surv(AgeEndpoint, event) ~ 1,
  data = cgm_surv
) |>
  ggsurvfit() +
  labs(
    x = "Age (years)",
    y = "Overall survival probability"
  ) +
  add_confidence_interval()
```

## Cox model

For the survival submodel, AIC was compared between a Cox proportional hazards model and Weibull, exponential, gaussian, logistic,lognormal and log logistic models. The Cox PH model had the lowest AIC and visual inspection of model residuals suggested good model fit. AIC and concordance was also better for the Cox model with age as the time variable compared to the Cox model with time from first CGM as the time variable.

We also compared the Cox model adjusted for sex, race/ethnicity, FDR status, and maximum AB group to a simplified model adjusted for sex only. Model performance was very similar, but the simplified model was better by BIC (479.2 vs. 488.0).

When debating whether to use age or time from first CGM as the time variable, AIC for the Cox model was better when using age but AIC was better for the linear model using time from first CGM. We decided that clinical usefulness should be the tiebreaker and therefore continued with age as the time variable in all models.

```{r}
# Cox model
cox_fit <- coxph(
  Surv(AgeEndpoint, event) ~ sex +
    Race_Ethn2 +
    screen_FDR_GP +
    maxAB_group,
  x = T,
  model = T,
  data = cgm_surv
)
cox_fit_simple <- coxph(
  Surv(AgeEndpoint, event) ~ sex,
  x = T,
  model = T,
  data = cgm_surv
)
BIC(cox_fit, cox_fit_simple)
cox_fit <- coxph(
  Surv(AgeEndpoint, event) ~ sex,
  x = T,
  model = T,
  data = cgm_surv
)
```

### Results

```{r}
summary(cox_fit)
```

### Diagnostics

```{r}
ggcoxdiagnostics(cox_fit, type = "martingale", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "deviance", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "score", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "schoenfeld", linear.predictions = F)
ggcoxdiagnostics(cox_fit, type = "partial", linear.predictions = F)
```

## Linear model results

Models with random slope and random intercept for participant were better than those with just a random intercept by BIC. The linear model using years from first CGM as the time variable was better than the model using age as the time variable by BIC.

```{r}
lme_fit_mean_ri <- lme(
  mean_glucose ~ Age,
  random = ~ 1 | ID,
  data = cgm_lmm
)
lme_fit_mean_rs <- lme(
  mean_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
BIC(lme_fit_mean_ri, lme_fit_mean_rs)
plot_df <- lme_fit_mean_rs$data
plot_df$pred <- predict(lme_fit_mean_rs)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = mean_glucose), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
# Final model
lme_fit_mean <- lme(
  mean_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
```

# Joint model results

```{r}
# There are too many models, so create a function to make sure each outcome is
# treated the same. For now this function only handles two longitudinal
# outcomes.
joint_fits <- function(
  outcomes,
  time = "Age",
  cox_adjust = c("sex"),
  lmm_adjust = NULL,
  cox_df = cgm_surv,
  lmm_df = cgm_lmm,
  jm_control = list(
    n_iter = 1e5,
    n_burnin = 2e3,
    n_thin = 8,
    n_chains = 4,
    cores = 4
  )
) {
  # Right now it may be unnecessary to re-fit the Cox model each time, but
  # put it in the function just in case.
  cox_f <- as.formula(paste0(
    "Surv(",
    time,
    "Endpoint, event)",
    "~",
    paste(cox_adjust, collapse = "+")
  ))
  cox_fit <- coxph(
    cox_f,
    x = T,
    model = T,
    data = cox_df
  )
  # Fit a random intercept and slope model for each outcome
  re_f <- as.formula(paste0("~", time, "|", "ID"))
  lme_mods <- lapply(outcomes, function(outcome) {
    lme_f <- as.formula(paste0(
      outcome,
      "~",
      paste0(c(time, lmm_adjust), collapse = "+")
    ))
    lme_fit <- lme(
      lme_f,
      random = re_f,
      data = cgm_lmm
    )
    return(lme_fit)
  })
  names(lme_mods) <- outcomes
  # For each outcome, fit the three different functional forms
  joint_mods <- lapply(outcomes, function(outcome) {
    joint_fit_value <- jm(
      cox_fit,
      lme_mods[[outcome]],
      time_var = time,
      id_var = "ID",
      control = jm_control
    )
    slope_ff <- list(as.formula(paste0("~", "slope(", outcome, ")")))
    names(slope_ff) <- outcome
    joint_fit_slope <- jm(
      cox_fit,
      lme_mods[[outcome]],
      time_var = time,
      id_var = "ID",
      control = jm_control,
      functional_forms = slope_ff
    )
    area_ff <- list(as.formula(paste0("~", "area(", outcome, ")")))
    names(area_ff) <- outcome
    joint_fit_area <- jm(
      cox_fit,
      lme_mods[[outcome]],
      time_var = time,
      id_var = "ID",
      control = jm_control,
      functional_forms = area_ff
    )
    return(list(
      "value" = joint_fit_value,
      "slope" = joint_fit_slope,
      "area" = joint_fit_area
    ))
  })
  names(joint_mods) <- outcomes
  # If only one outcome, just return all three
  if (length(outcomes) == 1) {
    joint_mods <- joint_mods[[1]]
    names(joint_mods) <- c("value", "slope", "area")
    return(joint_mods)
  } else if (length(outcomes) > 1) {
    # If multiple outcomes, do a lil' model selection.
    # Marginal DIC/WAIC is better because it evaluates the overall model fit
    # for the population by integrating out the random effects
    optimal_mods <- sapply(outcomes, function(outcome) {
      # I'm sure there's a more elegant way to do this but it works
      c("value", "slope", "area")[which.min(c(
        joint_mods[[outcome]][["value"]][["fit_stats"]][["marginal"]][["DIC"]],
        joint_mods[[outcome]][["slope"]][["fit_stats"]][["marginal"]][["DIC"]],
        joint_mods[[outcome]][["area"]][["fit_stats"]][["marginal"]][["DIC"]]
      ))]
    })
    # Compare models with the same functional forms for all outcomes to a model
    # with the optimal form for each outcome separately. This is pretty
    # inefficient because it fits all 3 funcitional form models for all outcomes
    # but I'd rather be consistent than efficient at this point.
    # First a model with value for both.
    joint_fit_value <- jm(
      cox_fit,
      lme_mods,
      time_var = time,
      id_var = "ID",
      control = jm_control
    )
    # Slope for both
    slope_ff <- sapply(outcomes, function(outcome) {
      as.formula(paste0("~", "slope(", outcome, ")"))
    })
    names(slope_ff) <- outcomes
    joint_fit_slope <- jm(
      cox_fit,
      lme_mods,
      time_var = time,
      id_var = "ID",
      control = jm_control,
      functional_forms = slope_ff
    )
    # Area for both
    area_ff <- sapply(outcomes, function(outcome) {
      as.formula(paste0("~", "area(", outcome, ")"))
    })
    names(area_ff) <- outcomes
    joint_fit_area <- jm(
      cox_fit,
      lme_mods,
      time_var = time,
      id_var = "ID",
      control = jm_control,
      functional_forms = area_ff
    )
    # With the optimal forms from above
    optimal_ff <- sapply(outcomes, function(outcome) {
      as.formula(paste0("~", optimal_mods[outcome], "(", outcome, ")"))
    })
    names(optimal_ff) <- outcomes
    joint_fit_optimal <- jm(
      cox_fit,
      lme_mods,
      time_var = time,
      id_var = "ID",
      control = jm_control,
      functional_forms = optimal_ff
    )
    # Return them all
    return(list(
      "value" = joint_fit_value,
      "slope" = joint_fit_slope,
      "area" = joint_fit_area,
      "optimal" = joint_fit_optimal
    ))
  }
}
```

## Mean glucose

```{r}
joint_models <- joint_fits(outcomes = "mean_glucose")
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(joint_models$value, joint_models$slope, joint_models$area)
```

```{r}
# Final model
mean_jm <- joint_models$value
```

## Glucose SD

```{r}
joint_models <- joint_fits(outcomes = "sd_glucose")
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(joint_models$value, joint_models$slope, joint_models$area)
```

```{r}
# Final model
sd_jm <- joint_models$value
```

## Glucose CV

```{r}
joint_models <- joint_fits(outcomes = "cv_glucose")
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(joint_models$value, joint_models$slope, joint_models$area)
```

```{r}
# Final model
cv_jm <- joint_models$area
```

## TAR140

```{r}
joint_models <- joint_fits(outcomes = "perc_time_over_140")
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(joint_models$value, joint_models$slope, joint_models$area)
```

```{r}
# Final model
tar_jm <- joint_models$value
```

## HbA1c

```{r}
joint_models <- joint_fits(outcomes = "hba1c")
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(joint_models$value, joint_models$slope, joint_models$area)
```

```{r}
# Final model
a1c_jm <- joint_models$area
```

## HbA1c and Mean

```{r}
joint_models <- joint_fits(outcomes = c("hba1c", "mean_glucose"))
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form: optimal

```{r}
summary(joint_models$optimal)
ggtraceplot(joint_models$optimal, parm = "alphas")
ggtraceplot(joint_models$optimal, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(
  joint_models$value,
  joint_models$slope,
  joint_models$area,
  joint_models$optimal
)
```

```{r}
# Final model
mean_a1c_jm <- joint_models$optimal
```

## HbA1c and SD

```{r}
joint_models <- joint_fits(outcomes = c("hba1c", "sd_glucose"))
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form: optimal

```{r}
summary(joint_models$optimal)
ggtraceplot(joint_models$optimal, parm = "alphas")
ggtraceplot(joint_models$optimal, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(
  joint_models$value,
  joint_models$slope,
  joint_models$area,
  joint_models$optimal
)
```

```{r}
# Final model
sd_a1c_jm <- joint_models$optimal
```

## HbA1c and CV

```{r}
joint_models <- joint_fits(outcomes = c("hba1c", "cv_glucose"))
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form: optimal

```{r}
summary(joint_models$optimal)
ggtraceplot(joint_models$optimal, parm = "alphas")
ggtraceplot(joint_models$optimal, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(
  joint_models$value,
  joint_models$slope,
  joint_models$area,
  joint_models$optimal
)
```

```{r}
# Final model
cv_a1c_jm <- joint_models$optimal
```

## HbA1c and TAR140

```{r}
joint_models <- joint_fits(outcomes = c("hba1c", "perc_time_over_140"))
```

### Functional form: value

```{r}
summary(joint_models$value)
ggtraceplot(joint_models$value, parm = "alphas")
ggtraceplot(joint_models$value, parm = "betas")
```

### Functional form: slope

```{r}
summary(joint_models$slope)
ggtraceplot(joint_models$slope, parm = "alphas")
ggtraceplot(joint_models$slope, parm = "betas")
```

### Functional form: area

```{r}
summary(joint_models$area)
ggtraceplot(joint_models$area, parm = "alphas")
ggtraceplot(joint_models$area, parm = "betas")
```

### Functional form: optimal

```{r}
summary(joint_models$optimal)
ggtraceplot(joint_models$optimal, parm = "alphas")
ggtraceplot(joint_models$optimal, parm = "betas")
```

### Functional form comparisons

```{r}
compare_jm(
  joint_models$value,
  joint_models$slope,
  joint_models$area,
  joint_models$optimal
)
```

```{r}
# Final model
tar_a1c_jm <- joint_models$optimal
```

# Dynamic predictions

```{r}
new_dat <- full_join(cgm_lmm, cgm_surv)
# Make predictions a function as well
dynamic_prediction <- function(df = new_dat, jm, id, dt = 1, ylabs, pos_ylab) {
  # Just one subject for now
  pred_dat <- df |> filter(ID == id)
  # Check group
  prog <- as.character(unique(pred_dat$Group)) == "Progressor"
  # Make a plot title
  if (prog) {
    plot_title <- paste(c(id, " (Progressed at Age ", round(unique(pred_dat$AgeEndpoint), 1), ")"), collapse = "")
  } else if (!prog) {
    plot_title <- paste(c(id, " (Did Not Progress)"), collapse = "")
  }
  # Set up times
  t0 <- unique(pred_dat$AgeEndpoint) - dt
  # Remove everything after t0, pretend we don't know if they progressed
  pred_dat <- pred_dat |>
    filter(Age <= t0) |>
    mutate(AgeEndpoint = max(Age), event = 0)
  # Predict!
  pred_long <- predict(
    jm,
    newdata = pred_dat,
    return_newdata = TRUE,
    process = "longitudinal",
    type = "mean_subject"
  )
  pred_surv <- predict(
    jm,
    newdata = pred_dat,
    process = "event",
    times = seq(t0, t0 + 1, length.out = 100),
    type_pred = "response",
    return_newdata = TRUE
  )
  plot(
    pred_long,
    pred_surv,
    outcomes = 1:2,
    xlab = "Age",
    ylab_long = ylabs,
    pos_ylab_long = pos_ylab,
    main = plot_title
  )
}
```

## HbA1c and CV

### Model results

```{r}
summary(cv_a1c_jm)
```

### Plots

```{r}
# Progressors
id <- "01-00375"
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and CV/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "04-02273"
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and CV/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "39390-3"
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and CV/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()
# Non-Progressors
id <- "01-00149"
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and CV/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "03-00456"
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and CV/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "98-00003"
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and CV/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = cv_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()
```

## HbA1c and TA140

### Model results

```{r}
summary(tar_a1c_jm)
```

### Plots

```{r}
# Progressors
id <- "01-00375"
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and TA140/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "04-02273"
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and TA140/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "39390-3"
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and TA140/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()
# Non-Progressors
id <- "01-00149"
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and TA140/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "03-00456"
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and TA140/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()

id <- "98-00003"
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
png(paste0(c("./Figures/HbA1c and TA140/", id, ".png"), collapse = ""), width = 2400, height = 1800, res = 300)
dynamic_prediction(id = id, jm = tar_a1c_jm, ylabs = c("HbA1c", "Glucose CV"), pos_ylab = c(5.5, 0.3))
dev.off()
```

### AUC table

```{r}
# Make a list of all the models we've looked at
models <- list(
  mean_jm,
  sd_jm,
  cv_jm,
  tar_jm,
  a1c_jm,
  mean_a1c_jm,
  sd_a1c_jm,
  cv_a1c_jm,
  tar_a1c_jm
)
names(models) <- c(
  "Mean",
  "SD",
  "CV",
  "TA140",
  "HbA1c",
  "HbA1c + Mean",
  "HbA1c + SD",
  "HbA1c + CV",
  "HbA1c + TA140"
)
# Make a grid for AUC testing
auc_grid <- expand.grid(
  start = seq(6, 29, by = 1),
  dt = 1:3
)
# Get AUC at various time ranges for each model
auc_grid_results <- apply(auc_grid, 1, function(r) {
  start <- as.numeric(r["start"])
  dt <- as.numeric(r["dt"])
  auc_results <- sapply(names(models), function(n) {
    tv <- tryCatch(
      tvROC(
        models[[n]],
        newdata = newdat,
        Tstart = start,
        Dt = dt
      ),
      error = function(e) {
        return(NA)
      }
    )
    if (length(tv) == 1) {
      return(NA)
    } else {
      return(tvAUC(tv)$auc)
    }
  })
  res <- data.frame(auc_results) |>
    rownames_to_column(var = "Model") |>
    rename(AUC = auc_results) |>
    mutate(up_to = start, window = dt)
  return(res)
})

t <- do.call(rbind, auc_grid_results)


auc_grid_results <- auc_grid_results |>
  pivot_wider(
    values_from = AUC,
    names_from = window
  ) |>
  group_split(up_to, .keep = F)
names(auc_grid_results) <- unique(auc_grid$start)
invisible(lapply(names(auc_grid_results), function(n) {
  cat("\n")
  cat("\n")
  cat("#### Using data up to", n, "years")
  cat("\n")
  cat("\n")
  print(kable(auc_grid_results[[n]]))
  cat("\n")
  cat("\n")
}))
```
