---
title: "Advanced CGM and OGTT: Joint Models"
author: "Tim Vigers"
date: "now"
date-format: "MMMM D, YYYY [at] hh:mm a"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(plotly)
library(nlme)
library(JMbayes2)
library(ggsurvfit)
library(survminer)
library(knitr)
base_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/home/tvigers/Documents/Data",
  "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-UW"
)
data_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/BDC/Andrea Steck/Advanced CGM and OGTT",
  "Tims-MacBook-Air.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Tims-Mac-mini.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Mac" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT"
)
home_dir <- paste0(base_dir, data_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
setwd(home_dir)
# Load data
# See BDC-Code/Andrea Steck/Advanced CGM and OGTT/create_analysis_dataset.R
load("./Data_Clean/analysis_dataset.RData")
# JM options
jm_control = list(n_thin = 2, n_chains = 4, cores = 4)
```

# Across the whole day

```{r}
#| fig-cap: Mean glucose over time
mean_plot <- ggplot(
  cgm_lmm,
  aes(x = Age, y = mean_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Mean Glucose (mg/dL)") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(mean_plot)
mean_plot <- ggplot(
  cgm_lmm,
  aes(x = TimeFromFirstCGM, y = mean_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Mean Glucose (mg/dL)") +
  xlab("Years From First CGM") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(mean_plot)
```

```{r}
#| fig-cap: Glucose SD over time
sd_plot <- ggplot(
  cgm_lmm,
  aes(x = Age, y = sd_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Glucose SD (mg/dL)") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(sd_plot)
sd_plot <- ggplot(
  cgm_lmm,
  aes(x = TimeFromFirstCGM, y = sd_glucose, color = ID)
) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.5) +
  ylab("Glucose SD (mg/dL)") +
  xlab("Years From First CGM") +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(sd_plot)
```

```{r}
survfit2(
  Surv(AgeEndpoint, event) ~ 1,
  data = cgm_surv
) |>
  ggsurvfit() +
  labs(
    x = "Age (years)",
    y = "Overall survival probability"
  ) +
  add_confidence_interval()
survfit2(
  Surv(EndTime, event) ~ 1,
  data = cgm_surv
) |>
  ggsurvfit() +
  labs(
    x = "Years From First CGM",
    y = "Overall survival probability"
  ) +
  add_confidence_interval()
```

## Cox model

For the survival submodel, AIC was compared between a Cox proportional hazards model and Weibull, exponential, gaussian, logistic,lognormal and log logistic models. The Cox PH model had the lowest AIC and visual inspection of model residuals suggested good model fit. AIC and concordance was also better for the Cox model with age as the time variable compared to the Cox model with time from first CGM as the time variable.

```{r}
# Cox model
cox_fit_age <- coxph(
  Surv(AgeEndpoint, event) ~ sex +
    Race_Ethn2 +
    screen_FDR_GP +
    maxAB_group,
  x = T,
  model = T,
  data = cgm_surv
)
cox_fit_time <- coxph(
  Surv(EndTime, event) ~ sex +
    Race_Ethn2 +
    screen_FDR_GP +
    maxAB_group,
  x = T,
  model = T,
  data = cgm_surv
)
AIC(cox_fit_age, cox_fit_time)
```

### Results

```{r}
summary(cox_fit_age)
summary(cox_fit_time)
```

### Diagnostics

#### Age as time variable

```{r}
ggcoxdiagnostics(cox_fit_age, type = "martingale", linear.predictions = F)
ggcoxdiagnostics(cox_fit_age, type = "deviance", linear.predictions = F)
ggcoxdiagnostics(cox_fit_age, type = "score", linear.predictions = F)
ggcoxdiagnostics(cox_fit_age, type = "schoenfeld", linear.predictions = F)
ggcoxdiagnostics(cox_fit_age, type = "partial", linear.predictions = F)
```

#### Years from first CGM as time variable

```{r}
ggcoxdiagnostics(cox_fit_time, type = "martingale", linear.predictions = F)
ggcoxdiagnostics(cox_fit_time, type = "deviance", linear.predictions = F)
ggcoxdiagnostics(cox_fit_time, type = "score", linear.predictions = F)
ggcoxdiagnostics(cox_fit_time, type = "schoenfeld", linear.predictions = F)
ggcoxdiagnostics(cox_fit_time, type = "partial", linear.predictions = F)
```

## Mean glucose

### Linear model results

Models with random slope and random intercept for participant were better than those with just a random intercept by AIC. The linear model using years from first CGM as the time variable was better than the model using age as the time variable by AIC.

```{r}
lme_fit_mean_rs_age <- lme(
  mean_glucose ~ Age,
  random = ~ Age | ID,
  data = cgm_lmm
)
lme_fit_mean_rs_time <- lme(
  mean_glucose ~ TimeFromFirstCGM,
  random = ~ TimeFromFirstCGM | ID,
  data = cgm_lmm
)
plot_df = lme_fit_mean_rs_age$data
plot_df$pred = predict(lme_fit_mean_rs_age)
ggplot(plot_df, aes(x = Age, color = ID)) +
  geom_point(aes(y = mean_glucose), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
plot_df = lme_fit_mean_rs_time$data
plot_df$pred = predict(lme_fit_mean_rs_time)
ggplot(plot_df, aes(x = TimeFromFirstCGM, color = ID)) +
  geom_point(aes(y = mean_glucose), alpha = 0.2) +
  theme(legend.position = "none") +
  geom_line(aes(y = pred))
```

### Linear model diagnostics

#### Age as time variable

```{r}
performance::check_model(lme_fit_mean_rs_age)
```

#### Years from first CGM as time variable

```{r}
performance::check_model(lme_fit_mean_rs_time)
```

### Joint model results

#### Age as time variable

```{r}
# Fit
joint_fit_mean_age <- jm(
  cox_fit_age,
  lme_fit_mean_rs_age,
  time_var = "Age",
  id_var = "ID",
  n_iter = 1e5,
  n_burnin = 1e3,
  control = jm_control
)
traceplot(joint_fit_mean_age)
summary(joint_fit_mean_age)
```

#### Years from first CGM as time variable

```{r}
joint_fit_mean_time <- jm(
  cox_fit_time,
  lme_fit_mean_rs_time,
  time_var = "TimeFromFirstCGM",
  id_var = "ID",
  n_iter = 1e5,
  n_burnin = 1e3,
  control = jm_control
)
traceplot(joint_fit_mean_time)
summary(joint_fit_mean_time)
```

```{r}
compare_jm(joint_fit_mean_age, joint_fit_mean_time)
```


```{r}
newdat = full_join(cgm_lmm, cgm_surv)
roc_age = tvROC(
  joint_fit_mean_age,
  newdata = newdat,
  Tstart = quantile(cgm_surv$AgeEndpoint, 0.75),
  Dt = 1
)
tvAUC(roc_age)
roc_time = tvROC(
  joint_fit_mean_time,
  newdata = newdat,
  Tstart = quantile(cgm_surv$EndTime, 0.75),
  Dt = 1
)
tvAUC(roc_time)
```