---
title: "Advanced CGM and OGTT: ROC Cross-Validation"
author: "Tim Vigers"
date: "now"
date-format: "MMMM D, YYYY [at] hh:mm a"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    embed-resources: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(caret)
library(pROC)
library(haven)
library(gtsummary)
library(knitr)
library(kableExtra)
base_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/home/tvigers/Documents/Data",
  "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-UW",
  "Mac-Studio" = "/Users/lpyle/Library/CloudStorage/OneDrive-SharedLibraries-UW",
  "MacBook-Pro-51.local" = "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW(2)"
)
data_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/BDC/Andrea Steck/Advanced CGM and OGTT",
  "Tims-MacBook-Air.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Tims-Mac-mini.local" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Mac" = "/UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "Mac-Studio" = "/Tim Vigers - UWMDI/Andrea Steck/Advanced CGM and OGTT",
  "MacBook-Pro-51.local" = "/Tim Vigers - UWMDI/Andrea Steck/Advanced CGM and OGTT"
)
home_dir <- paste0(base_dir, data_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

# Participants with CGM and OGTT (n = 152)

```{r}
# Set seed
set.seed(1017)
setwd(home_dir)
# Import data from Fran
df = read_sas("./Data_Raw/cgm_ogtt_a1c_bl_gt4days.sas7bdat")
stage3 = read_sas("./Data_Raw/stage3outcome152.sas7bdat")
df = left_join(df, stage3, by = join_by(ID))
# Create outcome variables
df$stage3case = factor(
  df$stage3case,
  levels = 0:1,
  labels = c("control", "case")
)
df$Case_T1D24 = factor(
  df$Case_T1D24,
  levels = 0:1,
  labels = c("control", "case")
)
# ROC analysis function
roc_analysis = function(
  data = df,
  outcome,
  predictor,
  metrics = c(
    "threshold",
    "sensitivity",
    "specificity",
    "ppv",
    "npv"
  )
) {
  # Formula
  f = as.formula(paste0(outcome, "~", predictor))
  # Create ROC curve
  r = roc(f, data = data, quiet = TRUE)
  # Plot
  auc = auc(r)
  best = coords(r, "best", best.method = "youden")
  cat("\n")
  cat("\n")
  cat("#### ROC Curve")
  cat("\n")
  cat("\n")
  plot.roc(r, print.thres = TRUE, print.auc = TRUE)
  cat("\n")
  cat("\n")
  # Bootstrap the threshold that maximizes the Youden index
  boots = data.frame(ci.coords(
    r,
    "best",
    best.method = "youden",
    best.policy = "random",
    ret = metrics,
    conf.level = 0.95,
    boot.n = 2000,
    boot.stratified = TRUE
  ))
  colnames(boots) = sub("\\.", "_", colnames(boots))
  boots = boots |>
    pivot_longer(
      everything(),
      names_to = c("Metric", ".value"),
      names_pattern = "(.*)_(.*)"
    )
  colnames(boots) = c("Metric", "2.5%ile", "Median", "97.5%ile")
  cat("\n")
  cat("\n")
  cat("#### Bootstrapped Youden Threshold")
  cat("\n")
  cat("\n")
  print(kable(boots, digits = 3))
  # All thresholds
  cat("\n")
  cat("\n")
  cat("#### All Thresholds")
  cat("\n")
  cat("\n")
  print(kable(coords(r, ret = metrics), digits = 3))
  cat("\n")
  cat("\n")
}
```

## Data Summary

```{r}
df |>
  select(-ID, -FirstID, -TESS_Study, -DateGiven, -where(is.Date), -starttime) |>
  tbl_summary()
```

## T1D

### HbA1c

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "A1C")
```

### Mean SG

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "avgsensorvalue")
```

### TA140

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "pt_above140")
```

### TA160

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "pt_above160")
```

### SD

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "stdevsensorvalue")
```

### CV

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "CVsensorvalue")
```

### MAGE

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "MAGE")
```

### MODD

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "MODD")
```

### Fasting glucose

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "Glu0")
```

### 2hr glucose 

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "Glu120")
```

### Glucose AUC

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "auc_glu")
```

### Fasting C-peptide

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "C_Peptide0")
```

### 2hr C-peptide

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "C_Peptide120")
```

### C-peptide AUC

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "auc_c_peptide")
```

### Peak C-peptide

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "max_c_peptide")
```

## Stage 3

### Mean SG

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "avgsensorvalue")
```

### TA140

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "pt_above140")
```

### TA160

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "pt_above160")
```

### SD

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "stdevsensorvalue")
```

### CV

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "CVsensorvalue")
```

### MAGE

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "MAGE")
```

### MODD

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "MODD")
```

### Fasting glucose

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "Glu0")
```

### 2hr glucose 

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "Glu120")
```

### Glucose AUC

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "auc_glu")
```

### Fasting C-peptide

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "C_Peptide0")
```

### 2hr C-peptide

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "C_Peptide120")
```

### C-peptide AUC

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "auc_c_peptide")
```

### Peak C-peptide

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "max_c_peptide")
```

# Larger group (n = 222)

```{r}
# Import data from Fran
df = read_sas("./Data_Raw/cgm_a1c_bl_gt4days2.sas7bdat")
# Create outcome variables
df$stage3case = df$stage3
df$stage3case[is.na(df$stage3case)] = 0
df$stage3case = factor(
  df$stage3case,
  levels = 0:1,
  labels = c("control", "case")
)
df$Case_T1D24 = factor(
  df$Case_T1D24,
  levels = 0:1,
  labels = c("control", "case")
)
```

## Data Summary

```{r}
df |>
  select(-ID, -FirstID, -TESS_Study, -DateGiven, -where(is.Date), -starttime) |>
  tbl_summary()
```

## T1D

### HbA1c

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "A1C")
```

### Mean SG

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "avgsensorvalue")
```

### TA140

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "pt_above140")
```

### TA160

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "pt_above160")
```

### SD

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "stdevsensorvalue")
```

### CV

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "CVsensorvalue")
```

### MAGE

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "MAGE")
```

### MODD

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "MODD")
```

### Fasting glucose

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "Glu0")
```

### 2hr glucose 

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "Glu120")
```

### Glucose AUC

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "auc_glu")
```

### Fasting C-peptide

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "C_Peptide0")
```

### 2hr C-peptide

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "C_Peptide120")
```

### C-peptide AUC

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "auc_c_peptide")
```

### Peak C-peptide

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "max_c_peptide")
```

## Stage 3

### HbA1c

```{r results="asis"}
roc_analysis(outcome = "Case_T1D24", predictor = "A1C")
```

### Mean SG

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "avgsensorvalue")
```

### TA140

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "pt_above140")
```

### TA160

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "pt_above160")
```

### SD

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "stdevsensorvalue")
```

### CV

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "CVsensorvalue")
```

### MAGE

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "MAGE")
```

### MODD

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "MODD")
```

### Fasting glucose

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "Glu0")
```

### 2hr glucose 

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "Glu120")
```

### Glucose AUC

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "auc_glu")
```

### Fasting C-peptide

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "C_Peptide0")
```

### 2hr C-peptide

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "C_Peptide120")
```

### C-peptide AUC

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "auc_c_peptide")
```

### Peak C-peptide

```{r results="asis"}
roc_analysis(outcome = "stage3case", predictor = "max_c_peptide")
```