---
title: "Trabecular Bone Samples T1DM"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(ggplot2)
library(performance)
library(knitr)
library(readxl)
library(nlme) # mixed models
library(readr)

# upload dataset and change names
bstm_t1d_raw <- read_csv("S:/Laura/BDC/Projects/Viral Shah/K23/bstm_t1d_raw.csv")
ht1d_raman = read_excel("S:/Laura/BDC/Projects/Viral Shah/Bone samples t1dm - ginger/2022-07-12b_hT1D_Raman_FullDataset-deidentified.xlsx", 
    col_types = c("numeric", "text", "text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))
# interested in coritcal bone
ht1d_raman_c = ht1d_raman %>% mutate(subjectid= `Subject #`, t1d_grp = `...3` ) %>% select(-c(`...3`,`Subject #`)) %>% filter(`Bone type`=="T")
```

```{r heterogeneity, include=FALSE}
# compress sd's down into one subj/row
ht1d_sd = ht1d_raman_c %>% group_by(subjectid, t1d_grp) %>% summarise(v2pa3_sd = sd(v2phos_amideIII),
                                                             cp_sd = sd(carbonate_phosphate),
                                                             cryst_sd = sd(crystallinity))

het_t1 = ht1d_sd %>% 
  select(t1d_grp, v2pa3_sd, cp_sd, cryst_sd, -subjectid) %>% 
  gather(key = variable, value = value, -t1d_grp) %>% 
  group_by(t1d_grp, variable) %>% 
  summarise(value = list(value)) %>% 
  spread(t1d_grp, value) %>% 
  group_by(variable) %>% 
  mutate(t_value = t.test(unlist(T1D), unlist(C))$statistic,
         p_value = t.test(unlist(T1D), unlist(C))$p.value
         )

het_t1 = het_t1 %>% filter(variable != "subjectid") %>% select(-c(T1D, C))
```

```{r mods, include=FALSE}
# phos amideIII
pa_mod = lme(v2phos_amideIII ~ t1d_grp ,
            random = ~1|subjectid,
            data = ht1d_raman_c,
            na.action = na.omit)

# carbphos
cp_mod = lme(carbonate_phosphate ~ t1d_grp ,
            random = ~1|subjectid,
            data = ht1d_raman_c,
            na.action = na.omit)

# cryst
cry_mod = lme(crystallinity ~ t1d_grp ,
            random = ~1|subjectid,
            data = ht1d_raman_c,
            na.action = na.omit)

# tables
# 2x2 table 
t2 = table(ht1d_raman$`Bone type`, ht1d_raman$...3)
# adequate number in each cell, chisq should be fine here
```

# Outstanding Data Remarks/Questions
note: n for groups is a bit small (n = 17 (10 t1d, 7 c))

# Methods
Only Subject's Trabecular bone measurements were used in primary analysis (n = 404 measurements from 17 subjects).

For v2phos, carbonatephos, and crystallinity, mixed models with random intercept for subject were fit with t1d status as a fixed effect. Model output included in code appendix. To assess heterogeneity, standard deviations of measures within subjects were aggregated and compared between groups using a t-test.

# Analysis

## 2x2 bonetype vs disease (added 7/13/2022) 

There is no significant difference in the proportions of bone types within the disease groups (p = 0.96)
```{r 2x2}
# counts
t2
# proportions of bone types
prop.table(t2, 1)

# chisq test
chisq.test(ht1d_raman$`Bone type`, ht1d_raman$...3)
```


## v2Phosphate:Amide III
### Group Differences
Adjusting for within subject correlation, there is no significant difference between T1d and Control groups (p = 0.75)
```{r phosamide plot, echo=FALSE}
ggplot(ht1d_raman_c, aes(x = v2phos_amideIII)) + 
  geom_histogram(aes(color = t1d_grp, fill = t1d_grp), bins = 30, alpha = 0.4) + theme_classic()
```

### Heterogeneity
There is no significant difference between t1d and control groups (p = 0.47)

## Carbonate:Phosphate
### Group Differences
Adjusting for within subject correlation, there is no significant difference between T1d and Control groups (p = 0.19)
```{r carbphos plot, echo=FALSE}
ggplot(ht1d_raman_c, aes(x = carbonate_phosphate)) + 
  geom_histogram(aes(color = t1d_grp, fill = t1d_grp), bins = 30, alpha = 0.4) + theme_classic()
```

### Heterogeneity
There is no significant difference between t1d and control groups (p = 0.82)

## Crystallinity
### Group Differences
Adjusting for within subject correlation, there is no significant difference between T1d and Control groups (p = 0.83)
```{r crystallinnity plot, echo=FALSE}
ggplot(ht1d_raman_c, aes(x = crystallinity)) + 
  geom_histogram(aes(color = t1d_grp, fill = t1d_grp), bins = 30, alpha = 0.4) + theme_classic()
```

### Heterogeneity
There is no significant difference between t1d and control groups (p = 0.24)

# Code Appendix
```{r mods sum}
summary(pa_mod)
summary(cp_mod)
summary(cry_mod)

kable(het_t1)
```