---
title: "BMT CGM Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(redcapAPI)
library(tableone)
library(skimr)
library(knitr)
library(AGD)
library(ega)
library(tidyverse)
```

```{r include=FALSE,cache=TRUE}
# REDCap API data import
source("/Users/timvigers/Documents/GitHub/BDC-Code/api_tokens.R")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = BMT_CGM)
data_full <- exportRecords(rcon)
```

```{r echo=FALSE}
# Remove participant 10
data_full <- data_full %>% filter(study_id != "10")
# Get demographic info
demographics <- data_full %>% group_by(study_id) %>% filter(!is.na(mrn))
# Combine separate race columns into one
levels(demographics$race___0) <- c(NA,"AI or AN")
levels(demographics$race___1) <- c(NA,"Asian")
levels(demographics$race___2) <- c(NA,"African American")
levels(demographics$race___3) <- c(NA,"NH or PI")
levels(demographics$race___4) <- c(NA,"White")
levels(demographics$race___5) <- c(NA,"Hispanic")
levels(demographics$race___6) <- c(NA,"Other")
demographics <- demographics %>% ungroup() %>% unite("Race",race___0,race___1,race___2,race___3,race___4,race___5,race___6)
demographics$Race <- gsub("NA","",demographics$Race)
demographics$Race <- gsub("_","",demographics$Race)
demographics$Race[which(demographics$race_multi == "yes")] <- "Multiple"
demographics$Race[which(demographics$Race == "")] <- "Unknown"
# Calculate age at transplant
demographics$AgeTx <- as.numeric(difftime(demographics$date_of_tx,demographics$date_of_birth,units = "days")) / 365.25
# Format/calculate variables for BMI calculation
demographics$sex <- demographics$gender
levels(demographics$sex) <- c("F","M")
# Calculate BMI and BMI percentile
demographics$bmi <- as.numeric(demographics$weight) / ((as.numeric(demographics$height)/100)^2)
demographics$bmi_z <- y2z(demographics$bmi,x = demographics$AgeTx,sex = demographics$sex,ref = get("cdc.bmi"))
demographics$bmi_percentile <- round(pnorm(demographics$bmi_z) * 100,3)
```

## Table 1: Descriptive Characteristics

```{r echo=FALSE}
dem_vars <- c("AgeTx","bmi_percentile","gender","Race","tanner_stage","tx_type","primary_dx_cat","spec_dx_cat")
t1 <- CreateTableOne(dem_vars,data = demographics)
t1 <- as.data.frame(print(t1, nonnormal = c("AgeTx","bmi_percentile"),printToggle = F,showAllLevels = T,missing = T))
t1 <- cbind(new = NA,t1)
colnames(t1) <- c("","Level","Overall","% Missing")
t1[,1] <- c("n","Age at HSCT (median [IQR])","BMI Percentile (median [IQR])",
            "Gender (%)","","Race (%)","","","","","",
            "Tanner Stage (%)","","","","","","HSCT Type (%)","",
            "Primary Diagnosis (%)","",
            "Specific Diagnosis (%)","","","","","","","","")
kable(t1,row.names = F)
```

## Table 2: Feasibility/Safety Tables

```{r echo=FALSE}
# Get CGM data
cgm_data <- data_full %>% filter(is.na(redcap_repeat_instrument)) %>%
  select(study_id,date_of_tx,
         sens1_place,plt1,transfuse1,bleed1,sens1_remove,sens1_reason,
         sens2_place,plt2,transfuse2,bleed2,sens2_remove,sens2_reason,
         sens3_place,plt3,transfuse3,bleed3,sens3_remove,sens3_reason,
         sens4_place,plt4,transfuse4,bleed4,sens4_remove,sens4_reason,
         sens5_place,plt5,transfuse5,bleed5,sens5_remove,sens5_reason,
         sens6_place,plt6,transfuse6,bleed6,sens6_remove,sens6_reason)

colnames(cgm_data) <- c("study_id","date_of_tx",
                        "sens_place.1","plt.1","transfuse.1","bleed.1","sens_remove.1","sens_reason.1",
                        "sens_place.2","plt.2","transfuse.2","bleed.2","sens_remove.2","sens_reason.2",
                        "sens_place.3","plt.3","transfuse.3","bleed.3","sens_remove.3","sens_reason.3",
                        "sens_place.4","plt.4","transfuse.4","bleed.4","sens_remove.4","sens_reason.4",
                        "sens_place.5","plt.5","transfuse.5","bleed.5","sens_remove.5","sens_reason.5",
                        "sens_place.6","plt.6","transfuse.6","bleed.6","sens_remove.6","sens_reason.6")

# Wide to long
cgm_data <- reshape(cgm_data,idvar = c("study_id","date_of_tx"),
                varying = colnames(cgm_data)[3:ncol(cgm_data)],
                direction = "long")
# Remove blank rows based on sensor placement date, format DF, calculate days of CGM wear
cgm_data <- cgm_data %>% filter(!is.na(sens_place)) %>% 
  arrange(as.numeric(study_id)) %>%
  mutate(days_worn_total = as.numeric(difftime(sens_remove,sens_place,units = 'days')),
         days_worn_pre_tx = 
           pmax(days_worn_total - pmax(as.numeric(difftime(sens_remove,date_of_tx,units = 'days')),0),0),
         days_worn_post_tx = 
           pmax(days_worn_total - pmax(as.numeric(difftime(date_of_tx,sens_place,units = 'days')),0),0))
# Platelet count variables
cgm_data$plt <- as.numeric(cgm_data$plt)
cgm_data$plt_bleed <- ifelse(cgm_data$bleed == "Yes",cgm_data$plt,NA)
cgm_data$plt_no_bleed <- ifelse(cgm_data$bleed == "No",cgm_data$plt,NA)
# Summary variables
cgm_summary <- cgm_data %>% group_by(study_id) %>%
  summarise(num_cgms = n(),
            total_days_worn = sum(days_worn_total),
            total_days_pre_tx = sum(days_worn_pre_tx,na.rm = T),
            total_days_post_tx = sum(days_worn_post_tx),
            last_cgm_removal_post_tx = 
              last(as.numeric(difftime(sens_remove,date_of_tx,units = 'days'))[which(sens_remove > date_of_tx)]))
# Remove participant 5 pre-HSCT data since they lost CGM close to translant
cgm_summary$total_days_pre_tx[5] <- NA
# AE and CGM refusal information
ae <- data_full %>% 
  select(study_id,adv_event:ae_attrib) %>% 
  filter(!is.na(adv_event)) %>%
  group_by(study_id) %>%
  summarise(bleeding = ifelse(sum(ae_type___4 == "Checked",na.rm = T) > 0,"Yes","No"),
            infection = ifelse(sum(ae_type___2 == "Checked",na.rm = T) > 0,"Yes","No"),
            skin_reaction = ifelse(sum((ae_type___0 == "Checked" | ae_type___1 == "Checked"),na.rm = T) > 0,"Yes","No"),
            other = ifelse(sum(ae_type___5 == "Checked",na.rm = T) > 0,"Yes","No"))
# Refusal
refuse <- data_full %>% select(study_id,refuse_cgm:refusal_reason___4) %>%
  filter(refuse_cgm == "Yes")
# Add AEs
cgm_summary <- full_join(cgm_summary,ae,by = "study_id") %>% arrange(as.numeric(study_id))
# Refusal
cgm_summary$refuse_cgm <- ifelse(cgm_summary$study_id %in% refuse$study_id,"Yes","No")
cgm_summary$time_to_refuse <- ifelse(cgm_summary$refuse_cgm == "Yes",cgm_summary$last_cgm_removal_post_tx,NA)
```

### By Patient

```{r echo=FALSE}
vars <- colnames(cgm_summary)[-1]
t2_by_patient <- CreateTableOne(vars,data = cgm_summary)
t2_by_patient <- as.data.frame(print(t2_by_patient, missing = T,printToggle = F,
                                     showAllLevels = T,
                                     nonnormal = c("num_cgms","total_days_worn",
                                                   "total_days_post_tx",
                                                   "last_cgm_removal_post_tx",
                                                   "time_to_refuse")))
t2_by_patient <- cbind(new = NA,t2_by_patient)
colnames(t2_by_patient) <- c("","Level","Overall","% Missing")
t2_by_patient[,1] <- c("n","Number of CGMs (median [IQR])",
                       "Total Days Worn (median [IQR])","Total Days Worn Pre-HSCT",
                       "Total Days Worn Post-HSCT (median [IQR])",
                       "Number of Days HSCT to Last CGM Removal (median [IQR])",
                       "AE - Bleeding (%)","","AE - Infection (%)",
                       "AE - Skin Reaction (%)","AE - Other (%)","",
                       "Refused CGM (%)","",
                       "Days to CGM Refusal (median [IQR])")
t2_by_patient$`% Missing` <- as.character(t2_by_patient$`% Missing`)
t2_by_patient$`% Missing`[nrow(t2_by_patient)] <- "0.0"
kable(t2_by_patient,row.names = F)
```

Number of CGMs worn was counted based on the number of non-missing sensor placement dates. Of the 84 CGMs with placement dates, 2 did not have removal dates (participant 5 placed 6/8/17 and participant 6 placed 8/7/17). Total days worn variables could not be calculated for these two participants, and the number of days from HSCT to last removal was calculated using the last CGM wear with a non-missing removal date.

Time to CGM refusal was calculated based on time from transplant to last CGM removal for those who refused CGM. All participants who refused CGM refused due to patient/parent preference:

```{r echo=FALSE}
kable(refuse)
```

### All CGMs

```{r echo=FALSE}
t2_all <- CreateTableOne(c("plt","plt_bleed","plt_no_bleed","bleed","sens_reason"),
                         data = cgm_data)
t2_all <- as.data.frame(print(t2_all,printToggle = F,nonnormal = c("plt","plt_bleed","plt_no_bleed"),
                              missing=T,showAllLevels = T))
t2_all <- cbind(new = NA,t2_all)
colnames(t2_all) <- c("","Level","Overall","% Missing")
t2_all[,1] <- c("n","Platelet Count at Time of CGM Placement - All (median [IQR])",
                "Platelet Count at CGM Placement - Bleeding (median [IQR])",
                "Platelet Count at CGM Placement - No Bleeding (median [IQR])",
                "Bleeding Events (%)","",
                "Sensor Removal Reason (%)","","","","","","","","")
t2_all$`% Missing` <- as.character(t2_all$`% Missing`)
t2_all$`% Missing`[3] <- "0.0"
kable(t2_all,row.names = F)
```

# CGM Glucose vs. Serum

```{r echo=FALSE,cache=TRUE}
# Read detailed CGM data and serum data
cgm <- read.csv("/Volumes/som-1/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Jenna Sopfe/BMT/Data_Cleaned/Prospective HSCT CGM Data Updated 6.24.19 FULL DATA SET.csv")
serum <- read.csv("/Volumes/som-1/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Jenna Sopfe/BMT/Data_Cleaned/Glucose_Monitoring_061019 edited 6.25.19.csv",na.strings = "") %>% 
  filter(!is.na(glucose_spec_taken_datetime))
# Format
cgm$Time <- lubridate::mdy_hm(cgm$Time)
cgm$Tx_date <- lubridate::mdy(cgm$Tx_date)
serum$glucose_spec_taken_datetime <- lubridate::dmy_hms(serum$glucose_spec_taken_datetime)
serum$date_of_tx <- lubridate::mdy(serum$date_of_tx)
colnames(cgm)[1] <- "study_id"
# Days from Tx
cgm$Days <- round(as.numeric(difftime(cgm$Time,cgm$Tx_date,units = "days")))
# Find CGM values closest to serum - really slow because of double for loop
for (r in 1:nrow(serum)) {
  id <- serum[r,"study_id"] # Match within study ID (some patients' wears overlap)
  temp <- cgm[which(cgm$study_id == id),]
  dates <- cgm_data[which(cgm_data$study_id == id),] # Get dates of CGM wear
  dates <- dates[which(!is.na(dates$sens_place) & !is.na(dates$sens_remove)),]
  date_list <- c()
  for (row in 1:nrow(dates)) {
    date_list <- c(date_list,dates$sens_place[row]:dates$sens_remove[row])
  }
  if (serum$glucose_spec_taken_datetime[r] %in% date_list){
    serum[r,"delete"] <- "N" # Exclude serum if not during CGM wear
  } else {
    serum[r,"delete"] <- "Y"
  }
  time <- temp$Time[which.min(abs(serum$glucose_spec_taken_datetime[r] - temp$Time))]
  serum[r,"Time"] <- time
}
# Remove serum values when patient was not wearing a CGM
serum <- serum[serum$delete == "N",]
# Join
cgm_serum <- left_join(serum,cgm,by=c("study_id","Time"))
cgm_serum <- cgm_serum %>% select(study_id,mrn,glucose_spec_taken_datetime,glucose_rslt,Time,Historic_Glucose,date_of_tx,Days)
colnames(cgm_serum) <- c("ID","MRN","Serum Time","Serum Value","CGM Time","CGM Value","Date of Tx","Days from Tx")
cgm_serum <- cgm_serum %>% arrange(ID,`Serum Time`)
# Calculate MARD
cgm_serum$`Absolute Diff.` <- abs(cgm_serum$`CGM Value` - cgm_serum$`Serum Value`)
cgm_serum$`Absolute Relative Diff.` <- cgm_serum$`Absolute Diff.` / cgm_serum$`Serum Value`
# Time periods
cgm_serum$`Time Period` <- cut(cgm_serum$`Days from Tx`,breaks = c(-7,0,15,29,43,60,Inf),
                               labels = c("Days -7 to -1","Days 0-14","Days 15-28","Days 29-42","Days 43-60","Days 60+"),right = F)
```

## Table 3: Continuous Glucose Monitoring and Blood Glucose Monitoring Metrics

```{r echo=FALSE}
vars <- c("CGM Value","Serum Value","Absolute Diff.","Absolute Relative Diff.")
t3 <- CreateTableOne(vars,data = cgm_serum)
kableone(t3)
```

## Table 4: Continuous Glucose Monitoring Accuracy by Time Period

```{r echo=FALSE,warning=FALSE}
vars <- c("Absolute Diff.","Absolute Relative Diff.")
t4 <- CreateTableOne(vars,strata = "Time Period",data = cgm_serum)
t4 <- as.data.frame(print(t4,printToggle = F))
t4 <- t(t4)
t4 <- t4[1:5,]
kable(t4)
```

## Figure 1: Glucose Distribution by Time Period

### a. CGM Values

```{r echo=FALSE}
fig1a <- ggplot(data = cgm_serum[!is.na(cgm_serum$`Time Period`),],aes(x = `Time Period`,y = `CGM Value`,na.r)) + geom_boxplot()
fig1a
```

### b. Serum Values

```{r echo=FALSE}
fig1b <- ggplot(data = cgm_serum[!is.na(cgm_serum$`Time Period`),],aes(x = `Time Period`,y = `Serum Value`,na.r)) + geom_boxplot()
fig1b
```

## Table 5: Clarke Error Grid Analysis (Serum as Reference)

```{r echo=FALSE}
zones <- getClarkeZones(cgm_serum$`Serum Value`,cgm_serum$`CGM Value`)
points <- table(factor(zones))
percentages <- round(table(factor(zones))/length(factor(zones))*100,3)
clark_table <- as.data.frame(rbind(points,percentages))
clark_table$Total <- rowSums(clark_table)
clark_table[3,] <- c(percentages[1] + percentages[2],rep("",ncol(clark_table)-1))
rownames(clark_table) <- c("Points","%","% A+B")
kable(clark_table)
```

## Figure 2: Clark Error Grid (Serum as Reference)

```{r echo=FALSE}
plotClarkeGrid(cgm_serum$`Serum Value`,cgm_serum$`CGM Value`)
```


# Questions for Jenna
- Did she mean boxplot instead of histogram?
- What to do about serum values more than 7 days prior to transplant?
- For CEGs, is serum the reference? Or CGM?