---
title: "Disglycemia in BMT Patients"
author: "Kristen Campbell & Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(redcapAPI)
library(AGD)
library(arsenal)
library(skimr)
library(knitr)
library(survival)
library(survminer)
library(survivalROC)
library(tdROC)
library(tidyverse)
```

```{r data import,echo=FALSE,cache=TRUE}
# REDCap API data import
api <- read.table("./Jenna Sopfe/api_token.txt",header = T,sep = "\t")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = as.character(api[1,1]))
data_full <- exportRecords(rcon)
# Remove participant 10
data_full <- data_full %>% filter(study_id != "10")
# Fill down Tx date
data_full <- data_full %>% group_by(study_id) %>% fill(date_of_tx) %>%
  ungroup()
# Format dates
datecols = c("date_death","date_of_tx","infxn_date","sens1_place","followup")
data_full[,datecols] <- lapply(data_full[,datecols],lubridate::ymd)
data_full$id <- as.character(data_full$study_id)
```

```{r df,echo=FALSE,cache=TRUE}
df <- data_full %>% filter(is.na(redcap_repeat_instrument))
# Combine separate race columns into one
levels(df$race___0) <- c(NA,"AI or AN")
levels(df$race___1) <- c(NA,"Asian")
levels(df$race___2) <- c(NA,"African American")
levels(df$race___3) <- c(NA,"NH or PI")
levels(df$race___4) <- c(NA,"White")
levels(df$race___5) <- c(NA,"Hispanic")
levels(df$race___6) <- c(NA,"Other")
df <- df %>% 
  unite("Race",race___0,race___1,race___2,race___3,race___4,
        race___5,race___6)
df$Race <- gsub("NA","",df$Race)
df$Race <- gsub("_","",df$Race)
df$Race[which(df$race_multi == "yes")] <- "Multiple"
df$Race[which(df$Race == "")] <- "Unknown"
# Calculate age at transplant
df$AgeTx <- as.numeric(difftime(df$date_of_tx,
                                df$date_of_birth,
                                units = "days")) / 365.25
# Format/calculate variables for BMI calculation
df$sex <- df$gender
levels(df$sex) <- c("F","M")
# Calculate BMI and BMI percentile
df$bmi <- as.numeric(df$weight) / 
  ((as.numeric(df$height)/100)^2)
df$bmi_z <- y2z(df$bmi,x = df$AgeTx,
                sex = df$sex,ref = get("cdc.bmi"))
df$bmi_percentile <- round(pnorm(df$bmi_z) * 100,3)
# BMI percentile groups
df$bmi_percentile_group <- 
  cut(df$bmi_percentile,
      breaks = c(0,85,95,100),
      labels = c("<85th %ile","85th-95th %ile",">= 95th %ile"),
      right = F)
# Infer Tanner stage
df$tanner_stage[df$AgeTx >= 18] <- "5"
df$tanner_stage[df$AgeTx <= 3] <- "1"
# Remove unnecessary columns
df <- df %>%
  select(id,date_of_tx,AgeTx,gender,tanner_stage,Race,gender,
         post_tx_steroids,bmi_percentile_group,tx_type,primary_dx_cat,
         spec_dx_cat,followup,date_death)
# Get earliest infection for each participant
infections <- data_full %>% 
  filter(redcap_repeat_instrument == "infections",
         redcap_repeat_instance == 1)
# Add to df
df = left_join(df,infections[,c("id","infxn_date")],by = "id")
# Event indicators
df$death <- factor(ifelse(!is.na(df$date_death),"Yes","No"))
df$infxn <- factor(ifelse(!is.na(df$infxn_date),"Yes","No"))
# Add censoring
df$infxn_date <- pmin(df$infxn_date,df$followup,na.rm = T)
df$date_death <- pmin(df$date_death,df$followup,na.rm = T)
# Time to event
df$time_to_infxn <- as.numeric(difftime(df$infxn_date,df$date_of_tx,
                                        units = "days"))
df$time_to_death <- as.numeric(difftime(df$date_death,df$date_of_tx,
                                        units = "days"))
```

# Table 1: Descriptive Characteristics

Patients aged 2 - 3 years were assumed to be Tanner stage 1, and patients aged 18+ were assumed to be Tanner stage 5. Participant 10 was excluded from all analyses.

```{r table 1,echo=FALSE,results='asis'}
table_one <- tableby(~ notest(AgeTx,"median","q1q3") + gender + 
                       tanner_stage + bmi_percentile_group + tx_type + 
                       primary_dx_cat + spec_dx_cat + infxn, 
                     data = df)
newlabels <- list(AgeTx = "Age at HSCT",Gender = "Sex",
                  bmi_percentile_group = "BMI Percentile",
                  `Transplant Type` = "HSCT Type",
                  `Primary Diagnosis Category` = "Primary Diagnosis",
                  `Specific Diagnosis Category` = "Specific Diagnosis",
                  infxn = "Infection")
summary(table_one,labelTranslations = newlabels,digits = 2)
```

# Table 2: Glycemic Control

```{r cgm summary,echo=FALSE,results='asis',cache=TRUE,message=FALSE}
# Import CGM
cgm = read.csv("./Jenna Sopfe/CGM Accuracy in BMT/Data_Cleaned/Prospective HSCT CGM Data Updated 6.24.19 FULL DATA SET.csv")
# Format columns
colnames(cgm)[1] <- "id"
cgm$id <- as.character(cgm$id)
cgm$Time = lubridate::mdy_hm(cgm$Time,tz = "UTC")
cgm$Tx_date = lubridate::mdy(cgm$Tx_date,tz = "UTC")
# Days from Tx
cgm$days_from_tx = 
  as.numeric(difftime(as.Date(cgm$Time),cgm$Tx_date,units = "days"))
cgm$time_cat = cut(cgm$days_from_tx,c(-Inf,-1,14,Inf),
                   labels = c("Day -7 to -1","Days 0 to 14","Day 14+"))
# Excursions above 140
cgm$exc_140 = ifelse(cgm$Historic_Glucose >= 140,1,0)
# Summary variables - excursions over 140 count if longer than 30 minutes
cgm_summ = cgm %>% group_by(id,time_cat) %>%
  summarise(avg = mean(Historic_Glucose,na.rm = T),
            cv = avg / sd(Historic_Glucose,na.rm = T),
            perc_70 = round(sum(Historic_Glucose < 70)/n(),3) * 100,
            perc_70_140 = round(sum(Historic_Glucose >= 70 & 
                                      Historic_Glucose <= 140)/n(),3)*100,
            perc_140 = round(sum(Historic_Glucose > 140)/n(),3) * 100,
            perc_200 = round(sum(Historic_Glucose > 200)/n(),3) * 100,
            num_exc_140 = 
              length(which(rle(exc_140)$lengths[which(rle(exc_140)$values == 1)] > 6)),
            n_obs = n(),
            num_days = max(abs(days_from_tx)))
# Summarize summary metrics
gluc_table = tableby(time_cat ~ avg + cv + kwt(perc_70,"median","q1q3") + 
                       kwt(perc_70_140,"median","q1q3") + 
                       kwt(perc_140,"median","q1q3") + kwt(perc_200,"median","q1q3") + 
                       kwt(num_exc_140,"median","q1q3") + 
                       kwt(n_obs,"median","q1q3") + kwt(num_days,"median","q1q3"),
                     data = cgm_summ)
summary(gluc_table,pfootnote = T,
        labelTranslations = 
          list(avg = "Mean SG (mg/dL)",cv = "CV",
               perc_70 = "% Time < 70 mg/dL",
               perc_70_140 = "% Time 70-140 mg/dL",
               perc_140 = "% Time > 140 mg/dL",
               perc_200 = "% Time > 200 mg/dL",
               num_exc_140 = "Number of Excursions > 140",
               n_obs = "Number of Observations",
               num_days = "Number of Days With CGM"))
# Add CGM summary metrics to df
# Long to wide
cgm_summ <- cgm_summ %>% 
  pivot_wider(id_cols = "id",names_from = time_cat,
              names_sep = " ",
              values_from = avg:num_days)
df <- left_join(df,cgm_summ,by = "id")
# Numeric outcome
df$infxn <- as.numeric(df$infxn)-1
df$death <- as.numeric(df$death)-1
```

# Survival Models

Age and Tanner 1 vs. Tanner 2-5 appear to violate proportional hazard (PH) assumptions.

```{r cutoff function and data,echo=FALSE,message=FALSE}
# tdROC cutoff function
opt_cutoff <- function(data = df,time = "time_to_infxn", event = "infxn",var,
                       nboot = 0,n.grid = 1000){
  temp <- data[,c(var,time,event)]
  temp <- as.data.frame(temp[complete.cases(temp),])
  if(grepl("14\\+",var)){temp <- temp[-c(which(temp[,time] <=14)),]}
  res <- as.data.frame(matrix(ncol = 3))
  colnames(res) <- c("Variable","Cutoff","AUC")
  res$Variable <- var
  if (nrow(temp) > 0) {
    fm <- tdROC(X=temp[,var],Y=temp[,time],delta = temp[,event], 
                tau = max(temp[,time]),n.grid = n.grid)
    auc <- as.numeric(fm$AUC2[1])
    fm <- as.data.frame(fm$ROC)
    fm$youd <- fm$sens + fm$spec - 1
    cut <- fm$grid[which(fm$youd == max(fm$youd))]
    cut <- cut[which(is.finite(cut))]
    res$Cutoff <- paste(round(min(cut)),"-",round(max(cut)))
    res$AUC <- round(auc,3)
  } else {
    res$Cutoff <- NA
    res$AUC <- NA
  }
  res
}
# Format data
df$tanner_stage_binary <- as.character(df$tanner_stage)
df$tanner_stage_binary[df$tanner_stage_binary=="unknown"] <- NA
df$tanner_stage_binary <- as.factor(df$tanner_stage_binary)
levels(df$tanner_stage_binary) <- c("Tanner 1","Tanner 2-5","Tanner 2-5",
                                    "Tanner 2-5")
```

## Time to First Infection

### Mean glucose

#### Results

```{r avg survival,echo=FALSE}
# Cox models 
# Days -7 to -1
avg_pre_cox <- coxph(Surv(time_to_infxn,infxn) ~ `avg Day -7 to -1`+
                       tx_type+post_tx_steroids,
                     data = df)
# Diagnostics
d <- cox.zph(avg_pre_cox)
#kable(d$table)
#ggcoxzph(d)
res <- broom::tidy(avg_pre_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
# Days 0 to 14
avg_imm_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `avg Days 0 to 14`+
                            tx_type+post_tx_steroids,
                          data = df)
# Diagnostics
d <- cox.zph(avg_imm_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(avg_imm_post_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
# Days 14+
avg_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `avg Day 14+`+
                        tx_type+post_tx_steroids,
                      data = df[which(df$time_to_infxn>14),])
# Diagnostics
d <- cox.zph(avg_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(avg_post_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
```

#### Cutoffs

```{r avg cuts,echo=FALSE}
l <- lapply(c("avg Day -7 to -1","avg Days 0 to 14","avg Day 14+"), function(x){
  opt_cutoff(var = x)
})
kable(do.call(rbind,l))
```

### Glucose CV

#### Results

```{r cv survival,echo=FALSE}
# Cox models 
# Days -7 to -1
cv_pre_cox <- coxph(Surv(time_to_infxn,infxn) ~ `cv Day -7 to -1`+
                      tx_type+post_tx_steroids,
                    data = df)
# Diagnostics
d <- cox.zph(cv_pre_cox)
#kable(d$table)
#ggcoxzph(d)
res <- broom::tidy(cv_pre_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
# Days 0 to 14
cv_imm_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `cv Days 0 to 14`+
                           tx_type+post_tx_steroids,
                         data = df)
# Diagnostics
d <- cox.zph(cv_imm_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(cv_imm_post_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
# Days 14+
cv_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `cv Day 14+`+
                       tx_type+post_tx_steroids,
                     data = df[which(df$time_to_infxn>14),])
# Diagnostics
d <- cox.zph(cv_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(cv_post_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
```

#### Cutoffs

```{r cv cuts,echo=FALSE}
l <- lapply(c("cv Day -7 to -1","cv Days 0 to 14","cv Day 14+"), function(x){
  opt_cutoff(var = x)
})
kable(do.call(rbind,l))
```

### Percent time > 140

#### Results

```{r perc survival,echo=FALSE}
# Cox models 
# Days -7 to -1
perc_140_pre_cox <- coxph(Surv(time_to_infxn,infxn) ~ `perc_140 Day -7 to -1`+
                            tx_type+post_tx_steroids,
                          data = df)
# Diagnostics
d <- cox.zph(perc_140_pre_cox)
#kable(d$table)
#ggcoxzph(d)
res <- broom::tidy(perc_140_pre_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
# Days 0 to 14
perc_140_imm_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `perc_140 Days 0 to 14`+
                                 tx_type+post_tx_steroids,
                               data = df)
# Diagnostics
d <- cox.zph(perc_140_imm_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(perc_140_imm_post_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
# Days 14+
perc_140_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `perc_140 Day 14+`+
                             tx_type+post_tx_steroids,
                           data = df[which(df$time_to_infxn>14),])
# Diagnostics
d <- cox.zph(perc_140_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(perc_140_post_cox)
res$term[2:nrow(res)] <- c("Allogeneic Tx","Post-Tx Steroids")
kable(res)
```

#### Cutoffs

```{r perc cuts,echo=FALSE}
l <- lapply(c("perc_140 Day -7 to -1","perc_140 Days 0 to 14","perc_140 Day 14+"), function(x){
  opt_cutoff(var = x)
})
kable(do.call(rbind,l))
```

### Notes

Mean glucose, glucose coefficient of variation, percent time above 140 mg/dL, and transplant type were not associated with increased risk of infection. Post-BMT steroids are strongly associated with infection. 

## Time to Death

There is not enough information to include both transplant type and steroid use in these models (the table below is simply to illustrate that there are no events in certain levels of these variables). When possible, post-transplant steroid use was kept in the model because it was associated with infection. However, it appeared to violate proportion

```{r echo=FALSE,results='asis'}
summary(tableby(death~interaction(tx_type,post_tx_steroids),df))
```

### Mean glucose

#### Results

```{r avg death,echo=FALSE}
# Cox models 
# Days -7 to -1
avg_pre_cox <- coxph(Surv(time_to_death,death) ~ `avg Day -7 to -1`+
                       tx_type,data = df)
# Diagnostics
d <- cox.zph(avg_pre_cox)
#kable(d$table)
#ggcoxzph(d)
res <- broom::tidy(avg_pre_cox)
res$term[2] <- "Allogeneic Tx"
kable(res)
# Days 0 to 14
avg_imm_post_cox <- coxph(Surv(time_to_death,death) ~ `avg Days 0 to 14`+
                            post_tx_steroids,
                          data = df)
# Diagnostics
d <- cox.zph(avg_imm_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(avg_imm_post_cox)
res$term[2] <- "Post-Tx Steroids"
kable(res)
# Days 14+
avg_post_cox <- coxph(Surv(time_to_death,death) ~ `avg Day 14+`+
                        post_tx_steroids,
                      data = df[which(df$time_to_death>14),])
# Diagnostics
d <- cox.zph(avg_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(avg_post_cox)
res$term[2] <- "Post-Tx Steroids"
kable(res)
```

#### Cutoffs

```{r avg cuts death,echo=FALSE}
l <- lapply(c("avg Day -7 to -1","avg Days 0 to 14","avg Day 14+"), function(x){
  opt_cutoff(var = x,time = "time_to_death", event = "death")
})
kable(do.call(rbind,l))
```

### Glucose CV

#### Results

```{r cv survival death,echo=FALSE}
# Cox models 
# Days -7 to -1
cv_pre_cox <- coxph(Surv(time_to_death,death) ~ `cv Day -7 to -1`+
                      post_tx_steroids,data = df)
# Diagnostics
d <- cox.zph(cv_pre_cox)
#kable(d$table)
#ggcoxzph(d)
res <- broom::tidy(cv_pre_cox)
res$term[2:nrow(res)] <- "Post-Tx Steroids"
kable(res)
# Days 0 to 14
cv_imm_post_cox <- coxph(Surv(time_to_death,death) ~ `cv Days 0 to 14`+
                           post_tx_steroids,data = df)
# Diagnostics
d <- cox.zph(cv_imm_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(cv_imm_post_cox)
res$term[2:nrow(res)] <- "Post-Tx Steroids"
kable(res)
# Days 14+
cv_post_cox <- coxph(Surv(time_to_death,death) ~ `cv Day 14+`+
                       post_tx_steroids,data = df[which(df$time_to_death>14),])
# Diagnostics
d <- cox.zph(cv_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(cv_post_cox)
res$term[2:nrow(res)] <- "Post-Tx Steroids"
kable(res)
```

#### Cutoffs

```{r cv cuts death,echo=FALSE}
l <- lapply(c("cv Day -7 to -1","cv Days 0 to 14","cv Day 14+"), function(x){
  opt_cutoff(var = x)
})
kable(do.call(rbind,l))
```

### Percent time > 140

#### Results

```{r perc survival death,echo=FALSE}
# Cox models 
# Days -7 to -1
perc_140_pre_cox <- coxph(Surv(time_to_death,death) ~ `perc_140 Day -7 to -1`+
                            post_tx_steroids,data = df)
# Diagnostics
d <- cox.zph(perc_140_pre_cox)
#kable(d$table)
#ggcoxzph(d)
res <- broom::tidy(perc_140_pre_cox)
res$term[2:nrow(res)] <- "Post-Tx Steroids"
kable(res)
# Days 0 to 14
perc_140_imm_post_cox <- coxph(Surv(time_to_death,death) ~ `perc_140 Days 0 to 14`+
                                 post_tx_steroids,data = df)
# Diagnostics
d <- cox.zph(perc_140_imm_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(perc_140_imm_post_cox)
res$term[2:nrow(res)] <- "Post-Tx Steroids"
kable(res)
# Days 14+
perc_140_post_cox <- coxph(Surv(time_to_death,death) ~ `perc_140 Day 14+`+
                             post_tx_steroids,data = df[which(df$time_to_death>14),])
# Diagnostics
d <- cox.zph(perc_140_post_cox)
# kable(d$table)
# ggcoxzph(d)
res <- broom::tidy(perc_140_post_cox)
res$term[2:nrow(res)] <- "Post-Tx Steroids"
kable(res)
```

#### Cutoffs

```{r perc cuts death,echo=FALSE}
l <- lapply(c("perc_140 Day -7 to -1","perc_140 Days 0 to 14","perc_140 Day 14+"), function(x){
  opt_cutoff(var = x)
})
kable(do.call(rbind,l))
```

### Notes

Mean glucose, glucose coefficient of variation, and percent time above 140 mg/dL were not associated with increased risk of death.