---
title: "Disglycemia in BMT Patients"
author: "Kristen Campbell & Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(redcapAPI)
library(AGD)
library(arsenal)
library(skimr)
library(knitr)
library(survival)
library(survminer)
library(survivalROC)
library(tdROC)
library(timeROC)
library(tidyverse)
```

```{r data import,echo=FALSE,cache=TRUE}
# REDCap API data import
api <- read.table("./Jenna Sopfe/api_token.txt",header = T,sep = "\t")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = as.character(api[1,1]))
data_full <- exportRecords(rcon)
# Remove participant 10
data_full <- data_full %>% filter(study_id != "10")
# Fill down Tx date
data_full <- data_full %>% group_by(study_id) %>% fill(date_of_tx) %>%
  ungroup()
# Format dates
datecols = c("date_death","date_of_tx","infxn_date","sens1_place","followup")
data_full[,datecols] <- lapply(data_full[,datecols],lubridate::ymd)
data_full$id <- as.character(data_full$study_id)
```

```{r df,echo=FALSE,cache=TRUE}
df <- data_full %>% filter(is.na(redcap_repeat_instrument))
# Combine separate race columns into one
levels(df$race___0) <- c(NA,"AI or AN")
levels(df$race___1) <- c(NA,"Asian")
levels(df$race___2) <- c(NA,"African American")
levels(df$race___3) <- c(NA,"NH or PI")
levels(df$race___4) <- c(NA,"White")
levels(df$race___5) <- c(NA,"Hispanic")
levels(df$race___6) <- c(NA,"Other")
df <- df %>% 
  unite("Race",race___0,race___1,race___2,race___3,race___4,
        race___5,race___6)
df$Race <- gsub("NA","",df$Race)
df$Race <- gsub("_","",df$Race)
df$Race[which(df$race_multi == "yes")] <- "Multiple"
df$Race[which(df$Race == "")] <- "Unknown"
# Calculate age at transplant
df$AgeTx <- as.numeric(difftime(df$date_of_tx,
                                df$date_of_birth,
                                units = "days")) / 365.25
# Format/calculate variables for BMI calculation
df$sex <- df$gender
levels(df$sex) <- c("F","M")
# Calculate BMI and BMI percentile
df$bmi <- as.numeric(df$weight) / 
  ((as.numeric(df$height)/100)^2)
df$bmi_z <- y2z(df$bmi,x = df$AgeTx,
                sex = df$sex,ref = get("cdc.bmi"))
df$bmi_percentile <- round(pnorm(df$bmi_z) * 100,3)
# BMI percentile groups
df$bmi_percentile_group <- 
  cut(df$bmi_percentile,
      breaks = c(0,85,95,100),
      labels = c("<85th %ile","85th-95th %ile",">= 95th %ile"),
      right = F)
# Infer Tanner stage
df$tanner_stage[df$AgeTx >= 18] <- "5"
df$tanner_stage[df$AgeTx <= 3] <- "1"
# Remove unnecessary columns
df <- df %>%
  select(id,date_of_tx,AgeTx,gender,tanner_stage,Race,gender,
         post_tx_steroids,bmi_percentile_group,tx_type,primary_dx_cat,
         spec_dx_cat,followup,date_death)
# Get earliest infection for each participant
infections <- data_full %>% 
  filter(redcap_repeat_instrument == "infections",
         redcap_repeat_instance == 1)
# Add to df
df = left_join(df,infections[,c("id","infxn_date")],by = "id")
# Event indicators
df$death <- factor(ifelse(!is.na(df$date_death),"Yes","No"))
df$infxn <- factor(ifelse(!is.na(df$infxn_date),"Yes","No"))
# Add censoring
df$infxn_date <- pmin(df$infxn_date,df$followup,na.rm = T)
df$date_death <- pmin(df$date_death,df$followup,na.rm = T)
# Time to event
df$time_to_infxn <- as.numeric(difftime(df$infxn_date,df$date_of_tx,
                                        units = "days"))
df$time_to_death <- as.numeric(difftime(df$date_death,df$date_of_tx,
                                        units = "days"))
```

# Table 1: Descriptive Characteristics

Patients aged 2 - 3 years were assumed to be Tanner stage 1, and patients aged 18+ were assumed to be Tanner stage 5. Participant 10 was excluded from all analyses.

```{r table 1,echo=FALSE,results='asis'}
table_one <- tableby(~ notest(AgeTx,"median","q1q3") + gender + 
                       tanner_stage + bmi_percentile_group + tx_type + 
                       primary_dx_cat + spec_dx_cat + infxn, 
                     data = df)
newlabels <- list(AgeTx = "Age at HSCT",Gender = "Sex",
                  bmi_percentile_group = "BMI Percentile",
                  `Transplant Type` = "HSCT Type",
                  `Primary Diagnosis Category` = "Primary Diagnosis",
                  `Specific Diagnosis Category` = "Specific Diagnosis",
                  infxn = "Infection")
summary(table_one,labelTranslations = newlabels,digits = 2)
```

# Table 2: Glycemic Control

```{r cgm summary,echo=FALSE,results='asis'}
# Import CGM
cgm = read.csv("./Jenna Sopfe/CGM Accuracy in BMT/Data_Cleaned/Prospective HSCT CGM Data Updated 6.24.19 FULL DATA SET.csv")
# Format columns
colnames(cgm)[1] <- "id"
cgm$id <- as.character(cgm$id)
cgm$Time = lubridate::mdy_hm(cgm$Time,tz = "UTC")
cgm$Tx_date = lubridate::mdy(cgm$Tx_date,tz = "UTC")
# Days from Tx
cgm$days_from_tx = 
  as.numeric(difftime(as.Date(cgm$Time),cgm$Tx_date,units = "days"))
cgm$time_cat = cut(cgm$days_from_tx,c(-Inf,-1,14,Inf),
                   labels = c("Day -7 to -1","Days 0 to 14","Day 14+"))
# Excursions above 140
cgm$exc_140 = ifelse(cgm$Historic_Glucose >= 140,1,0)
# Summary variables - excursions over 140 count if longer than 30 minutes
cgm_summ = cgm %>% group_by(id,time_cat) %>%
  summarise(avg = mean(Historic_Glucose,na.rm = T),
            cv = avg / sd(Historic_Glucose,na.rm = T),
            perc_70 = round(sum(Historic_Glucose < 70)/n(),3) * 100,
            perc_70_140 = round(sum(Historic_Glucose >= 70 & 
                                      Historic_Glucose <= 140)/n(),3)*100,
            perc_140 = round(sum(Historic_Glucose > 140)/n(),3) * 100,
            perc_200 = round(sum(Historic_Glucose > 200)/n(),3) * 100,
            num_exc_140 = 
              length(which(rle(exc_140)$lengths[which(rle(exc_140)$values == 1)] > 6)),
            n_obs = n(),
            num_days = max(abs(days_from_tx)))
# Summarize summary metrics
gluc_table = tableby(time_cat ~ avg + cv + kwt(perc_70,"median","q1q3") + 
                       kwt(perc_70_140,"median","q1q3") + 
                       kwt(perc_140,"median","q1q3") + kwt(perc_200,"median","q1q3") + 
                       kwt(num_exc_140,"median","q1q3") + 
                       kwt(n_obs,"median","q1q3") + kwt(num_days,"median","q1q3"),
                     data = cgm_summ)
summary(gluc_table,pfootnote = T,
        labelTranslations = 
          list(avg = "Mean SG (mg/dL)",cv = "CV",
               perc_70 = "% Time < 70 mg/dL",
               perc_70_140 = "% Time 70-140 mg/dL",
               perc_140 = "% Time > 140 mg/dL",
               perc_200 = "% Time > 200 mg/dL",
               num_exc_140 = "Number of Excursions > 140",
               n_obs = "Number of Observations",
               num_days = "Number of Days With CGM"))
# Add CGM summary metrics to df
# Long to wide
cgm_summ <- cgm_summ %>% 
  pivot_wider(id_cols = "id",names_from = time_cat,
              names_sep = " ",
              values_from = avg:num_days)
df <- left_join(df,cgm_summ,by = "id")
# Numeric outcome
df$infxn <- as.numeric(df$infxn)-1
df$death <- as.numeric(df$death)-1
```

# Survival Models

## Time to First Infection

### Mean glucose

Age appears to violate PH assumption, but Tanner stage doesn't. However, there are 0 participants with Tanner stage 2. Split Tanner into Tanner 1 vs. Tanner 2-5.

Post-BMT steroids are strongly associated with infection. Are people put on steroids in response to infection?

# FOR DAYS 14+ EXCLUDE EVENTS IN 0 - 14 DAY PERIOD

```{r survival,echo=FALSE,message=FALSE,fig.height=10,fig.width=10}
df$tanner_stage_binary <- df$tanner_stage
levels(df$tanner_stage_binary) <- c("unknown","Tanner 1",
                                    "Tanner 2-5","Tanner 2-5",
                                    "Tanner 2-5","Tanner 2-5")
# Cox models 
# Days -7 to -1
avg_pre_cox <- coxph(Surv(time_to_infxn,infxn) ~ `avg Day -7 to -1`+
                       tx_type+post_tx_steroids+tanner_stage_binary,
                     data = df)
# Diagnostics
d <- cox.zph(avg_pre_cox)
kable(d$table)
ggcoxzph(d)
# Days 0 to 14
avg_imm_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `avg Days 0 to 14`+
                            tx_type+post_tx_steroids+tanner_stage_binary,
                          data = df)
# Diagnostics
d <- cox.zph(avg_imm_post_cox)
kable(d$table)
ggcoxzph(d)
# Days 14+
avg_post_cox <- coxph(Surv(time_to_infxn,infxn) ~ `avg Day 14+`+
                        tx_type+post_tx_steroids+tanner_stage_binary,
                      data = df)
# Diagnostics
d <- cox.zph(avg_imm_post_cox)
kable(d$table)
ggcoxzph(d)
# Cutoffs
# Time dependent sensitivity and specificity
# timeROC package
res <- 
  SeSpPPVNPV(cutpoint=115,
             T=df$time_to_infxn,
             delta=df$infxn,marker=df$`avg Day -7 to -1`,
             cause=1,weighting="cox",
             other_markers = as.matrix(cbind(df$tanner_stage_binary,df$tx_type)),
             times=quantile(df$time_to_infxn,probs=seq(0.2,0.8,0.1)))
# tdROC package
temp <- df[,c("avg Day -7 to -1","time_to_infxn","infxn")]
temp <- temp[complete.cases(temp),]
fm <- tdROC(X=temp$`avg Day -7 to -1`,Y=temp$time_to_infxn, 
            delta = temp$infxn, tau = max(temp$time_to_infxn))
# survivalROC package
roc = 
  survivalROC(Stime=df$time_to_infxn,  
              status=df$infxn,      
              marker = df$`avg Day -7 to -1`,     
              predict.time = 365*2,span = 0.25*nrow(df)^(-0.20))
plot(roc$FP, roc$TP, type="l", xlim=c(0,1), ylim=c(0,1),   
     xlab=paste( "FP", "\n", "AUC = ",round(roc$AUC,3)), 
     ylab="TP",
     main="Mean Glucose Days -7 to -1, Method = NNE \n  Year = 2")
abline(0,1)
```