---
title: "Disglycemia in BMT Patients"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/peds/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(redcapAPI)
library(AGD)
library(arsenal)
library(skimr)
library(knitr)
library(splines)
library(JM)
library(patchwork)
library(tidyverse)
```

```{r data import,echo=FALSE,cache=TRUE}
# REDCap API data import
api <- read.table("./Jenna Sopfe/api_token.txt",header = T,sep = "\t")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = as.character(api[1,1]))
data_full <- exportRecords(rcon)
# Remove participant 10
data_full <- data_full %>% filter(study_id != "10")
# Fill down Tx date
data_full <- data_full %>% group_by(study_id) %>% fill(date_of_tx)
# Format dates
datecols = c("date_death","date_of_tx","infxn_date","sens1_place","followup")
data_full[,datecols] <- lapply(data_full[,datecols],lubridate::ymd)
```

```{r data cleaning,echo=FALSE}
# Get CGM data
cgm_data <- data_full %>% filter(is.na(redcap_repeat_instrument)) %>%
  select(study_id,date_of_tx,
         sens1_place,plt1,transfuse1,bleed1,sens1_remove,sens1_reason,
         sens2_place,plt2,transfuse2,bleed2,sens2_remove,sens2_reason,
         sens3_place,plt3,transfuse3,bleed3,sens3_remove,sens3_reason,
         sens4_place,plt4,transfuse4,bleed4,sens4_remove,sens4_reason,
         sens5_place,plt5,transfuse5,bleed5,sens5_remove,sens5_reason,
         sens6_place,plt6,transfuse6,bleed6,sens6_remove,sens6_reason)

colnames(cgm_data) <- c("study_id","date_of_tx",
                        "sens_place.1","plt.1","transfuse.1","bleed.1",
                        "sens_remove.1","sens_reason.1",
                        "sens_place.2","plt.2","transfuse.2","bleed.2",
                        "sens_remove.2","sens_reason.2",
                        "sens_place.3","plt.3","transfuse.3","bleed.3",
                        "sens_remove.3","sens_reason.3",
                        "sens_place.4","plt.4","transfuse.4","bleed.4",
                        "sens_remove.4","sens_reason.4",
                        "sens_place.5","plt.5","transfuse.5","bleed.5",
                        "sens_remove.5","sens_reason.5",
                        "sens_place.6","plt.6","transfuse.6","bleed.6",
                        "sens_remove.6","sens_reason.6")
cgm_data <- as.data.frame(cgm_data)
# Wide to long
cgm_data <- reshape(cgm_data,idvar = c("study_id","date_of_tx"),
                varying = colnames(cgm_data)[3:ncol(cgm_data)],
                direction = "long")
# Remove blank rows based on sensor placement date, format DF, calculate days of CGM wear
cgm_data <- cgm_data %>% filter(!is.na(sens_place)) %>% 
  arrange(as.numeric(study_id)) %>%
  mutate(days_worn_total = round(as.numeric(difftime(sens_remove,sens_place,
                                                     units = 'days'))),
         days_worn_pre_tx = 
           pmax(days_worn_total - 
                  pmax(round(as.numeric(difftime(sens_remove,date_of_tx,
                                                 units = 'days'))),0),0),
         days_worn_post_tx = 
           pmax(days_worn_total - 
                  pmax(round(as.numeric(difftime(date_of_tx,sens_place,
                                                 units = 'days'))),0),0))
# Platelet count variables
cgm_data$plt <- as.numeric(cgm_data$plt)
cgm_data$plt[cgm_data$transfuse == "Yes"] <- NA # Remove those with transfusion = "yes"
cgm_data$plt_bleed <- ifelse(cgm_data$bleed == "Yes",cgm_data$plt,NA)
cgm_data$plt_no_bleed <- ifelse(cgm_data$bleed == "No",cgm_data$plt,NA)
```

```{r demographics,echo=FALSE,cache=TRUE}
demographics <- data_full %>% filter(is.na(redcap_repeat_instrument))
# Combine separate race columns into one
levels(demographics$race___0) <- c(NA,"AI or AN")
levels(demographics$race___1) <- c(NA,"Asian")
levels(demographics$race___2) <- c(NA,"African American")
levels(demographics$race___3) <- c(NA,"NH or PI")
levels(demographics$race___4) <- c(NA,"White")
levels(demographics$race___5) <- c(NA,"Hispanic")
levels(demographics$race___6) <- c(NA,"Other")
demographics <- demographics %>% unite("Race",race___0,race___1,race___2,race___3,race___4,race___5,race___6)
demographics$Race <- gsub("NA","",demographics$Race)
demographics$Race <- gsub("_","",demographics$Race)
demographics$Race[which(demographics$race_multi == "yes")] <- "Multiple"
demographics$Race[which(demographics$Race == "")] <- "Unknown"
# Calculate age at transplant
demographics$AgeTx <- as.numeric(difftime(demographics$date_of_tx,
                                          demographics$date_of_birth,
                                          units = "days")) / 365.25
# Format/calculate variables for BMI calculation
demographics$sex <- demographics$gender
levels(demographics$sex) <- c("F","M")
# Calculate BMI and BMI percentile
demographics$bmi <- as.numeric(demographics$weight) / 
  ((as.numeric(demographics$height)/100)^2)
demographics$bmi_z <- y2z(demographics$bmi,x = demographics$AgeTx,
                          sex = demographics$sex,ref = get("cdc.bmi"))
demographics$bmi_percentile <- round(pnorm(demographics$bmi_z) * 100,3)
# BMI percentile groups
demographics$bmi_percentile_group <- cut(demographics$bmi_percentile,
                                         breaks = c(0,85,95,100),
                                         labels = c("<85th %ile","85th-95th %ile",
                                                    ">= 95th %ile"),
                                         right = F)
# Infer Tanner stage
demographics$tanner_stage[demographics$AgeTx >= 18] <- "5"
demographics$tanner_stage[demographics$AgeTx <= 3] <- "1"
```

# Table 1: Descriptive Characteristics

Patients aged 2 - 3 years were assumed to be Tanner stage 1, and patients aged 18+ were assumed to be Tanner stage 5. Particpant 10 was excluded from all analyses.

```{r table 1,echo=FALSE,results='asis'}
table_one <- tableby(~ notest(AgeTx,"median","q1q3") + gender + tanner_stage + 
                       bmi_percentile_group + tx_type + primary_dx_cat + spec_dx_cat, 
                     data = demographics)
newlabels <- list(AgeTx = "Age at HSCT",Gender = "Sex",
                  bmi_percentile_group = "BMI Percentile",
                  `Transplant Type` = "HSCT Type",
                  `Primary Diagnosis Category` = "Primary Diagnosis",
                  `Specific Diagnosis Category` = "Specific Diagnosis")
summary(table_one,labelTranslations = newlabels,digits = 2)
```

# Figure 1: Days from HSCT With CGM Data

```{r echo=FALSE}
# Filter out those with no removal or placement date
wear_plot_data <- cgm_data %>% filter(!is.na(sens_place) & !is.na(sens_remove))
# For each CGM wear, get the dates of each day the cgm was worn
wear_plot_data$dates <- NA
for (r in 1:nrow(wear_plot_data)) {
  wear_plot_data$dates[r] <- 
    paste0(seq(as.Date(wear_plot_data$sens_place[r]),
               as.Date(wear_plot_data$sens_remove[r]),by = "days"),collapse = ",")
}
# Split up each list of dates and make a new row for each day the CGM was worn
dates <- strsplit(wear_plot_data$dates, split = ",")
wear_plot_data <- data.frame(study_id = rep(wear_plot_data$study_id, sapply(dates, length)), 
                   day = unlist(dates))
# Format id for y axis
wear_plot_data$study_id <- as.numeric(as.character(wear_plot_data$study_id))
# Add in Tx date
wear_plot_data$date_of_tx <- 
  demographics$date_of_tx[match(wear_plot_data$study_id,
                                as.numeric(demographics$study_id))]
# Calculate number of days difference 
wear_plot_data$diff <- 
  as.numeric(difftime(wear_plot_data$day,wear_plot_data$date_of_tx,units = "days"))
# Remove duplicate rows
wear_plot_data <- unique(wear_plot_data)
# Plot
ggplot(wear_plot_data,aes(x = diff,y = study_id,group = study_id)) + 
  geom_point() + 
  ylab("Participant #") + xlab("Total Days From HSCT") + 
  theme_bw()
```

Each point indicates that the participant has CGM on a given day. Blank space indicates no CGM data.

# Joint Models

```{r infection joint data,echo=FALSE}
# Import CGM
cgm <- read.csv("./Jenna Sopfe/CGM Accuracy in BMT/Data_Cleaned/Prospective HSCT CGM Data Updated 6.24.19 FULL DATA SET.csv")
# Format columns
colnames(cgm)[1] <- "study_id"
cgm$Time <- lubridate::mdy_hm(cgm$Time,tz = "UTC")
cgm$Tx_date <- lubridate::mdy(cgm$Tx_date,tz = "UTC")
# Get earliest infection for each participant, censoring info, and covariates
events <- data_full %>% group_by(study_id) %>% 
  summarise(date_of_tx = date_of_tx[1],
            infxn_date = suppressWarnings(min(infxn_date,na.rm = T)),
            first_cgm = sens1_place[1],
            censor_date = followup[1])
events$event <- pmin(events$infxn_date,events$censor_date)
events$infection <- ifelse(!is.na(as.character(events$infxn_date)),1,0)
# Exogenous variables
events$study_id <- as.numeric(as.character(events$study_id))
demographics$study_id <- as.numeric(as.character(demographics$study_id))
dem_cov = c("sex","Race","AgeTx")
events = left_join(events,demographics[,c("study_id",dem_cov)],
                   by = "study_id")
# Combine race
events$Race <- ifelse(events$Race == "White","White","Non-white")
# Order
events <- events %>% arrange(study_id)
cgm <- cgm %>% arrange(study_id)
# Get event date and sex
cgm <- left_join(cgm,events[,c("study_id","event",dem_cov)],by = "study_id")
# Time from Tx and first CGM value
cgm <- cgm %>% group_by(study_id) %>% 
  mutate(time_from_tx = difftime(Time,Tx_date,units = "hours"),
         time_from_cgm_start = difftime(Time,Time[1],units = "hours"))
events <- events %>%
  mutate(time_to_event_from_tx = difftime(event,date_of_tx,units = "hours"),
         time_to_event_from_cgm_start = difftime(event,first_cgm,units = "hours"),
         time_first_cgm_to_tx = difftime(date_of_tx,first_cgm,units = "hours"))
# Average glucose by interval, filter truncate CGM data at event
cgm$interval <- lubridate::round_date(cgm$Time,unit = "12 hour")
cgm <- cgm %>% group_by(study_id,interval) %>%
  summarise(sg = mean(Historic_Glucose,na.rm = T),event = event[1]) %>%
  mutate(time_from_cgm_start = difftime(interval,interval[1],units = "hours")) %>%
  filter(interval < event)
```

## 12-hour Mean Glucose Over Time

```{r echo=FALSE,warning=FALSE,message=FALSE}
ggplot(cgm,aes(x = time_from_cgm_start, y = sg,group = study_id)) + 
  geom_line() + 
  theme_bw()
```

```{r infection joint models,echo=FALSE}
# Random intercept and random slope
# Tested polynomial models, but quadratic did not add to straight line model
# Random slope better than random slope and intercept
mixfit <- lme(sg ~ time_from_cgm_start,data = cgm,method = "ML",
              random = ~time_from_cgm_start|study_id)
# qqnorm(resid(mixfit)) looks iffy
# Spline model
knots_time = quantile(cgm$time_from_cgm_start,probs=c(.25, .5, .75))
mixfit_spline <- lme(sg ~ ns(time_from_cgm_start,df = 4),
                     data = cgm,method = "ML",random = ~time_from_cgm_start|study_id)
# Polynomial model
mixfit_poly <- lme(sg ~ poly(time_from_cgm_start,3),data = cgm,method = "ML",
              random = ~time_from_cgm_start|study_id)
# Survival part
survfit <- coxph(Surv(time_to_event_from_cgm_start,infection)~sex + Race + AgeTx,
                   data = events,x=T)
# Joint fit
jointFit <- jointModel(mixfit_spline, survfit, timeVar = "time_from_cgm_start",
                       method = "piecewise-PH-GH")
```

## Diagnostic Plots

### Splines

```{r echo=FALSE}
time_grid = seq(from = 0,to = max(cgm$time_from_cgm_start))
plot(cgm$time_from_cgm_start,cgm$sg,col="grey",main="Knots at Q1, Q2, and Q3",
     xlab = "Hours from CGM Start",ylab = "SG")
points(time_grid,
       predict(mixfit_spline,level = 0,
               newdata = list(time_from_cgm_start=time_grid)),
       col="darkgreen",lwd=2,type="l")
abline(v=knots_time,lty=2,col="darkgreen")
```

### Polynomial (Cubic)

```{r echo=FALSE}
time_grid = seq(from = 0,to = max(cgm$time_from_cgm_start))
plot(cgm$time_from_cgm_start,cgm$sg,col="grey",main="Cubic Model",
     xlab = "Hours from CGM Start",ylab = "SG")
points(time_grid,
       predict(mixfit_poly,level = 0,
               newdata = list(time_from_cgm_start=time_grid)),
       col="darkblue",lwd=2,type="l")
```

### Joint Model

Model code:

`r jointFit$call`

```{r echo=FALSE}
plot(jointFit)
```

## Model Results

```{r echo=FALSE}
kable(summary(jointFit)$`CoefTable-Long`,caption = "Longitudinal Process")
```

```{r echo=FALSE}
kable(summary(jointFit)$`CoefTable-Event`,caption = "Event Process")
```