---
title: "Disglycemia in BMT Patients"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/peds/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(redcapAPI)
library(arsenal)
library(skimr)
library(knitr)
library(pROC)
library(AGD)
library(tidyverse)
```

```{r data import,echo=FALSE,cache=TRUE}
# REDCap API data import
api <- read.table("./Jenna Sopfe/api_token.txt",header = T,sep = "\t")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = as.character(api[1,1]))
data_full <- exportRecords(rcon)
# Remove participant 10
data_full <- data_full %>% filter(study_id != "10")
# Fill down Tx date
data_full <- data_full %>% group_by(study_id) %>% fill(date_of_tx)
# Format dates
datecols = c("date_death","date_of_tx","infxn_date","sens1_place","followup")
data_full[,datecols] <- lapply(data_full[,datecols],lubridate::ymd)
```

```{r data cleaning,echo=FALSE}
# Get CGM data
cgm_data <- data_full %>% filter(is.na(redcap_repeat_instrument)) %>%
  select(study_id,date_of_tx,
         sens1_place,plt1,transfuse1,bleed1,sens1_remove,sens1_reason,
         sens2_place,plt2,transfuse2,bleed2,sens2_remove,sens2_reason,
         sens3_place,plt3,transfuse3,bleed3,sens3_remove,sens3_reason,
         sens4_place,plt4,transfuse4,bleed4,sens4_remove,sens4_reason,
         sens5_place,plt5,transfuse5,bleed5,sens5_remove,sens5_reason,
         sens6_place,plt6,transfuse6,bleed6,sens6_remove,sens6_reason)

colnames(cgm_data) <- c("study_id","date_of_tx",
                        "sens_place.1","plt.1","transfuse.1","bleed.1",
                        "sens_remove.1","sens_reason.1",
                        "sens_place.2","plt.2","transfuse.2","bleed.2",
                        "sens_remove.2","sens_reason.2",
                        "sens_place.3","plt.3","transfuse.3","bleed.3",
                        "sens_remove.3","sens_reason.3",
                        "sens_place.4","plt.4","transfuse.4","bleed.4",
                        "sens_remove.4","sens_reason.4",
                        "sens_place.5","plt.5","transfuse.5","bleed.5",
                        "sens_remove.5","sens_reason.5",
                        "sens_place.6","plt.6","transfuse.6","bleed.6",
                        "sens_remove.6","sens_reason.6")
cgm_data <- as.data.frame(cgm_data)
# Wide to long
cgm_data <- reshape(cgm_data,idvar = c("study_id","date_of_tx"),
                varying = colnames(cgm_data)[3:ncol(cgm_data)],
                direction = "long")
# Remove blank rows based on sensor placement date, format DF, calculate days of CGM wear
cgm_data <- cgm_data %>% filter(!is.na(sens_place)) %>% 
  arrange(as.numeric(study_id)) %>%
  mutate(days_worn_total = round(as.numeric(difftime(sens_remove,sens_place,
                                                     units = 'days'))),
         days_worn_pre_tx = 
           pmax(days_worn_total - 
                  pmax(round(as.numeric(difftime(sens_remove,date_of_tx,
                                                 units = 'days'))),0),0),
         days_worn_post_tx = 
           pmax(days_worn_total - 
                  pmax(round(as.numeric(difftime(date_of_tx,sens_place,
                                                 units = 'days'))),0),0))
# Platelet count variables
cgm_data$plt <- as.numeric(cgm_data$plt)
cgm_data$plt[cgm_data$transfuse == "Yes"] <- NA # Remove those with transfusion = "yes"
cgm_data$plt_bleed <- ifelse(cgm_data$bleed == "Yes",cgm_data$plt,NA)
cgm_data$plt_no_bleed <- ifelse(cgm_data$bleed == "No",cgm_data$plt,NA)
```

```{r demographics,echo=FALSE,cache=TRUE}
demographics <- data_full %>% filter(is.na(redcap_repeat_instrument))
# Combine separate race columns into one
levels(demographics$race___0) <- c(NA,"AI or AN")
levels(demographics$race___1) <- c(NA,"Asian")
levels(demographics$race___2) <- c(NA,"African American")
levels(demographics$race___3) <- c(NA,"NH or PI")
levels(demographics$race___4) <- c(NA,"White")
levels(demographics$race___5) <- c(NA,"Hispanic")
levels(demographics$race___6) <- c(NA,"Other")
demographics <- demographics %>% unite("Race",race___0,race___1,race___2,race___3,race___4,race___5,race___6)
demographics$Race <- gsub("NA","",demographics$Race)
demographics$Race <- gsub("_","",demographics$Race)
demographics$Race[which(demographics$race_multi == "yes")] <- "Multiple"
demographics$Race[which(demographics$Race == "")] <- "Unknown"
# Calculate age at transplant
demographics$AgeTx <- as.numeric(difftime(demographics$date_of_tx,
                                          demographics$date_of_birth,
                                          units = "days")) / 365.25
# Format/calculate variables for BMI calculation
demographics$sex <- demographics$gender
levels(demographics$sex) <- c("F","M")
# Calculate BMI and BMI percentile
demographics$bmi <- as.numeric(demographics$weight) / 
  ((as.numeric(demographics$height)/100)^2)
demographics$bmi_z <- y2z(demographics$bmi,x = demographics$AgeTx,
                          sex = demographics$sex,ref = get("cdc.bmi"))
demographics$bmi_percentile <- round(pnorm(demographics$bmi_z) * 100,3)
# BMI percentile groups
demographics$bmi_percentile_group <- cut(demographics$bmi_percentile,
                                         breaks = c(0,85,95,100),
                                         labels = c("<85th %ile","85th-95th %ile",
                                                    ">= 95th %ile"),
                                         right = F)
# Infer Tanner stage
demographics$tanner_stage[demographics$AgeTx >= 18] <- "5"
demographics$tanner_stage[demographics$AgeTx <= 3] <- "1"
# Numeric study id
demographics$study_id = as.numeric(as.character(demographics$study_id))
# Get earliest infection for each participant, censoring info, and covariates
events = data_full %>% group_by(study_id) %>% 
  summarise(infxn_date = suppressWarnings(min(infxn_date,na.rm = T)),
            death_date = suppressWarnings(min(date_death,na.rm = T)))
events$Infection = factor(ifelse(!is.na(as.character(events$infxn_date)),1,0),
                             levels = 0:1,labels = c("No","Yes"))
events$Death = factor(ifelse(!is.na(as.character(events$death_date)),1,0),
                      levels = 0:1,labels = c("No","Yes"))
events$study_id = as.numeric(as.character(events$study_id))
events = events %>% arrange(study_id) %>% dplyr::select(study_id,Infection,Death)
# Add to demographics
demographics = left_join(demographics,events,by = "study_id")
```

# Table 1a: Descriptive Characteristics by Event Status - Infection

Patients aged 2 - 3 years were assumed to be Tanner stage 1, and patients aged 18+ were assumed to be Tanner stage 5. Particpant 10 was excluded from all analyses.

```{r table 1a,echo=FALSE,results='asis'}
table_one <- tableby(Infection ~ notest(AgeTx,"median","q1q3") + gender + 
                       tanner_stage + bmi_percentile_group + tx_type + 
                       primary_dx_cat + spec_dx_cat, 
                     data = demographics)
newlabels <- list(AgeTx = "Age at HSCT",Gender = "Sex",
                  bmi_percentile_group = "BMI Percentile",
                  `Transplant Type` = "HSCT Type",
                  `Primary Diagnosis Category` = "Primary Diagnosis",
                  `Specific Diagnosis Category` = "Specific Diagnosis")
summary(table_one,labelTranslations = newlabels,digits = 2)
```

# Table 1a: Descriptive Characteristics by Event Status - Death

```{r table 1b,echo=FALSE,results='asis'}
table_one <- tableby(Death ~ notest(AgeTx,"median","q1q3") + gender + 
                       tanner_stage + bmi_percentile_group + tx_type + 
                       primary_dx_cat + spec_dx_cat, 
                     data = demographics)
newlabels <- list(AgeTx = "Age at HSCT",Gender = "Sex",
                  bmi_percentile_group = "BMI Percentile",
                  `Transplant Type` = "HSCT Type",
                  `Primary Diagnosis Category` = "Primary Diagnosis",
                  `Specific Diagnosis Category` = "Specific Diagnosis")
summary(table_one,labelTranslations = newlabels,digits = 2)
```

# Figure 1: Days from HSCT With CGM Data

```{r echo=FALSE,dpi=600}
# Filter out those with no removal or placement date
wear_plot_data <- cgm_data %>% filter(!is.na(sens_place) & !is.na(sens_remove))
# For each CGM wear, get the dates of each day the cgm was worn
wear_plot_data$dates <- NA
for (r in 1:nrow(wear_plot_data)) {
  wear_plot_data$dates[r] <- 
    paste0(seq(as.Date(wear_plot_data$sens_place[r]),
               as.Date(wear_plot_data$sens_remove[r]),by = "days"),collapse = ",")
}
# Split up each list of dates and make a new row for each day the CGM was worn
dates <- strsplit(wear_plot_data$dates, split = ",")
wear_plot_data <- data.frame(study_id = rep(wear_plot_data$study_id, sapply(dates, length)), 
                   day = unlist(dates))
# Format id for y axis
wear_plot_data$study_id <- as.numeric(as.character(wear_plot_data$study_id))
# Add in Tx date
wear_plot_data$date_of_tx <- 
  demographics$date_of_tx[match(wear_plot_data$study_id,
                                as.numeric(demographics$study_id))]
# Calculate number of days difference 
wear_plot_data$diff <- 
  as.numeric(difftime(wear_plot_data$day,wear_plot_data$date_of_tx,units = "days"))
# Remove duplicate rows
wear_plot_data <- unique(wear_plot_data)
# Plot
ggplot(wear_plot_data,aes(x = diff,y = study_id,group = study_id)) + 
  geom_point() + 
  ylab("Participant #") + xlab("Total Days From HSCT") + 
  theme_bw()
```

Each point indicates that the participant has CGM on a given day. Blank space indicates no CGM data.

# Table 2: Glycemic Control

```{r cgm summary,echo=FALSE,results='asis'}
# Import CGM
cgm = read.csv("./Jenna Sopfe/CGM Accuracy in BMT/Data_Cleaned/Prospective HSCT CGM Data Updated 6.24.19 FULL DATA SET.csv")
# Format columns
colnames(cgm)[1] <- "study_id"
cgm$Time = lubridate::mdy_hm(cgm$Time,tz = "UTC")
cgm$Tx_date = lubridate::mdy(cgm$Tx_date,tz = "UTC")
# Days from Tx
cgm$days_from_tx = as.numeric(difftime(as.Date(cgm$Time),cgm$Tx_date,units = "days"))
cgm$time_cat = cut(cgm$days_from_tx,c(-Inf,-1,14,Inf),
                   labels = c("Day -7 to -1","Days 0 to 14","Day 14+"))
# Excursions above 140
cgm$exc_140 = ifelse(cgm$Historic_Glucose >= 140,1,0)
# Summary variables - excursions over 140 count if longer than 30 minutes
cgm_summ = cgm %>% group_by(study_id,time_cat) %>%
  summarise(avg = mean(Historic_Glucose,na.rm = T),
            cv = avg / sd(Historic_Glucose,na.rm = T),
            perc_70 = round(sum(Historic_Glucose < 70)/n(),3) * 100,
            perc_70_140 = round(sum(Historic_Glucose >= 70 & 
                                      Historic_Glucose <= 140)/n(),3) * 100,
            perc_140 = round(sum(Historic_Glucose > 140)/n(),3) * 100,
            perc_200 = round(sum(Historic_Glucose > 200)/n(),3) * 100,
            num_exc_140 = 
              length(which(rle(exc_140)$lengths[which(rle(exc_140)$values == 1)] > 6)),
            n_obs = n(),num_days = max(abs(days_from_tx)))
# Summarize summary metrics
gluc_table = tableby(time_cat ~ avg + cv + kwt(perc_70,"median","q1q3") + 
                       kwt(perc_70_140,"median","q1q3") + 
                       kwt(perc_140,"median","q1q3") + kwt(perc_200,"median","q1q3") + 
                       kwt(num_exc_140,"median","q1q3") + 
                       kwt(n_obs,"median","q1q3") + kwt(num_days,"median","q1q3"),
                     data = cgm_summ)
summary(gluc_table,pfootnote = T,
        labelTranslations = list(avg = "Mean SG (mg/dL)",cv = "CV",
                                 perc_70 = "% Time < 70 mg/dL",
                                 perc_70_140 = "% Time 70-140 mg/dL",
                                 perc_140 = "% Time > 140 mg/dL",
                                 perc_200 = "% Time > 200 mg/dL",
                                 num_exc_140 = "Number of Excursions > 140",
                                 n_obs = "Number of Observations",
                                 num_days = "Number of Days With CGM"))
```

# ROC Analysis

```{r roc data,echo=FALSE}
# Add event data
cgm_summ = left_join(cgm_summ,events,by = "study_id")
# Split by timepoint
pre_tx = cgm_summ %>% filter(time_cat == "Day -7 to -1")
imm_post_tx = cgm_summ %>% filter(time_cat == "Days 0 to 14")
post_tx = cgm_summ %>% filter(time_cat == "Day 14+")
```

```{r roc function,echo=FALSE}
roc_func = function(outcome,predictor,predictor_name,
                    leg_x = 1.32,leg_y = 0.95,best = "youden",
                    ret = c("threshold","specificity","sensitivity")){
  form = as.formula(paste0(outcome,"~",predictor))
  # ROC curves
  pre_roc = roc(form,data = pre_tx,quiet = T)
  imm_post_roc = roc(form,data = imm_post_tx,quiet = T)
  post_roc = roc(form,data = post_tx,quiet = T)
  # Plot
  plot(pre_roc, col = "blue",main = predictor_name)
  plot(imm_post_roc,col = "dark green",add = T)
  plot(post_roc,col = "red",add = T)
  legend(leg_x,leg_y,legend=levels(cgm_summ$time_cat),
         col=c("blue","dark green","red"), lty=1)
  # Format results
  res_pre = coords(pre_roc,"best",transpose = F,best.method = best)
  res_pre$time = levels(cgm_summ$time_cat)[1]
  res_pre$auc = as.numeric(pre_roc$auc)
  res_imm_post = coords(imm_post_roc,"best",transpose = F,best.method = best)
  res_imm_post$time = levels(cgm_summ$time_cat)[2]
  res_imm_post$auc = as.numeric(imm_post_roc$auc)
  res_post = coords(post_roc,"best",transpose = F,best.method = best)
  res_post$time = levels(cgm_summ$time_cat)[3]
  res_post$auc = as.numeric(post_roc$auc)
  results = bind_rows(res_pre,res_imm_post,res_post)
  results = results[,c("time","threshold","specificity","sensitivity","auc")]
  results[,2:5] = lapply(results[,2:5],function(x){round(x,2)})
  kable(results,row.names = T)
}
```

## Infection

Optimal cutoffs for each CGM metric...

```{r infection roc,echo=FALSE}
# ROC for infection and each CGM metric
roc_func(outcome = "Infection",predictor = "avg",
         predictor_name = "Average SG (mg/dL)")

roc_func(outcome = "Infection",predictor = "cv",
         predictor_name = "CV")

roc_func(outcome = "Infection",predictor = "perc_70",
         predictor_name = "% Time < 70 mg/dL")

roc_func(outcome = "Infection",predictor = "perc_70_140",
         predictor_name = "% Time 70-140 mg/dL")

roc_func(outcome = "Infection",predictor = "perc_140",
         predictor_name = "% Time > 140 mg/dL")

roc_func(outcome = "Infection",predictor = "perc_200",
         predictor_name = "% Time > 200 mg/dL")

roc_func(outcome = "Infection",predictor = "num_exc_140",
         predictor_name = "Number of Excursions > 140")
```

## Death

```{r death roc,echo=FALSE}
# ROC for death and each CGM metric
roc_func(outcome = "Death",predictor = "avg",
         predictor_name = "Average SG (mg/dL)")

roc_func(outcome = "Death",predictor = "cv",
         predictor_name = "CV")

roc_func(outcome = "Death",predictor = "perc_70",
         predictor_name = "% Time < 70 mg/dL")

roc_func(outcome = "Death",predictor = "perc_70_140",
         predictor_name = "% Time 70-140 mg/dL")

roc_func(outcome = "Death",predictor = "perc_140",
         predictor_name = "% Time > 140 mg/dL")

roc_func(outcome = "Death",predictor = "perc_200",
         predictor_name = "% Time > 200 mg/dL")

roc_func(outcome = "Death",predictor = "num_exc_140",
         predictor_name = "Number of Excursions > 140")
```

