---
title: "CGM and BMT"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(redcapAPI)
library(lubridate)
library(plotly)
library(reshape2)
library(knitr)
library(tidyverse)
library(arsenal)
knitr::opts_chunk$set(echo = F)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/BDC/Jenna Sopfe")
```

```{r data import}
# REDCap import
api <- read.table("./api_token.txt",header = T,sep = "\t")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = as.character(api[1,1]))
data_full <- exportRecords(rcon)
# Remove participant 10
data_full <- data_full %>% filter(study_id != "10")
# Fill down Tx date
data_full <- data_full %>% group_by(study_id) %>% fill(date_of_tx)
# Format dates
datecols = c("time_engraft","time_plt_engraft")
data_full[,datecols] <- lapply(data_full[,datecols],lubridate::ymd)
data_full$study_id = as.numeric(data_full$study_id)
# Calculate BMI
data_full$bmi = data_full$weight/((data_full$height/100)^2)
# Count infections, TPN
data_full = data_full %>% group_by(study_id) %>% 
  mutate(infections = sum(!is.na(infxn_type)))
data_full$infections_yn = ifelse(data_full$infections>0,"Yes","No")
# Clean
cgm = read.csv("./CGM Accuracy in BMT/Data_Cleaned/Prospective HSCT CGM Data Updated 6.24.19 FULL DATA SET.csv",na.strings = "")
cgm$Time = mdy_hm(cgm$Time)
cgm$Tx_date = mdy(cgm$Tx_date)
cgm$dx_cat = factor(cgm$dx_cat)
cgm$tx_type = factor(cgm$tx_type)
# Get other dates
dates = data_full %>% select(study_id,bmi,all_of(datecols)) %>% 
  group_by(study_id) %>% fill(all_of(datecols),.direction = "updown") %>%
  filter(row_number() == 1)
cgm = left_join(cgm,dates,by = c("SubjectID" = "study_id"))
cgm$time_engraft_minus_3 = cgm$time_engraft - 3
# Time since engraftment
cgm$time_from_engraft = as.numeric(difftime(cgm$Time,cgm$time_engraft_minus_3,units = "mins"))
# Wide format for SAS
cgm_wide = cgm %>% filter(time_from_engraft >= 0) %>%
  group_by(SubjectID) %>% mutate(n = row_number()) %>% 
  select(SubjectID,tx_type,dx_cat,bmi,Historic_Glucose,time_from_engraft,n) %>%
  pivot_wider(names_from = n,
              values_from = c("Historic_Glucose","time_from_engraft")) %>%
  filter(!is.na(tx_type) & !is.na(dx_cat)) %>% ungroup()
# Write CSV for checking with SAS
write.csv(cgm_wide,file = "./CGM and BMT FDA/Data_Cleaned/cleaned_for_traj.csv",
          row.names = F,na = "")
# Convert to mmol/L
cgm$mmol = cgm$Historic_Glucose / 18.0182
# Split by ID and sensor number
cgm_list = split.data.frame(cgm,list(cgm$SubjectID))
cgm_list = lapply(cgm_list, function(df){
  df = df %>% arrange(Time)
  df
})
tg3.9 = lapply(cgm_list, function(df){
  mean(df$mmol < 3.9,na.rm = T) * 24
}) 
tg8.9 = lapply(cgm_list, function(df){
  mean(df$mmol > 8.9,na.rm = T) * 24
})
mbg = lapply(cgm_list, function(df){
  mean(df$mmol,na.rm = T)
})
r = lapply(cgm_list, function(df){
  abs(diff(range(df$mmol,na.rm = T)))
})
modd = lapply(cgm_list, function(df){
  df$round_time = lubridate::round_date(df$Time,unit = "5 minutes")
  df$round_time = sapply(strsplit(as.character(df$round_time)," "),"[[",2)
  modd = df %>% group_by(round_time) %>% 
    mutate(t_diff = abs(mmol - lag(mmol)))
  modd = mean(modd$t_diff,na.rm = T)
  modd
})
# Q score for each sensor - calculate hours per day from percent time
q_scores = lapply(cgm_list, function(df){
  tg3.9 = mean(df$mmol < 3.9,na.rm = T) * 24
  tg8.9 = mean(df$mmol > 8.9,na.rm = T) * 24
  mbg = mean(cgm$mmol,na.rm = T)
  r = abs(diff(range(cgm$mmol,na.rm = T)))
  df$round_time = lubridate::round_date(df$Time,unit = "5 minutes")
  df$round_time = sapply(strsplit(as.character(df$round_time)," "),"[[",2)
  modd = df %>% group_by(round_time) %>% 
    mutate(t_diff = abs(mmol - lag(mmol)))
  modd = mean(modd$t_diff,na.rm = T)
  q = 8 + (mbg - 7.8)/1.7 + (r - 7.5)/2.9 + (tg3.9-0.6)/1.2 + (tg8.9 - 6.2)/5.7 + 
    (modd - 1.8)/0.9
  return(q)
})
q_scores = data.frame(do.call(rbind,q_scores))
colnames(q_scores) = "q_score"
q_scores$study_id = as.numeric(rownames(q_scores))
# Mean, SD, and CV
means = lapply(cgm_list, function(df){
  return(mean(df$Historic_Glucose,na.rm = T))
})
means = data.frame(do.call(rbind,means))
colnames(means) = "mean_glucose"
sds = lapply(cgm_list, function(df){
  return(sd(df$Historic_Glucose,na.rm = T))
})
sds = data.frame(do.call(rbind,sds))
colnames(sds) = "sd_glucose"
cvs = lapply(cgm_list, function(df){
  m = mean(df$Historic_Glucose,na.rm = T)
  s = sd(df$Historic_Glucose,na.rm = T)
  return(s/m)
})
cvs = data.frame(do.call(rbind,cvs))
colnames(cvs) = "cv_glucose"
# Add to Q scores
q_scores = data.frame(do.call(cbind,c(q_scores,means,sds,cvs)))
# Merge into full data
data_full = left_join(data_full,q_scores,by = "study_id")
# Get each person's first row, fix race, etc.
df = data_full %>% group_by(study_id) %>% filter(is.na(redcap_repeat_instance))
levels(df$race___0) <- c(NA,"AI or AN")
levels(df$race___1) <- c(NA,"Asian")
levels(df$race___2) <- c(NA,"African American")
levels(df$race___3) <- c(NA,"NH or PI")
levels(df$race___4) <- c(NA,"White")
levels(df$race___5) <- c(NA,"Hispanic")
levels(df$race___6) <- c(NA,"Other")
df <- df %>% ungroup() %>% 
  unite("Race",race___0,race___1,race___2,race___3,race___4,race___5,race___6)
df$Race <- gsub("NA","",df$Race)
df$Race <- gsub("_","",df$Race)
df$Race[which(df$race_multi == "yes")] <- "Multiple"
df$Race[which(df$Race == "")] <- "Unknown"
df$AgeTx <- 
  as.numeric(difftime(df$date_of_tx,df$date_of_birth,
                      units = "days")) / 365.25
```

# Table 1: Participant Characteristics

Participant 21 has a high Q score, but this appears to be correct and largely due to spending a lot of time below 3.9 mmol/L.

```{r results='asis'}
t1 = tableby(~notest(q_score,"N", "median", "q1q3")+ notest(mean_glucose,"N", "median", "q1q3") + notest(sd_glucose,"N", "median", "q1q3") +notest(cv_glucose,"N", "median", "q1q3") + AgeTx+gender+Race+tanner_stage+tx_type+
               primary_dx_cat+spec_dx_cat+prep_regimen+sec_icu+factor(infections)+
               post_tx_tpn,
             data = df)
summary(t1,labelTranslations = list(AgeTx = "Age at Tx",q_score = "Q Score",
                                    `factor(infections)` = "Num. Infections"))
```

# Q score component quantiles

```{r results='asis'}
li = list(tg3.9,tg8.9,mbg,r,modd)
names(li) = c("tg<3.9","tg>8.9","MBG","range","MODD")
l = lapply(li, function(x){
  d = unlist(x)
  quantile(d)
})
l = as.data.frame(l)
kable(l,digits = 2)
```

# Q Score by Group

```{r}
# Any ICU admission
df$any_icu = factor(rowSums(df[,c("prim_icu","sec_icu")]=="Yes") > 0)
levels(df$any_icu) = c("No","Yes")
```

Q score was compared between groups using a Wilcoxon rank sum test.

```{r comparison function}
compare = function(data,group){
  # Test
  f = as.formula(paste0("q_score ~ ",group))
  w = wilcox.test(f,data = data)
  p = w$p.value
  p_form = format.pval(p,digits = 3,eps = 0.001)
  # Plot
  plot = ggplot(data,aes_string(x = group,y = "q_score",fill = group)) + 
    geom_boxplot() + ylab("Q score") + theme_bw() + theme(legend.position = "none")
  # Print results
  cat("\n")
  cat("\n")
  cat(paste("##",group))
  cat("\n")
  cat("\n")
  print(plot)
  cat("\n")
  cat("\n")
  # Medians
  data %>% group_by_(group) %>% 
    summarise(median = median(q_score,na.rm=T),
              q1q3 = paste0(round(quantile(q_score,na.rm=T)[2],2),
                            " - ",
                            round(quantile(q_score,na.rm=T)[4],2))) %>% 
    kable(.,digits = 2) %>% print(.)
  cat("\n")
  cat("\n")
  if(p > 0.05){
    cat(paste0("Q score was not significantly different by ",group," (p = ",p_form,")."))
  } else if (p <= 0.05){
    cat(paste0("Q score was significantly different by ",group," (p = ",p_form,")."))
  } else {
    cat("Test failed.")
  }
}
```

```{r results='asis'}
vars = c("tx_type","primary_dx_cat","prep_regimen","any_icu","infections_yn","post_tx_tpn")
for (v in vars) {
  compare(df,v)
}
```

# Mean Glucose by Group

Mean glucose was compared between groups using a Wilcoxon rank sum test.

```{r}
compare = function(data,group){
  # Test
  f = as.formula(paste0("mean_glucose ~ ",group))
  w = wilcox.test(f,data = data)
  p = w$p.value
  p_form = format.pval(p,digits = 3,eps = 0.001)
  # Plot
  plot = ggplot(data,aes_string(x = group,y = "mean_glucose",fill = group)) + 
    geom_boxplot() + ylab("Mean glucose") + theme_bw() + theme(legend.position = "none")
  # Print results
  cat("\n")
  cat("\n")
  cat(paste("##",group))
  cat("\n")
  cat("\n")
  print(plot)
  cat("\n")
  cat("\n")
  # Medians
  data %>% group_by_(group) %>% 
    summarise(median = median(mean_glucose,na.rm=T),
              q1q3 = paste0(round(quantile(mean_glucose,na.rm=T)[2],2),
                            " - ",
                            round(quantile(mean_glucose,na.rm=T)[4],2))) %>% 
    kable(.,digits = 2) %>% print(.)
  cat("\n")
  cat("\n")
  if(p > 0.05){
    cat(paste0("Mean glucose was not significantly different by ",group," (p = ",p_form,")."))
  } else if (p <= 0.05){
    cat(paste0("Mean glucose was significantly different by ",group," (p = ",p_form,")."))
  } else {
    cat("Test failed.")
  }
}
```

```{r results='asis'}
vars = c("tx_type","primary_dx_cat","prep_regimen","any_icu","infections_yn","post_tx_tpn")
for (v in vars) {
  compare(df,v)
}
```

# Glucose SD by Group

Glucose SD was compared between groups using a Wilcoxon rank sum test.

```{r}
compare = function(data,group){
  # Test
  f = as.formula(paste0("sd_glucose ~ ",group))
  w = wilcox.test(f,data = data)
  p = w$p.value
  p_form = format.pval(p,digits = 3,eps = 0.001)
  # Plot
  plot = ggplot(data,aes_string(x = group,y = "sd_glucose",fill = group)) + 
    geom_boxplot() + ylab("Glucose SD") + theme_bw() + theme(legend.position = "none")
  # Print results
  cat("\n")
  cat("\n")
  cat(paste("##",group))
  cat("\n")
  cat("\n")
  print(plot)
  cat("\n")
  cat("\n")
  # Medians
  data %>% group_by_(group) %>% 
    summarise(median = median(sd_glucose,na.rm=T),
              q1q3 = paste0(round(quantile(sd_glucose,na.rm=T)[2],2),
                            " - ",
                            round(quantile(sd_glucose,na.rm=T)[4],2))) %>% 
    kable(.,digits = 2) %>% print(.)
  cat("\n")
  cat("\n")
  if(p > 0.05){
    cat(paste0("Glucose SD was not significantly different by ",group," (p = ",p_form,")."))
  } else if (p <= 0.05){
    cat(paste0("Glucose SD was significantly different by ",group," (p = ",p_form,")."))
  } else {
    cat("Test failed.")
  }
}
```

```{r results='asis'}
vars = c("tx_type","primary_dx_cat","prep_regimen","any_icu","infections_yn","post_tx_tpn")
for (v in vars) {
  compare(df,v)
}
```

# Glucose CV by Group

Glucose CV was compared between groups using a Wilcoxon rank sum test.

```{r}
compare = function(data,group){
  # Test
  f = as.formula(paste0("cv_glucose ~ ",group))
  w = wilcox.test(f,data = data)
  p = w$p.value
  p_form = format.pval(p,digits = 3,eps = 0.001)
  # Plot
  plot = ggplot(data,aes_string(x = group,y = "cv_glucose",fill = group)) + 
    geom_boxplot() + ylab("Glucose CV") + theme_bw() + theme(legend.position = "none")
  # Print results
  cat("\n")
  cat("\n")
  cat(paste("##",group))
  cat("\n")
  cat("\n")
  print(plot)
  cat("\n")
  cat("\n")
  # Medians
  data %>% group_by_(group) %>% 
    summarise(median = median(cv_glucose,na.rm=T),
              q1q3 = paste0(round(quantile(cv_glucose,na.rm=T)[2],2),
                            " - ",
                            round(quantile(cv_glucose,na.rm=T)[4],2))) %>% 
    kable(.,digits = 2) %>% print(.)
  cat("\n")
  cat("\n")
  if(p > 0.05){
    cat(paste0("Glucose CV was not significantly different by ",group," (p = ",p_form,")."))
  } else if (p <= 0.05){
    cat(paste0("Glucose CV was significantly different by ",group," (p = ",p_form,")."))
  } else {
    cat("Test failed.")
  }
}
```

```{r results='asis'}
vars = c("tx_type","primary_dx_cat","prep_regimen","any_icu","infections_yn","post_tx_tpn")
for (v in vars) {
  compare(df,v)
}
```

