---
title: "BMT and CGM"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Work/Jenna Sopfe")
library(redcapAPI)
library(tableone)
library(arsenal)
library(skimr)
library(knitr)
library(AGD)
library(ega)
library(blandr)
library(tidyverse)
```

```{r include=FALSE,cache=TRUE}
# REDCap API data import
api <- read.table("./api_token.txt",header = T,sep = "\t")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = as.character(api[1,1]))
data_full <- exportRecords(rcon)
```

```{r }
# Import Jenna and Greg's visual analysis
patterns = read.csv("./Disglycemia in BMT/Data_Cleaned/graph_patterns.csv",
                    na.strings = c("","NA"))
keep = patterns$Subject[patterns$Suffucient.Data == "Y"]
keep = keep[!is.na(keep)]
patterns$pre.HSCT.low.then.increases.post.day.0 = NULL
patterns$pre.HSCT.no.change..increase.post.HSCT = NULL
# Remove participant 10
data_full <- data_full %>% filter(study_id %in% keep)
# Get demographic info
demographics <- data_full %>% group_by(study_id) %>% filter(!is.na(mrn))
# Combine separate race columns into one
levels(demographics$race___0) <- c(NA,"AI or AN")
levels(demographics$race___1) <- c(NA,"Asian")
levels(demographics$race___2) <- c(NA,"African American")
levels(demographics$race___3) <- c(NA,"NH or PI")
levels(demographics$race___4) <- c(NA,"White")
levels(demographics$race___5) <- c(NA,"Hispanic")
levels(demographics$race___6) <- c(NA,"Other")
demographics <- demographics %>% ungroup() %>% unite("Race",race___0,race___1,race___2,race___3,race___4,race___5,race___6)
demographics$Race <- gsub("NA","",demographics$Race)
demographics$Race <- gsub("_","",demographics$Race)
demographics$Race[which(demographics$race_multi == "yes")] <- "Multiple"
demographics$Race[which(demographics$Race == "")] <- "Unknown"
# Calculate age at transplant
demographics$AgeTx <- as.numeric(difftime(demographics$date_of_tx,demographics$date_of_birth,units = "days")) / 365.25
# Format/calculate variables for BMI calculation
demographics$sex <- demographics$gender
levels(demographics$sex) <- c("F","M")
# Calculate BMI and BMI percentile
demographics$bmi <- as.numeric(demographics$weight) / ((as.numeric(demographics$height)/100)^2)
demographics$bmi_z <- y2z(demographics$bmi,x = demographics$AgeTx,sex = demographics$sex,ref = get("cdc.bmi"))
demographics$bmi_percentile <- round(pnorm(demographics$bmi_z) * 100,3)
# BMI percentile groups
demographics$bmi_percentile_group <- cut(demographics$bmi_percentile,
                                         breaks = c(0,85,95,100),
                                         labels = c("<85th %ile","85th-95th %ile",
                                                    ">= 95th %ile"),
                                         right = F)
```

## Table 1: Descriptive Characteristics

Participants without sufficient data excluded from the following analyses. 

```{r }
dem_vars <- c("AgeTx","gender","Race","tanner_stage","bmi_percentile_group","tx_type","primary_dx_cat","spec_dx_cat")
t1 <- CreateTableOne(dem_vars,data = demographics)
t1 <- as.data.frame(print(t1, nonnormal = c("AgeTx"),printToggle = F,showAllLevels = T))
t1 <- cbind(new = NA,t1)
colnames(t1) <- c("","Level","Overall")
t1[,1] <- c("n","Age at HSCT (median [IQR])",
            "Gender (%)","","Race (%)","","","","","",
            "Tanner Stage (%)","","","","","",
            "BMI Percentile","","",
            "HSCT Type (%)","",
            "Primary Diagnosis (%)","",
            "Specific Diagnosis (%)","","","","","","","","")
kable(t1,row.names = F)
```

# Pattern Analysis

Unfortunately, many of the graph patterns do not have enough information to analyze. For example, both participants with peri-HSCT increases had allogenic HSCTs, so it is not really possible to examine trends by transplant type. Often cases like this will result in a p value of 1.

Because we are only comparing two groups (i.e. "pattern" vs. "no pattern"), a linear model ANOVA is the same as a t test.

```{r}
patterns = patterns[patterns$Suffucient.Data == "Y",]
outcome_vars = colnames(patterns)[7:ncol(patterns)]
# Add variables
patterns$bmi_percentile = demographics$bmi_percentile[match(patterns$Subject,demographics$study_id)]
patterns$primary_dx_cat = demographics$primary_dx_cat[match(patterns$Subject,demographics$study_id)]
# Steroid use within 1 week engraftment
data_full$steroid_post_engraft = NA
data_full$steroid_post_engraft[!is.na(data_full$steroid_dates)] = "No"
data_full$steroid_post_engraft[18] = "Yes"
data_full$steroid_post_engraft[59] = "Yes"
yes = as.character(data_full$study_id[which(data_full$steroid_post_engraft == "Yes")])
no = as.character(data_full$study_id[which(data_full$steroid_post_engraft == "No")])
patterns$steroid_post_engraft = "No"
patterns$steroid_post_engraft[patterns$Subject %in% yes] = "Yes"

trend_vars = c(Age = "Age",bmi_percentile = "BMI %ile",HSCT.Type = "HSCT type",
               Prep.Regimen = "Prep. regimen",primary_dx_cat = "Primary diagnosis")

right_form = paste0(names(trend_vars),collapse = "+")
```

## Peri-HSCT increase

```{r results='asis'}
patterns$peri.HSCT.increase = factor(patterns$peri.HSCT.increase,labels = c("No","Yes"))
form = as.formula(paste0("peri.HSCT.increase~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## Pre-HSCT increased until day 0

```{r results='asis'}
patterns$Pre.HSCT.increased.until.day.0 = factor(patterns$Pre.HSCT.increased.until.day.0,labels = c("No","Yes"))
form = as.formula(paste0("Pre.HSCT.increased.until.day.0~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## Pre-HSCT high/rise with fall pre day-0

```{r results='asis'}
patterns$pre.HSCT.high.rise.with.fall.pre.day.0 = 
  factor(patterns$pre.HSCT.high.rise.with.fall.pre.day.0,labels = c("No","Yes"))
form = as.formula(paste0("pre.HSCT.high.rise.with.fall.pre.day.0~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## No identifiable peri-HSCT pattern

```{r results='asis'}
patterns$no.identifiable.peri.HSCT.pattern = 
  factor(patterns$no.identifiable.peri.HSCT.pattern,labels = c("No","Yes"))
form = as.formula(paste0("no.identifiable.peri.HSCT.pattern~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## Other

```{r results='asis'}
patterns$other = 
  factor(patterns$other,labels = c("No","Yes"))
form = as.formula(paste0("other~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## Pre-engraftment increase then decrease post-eng.

```{r results='asis'}
patterns$pre.engraftment.increase.then.decrease.post.eng. = 
  factor(patterns$pre.engraftment.increase.then.decrease.post.eng.,labels = c("No","Yes"))
form = as.formula(paste0("pre.engraftment.increase.then.decrease.post.eng.~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## Peri-engraftment increase

```{r}
trend_vars = c(Age = "Age",bmi_percentile = "BMI %ile",HSCT.Type = "HSCT type",
               Prep.Regimen = "Prep. regimen",primary_dx_cat = "Primary diagnosis",
               steroid_post_engraft = "Post-transplant steroids within 1 week of engraftment")

right_form = paste0(names(trend_vars),collapse = "+")
```

```{r results='asis'}
patterns$peri.engraftment.increase = 
  factor(patterns$peri.engraftment.increase,labels = c("No","Yes"))
form = as.formula(paste0("peri.engraftment.increase~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## Peri-engraftment decrease

```{r results='asis'}
patterns$peri.engraftment.decrease = 
  factor(patterns$peri.engraftment.decrease,labels = c("No","Yes"))
form = as.formula(paste0("peri.engraftment.decrease~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## No identifiable peri-engraft pattern

```{r results='asis'}
patterns$no.identifiable.peri.engraft.pattern = 
  factor(patterns$no.identifiable.peri.engraft.pattern,labels = c("No","Yes"))
form = as.formula(paste0("no.identifiable.peri.engraft.pattern~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```

## Peri-engraft other

```{r results='asis'}
patterns$peri.engraft.other = 
  factor(patterns$peri.engraft.other,labels = c("No","Yes"))
form = as.formula(paste0("peri.engraft.other~",right_form))
hsct = tableby(form,data = patterns,cat.test = "fe")
summary(hsct,labelTranslations = trend_vars,pfootnote = T)
```
