---
title: "HCL Overnight Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(tools)
library(arsenal)
library(cgmanalysis)
library(car)
library(skimr)
library(Hmisc)
library(knitr)
library(lme4)
library(nlme)
library(MASS)
library(tidyverse)
library(broom)
library(broom.mixed)
library(emmeans)
knitr::opts_chunk$set(echo = TRUE)
home_dir = ifelse(.Platform$OS.type != "unix",
                  "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/Nocturnal Alarms/",
                  "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/Nocturnal Alarms/")
knitr::opts_knit$set(root.dir = home_dir)
```

```{r echo=FALSE,include=FALSE}
# Import Cari's data
source("~/GitHub/BDC-Code/Erin Cobry/Nocturnal Alarms/AdvancedClosedLoopCl_R_2020-01-05_1101.r")
# Demographics
demographics <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(child_ya))
# remove missing
no_data <- c("27","30","38","41")
demographics <- demographics[-c(which(demographics$record_id %in% no_data)),]
# Manually corrected a few dates in CSV file.
# Read in (manually converted to long format from Erin's Excel spreadsheet)
overnight_alarms <- read.csv("./Data_Cleaned/nocturnal_alarms.csv",
                             na.strings = c("","No Data"))
overnight_alarms = overnight_alarms[overnight_alarms$num_nights > 0,]
# Match IDs and filter
dates = read.csv("./Data_Cleaned/subject_dates.csv")
overnight_alarms$id = dates$id[match(gsub(" ","",overnight_alarms$id),tolower(gsub(" ","",dates$name)))]
overnight_alarms <- overnight_alarms %>% filter(!(id %in% no_data))
# Format date columns
overnight_alarms[,c("start_date","end_date")] <- 
  lapply(overnight_alarms[,c("start_date","end_date")], 
         function(x) {lubridate::mdy(as.character(x))})
# Numeric timepoints in months
overnight_alarms$numeric_time <- 
  car::recode(overnight_alarms$timepoint,
              "'Baseline' = 0;'Post' = 0.5;'T1' = 3;'T2' = 6;'T3' = 9;'T4' = 12")
overnight_alarms$numeric_time <- 
  as.numeric(as.character(overnight_alarms$numeric_time))
# Alarms per day
alarm_vars = c("num_alarms","threshold_alarms","maintenance_alarms","hcl_alarms",
               "pump_alarms","other_alarms")
overnight_alarms = overnight_alarms %>%
  mutate(across(all_of(alarm_vars),.fns = ~./num_nights,.names = "{col}_per_night"))
 # Get Erin's participants 
demographics <- demographics[which(demographics$record_id %in% overnight_alarms$id),]
# Combine CGM and pump hx levels
levels(demographics$demographics_cgmhx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
levels(demographics$demographics_pumphx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
```

# Descriptive Characteristics

```{r echo=FALSE,results='asis'}
# Labels
labels <- list(demographics_t1d_duration = "Diabetes Dur. (years)", 
               demographics_ethnicity.factor = "Ethnicity",
               demographics_sex.factor = "Sex",
               demographics_cgmhx.factor = "CGM History", 
               demographics_pumphx.factor = "Pump History")
# Table 1
t1 <- tableby(~ hba1c + kwt(demographics_age,"median","q1q3","range") + 
                kwt(demographics_t1d_duration,"median","q1q3") + 
                demographics_sex.factor + demographics_ethnicity.factor +
                demographics_cgmhx.factor + demographics_pumphx.factor,
              data = demographics)
# Print
summary(t1,labelTranslations = labels)
```

```{r echo=FALSE,include=FALSE}
# Read in CGM data
cgm <- read.csv("./Data_Cleaned/cgm_variables.csv",stringsAsFactors = F)
# ID and timepoint
cgm$id <- as.numeric(gsub("[[:alpha:]]","",sapply(strsplit(cgm$subject_id,"_"),`[[`,1)))
cgm$timepoint <- sapply(strsplit(cgm$subject_id,"_"),`[[`,2)
cgm$timepoint[cgm$timepoint == "week 2"] = "Post"
# Select columns
cgm <- cgm %>% select(id,timepoint,nighttime_avg_sens_glucose,
                      percent_time_70_180_night,percent_time_under_70_night)
# Merge
overnight_alarms <- left_join(overnight_alarms,cgm,by = c("id","timepoint"))
# Read in corrected AM data
am_correct <- read.csv("./Data_Cleaned/REDCap AM Data.csv")
am_correct$timepoint[am_correct$timepoint == "week 2"] = "Post"
# Add to overnight alarms
overnight_alarms <- left_join(overnight_alarms,am_correct,by = c("id", "timepoint"))
# Order timepoints
overnight_alarms$timepoint <- factor(overnight_alarms$timepoint,
                                     levels = c("Baseline","Post","T1","T2","T3","T4" ))
levels(overnight_alarms$timepoint) <- c("Baseline","Week 2",
                                        "T1","T2","T3","T4" )
overnight_alarms = overnight_alarms %>% arrange(id,timepoint)
# Remove discontinuers
discont = overnight_alarms %>% group_by(id) %>% 
  summarise(discont = any(sensor_wear <= 10,na.rm = T)) %>%
  filter(discont == T) %>% .$id
```

Discontinuers were defined as any participant who had $\leq$ 10% sensor wear at any timepoint. These participants were excluded from all models below. 

# Edits

Greg’s comments:
-	Could we look at adjusting for time with sensor wear (since we already did time in auto mode)? 

-	Greg also mentioned doing a comparison of the alarms at 3, 6, 9, and 12 months compared to 2 weeks. Would you mind looking at this as well? 

Laurel’s comments:
-	She brings up an interesting point about excluding the discontinuers. This may also help address Greg’s comments about adjusting for the sensor use. They previously defined discontinuation (of the HCL system) as those with <10% use. I think we could use the same to look at sensor use (<10% sensor use). So I wonder if we could look at each time point again but this time exclude anyone that had <10% sensor use. 

-	And then look at those who met criteria for discontinuation to see how many alerts they had at the 2 weeks mark vs those who were not discontinuers. 


# Alarm Frequency

## Continuers

```{r echo=FALSE}
alarm_vars_per_night = paste0(alarm_vars,"_per_night")
alarm_table <- overnight_alarms %>% group_by(timepoint) %>%
  filter(!id %in% discont) %>%
  summarise(across(all_of(c(alarm_vars_per_night,"sensor_wear","time_am")),
                   .fns = ~round(median(.,na.rm = T),3)))
colnames(alarm_table)[2:ncol(alarm_table)] = 
  paste0("Median ",toTitleCase(gsub("_"," ",colnames(alarm_table)[2:ncol(alarm_table)])))
kable(alarm_table)
```

## Discontinuers

```{r echo=FALSE}
alarm_table <- overnight_alarms %>% group_by(timepoint) %>%
  filter(id %in% discont) %>%
  summarise(across(all_of(c(alarm_vars_per_night,"sensor_wear","time_am")),
                   .fns = ~round(median(.,na.rm = T),3)))
colnames(alarm_table)[2:ncol(alarm_table)] = 
  paste0("Median ",toTitleCase(gsub("_"," ",colnames(alarm_table)[2:ncol(alarm_table)])))
kable(alarm_table)
```

# Longitudinal Analysis

```{r echo=FALSE}
overnight_alarms = overnight_alarms %>% filter(!id %in% discont)
```

Total alarms per night were analyzed using a linear mixed model with random intercept for participant and continuous AR1 correlation structure. All other outcomes were analyzed using a generalized Poisson mixed model with random intercept for subject and continuous AR1 correlation structure.

## Alarms per Night By Timepoint

### Model Results - Unadjusted

#### Comparison to Baseline

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- lme(num_alarms_per_night ~ timepoint,random = ~1|id,
                   data = overnight_alarms,correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- sub("timepoint","",results$term)
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a statistically significant effect of timepoint on the rate of alarms per night (overall p = `r round(an_res["timepoint",3],3)`).

#### Mean at Each Timepoint

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Alarms per Night Over Time

```{r echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = numeric_time, y = num_alarms_per_night,group = id)) + 
  geom_line(alpha=0.1) + 
  geom_line(data=means,aes(x=numeric_time,y=as.numeric(emmean),group=1),
            size = 1,inherit.aes = F) + 
  theme_bw() +xlab("Timepoint") + ylab("Alarms per Night") + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_x_continuous(breaks=c(0,0.5,3,6,9,12),
                   labels=c("Baseline","Week 2","3 Month","6 Month","9 Month","12 Month"))
alarms_over_time
```

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 = lme(num_alarms_per_night ~ timepoint + time_am,random = ~1|id,na.action = na.omit,
                   data = overnight_alarms,correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was no effect of timepoint on the rate of alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was a statistically significant effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["time_am",3],3)`).

#### LS Mean at Each Timepoint

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Sensor Wear

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 = lme(num_alarms_per_night ~ timepoint + sensor_wear,random = ~1|id,
                  na.action = na.omit,data = overnight_alarms,
                  correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Sensor Wear" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of alarms per night after adjusting for sensor wear (overall p = `r round(an_res["timepoint",3],3)`). There was not a statistically significant effect of sensor wear on the rate of alarms per night (overall p = `r round(an_res["sensor_wear",3],3)`).

#### LS Mean at Each Timepoint

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Number of Observations at Each Timepoint

```{r echo=FALSE,message=FALSE}
alarm_obs <- overnight_alarms %>% group_by(timepoint) %>% count(!is.na(num_alarms_per_night))
kable(alarm_obs[,c("timepoint","n")])
```

## Threshold Alarms per Night By Timepoint

### Plot

```{r echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = threshold_alarms_per_night,group = id)) + 
  geom_line(aes(color=factor(id))) + theme_bw() + 
  xlab("Timepoint") + ylab("Threshold Alarms per Night") + 
  theme(legend.position = "none")
alarms_over_time
```

### Model Results - Unadjusted

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(threshold_alarms ~ timepoint + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T,exponentiate = T)
results$term <- sub("timepoint","",results$term)
results[,c("estimate","conf.low","conf.high")] = 
  lapply(results[,c("estimate","conf.low","conf.high")],exp)
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a statistically significant effect of timepoint on the rate of threshold alarms per night (overall p = `r round(an_res["timepoint",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(threshold_alarms ~ timepoint + time_am + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = 
  lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of threshold alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was not a  statistically significant effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["time_am",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Sensor Wear

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(threshold_alarms ~ timepoint + sensor_wear + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Sensor Wear" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = 
  lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was not a significant effect of timepoint on the rate of threshold alarms per night after adjusting for sensor wear (overall p = `r round(an_res["timepoint",3],3)`). There was also not a statistically significant effect of sensor wear on the rate of alarms per night (overall p = `r round(an_res["sensor_wear",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Number of Observations at Each Timepoint

```{r echo=FALSE,message=FALSE}
alarm_obs <- overnight_alarms %>% group_by(timepoint) %>% count(!is.na(threshold_alarms))
kable(alarm_obs[,c("timepoint","n")])
```

## Maintenance Alarms per Night By Timepoint

### Plot

```{r echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = maintenance_alarms_per_night,group = id)) + 
  geom_line(aes(color=factor(id))) + theme_bw() + 
  xlab("Timepoint") + ylab("Maintenance Alarms per Night") + 
  theme(legend.position = "none")
alarms_over_time
```

### Model Results - Unadjusted

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(maintenance_alarms ~ timepoint + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- sub("timepoint","",results$term)
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was an effect of timepoint on the rate of maintenance alarms per night (overall p = `r round(an_res["timepoint",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(maintenance_alarms ~ timepoint + time_am + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was an effect of timepoint on the rate of maintenance alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was also an effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["time_am",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Sensor Wear

Model would not converge.

```{r echo=FALSE,message=FALSE,eval=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(maintenance_alarms ~ timepoint + sensor_wear + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```
### Number of Observations at Each Timepoint

```{r echo=FALSE,message=FALSE}
alarm_obs <- overnight_alarms %>% group_by(timepoint) %>% count(!is.na(maintenance_alarms))
kable(alarm_obs[,c("timepoint","n")])
```

## HCL Alarms per Night By Timepoint

### Plot

```{r echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = hcl_alarms_per_night,group = id)) + 
  geom_line(aes(color=factor(id))) + theme_bw() + 
  xlab("Timepoint") + ylab("HCL Alarms per Night") + 
  theme(legend.position = "none")
alarms_over_time
```

### Model Results - Unadjusted

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(hcl_alarms ~ timepoint + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- sub("timepoint","",results$term)
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of HCL alarms per night (overall p < 0.001). The estimates presented above are interpreted as incidence rate ratios.

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(maintenance_alarms ~ timepoint + time_am + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was an effect of timepoint on the rate of maintenance alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was also an effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["time_am",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(hcl_alarms ~ timepoint + time_am + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of HCL alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was also a significant effect of time in AM on the rate of alarms per night (overall p < 0.001). The estimates presented above are interpreted as incidence rate ratios.

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(maintenance_alarms ~ timepoint + time_am + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was an effect of timepoint on the rate of maintenance alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was also an effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["time_am",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Sensor Wear

Model did not converge

```{r echo=FALSE,message=FALSE,eval=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(maintenance_alarms ~ timepoint + sensor_wear + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1,type = 3))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

### Number of Observations at Each Timepoint

```{r echo=FALSE,message=FALSE}
alarm_obs <- overnight_alarms %>% group_by(timepoint) %>% count(!is.na(hcl_alarms))
kable(alarm_obs[,c("timepoint","n")])
```

## Pump Alarms per Night By Timepoint

### Plot

```{r echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = pump_alarms_per_night,group = id)) + 
  geom_line(aes(color=factor(id))) + theme_bw() + 
  xlab("Timepoint") + ylab("Pump Alarms per Night") + 
  theme(legend.position = "none")
alarms_over_time
```

### Model Results - Unadjusted

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(pump_alarms ~ timepoint + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- sub("timepoint","",results$term)
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of pump alarms per night (overall p = `r round(an_res["timepoint",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(pump_alarms ~ timepoint + time_am + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of pump alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was also a significant effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["time_am",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Sensor Wear

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(pump_alarms ~ timepoint + sensor_wear + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Sensor Wear")
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was not a significant effect of timepoint on the rate of pump alarms per night after adjusting for sensor wear (overall p = `r round(an_res["timepoint",3],3)`). There was also not a significant effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["sensor_wear",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Number of Observations at Each Timepoint

```{r echo=FALSE,message=FALSE}
alarm_obs <- overnight_alarms %>% group_by(timepoint) %>% count(!is.na(pump_alarms))
kable(alarm_obs[,c("timepoint","n")])
```

## Other Alarms per Night By Timepoint

### Plot

```{r echo=FALSE,warning=FALSE,dpi=600}
alarms_over_time <- 
  ggplot(overnight_alarms,aes(x = timepoint, y = other_alarms_per_night,group = id)) + 
  geom_line(aes(color=factor(id))) + theme_bw() + 
  xlab("Timepoint") + ylab("Other Alarms per Night") + 
  theme(legend.position = "none")
alarms_over_time
```

### Model Results - Unadjusted

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(other_alarms ~ timepoint + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- sub("timepoint","",results$term)
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of other alarms per night (overall p = `r round(an_res["timepoint",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Time in AM

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(other_alarms ~ timepoint + time_am + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of other alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was not significant effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["time_am",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Model Results - Adjusted for Sensor Wear

```{r echo=FALSE,message=FALSE}
# Random intercept (not enough data for a random slope)
mod_ri_car1 <- glmmPQL(other_alarms ~ timepoint + sensor_wear + offset(log(num_nights)),
                     random = ~1|id,family = "poisson",
                     data = overnight_alarms,
                     correlation = corCAR1(form = ~numeric_time))
results <- tidy(mod_ri_car1,effects = "fixed",conf.int = T)
results$term <- c("(Intercept)","Week 2","T1","T2","T3","T4","% Time in AM" )
results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
results[,c("estimate","conf.low","conf.high")] = lapply(results[,c("estimate","conf.low","conf.high")],exp)
# Anova
an_res <- as.data.frame(Anova(mod_ri_car1))
# Print
kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3)
```

There was a significant effect of timepoint on the rate of other alarms per night after adjusting for time in AM (overall p = `r round(an_res["timepoint",3],3)`). There was not significant effect of time in AM on the rate of alarms per night (overall p = `r round(an_res["sensor_wear",3],3)`). The estimates presented above are interpreted as incidence rate ratios.

#### LS Mean at Each Timepoint

Results below are presented on the log scale.

```{r echo=FALSE}
emm1 = emmeans(mod_ri_car1,~timepoint)
means = data.frame(emm1)
kable(means)
means$numeric_time = c(0,0.5,3,6,9,12)
```

#### Comparison to week 2

```{r echo=FALSE}
base = c(1,0,0,0,0,0)
w2 = c(0,1,0,0,0,0)
t1 = c(0,0,1,0,0,0)
t2 = c(0,0,0,1,0,0)
t3 = c(0,0,0,0,1,0)
t4 = c(0,0,0,0,0,1)
con = contrast(emm1,method = list("Baseline to Week 2"= w2 - base,
                            "Week 2 to T1"=t1 - w2,
                            "Week 2 to T2"=t2 - w2,
                            "Week 2 to T3"=t3 - w2,
                            "Week 2 to T3"=t4 - w2))
kable(con,digits = 3)
```

### Number of Observations at Each Timepoint

```{r echo=FALSE,message=FALSE}
alarm_obs <- overnight_alarms %>% group_by(timepoint) %>% count(!is.na(other_alarms))
kable(alarm_obs[,c("timepoint","n")])
```