---
title: "HCL Overnight Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
library(tools)
library(arsenal)
library(cgmanalysis)
library(car)
library(skimr)
library(Hmisc)
library(knitr)
library(lmerTest)
library(nlme)
library(performance)
library(MASS)
library(tidyverse)
library(broom)
library(broom.mixed)
library(lubridate)
library(emmeans)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r echo=FALSE,include=FALSE}
# Import Cari's data
source("~/GitHub/BDC-Code/Erin Cobry/Nocturnal Alarms/AdvancedClosedLoopCl_R_2020-01-05_1101.r")
# Demographics
demographics <- data %>% 
  filter(redcap_event_name %in% c("baseline_arm_1","baseline_arm_2"),
         !is.na(child_ya))
# remove missing
no_data <- c("27","30","38","41")
demographics <- demographics[-c(which(demographics$record_id %in% no_data)),]
# Manually corrected a few dates in CSV file.
# Read in (manually converted to long format from Erin's Excel spreadsheet)
alarms <- read.csv("./Erin Cobry/Nocturnal Alarms/Data_Cleaned/alarms_per_hour.csv",na.strings = "")
alarms$timepoint = as.factor(alarms$timepoint)
levels(alarms$timepoint) = c("Baseline","T1","Post","Post","Post",
                                       "Post","Baseline","Baseline","T1","T2","T3","T4")
# Match IDs and filter
dates = read.csv("./Erin Cobry/Nocturnal Alarms/Data_Cleaned/subject_dates.csv")
alarms$id = dates$id[match(gsub(" ","",alarms$id),tolower(gsub(" ","",dates$name)))]
alarms <- alarms %>% filter(!(id %in% no_data))
# Format date columns
alarms[,c("start_date","end_date")] <- 
  lapply(alarms[,c("start_date","end_date")], 
         function(x) {lubridate::mdy(as.character(x))})
# Numeric timepoints in months
alarms$numeric_time <- 
  car::recode(alarms$timepoint,
              "'Baseline' = 0;'Post' = 0.5;'T1' = 3;'T2' = 6;'T3' = 9;'T4' = 12")
alarms$numeric_time <- 
  as.numeric(as.character(alarms$numeric_time))
# Calculate alarms per hour
alarms = alarms %>% 
  select(id:end_date,numeric_time,everything())
# Get Erin's participants 
demographics <- demographics[which(demographics$record_id %in% alarms$id),]
# Combine CGM and pump hx levels
levels(demographics$demographics_cgmhx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
levels(demographics$demographics_pumphx.factor) <-
  c("< 1 month","> 1 month","> 1 month","> 1 month","> 1 month","> 1 month")
# Time from AM start and early vs. late adopters
demographics$early_late = cut(ymd(demographics$automode_start),breaks = 2,
                              labels = c("Early Adopter","Late Adopter"))
demographics$id = as.numeric(demographics$record_id)
alarms = left_join(alarms,demographics[,c("id","automode_start","early_late")],by = "id")
alarms$automode_start = ymd(alarms$automode_start)
alarms$days_from_am_start = 
  as.numeric(difftime(alarms$start_date,
                      alarms$automode_start,units = "days"))
```

# Descriptive Characteristics

Participants were split into two equal groups based on AM start date. Due to several duplicated AM start dates, the groups were not exactly even but are reasonably close. 

```{r echo=FALSE,results='asis'}
# Labels
labels <- list(demographics_t1d_duration = "Diabetes Dur. (years)", 
               demographics_ethnicity.factor = "Ethnicity",
               demographics_sex.factor = "Sex",
               demographics_cgmhx.factor = "CGM History", 
               demographics_pumphx.factor = "Pump History")
# Table 1
t1 <- tableby(early_late ~ hba1c + kwt(demographics_age,"median","q1q3","range") + 
                kwt(demographics_t1d_duration,"median","q1q3") + 
                demographics_sex.factor + demographics_ethnicity.factor +
                demographics_cgmhx.factor + demographics_pumphx.factor,
              data = demographics)
# Print
summary(t1,labelTranslations = labels)
```

```{r echo=FALSE,include=FALSE}
# Read in CGM data
cgm <- read.csv("./Erin Cobry/Nocturnal Alarms/Data_Cleaned/cgm_variables.csv",stringsAsFactors = F)
# ID and timepoint
cgm$id <- as.numeric(gsub("[[:alpha:]]","",sapply(strsplit(cgm$subject_id,"_"),`[[`,1)))
cgm$timepoint <- sapply(strsplit(cgm$subject_id,"_"),`[[`,2)
cgm$timepoint[cgm$timepoint == "week 2"] = "Post"
# Select columns
cgm <- cgm %>% select(id,timepoint,nighttime_avg_sens_glucose,
                      percent_time_70_180_night,percent_time_under_70_night)
# Merge
alarms <- left_join(alarms,cgm,by = c("id", "timepoint"))
# Read in corrected AM data
am_correct <- read.csv("./Erin Cobry/Nocturnal Alarms/Data_Cleaned/REDCap AM Data.csv")
am_correct$timepoint[am_correct$timepoint == "week 2"] = "Post"
# Add to overnight alarms
alarms <- left_join(alarms,am_correct,by = c("id", "timepoint"))
# Order timepoints
alarms$timepoint <- factor(alarms$timepoint,levels = c("Baseline","Post","T1","T2","T3","T4" ))
levels(alarms$timepoint) <- c("Baseline","Week 2","T1","T2","T3","T4" )
alarms = alarms %>% arrange(id,timepoint)
# Calc
# Remove discontinuers
discont = alarms %>% group_by(id) %>% 
  summarise(discont = any(sensor_wear <= 10,na.rm = T)) %>%
  filter(discont == T) %>% .$id
```

Discontinuers were defined as any participant who had $\leq$ 10% sensor wear at any timepoint. These participants (a total of `r length(discont)`) were excluded from all models below. 

# Longitudinal Analysis

```{r echo=FALSE}
alarms = alarms %>% filter(!id %in% discont)
```

All alarms were analyzed using a linear mixed model with random intercept for participant.

```{r}
# Model function
fit_mod = function(outcome,covariates = NULL,df = alarms){
  # Random intercept (not enough data for a random slope). CAR1 structure didn't
  # seem to make a difference
  if(!is.null(covariates)){
    f = as.formula(paste0(outcome,"~ timepoint*early_late+",
                          paste0(covariates,collapse = "+"),"+(1|id)"))
  } else {
    f = as.formula(paste0(outcome,"~ timepoint*early_late+(1|id)"))
  }
  list2env(list(f=f,df=df),.GlobalEnv)
  mod_ri <- lmer(f,data = df)
  results <- tidy(mod_ri,effects = "fixed",conf.int = T)
  results$term <- sub("timepoint","",results$term)
  results$term <- sub("early_late","",results$term)
  results$term <- sub("sensor_wear","Sensor Wear",results$term)
  results$term <- sub("time_am","Time in AM",results$term)
  results$p.value <- format.pval(results$p.value,eps = 0.001,digits = 3)
  # Anova
  an_res <- as.data.frame(Anova(mod_ri,type = 3))
  rownames(an_res) <- sub("timepoint","Timepoint",rownames(an_res))
  rownames(an_res) <- sub("early_late","Early vs. Late Adopter",rownames(an_res))
  rownames(an_res) <- sub("sensor_wear","Sensor Wear",rownames(an_res))
  rownames(an_res) <- sub("time_am","Time in AM",rownames(an_res))
  an_res$`Pr(>Chisq)` <- format.pval(an_res$`Pr(>Chisq)`,eps = 0.001,digits = 3)
  list2env(list(an_res=an_res),.GlobalEnv)
  # Print
  cat("\n")
  cat("#### Model Results")
  cat("\n")
  print(kable(results[,c("term","estimate","conf.low","conf.high","p.value")],digits = 3))
  cat("\n")
  cat("#### Type III Tests of Fixed Effects")
  cat("\n")
  print(kable(an_res,digits = 3))
  cat("\n")
}
```

## Total alarms per hour

### Unadjusted

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "num_alarms_night")
```

### Adjusted for time in AM

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "num_alarms_night",covariates = "time_am")
```

### Adjusted for sensor wear

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "num_alarms_night",covariates = "sensor_wear")
```

## Threshold alarms per hour

### Unadjusted

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "threshold_alarms_night")
```

### Adjusted for time in AM

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "threshold_alarms_night",covariates = "time_am")
```

### Adjusted for sensor wear

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "threshold_alarms_night",covariates = "sensor_wear")
```

## Maintenance alarms per hour

### Unadjusted

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "maintenance_alarms_night")
```

### Adjusted for time in AM

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "maintenance_alarms_night",covariates = "time_am")
```

### Adjusted for sensor wear

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "maintenance_alarms_night",covariates = "sensor_wear")
```

## HCL alarms per hour

### Unadjusted

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "hcl_alarms_night")
```

### Adjusted for time in AM

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "hcl_alarms_night",covariates = "time_am")
```

### Adjusted for sensor wear

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "hcl_alarms_night",covariates = "sensor_wear")
```

## Pump alarms per hour

### Unadjusted

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "pump_alarms_night")
```

### Adjusted for time in AM

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "pump_alarms_night",covariates = "time_am")
```

### Adjusted for sensor wear

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "pump_alarms_night",covariates = "sensor_wear")
```

## Other alarms per hour

### Unadjusted

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "other_alarms_night")
```

### Adjusted for time in AM

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "other_alarms_night",covariates = "time_am")
```

### Adjusted for sensor wear

```{r echo=FALSE,results='asis'}
fit_mod(outcome = "other_alarms_night",covariates = "sensor_wear")
```

# Figure

```{r fig.width=10,fig.height=8}
# Just relevant variables, melt to longer
plot_df = alarms %>% select(id,timepoint,ends_with("am"),-time_am) %>%
  pivot_longer(num_alarms_10pm_12am:other_alarms_4am_6am,
               names_to = c("type", "time"),names_pattern = "(.*_.*)_(.*_.*)") %>%
  filter(type != "num_alarms")
plot_df$time = factor(plot_df$time,levels = c("10pm_12am","12am_2am","2am_4am","4am_6am"),
                      labels = c("10pm-12am","12am-2am","2am-4am","4am-6am"))
plot_df$type = factor(plot_df$type,levels = c("hcl_alarms","maintenance_alarms","other_alarms",
                                              "pump_alarms","threshold_alarms"),
                      labels = c("HCL Alarms","Maintenance Alarms",
                                 "Other Alarms","Pump Alarms","Threshold Alarms"))
# Bar chart
ggplot(plot_df,aes(x=time,y=value,fill=type)) + 
  geom_boxplot() +
  scale_fill_brewer(palette = "Set2",name = "") +
  theme_bw() + xlab("Time") + ylab("Alarms Per Hour") +
  facet_wrap(~ timepoint)
```