---
title: ''
output:
  pdf_document: default
header-includes:
  - \usepackage{caption}
  - \pagenumbering{gobble}
---
\captionsetup[table]{labelformat=empty}

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
#read
A1Cs <- readxl::read_excel('S:/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Raw/Venus and POC A1Cs_corrections Updated 5-10-23.xlsx')
dates <- readr::read_csv('S:/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/PICLS/Data_Raw/picls_date_check.csv')
```

```{r clean data, include=FALSE}
#drop comment columns, clean a few names, format date
A1Cs=A1Cs[,-c(6,7)] %>%
rename('A1C'='A1C Venousalue') %>% 
  mutate(Source=recode(Source, Venous = 'Venous', Venousenous = 'Venous', POC =  'POC' ),
         `Date Taken`=as.Date(`Date Taken`, format='%Y-%m-%d'))

dates=dates %>% 
  mutate(edd=as.Date(edd, format='%m/%d/%Y'),
         edd_derived_date=as.Date(edd_derived_date,format='%m/%d/%Y')) %>%
  mutate(edd_derived_date_D6=edd_derived_date+6) %>% 
  rename('edd_derived_date_D0'='edd_derived_date')

#new df of rows where A1C date and gestational date line up
#join by ID and whether A1C date taken falls into D0-D6 range
A1Cs_gestdates=A1Cs %>% 
  left_join(dates, by=c('Study ID'='ID', NULL)) %>% 
  filter(`Date Taken`>=edd_derived_date_D0 & `Date Taken`<=edd_derived_date_D6)
```

```{r tables, echo=FALSE}
##count number of subjects with data at visits of interest

#venous only
venous_A1Cs_gestdates=A1Cs_gestdates %>%
  filter(Source == 'Venous')

venous_counts=venous_A1Cs_gestdates %>% 
  group_by(`Study ID`) %>% 
  summarize(Visit_4=sum(week>=6 & week<=7),
            Visit_5=sum(week>=8 & week<=11),
            Visit_6=sum(week>=14 & week<=17),
            Visit_7=sum(week>=18 & week<=19),
            Visit_8=sum(week>=20 & week<=23),
            Visit_9=sum(week>=24 & week<=27),
            Visit_10=sum(week>=30 & week<=31),
            Visits_11thru12=sum(week>=32 & week<=35),
            Visits_13thru14=sum(week>=36 & week<=39))

venous_counts %>% 
  summarize('Visit 4 (6w0d-7w6d)'=sum(Visit_4 != 0),
            'Visit 5 (8w0d-11w6d)'=sum(Visit_5 != 0),
            'Visit 6 (14w0d-17w6d)'=sum(Visit_6 != 0),
            'Visit 7 (18w0d-19w6d)'=sum(Visit_7 != 0),
            'Visit 8 (20w0d-23w6d)'=sum(Visit_8 != 0),
            'Visit 9 (24w0d-27w6d)'=sum(Visit_9 != 0),
            'Visit 10 (30w0d-31w6d)'=sum(Visit_10 != 0),
            'Visits 11-12 (32w0d-35w6d)'=sum(Visits_11thru12 != 0),
            'Visits 13-14 (36w0d-39w6d)'=sum(Visits_13thru14 != 0)) %>% t() %>% 
  kable(caption='Number of Subjects with A1C Values within Study Visit Window (Venous Only):', booktabs=TRUE, linesep='',
        col.names='n') %>% 
  kable_styling(latex_options='HOLD_position')

#venous + POC
POCvenous_counts=A1Cs_gestdates %>% 
  group_by(`Study ID`) %>% 
  summarize(Visit_4=sum(week>=6 & week<=7),
            Visit_5=sum(week>=8 & week<=11),
            Visit_6=sum(week>=14 & week<=17),
            Visit_7=sum(week>=18 & week<=19),
            Visit_8=sum(week>=20 & week<=23),
            Visit_9=sum(week>=24 & week<=27),
            Visit_10=sum(week>=30 & week<=31),
            Visits_11thru12=sum(week>=32 & week<=35),
            Visits_13thru14=sum(week>=36 & week<=39))

POCvenous_counts %>% 
  summarize('Visit 4 (6w0d-7w6d)'=sum(Visit_4 != 0),
            'Visit 5 (8w0d-11w6d)'=sum(Visit_5 != 0),
            'Visit 6 (14w0d-17w6d)'=sum(Visit_6 != 0),
            'Visit 7 (18w0d-19w6d)'=sum(Visit_7 != 0),
            'Visit 8 (20w0d-23w6d)'=sum(Visit_8 != 0),
            'Visit 9 (24w0d-27w6d)'=sum(Visit_9 != 0),
            'Visit 10 (30w0d-31w6d)'=sum(Visit_10 != 0),
            'Visits 11-12 (32w0d-35w6d)'=sum(Visits_11thru12 != 0),
            'Visits 13-14 (36w0d-39w6d)'=sum(Visits_13thru14 != 0)) %>% t() %>% 
  kable(caption='Number of Subjects with A1C Values within Study Visit Window (Venous + POC):', booktabs=TRUE, linesep='',
        col.names='n') %>% 
  kable_styling(latex_options='HOLD_position')
```
*Notes:*

* Gestational weeks corresponding to each subject's study dates were calculated based on the EDD date. Subjects who did not have an EDD date are not included in the above summaries; the summaries are only based on the 23 subjects with EDD dates.

* Example: Referencing the Venous only table, there are 4 out of 23 subjects who have at least one venous A1C value that falls within the Visit 4 window.