---
title: "Tech Satisfaction after HCL"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)


# 3 different data sets
HCL_DATA_2024 <- read_csv("S:/Laura/BDC/Projects/Cari Berget/Berget Trojanowski distress/HCLClinicalObservati_DATA_2024.csv")
ClinicalObservationC_DATA_2024 <- read_csv("S:/Laura/BDC/Projects/Cari Berget/Berget Trojanowski distress/ClinicalObservationC_DATA_2024-03-05_0920.csv")
ClinicalObservationP_DATA_2024 <- read_csv("S:/Laura/BDC/Projects/Cari Berget/Berget Trojanowski distress/ClinicalObservationP_DATA_2024-03-05_0931.csv")
```

```{r data clean, include = FALSE}
################## Clinical Obs - Child/ Adult ##########################

# surveys - paid c need to be coded 0-4, subtract by 1 before scoring, paid y ready
# baseline
clinicalc_bl = ClinicalObservationC_DATA_2024 %>% filter(timepoint == 0) %>% select(record_id, timepoint,redcap_event_name,demographics_hcl, demographics_age, demographics_ethnicity,
                                                                                    demographics_race, demographics_sex, demographics_insurance, demographics_t1d_duration, hba1c,
                                                                                    ya_paid1:ya_paid20, c_paid1:c_paid20, inspire_b1:inspire_b17)

# sub 1 from cpaid then reverse score, remove 5 from inspires
clinicalc_bl = clinicalc_bl %>% mutate(across(c_paid1:c_paid20, ~.-1))
clinicalc_bl = clinicalc_bl %>% mutate(across(c_paid1:c_paid20, ~4-.))

remove5 = function(x){ifelse(x==5, NA_real_, x)}
clinicalc_bl = clinicalc_bl %>% mutate(across(inspire_b1:inspire_b17, remove5))

# score paid and inspire bl  
clinicalc_bl = clinicalc_bl %>% mutate(c_paid =  select(., c_paid1:c_paid20) %>% rowMeans(na.rm=T),
          c_paid_score =  c_paid*25,
          ya_paid =  select(., ya_paid1:ya_paid20) %>% rowSums(),
          ya_paid_score =  ya_paid*1.25,
          paid_score = ifelse(is.na(ya_paid_score), c_paid_score, ya_paid_score),
          inspire_score = select(., inspire_b1:inspire_b17) %>% rowMeans(na.rm=T)*25)
clinicalc_bl = clinicalc_bl %>% select(record_id, timepoint,redcap_event_name,demographics_hcl, demographics_age, demographics_ethnicity,                                                                                    demographics_race, demographics_sex, demographics_insurance, demographics_t1d_duration, hba1c,paid_score, inspire_score) 
clinicalc_bl = clinicalc_bl %>% mutate(baseline_paid = paid_score)

# followup
clinicalc_f = ClinicalObservationC_DATA_2024 %>% filter(timepoint != 0) %>% select(record_id, timepoint,redcap_event_name, hba1c,
                                                                                    ya_paid1:ya_paid20, c_paid1:c_paid20, inspire_f1:inspire_f17)

# sub 1 from cpaid then reverse score, remove 5 from inspires
clinicalc_f = clinicalc_f %>% mutate(across(c_paid1:c_paid20, ~.-1))
clinicalc_f = clinicalc_f %>% mutate(across(c_paid1:c_paid20, ~4-.))

clinicalc_f = clinicalc_f %>% mutate(across(inspire_f1:inspire_f17, remove5))

# score paid and inspire bl  
clinicalc_f = clinicalc_f %>% mutate(c_paid =  select(., c_paid1:c_paid20) %>% rowMeans(na.rm=T),
          c_paid_score =  c_paid*25,
          ya_paid =  select(., ya_paid1:ya_paid20) %>% rowSums(),
          ya_paid_score =  ya_paid*1.25,
          paid_score = ifelse(is.na(ya_paid_score), c_paid_score, ya_paid_score),
          inspire_score = select(., inspire_f1:inspire_f17) %>% rowMeans(na.rm=T)*25)
clinicalc_f = clinicalc_f %>% select(record_id, timepoint,redcap_event_name, hba1c,paid_score, inspire_score)

# merge base and FU
clinical_c = full_join(clinicalc_bl, clinicalc_f)
clinical_c = clinical_c %>% group_by(record_id) %>% fill(demographics_hcl:demographics_t1d_duration, .direction = "down")
# ignore adult arm record id 181-184
clinical_c = clinical_c %>% filter(!(record_id %in% c(181,182,183,184)))
# only include controliq=1, exclude medtronic (0)
clinical_c = clinical_c %>% filter(demographics_hcl == 1)
rm(clinicalc_bl, clinicalc_f, ClinicalObservationC_DATA_2024)
#########################################################################
################## Clinical Obs - Parent       ##########################
# surveys - paid c need to be coded 0-4, subtract by 1 before scoring, paid y ready
# baseline
clinicalp_bl = ClinicalObservationP_DATA_2024 %>% filter(p_timepoint == 0) %>% select(record_id, p_timepoint,redcap_event_name,
                                                                                    p_paid1:p_paid18, p_b_inspire1:p_b_inspire21)

# sub 1 from ppaid then reverse score, remove 5 from inspires
clinicalp_bl = clinicalp_bl %>% mutate(across(p_paid1:p_paid18, ~.-1))
clinicalp_bl = clinicalp_bl %>% mutate(across(p_paid1:p_paid18, ~4-.))

clinicalp_bl = clinicalp_bl %>% mutate(across(p_b_inspire1:p_b_inspire21, remove5))

# score paid and inspire bl  
clinicalp_bl = clinicalp_bl %>% mutate(p_paid =  select(., p_paid1:p_paid18) %>% rowMeans(na.rm=T),
          p_paid_score =  p_paid*25,
          p_inspire_score = select(., p_b_inspire1:p_b_inspire21) %>% rowMeans(na.rm=T)*25)
clinicalp_bl = clinicalp_bl %>% select(record_id, p_timepoint,redcap_event_name,p_paid_score, p_inspire_score) 
clinicalp_bl = clinicalp_bl %>% mutate(baseline_p_paid = p_paid_score)

# followup
clinicalp_f = ClinicalObservationP_DATA_2024 %>% select(record_id,redcap_event_name, p_inspire1_t1:p_inspire21_t1,
                                                            p_inspire1_t2:p_inspire21_t2,p_inspire1_t3:p_inspire21_t3,
                                                            p_inspire1_t4:p_inspire21_t4,p_inspire1_t4_suvey5:p_inspire21_t4_suvey5, 
                                                        p_paid1_t1:p_paid18_t1,p_paid1_t2:p_paid18_t2,p_paid1_t3:p_paid18_t3,
                                                        p_paid1_t4:p_paid18_t4,p_paid1_t4_suvey5:p_paid18_t4_suvey5)

# sub 1 from cpaid then reverse score, remove 5 from inspires
clinicalp_f = clinicalp_f %>% mutate(across(p_paid1_t1:p_paid18_t4_suvey5, ~.-1))
clinicalp_f = clinicalp_f %>% mutate(across(p_paid1_t1:p_paid18_t4_suvey5, ~4-.))

clinicalp_f = clinicalp_f %>% mutate(across(p_inspire1_t1:p_inspire21_t4_suvey5, remove5))

# score paid and inspire bl  
clinicalp_f = clinicalp_f %>% mutate(p_paid1 =  select(., p_paid1_t1:p_paid18_t1) %>% rowMeans(na.rm=T),
          p_paid_score1 =  p_paid1*25,
          p_paid2 =  select(., p_paid1_t2:p_paid18_t2) %>% rowMeans(na.rm=T),
          p_paid_score2 =  p_paid2*25,
          p_paid3=  select(., p_paid1_t3:p_paid18_t3) %>% rowMeans(na.rm=T),
          p_paid_score3 =  p_paid3*25,
          p_paid4 =  select(., p_paid1_t4:p_paid18_t4) %>% rowMeans(na.rm=T),
          p_paid_score4 =  p_paid4*25,
          p_paid5 =  select(., p_paid1_t4_suvey5:p_paid18_t4_suvey5) %>% rowMeans(na.rm=T),
          p_paid_score5 =  p_paid5*25,
          
          p_inspire_score1 = select(., p_inspire1_t1:p_inspire21_t1) %>% rowMeans(na.rm=T)*25,
          p_inspire_score2 = select(., p_inspire1_t2:p_inspire21_t2) %>% rowMeans(na.rm=T)*25,
          p_inspire_score3 = select(., p_inspire1_t3:p_inspire21_t3) %>% rowMeans(na.rm=T)*25,
          p_inspire_score4 = select(., p_inspire1_t4:p_inspire21_t4) %>% rowMeans(na.rm=T)*25,
          p_inspire_score5 = select(., p_inspire1_t4_suvey5:p_inspire21_t4_suvey5) %>% rowMeans(na.rm=T)*25)

clinicalp_f = clinicalp_f %>% select(record_id, p_paid_score1,p_paid_score2,p_paid_score3,p_paid_score4,p_paid_score5,
                                     p_inspire_score1,p_inspire_score2,p_inspire_score3,p_inspire_score4,p_inspire_score5)

# wide to long, add p_timepoint and timepoint vars
clinicalp_f_paid = clinicalp_f %>% select(record_id, p_paid_score1:p_paid_score5) %>% pivot_longer(cols=!record_id,
                                                                                                   names_to ="timepoint",
                                                                                                   names_prefix = "p_paid_score",
                                                                                                   values_to = "p_paid_score")
clinicalp_f_ins = clinicalp_f %>% select(record_id, p_inspire_score1:p_inspire_score5) %>% pivot_longer(cols=!record_id,
                                                                                                   names_to ="timepoint",
                                                                                                   names_prefix = "p_inspire_score",
                                                                                                   values_to = "p_inspire_score")
clinicalp_f_scores = full_join(clinicalp_f_ins,clinicalp_f_paid)
clinicalp_f_scores$p_timepoint = as.numeric(clinicalp_f_scores$timepoint)
# merge base and FU
clinical_p = full_join(clinicalp_bl, clinicalp_f_scores) %>% unique()
clinical_p = clinical_p %>% group_by(record_id) %>% fill(baseline_p_paid, .direction = "down")

# only include controliq=1, exclude medtronic (0)
# control iq users start at 120, ignore all else
clinical_p = clinical_p %>% filter(!(record_id %in% c("1a")))
clinical_p$record_id = as.numeric(clinical_p$record_id)
clinical_p = clinical_p %>% filter(record_id >= 120) %>% select(record_id, p_timepoint, p_paid_score,p_inspire_score,baseline_p_paid)

rm(clinicalp_bl, clinicalp_f, ClinicalObservationP_DATA_2024, clinicalp_f_ins,clinicalp_f_paid,clinicalp_f_scores)
#########################################################################
# join clinical obs
clinical = full_join(clinical_c, clinical_p, by = c("record_id", "timepoint" = "p_timepoint")) %>% unique()
clinical = clinical %>% select(record_id, timepoint, demographics_hcl:demographics_t1d_duration, hba1c:baseline_p_paid)
clinical = clinical %>% group_by(record_id) %>% fill(demographics_hcl:demographics_t1d_duration, .direction = "down")
rm(clinical_c,clinical_p)
########################################################################

############################OMNIPOD 5 COHORT################################


# CHILD BASELINE
op5c_bl = HCL_DATA_2024 %>% select(record_id, redcap_event_name, demo_c_age, demo_c_insurance,demo_c_hcl,demo_c_ethnicity,demo_c_race,demo_c_sex,demo_c_t1d_duration,gly_a1c,
                                   c_b_inspire1:c_b_inspire17, c_b_paid1:c_b_paid20, 
                                           ya_b_header_timepoint:ya_b_inspire17) %>% filter(redcap_event_name == "child_bl_arm_1")
# score surveys
op5c_bl =op5c_bl %>% mutate(cb_inspire_total = select(., c_b_inspire1:c_b_inspire17) %>% rowMeans()*25, 
                                               ya_b_inspire_total = select(., ya_b_inspire1:ya_b_inspire17) %>% rowMeans()*25,
                                               inspire_score = ifelse(demo_c_age < 13, cb_inspire_total, ya_b_inspire_total), # INSPIRE TOTAL
                                               
                                               
                                               cb_paid =  select(., c_b_paid1:c_b_paid20) %>% rowSums(),
                                               cb_paid_score = (80 - cb_paid)*25/20,
                                               ya_paid =  select(., ya_b_paid1:ya_b_paid20) %>% rowSums(),
                                               ya_paid_score = (80 - ya_paid)*25/20,
                                               paid_score = ifelse(demo_c_age < 13, cb_paid_score, ya_paid_score), # PAID
                            timepoint = 0,
                            baseline_paid = paid_score
                                   )  %>% select(record_id, timepoint,redcap_event_name, demo_c_age:gly_a1c, inspire_score, paid_score, baseline_paid)

# CHILD FU
op5c_fu = HCL_DATA_2024 %>% select(record_id,redcap_event_name,gly_a1c,c_f_timepoint:c_f_inspire17,
                                           ya_f_header_timepoint:ya_f_inspire17, seeds_fu1:seeds_fu20) %>% filter(redcap_event_name %in% c("child_m1_arm_1","child_m3_arm_1","child_m6_arm_1","child_m9_arm_1", "child_m12_arm_1"))
op5c_fu = left_join(op5c_fu, op5c_bl %>% select(record_id, demo_c_age) %>% unique())

op5c_fu =op5c_fu %>% mutate(cb_inspire_total = select(., c_f_inspire1:c_f_inspire17) %>% rowMeans()*25, 
                                               ya_f_inspire_total = select(., ya_f_inspire1:ya_f_inspire17) %>% rowMeans()*25,
                                               inspire_score = ifelse(demo_c_age < 13, cb_inspire_total, ya_f_inspire_total), # INSPIRE TOTAL
                                               
                                              
                                               cb_paid =  select(., c_f_paid1:c_f_paid20) %>% rowSums(),
                                               cb_paid_score = (80 - cb_paid)*25/20,
                                               ya_paid =  select(., ya_f_paid1:ya_f_paid20) %>% rowSums(),
                                               ya_paid_score = (80 - ya_paid)*25/20,
                                               paid_score = ifelse(demo_c_age < 13, cb_paid_score, ya_paid_score) # PAID
                                               

                                   )  %>% select(record_id, redcap_event_name, demo_c_age,gly_a1c, inspire_score, paid_score)

op5c_fu = op5c_fu %>% mutate(timepoint = case_when(redcap_event_name == "child_m1_arm_1" ~ 1,
                                                    redcap_event_name == "child_m3_arm_1" ~ 2,
                                                    redcap_event_name == "child_m6_arm_1" ~ 3,
                                                    redcap_event_name == "child_m9_arm_1" ~ 4,
                                                    redcap_event_name == "child_m12_arm_1" ~ 5))
# merge child
op5c = full_join(op5c_bl, op5c_fu) %>% unique() %>%mutate(hba1c = gly_a1c) %>% select(-redcap_event_name, - gly_a1c)
rm(op5c_bl,op5c_fu)
# PARENT BL
op5p_bl = HCL_DATA_2024 %>% select(record_id,redcap_event_name, p_b_inspire1:p_b_inspire21, p_b_hfs_behave1:p_b_hfs_worry26,
                                       p_b_paid1:p_b_paid18) %>% filter(redcap_event_name == "parent_bl_arm_1")

op5p_bl = op5p_bl %>% mutate(inspire_score_p = select(., p_b_inspire1:p_b_inspire21) %>% rowMeans()*25, # inspire

                                               paid_p =  select(., p_b_paid1:p_b_paid18) %>% rowSums(),
                                               paid_score_p = (72-paid_p)*25/18 ,
                             
                             timepoint = 0, 
                             baseline_p_paid = paid_score_p
                                   )  %>% select(record_id, timepoint,redcap_event_name, inspire_score_p,paid_score_p, baseline_p_paid)

# PARENT FU
op5p_fu = HCL_DATA_2024 %>% select(record_id,redcap_event_name, p_f_inspire1:p_f_inspire21, p_f_hfs_behave1:p_f_hfs_worry26,
                                       p_f_paid1:p_f_paid18) %>% filter(redcap_event_name %in% c("parent_m1_arm_1","parent_m3_arm_1","parent_m6_arm_1","parent_m9_arm_1", "parent_m12_arm_1"))

op5p_fu = op5p_fu %>% mutate(inspire_score_p = select(., p_f_inspire1:p_f_inspire21) %>% rowMeans()*25, 
                                      
                                               
                                               paid_p =  select(., p_f_paid1:p_f_paid18) %>% rowSums(),
                                               paid_score_p = (72-paid_p)*25/18 
                                   )  %>% select(record_id, redcap_event_name, inspire_score_p,paid_score_p)
op5p_fu = op5p_fu %>% mutate(timepoint = case_when(redcap_event_name == "parent_m1_arm_1" ~ 1,
                                                    redcap_event_name == "parent_m3_arm_1" ~ 2,
                                                    redcap_event_name == "parent_m6_arm_1" ~ 3,
                                                    redcap_event_name == "parent_m9_arm_1" ~ 4,
                                                    redcap_event_name == "parent_m12_arm_1" ~ 5))

op5p = full_join(op5p_bl, op5p_fu)
rm(op5p_bl,op5p_fu)
# merge datasets
op5 = full_join(op5c, op5p)
rm(op5c, op5p, HCL_DATA_2024)

op5$demographics_age = op5$demo_c_age
op5$demographics_ethnicity = op5$demo_c_ethnicity
op5$demographics_hcl = op5$demo_c_hcl
op5$demographics_race = op5$demo_c_race
op5$demographics_sex = op5$demo_c_sex
op5$demographics_insurance = op5$demo_c_insurance + 1
op5$demographics_t1d_duration = op5$demo_c_t1d_duration

# record_id, timepoint,redcap_event_name, demographics_t1d_duration, hba1c,

# child outcomes
analysis = full_join(analysis_gly_t1, psy_soc)

analysis$redcap_event_name = factor(analysis$redcap_event_name, levels = c("child_bl_arm_1","child_m1_arm_1","child_m3_arm_1","child_m6_arm_1","child_m9_arm_1", "child_m12_arm_1"),
                                    labels = c("Baseline", "Month 1", "Month 3", "Month 6","Month 9", "Month 12"))

```

-	Two separate mixed models (one for kids and one for parents) with tech satisfaction as a time-varying covariate predicting diabetes distress at each time point
-	Model covariates: baseline distress, A1C, sex, T1D duration
-	Hypothesis: Those with higher satisfaction will have lower diabetes distress.
-	Scoring guides for INSPIRE (tech satisfaction) and PAID (diabetes distress) appear at the end of this document
