---
title: "PICLS Data Explore"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(Hmisc)

# data import
picls = read_csv("S:/Laura/BDC/Projects/Janet Snell-Bergeon/PICLS/PICLS_CS_FORMS.csv")
preconception = picls %>% select(pid:redcap_repeat_instance,a1c_precon:preconception_complete)
smbg_q_ri =  picls %>% select(pid:redcap_repeat_instance,day1_basal_pm_run_in:smbg_questionnaire_run_in_complete)
smbg_q_ra = picls %>% select(pid:redcap_repeat_instance,day1_brkft_time:smbg_questionnaire_randomization_complete)
hfs = picls %>% select(pid:redcap_repeat_instance,fear_beh_bedtime_snacks:hypoglycemic_fear_survey_complete)
health_sf36 = picls %>% select(pid:redcap_repeat_instance,health_general:health_survey_sf36_complete)
insulin = picls %>% select(pid:redcap_repeat_instance,ins_complicated:insulin_delivery_satisfaction_survey_complete)
glucose = picls %>% select(pid:redcap_repeat_instance,gluc_visit_date: glucose_monitoring_satisfaction_survey_complete)
acet = picls %>% select(pid:redcap_repeat_instance,acet_time:acetaminophen_complete)
qol = picls %>% select(pid:redcap_repeat_instance,qol_consent:qol_complete)
```

FORMS to do: 
SMBG Questionnaire runin
SMBG Questionnaire rand
HFS
Health Survey 36
Insulin Delivery Satisfatcion survey
glucose monitoring satisfacion survey
qol

# preconception

34 subj
precon complete: n = 32, 1 incomplete, 13 unverified, 18 complt

a1c: n = 30 (4 missing?) range 6-13, most around 7
weight (kg): n = 30, range: 58-113, most around 70s
sbp: n = 30, range : 92-150, most around 110
dbp: n = 30, 58- 93, most around 70-80, 3 at 87 - char var orig
hr: n = 29, 57-153; 1 subj at 126,153, text says max should be 150
ins delivery: n = 30 only mdi (2) and pump (28)
ins basal type : n = 31, all 1,3,7,8,9
ins basal dose: n = 27, 12-58, most around 20-40
ins bol t: n = 31, only 1 2 3
ins bol dose: n = 23, 8-39, most arounf 13-30

```{r precon, echo=FALSE}
# preconception
preconception$dbp_precon = as.numeric(preconception$dbp_precon)
describe(preconception)
```

# acetaminophen

n = 177, of 34 subj, 18 unverified, 159 complete
all times of day used, n = 168

```{r acet}
# acetaminophen
describe(acet)
```

# qol


```{r qol}
describe(qol)
```