---
title: "PICLS Data Explore"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(Hmisc)

# data import
picls = read_csv("S:/Laura/BDC/Projects/Janet Snell-Bergeon/PICLS/PICLS_CS_FORMS.csv")
preconception = picls %>% select(pid:redcap_repeat_instance,a1c_precon:preconception_complete)
smbg_q_ri =  picls %>% select(pid:redcap_repeat_instance,day1_basal_pm_run_in:smbg_questionnaire_run_in_complete)
smbg_q_ra = picls %>% select(pid:redcap_repeat_instance,day1_brkft_time:smbg_questionnaire_randomization_complete)
hfs = picls %>% select(pid:redcap_repeat_instance,fear_beh_bedtime_snacks:hypoglycemic_fear_survey_complete)
health_sf36 = picls %>% select(pid:redcap_repeat_instance,health_general:health_survey_sf36_complete)
insulin = picls %>% select(pid:redcap_repeat_instance,ins_complicated:insulin_delivery_satisfaction_survey_complete)
glucose = picls %>% select(pid:redcap_repeat_instance,gluc_visit_date: glucose_monitoring_satisfaction_survey_complete)
acet = picls %>% select(pid:redcap_repeat_instance,acet_time:acetaminophen_complete)
qol = picls %>% select(pid:redcap_repeat_instance,qol_consent:qol_complete)


# lauras
picls2 = read_csv("S:/Laura/BDC/Projects/Janet Snell-Bergeon/PICLS/PICLS_LP_FORMS.csv")
exemption = picls2 %>% select(pid:redcap_repeat_instance, incl___1:exemption_waiver_complete)
pregvis = picls2 %>% select(pid:redcap_repeat_instance,preg_visit_num:pregnancy_visits_complete)
igm = picls2 %>% select(pid:redcap_repeat_instance,interim_mngmt_change___1:interim_glucose_management_complete)
foodlog = picls2 %>% select(pid:redcap_repeat_instance, food_gest_wks:food_log_complete)
nc = picls2 %>% select(pid:redcap_repeat_instance, nutr_extra_reason_2___1:nutrition_counseling_complete)
ret = picls2 %>% select(pid:redcap_repeat_instance, retinop_precon_right:retinopathy_complete)
labs = picls2 %>% select(pid:redcap_repeat_instance,trimester:labs_complete)

```


# exemption
n = 20, all completed
```{r exem}
describe(exemption)
```

# preg visits

120a bmi recorded at 60.6, next highest is in 40s


```{r preg}
describe(pregvis)
```


# interim glucose management

n = 547 complete subj visits

```{r igm}
describe(igm)
```

# food log
n = 75 completed the food log (3 unverified)
56 had gest wks, days
74 had food day 1,3, 72 had day 2


```{r food}
describe(foodlog)
```

# nutrition counseling
n = 24 (4 unverified)

all complete cases

```{r nc}
describe(nc)
```

# retinopathy

mostly missing data, n = 22, 1 unv 1 incomplt

most cells are missing in this data frame
```{r ret}
describe(ret)
```

# labs

n = 25, all unverified, seems ok

most cells are missing in this data frame
```{r labs}
describe(labs)
```


# preconception

34 subj
precon complete: n = 32, 1 incomplete, 13 unverified, 18 complt

a1c: n = 30 (4 missing?) range 6-13, most around 7
weight (kg): n = 30, range: 58-113, most around 70s
sbp: n = 30, range : 92-150, most around 110
dbp: n = 30, 58- 93, most around 70-80, 3 at 87 - char var orig
hr: n = 29, 57-153; 1 subj at 126,153, text says max should be 150
ins delivery: n = 30 only mdi (2) and pump (28)
ins basal type : n = 31, all 1,3,7,8,9
ins basal dose: n = 27, 12-58, most around 20-40
ins bol t: n = 31, only 1 2 3
ins bol dose: n = 23, 8-39, most arounf 13-30

```{r precon, echo=FALSE}
# preconception
preconception$dbp_precon = as.numeric(preconception$dbp_precon)
describe(preconception)
```

# acetaminophen

n = 177, of 34 subj, 18 unverified, 159 complete
all times of day used, n = 168

```{r acet}
# acetaminophen
describe(acet)
```

# qol

n = 21 for consent and 22 for consent dates,
17-19 for rest of measured variables
```{r qol}
describe(qol)
```

# glucose monitoring satisfacion survey

https://behavioraldiabetes.org/xwp/wp-content/uploads/2015/12/GMSS-T1D.pdf
sum up the items, reverse code 2-7, 11-13, 15

range should be between 15 and 75
```{r glucose }
# lets just see how many scores we have and a range before reverse coding
glucose = glucose %>% mutate(
  total = gluc_satisfied + gluc_less_restricted + gluc_frustrated + gluc_spontaneous + gluc_more_open +
    (6 - gluc_think)+ (6 - gluc_time)+ (6 - gluc_inaccurate)+ (6 - gluc_worry)+ (6 - gluc_hassle)+ (6 - gluc_trust)+ (6 - gluc_irritations)+ (6 - gluc_nonsense)+ (6 - gluc_depressed)+ (6 - gluc_painful))

describe(glucose)
```

# Health Survey 36

92 surveys taken, 1 missing for health emotions less work

im only counting 35 questions, i think one of the physical is missing
only 9 physical questions; i think there should be 10? 3-12

```{r health sf36}
# health36 does bunch of different sub scores

# physical: 3:12
# health_sf36 = health_sf36 %>% mutate(
#   physical_tot = (health_lim_vigor - 1)*50 +  (health_lim_vigor - 1)*50 +
# )

describe(health_sf36)
```

# insulin delivery satisfatcion survey

https://behavioraldiabetes.org/xwp/wp-content/uploads/2015/12/Insulin-Device-Satisfaction-Survey.pdf
sum up the items, reverse code 1, 3,4,7-13, rnage should be 14-70

n = 93, all complete


```{r insulin delivery }

insulin = insulin %>% mutate(
  total_idss = (6 -ins_complicated) + (ins_control_glucose) + (6 -ins_embarrass) + (6 -ins_hassle_use) + (ins_control_diabetes) + (ins_works_well) + (6-ins_expensive)
  + (6-ins_spontaneous)+ (6-ins_many_pieces)+ (6-ins_little_benefit)+ (6-ins_inconvenient)+ (6-ins_time)+ (6-ins_hassle_carry)+ (ins_positive)
)

describe(insulin)
```

# hypoglycemeia fear survey

n = 89, 3 incomplete, all totals in range though

https://olucdenver-my.sharepoint.com/:w:/r/personal/laura_pyle_cuanschutz_edu/_layouts/15/Doc.aspx?sourcedoc=%7B879447E4-3197-4273-B46D-5A536009FD83%7D&file=670G%20Clinical%20Observation%20Study_Survey%20scoring%20instructions_CIQ%2012%20months.docx&action=default&mobileredirect=true

27 questions, so we sum all up?

```{r hfs}
hfs = hfs %>% mutate(hfs_tot = rowSums(.[5:31]) - 26) # subtract 1 for each response(26 questions)

describe(hfs)
```


# smbg

```{r smbg}

```