---
title: "Insulin Cost Analysis"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(MASS)
library(table1)
library(readxl)
# data import
Insulin_cost_whole_sample <- read_excel("S:/Laura/BDC/Projects/Lauren Waterman/Insulin cost analysis/Insulin_cost_whole sample.xlsx", 
    sheet = "Insulin_whole sample")

```

```{r dc, include = F}

Insulin_cost_whole_sample$friends_family =  Insulin_cost_whole_sample$`Friends/Family (binary: 0=0, 1=any)`
analysis = Insulin_cost_whole_sample %>% replace(.=="NULL", NA)%>%mutate(insulin_cost = case_when(`Cost per month` == "NULL" ~ NA_real_,
                                                                         TRUE ~ as.numeric(`Cost per month`)),
                                                insurance = case_when(`insurance_type (3=private, 2=public, 1=tricare, 4=other, 5=none, 6=unsure)` == 3 ~ "Private",
                                                                      `insurance_type (3=private, 2=public, 1=tricare, 4=other, 5=none, 6=unsure)` == 2 ~ "Public",
                                                                      `insurance_type (3=private, 2=public, 1=tricare, 4=other, 5=none, 6=unsure)` == 1 ~ "Tricare",
                                                                      `insurance_type (3=private, 2=public, 1=tricare, 4=other, 5=none, 6=unsure)` == 4 ~ "Other",
                                                                      `insurance_type (3=private, 2=public, 1=tricare, 4=other, 5=none, 6=unsure)` == 6 ~ "Unsure"),
                                                race = case_when(`race (0=w, 1=b, 2=Asian, 3=native hawaiian, 4=american indian, 5=multiple, 6=?)` == 0 ~"White",
                                                                 `race (0=w, 1=b, 2=Asian, 3=native hawaiian, 4=american indian, 5=multiple, 6=?)` == 1 ~"Black",
                                                                 `race (0=w, 1=b, 2=Asian, 3=native hawaiian, 4=american indian, 5=multiple, 6=?)` == 2 ~"Asian",
                                                                 `race (0=w, 1=b, 2=Asian, 3=native hawaiian, 4=american indian, 5=multiple, 6=?)` == 3 ~"Native Hawaiian",
                                                                 `race (0=w, 1=b, 2=Asian, 3=native hawaiian, 4=american indian, 5=multiple, 6=?)` == 4 ~"American Indian",
                                                                 `race (0=w, 1=b, 2=Asian, 3=native hawaiian, 4=american indian, 5=multiple, 6=?)` == 5 ~"Multiple",
                                                                 `race (0=w, 1=b, 2=Asian, 3=native hawaiian, 4=american indian, 5=multiple, 6=?)` == 6 ~"Unknown"),
                                                sex = case_when(`gender (1=m, 2=f, 3=o)` == 1 ~ "Male",
                                                                `gender (1=m, 2=f, 3=o)` == 2 ~ "Female",
                                                                `gender (1=m, 2=f, 3=o)` == 3 ~ "Other"),
                                                ethnicity = case_when(`ethnicity (0=H, 1=NH)`== 0 ~ "Hispanic",
                                                                      `ethnicity (0=H, 1=NH)` == 1 ~ "Nonhispanic"),
                                                diabetes_duration = `Diabetes duration`,
                                                emerging_adult = `Emerging Adult`,
                                                instate = `instate (1=yes)`,
                                                ration1 = `Run out (binary: 0=0, 1=any)`,
                                                ration2 = friends_family,
                                                ration3 = `Not filled Rx: cost (binary: 0=0, 1=any)`,
                                                ration4 = `Running out (binary: 0=0, 1=any)`, 
                                                ration5 = `Paid case (binary: 0=0, 1=any)`,
                                                ration6 = `Black market insulin (binary: 0=0, 1=any)`,
                                                ration7 = `Ration due to cost (binary: 0=0, 1=any)`) %>%  filter(!is.na(record_id))
analysis[99:105] = lapply(analysis[99:105], as.numeric)
analysis = analysis %>% mutate(rationsum = rowSums(across(ration1:ration7)),
                               rationing = case_when(rationsum == 0 ~ "Non Rationer",
                                                     rationsum > 0 ~ "Rationer"))
analysis$TIR = as.numeric(analysis$TIR)

# likerts 

# aim 6
analysis$walmart = 1 - (as.numeric(analysis$`did you know about walmart? (1=y, 2=n)`) -1)
table(analysis$walmart)
```

```{r function, include = F}
# fit model fun
fit_mod = function(outcome,compgroup,df,age=FALSE){

  if(age==FALSE){
    # Fit model
    
    f = as.formula(paste0(outcome,"~",compgroup,"+ sex + diabetes_duration + race + ethnicity + insurance"))
    mod = lm(f, data = df, na.action = na.omit)
    summary(mod)
  }
    else{
    # Fit model
    
    f = as.formula(paste0(outcome,"~",compgroup,"+ sex + diabetes_duration + race + ethnicity + insurance + age"))
    mod = lm(f, data = df, na.action = na.omit)
    summary(mod)
  }
}

fit_lr_mod = function(outcome,compgroup,df){
  print(outcome)
    f = as.formula(paste0(outcome,"~",compgroup,"+ sex + diabetes_duration + race + ethnicity + insurance"))
    mod = glm(f, data = df, na.action = na.omit, family = binomial(link='logit'))
    print(summary(mod))
    print("Model Coefficient Odds Ratios")
    print(exp(coef(mod)))
}

fit_lik_lr_mod = function(outcome,compgroup,df){
  df = df %>% filter(paste0(outcome) != "2") %>% mutate(newoutcome = as.numeric(paste0(outcome)) - 1)
  
    f = as.formula(paste0("newoutcome","~",compgroup,"+ sex + diabetes_duration + race + ethnicity + insurance"))
    mod = glm(f, data = df, na.action = na.omit, family = binomial(link='logit'))
    summary(mod)
}


change_likert = function(x){as.factor(as.numeric(x) - 1)}
nm1 <- grep("(binary 1=D, 2=A, 3=NS)", names(analysis), value = TRUE)
analysis[nm1] <- lapply(analysis[nm1], change_likert)
```
# Outstanding Data Remarks

- Record ID 12 has out of pocket costs and some other columns as "NULL" --> set to NA
- Wasn't sure how to name the rationing columns so i went with the first few words of the column name in the frequency tables

# Methods

For all models, adjusting for sex, diabetes duration, race/eth, insurance, when applicable add in age

"is there a difference in out of pocket costs between groups?"

aim 1 and 2: linear models with above covariates

aim 4,5,6: logistic regression with above covariates


# Analysis

## Demographics by subgrouping

### Emerging Adults

```{r}
table1(~sex + race + ethnicity + diabetes_duration + insurance + pump_user+ age + HbA1c|emerging_adult, data = analysis)
```

### Instate

```{r}
table1(~sex + race + ethnicity + diabetes_duration + insurance + pump_user+ age + HbA1c|instate, data = analysis %>% filter(!is.na(instate)))

```


### Rationing

```{r}
table1(~sex + race + ethnicity + diabetes_duration + insurance + pump_user + age + HbA1c|rationing, data = analysis %>% filter(!is.na(rationing)))

```

## Aim 1: Comparisons of out of pocket costs (col e)

### Emerging adults vs parents/guardians (col d)

There is no significant difference in out of pocket costs between Emerging adults and Parents/Guardians (p = 0.98)

```{r}
fit_mod("insulin_cost", "emerging_adult", df = analysis, age=FALSE )
```

### instate v out of state (col i)

There is no significant difference in out of pocket costs between instate and out of state (p = 0.135)


```{r}
fit_mod("insulin_cost", "instate", df = analysis, age=FALSE )
```

### rationers

There is a significant difference in oop costs between rationers and non rationers (p = 0.004). On average, Rationers spend 52.14 (SE = 17.70) higher than nonrationers. 

```{r}
fit_mod("insulin_cost", "rationing", df = analysis, age=FALSE )
```

## Aim 2: hba1c and TIR

### Emerging adults vs parents/guardians (col d)

There is a significant difference in TIR between emerging adults and parents/guardians (p = 0.01). On average, Emerging adults have 14.4 (SE = 5.7) higher TIR than parents/guardians. 
There is a significant difference in a1c between emerging adults and parents/guardians (p = 0.004). On average, Emerging adults have .95(SE = .32) lower a1c than parents/guardians. 

```{r}
fit_mod("TIR", "emerging_adult", df = analysis, age=TRUE )
```

```{r}
fit_mod("HbA1c", "emerging_adult", df = analysis, age=TRUE )
```

### instate v out of state (col i)

There is no significant difference in TIR between instate and out of state subjects (p = 0.39).
There is no significant difference in a1c between instate and out of state subjects (p = 0.37).

```{r}
fit_mod("TIR", "instate", df = analysis, age=TRUE )
```

```{r}
fit_mod("HbA1c", "instate", df = analysis, age=TRUE )
```

### rationers

There is a significant difference in TIR between rationers and nonrationers (p = 0.02). On average, rationers have 9.78 (SE = 4.03) lower TIR than nonrationers. 
There is a significant difference in a1c between rationers and nonrationers (p = 0.003). On average, rationers have 0.69 (SE =.23) higher TIR than nonrationers. 


```{r}
fit_mod("TIR", "rationing", df = analysis, age=TRUE )
```

```{r}
fit_mod("HbA1c", "rationing", df = analysis, age=TRUE )
```

## Aim 4: comparing rationing response among subgroups

### emerging adult

Running out: emerging adult is associated with Running out p = 0.03** Emerging adults are more likely than parents/guardians to experience Running out by a factor of 4.1 times.

Other rationing variables not significantly different for emerging adults (see results below)

```{r}
ea_t = analysis %>% group_by(emerging_adult) %>% summarise(n = n(),
                                                    run_out = sum(ration1, na.rm = T),
                                                    friends_family = sum(ration2, na.rm = T),
                                                    not_filled_rx = sum(ration3, na.rm = T),
                                                    running_out = sum(ration4, na.rm = T),
                                                    paid_case = sum(ration5, na.rm = T),
                                                    black_market = sum(ration6, na.rm = T),
                                                    ration_cost = sum(ration7, na.rm = T))
kable(ea_t)

print("Run Out")
fit_lr_mod("ration1", "emerging_adult", df = analysis)

print("Friends/Family")
fit_lr_mod("ration2", "emerging_adult", df = analysis)

print("Not Filled Rx")
fit_lr_mod("ration3", "emerging_adult", df = analysis)

print("Running Out")
fit_lr_mod("ration4", "emerging_adult", df = analysis)

print("Paid Case")
fit_lr_mod("ration5", "emerging_adult", df = analysis)

print("Black Market")
fit_lr_mod("ration6", "emerging_adult", df = analysis)

print("Ration due to Cost")
fit_lr_mod("ration7", "emerging_adult", df = analysis)

```

### instate

Not filled Rx: instate/outstate is associated with not filling rx (p = 0.01)
Running out:instate/outstate is associated with running out (p = 0.048)
Paid Case: instate/outstate is associated with paid case (p = 0.006)
ration due to cost: instate/outstate is associated with ration due to cost (p = 0.047)

```{r}
analysis %>% group_by(instate) %>% summarise(n = n(),
                                                    run_out = sum(ration1, na.rm = T),
                                                    friends_family = sum(ration2, na.rm = T),
                                                    not_filled_rx = sum(ration3, na.rm = T),
                                                    running_out = sum(ration4, na.rm = T),
                                                    paid_case = sum(ration5, na.rm = T),
                                                    black_market = sum(ration6, na.rm = T),
                                                    ration_cost = sum(ration7, na.rm = T))

print("Run Out")
fit_lr_mod("ration1", "instate", df = analysis)

print("Friends/Family")
fit_lr_mod("ration2", "instate", df = analysis)

print("Not Filled Rx")
fit_lr_mod("ration3", "instate", df = analysis)

print("Running Out")
fit_lr_mod("ration4", "instate", df = analysis)

print("Paid Case")
fit_lr_mod("ration5", "instate", df = analysis)

print("Black Market")
fit_lr_mod("ration6", "instate", df = analysis)

print("Ration due to Cost")
fit_lr_mod("ration7", "instate", df = analysis)
```

## likerts

For this section, to run the logistic regressions I subtracted 1 from the value to get a 0 and 1 value for D and A (filtering out Not sure answers)

### emerging adult frequencies

Significant associations:

Plan if price more than expected (p = 0.04)


```{r}
with(analysis, table(emerging_adult, `Too expensive (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`Insulin cap effective (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`PA issues (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`Skipped bills (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`Plan if price more than expected (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`Know who to contact (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(emerging_adult,`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)`))


fit_lr_mod("`Too expensive (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Too expensive (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Insulin cap effective (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Insulin cap effective (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`PA issues (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`PA issues (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Skipped bills (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Skipped bills (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Plan if price more than expected (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Plan if price more than expected (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Know who to contact (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Know who to contact (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)`", "emerging_adult", df = analysis %>% filter(`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)` != "2"))
```


### instate

Significant associations:

Too Expensive (p = 0.012)
Skipped Bills (p = 0.001)
Avoided appts due to cost (p = 0.0003)


```{r}
with(analysis, table(instate, `Too expensive (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`Insulin cap effective (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`PA issues (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`Skipped bills (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`Plan if price more than expected (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`Know who to contact (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(instate,`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)`))


fit_lr_mod("`Too expensive (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Too expensive (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Insulin cap effective (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Insulin cap effective (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`PA issues (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`PA issues (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Skipped bills (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Skipped bills (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Plan if price more than expected (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Plan if price more than expected (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Know who to contact (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Know who to contact (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)`", "instate", df = analysis %>% filter(`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)` != "2"))


```

### rationers

Significant associations:

Too Expensive (p < 0.0001)
Insulin Unaffordable next 3 Months ( p < 0.0001)
Skipped Bills (p = 0.002)
Plan if Price more than expected (p = 0.003)
Avoided appts due to cost (p = 0.011)

```{r}
with(analysis, table(rationing, `Too expensive (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`Insulin cap effective (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`PA issues (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`Skipped bills (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`Plan if price more than expected (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`Know who to contact (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)`))
with(analysis, table(rationing,`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)`))



fit_lr_mod("`Too expensive (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Too expensive (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Insulin unaffordable next 3 months (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Insulin cap effective (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Insulin cap effective (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`PA issues (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`PA issues (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Skipped bills (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Skipped bills (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Plan if price more than expected (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Plan if price more than expected (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Know who to contact (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Know who to contact (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Comfortable with BDC team (binary 1=D, 2=A, 3=NS)` != "2"))
fit_lr_mod("`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)`", "rationing", df = analysis %>% filter(`Avoided appts due to cost (binary 1=D, 2=A, 3=NS)` != "2"))
```
## Insulin Access

```{r, include=FALSE}
# dichotomize the other variables
analysis = analysis %>% mutate(walmart_lr = case_when(`did you know about walmart? (1=y, 2=n)` == 2 ~ 0,
                                                      `did you know about walmart? (1=y, 2=n)` == 1 ~ 1),
                               know_insulin = case_when(`what company makes your insulin (1=lily, 2=Novo, 5=?` == 5 ~ 0,
                                                        `what company makes your insulin (1=lily, 2=Novo, 5=?` %in% c(1,2) ~ 1),
                               know_programs = case_when(`tried applying for programs (1=n, didn't know, 2= no but don't need, 3=y but not qualify, 4=received)` %in% c(1,2) ~ 0,
                                                         `tried applying for programs (1=n, didn't know, 2= no but don't need, 3=y but not qualify, 4=received)` %in% c(3,4) ~ 1))

```

### Emerging Adults

*insufficient model convergence for knowing programs (combining 1,2 and 3,4) ~ not a ton in the 3,4 group, similar for lily/novo sub popns

```{r}
table1(~factor(`did you know about walmart? (1=y, 2=n)`) +
         factor(`what company makes your insulin (1=lily, 2=Novo, 5=?`) +
         factor(`tried applying for programs (1=n, didn't know, 2= no but don't need, 3=y but not qualify, 4=received)`)|emerging_adult, data = analysis, overall = F )

table1(~factor(`Know about lily (0=n)`)|emerging_adult, data = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==1) )
table1(~factor(`Know about novo (0=n)`)|emerging_adult, data = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==2) )

fit_lr_mod("walmart_lr", "emerging_adult", df = analysis)
fit_lr_mod("know_insulin", "emerging_adult", df = analysis)
#fit_lr_mod("know_programs", "emerging_adult", df = analysis)

# 
# fit_lr_mod("`Know about lily (0=n)`", "emerging_adult", df = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==1))
# fit_lr_mod("`Know about novo (0=n)`", "emerging_adult", df = analysis%>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==2))

```



### In State

*insufficient model convergence for knowing programs (combining 1,2 and 3,4) ~ not a ton in the 3,4 group, similar for lily/novo sub popns


```{r}
table1(~factor(`did you know about walmart? (1=y, 2=n)`) +
         factor(`what company makes your insulin (1=lily, 2=Novo, 5=?`) +
         factor(`tried applying for programs (1=n, didn't know, 2= no but don't need, 3=y but not qualify, 4=received)`)|instate, data = analysis %>% filter(!is.na(instate)), overall = F )

table1(~factor(`Know about lily (0=n)`)|instate, data = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==1)%>% filter(!is.na(instate)) )
table1(~factor(`Know about novo (0=n)`)|instate, data = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==2)%>% filter(!is.na(instate)) )

fit_lr_mod("walmart_lr", "instate", df = analysis)
fit_lr_mod("know_insulin", "instate", df = analysis)
# fit_lr_mod("know_programs", "instate", df = analysis)
# 
# 
# fit_lr_mod("`Know about lily (0=n)`", "instate", df = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==1))
# fit_lr_mod("`Know about novo (0=n)`", "instate", df = analysis%>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==2))
```

### Rationers

*insufficient model convergence for knowing programs (combining 1,2 and 3,4) ~ not a ton in the 3,4 group, similar for lily sub popns


```{r}
table1(~factor(`did you know about walmart? (1=y, 2=n)`) +
         factor(`what company makes your insulin (1=lily, 2=Novo, 5=?`) +
         factor(`tried applying for programs (1=n, didn't know, 2= no but don't need, 3=y but not qualify, 4=received)`)|rationing, data = analysis%>% filter(!is.na(rationing)), overall = F )

table1(~factor(`Know about lily (0=n)`)|rationing, data = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==1)%>% filter(!is.na(rationing)) )
table1(~factor(`Know about novo (0=n)`)|rationing, data = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==2) %>% filter(!is.na(rationing)))

fit_lr_mod("walmart_lr", "rationing", df = analysis)
fit_lr_mod("know_insulin", "rationing", df = analysis)
# fit_lr_mod("know_programs", "rationing", df = analysis)
# 
# 
# fit_lr_mod("`Know about lily (0=n)`", "rationing", df = analysis %>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==1))
fit_lr_mod("`Know about novo (0=n)`", "rationing", df = analysis%>% filter(`what company makes your insulin (1=lily, 2=Novo, 5=?` ==2))
```