---
title: "6 Month Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(redcapAPI)
library(tidyverse)
library(knitr)
library(tableone)
library(readxl)
library(nlme)
library(gridExtra)
source("/Users/timvigers/Documents/GitHub/Tim-and-Laura/tim_R_functions.R")
# Check OS and alter file path accordingly.
if (.Platform$OS.type == "windows") {pathstart <- "//ucdenver.pvt/"} else if (.Platform$OS.type == "unix"){pathstart <- "/Volumes/"}
```

```{r echo=FALSE,cache=TRUE,include=FALSE}
# REDCap API data import
source("/Users/timvigers/Documents/GitHub/BDC-Code/api_tokens.R")
# YA
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = YA670g)
ya_full <- exportRecords(rcon)
# Parents
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = Parent670g)
parent_full <- exportRecords(rcon)
# Child
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = Child670g)
# Read in baseline glycemic data
baseline_glycemic_csv <- read.csv("/Volumes/som-1/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Cari Berget/6 Month Manuscript/Data_Cleaned/670GChild_BaselineGlycemicData_14JAN2019.csv")
```

```{r echo=FALSE,cache=TRUE,include=FALSE}
child_full <- read.csv("/Volumes/som-1/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Cari Berget/6 Month Manuscript/Data_Cleaned/670GChild_DATA_2019-05-06.csv")
# Read in 
# Format child glycemic
child_full$record_id <- as.character(child_full$record_id)
child_full$c_t1_date_m1 <- lubridate::mdy(as.character(child_full$c_t1_date_m1))
child_full$automode_start <- lubridate::mdy(as.character(child_full$automode_start))
child_full$c_t1_date <- lubridate::mdy(as.character(child_full$c_t1_date))
child_full$c_t2_date <- lubridate::mdy(as.character(child_full$c_t2_date))
# Demographics
demographics_child <- child_full %>%
  select(record_id,automode_start,demographics_t1d_duration,demographics_age:demographics_cgmhx)
demographics_ya <- ya_full %>%
  select(record_id,ya_automoste_start,demographics_t1d_duration,ya_demographics_age:ya_demographics_cgmhx)
demographics_ya[,3:ncol(demographics_ya)] <- lapply(demographics_ya[,3:ncol(demographics_ya)],as.numeric)
colnames(demographics_ya) <- sub("ya_","",colnames(demographics_ya))
colnames(demographics_ya)[2] <- "automode_start"
# Filter YA out of child data
ya <- as.character(demographics_ya$record_id)
demographics_child <- demographics_child[-c(which(demographics_child$record_id %in% ya)),]
child <- as.character(demographics_child$record_id)
demographics <- rbind(demographics_child,demographics_ya) %>%
  arrange(as.numeric(record_id))
# Never trained and withdrawn
withdrawn <- c("41","50") # Per Cari
never_trained <- demographics$record_id[which(is.na(demographics$automode_start))]
never_trained <- never_trained[which(!(never_trained %in% withdrawn))]
# Age group
demographics$age_group <- cut(demographics$demographics_age, 
                              breaks = c(0,14,18,Inf),
                              labels = c("< 14","14 - 17","18 +"),right = F)
# Format glycemic data
glycemic_colnames <- c("record_id","days","hba1c","am_time","mm_time",
                       "sensor_wear","sensor_u54","sensor_55_69",
                       "sensor_70_180","sensor_181_250","sensor_g250",
                       "mean_sg","sd","bg_checks","calibrations",
                       "tdd","basal","bolus","amexits",
                       "amexit_day","amexit_hyper","amexit_hypo",
                       "amexit_manual","amexit_other")
# Baseline
baseline_hba1c <- child_full %>%
  select(record_id,hba1c_baseline)
baseline_hba1c$record_id <- as.integer(baseline_hba1c$record_id)
baseline_glycemic <- left_join(baseline_glycemic_csv,baseline_hba1c, by = "record_id")
colnames(baseline_glycemic) <- sub("_b","",colnames(baseline_glycemic))
colnames(baseline_glycemic)[ncol(baseline_glycemic)] <- "hba1c"
baseline_glycemic[,c(glycemic_colnames[which(!(glycemic_colnames %in% colnames(baseline_glycemic)))])] <- NA
baseline_glycemic <- baseline_glycemic[glycemic_colnames]
baseline_glycemic <- as.data.frame(lapply(baseline_glycemic, as.numeric))
baseline_glycemic$Timepoint <- "B"
# M1
m1_glycemic_child <- child_full %>%
  select(record_id,c_t1_date_m1,hba1c_m1,am_time_m1:amexits_other_m1) %>%
  filter(record_id %in% child)
m1_glycemic_child <- left_join(m1_glycemic_child,
                               demographics[,c("record_id","automode_start")],
                               by = "record_id")
m1_glycemic_child$days <- as.numeric(difftime(m1_glycemic_child$c_t1_date_m1,m1_glycemic_child$automode_start))
m1_glycemic_child <- m1_glycemic_child %>% select(record_id,days,hba1c_m1:amexits_other_m1)
colnames(m1_glycemic_child) <- glycemic_colnames
m1_glycemic_ya <- ya_full %>% 
  select(record_id,ya_t1_date_m1,ya_hba1c_t1_m1,ya_t1_am_time_m1:ya_t1_amexit_other_m1) %>%
  filter(record_id %in% ya)
m1_glycemic_ya <- left_join(m1_glycemic_ya,
                               demographics[,c("record_id","automode_start")],
                               by = "record_id")
m1_glycemic_ya$days <- as.numeric(difftime(m1_glycemic_ya$ya_t1_date_m1,m1_glycemic_ya$automode_start))
m1_glycemic_ya <- m1_glycemic_ya %>% select(record_id,days,ya_hba1c_t1_m1:ya_t1_amexit_other_m1)
colnames(m1_glycemic_ya) <- glycemic_colnames
m1_glycemic <- rbind(m1_glycemic_child,m1_glycemic_ya)
m1_glycemic <- as.data.frame(lapply(m1_glycemic, as.numeric))
m1_glycemic$Timepoint <- "M1"
# T1
t1_glycemic_child <- child_full %>%
  select(record_id,c_t1_date,c_hba1c_baseline_t1,c_t1_am_time:c_t2_amexits_other) %>%
  filter(record_id %in% child)
t1_glycemic_child <- left_join(t1_glycemic_child,
                               demographics[,c("record_id","automode_start")],
                               by = "record_id")
t1_glycemic_child$days <- as.numeric(difftime(t1_glycemic_child$c_t1_date,t1_glycemic_child$automode_start))
t1_glycemic_child <- t1_glycemic_child %>% select(record_id,days,c_hba1c_baseline_t1:c_t2_amexits_other)
colnames(t1_glycemic_child) <- glycemic_colnames
t1_glycemic_ya <- ya_full %>% 
  select(record_id,ya_t1_date,ya_hba1c_t1,ya_t1_am_time:ya_t1_amexit_other) %>%
  filter(record_id %in% ya)
t1_glycemic_ya <- left_join(t1_glycemic_ya,
                               demographics[,c("record_id","automode_start")],
                               by = "record_id")
t1_glycemic_ya$days <- as.numeric(difftime(t1_glycemic_ya$ya_t1_date,t1_glycemic_ya$automode_start))
t1_glycemic_ya <- t1_glycemic_ya %>% select(record_id,days,ya_hba1c_t1:ya_t1_amexit_other)
colnames(t1_glycemic_ya) <- glycemic_colnames
t1_glycemic <- rbind(t1_glycemic_child,t1_glycemic_ya) %>%
  arrange(as.numeric(record_id))
t1_glycemic <- as.data.frame(lapply(t1_glycemic, as.numeric))
t1_glycemic$Timepoint <- "T1"
# T2
t2_glycemic_child <- child_full %>%
  select(record_id,c_t2_date,c_hba1c_baseline_t2,c_t2_am_time:c_t2_amexit_other) %>%
  filter(record_id %in% child)
t2_glycemic_child <- left_join(t2_glycemic_child,
                               demographics[,c("record_id","automode_start")],
                               by = "record_id")
t2_glycemic_child$days <- as.numeric(difftime(t2_glycemic_child$c_t2_date,t2_glycemic_child$automode_start))
t2_glycemic_child <- t2_glycemic_child %>% select(record_id,days,c_hba1c_baseline_t2:c_t2_amexit_other)
colnames(t2_glycemic_child) <- glycemic_colnames
t2_glycemic_ya <- ya_full %>% 
  select(record_id,ya_t2_date,ya_hba1c_t2,ya_t2_am_time:ya_t2_amexit_other) %>%
  filter(record_id %in% ya)
t2_glycemic_ya <- left_join(t2_glycemic_ya,
                               demographics[,c("record_id","automode_start")],
                               by = "record_id")
t2_glycemic_ya$days <- as.numeric(difftime(t2_glycemic_ya$ya_t2_date,t2_glycemic_ya$automode_start))
t2_glycemic_ya <- t2_glycemic_ya %>% select(record_id,days,ya_hba1c_t2:ya_t2_amexit_other)
colnames(t2_glycemic_ya) <- glycemic_colnames
t2_glycemic <- rbind(t2_glycemic_child,t2_glycemic_ya) %>%
  arrange(as.numeric(record_id))
t2_glycemic <- as.data.frame(lapply(t2_glycemic, as.numeric))
t2_glycemic$Timepoint <- "T2"
# Bind timepoints, 
glycemic_data <- bind_rows(baseline_glycemic,m1_glycemic,t1_glycemic,t2_glycemic) 
# Baseline HbA1c group
glycemic_data <- glycemic_data %>%
  group_by(record_id) %>%
  mutate(hba1c_baseline = hba1c[1]) %>%
  mutate(hba1c_clinical = cut(hba1c_baseline,breaks = c(0,7.5,9.0,Inf),
                              labels = c("Low Baseline HbA1c","Medium Baseline HbA1c","High Baseline HbA1c"))) %>%
  arrange(record_id,Timepoint) %>%
  select(record_id,Timepoint,days,hba1c:hba1c_clinical) %>%
  filter(!(record_id %in% never_trained) & !(record_id %in% withdrawn))
# Discontinuers
t1_discont <- glycemic_data %>%
  filter(Timepoint == "T1", am_time < 10) %>%
  .$record_id
# Remove participant 10 from discontinuers, because she had AM > 10 at T2.
t1_discont <- t1_discont[-c(which(t1_discont == "10"))]
# T2 discontinuers
t2_discont <- glycemic_data %>%
  filter(Timepoint == "T2", am_time < 10) %>%
  .$record_id
t2_discont <- t2_discont[which(!(t2_discont %in% t1_discont))]
```

```{r echo=FALSE,include=FALSE,cache=TRUE}
# Format survey data
# Child
child_surveys_b <- child_full %>%
  select(record_id,timepoint_baselinesurvey:c_paid20)
colnames(child_surveys_b)[2] <- "Timepoint"
child_surveys_1 <- child_full %>%
  select(record_id,timepoint_survey1:c_paid20_t1)
colnames(child_surveys_1) <- sub("_t1","",colnames(child_surveys_1))
colnames(child_surveys_1)[2] <- "Timepoint"
child_surveys_2 <- child_full %>%
  select(record_id,timepoint_survey2:c_paid20_t2)
colnames(child_surveys_2) <- sub("_t2","",colnames(child_surveys_2))
colnames(child_surveys_2)[2] <- "Timepoint"
child_surveys_3 <- child_full %>%
  select(record_id,timepoint_survey3:c_paid20_t3)
colnames(child_surveys_3) <- sub("_t3","",colnames(child_surveys_3))
colnames(child_surveys_3)[2] <- "Timepoint"
child_surveys_4 <- child_full %>%
  select(record_id,timepoint_survey4:c_paid20_t4)
colnames(child_surveys_4) <- sub("_t4","",colnames(child_surveys_4))
colnames(child_surveys_4)[2] <- "Timepoint"
child_surveys <- bind_rows(child_surveys_b,child_surveys_1,child_surveys_2,
                           child_surveys_3,child_surveys_4)
child_surveys <- child_surveys %>%
  arrange(as.numeric(record_id),Timepoint)
colnames(child_surveys) <- sub("c_","",colnames(child_surveys))
colnames(child_surveys) <- sub("hfs_","",colnames(child_surveys))
# YA
ya_surveys_b <- ya_full %>%
  select(record_id,timepoint_baselinesurvey,ya_hfs_behave1_b:ya_hfs_worry18_b,
         ya_paid1_base:ya_paid20_base)
ya_surveys_b[,3:ncol(ya_surveys_b)] <- lapply(ya_surveys_b[,3:ncol(ya_surveys_b)],as.numeric)
colnames(ya_surveys_b)[2] <- "Timepoint"
colnames(ya_surveys_b)[3:ncol(ya_surveys_b)] <- 
  substr(colnames(ya_surveys_b)[3:ncol(ya_surveys_b)],1,
         nchar(colnames(ya_surveys_b)[3:ncol(ya_surveys_b)])-2)
colnames(ya_surveys_b) <- sub("_ba","",colnames(ya_surveys_b))
ya_surveys_1 <- ya_full %>%
  select(record_id,timepoint_survey1,ya_hfs_behave1_t1:ya_hfs_worry18_t1,
         ya_paid1_t1:ya_paid20_t1)
ya_surveys_1[,3:ncol(ya_surveys_1)] <- lapply(ya_surveys_1[,3:ncol(ya_surveys_1)],as.numeric)
colnames(ya_surveys_1) <- sub("_t1","",colnames(ya_surveys_1))
colnames(ya_surveys_1)[which(colnames(ya_surveys_1) == "ya_hfs_worry8_b")] <- "ya_hfs_worry8"
colnames(ya_surveys_1)[2] <- "Timepoint"
ya_surveys_2 <- ya_full %>%
  select(record_id,timepoint_survey2,ya_hfs_behave1_t2:ya_hfs_worry18_t2,
         ya_paid1_t2:ya_paid20_t2)
ya_surveys_2[,3:ncol(ya_surveys_2)] <- lapply(ya_surveys_2[,3:ncol(ya_surveys_2)],as.numeric)
colnames(ya_surveys_2) <- sub("_t2","",colnames(ya_surveys_2))
colnames(ya_surveys_2) <- sub("_b_t1","",colnames(ya_surveys_2))
colnames(ya_surveys_2)[2] <- "Timepoint"
ya_surveys_3 <- ya_full %>%
  select(record_id,timepoint_survey3,ya_hfs_behave1_t3:ya_hfs_worry18_t3,
         ya_paid1_t3:ya_paid20_t3)
ya_surveys_3[,3:ncol(ya_surveys_3)] <- lapply(ya_surveys_3[,3:ncol(ya_surveys_3)],as.numeric)
colnames(ya_surveys_3) <- sub("_t3","",colnames(ya_surveys_3))
colnames(ya_surveys_3) <- sub("_3","",colnames(ya_surveys_3))
colnames(ya_surveys_3)[2] <- "Timepoint"
ya_surveys_4 <- ya_full %>%
  select(record_id,timepoint_survey4,ya_hfs_behave1_t4:ya_hfs_worry18_t4,
         ya_paid1_t4:ya_paid20_t4)
ya_surveys_4[,3:ncol(ya_surveys_4)] <- lapply(ya_surveys_4[,3:ncol(ya_surveys_4)],as.numeric)
colnames(ya_surveys_4) <- sub("_t4","",colnames(ya_surveys_4))
colnames(ya_surveys_4)[2] <- "Timepoint"
ya_surveys <- bind_rows(ya_surveys_b,ya_surveys_1,ya_surveys_2,
                           ya_surveys_3,ya_surveys_4)
colnames(ya_surveys) <- sub("ya_","",colnames(ya_surveys))
colnames(ya_surveys) <- sub("hfs_","",colnames(ya_surveys))
ya_surveys <- ya_surveys %>%
  arrange(as.numeric(record_id),Timepoint)
# Parents
parent_surveys_b <- parent_full %>%
  select(record_id,timepoint_baselinesurvey:p_hfs_worry26,p_paid1:p_paid18)
colnames(parent_surveys_b)[2] <- "Timepoint"
parent_surveys_1 <- parent_full %>%
  select(record_id,timepoint_survey1:p_hfs_worry26_t1,p_paid1_t1:p_paid18_t1)
colnames(parent_surveys_1) <- sub("_t1","",colnames(parent_surveys_1))
colnames(parent_surveys_1)[2] <- "Timepoint"
parent_surveys_2 <- parent_full %>%
  select(record_id,timepoint_survey2:p_hfs_worry26_t2,p_paid1_t2:p_paid18_t2)
colnames(parent_surveys_2) <- sub("_t2","",colnames(parent_surveys_2))
colnames(parent_surveys_2)[2] <- "Timepoint"
parent_surveys_3 <- parent_full %>%
  select(record_id,timepoint_survey3:p_hfs_worry26_t3,p_paid1_t3:p_paid18_t3)
colnames(parent_surveys_3) <- sub("_t3","",colnames(parent_surveys_3))
colnames(parent_surveys_3)[2] <- "Timepoint"
parent_surveys_4 <- parent_full %>%
  select(record_id,timepoint_survey4:p_hfs_worry26_t4,p_paid1_t4:p_paid18_t4)
colnames(parent_surveys_4) <- sub("_t4","",colnames(parent_surveys_4))
colnames(parent_surveys_4)[2] <- "Timepoint"
parent_surveys <- bind_rows(parent_surveys_b,parent_surveys_1,parent_surveys_2,
                           parent_surveys_3,parent_surveys_4)
colnames(parent_surveys) <- sub("p_","",colnames(parent_surveys))
colnames(parent_surveys) <- sub("hfs_","",colnames(parent_surveys))
parent_surveys <- parent_surveys %>%
  filter(!(record_id %in% c("1a","74a"))) %>% # exclude 1a and 74a per Cari
  arrange(as.numeric(record_id),Timepoint)
```

```{r eval=FALSE,include=FALSE}
# Find those with an M1 but no T1
# Per Cari, we should use M1 surveys as T1 for those with T2 and M1 but no T2. This was fixed in REDCap. Record IDs emailed to Cari for her records. Each of these should have 0 rows now.
m1_not1_child <- child_surveys %>%
  group_by(record_id) %>%
  filter(any(Timepoint  == "month 1")) %>%
  summarise(Timepoints = paste(Timepoint, collapse = ", ")) %>%
  filter(Timepoints == "baseline, month 1, Time 2")
m1_not1_ya <- ya_surveys %>%
  group_by(record_id) %>%
  filter(any(Timepoint  == "Month 1")) %>%
  summarise(Timepoints = paste(Timepoint, collapse = ", ")) %>%
  filter(Timepoints == "baseline, Month 1, Time 2")
m1_not1_parent <- parent_surveys %>%
  group_by(record_id) %>%
  filter(any(Timepoint  == "month 1")) %>%
  summarise(Timepoints = paste(Timepoint, collapse = ", ")) %>%
  filter(Timepoints == "baseline, month 1, Time 2")
```

```{r echo=FALSE,cache=TRUE,include=FALSE}
# Survey scoring
# Subtract 1 to code from 0-4
child_surveys[,3:ncol(child_surveys)] <- lapply(child_surveys[,3:ncol(child_surveys)],function(x){as.numeric(x)-1})
ya_surveys[,3:ncol(ya_surveys)] <- lapply(ya_surveys[,3:ncol(ya_surveys)],function(x){as.numeric(x)-1})
parent_surveys[,3:ncol(parent_surveys)] <- lapply(parent_surveys[,3:ncol(parent_surveys)],function(x){as.numeric(x)-1})
# Split into PAID and HFS subscales
# Child
child_surveys_paid <- child_surveys %>%
  select(record_id,Timepoint,paid1:paid20)
child_surveys_behave <- child_surveys %>%
  select(record_id,Timepoint,behave1:behave10)
child_surveys_worry <- child_surveys %>%
  select(record_id,Timepoint,worry11:worry25)
# YA
ya_surveys_paid <- ya_surveys %>%
  select(record_id,Timepoint,paid1:paid20)
ya_surveys_behave <- ya_surveys %>%
  select(record_id,Timepoint,behave1:behave15)
ya_surveys_worry <- ya_surveys %>%
  select(record_id,Timepoint,worry1:worry18)
# Parent
parent_surveys_paid <- parent_surveys %>%
  select(record_id,Timepoint,paid1:paid18)
parent_surveys_behave <- parent_surveys %>%
  select(record_id,Timepoint,behave1:behave11)
parent_surveys_worry <- parent_surveys %>%
  select(record_id,Timepoint,worry12:worry26)
# Replace missing data with means if > 75% answered
# Children
# PAID
missing_threshold <- ceiling(length(3:ncol(child_surveys_paid)) * 0.25)
for (r in 1:nrow(child_surveys_paid)) { # Couldn't get apply to work for some reason
  x <- child_surveys_paid[r,3:ncol(child_surveys_paid)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  child_surveys_paid[r,3:ncol(child_surveys_paid)] <- x
}
# Behave
missing_threshold <- ceiling(length(3:ncol(child_surveys_behave)) * 0.25)
for (r in 1:nrow(child_surveys_behave)) { # Couldn't get apply to work for some reason
  x <- child_surveys_behave[r,3:ncol(child_surveys_behave)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  child_surveys_behave[r,3:ncol(child_surveys_behave)] <- x
}
# Worry
missing_threshold <- ceiling(length(3:ncol(child_surveys_worry)) * 0.25)
for (r in 1:nrow(child_surveys_worry)) { # Couldn't get apply to work for some reason
  x <- child_surveys_worry[r,3:ncol(child_surveys_worry)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  child_surveys_worry[r,3:ncol(child_surveys_worry)] <- x
}
# YA
#PAID
missing_threshold <- ceiling(length(3:ncol(ya_surveys_paid)) * 0.25)
for (r in 1:nrow(ya_surveys_paid)) { # Couldn't get apply to work for some reason
  x <- ya_surveys_paid[r,3:ncol(ya_surveys_paid)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  ya_surveys_paid[r,3:ncol(ya_surveys_paid)] <- x
}
# Behave
missing_threshold <- ceiling(length(3:ncol(ya_surveys_behave)) * 0.25)
for (r in 1:nrow(ya_surveys_behave)) { # Couldn't get apply to work for some reason
  x <- ya_surveys_behave[r,3:ncol(ya_surveys_behave)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  ya_surveys_behave[r,3:ncol(ya_surveys_behave)] <- x
}
# Worry
missing_threshold <- ceiling(length(3:ncol(ya_surveys_worry)) * 0.25)
for (r in 1:nrow(ya_surveys_worry)) { # Couldn't get apply to work for some reason
  x <- ya_surveys_worry[r,3:ncol(ya_surveys_worry)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  ya_surveys_worry[r,3:ncol(ya_surveys_worry)] <- x
}
# Parents
# PAID
missing_threshold <- ceiling(length(3:ncol(parent_surveys_paid)) * 0.25)
for (r in 1:nrow(parent_surveys_paid)) { # Couldn't get apply to work for some reason
  x <- parent_surveys_paid[r,3:ncol(parent_surveys_paid)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  parent_surveys_paid[r,3:ncol(parent_surveys_paid)] <- x
}
# Behave
missing_threshold <- ceiling(length(3:ncol(parent_surveys_behave)) * 0.25)
for (r in 1:nrow(parent_surveys_behave)) { # Couldn't get apply to work for some reason
  x <- parent_surveys_behave[r,3:ncol(parent_surveys_behave)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  parent_surveys_behave[r,3:ncol(parent_surveys_behave)] <- x
}
# Worry
missing_threshold <- ceiling(length(3:ncol(parent_surveys_worry)) * 0.25)
for (r in 1:nrow(parent_surveys_worry)) { # Couldn't get apply to work for some reason
  x <- parent_surveys_worry[r,3:ncol(parent_surveys_worry)]
  replacement <- mean(as.numeric(x),na.rm = T)
  num_missing <- length(which(is.na(x)))
  if (num_missing<=missing_threshold && num_missing>0) {
    x[which(is.na(x))] <- replacement
    } else if (num_missing > missing_threshold) {
      x <- rep(NA,length(x))
    }
  parent_surveys_worry[r,3:ncol(parent_surveys_worry)] <- x
}
# Score children
# PAID
child_surveys_paid$paid_score <- 
  apply(child_surveys_paid[3:ncol(child_surveys_paid)],1,
        function(x) (mean(4-x))*25)
# Behave
child_surveys_behave$behave_score <- 
  apply(child_surveys_behave[3:ncol(child_surveys_behave)],1,sum)
# Worry
child_surveys_worry$worry_score <- 
  apply(child_surveys_worry[3:ncol(child_surveys_worry)],1,sum)

# Score YA
# PAID
ya_surveys_paid$paid_score <- 
  apply(ya_surveys_paid[3:ncol(ya_surveys_paid)],1,
        function(x) (sum(x))*1.25)
# Behave
ya_surveys_behave$behave_score <- 
  apply(ya_surveys_behave[3:ncol(ya_surveys_behave)],1,sum)
# Worry
ya_surveys_worry$worry_score <- 
  apply(ya_surveys_worry[3:ncol(ya_surveys_worry)],1,sum)

# Score parents
# PAID
parent_surveys_paid$paid_score <- 
  apply(parent_surveys_paid[3:ncol(parent_surveys_paid)],1,
        function(x) (mean(4-x))*25)
# Behave
parent_surveys_behave$behave_score <- 
  apply(parent_surveys_behave[3:ncol(parent_surveys_behave)],1,sum)
# Worry
parent_surveys_worry$worry_score <- 
  apply(parent_surveys_worry[3:ncol(parent_surveys_worry)],1,sum)
```

# Methods

Linear mixed effects models were used to examine change in HbA1c over time. Three models were compared using Aikaike’s Information Criterion to select the best model:

1.	A random intercept model with visit number (baseline, T1, T2) as a categorical time variable.  No adjustment for baseline.
2.	A random intercept model with visit number as a continuous time variable, without adjustment for baseline.
3.	A random intercept and random slope model, without adjustment for baseline.

The random intercept model with time treated as a categorical variable was the best model. Originally the models were adjusted for baseline value (as we did for the abstract), but it was decided that this adjustment does not make sense given the question this study wants to answer, in addition to making the model results difficult to interpret.

The HbA1c model included an interaction term for baseline HbA1c group with timepoint, but the CGM variable models did not. We examined the interaction effect of age and timepoint on HbA1c, but it was not significant. Differences between timepoints for each baseline HbA1c group were compared using linear contrasts.

Discontinuation was defined as AM time < 10 at either T1 or T2. This analysis excluded those never trained on automode, and all timepoints with < 10 AM time.

# Results

```{r echo=FALSE,include=FALSE}
# Table 1
dem_vars <- c("hba1c_baseline","demographics_t1d_duration","demographics_age",
              "age_group","hba1c_clinical","demographics_ethnicity","demographics_race",
              "demographics_sex","demographics_insurance","demographics_pumphx",
              "demographics_cgmhx")
# Get days from AM
days <- glycemic_data %>%
  select(record_id,Timepoint,days)
days <- spread(days,Timepoint,days)
colnames(days) <- c("record_id","baseline_a1c_to_am","am_to_m1","am_to_t1","am_to_t2")
days$record_id <- as.character(days$record_id)
# Add to demographics
demographics <- left_join(demographics,days,by = "record_id")
demographics <- left_join(demographics,child_full[,c("record_id","hba1c_date_b","hba1c_baseline")])
demographics$hba1c_clinical <- cut(demographics$hba1c_baseline,breaks = c(0,7.5,9.0,Inf),
                              labels = c("Low Baseline HbA1c","Medium Baseline HbA1c","High Baseline HbA1c"))
demographics$hba1c_date_b <- lubridate::mdy(demographics$hba1c_date_b)
demographics$baseline_a1c_to_am <- as.numeric(difftime(demographics$automode_start,demographics$hba1c_date_b,units = "days"))
# Table 1
demographics$continuation <- ifelse(demographics$record_id %in% t1_discont,"Discontinued T1",
                               ifelse(demographics$record_id %in% t2_discont,"Discontinued T2",
                                      ifelse(demographics$record_id %in% never_trained,"Never Trained","Continued")))
demographics$trained <- ifelse(demographics$continuation == "Never Trained","Never Trained","Trained")
# Format
demographics[c("demographics_ethnicity","demographics_race","demographics_sex",
               "demographics_insurance","demographics_pumphx",
               "demographics_cgmhx")] <- lapply(demographics[c("demographics_ethnicity","demographics_race","demographics_sex",
               "demographics_insurance","demographics_pumphx",
               "demographics_cgmhx")], as.factor)
# Variables
dem_vars <- c("baseline_a1c_to_am","am_to_m1","am_to_t1","am_to_t2",dem_vars)
# Normality check
nonnormal <- norm.check(demographics,c("hba1c_baseline","baseline_a1c_to_am",
                                       "am_to_m1","am_to_t1","am_to_t2",
                                       "demographics_age","demographics_t1d_duration"))
# Trained vs. never
t1a <- CreateTableOne(dem_vars,data = demographics,strata = "trained")
t1a <- print(t1a,nonnormal = nonnormal,printToggle = F)
# Continuers
t1b <- CreateTableOne(dem_vars,data = demographics,strata = "continuation")
t1b <- print(t1b,nonnormal = nonnormal,printToggle = F)
```

### Table 1a: Descriptive statistics by AM training status

```{r echo=FALSE}
kable(t1a)
```

### Table 1b: Descriptive statistics by continuation status

```{r echo=FALSE}
kable(t1b)
```

```{r echo=FALSE,include=FALSE}
# 6 month glycemic analysis
# If participant discontinued at T1, remove T1 and T2, if dicontinued at T2, remove just T2. Also remove those never trained.
glycemic_data_6_month <- glycemic_data %>% 
  filter(!(record_id %in% never_trained)) %>%
  filter(!(record_id %in% t1_discont & (Timepoint == "T1" | Timepoint == "T2"))) %>%
  filter(!(record_id %in% t2_discont & Timepoint == "T2"))
glycemic_data_6_month$record_id <- as.character(glycemic_data_6_month$record_id)
glycemic_data_6_month <- left_join(glycemic_data_6_month,demographics[,c("record_id","age_group")])
# Remove M1, group by clinical A1c cutoffs at baseline
data_no_m1 <- glycemic_data_6_month %>%
  filter(Timepoint != "M1")
# Write file for SAS contrasts
write.csv(data_no_m1,file = "/Users/timvigers/Desktop/cari.csv",
          row.names = F,na="")
# Remove B (e.g. for % time AM models etc.)
data_no_b <- glycemic_data_6_month %>%
  filter(Timepoint != "B")
# A1c plot
a1c_overall <- 
  ggplot(data_no_m1,aes_string(x = "Timepoint",y = "hba1c")) + 
  geom_point(size = 0.2) +
  stat_summary(fun.y=mean, geom="point") +
  stat_summary(fun.y=mean, geom="line", aes(group = 1)) +
  stat_summary(aes(label=round(..y..,3)), fun.y=mean, geom="text",hjust=1.2,vjust = 1.2)+
  xlab("Timepoint") + 
  ylab("HbA1c (%)") +
  theme(legend.title=element_blank())
```

### Figure 1: Mean HbA1c by Timepoint

```{r echo=FALSE,warning=FALSE,dpi=600}
a1c_overall
```

### Table 2: A1c Mixed Models by Baseline HbA1c Group

```{r include=FALSE}
# A1c mixed models
a1c_age_mod <- lme(hba1c ~ Timepoint*age_group,random=~1|record_id,data = data_no_m1,na.action = na.omit)
a1c_glyc_control_mod <- lme(hba1c ~ Timepoint*hba1c_clinical,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# Means models
a1c_age_mod_means <- lme(hba1c ~ Timepoint:age_group-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
a1c_glyc_control_mod_means <- lme(hba1c ~ Timepoint:hba1c_clinical-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# Contrasts
# cmat.emmc <- rbind(c(0, 0, 1, 0, 0, 0, 0, 0, -1),
#               c(1, 0, 0, 0, 0, 0, -1, 0, 0),
#               c(0, 1, 0, 0, 0, 0, 0, -1, 0),
#               c(0, 0, 0, 0, 0, 1, 0, 0, -1),
#               c(0, 0, 0, 1, 0, 0, -1, 0, 0),
#               c(0, 0, 0, 0, 1, 0, 0, -1, 0))
# a1c_glyc_emm <- emmeans(a1c_glyc_control_mod,~Timepoint:hba1c_clinical)
# contrast(a1c_glyc_emm,"cmat")
```

```{r echo=FALSE}
# Import results from contrasts in SAS.
estimates <- read_excel(paste0(pathstart,"som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Cari Berget/6 Month Manuscript/Reports/estimates.xlsx"))
means <- read_excel(paste0(pathstart,"som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Cari Berget/6 Month Manuscript/Reports/means.xlsx"))
```

```{r echo=FALSE}
kable(means[,c("Timepoint","hba1c_clinical","Estimate","StdErr","Probt")],caption = "Group Means by Time Point")
kable(anova.lme(a1c_glyc_control_mod),caption = "Type 3 Tests of Fixed Effects")
kable(estimates[,c("Label","Estimate","StdErr","Probt")],caption = "Timepoint Differences by HbA1c Group")
```

### HbA1c mixed model interpretation
The "Group Means by Timepoint" table shows the average HbA1c at each timepoint, for each HbA1c group. The p values in this table ("Probt") refer to whether the mean HbA1c is close to 0, so they can be ignored for now. 

The type 3 tests of fixed effects show that timepoint, HbA1c group, and the interaction between timepoint and HbA1c group were each significant overall.

"Timepoint Differences by HbA1c Group" shows whether time 1 or time 2 were significantly different from baseline, by HbA1c group (Low = (0,7.5], Medium = (7.5,9], High = (9,Inf]). So on average, the low A1c group increased by 0.098 from baseline to visit 1, but it was not statistically significant. On average the high A1c group decreased by 0.994 from baseline to visit 1, which was statistically significant (p < 0.0001).

```{r echo=FALSE,include=FALSE,cache=TRUE}
# Per Cari Berget 5/1/19:
# "Subjects 4,8,16,17,21 completed the child survey versions in the baseline survey form and adult surveys for the rest of the time points.  The young adult surveys are the correct ones"
# These are all included in the defined YA group anyway.

# 6 month survey analysis
child_surveys_6_month <- child_surveys_paid %>%
  select(record_id,Timepoint,paid_score) %>%
  left_join(child_surveys_behave[,c("record_id","Timepoint","behave_score")],
            by=c("record_id", "Timepoint")) %>%
  left_join(child_surveys_worry[,c("record_id","Timepoint","worry_score")],
            by=c("record_id", "Timepoint")) %>%
  mutate(total_hfs_score = worry_score + behave_score) %>%
  filter(Timepoint %in% c(0,2,3))
child_surveys_6_month$Timepoint <- as.factor(child_surveys_6_month$Timepoint)
levels(child_surveys_6_month$Timepoint) <- list("B"=0,"T1"=2,"T2"=3)
child_surveys_6_month$Group <- "Child"
# Z scores
child_surveys_6_month$worry_score_z <- (child_surveys_6_month$worry_score - mean(child_surveys_6_month$worry_score, na.rm = T)) / sd(child_surveys_6_month$worry_score, na.rm = T)
child_surveys_6_month$behave_score_z <- (child_surveys_6_month$behave_score - mean(child_surveys_6_month$behave_score, na.rm = T)) / sd(child_surveys_6_month$behave_score, na.rm = T)

ya_surveys_6_month <- ya_surveys_paid %>%
  select(record_id,Timepoint,paid_score) %>%
  left_join(ya_surveys_behave[,c("record_id","Timepoint","behave_score")],
            by=c("record_id", "Timepoint")) %>%
  left_join(ya_surveys_worry[,c("record_id","Timepoint","worry_score")],
            by=c("record_id", "Timepoint")) %>%
  mutate(total_hfs_score = worry_score + behave_score) %>%
  filter(Timepoint %in% c("baseline","Time 1","Time 2"))
ya_surveys_6_month$Timepoint <- ifelse(ya_surveys_6_month$Timepoint == "baseline",0,
                                       ifelse(ya_surveys_6_month$Timepoint == "Time 1",2,3))
ya_surveys_6_month$Timepoint <- as.factor(ya_surveys_6_month$Timepoint)
levels(ya_surveys_6_month$Timepoint) <- list("B"=0,"T1"=2,"T2"=3)
ya_surveys_6_month$Group <- "YA"
# Z scores
ya_surveys_6_month$worry_score_z <- (ya_surveys_6_month$worry_score - mean(ya_surveys_6_month$worry_score, na.rm = T)) / sd(ya_surveys_6_month$worry_score, na.rm = T)
ya_surveys_6_month$behave_score_z <- (ya_surveys_6_month$behave_score - mean(ya_surveys_6_month$behave_score, na.rm = T)) / sd(ya_surveys_6_month$behave_score, na.rm = T)

parent_surveys_6_month <- parent_surveys_paid %>%
  select(record_id,Timepoint,paid_score) %>%
  left_join(parent_surveys_behave[,c("record_id","Timepoint","behave_score")],
            by=c("record_id", "Timepoint")) %>%
  left_join(parent_surveys_worry[,c("record_id","Timepoint","worry_score")],
            by=c("record_id", "Timepoint")) %>%
  mutate(total_hfs_score = worry_score + behave_score) %>%
  filter(Timepoint %in% c("baseline","Time 1","Time 2"))
parent_surveys_6_month$Timepoint <- as.numeric(parent_surveys_6_month$Timepoint) -1
parent_surveys_6_month$Group <- "Parents"
parent_surveys_6_month$Timepoint <- as.factor(parent_surveys_6_month$Timepoint)
levels(parent_surveys_6_month$Timepoint) <- list("B"=0,"T1"=2,"T2"=3)
parent_surveys_6_month$record_id <- paste0(parent_surveys_6_month$record_id,"p")
# Z scores
parent_surveys_6_month$worry_score_z <- (parent_surveys_6_month$worry_score - mean(parent_surveys_6_month$worry_score, na.rm = T)) / sd(parent_surveys_6_month$worry_score, na.rm = T)
parent_surveys_6_month$behave_score_z <- (parent_surveys_6_month$behave_score - mean(parent_surveys_6_month$behave_score, na.rm = T)) / sd(parent_surveys_6_month$behave_score, na.rm = T)
# Combine all survey data
all_surveys_6_month <- bind_rows(child_surveys_6_month,ya_surveys_6_month,parent_surveys_6_month)
```

```{r echo=FALSE}
# CGM data models
# TIR
tir_u54_mod <- lme(sensor_u54 ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
tir_u54_mod_means <- lme(sensor_u54 ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)

tir_55_69_mod <- lme(sensor_55_69 ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
tir_55_69_mod_means <- lme(sensor_55_69 ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)

tir_70_180_mod <- lme(sensor_70_180 ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
tir_70_180_mod_means <- lme(sensor_70_180 ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)

tir_181_250_mod <- lme(sensor_181_250 ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
tir_181_250_mod_means <- lme(sensor_181_250 ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)

tir_g250_mod <- lme(sensor_g250 ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
tir_g250_mod_means <- lme(sensor_g250 ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# AM time
am_mod <- lme(am_time ~ Timepoint,random=~1|record_id,data = data_no_b,na.action = na.omit)
am_mod_means <- lme(am_time ~ Timepoint-1,random=~1|record_id,data = data_no_b,na.action = na.omit)
# Sensor wear
sensor_wear_mod <- lme(sensor_wear ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
sensor_wear_mod_means <- lme(sensor_wear ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# AM exits
amexit_day_mod <- lme(amexit_day ~ Timepoint,random=~1|record_id,data = data_no_b,na.action = na.omit)
amexit_day_mod_means <- lme(amexit_day ~ Timepoint-1,random=~1|record_id,data = data_no_b,na.action = na.omit)

# BG Checks
bg_checks_mod <- lme(bg_checks ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
bg_checks_mod_means <- lme(bg_checks ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# Calibrations
calibrations_mod <- lme(calibrations ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
calibrations_mod_means <- lme(calibrations ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# TDD
tdd_mod <- lme(tdd ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
tdd_mod_means <- lme(tdd ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# Basal
basal_mod <- lme(basal ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
basal_mod_means <- lme(basal ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# Bolus
bolus_mod <- lme(bolus ~ Timepoint,random=~1|record_id,data = data_no_m1,na.action = na.omit)
bolus_mod_means <- lme(bolus ~ Timepoint-1,random=~1|record_id,data = data_no_m1,na.action = na.omit)
# Survey scores
paid_mod <- lme(paid_score ~ Timepoint,random=~1|record_id,data = all_surveys_6_month,na.action = na.omit)
paid_mod_means <- lme(paid_score ~ Timepoint-1,random=~1|record_id,data = all_surveys_6_month,na.action = na.omit) # Slopes for family?
behave_mod <- lme(behave_score_z ~ Timepoint,random=~1|record_id,data = all_surveys_6_month,na.action = na.omit)
behave_mod_means <- lme(behave_score_z ~ Timepoint-1,random=~1|record_id,data = all_surveys_6_month,na.action = na.omit)
worry_mod <- lme(worry_score_z ~ Timepoint,random=~1|record_id,data = all_surveys_6_month,na.action = na.omit)
worry_mod_means <- lme(worry_score_z ~ Timepoint-1,random=~1|record_id,data = all_surveys_6_month,na.action = na.omit)
```

## Table 3: CGM Variable Mixed Models Summary

```{r echo=FALSE}
# Make table
mod_summary <- as.data.frame(matrix(nrow = 16,ncol = 6))
colnames(mod_summary) <- c("Baseline","Month 1","Time 1","T1 P value","Time 2","T2 P value")
rownames(mod_summary) <- c("TIR Under 54","TIR 55-69","TIR 70-180",
                           "TIR 181-250","TIR >250","AM Time","Sensor Wear",
                           "AM Exits Per Day","BG Checks","Calibrations","TDD",
                           "Basal","Bolus","PAID","HFS Behave Z Score","HFS Worry Z Score")
# TIR
tir_u54 <- paste0(round(summary(tir_u54_mod_means)$tTable[,1],3)," (",
              round(summary(tir_u54_mod_means)$tTable[,2],3),")")
mod_summary["TIR Under 54","Baseline"] <- tir_u54[1]
mod_summary["TIR Under 54","Time 1"] <- tir_u54[2]
mod_summary["TIR Under 54","T1 P value"] <- summary(tir_u54_mod)$tTable[2,5]
mod_summary["TIR Under 54","Time 2"] <- tir_u54[3]
mod_summary["TIR Under 54","T2 P value"] <- summary(tir_u54_mod)$tTable[3,5]

tir_55_69 <- paste0(round(summary(tir_55_69_mod_means)$tTable[,1],3)," (",
              round(summary(tir_55_69_mod_means)$tTable[,2],3),")")
mod_summary["TIR 55-69","Baseline"] <- tir_55_69[1]
mod_summary["TIR 55-69","Time 1"] <- tir_55_69[2]
mod_summary["TIR 55-69","T1 P value"] <- summary(tir_55_69_mod)$tTable[2,5]
mod_summary["TIR 55-69","Time 2"] <- tir_55_69[3]
mod_summary["TIR 55-69","T2 P value"] <- summary(tir_55_69_mod)$tTable[3,5]

tir_70_180 <- paste0(round(summary(tir_70_180_mod_means)$tTable[,1],3)," (",
              round(summary(tir_70_180_mod_means)$tTable[,2],3),")")
mod_summary["TIR 70-180","Baseline"] <- tir_70_180[1]
mod_summary["TIR 70-180","Time 1"] <- tir_70_180[2]
mod_summary["TIR 70-180","T1 P value"] <- summary(tir_70_180_mod)$tTable[2,5]
mod_summary["TIR 70-180","Time 2"] <- tir_70_180[3]
mod_summary["TIR 70-180","T2 P value"] <- summary(tir_70_180_mod)$tTable[3,5]

tir_181_250 <- paste0(round(summary(tir_181_250_mod_means)$tTable[,1],3)," (",
              round(summary(tir_181_250_mod_means)$tTable[,2],3),")")
mod_summary["TIR 181-250","Baseline"] <- tir_181_250[1]
mod_summary["TIR 181-250","Time 1"] <- tir_181_250[2]
mod_summary["TIR 181-250","T1 P value"] <- summary(tir_181_250_mod)$tTable[2,5]
mod_summary["TIR 181-250","Time 2"] <- tir_181_250[3]
mod_summary["TIR 181-250","T2 P value"] <- summary(tir_181_250_mod)$tTable[3,5]

tir_g250 <- paste0(round(summary(tir_g250_mod_means)$tTable[,1],3)," (",
              round(summary(tir_g250_mod_means)$tTable[,2],3),")")
mod_summary["TIR >250","Baseline"] <- tir_g250[1]
mod_summary["TIR >250","Time 1"] <- tir_g250[2]
mod_summary["TIR >250","T1 P value"] <- summary(tir_g250_mod)$tTable[2,5]
mod_summary["TIR >250","Time 2"] <- tir_g250[3]
mod_summary["TIR >250","T2 P value"] <- summary(tir_g250_mod)$tTable[3,5]

# AM time
am <- paste0(round(summary(am_mod_means)$tTable[,1],3)," (",
             round(summary(am_mod_means)$tTable[,2],3),")")
mod_summary["AM Time","Month 1"] <- am[1]
mod_summary["AM Time","Time 1"] <- am[2]
mod_summary["AM Time","T1 P value"] <- summary(am_mod)$tTable[2,5]
mod_summary["AM Time","Time 2"] <- am[3]
mod_summary["AM Time","T2 P value"] <- summary(am_mod)$tTable[3,5]

# Sensor
sensor_wear <- paste0(round(summary(sensor_wear_mod_means)$tTable[,1],3)," (",
                      round(summary(sensor_wear_mod_means)$tTable[,2],3),")")
mod_summary["Sensor Wear","Baseline"] <- sensor_wear[1]
mod_summary["Sensor Wear","Time 1"] <- sensor_wear[2]
mod_summary["Sensor Wear","T1 P value"] <- summary(sensor_wear_mod)$tTable[2,5]
mod_summary["Sensor Wear","Time 2"] <- sensor_wear[3]
mod_summary["Sensor Wear","T2 P value"] <- summary(sensor_wear_mod)$tTable[3,5]

# AM Exits Per Day
amexit_day <- paste0(round(summary(amexit_day_mod_means)$tTable[,1],3)," (",
                     round(summary(amexit_day_mod_means)$tTable[,2],3),")")
mod_summary["AM Exits Per Day","Month 1"] <- amexit_day[1]
mod_summary["AM Exits Per Day","Time 1"] <- amexit_day[2]
mod_summary["AM Exits Per Day","T1 P value"] <- summary(amexit_day_mod)$tTable[2,5]
mod_summary["AM Exits Per Day","Time 2"] <- amexit_day[3]
mod_summary["AM Exits Per Day","T2 P value"] <- summary(amexit_day_mod)$tTable[3,5]

# BG Checks
bg_checks <- paste0(round(summary(bg_checks_mod_means)$tTable[,1],3)," (",
              round(summary(bg_checks_mod_means)$tTable[,2],3),")")
mod_summary["BG Checks","Baseline"] <- bg_checks[1]
mod_summary["BG Checks","Time 1"] <- bg_checks[2]
mod_summary["BG Checks","T1 P value"] <- summary(bg_checks_mod)$tTable[2,5]
mod_summary["BG Checks","Time 2"] <- bg_checks[3]
mod_summary["BG Checks","T2 P value"] <- summary(bg_checks_mod)$tTable[3,5]
# Calibrations
calibrations <- paste0(round(summary(calibrations_mod_means)$tTable[,1],3)," (",
              round(summary(calibrations_mod_means)$tTable[,2],3),")")
mod_summary["Calibrations","Baseline"] <- calibrations[1]
mod_summary["Calibrations","Time 1"] <- calibrations[2]
mod_summary["Calibrations","T1 P value"] <- summary(calibrations_mod)$tTable[2,5]
mod_summary["Calibrations","Time 2"] <- calibrations[3]
mod_summary["Calibrations","T2 P value"] <- summary(calibrations_mod)$tTable[3,5]
# TDD
tdd <- paste0(round(summary(tdd_mod_means)$tTable[,1],3)," (",
              round(summary(tdd_mod_means)$tTable[,2],3),")")
mod_summary["TDD","Baseline"] <- tdd[1]
mod_summary["TDD","Time 1"] <- tdd[2]
mod_summary["TDD","T1 P value"] <- summary(tdd_mod)$tTable[2,5]
mod_summary["TDD","Time 2"] <- tdd[3]
mod_summary["TDD","T2 P value"] <- summary(tdd_mod)$tTable[3,5]
# Basal
basal <- paste0(round(summary(basal_mod_means)$tTable[,1],3)," (",
              round(summary(basal_mod_means)$tTable[,2],3),")")
mod_summary["Basal","Baseline"] <- basal[1]
mod_summary["Basal","Time 1"] <- basal[2]
mod_summary["Basal","T1 P value"] <- summary(basal_mod)$tTable[2,5]
mod_summary["Basal","Time 2"] <- basal[3]
mod_summary["Basal","T2 P value"] <- summary(basal_mod)$tTable[3,5]
# Bolus
bolus <- paste0(round(summary(bolus_mod_means)$tTable[,1],3)," (",
              round(summary(bolus_mod_means)$tTable[,2],3),")")
mod_summary["Bolus","Baseline"] <- bolus[1]
mod_summary["Bolus","Time 1"] <- bolus[2]
mod_summary["Bolus","T1 P value"] <- summary(bolus_mod)$tTable[2,5]
mod_summary["Bolus","Time 2"] <- bolus[3]
mod_summary["Bolus","T2 P value"] <- summary(bolus_mod)$tTable[3,5]
# PAID
paid <- paste0(round(summary(paid_mod_means)$tTable[,1],3)," (",
              round(summary(paid_mod_means)$tTable[,2],3),")")
mod_summary["PAID","Baseline"] <- paid[1]
mod_summary["PAID","Time 1"] <- paid[2]
mod_summary["PAID","T1 P value"] <- summary(paid_mod)$tTable[2,5]
mod_summary["PAID","Time 2"] <- paid[3]
mod_summary["PAID","T2 P value"] <- summary(paid_mod)$tTable[3,5]
# HFS Behave
behave <- paste0(round(summary(behave_mod_means)$tTable[,1],3)," (",
              round(summary(behave_mod_means)$tTable[,2],3),")")
mod_summary["HFS Behave Z Score","Baseline"] <- behave[1]
mod_summary["HFS Behave Z Score","Time 1"] <- behave[2]
mod_summary["HFS Behave Z Score","T1 P value"] <- summary(behave_mod)$tTable[2,5]
mod_summary["HFS Behave Z Score","Time 2"] <- behave[3]
mod_summary["HFS Behave Z Score","T2 P value"] <- summary(behave_mod)$tTable[3,5]
# PAID
worry <- paste0(round(summary(worry_mod_means)$tTable[,1],3)," (",
              round(summary(worry_mod_means)$tTable[,2],3),")")
mod_summary["HFS Worry Z Score","Baseline"] <- worry[1]
mod_summary["HFS Worry Z Score","Time 1"] <- worry[2]
mod_summary["HFS Worry Z Score","T1 P value"] <- summary(worry_mod)$tTable[2,5]
mod_summary["HFS Worry Z Score","Time 2"] <- worry[3]
mod_summary["HFS Worry Z Score","T2 P value"] <- summary(worry_mod)$tTable[3,5]
# Adjust p values, format
mod_summary$`T1 P value` <- format.pval(p.adjust(mod_summary$`T1 P value`,"fdr"),
                                        eps = 0.001,digits = 2)
mod_summary$`T2 P value` <- format.pval(p.adjust(mod_summary$`T2 P value`,"fdr"),
                                        eps = 0.001,digits = 2)
# Print
options(knitr.kable.NA = '')
kable(mod_summary)
```

Reported as mean (standard error). All continuers, no interaction by HbA1c group included in these models. P values adjusted for multiple comparisons using the false discovery rate (FDR) method. 



```{r echo=FALSE,warning=FALSE,dpi=600}
# Plot
# PAID
paid_plot_data <- rbind(child_surveys_6_month[,c("record_id","Timepoint","Group","paid_score")],
                        ya_surveys_6_month[,c("record_id","Timepoint","Group","paid_score")],
                        parent_surveys_6_month[,c("record_id","Timepoint","Group","paid_score")])

paid_plot <- ggplot(subset(paid_plot_data),aes(x=Timepoint,y=paid_score,group=record_id))+
  geom_point(size = 0.2) +
  stat_summary(fun.y=mean, geom="point",aes(group = Group,colour = Group)) +
  stat_summary(fun.y=mean, geom="line", aes(group = Group,colour = Group)) +
  xlab("Timepoint") + 
  ylab("PAID Score") +
  theme(legend.title=element_blank())
# Behave
behave_plot_data <- rbind(child_surveys_6_month[,c("record_id","Timepoint","Group","behave_score_z")],
                        ya_surveys_6_month[,c("record_id","Timepoint","Group","behave_score_z")],
                        parent_surveys_6_month[,c("record_id","Timepoint","Group","behave_score_z")])
behave_plot_data$Timepoint <- as.factor(behave_plot_data$Timepoint)
levels(behave_plot_data$Timepoint) <- list("B"=0,"T1"=2,"T2"=3)
behave_plot <- ggplot(subset(behave_plot_data),aes(x=Timepoint,y=behave_score_z,group=record_id))+
  geom_point(size = 0.2) +
  stat_summary(fun.y=mean, geom="point",aes(group = Group,colour = Group)) +
  stat_summary(fun.y=mean, geom="line", aes(group = Group,colour = Group)) +
  xlab("Timepoint") + 
  ylab("Behave Z Score") +
  theme(legend.title=element_blank())
# Worry
worry_plot_data <- rbind(child_surveys_6_month[,c("record_id","Timepoint","Group","worry_score_z")],
                        ya_surveys_6_month[,c("record_id","Timepoint","Group","worry_score_z")],
                        parent_surveys_6_month[,c("record_id","Timepoint","Group","worry_score_z")])
worry_plot_data$Timepoint <- as.factor(worry_plot_data$Timepoint)
levels(worry_plot_data$Timepoint) <- list("B"=0,"T1"=2,"T2"=3)
worry_plot <- ggplot(subset(worry_plot_data),aes(x=Timepoint,y=worry_score_z,group=record_id))+
  geom_point(size = 0.2) +
  stat_summary(fun.y=mean, geom="point",aes(group = Group,colour = Group)) +
  stat_summary(fun.y=mean, geom="line", aes(group = Group,colour = Group)) +
  xlab("Timepoint") + 
  ylab("Worry Z Score") +
  theme(legend.title=element_blank())
grid.arrange(paid_plot,behave_plot,worry_plot, ncol=2)
```