---
title: "Dexcom Report and CGM Metrics: Power Calculations"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_knit$set(root.dir = "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(knitr)
library(pwr)
```

# HEADER

The goal of this analysis is to assess the correlation between Dexcom Clarity report features (rebound hyperglycemia, sustained hyperglycemia, etc.) with International Consensus CGM metrics (time in range, SD, etc.).

Because these data have already been collected, we calculated the detectable linear correlation coefficient for a selection of sample sizes and levels of power ($\alpha = 0.05$ for all):

```{r}
# Define possible n values
ns = seq(100,300,by = 25)
# Make results data frame
power = data.frame(matrix(ncol = 5,nrow = length(ns)))
colnames(power) = c("n","R (80% Power)","R (85% Power)",
                    "R (90% Power)","R (95% Power)")
power$n = ns
# Fill in
power$`R (80% Power)` = sapply(ns,function(x){
  pwr.r.test(n = x,sig.level = 0.05,power = 0.8)$r
  })
power$`R (85% Power)` = sapply(ns,function(x){
  pwr.r.test(n = x,sig.level = 0.05,power = 0.85)$r
  })
power$`R (90% Power)` = sapply(ns,function(x){
  pwr.r.test(n = x,sig.level = 0.05,power = 0.9)$r
  })
power$`R (95% Power)` = sapply(ns,function(x){
  pwr.r.test(n = x,sig.level = 0.05,power = 0.95)$r
  })
# Print
kable(power,digits = 3)
```
