---
title: "Control IQ 12 Month Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
bibliography: "C:/Users/timvigers/Dropbox/Miscellaneous/references.bib"
csl: "C:/Users/timvigers/Dropbox/Miscellaneous/multidisciplinary-digital-publishing-institute.csl"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(lmerTest)
library(performance)
library(broom.mixed)
library(emmeans)
library(tidyverse)
library(knitr)
home_dir = 
  ifelse(.Platform$OS.type != "unix",
         "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Cari Berget/Control IQ 12 Month MS",
         "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Cari Berget/Control IQ 12 Month MS")
knitr::opts_knit$set(root.dir = home_dir)
knitr::opts_chunk$set(echo = FALSE,results='asis',fig.width=12,fig.height=10)
```

```{r data,include=FALSE}
# Import
source("C:/Users/timvigers/Documents/GitHub/BDC-Code/Cari Berget/Control IQ 12 month analysis/import_peds_data.r")
source("C:/Users/timvigers/Documents/GitHub/BDC-Code/Cari Berget/Control IQ 12 month analysis/import_parent_data.r")
# Survey scoring function
source("C:/Users/timvigers/Documents/GitHub/BDC-Code/Cari Berget/Scoring/scoring_function.R")
# Format date columns
date_cols = c("date","date_visit","demographics_diabetesdx","demographics_dob")
data[,date_cols] = lapply(data[,date_cols],lubridate::ymd)
# Fill down by participant, baseline HbA1c group
data = data %>% group_by(record_id) %>% 
  fill(demographics_diabetesdx,demographics_dob,demographics_sex.factor,demographics_age)
# Calculate variables
data = data %>% filter(!is.na(gyl_timepoint.factor),record_id >= 119) %>%
  mutate(age = as.numeric(difftime(date,demographics_dob,units = "days"))/365.25,
         sensor_over_180 = sensor_181_250 + sensor_g250,
         sensor_under_70 = sensor_54_69 + sensor_u54) %>%
  group_by(record_id) %>%
  mutate(hba1c_group = cut(hba1c[row_number() == 1],c(-Inf,7,9,Inf),
                           labels = c("<7%","7-9%",">=9%"),right = F),
         age_group = cut(demographics_age[row_number() == 1],c(-Inf,14,18,Inf),
                         labels = c("<14","14-17",">=18"),right = F))
# Score surveys
data = peds_scores(data)
parent_data = parent_scores(parent_data)
# Parent surveys to long
parent_data = parent_data %>% select(record_id,p_paid_b:p_tech_b) %>%
  pivot_longer(p_paid_b:p_tech_b,
               names_to = c("survey","gyl_timepoint.factor"),
               names_pattern = "p_(.*)_(.*)") %>%
  pivot_wider(names_from = "survey",names_prefix = "p_")
parent_data$gyl_timepoint.factor =
  factor(parent_data$gyl_timepoint.factor,
         levels = c("b","m1","t1","t2","t3","t4"),
         labels = c("Baseline","Month 1/ Training F/U","3 Months",
                    "6 Months","9 Months","12 Months"))
# Dicontinuers
discontinuers = unique(data$record_id[data$time_am < 10])
discontinuers = discontinuers[!is.na(discontinuers)]
# Numeric time
data$numeric_time = data$gyl_timepoint.factor
levels(data$numeric_time) = c(0,1,3,6,9,12)
data$numeric_time = as.numeric(as.character(data$numeric_time))
# Select relevant columns, exclude discontinuers
data = data %>% 
  filter(!record_id %in% discontinuers,!grepl("arm_3",redcap_event_name)) %>%
  select(record_id,gyl_timepoint.factor,numeric_time,age_group,hba1c_group,hba1c,
         demographics_sex.factor,time_am,sensor_wear,tdd,tdd_bolus,tdd_basal,
         dailymealbolus,sleep,exercise,sensor_mean,sensor_70_180,
         sensor_over_180,sensor_g250,sensor_under_70,sensor_u54,
         all_of(peds_surveys))
# Add parent scores
data$record_id = as.character(data$record_id)
parent_data$record_id = as.character(parent_data$record_id)
data = left_join(data,parent_data,by = c("record_id", "gyl_timepoint.factor"))
# Write cleaned CSV
write.csv(data,file = "./Data_Cleaned/ciq_analysis_dataset.csv",row.names = F,na = "")
# Clean workspace
rm(list = ls()[-which(ls() %in% c("data","discontinuers"))])
```

# Table 1: Participant Characteristics at Baseline

A total of `r length(discontinuers)` participants were excluded due to HCL use <10% at 1 or more visits. 

```{r table 1}
t1 = tableby(~age_group + hba1c + hba1c_group + demographics_sex.factor + 
               sensor_wear + tdd + tdd_bolus + tdd_basal + dailymealbolus + 
               sensor_mean + sensor_70_180 + sensor_over_180 + 
               sensor_g250 + sensor_under_70 + sensor_u54 + c_paid + c_hfs_maintain + 
               c_hfs_helpless + c_hfs_worry + c_inspire + ya_paid + ya_hfs_behave + 
               ya_hfs_worry + adult_inspire + c_tech + p_paid + p_hfs_maintain + 
               p_hfs_helpless + p_hfs_worry + p_inspire + p_tech,
             data = data[data$gyl_timepoint.factor == "Baseline",])
summary(t1,labelTranslations = 
          list(age_group = "Age Group",hba1c_group = "HbA1c Group",
               demographics_sex.factor = "Sex"))
```

# HbA1c 

## Plot

```{r message=FALSE}
hba1c_time_plot = ggplot(data[!is.na(data$hba1c) & !is.na(data$hba1c_group),],
                         aes(x = numeric_time,y = hba1c,group = record_id)) + 
  geom_point() + geom_line(aes(color = hba1c_group),alpha = 0.2) + 
  geom_smooth(aes(group = hba1c_group,color = hba1c_group),size = 2) + 
  theme_bw() + scale_x_continuous(breaks = c(0,1,3,6,9,12)) + 
  xlab("Months") + ylab("HbA1c") + labs(color='HbA1c Group') 
hba1c_time_plot
```

Bold lines indicate loess smoothed trajectories for each group, and gray shading indicates the standard error of these trajectories.

# Methods

## Model selection

Time was treated as numeric because timepoints are not evenly spaced, and the interpretation of interactions between a large number of categories is difficult. The final model structure was selected using mean sensor glucose as the outcome, and the same model structure was used for all following models for consistency. Although better by AIC, some random slope models did not converge, so a random intercept for participant was used in all models.

All analyses were performed using R version `r paste0(R.version$major,".",R.version$minor)` [@rcoreteamLanguageEnvironmentStatistical2021]. Linear mixed models were fit using the `lme4` [@batesFittingLinearMixedEffects2014] package and all model assumptions were assessed visually using the `performance` package[@ludeckePerformancePackageAssessment2021]. 

```{r model selection,include=FALSE}
# Remove outlier values
mod_data = data %>% filter(sensor_mean <= 400)
# Use numeric time since the M1 visit is not equally spaced. Also, interpreting the 
# interactions between categorical time and categorical HbA1c group is a nightmare.
ri_mod = lmer(sensor_mean ~ numeric_time*hba1c_group + (1|record_id),data = mod_data)
ri_rs_mod = lmer(sensor_mean ~ numeric_time*hba1c_group + (numeric_time|record_id),
                 data = mod_data)
```

# Results

```{r model function}
fit_mod = function(df = mod_data,outcome,group = NULL,
                   family = "gaussian",diagnostics = F){
  f = as.formula(paste0(outcome,"~","numeric_time*",group,"+(1|record_id)"))
  # Fit
  mod = lmer(f,data = df)
  # Print diagnostics if necessary
  if(diagnostics){
    mod_check = check_model(mod)
    print(mod_check)
  }
  # Trend over time for groups
  t = as.formula(paste0(group," ~ numeric_time"))
  trends = emmip(mod,t,cov.reduce = range)+
    ggtitle("Estimated Change Over Time") + 
    xlab("Months From Baseline") + theme_bw() + 
    scale_x_continuous(breaks = c(0,1,3,6,9,12))
  print(trends)
  # Print table
  coeffs = tidy(mod,"fixed") %>% select(term:p.value)
  kable(coeffs)
}
```

## Sensor mean

### Full cohort

```{r}
fit_mod(outcome = "sensor_mean")
```

### By HbA1c group

```{r}
fit_mod(outcome = "sensor_mean",group = "hba1c_group")
```

### By age group

```{r}
fit_mod(outcome = "sensor_mean",group = "age_group")
```

# References
