---
title: "Control IQ 12 Month Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(lmerTest)
library(censReg)
library(plm)
library(performance)
library(broom.mixed)
library(emmeans)
library(tidyverse)
library(knitr)
home_dir = 
  ifelse(.Platform$OS.type != "unix",
         "B:/Projects/Cari Berget/Control IQ 12 Month MS",
         "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Cari Berget/Control IQ 12 Month MS")
knitr::opts_knit$set(root.dir = home_dir)
knitr::opts_chunk$set(echo = FALSE,message = FALSE,results='asis',fig.width=12,fig.height=10)
```

```{r data,include=FALSE}
# Import
source("~/GitHub/BDC-Code/Cari Berget/Control IQ 12 month analysis/import_peds_data.r")
source("~/GitHub/BDC-Code/Cari Berget/Control IQ 12 month analysis/import_parent_data.r")
# Survey scoring function
source("~/GitHub/BDC-Code/Cari Berget/Scoring/scoring_function.R")
# Fill down by participant, baseline HbA1c group
data = data %>% group_by(record_id) %>% 
  fill(demographics_diabetesdx,demographics_dob,demographics_sex.factor,demographics_age)
# Calculate variables
data = data %>% filter(!is.na(gyl_timepoint.factor),record_id >= 119) %>%
  mutate(sensor_over_180 = sensor_181_250 + sensor_g250,
         sensor_under_70 = sensor_54_69 + sensor_u54) %>%
  group_by(record_id) %>%
  mutate(hba1c_group = cut(hba1c[row_number() == 1],c(-Inf,7,9,Inf),
                           labels = c("<7%","7-9%",">=9%"),right = F),
         age_group = cut(demographics_age[row_number() == 1],c(-Inf,14,18,Inf),
                         labels = c("<14","14-17",">=18"),right = F))
# Score surveys
data = peds_scores(data)
parent_data = parent_scores(parent_data)
# Parent surveys to long
parent_data = parent_data %>% select(record_id,p_paid_b:p_tech_b) %>%
  pivot_longer(p_paid_b:p_tech_b,
               names_to = c("survey","gyl_timepoint.factor"),
               names_pattern = "p_(.*)_(.*)") %>%
  pivot_wider(names_from = "survey",names_prefix = "p_")
parent_data$gyl_timepoint.factor =
  factor(parent_data$gyl_timepoint.factor,
         levels = c("b","m1","t1","t2","t3","t4"),
         labels = c("Baseline","Month 1/ Training F/U","3 Months",
                    "6 Months","9 Months","12 Months"))
# Dicontinuers
discontinuers = unique(data$record_id[data$time_am < 10])
discontinuers = discontinuers[!is.na(discontinuers)]
# Numeric time
data$numeric_time = data$gyl_timepoint.factor
levels(data$numeric_time) = c(0,1,3,6,9,12)
data$numeric_time = as.numeric(as.character(data$numeric_time))
# Select relevant columns, exclude discontinuers
data = data %>% 
  filter(!record_id %in% discontinuers,!grepl("arm_3",redcap_event_name)) %>%
  select(record_id,gyl_timepoint.factor,numeric_time,demographics_age,age_group,hba1c_group,hba1c,
         demographics_sex.factor,time_am,sensor_wear,tdd,tdd_bolus,tdd_basal,
         dailymealbolus,sleep,exercise,sensor_mean,sensor_70_180,
         sensor_over_180,sensor_g250,sensor_under_70,sensor_u54,
         all_of(peds_surveys))
# Add parent scores
data$record_id = as.character(data$record_id)
parent_data$record_id = as.character(parent_data$record_id)
data = left_join(data,parent_data,by = c("record_id", "gyl_timepoint.factor"))
# Write cleaned CSV
write.csv(data,file = "./Data_Cleaned/ciq_analysis_dataset.csv",row.names = F,na = "")
# Clean workspace
rm(list = ls()[-which(ls() %in% c("data","discontinuers"))])
```

# Table 1: Participant Characteristics at Baseline

A total of `r length(discontinuers)` participants were excluded due to HCL use <10% at 1 or more visits. 

```{r table 1}
t1 = tableby(~ demographics_age + age_group + hba1c + hba1c_group + demographics_sex.factor + 
               sensor_wear + tdd + tdd_bolus + tdd_basal + dailymealbolus + 
               sensor_mean + sensor_70_180 + sensor_over_180 + 
               sensor_g250 + sensor_under_70 + sensor_u54 + c_paid + c_hfs_maintain + 
               c_hfs_helpless + c_hfs_worry + c_inspire + ya_paid + ya_hfs_behave + 
               ya_hfs_worry + adult_inspire + c_tech + p_paid + p_hfs_maintain + 
               p_hfs_helpless + p_hfs_worry + p_inspire + p_tech,
             data = data[data$gyl_timepoint.factor == "Baseline",])
summary(t1,labelTranslations = 
          list(age_group = "Age Group",hba1c_group = "HbA1c Group",
               demographics_sex.factor = "Sex"))
```

# Methods

## Model selection

Time was treated as numeric because timepoints are not evenly spaced, and the interpretation of interactions between a large number of categories is difficult. The final model structure was selected using mean sensor glucose as the outcome, and the same model structure was used for all following models for consistency. Although better by AIC, some random slope models did not converge, so a random intercept for participant was used in all models to account for within-subject autocorrelation.

All analyses were performed using R version `r paste0(R.version$major,".",R.version$minor)` [@rcoreteamLanguageEnvironmentStatistical2021]. Linear mixed models were fit using the `lme4` [@batesFittingLinearMixedEffects2014] package and all model assumptions were assessed visually using the `performance` package [@ludeckePerformancePackageAssessment2021]. 

```{r model selection,include=FALSE}
# Remove outlier values
mod_data = data %>% filter(sensor_mean <= 400)
# Use numeric time since the M1 visit is not equally spaced. Also, interpreting the 
# interactions between categorical time and categorical HbA1c group is a nightmare.
ri_mod = lmer(sensor_mean ~ numeric_time*hba1c_group + (1|record_id),data = mod_data)
ri_rs_mod = lmer(sensor_mean ~ numeric_time*hba1c_group + (numeric_time|record_id),
                 data = mod_data)
```

```{r model function}
fit_mod = function(df = mod_data,outcome,group = NULL,y_axis = NULL,
                   diagnostics = F,baseline = T){
  # Is there data at baseline?
  if (!baseline){
    df = df[df$gyl_timepoint.factor != "Baseline",]
    df$gyl_timepoint.factor = droplevels(df$gyl_timepoint.factor)
  }
  # Equation formula
  f = as.formula(paste0(outcome,"~","gyl_timepoint.factor*",group,"+(1|record_id)"))
  # Fit
  mod = lmer(f,data = df)
  # Print type 3 test
  type3 = data.frame(car::Anova(mod))
  rownames(type3) = sub("gyl_timepoint.factor","Timepoint",rownames(type3))
  rownames(type3) = sub("hba1c_group","HbA1c Group",rownames(type3))
  rownames(type3) = sub("age_group","Age Group",rownames(type3))
  print(kable(type3,digits = 3,caption = "Type 3 Tests of Fixed Effects"))
  # Contrasts
  if (!is.null(group)){
    # Group and timepoint means
    f_means = as.formula(paste0("~","gyl_timepoint.factor*",group))
    # Contrasts
    if (!baseline){
      if(nrow(data.frame(emmeans(mod,f_means))) == 15){
        m1_low = c(1,rep(0,14))
        m3_low = c(0,1,rep(0,13))
        m6_low = c(rep(0,2),1,rep(0,12))
        m9_low = c(rep(0,3),1,rep(0,11))
        m12_low = c(rep(0,4),1,rep(0,10))
        
        m1_med = c(rep(0,5),1,rep(0,9))
        m3_med = c(rep(0,6),1,rep(0,8))
        m6_med = c(rep(0,7),1,rep(0,7))
        m9_med = c(rep(0,8),1,rep(0,6))
        m12_med = c(rep(0,9),1,rep(0,5))
        
        m1_high = c(rep(0,10),1,rep(0,4))
        m3_high = c(rep(0,11),1,rep(0,3))
        m6_high = c(rep(0,12),1,rep(0,2))
        m9_high = c(rep(0,13),1,rep(0,1))
        m12_high = c(rep(0,14),1)
        con = emmeans(mod,f_means,
                    contr = list("M1 to M3 Low Group" = m3_low - m1_low,
                                 "M3 to M6 Low Group" = m6_low - m3_low,
                                 "M3 to M9 Low Group" = m9_low - m3_low,
                                 "M3 to M12 Low Group" = m12_low - m3_low,
                                 "M1 to M3 Med. Group" = m3_med - m1_med,
                                 "M3 to M6 Med. Group" = m6_med - m3_med,
                                 "M3 to M9 Med. Group" = m9_med - m3_med,
                                 "M3 to M12 Med. Group" = m12_med - m3_med,
                                 "M1 to M3 High Group" = m3_high - m1_high,
                                 "M3 to M6 High Group" = m6_high - m3_high,
                                 "M3 to M9 High Group" = m9_high - m3_high,
                                 "M3 to M12 High Group" = m12_high - m3_high,
                                 "M1 to M3 Low vs. High" = (m3_low - m1_low) - (m3_high - m1_high),
                                 "M1 to M3 Med. vs. High" = (m3_med - m1_med) - (m3_high - m1_high),
                                 "M1 to M3 Low vs. Med." = (m3_low - m1_low) - (m3_med - m1_med),
                                 "M3 to M6 Low vs. High" = (m6_low - m3_low) - (m6_high - m3_high),
                                 "M3 to M6 Med. vs. High" = (m6_med - m3_med) - (m6_high - m3_high),
                                 "M3 to M6 Low vs. Med." = (m6_low - m3_low) - (m6_med - m3_med),
                                 "M3 to M9 Low vs. High" = (m9_low - m3_low) - (m9_high - m3_high),
                                 "M3 to M9 Med. vs. High" = (m9_med - m3_med) - (m9_high - m3_high),
                                 "M3 to M9 Low vs. Med." = (m9_low - m3_low) - (m9_med - m3_med),
                                 "M3 to M12 Low vs. High" = (m12_low - m3_low) - (m12_high - m3_high),
                                 "M3 to M12 Med. vs. High" = (m12_med - m3_med) - (m12_high - m3_high),
                                 "M3 to M12 Low vs. Med." = (m12_low - m3_low) - (m12_med - m3_med)))
      } else if (nrow(data.frame(emmeans(mod,f_means))) == 10){
        m1_low = c(1,rep(0,9))
        m3_low = c(0,1,rep(0,8))
        m6_low = c(rep(0,2),1,rep(0,7))
        m9_low = c(rep(0,3),1,rep(0,6))
        m12_low = c(rep(0,4),1,rep(0,5))
        
        m1_high = c(rep(0,5),1,rep(0,4))
        m3_high = c(rep(0,6),1,rep(0,3))
        m6_high = c(rep(0,7),1,rep(0,2))
        m9_high = c(rep(0,8),1,rep(0,1))
        m12_high = c(rep(0,9),1)
        
        con = emmeans(mod,f_means,
                    contr = list("M1 to M3 Low Group" = m3_low - m1_low,
                                 "M3 to M6 Low Group" = m6_low - m3_low,
                                 "M3 to M9 Low Group" = m9_low - m3_low,
                                 "M3 to M12 Low Group" = m12_low - m3_low,
                                 "M1 to M3 High Group" = m3_high - m1_high,
                                 "M3 to M6 High Group" = m6_high - m3_high,
                                 "M3 to M9 High Group" = m9_high - m3_high,
                                 "M3 to M12 High Group" = m12_high - m3_high,
                                 "M1 to M3 Low vs. High" = (m3_low - m1_low) - (m3_high - m1_high),
                                 "M3 to M6 Low vs. High" = (m6_low - m3_low) - (m6_high - m3_high),
                                 "M3 to M9 Low vs. High" = (m9_low - m3_low) - (m9_high - m3_high),
                                 "M3 to M12 Low vs. High" = (m12_low - m3_low) - (m12_high - m3_high)))
      }
    } else {
      if(nrow(data.frame(emmeans(mod,f_means))) == 18) {
        base_low = c(1,rep(0,17))
        m3_low = c(rep(0,2),1,rep(0,15))
        m6_low = c(rep(0,3),1,rep(0,14))
        m9_low = c(rep(0,4),1,rep(0,13))
        m12_low = c(rep(0,5),1,rep(0,12))
        
        base_med = c(rep(0,6),1,rep(0,11))
        m3_med = c(rep(0,8),1,rep(0,9))
        m6_med = c(rep(0,9),1,rep(0,8))
        m9_med = c(rep(0,10),1,rep(0,7))
        m12_med = c(rep(0,11),1,rep(0,6))
        
        base_high = c(rep(0,12),1,rep(0,5))
        m3_high = c(rep(0,14),1,rep(0,3))
        m6_high = c(rep(0,15),1,rep(0,2))
        m9_high = c(rep(0,16),1,0)
        m12_high = c(rep(0,17),1)
        
        con = emmeans(mod,f_means,
                      contr = list("Baseline to M3 Low Group" = m3_low - base_low,
                                   "M3 to M6 Low Group" = m6_low - m3_low,
                                   "M3 to M9 Low Group" = m9_low - m3_low,
                                   "M3 to M12 Low Group" = m12_low - m3_low,
                                   "Baseline to M3 Med. Group" = m3_med - base_med,
                                   "M3 to M6 Med. Group" = m6_med - m3_med,
                                   "M3 to M9 Med. Group" = m9_med - m3_med,
                                   "M3 to M12 Med. Group" = m12_med - m3_med,
                                   "Baseline to M3 High Group" = m3_high - base_high,
                                   "M3 to M6 High Group" = m6_high - m3_high,
                                   "M3 to M9 High Group" = m9_high - m3_high,
                                   "M3 to M12 High Group" = m12_high - m3_high,
                                   "Baseline to M3 Low vs. High" = (m3_low - base_low) - (m3_high - base_high),
                                   "Baseline to M3 Med. vs. High" = (m3_med - base_med) - (m3_high - base_high),
                                   "Baseline to M3 Low vs. Med." = (m3_low - base_low) - (m3_med - base_med),
                                   "M3 to M6 Low vs. High" = (m6_low - m3_low) - (m6_high - m3_high),
                                   "M3 to M6 Med. vs. High" = (m6_med - m3_med) - (m6_high - m3_high),
                                   "M3 to M6 Low vs. Med." = (m6_low - m3_low) - (m6_med - m3_med),
                                   "M3 to M9 Low vs. High" = (m9_low - m3_low) - (m9_high - m3_high),
                                   "M3 to M9 Med. vs. High" = (m9_med - m3_med) - (m9_high - m3_high),
                                   "M3 to M9 Low vs. Med." = (m9_low - m3_low) - (m9_med - m3_med),
                                   "M3 to M12 Low vs. High" = (m12_low - m3_low) - (m12_high - m3_high),
                                   "M3 to M12 Med. vs. High" = (m12_med - m3_med) - (m12_high - m3_high),
                                   "M3 to M12 Low vs. Med." = (m12_low - m3_low) - (m12_med - m3_med)))
        
      } else if (nrow(data.frame(emmeans(mod,f_means))) == 15) {
        base_low = c(1,rep(0,14))
        m3_low = c(0,1,rep(0,13))
        m6_low = c(rep(0,2),1,rep(0,12))
        m9_low = c(rep(0,3),1,rep(0,11))
        m12_low = c(rep(0,4),1,rep(0,10))
        
        base_med = c(rep(0,5),1,rep(0,9))
        m3_med = c(rep(0,6),1,rep(0,8))
        m6_med = c(rep(0,7),1,rep(0,7))
        m9_med = c(rep(0,8),1,rep(0,6))
        m12_med = c(rep(0,9),1,rep(0,5))
        
        base_high = c(rep(0,10),1,rep(0,4))
        m3_high = c(rep(0,11),1,rep(0,3))
        m6_high = c(rep(0,12),1,rep(0,2))
        m9_high = c(rep(0,13),1,rep(0,1))
        m12_high = c(rep(0,14),1)
        
        con = emmeans(mod,f_means,
                      contr = list("Baseline to M3 Low Group" = m3_low - base_low,
                                   "M3 to M6 Low Group" = m6_low - m3_low,
                                   "M3 to M9 Low Group" = m9_low - m3_low,
                                   "M3 to M12 Low Group" = m12_low - m3_low,
                                   "Baseline to M3 Med. Group" = m3_med - base_med,
                                   "M3 to M6 Med. Group" = m6_med - m3_med,
                                   "M3 to M9 Med. Group" = m9_med - m3_med,
                                   "M3 to M12 Med. Group" = m12_med - m3_med,
                                   "Baseline to M3 High Group" = m3_high - base_high,
                                   "M3 to M6 High Group" = m6_high - m3_high,
                                   "M3 to M9 High Group" = m9_high - m3_high,
                                   "M3 to M12 High Group" = m12_high - m3_high,
                                   "Baseline to M3 Low vs. High" = (m3_low - base_low) - (m3_high - base_high),
                                   "Baseline to M3 Med. vs. High" = (m3_med - base_med) - (m3_high - base_high),
                                   "Baseline to M3 Low vs. Med." = (m3_low - base_low) - (m3_med - base_med),
                                   "M3 to M6 Low vs. High" = (m6_low - m3_low) - (m6_high - m3_high),
                                   "M3 to M6 Med. vs. High" = (m6_med - m3_med) - (m6_high - m3_high),
                                   "M3 to M6 Low vs. Med." = (m6_low - m3_low) - (m6_med - m3_med),
                                   "M3 to M9 Low vs. High" = (m9_low - m3_low) - (m9_high - m3_high),
                                   "M3 to M9 Med. vs. High" = (m9_med - m3_med) - (m9_high - m3_high),
                                   "M3 to M9 Low vs. Med." = (m9_low - m3_low) - (m9_med - m3_med),
                                   "M3 to M12 Low vs. High" = (m12_low - m3_low) - (m12_high - m3_high),
                                   "M3 to M12 Med. vs. High" = (m12_med - m3_med) - (m12_high - m3_high),
                                   "M3 to M12 Low vs. Med." = (m12_low - m3_low) - (m12_med - m3_med)))
      } else if (nrow(data.frame(emmeans(mod,f_means))) == 10){
        base_low = c(1,rep(0,9))
        m3_low = c(0,1,rep(0,8))
        m6_low = c(rep(0,2),1,rep(0,7))
        m9_low = c(rep(0,3),1,rep(0,6))
        m12_low = c(rep(0,4),1,rep(0,5))
        
        base_high = c(rep(0,5),1,rep(0,4))
        m3_high = c(rep(0,6),1,rep(0,3))
        m6_high = c(rep(0,7),1,rep(0,2))
        m9_high = c(rep(0,8),1,rep(0,1))
        m12_high = c(rep(0,9),1)
        con = emmeans(mod,f_means,
                      contr = list("Baseline to M3 Low Group" = m3_low - base_low,
                                   "M3 to M6 Low Group" = m6_low - m3_low,
                                   "M3 to M9 Low Group" = m9_low - m3_low,
                                   "M3 to M12 Low Group" = m12_low - m3_low,
                                   "Baseline to M3 High Group" = m3_high - base_high,
                                   "M3 to M6 High Group" = m6_high - m3_high,
                                   "M3 to M9 High Group" = m9_high - m3_high,
                                   "M3 to M12 High Group" = m12_high - m3_high,
                                   "Baseline to M3 Low vs. High" = (m3_low - base_low) - (m3_high - base_high),
                                   "M3 to M6 Low vs. High" = (m6_low - m3_low) - (m6_high - m3_high),
                                   "M3 to M9 Low vs. High" = (m9_low - m3_low) - (m9_high - m3_high),
                                   "M3 to M12 Low vs. High" = (m12_low - m3_low) - (m12_high - m3_high)))
      }
    }
    # Plot
    means = data.frame(con$emmeans)
    means = means[!is.na(means$emmean),]
    mean_plot = ggplot(means,aes_string(y = "emmean",x = "gyl_timepoint.factor",
                                        group = group,color = group)) + 
      geom_line() + xlab("Timepoint") + ylab(outcome) + theme_bw()
  } else {
    # Timepoint means
    f_means = as.formula(paste0("~","gyl_timepoint.factor"))
    if (!baseline){
      m1 = c(1,rep(0,4))
      m3 = c(0,1,rep(0,3))
      m6 = c(rep(0,2),1,rep(0,2))
      m9 = c(rep(0,3),1,rep(0,1))
      m12 = c(rep(0,4),1)
      # Contrasts
      con = emmeans(mod,f_means,
                    contr = list("M1 to M3" = m3 - m1,
                                 "M3 to M6" = m6 - m3,
                                 "M3 to M9" = m9 - m3,
                                 "M3 to M12" = m12 - m3))
    } else {
      if(nrow(summary(mod)$coefficients) == 6) {
        base = c(1,rep(0,5))
        m3 = c(rep(0,2),1,rep(0,3))
        m6 = c(rep(0,3),1,rep(0,2))
        m9 = c(rep(0,4),1,rep(0,1))
        m12 = c(rep(0,5),1)
      } else {
        base = c(1,rep(0,4))
        m3 = c(rep(0,1),1,rep(0,3))
        m6 = c(rep(0,2),1,rep(0,2))
        m9 = c(rep(0,3),1,rep(0,1))
        m12 = c(rep(0,4),1)
      }
      # Contrasts
      con = emmeans(mod,f_means,
                    contr = list("Baseline to M3" = m3 - base,
                                 "M3 to M6" = m6 - m3,
                                 "M3 to M9" = m9 - m3,
                                 "M3 to M12" = m12 - m3))
    }
    # Plot
    means = data.frame(con$emmeans)
    means = means[!is.na(means$emmean),]
    mean_plot = ggplot(means,aes_string(y = "emmean",x = "gyl_timepoint.factor")) + 
      geom_line(aes(group = 1)) + xlab("Timepoint") + ylab(outcome) + theme_bw()
  }
  print(kable(con$emmeans,digits = 3,caption = "Means"))
  print(kable(con$contrasts,digits = 3,caption = "Comparisons"))
  print(mean_plot)
  # Print diagnostics if necessary
  if(diagnostics){
    mod_check = check_model(mod)
    try(print(mod_check))
  }
}
```

Type 3 tests of fixed effect indicate the overall effect for a categorical variable, and in the coefficient tables ":" indicates an interaction term. So, for example, a significant p value in the "Months:HbA1c group" row of the type 3 table indicates an overall significant interaction between time and HbA1c group. 

```{r}
analysis_vars = c("hba1c","sensor_mean","sensor_70_180","sensor_over_180",
                  "sensor_g250","sensor_under_70","sensor_u54","time_am",
                  "sensor_wear","tdd","tdd_basal","tdd_bolus","dailymealbolus",
                  "sleep","exercise","c_paid","p_paid","c_inspire","p_inspire",
                  "c_hfs_helpless","p_hfs_helpless","c_hfs_maintain","p_hfs_maintain",
                  "c_hfs_worry","p_hfs_worry")
for (v in analysis_vars){
  b = ifelse(all(is.na(mod_data[mod_data$gyl_timepoint.factor=="Baseline",v])),F,T)
  cat("\n")
  cat(paste0("# ",v))
  cat("\n")
  cat("## Full cohort")
  fit_mod(outcome = v,baseline = b)
  cat("\n")
  cat("\n")
  cat("## By HbA1c group ")
  cat("\n")
  fit_mod(outcome = v,group = "hba1c_group",baseline = b)
  cat("\n")
  cat("\n")
  cat("## By age group ")
  cat("\n")
  fit_mod(outcome = v,group = "age_group",baseline = b)
  cat("\n")
}
```
