---
title: "Optimal Duration of GRI"
author: "Casey Sakamoto"
date: "4/14/2022"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(readxl) # data import
library(knitr) # tables
library(tidyverse) # data manip
library(ggplot2) # plots
library(table1)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Viral Shah/GRI"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/GRI"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/GRI"
}
knitr::opts_knit$set(root.dir = home_dir)

setwd(home_dir)
rm(home_dir)
```

```{r data import, include=FALSE}
# subjid is firstlast, columns for 7, 14, 30, 60 , 90 day gri
gri = read_csv("S:/Laura/BDC/Projects/Kaan Akturk/CGM Sampling/Data_Cleaned/gri.csv")
gri_val = c("day7_gri","day14_gri","day30_gri","day60_gri","day90_gri")

# subject info
subject_info = read_csv("S:/Laura/BDC/Projects/Kaan Akturk/CGM Sampling/Data_Cleaned/subject_info.csv")
baseline_char = read_excel("S:/Laura/BDC/Projects/Viral Shah/GRI/Data Clean/baseline_char.xlsx", 
    na = "NULL")
# compute correlations
gri_cor = cor(gri[,gri_val], use = "pairwise.complete.obs")

# put into data frame and compute r^2
gri_cor = as.data.frame(gri_cor[,5])
gri_cor = rename(gri_cor, "r" =`gri_cor[, 5]`)
gri_cor$r_sq = gri_cor$r^2
gri_cor$Day = factor(c("7","14","30","60","90"), levels = c("7", "14","30", "60", "90"))

# For demog table want age, sex, diabetes duration, a1c, bmi
# dont have t1d or bmi yet
subject_info = subject_info %>% mutate(id = tolower(paste0(FirstName, LastName))) %>% select(id, Gender, Age, MostRecentA1C) %>% 
  filter(id %in% gri$id)
baseline_char = baseline_char %>% mutate(id = tolower(paste0(FirstName, LastName))) %>% select(id, Gender, Age, MostRecentA1C, DiabetesDuration_MostRecentVisitDate, BMI) %>% 
  filter(id %in% gri$id)
# not sure why theres some missing data from the data w baseline char and tims data, so ill merge them
subject_info = full_join(subject_info, baseline_char)

gri = full_join(subject_info, gri)
gri[gri$id == "cortneyfernandez", "Gender"] = "F"
gri[gri$id == "cortneyfernandez", "Age"] = 41.41
gri[gri$id == "cortneyfernandez", "MostRecentA1C"] = 6.6
gri[gri$id == "cortneyfernandez", "DiabetesDuration_MostRecentVisitDate"] = 12.36
gri[gri$id == "cortneyfernandez", "BMI"] = 21.87

gri[gri$id == "lynellrice", "Gender"] = "F"
gri[gri$id == "lynellrice", "Age"] = 41.75
gri[gri$id == "lynellrice", "MostRecentA1C"] = 6.5
gri[gri$id == "lynellrice", "DiabetesDuration_MostRecentVisitDate"] = 20.99
gri[gri$id == "lynellrice", "BMI"] = 28.13

```


# Background
The objective is to evaluate optimal duration of CGM data for GRI measurement. This is similar to work previously published to figure out optimal duration of CGM data for TIR and A1c (ref attached). 

# Outstanding Data Questions/Remarks
N/A

# Methods
R2 and R of GRI over 7-days, 14-days, 30-days, 60-days and 90-days vs 90 days GRI were calculated and plotted using the ggplot2 package in r. 

# Analysis
```{r table manip}
label(gri$MostRecentA1C) = "Most Recent A1C"
label(gri$Gender) = "Sex"
label(gri$DiabetesDuration_MostRecentVisitDate) = "Diabetes Duration"

rndr <- function(x, name, ...) {
    if (!is.numeric(x)) return(render.categorical.default(x))
    what <- switch(name,
        Age = "Median [Q1, Q3]",
        DiabetesDuration_MostRecentVisitDate  = "Median [Q1, Q3]",
        MostRecentA1C = "Median [Q1, Q3]",
        BMI = "Median [Q1, Q3]")
    parse.abbrev.render.code(c("", what))(x)
}

demog_table = table1( ~ Age + Gender + DiabetesDuration_MostRecentVisitDate + MostRecentA1C + BMI, render = rndr,data = gri)
```

```{r baseline char}
demog_table
```

Table 1 of the baseline characteristics added in 4/21/2022


```{r analysis}
ggplot(data=gri_cor, aes(x=Day, y = r)) + 
  geom_point() +
  scale_y_continuous(name = "R", sec.axis = sec_axis(~.^2, name = "R2")) +
  theme_classic()
```

We observe a very strong linear relationship between the 7 day GRI and 90 day GRI (r = 0.892). Increasing the days to 14 from 7, we observe a 0.046 increase in the r coefficient. Increasing the days to 30 from 14 we observe a 0.029 increase in the r coefficient. Increasing the days to 60 from 30 we observe a 0.023 increase in the r coefficient.

```{r corr table}
kable(gri_cor[,1:2], digits = 3)
```