---
title: "A1c and TIR"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(knitr)
library(tableone)
library(dplyr)
library(tidyr)
library(stringr)
library(car)

alldata <- read.csv("B:\\Projects\\Todd Alonso\\A1c and TIR\\Raw data\\Peds_T1D_5groups_5_11_withDKAcol.csv",na.strings = c(" ","","NULL","#VALUE","#VALUE!","?"))

### LIMIT TO >= 70% SENSOR WEAR...SHOULD BE N=1952
alldata$CGMUsePercent <- ifelse(alldata$CGMUsePercent %in% c("attemptedboth","x"),NA,alldata$CGMUsePercent)
alldata$CGMUsePercent <- as.numeric(as.character(alldata$CGMUsePercent))
data <- alldata %>% filter(CGMUsePercent>=70)
data <- data %>% filter(OnCGM==1)

# fix DKA variable
data$DKA.at.onset <- str_trim(data$DKA.at.onset)
data$DKA.at.onset <- ifelse(data$DKA.at.onset=="UNKN",NA,data$DKA.at.onset)

# recode any A1c >14 to 14
data$A1C_Value <- ifelse(data$A1C_Value>14,14,data$A1C_Value)

# factors
data$Age_Groups <- as.factor(data$Age_Groups)
data$Gender <- as.factor(data$Gender)
data$English <- as.factor(data$English)
data$Spanish <- as.factor(data$Spanish)
data$OnCGM <- as.factor(data$OnCGM)
data$OnInsulinPump <- as.factor(data$OnInsulinPump)
data$Rural <- as.factor(data$Rural)
data$OnLoop <- as.factor(data$OnLoop)

# age group 6-11
data$age_6_11 <- ifelse(data$Age_LastVisitDate>=6 & data$Age_LastVisitDate<12,1,0)

# df with HbA1c >=14 excluded
data_a1clt14 <- data %>% filter(A1C_Value<14)

# separate dfs by DKA at diagnosis
dkayes <- data %>% filter(DKA.at.onset=="Y")
dkano <- data %>% filter(DKA.at.onset=="N")
dkayes_a1clt14 <- dkayes %>% filter(A1C_Value<14)
dkano_a1clt14 <- dkano %>% filter(A1C_Value<14)

# raw correlation
corsave <- cor(data$A1C_Value,data$CGMInRange,use = "complete.obs")
corsave_dkayes <- cor(dkayes$A1C_Value,dkayes$CGMInRange,use = "complete.obs")
corsave_dkano <- cor(dkano$A1C_Value,dkano$CGMInRange,use = "complete.obs")

# raw correlation A1c<14
corsave_a1clt14 <- cor(data_a1clt14$A1C_Value,data_a1clt14$CGMInRange,use = "complete.obs")
corsave_dkayes_a1clt14 <- cor(dkayes_a1clt14$A1C_Value,dkayes_a1clt14$CGMInRange,use = "complete.obs")
corsave_dkano_a1clt14 <- cor(dkano_a1clt14$A1C_Value,dkano_a1clt14$CGMInRange,use = "complete.obs")

# linear models - all patients, DKA at onset, no DKA at onset
mod <- glm(data$CGMInRange~data$A1C_Value)
mod_dkayes <- glm(dkayes$CGMInRange~dkayes$A1C_Value)
mod_dkano <- glm(dkano$CGMInRange~dkano$A1C_Value)

# linear models - all patients, DKA at onset, no DKA at onset
# A1c <14
mod_a1clt14 <- glm(data_a1clt14$CGMInRange~data_a1clt14$A1C_Value)
mod_dkayes_a1clt14 <- glm(dkayes_a1clt14$CGMInRange~dkayes_a1clt14$A1C_Value)
mod_dkano_a1clt14 <- glm(dkano_a1clt14$CGMInRange~dkano_a1clt14$A1C_Value)

# checking the relationship in those with percent use <70
lowuse <- alldata[alldata$CGMUsePercent<70,]
corsavelow <- cor(lowuse$A1C_Value,lowuse$CGMInRange,use = "complete.obs")
modlow <- glm(lowuse$CGMInRange~lowuse$A1C_Value)


```

# Background

The purpose of this analysis is to examine the association between HbA1c and CGM time in range, as well as the influence of age, duration since diagnosis, and DKA at diagnosis.

# Methods

Patient recorded as not on CGM or with CGM use >=70% were excluded.  Pearson's correlation coefficient and generalized linear models were used to examine the relationship between HbA1c and CGM time in range.

# Results

## HbA1c values >14% included, recoded as 14%

### Table of descriptive statistics

```{r echo=FALSE, include=F}
# table of descriptive statistics
t1 <- CreateTableOne(data=data,vars=c("Age_LastVisitDate","Age_Groups","Gender","DiabetesDuration_LastVisitDate","Race_Ethnicity",
                                 "English","Spanish","A1C_Value","OnCGM","CGMUsePercent","CGMMeanGlucose",
                                 "CGMHighPercent","CGMLowPercent","CGMInRange","OnInsulinPump","OnLoop",
                                 "Rural","InsuranceCategory","DKA.at.onset"))
t1 <- print(t1)
```

```{r echo=FALSE, comment=""}
kable(t1)
```

### Association of DKA at onset and HbA1c

```{r echo=FALSE, include=F}
t2 <- CreateTableOne(data=data,vars="A1C_Value",strata = "DKA.at.onset")
t2 <- print(t2,nonnormal = "A1C_Value")
```

```{r echo=FALSE, comment=""}

kable(t2)
```

### Scatterplot and unadjusted model 

The unadjusted correlation was `r corsave`.  The results of the unadjusted model are shown below.  For each one unit increase in HbA1c, CGM TIR decreased by `r mod$coefficients[2]`.

```{r echo=FALSE, comment=""}
plot(data$A1C_Value,data$CGMInRange)

summary(mod)
```

### Correlation in ages 6-11

The overall correlation in patients aged 6-11 was `r cor(data[data$age_6_11==1,]$A1C_Value,data[data$age_6_11==1,]$CGMInRange,use = "complete.obs")`.  
```{r echo=FALSE, comment=""}
plot(data[data$age_6_11==1,]$A1C_Value,data[data$age_6_11==1,]$CGMInRange)
```

The correlation in patients aged 6-11 with DKA at diagnosis was `r cor(data[data$age_6_11==1 & data$DKA.at.onset=='Y',]$A1C_Value,data[data$age_6_11==1 & data$DKA.at.onset=='Y',]$CGMInRange,use = "complete.obs")` and the correlation in patients 6-11 without DKA at diagnosis was `r cor(data[data$age_6_11==1 & data$DKA.at.onset=='N',]$A1C_Value,data[data$age_6_11==1 & data$DKA.at.onset=='N',]$CGMInRange,use = "complete.obs")`.

### Model with interaction of HbA1c with age group, duration, and DKA at onset

Results of a model with interactions between HbA1c with age group, diabetes duration, and DKA at onset are shown below.  The interaction between HbA1c and age group was not significant, while the interaction between HbA1c and DKA at onset was almost significant.  The interaction between HbA1c and diabetes duration was highly significant. 

```{r echo=FALSE, comment=""}
modint <- glm(data=data,CGMInRange~A1C_Value * Age_Groups + A1C_Value * DiabetesDuration_LastVisitDate + A1C_Value*DKA.at.onset)

Anova(modint,type="III")
summary(modint)
```

### Model with interaction of HbA1c with duration and DKA at onset

Because the interaction between HbA1c and age group was not significant, it was removed from the model.  The reduced model results are shown below.

```{r echo=FALSE, comment=""}
modint2 <- glm(data=data,CGMInRange~A1C_Value + Age_Groups + A1C_Value * DiabetesDuration_LastVisitDate + A1C_Value*DKA.at.onset)

Anova(modint2,type="III")
summary(modint2)

```

### Association by DKA at diagnosis

The interaction of HbA1c and DKA at diagnosis was almost significant.  Patients were stratified by DKA at diagnosis and the unadjusted associations between HbA1c and CGM TIR are shown below. The slope of the relationship between HbA1c and CGM TIR was steeper in those with DKA at diagnosis.

#### DKA at diagnosis

The unadjusted correlation in those with DKA at diagnosis was `r corsave_dkayes`.  

```{r echo=FALSE, comment=""}
plot(dkayes$A1C_Value,dkayes$CGMInRange)

summary(mod_dkayes)
```

#### No DKA at diagnosis

The unadjusted correlation in those without DKA at diagnosis was `r corsave_dkano`.

```{r echo=FALSE, comment=""}
plot(dkano$A1C_Value,dkano$CGMInRange)

summary(mod_dkano)
```

## HbA1c values >=14% excluded

### Table of descriptive statistics

```{r echo=FALSE, include=F}
# table of descriptive statistics
t1 <- CreateTableOne(data=data_a1clt14,vars=c("Age_LastVisitDate","Age_Groups","Gender","DiabetesDuration_LastVisitDate","Race_Ethnicity",
                                 "English","Spanish","A1C_Value","OnCGM","CGMUsePercent","CGMMeanGlucose",
                                 "CGMHighPercent","CGMLowPercent","CGMInRange","OnInsulinPump","OnLoop",
                                 "Rural","InsuranceCategory","DKA.at.onset"))
t1 <- print(t1)
```

```{r echo=FALSE, comment=""}
kable(t1)
```

### Association of DKA at onset and HbA1c

```{r echo=FALSE, include=F}
t2 <- CreateTableOne(data=data_a1clt14,vars="A1C_Value",strata = "DKA.at.onset")
t2 <- print(t2,nonnormal = "A1C_Value")
```

```{r echo=FALSE, comment=""}

kable(t2)
```

### Scatterplot and unadjusted model 

The unadjusted correlation was `r corsave_a1clt14`.  The results of the unadjusted model are shown below.  For each one unit increase in HbA1c, CGM TIR decreased by `r mod_a1clt14$coefficients[2]`.

```{r echo=FALSE, comment=""}
plot(data_a1clt14$A1C_Value,data_a1clt14$CGMInRange)

summary(mod_a1clt14)
```

### Correlation in ages 6-11

The overall correlation in patients aged 6-11 was `r cor(data_a1clt14[data_a1clt14$age_6_11==1,]$A1C_Value,data_a1clt14[data_a1clt14$age_6_11==1,]$CGMInRange,use = "complete.obs")`.  
```{r echo=FALSE, comment=""}
plot(data_a1clt14[data_a1clt14$age_6_11==1,]$A1C_Value,data_a1clt14[data_a1clt14$age_6_11==1,]$CGMInRange)
```

The correlation in patients aged 6-11 with DKA at diagnosis was `r cor(data_a1clt14[data_a1clt14$age_6_11==1 & data_a1clt14$DKA.at.onset=='Y',]$A1C_Value,data_a1clt14[data_a1clt14$age_6_11==1 & data_a1clt14$DKA.at.onset=='Y',]$CGMInRange,use = "complete.obs")` and the correlation in patients 6-11 without DKA at diagnosis was `r cor(data_a1clt14[data_a1clt14$age_6_11==1 & data_a1clt14$DKA.at.onset=='N',]$A1C_Value,data_a1clt14[data_a1clt14$age_6_11==1 & data_a1clt14$DKA.at.onset=='N',]$CGMInRange,use = "complete.obs")`.

### Model with interaction of HbA1c with age group, duration, and DKA at onset

Results of a model with interactions between HbA1c with age group, diabetes duration, and DKA at onset are shown below.  The interactions between HbA1c and age group and between HbA1c and DKA at onset were not significant.  The interaction between HbA1c and diabetes duration was highly significant. 

```{r echo=FALSE, comment=""}
modint <- glm(data=data_a1clt14,CGMInRange~A1C_Value * Age_Groups + A1C_Value * DiabetesDuration_LastVisitDate + A1C_Value*DKA.at.onset)

Anova(modint,type="III")
summary(modint)
```

### Model with interaction of HbA1c with duration and DKA at onset

Because the interaction between HbA1c and age group was not significant, it was removed from the model.  The reduced model results are shown below.

```{r echo=FALSE, comment=""}
modint2 <- glm(data=data_a1clt14,CGMInRange~A1C_Value + Age_Groups + A1C_Value * DiabetesDuration_LastVisitDate + A1C_Value*DKA.at.onset)

Anova(modint2,type="III")
summary(modint2)

```

### Association by DKA at diagnosis

Although the interaction of HbA1c and DKA at diagnosis was not significant after excluding A1c values >=14%, these results are included to be consistent with the analyses of the dataset with all A1c values.  Patients were stratified by DKA at diagnosis and the unadjusted associations between HbA1c and CGM TIR are shown below. The slope of the relationship between HbA1c and CGM TIR was steeper in those with DKA at diagnosis.

#### DKA at diagnosis

The unadjusted correlation in those with DKA at diagnosis was `r corsave_dkayes`.  

```{r echo=FALSE, comment=""}
plot(dkayes_a1clt14$A1C_Value,dkayes_a1clt14$CGMInRange)

summary(mod_dkayes_a1clt14)
```

#### No DKA at diagnosis

The unadjusted correlation in those without DKA at diagnosis was `r corsave_dkano`.

```{r echo=FALSE, comment=""}
plot(dkano_a1clt14$A1C_Value,dkano_a1clt14$CGMInRange)

summary(mod_dkano_a1clt14)
```
