---
title: "ASK SARS-COV-2 ab (ADA 2022)"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(tableone)
library(knitr)
library(car)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

data <- read.csv("B:\\Projects\\Marian Rewers\\ASK\\Data raw\\ASK_Dataset_Abstract_CovidScreening_2022-01-07.csv",na.strings = c("NA",""," "))
data$DM_pos <- ifelse(data$MAB=="Yes" | data$SBB=="Yes","YES","NO")
data$DM_pos_01 <- ifelse(data$DM_pos=="YES",1,
                                ifelse(!is.na(data$DM_pos),0,NA))
data$ECL_COVID_Out_01 <- ifelse(data$ECL_COVID_Out=="Positive",1,
                                ifelse(!is.na(data$ECL_COVID_Out),0,NA))

# categorize age
data$age_cat <- ifelse(is.na(data$AgeAtScreen),NA,
                       ifelse(data$AgeAtScreen<5, "0 - 4.9 yrs",
                              ifelse(data$AgeAtScreen<12,"5 - 11.9 yrs", "12+ yrs")))

t1 <- CreateTableOne(vars=c("DM_pos","MAB","SBB"),strata = "ECL_COVID_Out", data=data)
t1 <- print(t1,exact=c("DM_pos","MAB","SBB"))

t2 <- CreateTableOne(vars=c("DM_pos","MAB","SBB"),strata = "age_cat", data=data)
t2 <- print(t2,exact=c("DM_pos","MAB","SBB"))

mod1 <- glm(data=data, DM_pos_01 ~ ECL_COVID_Out_01, family="binomial")
or1 <- exp(cbind(OR = coef(mod1), confint(mod1)))

mod2 <- glm(data=data, DM_pos_01 ~ ECL_COVID_Out_01 + age_cat + Sex + T1D_Fam + NHW, family="binomial")
or2 <- exp(cbind(OR = coef(mod2), confint(mod2)))

```

# Table 1.  

The p-value tests whether there is a significant association between each of the DM Ab variables and SARS-Cov-2 positivity.  None of the associations were statistically significant.

```{r echo=FALSE}
kable(t1)
```

# Table 2.  

The p-value tests whether there is a significant association between each of the DM Ab variables and age category.  None of the associations were statistically significant.

```{r echo=FALSE}
kable(t2)
```

# Unadjusted logistic regression model

The p-value (labeled "Pr(>Chisq)") tests whether there is a significant association between DM Ab positivity and SARS-Cov-2 positivity.  The association was not statistically significant.  The odds ratio was `r round(or1[2,1],2)`, which means that people who tested positive for COVID-19 antibodies had `r round(or1[2,1],2)` times the odds of having positive DM antibodies, although this was not significant.


```{r echo=FALSE,comment=""}
Anova(mod1,type = "III")
or1
```

# Adjusted logistic regression model

The p-value (labeled "Pr(>Chisq)") for each term in the model tests whether there is a significant association between each variable and DM antibody positivity.  

The odds ratio for SARS-Cov-2 positivity was `r round(or2[2,1],2)`, which means that people with SARS-Cov-2 positivity had `r round(or2[2,1],2)` times the odds of having positive DM antibodies antibodies, although this was not significant.

```{r echo=FALSE,comment=""}
Anova(mod2,type = "III")
or2
```

