---
title: "Tandem Patient Reported Outcomes (PROs)"
author: "Kristen Campbell"
date: "May 15, 2020"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
setwd("C:/Users/campbkri/Documents/GitHub/BDC-Code/Laurel Messer/Tandem/Code/4times")
source("S:/Shared Material/Shared Code/R/temp_table1.R")

##functions specifically for this project
source("00_data_4times.R")
source("01_survey_4times.R")
source("03c_models_all.R")
dat.all<-dat
dat.all$in_models<-1
dat.all$in_models[dat.all$baseline_complete_yn==0]<-0
dat.all$in_models[is.na(dat.all$Baseline_A1C)]<-0
dat.all$in_models<-as.factor(dat.all$in_models)
label(dat.all$in_models)<-"In 4time Models (has baseline survey and baseline A1c)"

dat<-subset(dat.all,dat.all$in_models==1)
#demographics table:
demo<-final_table(dat.all,c('B_RESPONDENT','BaselineAGE','duration_of_diabetes_at_baseline_years',
                        'cgm_yn','method_cat'),
                     dat.all$in_models,margin=2,single=F,2,col.names=T, summary.stat='mean')

demo.type<-final_table(dat,c('Gender','B_RESPONDENT','BaselineAGE','duration_of_diabetes_at_baseline_years',
                        'Baseline_A1C','cgm_yn'),
                     dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='mean')
demo.type
demo.type<-demo.type[,-2]

##survey data:
missing<-final_table(dat.all,c('baseline_complete_yn','post2m_complete_yn', 'post4m_complete_yn','post6m_complete_yn',
                               'allcomplete','cohort_complete'),
                     rep(1:nrow(dat.all)),margin=2,single=T,2,col.names=T, summary.stat='median')
missing

factors<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="baseline_factor1"),
                                         which(colnames(dat)=="mid_factor1"),
                                         which(colnames(dat)=="post2m_factor1"),
                                         which(colnames(dat)=="post4m_factor1"),
                                        which(colnames(dat)=="post6m_factor1"),
                                        which(colnames(dat)=="baseline_factor2"),
                                        which(colnames(dat)=="mid_factor2"),
                                         which(colnames(dat)=="post2m_factor2"),
                                         which(colnames(dat)=="post4m_factor2"),
                                        which(colnames(dat)=="post6m_factor2"))]),
                     dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')
factors



# #opt in data:
# 
# opt_in<-final_table(dat,c('Baseline_SurveyStudy_OptIn','post2m_BIQuse','post4m_BIQuse','post6m_BIQuse',
#                           'post2m_BIQ_YES','post4m_BIQ_YES','post6m_BIQ_YES'),
#                      rep(1:nrow(dat)),margin=2,single=F,2,col.names=T, summary.stat='median')
# opt_in
#missing data:
# missing<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="baseline_num_complete"):
#                                         which(colnames(dat)=="allcomplete"))]),
#                      rep(1:nrow(dat)),margin=2,single=F,2,col.names=T, summary.stat='median')
# missing

#baseline survey summary:
baseline_surv<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="Baseline_1"):
                                        which(colnames(dat)=="Baseline_12"))]),
                     dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')

post2mo_surv<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="post2m_1"):
                                        which(colnames(dat)=="post2m_12"))]),
                     dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')

post4mo_surv<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="post4m_1"):
                                        which(colnames(dat)=="post4m_12"))]),
                     dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')

post6mo_surv<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="post6m_1"):
                                        which(colnames(dat)=="post6m_12"))]),
                     dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')
# post6mo_surv

####Models:

```


##Summary Stats
**Table 1: Patient Demographics - By Method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(demo.type, align = c("l", "l"))
```

##Missing Data
**Table 2a: Survey Missing Data**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(missing, align = c("l", "l"))
```

There are `r nrow(dat)` patients that met the final inclusion criteria for the analysis. Of those `r nrow(dat)` patients, `r table(dat$mid_point)[1]` (`r round(table(dat$mid_point)[1]/nrow(dat)*100,2)`%) finished a 2 month survey, so that time point was used as the mid-point.  `r table(dat$mid_point)[2]` (`r round(table(dat$mid_point)[2]/nrow(dat)*100,2)`%) did not have a 2 month survey, but did have a 4 month survey, so this was used as the mid-point.

##Satisfaction Factor
**Table 3a: Factor summary stats, by previous method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(mod1_data, align = c("l", "l"))
```

**Table 3b: Changes over time by method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(mod1, align = c("l", "l"))
```

##Diabetes Burden Factor
**Table 3a: Factor summary stats, by previous method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(mod2_data, align = c("l", "l"))
```

**Table 3b: Changes over time by method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(mod2, align = c("l", "l"))
```


##Survey Data - results
**Table 6: Factor summary stats, by previous method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(factors, align = c("l", "l"))
```

##Supplement:
**Table S3: Patient Demographics - Inclusion/Exclusion**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(demo, align = c("l", "l"))
```

**Table S4: Individual Survey Items:**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(baseline_surv, align = c("l", "l"))
```


```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(post2mo_surv, align = c("l", "l"))
```

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(post4mo_surv, align = c("l", "l"))
```

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(post6mo_surv, align = c("l", "l"))
```
