---
title: "JDRF Behavioral Intervention"
author: "Tim Vigers, Casey Sakamoto"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    number-sections: true
    toc-float: true
    code-fold: true
    self-contained: true
    fig-height: 10
    fig-width: 10
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: 
  markdown: 
    wrap: 72
---

```{r libraries and import, include = F}
#| include: false
library(tidyverse)
library(plotly)
library(redcapAPI)
library(gtsummary)
library(lmerTest)
library(GLMMadaptive)
library(emmeans)
library(broom.mixed)
library(performance)
library(kableExtra)
unlockREDCap(c(rcon = "JDRF Behavioral Intervention"),
  keyring = "API_KEYs",
  envir = globalenv(),
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = globalenv())
# Review invalid data
Records <- exportRecordsTyped(rcon)
reviewInvalidRecords(Records)

# api k:  0FB8B28FB5F4D22A98C5090488CE0BCB
```

```{r data cleaning, include=FALSE}
#| message: false
# Add treatment group to glycemic data
db_glycemic_data <- left_join(
  db_glycemic_data,
  db_randomization_form %>%
    filter(!is.na(treatment_group)) %>%
    select(record_id, treatment_group)
)
# Everyone else is considered "observe clinically"
db_glycemic_data$treatment_group <- as.character(db_glycemic_data$treatment_group)
db_glycemic_data$treatment_group[is.na(db_glycemic_data$treatment_group)] <-
  "Observe Clincically"
db_glycemic_data$treatment_group <- factor(db_glycemic_data$treatment_group,
  levels = c("Observe Clinically", "Behavioral Intervention", "Care+ (Control)")
)
# If they are missing gly_timepoint they don't have glycemic data at that visit
db_glycemic_data <- db_glycemic_data %>% filter(!is.na(gly_timepoint))
```

# Table 1

```{r, echo = FALSE}
#| label: tbl-1
#| tbl-cap: Participant Characteristics
#| # add in group
db_tgroups = db_glycemic_data %>% select(record_id, treatment_group) %>% unique()
db_demographics = left_join(db_demographics, db_tgroups)
#|
db_demographics %>% filter(redcap_event_name == "Baseline (Arm 1: Optimal Use)")  %>%
  select(redcap_event_name, treatment_group,demo_hcl, demo_hcl_2, demo_age:demo_predict) %>%
  mutate(
    redcap_event_name = droplevels(redcap_event_name),
    demo_probability = as.numeric(demo_probability)
  ) %>%
  tbl_summary(
    by = treatment_group,
    missing_text = "Missing",
    type = list(
      demo_age ~ "continuous", demo_probability ~ "continuous"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})")
  ) %>% add_overall() %>%
  as_kable_extra()
```

# APPENDED MAY 2024: SURVEY ANALYSIS

STILL CODING/WRTING UP SURVEY ANALYSES

o At baseline we
conducted the following surveys: Paid-PEDS, Inspire, Hypoglycemia Fear
Survey, Diabetes Technology Questionnaire. We need to see how the
baseline scores for each of these tools compared between the 3 groups. 

o In the intervention group we also gave an "Intervention" survey which
included questions about the sessions as well as Diabetes Centered
Family Conflict questions. I'd like to look at each question for this
one in addition to composite scores

# APPENDED MAY 2024 -- BOLUS ANALYSIS

• Glycemic Data (Table 2) -- Needs some Pump data as well

```{r bolus an}
# want to see if behavioral intervention impacted bolus behavior and how bolus affected TIR
# mediation function using lavaan package
library(lavaan)

# gly_dailybolus and gly_dailymealbolus the vars
df = db_glycemic_data %>% select(record_id, gly_timepoint, treatment_group, gly_dailybolus, gly_sensor_70_180) %>%  drop_na()
df$treatment_group = factor(df$treatment_group, levels = c("Care+ (Control)", "Behavioral Intervention"))
mediation_model = "
# mediator 
gly_dailybolus ~ a*treatment_group
gly_sensor_70_180 ~ b*gly_dailybolus
# direct effect
gly_sensor_70_180 ~ c*treatment_group
# indirect effect
ab := a*b
# total effect
total := c + (a*b)
"
fit = sem(mediation_model, data = df)
summary(fit)
parameterestimates(fit) %>% kable()
```


Question:"Did the BI impact bolusing behavior and if so did that drive improvement in TIR?" 


## Writeup

Remarks: In the main analysis LMM, we see that BI was not significantly associated with TIR 70-180 (p = 0.1).

Treatment group (Care + vs Behavioral Intervention) was not associated with avg number of daily boluses used (var gly_dailybolus) (p = 0.8).

Adjusting for treatment group, every 1-unit increase in average daily bolus use was associated with a 1.03 (S.E. = 0.37) higher TIR (p = 0.01).

Treatment group was not associated with TIR indirectly through increases in bolus use (95% CI:[-1.33, 1.03], p = 0.801). **indirect effect ab

There was no sufficient evidence treatment group was associated with TIR independent of its association with number of boluses used (p = 0.250)

# Glycemic data

## Groups by timepoint

```{r gbt}
db_glycemic_data %>%
  group_by(gly_timepoint) %>%
  summarise(
    n = length(unique(record_id)),
    `n Observe Clincically` = sum(treatment_group == "Observe Clincically"),
    `n Behavioral Intervention` = sum(treatment_group == "Behavioral Intervention"),
    `n Care+ (Control)` = sum(treatment_group == "Care+ (Control)")
  ) %>%
  kable()
# # Write for checking
# write.csv(
#   db_glycemic_data %>%
#     select(
#       record_id, redcap_event_name, redcap_survey_identifier,
#       treatment_group, everything()
#     ),
#   file = "~/cleaned_glycemic_data.csv", row.names = F, na = ""
# )
```

## Sensor wear

### Plot

```{r sw}
v <- "gly_sensorwear"
df <- db_glycemic_data %>%  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%  drop_na()

p <- ggplot(df, aes(  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r sw m}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r sw r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r sw me}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor under 54

### Plot

```{r su54}
v <- "gly_sensor_u54"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

Because the majority of people did not have any time under 54 mg/dL, we
used a hurdle model in which the first part of the model is the
probability of a zero value and the second part is the probability of
the non-zero values.

```{r su54 m}
f <- as.formula(paste0(v, " ~ gly_timepoint*treatment_group"))
mod <- mixed_model(f,
  random = ~ 1 | record_id, data = df,
  family = hurdle.lognormal(), n_phis = 1, zi_fixed = ~1
)
```

#### Results

```{r su54 r}
kable(tidy(mod), digits = 3)
```

#### Means

These model means do not really make sense since we can't have negative
values for this outcome, so this model likely needs additional work.

```{r su54 me}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor Under 70

### Plot

```{r}
db_glycemic_data$gly_sensor_u70 = db_glycemic_data$gly_sensor_u54 + db_glycemic_data$gly_sensor_54_69
v <- "gly_sensor_u70"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

Because the majority of people did not have any time in the range 54 -
69 mg/dL, we used a hurdle model in which the first part of the model is
the probability of a zero value and the second part is the probability
of the non-zero values.

```{r}
f <- as.formula(paste0(v, " ~ gly_timepoint*treatment_group"))
mod <- mixed_model(f,
  random = ~ 1 | record_id, data = df,
  family = hurdle.lognormal(), n_phis = 1, zi_fixed = ~1
)
```

#### Results

```{r}
kable(tidy(mod), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor 70 - 180

### Plot

```{r}
v <- "gly_sensor_70_180"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor Over 180

### Plot

```{r}
db_glycemic_data$gly_sensor_g181 = db_glycemic_data$gly_sensor_g250 + db_glycemic_data$gly_sensor_181_250
v <- "gly_sensor_g181"

df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor \> 250

### Plot

```{r}
v <- "gly_sensor_g250"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Mean sensor

### Plot

```{r}
v <- "gly_sensor_mean"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor SD

### Plot

```{r}
v <- "gly_sensor_sd"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint | treatment_group)), digits = 3)
```
