---
title: "JDRF Behavioral Intervention"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    number-sections: true
    toc-float: true
    code-fold: true
    self-contained: true
    fig-height: 10
    fig-width: 10
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/endnote.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

```{r libraries and import}
#| include: false
library(tidyverse)
library(plotly)
library(redcapAPI)
library(gtsummary)
library(lmerTest)
library(GLMMadaptive)
library(emmeans)
library(broom.mixed)
library(performance)
library(kableExtra)
unlockREDCap(c(rcon = "JDRF Behavioral Intervention"),
  keyring = "API_KEYs",
  envir = globalenv(),
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = globalenv())
# Review invalid data
Records <- exportRecordsTyped(rcon)
reviewInvalidRecords(Records)
```

```{r data cleaning}
#| message: false
# Add treatment group to glycemic data
db_glycemic_data <- left_join(
  db_glycemic_data,
  db_randomization_form %>%
    filter(!is.na(treatment_group)) %>%
    select(record_id, treatment_group)
)
# Everyone else is considered "observe clinically"
db_glycemic_data$treatment_group <- as.character(db_glycemic_data$treatment_group)
db_glycemic_data$treatment_group[is.na(db_glycemic_data$treatment_group)] <-
  "Observe Clincically"
db_glycemic_data$treatment_group <- factor(db_glycemic_data$treatment_group,
  levels = c("Observe Clincically", "Behavioral Intervention", "Care+ (Control)")
)
# If they are missing gly_timepoint they don't have glycemic data at that visit
db_glycemic_data <- db_glycemic_data %>% filter(!is.na(gly_timepoint))
```

# Table 1

```{r}
#| label: tbl-1
#| tbl-cap: Participant Characteristics
db_demographics %>%
  select(redcap_event_name, demo_hcl, demo_hcl_2, demo_age:demo_predict) %>%
  mutate(
    redcap_event_name = droplevels(redcap_event_name),
    demo_probability = as.numeric(demo_probability)
  ) %>%
  tbl_summary(
    missing_text = "Missing",
    type = list(
      demo_age ~ "continuous", demo_probability ~ "continuous"
    )
  ) %>%
  as_kable_extra()
```

# Glycemic data

## Groups by timepoint

```{r}
db_glycemic_data %>%
  group_by(gly_timepoint) %>%
  summarise(
    n = length(unique(record_id)),
    `n Observe Clincically` = sum(treatment_group == "Observe Clincically"),
    `n Behavioral Intervention` = sum(treatment_group == "Behavioral Intervention"),
    `n Care+ (Control)` = sum(treatment_group == "Care+ (Control)")
  ) %>%
  kable()
# Write for checking
write.csv(
  db_glycemic_data %>%
    select(
      record_id, redcap_event_name, redcap_survey_identifier,
      treatment_group, everything()
    ),
  file = "~/cleaned_glycemic_data.csv", row.names = F, na = ""
)
```

## HbA1c

### Plot

```{r}
v <- "gly_a1c"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
#| message: false
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor wear

### Plot

```{r}
v <- "gly_sensorwear"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor under 54

### Plot

```{r}
v <- "gly_sensor_u54"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

Because the majority of people did not have any time under 54 mg/dL, we used a hurdle model in which the first part of the model is the probability of a zero value and the second part is the probability of the non-zero values. 

```{r}
f <- as.formula(paste0(v, " ~ gly_timepoint*treatment_group"))
mod <- mixed_model(f,
  random = ~ 1 | record_id, data = df,
  family = hurdle.lognormal(), n_phis = 1, zi_fixed = ~1
)
```

#### Results

```{r}
kable(tidy(mod), digits = 3)
```

#### Means

These model means do not really make sense since we can't have negative values for this outcome, so this model likely needs additional work.

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor 54 - 69

### Plot

```{r}
v <- "gly_sensor_54_69"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

Because the majority of people did not have any time in the range 54 - 69 mg/dL, we used a hurdle model in which the first part of the model is the probability of a zero value and the second part is the probability of the non-zero values. 

```{r}
f <- as.formula(paste0(v, " ~ gly_timepoint*treatment_group"))
mod <- mixed_model(f,
  random = ~ 1 | record_id, data = df,
  family = hurdle.lognormal(), n_phis = 1, zi_fixed = ~1
)
```

#### Results

```{r}
kable(tidy(mod), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor 70 - 180

### Plot

```{r}
v <- "gly_sensor_70_180"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor 181 - 250

### Plot

```{r}
v <- "gly_sensor_181_250"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor > 250

### Plot

```{r}
v <- "gly_sensor_g250"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Mean sensor

### Plot

```{r}
v <- "gly_sensor_mean"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint * treatment_group)), digits = 3)
```

## Sensor SD

### Plot

```{r}
v <- "gly_sensor_sd"
df <- db_glycemic_data %>%
  select(record_id, gly_timepoint, treatment_group, all_of(v)) %>%
  drop_na()
p <- ggplot(df, aes(
  x = gly_timepoint, y = !!sym(v), group = record_id, color = treatment_group
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.title.x = element_blank())
ggplotly(p)
```

### Model

```{r}
f <- as.formula(paste0(v, "~gly_timepoint*treatment_group+(1|record_id)"))
mod <- lmer(f, data = db_glycemic_data)
```

#### Results

```{r}
kable(tidy(mod, effects = "fixed"), digits = 3)
```

#### Means

```{r}
kable(data.frame(emmeans(mod, ~ gly_timepoint | treatment_group)), digits = 3)
```
