---
title: "Analysis for Bone Health Manuscript"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
# Load necessary libraries.
library(tableone)
library(nortest)
library(emmeans)
library(knitr)
source('~/Documents/GitHub/Tim-and-Laura/tim_R_functions.R')
```

```{r echo=FALSE, include=FALSE}
# Define all the variables that don't require radius or tibia length adjustment.
no.adjust <- c("BMD1.3",
               "BMDLumbarTotal",
               "BMDNeck",
               "BMDTotalHip",
               "LengthR",
               "LengthT",
               "TBS.L1L4",
               "BSI_Crt_R4",
               "BSI_R4",
               "BSI_SubCrt_R4",
               "BSI_Trab_R4",
               "SSI_R33",
               "Imax_RADIUS",
               "Imin_RADIUS",
               "Imax_Imin_RADIUS",
               "Tot_A_R4",
               "Tot_Den_R4",
               "TRAB_A_R4",
               "Trab_den_R4",
               "CRT_A_R4",
               "CrtDen_R4",
               "CRT_Den_R33",
               "CRT_THK_R33",
               "EndoC_R33",
               "PeriC_R33",
               "BSI_Crt_T4",
               "BSI_SubCrt_T4",
               "BSI_T4",
               "BSI_Trab_T4",
               "SSI_T38",
               "SSI_T66",
               "Imax_TIB_38",
               "Imax_TIB_66",
               "Imin_TIB_38",
               "Imin_TIB_66",
               "Imax_Imin_TIB_38",
               "Imax_Imin_TIB_66",
               "Trab_A_T4",
               "Trab_Den_T4",
               "Crt_A_T4",
               "Crt_A_T38",
               "Crt_A_T66",
               "Crt_Den_T4",
               "Crt_Den_T38",
               "Crt_Den_T66",
               "Crt_thk_T38",
               "Crt_thk_T66",
               "EndoC_T38",
               "EndoC_T66",
               "PeriC_t38",
               "PeriC_t66",
               "Tot_A_T4",
               "Tot_Den_T4")
no.adjust <- no.adjust[order(no.adjust)]
# Define radius variables that need to be adjusted by length.
radius.adjust <- c("BSI_Crt_R4",
                   "BSI_R4",
                   "BSI_SubCrt_R4",
                   "BSI_Trab_R4",
                   "SSI_R33",
                   "Imax_RADIUS",
                   "Imin_RADIUS",
                   "Imax_Imin_RADIUS",
                   "Tot_A_R4",
                   "Tot_Den_R4",
                   "TRAB_A_R4",
                   "Trab_den_R4",
                   "CRT_A_R4",
                   "CrtDen_R4",
                   "CRT_Den_R33",
                   "CRT_THK_R33",
                   "EndoC_R33",
                   "PeriC_R33")
radius.adjust <- radius.adjust[order(radius.adjust)]
# Define tibia variables that need to be adjusted by length.
tibia.adjust <- c("BSI_Crt_T4",
                  "BSI_SubCrt_T4",
                  "BSI_T4",
                  "BSI_Trab_T4",
                  "SSI_T38",
                  "SSI_T66",
                  "Imax_TIB_38",
                  "Imax_TIB_66",
                  "Imin_TIB_38",
                  "Imin_TIB_66",
                  "Imax_Imin_TIB_38",
                  "Imax_Imin_TIB_66",
                  "Trab_A_T4",
                  "Trab_Den_T4",
                  "Crt_A_T4",
                  "Crt_A_T38",
                  "Crt_A_T66",
                  "Crt_Den_T4",
                  "Crt_Den_T38",
                  "Crt_Den_T66",
                  "Crt_thk_T38",
                  "Crt_thk_T66",
                  "EndoC_T38",
                  "EndoC_T66",
                  "PeriC_t38",
                  "PeriC_t66",
                  "Tot_A_T4",
                  "Tot_Den_T4")
tibia.adjust <- tibia.adjust[order(tibia.adjust)]
# Define all the variables of interest.
all.variables <- c(no.adjust,radius.adjust,tibia.adjust)
all.variables <- all.variables[order(all.variables)]
```
    
```{r echo=FALSE}
# Data management
# Check OS.
pathstart <- os.check()
filename <- paste(pathstart,"som/RI Biostatistics Core/Shared/Laura Kaci projects/Bone studies/pQCT/DATA/QTable1_BaselineData_9.17.18.csv",sep="")
# Read in.
pqct <- read.csv(filename,na.strings = c("No record","Incomplete"))
# Remove people without pQCT data and controls with A1c > 7%.
pqct <- pqct[-c(which(pqct$Diabetic == "No" & pqct$HbA1c >= 7.0)),]
pqct <- pqct[which(!is.na(pqct$LengthT) | !is.na(pqct$LengthR)),]
# Set 0 values in Endo_C_r4 data to NA, per emails with Janet and Viral.
pqct$Endo_C_r4[pqct$Endo_C_r4 == 0] <- NA
# Merge left and right Neck, Hip, HRT, Parental fractures. Convert variables 
# coded as 1 and 2 to TRUE/FALSE.
pqct$BMDNeck <- pmin(pqct$BMDRNeck,pqct$BMDLNeck,na.rm = TRUE)
pqct$BMDTotalHip <- pmin(pqct$BMDTotalLHip,pqct$BMDTotalRHip,na.rm = TRUE)
pqct$HRT_ever <- pmin(pqct$HRT,pqct$HRT_Curr,na.rm = TRUE)
pqct$HRT_ever[which(pqct$HRT_ever == 2)] <- 0
pqct$HRT_ever <- as.logical(pqct$HRT_ever)
pqct$HTN[which(pqct$HTN == 2)] <- 0
pqct$HTN <- as.logical(pqct$HTN)
pqct$Nephropathy[which(pqct$Nephropathy == 2)] <- 0
pqct$Nephropathy <- as.logical(pqct$Nephropathy)
pqct$Neuropathy[which(pqct$Neuropathy == 2)] <- 0
pqct$Neuropathy <- as.logical(pqct$Neuropathy)
pqct$Retinopathy[which(pqct$Retinopathy == 2)] <- 0
pqct$Retinopathy <- as.logical(pqct$Retinopathy)
pqct$Statin[which(pqct$Statin == 2)] <- 0
pqct$Statin <- as.logical(pqct$Statin)
pqct$HipParent <- as.logical(pmax(pqct$HipMother,pqct$HipFather))
pqct$FractureBone[which(pqct$FractureBone == 2)] <- 0
pqct$FractureBone <- as.logical(pqct$FractureBone)
# Add age at T1D diagnosis variable.
pqct$age.dx.group <- NA
pqct$age.dx.group[which(pqct$Age-pqct$T1DDuration >= 20)] <- "Greater than 20"
pqct$age.dx.group[which(pqct$Age-pqct$T1DDuration < 20)] <- "Less than 20"
```

```{r echo=FALSE,include=FALSE}
table.one.vars <- c("ACR","Age","BMI","calc","CalciumCorrected","CTX1","eGFR",
                    "FractureQty","HbA1c","IGF1","InsulinDose","LengthR",
                    "LengthT","Muscle_A","IMAT","Menopause_yrs",
                    "Osteocalcin.1","P1NP","PhosphorusSerum","PTH",
                    "T1DDuration","TtlFatMass","TtlFatPerc","TtlLeanMass",
                    "UCa_Cr","VitD","FractureBone","HipNeither","HipParent",
                    "HRT_ever","HTN","Statin","Nephropathy","Neuropathy",
                    "Retinopathy")
# Table 1 by T1D status.
# Test continuous variables for normality and find variables that require an exact 
# test.
nonnormal.vars <- NULL
exact.vars <- NULL
for (var in 1:length(table.one.vars)) {
  if (typeof(pqct[,table.one.vars[var]]) == "logical"){
    group1count <- sum(pqct[which(pqct$Diabetic == "Yes"),table.one.vars[var]],
                       na.rm = TRUE)
    group2count <- sum(pqct[which(pqct$Diabetic == "No"),table.one.vars[var]],
                       na.rm = TRUE)
    if (group1count < 5 | group2count < 5) {
      exact.vars <- c(exact.vars,table.one.vars[var])
    } else {
      exact.vars <- exact.vars
    }
  } else {
    ks <- lillie.test(pqct[,table.one.vars[var]])
    if (ks$p.value <= 0.05) {
      nonnormal.vars <- c(nonnormal.vars,table.one.vars[var])
    }
  }
}

table.one.t1d <- CreateTableOne(table.one.vars,strata = "Diabetic",pqct,
                            includeNA = FALSE)
table.one.t1d <- print(table.one.t1d,exact = exact.vars,
                       nonnormal = nonnormal.vars)
table.one.t1d <- as.data.frame(table.one.t1d)
# Table 1 by age at T1D diagnosis.
# Test continuous variables for normality and find variables that require an exact 
# test.
nonnormal.vars <- NULL
exact.vars <- NULL
for (var in 1:length(table.one.vars)) {
  if (typeof(pqct[,table.one.vars[var]]) == "logical"){
    group1count <- 
      sum(pqct[which(pqct$age.dx.group == "Greater than 20"),table.one.vars[var]],
          na.rm = TRUE)
    group2count <- 
      sum(pqct[which(pqct$age.dx.group == "Less than 20"),table.one.vars[var]],
          na.rm = TRUE)
    if (group1count < 5 | group2count < 5) {
      exact.vars <- c(exact.vars,table.one.vars[var])
    } else {
      exact.vars <- exact.vars
    }
  } else {
    ks <- lillie.test(pqct[,table.one.vars[var]])
    if (ks$p.value <= 0.05) {
      nonnormal.vars <- c(nonnormal.vars,table.one.vars[var])
    }
  }
}
table.one.age.group <- CreateTableOne(table.one.vars,strata = "age.dx.group",pqct,
                            includeNA = FALSE)
table.one.age.group <- print(table.one.age.group,nonnormal = nonnormal.vars,
                             exact = exact.vars)
table.one.age.group <- as.data.frame(table.one.age.group)
```

## Table 1a
Descriptive statistics by diabetic status. Normal variables were compared using an unpaired t-test, and non-normal variables were compared using the Kruskal-Wallis Rank Sum Test (test = "nonnorm"). Categorical variables were compared using a chi-squared test, and low-count variables were compared using Fisher's Exact Test (test = "exact").

```{r echo=FALSE}
kable(table.one.t1d)
```

## Table 1b
Descriptive statistics by age at diagnosis. Normal variables were compared using an unpaired t-test, and non-normal variables were compared using the Kruskal-Wallis Rank Sum Test (test = "nonnorm"). Categorical variables were compared using a chi-squared test, and low-count variables were compared using Fisher's Exact Test (test = "exact").

```{r echo=FALSE}
kable(table.one.age.group)
```

# Primary Aim: Differences by T1D Status
```{r echo=FALSE}
# Make a function for running lm by diabetic status, with alterable adjustment variables and dataset.
lm.by.t1d <- function(var, adj, dataframe) {
  adj <- paste(adj,collapse = " + ")
  formula <- paste(var,"~","Diabetic+",adj,sep = "")
  fit <- lm(as.formula(formula),dataframe)
  ls.means <- as.data.frame(emmeans(fit,"Diabetic"))
  ls.means <- ls.means[,1:3]
  ls.means$pval <- format.pval(summary(fit)$coef[2,4],eps = 0.001,digits = 3)
  return(ls.means)
}
```

```{r echo=FALSE}
# Run all variables that don't need to be length-adjusted.
t1d.age.results <- data.frame(matrix(nrow = (length(no.adjust)),ncol = 6))
colnames(t1d.age.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(no.adjust)) {
  table <- lm.by.t1d(no.adjust[v],"Age",pqct)
  t1d.age.results[v,1] <- no.adjust[v]
  t1d.age.results[v,2] <- round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.results[v,3] <- round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.results[v,4] <- round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.results[v,5] <- round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.results[v,6] <- table$pval[1]
}
```

## Table 2a
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age.
```{r echo=FALSE}
kable(t1d.age.results)
```

```{r echo=FALSE}
# Run all radius variables.
t1d.age.radius.results <- data.frame(matrix(nrow = (length(radius.adjust)),ncol = 6))
colnames(t1d.age.radius.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(radius.adjust)) {
  table <- lm.by.t1d(radius.adjust[v],c("Age","LengthR"),pqct)
  t1d.age.radius.results[v,1] <- 
    radius.adjust[v]
  t1d.age.radius.results[v,2] <- 
    round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.radius.results[v,3] <- 
    round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.radius.results[v,4] <- 
    round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.radius.results[v,5] <- 
    round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.radius.results[v,6] <- 
    table$pval[1]
}
```

## Table 2b
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age and radius length.
```{r echo=FALSE}
kable(t1d.age.radius.results)
```

```{r echo=FALSE}
# Run all tibia variables.
t1d.age.tibia.results <- data.frame(matrix(nrow = (length(tibia.adjust)),ncol = 6))
colnames(t1d.age.tibia.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(tibia.adjust)) {
  table <- lm.by.t1d(tibia.adjust[v],c("Age","LengthT"),pqct)
  t1d.age.tibia.results[v,1] <- 
    tibia.adjust[v]
  t1d.age.tibia.results[v,2] <- 
    round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.tibia.results[v,3] <- 
    round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.tibia.results[v,4] <- 
    round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.tibia.results[v,5] <- 
    round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.tibia.results[v,6] <- 
    table$pval[1]
}
```

## Table 2c
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age and tibia length.
```{r echo=FALSE}
kable(t1d.age.tibia.results)
```


```{r echo=FALSE}
# Run all variables that don't need to be length-adjusted.
t1d.age.men.results <- data.frame(matrix(nrow = (length(no.adjust)),ncol = 6))
colnames(t1d.age.men.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(no.adjust)) {
  table <- lm.by.t1d(no.adjust[v],c("Age","Menopause_yrs"),pqct)
  t1d.age.men.results[v,1] <- no.adjust[v]
  t1d.age.men.results[v,2] <- round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.men.results[v,3] <- round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.men.results[v,4] <- round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.men.results[v,5] <- round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.men.results[v,6] <- table$pval[1]
}
```

## Table 3a
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age and years of menopause.
```{r echo=FALSE}
kable(t1d.age.men.results)
```

```{r echo=FALSE}
# Run all radius variables.
t1d.age.men.radius.results <- data.frame(matrix(nrow = (length(radius.adjust)),ncol = 6))
colnames(t1d.age.men.radius.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(radius.adjust)) {
  table <- lm.by.t1d(radius.adjust[v],c("Age","Menopause_yrs","LengthR"),pqct)
  t1d.age.men.radius.results[v,1] <- 
    radius.adjust[v]
  t1d.age.men.radius.results[v,2] <- 
    round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.men.radius.results[v,3] <- 
    round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.men.radius.results[v,4] <- 
    round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.men.radius.results[v,5] <- 
    round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.men.radius.results[v,6] <- 
    table$pval[1]
}
```

## Table 3b
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age, years of menopause, and radius length.
```{r echo=FALSE}
kable(t1d.age.men.radius.results)
```

```{r echo=FALSE}
# Run all tibia variables.
t1d.age.men.tibia.results <- 
  data.frame(matrix(nrow = (length(tibia.adjust)),ncol = 6))
colnames(t1d.age.men.tibia.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(tibia.adjust)) {
  table <- lm.by.t1d(tibia.adjust[v],c("Age","Menopause_yrs","LengthT"),pqct)
  t1d.age.men.tibia.results[v,1] <- 
    tibia.adjust[v]
  t1d.age.men.tibia.results[v,2] <- 
    round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.men.tibia.results[v,3] <- 
    round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.men.tibia.results[v,4] <- 
    round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.men.tibia.results[v,5] <- 
    round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.men.tibia.results[v,6] <- 
    table$pval[1]
}
```

## Table 3c
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age, years of menopause, and tibia length.
```{r echo=FALSE}
kable(t1d.age.men.tibia.results)
```

```{r echo=FALSE}
# Run all variables that don't need to be length-adjusted, add BMI and remove 
# menopause years.
t1d.age.bmi.results <- data.frame(matrix(nrow = (length(no.adjust)),ncol = 6))
colnames(t1d.age.bmi.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(no.adjust)) {
  table <- lm.by.t1d(no.adjust[v],c("Age","BMI"),pqct)
  t1d.age.bmi.results[v,1] <- no.adjust[v]
  t1d.age.bmi.results[v,2] <- round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.bmi.results[v,3] <- round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.bmi.results[v,4] <- round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.bmi.results[v,5] <- round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.bmi.results[v,6] <- table$pval[1]
}
```

## Table 4a
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age and BMI.
```{r echo=FALSE}
kable(t1d.age.bmi.results)
```

```{r echo=FALSE}
# Run all radius variables, add BMI and remove menopause years.
t1d.age.bmi.radius.results <- data.frame(matrix(nrow = (length(radius.adjust)),ncol = 6))
colnames(t1d.age.bmi.radius.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(radius.adjust)) {
  table <- lm.by.t1d(radius.adjust[v],c("Age","BMI","LengthR"),pqct)
  t1d.age.bmi.radius.results[v,1] <- 
    radius.adjust[v]
  t1d.age.bmi.radius.results[v,2] <- 
    round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.bmi.radius.results[v,3] <- 
    round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.bmi.radius.results[v,4] <- 
    round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.bmi.radius.results[v,5] <- 
    round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.bmi.radius.results[v,6] <- 
    table$pval[1]
}
```

## Table 4b
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age, BMI, and radius length.
```{r echo=FALSE}
kable(t1d.age.bmi.radius.results)
```

```{r echo=FALSE}
# Run all tibia variables, add BMI and remove menopause years.
t1d.age.bmi.tibia.results <- data.frame(matrix(nrow = (length(tibia.adjust)),ncol = 6))
colnames(t1d.age.bmi.tibia.results) <- c("Variable","Mean (T1D+)","SE (T1D+)",
                           "Mean (T1D-)","SE (T1D-)","P Value")
for(v in 1:length(tibia.adjust)) {
  table <- lm.by.t1d(tibia.adjust[v],c("Age","BMI","LengthT"),pqct)
  t1d.age.bmi.tibia.results[v,1] <- 
    tibia.adjust[v]
  t1d.age.bmi.tibia.results[v,2] <- 
    round(table$emmean[which(table$Diabetic == "Yes")],3)
  t1d.age.bmi.tibia.results[v,3] <- 
    round(table$SE[which(table$Diabetic == "Yes")],3)
  t1d.age.bmi.tibia.results[v,4] <- 
    round(table$emmean[which(table$Diabetic == "No")],3)
  t1d.age.bmi.tibia.results[v,5] <- 
    round(table$SE[which(table$Diabetic == "No")],3)
  t1d.age.bmi.tibia.results[v,6] <- 
    table$pval[1]
}
```

## Table 4c
Results for Least Square means +/- standard error for variables by diabetic status adjusting for age, BMI, and tibia length.
```{r echo=FALSE}
kable(t1d.age.bmi.tibia.results)
```

# Secondary Aim: Differences by Age at T1D Diagnosis 
```{r echo=FALSE}
# Make a new function for running lm by diabetic status, with alterable 
# adjustment variables and dataset.
lm.by.age.dx <- function(var, adj, dataframe) {
  adj <- paste(adj,collapse = " + ")
  formula <- paste(var,"~","age.dx.group +",adj,sep = "")
  fit <- lm(as.formula(formula),dataframe)
  ls.means <- as.data.frame(emmeans(fit,"age.dx.group"))
  ls.means <- ls.means[,1:3]
  ls.means$pval <- format.pval(summary(fit)$coef[2,4],eps = 0.001,digits = 3)
  return(ls.means)
}
```

```{r echo=FALSE}
# Run all variables that don't need to be length-adjusted.
age.group.results <- data.frame(matrix(nrow = (length(no.adjust)),ncol = 6))
colnames(age.group.results) <- c("Variable","Mean (> 20 yrs)","SE (> 20 yrs)",
                           "Mean (< 20 yrs)","SE (< 20 yrs)","P Value")
for(v in 1:length(no.adjust)) {
  table <- lm.by.age.dx(no.adjust[v],"Age",pqct)
  age.group.results[v,1] <- 
    no.adjust[v]
  age.group.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.results[v,6] <- 
    table$pval[1]
}
```

## Table 5a
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age.
```{r echo=FALSE}
kable(age.group.results)
```

```{r echo=FALSE}
# Run all radius variables.
age.group.radius.results <- 
  data.frame(matrix(nrow = (length(radius.adjust)),ncol = 6))
colnames(age.group.radius.results) <- c("Variable","Mean (> 20 yrs)",
                                        "SE (> 20 yrs)","Mean (< 20 yrs)",
                                        "SE (< 20 yrs)","P Value")
for(v in 1:length(radius.adjust)) {
  table <- lm.by.age.dx(radius.adjust[v],c("Age","LengthR"),pqct)
  age.group.radius.results[v,1] <- 
    radius.adjust[v]
  age.group.radius.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.radius.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.radius.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.radius.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.radius.results[v,6] <- 
    table$pval[1]
}
```

## Table 5b
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age and radius length.
```{r echo=FALSE}
kable(age.group.radius.results)
```

```{r echo=FALSE,include=FALSE}
# Run all tibia variables.
age.group.tibia.results <- 
  data.frame(matrix(nrow = (length(tibia.adjust)),ncol = 6))
colnames(age.group.tibia.results) <- c("Variable","Mean (> 20 yrs)",
                                        "SE (> 20 yrs)","Mean (< 20 yrs)",
                                        "SE (< 20 yrs)","P Value")
for(v in 1:length(tibia.adjust)) {
  table <- lm.by.age.dx(tibia.adjust[v],c("Age","LengthT"),pqct)
  age.group.tibia.results[v,1] <- 
    tibia.adjust[v]
  age.group.tibia.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.tibia.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.tibia.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.tibia.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.tibia.results[v,6] <- 
    table$pval[1]
}
```

## Table 5c
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age and tibia length.
```{r echo=FALSE}
kable(age.group.tibia.results)
```

```{r echo=FALSE}
# Run all variables that don't need to be length-adjusted.
age.group.results <- data.frame(matrix(nrow = (length(no.adjust)),ncol = 6))
colnames(age.group.results) <- c("Variable","Mean (> 20 yrs)","SE (> 20 yrs)",
                           "Mean (< 20 yrs)","SE (< 20 yrs)","P Value")
for(v in 1:length(no.adjust)) {
  table <- lm.by.age.dx(no.adjust[v],c("Age","Menopause_yrs"),pqct)
  age.group.results[v,1] <- 
    no.adjust[v]
  age.group.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.results[v,6] <- 
    table$pval[1]
}
```

## Table 6a
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age and years of menopause.
```{r echo=FALSE}
kable(age.group.results)
```

```{r echo=FALSE}
# Run all radius variables.
age.group.radius.results <- 
  data.frame(matrix(nrow = (length(radius.adjust)),ncol = 6))
colnames(age.group.radius.results) <- c("Variable","Mean (> 20 yrs)",
                                        "SE (> 20 yrs)","Mean (< 20 yrs)",
                                        "SE (< 20 yrs)","P Value")
for(v in 1:length(radius.adjust)) {
  table <- lm.by.age.dx(radius.adjust[v],c("Age","Menopause_yrs","LengthR"),pqct)
  age.group.radius.results[v,1] <- 
    radius.adjust[v]
  age.group.radius.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.radius.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.radius.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.radius.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.radius.results[v,6] <- 
    table$pval[1]
}
```

## Table 6b
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age, years of menopause, and radius length.
```{r echo=FALSE}
kable(age.group.radius.results)
```

```{r echo=FALSE,include=FALSE}
# Run all tibia variables.
age.group.tibia.results <- 
  data.frame(matrix(nrow = (length(tibia.adjust)),ncol = 6))
colnames(age.group.tibia.results) <- c("Variable","Mean (> 20 yrs)",
                                        "SE (> 20 yrs)","Mean (< 20 yrs)",
                                        "SE (< 20 yrs)","P Value")
for(v in 1:length(tibia.adjust)) {
  table <- lm.by.age.dx(tibia.adjust[v],c("Age","Menopause_yrs","LengthT"),pqct)
  age.group.tibia.results[v,1] <- 
    tibia.adjust[v]
  age.group.tibia.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.tibia.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.tibia.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.tibia.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.tibia.results[v,6] <- 
    table$pval[1]
}
```

## Table 6c
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age, years of menopause, and tibia length.
```{r echo=FALSE}
kable(age.group.tibia.results)
```

```{r echo=FALSE}
# Run all variables that don't need to be length-adjusted.
age.group.bmi.results <- data.frame(matrix(nrow = (length(no.adjust)),ncol = 6))
colnames(age.group.bmi.results) <- c("Variable","Mean (> 20 yrs)","SE (> 20 yrs)",
                           "Mean (< 20 yrs)","SE (< 20 yrs)","P Value")
for(v in 1:length(no.adjust)) {
  table <- lm.by.age.dx(no.adjust[v],c("Age","BMI"),pqct)
  age.group.bmi.results[v,1] <- 
    no.adjust[v]
  age.group.bmi.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.bmi.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.bmi.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.bmi.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.bmi.results[v,6] <- 
    table$pval[1]
}
```

## Table 7a
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age and BMI.
```{r echo=FALSE}
kable(age.group.bmi.results)
```

```{r echo=FALSE}
# Run all radius variables.
age.group.bmi.radius.results <- 
  data.frame(matrix(nrow = (length(radius.adjust)),ncol = 6))
colnames(age.group.bmi.radius.results) <- c("Variable","Mean (> 20 yrs)",
                                        "SE (> 20 yrs)","Mean (< 20 yrs)",
                                        "SE (< 20 yrs)","P Value")
for(v in 1:length(radius.adjust)) {
  table <- lm.by.age.dx(radius.adjust[v],c("Age","BMI","LengthR"),pqct)
  age.group.bmi.radius.results[v,1] <- 
    radius.adjust[v]
  age.group.bmi.radius.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.bmi.radius.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.bmi.radius.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.bmi.radius.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.bmi.radius.results[v,6] <- 
    table$pval[1]
}
```

## Table 7b
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age, BMI, and radius length.
```{r echo=FALSE}
kable(age.group.bmi.radius.results)
```

```{r echo=FALSE}
# Run all tibia variables.
age.group.bmi.tibia.results <- 
  data.frame(matrix(nrow = (length(tibia.adjust)),ncol = 6))
colnames(age.group.bmi.tibia.results) <- c("Variable","Mean (> 20 yrs)",
                                        "SE (> 20 yrs)","Mean (< 20 yrs)",
                                        "SE (< 20 yrs)","P Value")
for(v in 1:length(tibia.adjust)) {
  table <- lm.by.age.dx(tibia.adjust[v],c("Age","BMI","LengthT"),pqct)
  age.group.bmi.tibia.results[v,1] <- 
    tibia.adjust[v]
  age.group.bmi.tibia.results[v,2] <- 
    round(table$emmean[which(table$age.dx.group == "Greater than 20")],3)
  age.group.bmi.tibia.results[v,3] <- 
    round(table$SE[which(table$age.dx.group == "Greater than 20")],3)
  age.group.bmi.tibia.results[v,4] <- 
    round(table$emmean[which(table$age.dx.group == "Less than 20")],3)
  age.group.bmi.tibia.results[v,5] <- 
    round(table$SE[which(table$age.dx.group == "Less than 20")],3)
  age.group.bmi.tibia.results[v,6] <- 
    table$pval[1]
}
```

## Table 7c
Results for Least Square means +/- standard error for variables by age at diagnosis adjusting for age, BMI, and tibia length.
```{r echo=FALSE}
kable(age.group.bmi.tibia.results)
```