---
title: "QCT Pilot Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE,include=FALSE}
library(knitr)
library(pander)
library(psych)
library(Hmisc)
library(corrplot)
source('~/Documents/GitHub/Tim-and-Laura/tim_R_functions.R')
panderOptions("table.continues","")
```

```{r echo=FALSE}
# Data management
# Check OS.
pathstart <- os.check()
filename1 <- paste(pathstart,"som/PEDS/RI Biostatistics Core/Shared/Laura Tim projects/Viral Shah/QCT Pilot Analysis/Data_Cleaned/T1D Femoral Neck Structure.csv",sep="")
filename2 <- paste(pathstart,"som/PEDS/RI Biostatistics Core/Shared/Laura Tim projects/Viral Shah/QCT Pilot Analysis/Data_Cleaned/T1D Pilot QCT Data.csv",sep="")
# Read in.
qct1 <- read.csv(filename1)
qct2 <- read.csv(filename2)
qct <- merge(qct1,qct2,by = "StudyID")
# Read in pQCT data.
filename <- paste(pathstart,"som/PEDS/RI Biostatistics Core/Shared/Laura Kaci projects/Bone studies/pQCT/DATA/QTable1_BaselineData_9.17.18.csv",sep="")
# Read in.
pqct <- read.csv(filename,na.strings = c("No record","Incomplete"))
# Remove people without pQCT data and controls with A1c > 7%.
pqct <- pqct[-c(which(pqct$Diabetic == "No" & pqct$HbA1c >= 7.0)),]
pqct <- pqct[which(!is.na(pqct$LengthT) | !is.na(pqct$LengthR)),]
# Set 0 values in Endo_C_r4 data to NA, per emails with Janet and Viral.
pqct$Endo_C_r4[pqct$Endo_C_r4 == 0] <- NA
# Merge left and right Neck, Hip, HRT, Parental fractures. Convert variables 
# coded as 1 and 2 to TRUE/FALSE.
pqct$BMDNeck <- pmin(pqct$BMDRNeck,pqct$BMDLNeck,na.rm = TRUE)
pqct$BMDTotalHip <- pmin(pqct$BMDTotalLHip,pqct$BMDTotalRHip,na.rm = TRUE)
pqct$HRT_ever <- pmin(pqct$HRT,pqct$HRT_Curr,na.rm = TRUE)
pqct$HRT_ever[which(pqct$HRT_ever == 2)] <- 0
pqct$HRT_ever <- as.logical(pqct$HRT_ever)
pqct$HTN[which(pqct$HTN == 2)] <- 0
pqct$HTN <- as.logical(pqct$HTN)
pqct$Nephropathy[which(pqct$Nephropathy == 2)] <- 0
pqct$Nephropathy <- as.logical(pqct$Nephropathy)
pqct$Neuropathy[which(pqct$Neuropathy == 2)] <- 0
pqct$Neuropathy <- as.logical(pqct$Neuropathy)
pqct$Retinopathy[which(pqct$Retinopathy == 2)] <- 0
pqct$Retinopathy <- as.logical(pqct$Retinopathy)
pqct$Statin[which(pqct$Statin == 2)] <- 0
pqct$Statin <- as.logical(pqct$Statin)
pqct$HipParent <- as.logical(pmax(pqct$HipMother,pqct$HipFather))
pqct$FractureBone[which(pqct$FractureBone == 2)] <- 0
pqct$FractureBone <- as.logical(pqct$FractureBone)
# Add age at T1D diagnosis variable.
pqct$age.dx.group <- NA
pqct$age.dx.group[which(pqct$Age-pqct$T1DDuration >= 20)] <- "Greater than 20"
pqct$age.dx.group[which(pqct$Age-pqct$T1DDuration < 20)] <- "Less than 20"
# Remove _F from study IDs
pqct$StudyID <- sub("_F","",pqct$StudyID)
# Merge pQCT and QCT data.
dat <- merge(qct,pqct,by = "StudyID")
dat$avglumbarvBMD <- rowMeans(dat[,c("L1BMD","L2BMD")])
colnames(dat) <- tolower(colnames(dat))
```

```{r echo=FALSE}
# Define pQCT thickness variables (adjust case for easier matching)
r <- c("crt_thk_r33","Crt_thk_T38","CRT_THK_T66",
       "Crt_A_R4","Crt_A_R33","CRT_A_t4","CRT_A_T38","CRT_A_T66")
r <- tolower(r)
# Define QCT variables.
c <- c("average.cortical.thickness","cortical.area")
c <- tolower(c)
# Correlate.
thick.area <- as.matrix(cbind(dat[,r],dat[,c]))
thick.area.corr <- corr.test(thick.area[,c],thick.area[,r],
                             adjust = "none")

# Define pQCT density variables (adjust case for easier matching)
r <- c("crt_den_t4","CRT_den_R33","Crt_Den_T38")
r <- tolower(r)
# Define QCT variables.
c <- c("average.cortical.BMD","Neck.cortical.bmd","Troch.cortical.BMD",
       "Intertroch.cortical.bmd","Total.hip.Cortical.BMD")
c <- tolower(c)
# Correlate.
dense.vbmd <- as.matrix(cbind(dat[,r],dat[,c]))
dense.vbmd.corr <- corr.test(dense.vbmd[,c],dense.vbmd[,r],
                             adjust = "none")

# Define pQCT density variables (adjust case for easier matching)
r <- c("tot_den_R4","Tot_Den_T4")
r <- tolower(r)
# Define QCT variables.
c <- c("Total.BMD","l1bmd","l2bmd","neck.integral.bmd","avglumbarvbmd",
       "troch.integral.bmd","intertroch.integral.bmd",
       "total.hip.Integral.bmd")
c <- tolower(c)
# Correlate.
total.dense.vbmd <- as.matrix(cbind(dat[,r],dat[,c]))
total.dense.vbmd.corr <- corr.test(total.dense.vbmd[,c],total.dense.vbmd[,r],
                                   adjust = "none")
```

Table 1a: Pearson correlations of pQCT and QCT measures with cortical area and thickness.
```{r echo=FALSE}
# Bold r values with significant p values. Round to two digits.
bold <- which(thick.area.corr$p < 0.05,arr.ind = T)
thick.area.print.r <- round(thick.area.corr$r,2)
thick.area.print.r[bold] <- paste("**",thick.area.print.r[bold],"**",sep = "")
kable(thick.area.print.r)
```
\pagebreak

Table 1b: Unadjusted P-values for correlations of pQCT and QCT measures with cortical area and thickness.
```{r echo=FALSE}
# Bold r values with significant p values. Round to two digits.
thick.area.print.p <- round(thick.area.corr$p,2)
thick.area.print.p[bold] <- paste("**",thick.area.print.p[bold],"**",sep = "")
kable(thick.area.print.p)
```
\pagebreak

Table 2a: Pearson correlations of pQCT and QCT measures cortical vBMD.
```{r echo=FALSE}
# Bold r values with significant p values. Round to two digits.
bold <- which(dense.vbmd.corr$p < 0.05,arr.ind = T)
dense.vbmd.print.r <- round(dense.vbmd.corr$r,2)
dense.vbmd.print.r[bold] <- paste("**",dense.vbmd.print.r[bold],"**",sep = "")
kable(dense.vbmd.print.r)
```
\pagebreak

Table 2b: Unadjusted P-values for correlations of pQCT and QCT measures cortical vBMD
```{r echo=FALSE}
# Bold r values with significant p values. Round to two digits.
dense.vbmd.print.p <- round(dense.vbmd.corr$p,2)
dense.vbmd.print.p[bold] <- paste("**",dense.vbmd.print.p[bold],"**",sep = "")
kable(dense.vbmd.print.p)
```
\pagebreak

Table 3a: Pearson correlations of total density (vBMD) measures for pQCT and QCT.
```{r echo=FALSE}
# Bold r values with significant p values. Round to two digits.
bold <- which(total.dense.vbmd.corr$p < 0.05,arr.ind = T)
total.dense.vbmd.print.r <- round(total.dense.vbmd.corr$r,2)
total.dense.vbmd.print.r[bold] <- paste("**",total.dense.vbmd.print.r[bold],"**",sep = "")
kable(total.dense.vbmd.print.r)
```
\pagebreak

Table 3b: Unadjusted P-values for correlations of total density (vBMD) measures for pQCT and QCT
```{r echo=FALSE}
# Bold r values with significant p values. Round to two digits.
total.dense.vbmd.print.p <- round(total.dense.vbmd.corr$p,2)
total.dense.vbmd.print.p[bold] <- paste("**",total.dense.vbmd.print.p[bold],"**",sep = "")
kable(total.dense.vbmd.print.p)
```
\pagebreak

Figure 1: Correlogram of pQCT and QCT measures with cortical area and thickness.

```{r echo=FALSE}
corrplot(thick.area.corr$r,cl.ratio = 0.5,cl.pos = "b")
```
\pagebreak

Figure 2: Correlogram of pQCT and QCT measures cortical vBMD.

```{r echo=FALSE}
corrplot(dense.vbmd.corr$r,cl.ratio = 0.3)
```
\pagebreak

Figure 3: Correlogram of total density (vBMD) measures for pQCT and QCT.

```{r echo=FALSE}
corrplot(total.dense.vbmd.corr$r,cl.ratio = 0.8)
```