---
title: "Insulitis manuscript"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:   
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
# required libraries
library(purrr)
library(tableone)
library(dplyr)
library(plyr)
library(zoo)
library(nlme)
library(lme4)
library(emmeans)
library(knitr)
library(car)
library(tidyr)
library(stringr)

```

```{r, include=F}
# read in data

bT_interactions_new <- read.csv("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Rachel Friedman/early stage b-T interactions new.csv")

bT_interactions_old <- read.csv("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Rachel Friedman/early stage b-T interactions old.csv")

islet_wise_new <- read.csv("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Rachel Friedman/islet wise insulitis progression new.csv")

islet_wise_old <- read.csv("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Rachel Friedman/islet wise insulitis progression old.csv")


```

# Methods

A Wilcoxon test was used to compare the a-linked b-T and non a-linked b-T.

To evaluate whether an exponential or linear model was a better fit to the T/b cell ratios and T-cell density, both types of models were fit. There are two metrics that can be used to compare the model fit:

1) Adjusted R2: higher is better

2) AIC: smaller is better

# Results

## Item 2(B) - early stage b-T interactions - NEW DATA

```{r, include=TRUE, results="hide"}
t <- CreateTableOne(data=bT_interactions_new, strata = "linked_nonlinked", vars = "result")
t <- print(t, nonnormal = "result")
```

```{r}
kable(t)
```

## Item 2(B) - early stage b-T interactions - OLD DATA

```{r, include=TRUE, results="hide"}
t <- CreateTableOne(data=bT_interactions_old, strata = "linked_nonlinked", vars = "result")
t <- print(t, nonnormal = "result")
```

```{r}
kable(t)
```

## Comparing exponential and linear fits - NEW DATA

### T/b cell ratios

```{r}
mod_lin <- lm(islet_wise_new$T..β..cell.ratios ~ islet_wise_new$Islet.rank)
summary(mod_lin)
AIC(mod_lin)

mod_exp <- lm(log(islet_wise_new$T..β..cell.ratios) ~ islet_wise_new$Islet.rank)
summary(mod_exp)
AIC(mod_exp)
```

The adjusted R2 for the linear model was `r summary(mod_lin)$adj.r.squared`. The adjusted R2 for the exponential model was `r summary(mod_exp)$adj.r.squared`. The AIC for the linear model was `r AIC(mod_lin)` and the AIC for the exponential model was `r AIC(mod_exp)`. The exponential model is a better fitting model.

### T-cell density

```{r}
mod_lin <- lm(islet_wise_new$T..cell.density..mm2 ~ islet_wise_new$Islet.rank)
summary(mod_lin)
AIC(mod_lin)

mod_exp <- lm(log(islet_wise_new$T..cell.density..mm2) ~ islet_wise_new$Islet.rank)
summary(mod_exp)
AIC(mod_exp)
```

The adjusted R2 for the linear model was `r summary(mod_lin)$adj.r.squared`. The adjusted R2 for the exponential model was `r summary(mod_exp)$adj.r.squared`. The AIC for the linear model was `r AIC(mod_lin)` and the AIC for the exponential model was `r AIC(mod_exp)`. The exponential model is a better fitting model.

## Comparing exponential and linear fits - OLD DATA

### T/b cell ratios

```{r}
mod_lin <- lm(islet_wise_old$T..β..cell.ratios ~ islet_wise_old$Islet.rank)
summary(mod_lin)
AIC(mod_lin)

mod_exp <- lm(log(islet_wise_old$T..β..cell.ratios) ~ islet_wise_old$Islet.rank)
summary(mod_exp)
AIC(mod_exp)
```

The adjusted R2 for the linear model was `r summary(mod_lin)$adj.r.squared`. The adjusted R2 for the exponential model was `r summary(mod_exp)$adj.r.squared`. The AIC for the linear model was `r AIC(mod_lin)` and the AIC for the exponential model was `r AIC(mod_exp)`. The exponential model is a better fitting model.

### T-cell density

```{r}
mod_lin <- lm(islet_wise_old$T..cell.density..mm2 ~ islet_wise_old$Islet.rank)
summary(mod_lin)
AIC(mod_lin)

mod_exp <- lm(log(islet_wise_old$T..cell.density..mm2) ~ islet_wise_old$Islet.rank)
summary(mod_exp)
AIC(mod_exp)
```

The adjusted R2 for the linear model was `r summary(mod_lin)$adj.r.squared`. The adjusted R2 for the exponential model was `r summary(mod_exp)$adj.r.squared`. The AIC for the linear model was `r AIC(mod_lin)` and the AIC for the exponential model was `r AIC(mod_exp)`. The exponential model is a better fitting model.