---
title: "A1c trends at BDC"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(knitr)
library(readxl)
library(tableone)
library(dplyr)
library(tidyr)
library(nlme)
library(ggplot2)
library(performance)
library(car)
library(lme4)

patientdata <- read_excel("T:\\Todd Alonso\\A1c trends at BDC\\A1cTrendsAtBDC_PatientLevel_03162021_stupdates_mas072321.xlsx",na=c("","NA","NULL"))
patientdata <- unique(patientdata)
# no dups

# fix onset date for 1 patient
patientdata[patientdata$EPICMRN==1412180,]$OnsetDate <- as.Date("2008-11-08")

visitdata <- read_excel("T:\\Todd Alonso\\A1c trends at BDC\\A1cTrendsAtBDC_VisitLevel_03162021_stupdates.xlsx",na=c("","NA","NULL"))
visitdata <- unique(visitdata)
visitdata$FirstName <- NULL
visitdata$LastName <- NULL
visitdata$DOB <- NULL
visitdata$OnsetDate <- NULL
# no dups

# create one dataset using only those in both dataframes
alldata <- merge(visitdata,patientdata,by="EPICMRN",all.x = FALSE,all=FALSE)

# create variable for visit year
alldata$visityear <- format(alldata$VisitDate,format = "%Y")

# exclude visityear 2011 and 2012
alldata <- alldata[!alldata$visityear %in% c(2011,2012),]

# exclude visityear 2020
alldata <- alldata[!alldata$visityear==2020,]

# create simpler race/ethnicity variable
alldata$Race_Ethnicity_combined <- ifelse(alldata$Race_Ethnicity=="Hispanic","Hispanic",
                                          ifelse(alldata$Race_Ethnicity=="Non-Hispanic White","Non-Hispanic White","Other"))
alldata$Race_Ethnicity_combined <- as.factor(alldata$Race_Ethnicity_combined)
alldata$Race_Ethnicity_combined <- relevel(alldata$Race_Ethnicity_combined,ref="Non-Hispanic White")

# fix language
alldata$Language <- ifelse(alldata$Language=="ENGLISH","English",alldata$Language)
alldata$Language <- ifelse(alldata$Language=="SPANISH","Spanish",alldata$Language)
alldata$Language <- ifelse(alldata$Language=="Other - please contact MI Department at x79800","Other",alldata$Language)
alldata$Language_combined <- ifelse(alldata$Language=="English","English",
                                    ifelse(alldata$Language=="Spanish","Spanish","Other"))

# create variable with mean A1c per year for each patient
a1c <- alldata[,c("EPICMRN","visityear","A1c_Value")]
a1c_year <- a1c %>% group_by(EPICMRN,visityear) %>% mutate(A1c_Mean = mean(A1c_Value))  %>% filter(row_number()==1)
a1c_year <- a1c_year[,c("EPICMRN","visityear","A1c_Mean")]
alldata <- merge(alldata,a1c_year,by=c("EPICMRN","visityear"),all.x = T,all.y=T)

# categorize the mean A1c value for the year as <6%, 6-12%, or 12%
alldata$A1c_Mean_cat <- ifelse(is.na(alldata$A1c_Mean),NA,
                               ifelse(alldata$A1c_Mean<6,"<6%",
                               ifelse(alldata$A1c_Mean>12,">12%","6-12%")))

# categorize the mean A1c value for the year as <6%, 6-12%, or 12%
alldata$A1c_Mean_cat2 <- ifelse(is.na(alldata$A1c_Mean),NA,
                               ifelse(alldata$A1c_Mean<7,"<7%",
                               ifelse(alldata$A1c_Mean>=7 & alldata$A1c_Mean<9,"7 - <9%",
                                      ifelse(alldata$A1c_Mean>=9 & alldata$A1c_Mean<12,"9 - <12%",">=12%"))))

# categorical variable for mean A1c >=12
alldata$A1c_Mean_cat12 <- ifelse(alldata$A1c_Mean_cat==">12%",1,0)

# count number of visits per year
visittemp <- alldata[,c("EPICMRN","visityear")]
visitcount <- visittemp %>% group_by(EPICMRN,visityear) %>% summarise(n = n()) 
colnames(visitcount) <- c("EPICMRN","visityear","visitcount")
alldata <- merge(alldata,visitcount,by=c("EPICMRN","visityear"),all.x=T,all.y=T)

# find first visit year
firstvisit <- alldata[,c("EPICMRN","visityear")]
firstvisit <- firstvisit %>% group_by(EPICMRN) %>% filter(row_number()==1)
colnames(firstvisit) <- c("EPICMRN","firstvisit")
#alldata <- merge(alldata,firstvisit,by="EPICMRN",all.x = T,all.y = T)

# collapse insurance
alldata$InsuranceCategory_combined <- ifelse(alldata$InsuranceCategory=="Medicaid","Public",
                                             ifelse(alldata$InsuranceCategory %in% c("Private","Military Plans"),"Private","Unknown"))

# create new variable for pump/no pump
alldata$pump <- as.factor(ifelse(is.na(alldata$InsulinRegimen),NA,ifelse(alldata$InsulinRegimen=="Insulin Pump",1,0)))

# create categorical age variable
alldata$a1c_agecat <- ifelse(is.na(alldata$Age_A1cResultDate),NA,
                             ifelse(alldata$Age_A1cResultDate<6,"<6 years",
                                    ifelse(alldata$Age_A1cResultDate>=6 & alldata$Age_A1cResultDate<12,"6 - <12 years",
                                           ifelse(alldata$Age_A1cResultDate>=12 & alldata$Age_A1cResultDate<18,"12 - <18 years",
                                                  ifelse(alldata$Age_A1cResultDate>= 18 & alldata$Age_A1cResultDate<22,"18 - <22 years",">=22 years")))))
  
# sort by EPICMRN and visit data
alldata <- alldata[with(alldata,order(EPICMRN,VisitDate)),]

# create dataset for each year with patients who had visits in that year
#dat2011 <- alldata[alldata$visityear==2011,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
#dat2012 <- alldata[alldata$visityear==2012,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2013 <- alldata[alldata$visityear==2013,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2014 <- alldata[alldata$visityear==2014,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2015 <- alldata[alldata$visityear==2015,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2016 <- alldata[alldata$visityear==2016,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2017 <- alldata[alldata$visityear==2017,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2018 <- alldata[alldata$visityear==2018,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2019 <- alldata[alldata$visityear==2019,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
#dat2020 <- alldata[alldata$visityear==2020,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
#demo_by_year <- rbind(dat2013,dat2014,dat2015,dat2016,dat2017,dat2018,dat2019,dat2020)
demo_by_year <- rbind(dat2013,dat2014,dat2015,dat2016,dat2017,dat2018,dat2019)

# make a new df that will have every year for every patient
# create variable for whether the patient was seen at least once in the last year, starting at their first visit
# prior to first visit, will be NA
alldata_expand <- alldata %>% ungroup() %>% complete(EPICMRN, visityear)
alldata_expand <- merge(alldata_expand,firstvisit,by="EPICMRN",all.x=T,all.y=T)
alldata_expand$visit_in_year <- NA
alldata_expand$visit_in_year <- ifelse(alldata_expand$visityear<alldata_expand$firstvisit,NA,
                         ifelse(alldata_expand$visitcount %in% c(0,NA),0,1))
alldata_expand$visit_in_year <- as.factor(alldata_expand$visit_in_year)

# create variable for having a visit in the past year
x <- alldata_expand[,c("EPICMRN","visityear","visit_in_year")]
x <- x %>% group_by(EPICMRN,visityear) %>% filter(row_number()==1)
x$visit_in_year <- as.numeric(as.character(x$visit_in_year))
x$visityear <- as.numeric(x$visityear)
x <- x %>% ungroup() %>% group_by(EPICMRN) %>% mutate(visit_last_year=lag(visit_in_year, order_by = visityear))
x$visit_last_year <- as.factor(x$visit_last_year)
x$visit_in_year <- NULL
alldata_expand <- merge(alldata_expand,x,by=c("EPICMRN","visityear"),all.x = T,all.y = T)

# create table of patients having at least one visit in each year
dat2013_expand <- alldata_expand[alldata_expand$visityear==2013,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2014_expand <- alldata_expand[alldata_expand$visityear==2014,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2015_expand <- alldata_expand[alldata_expand$visityear==2015,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2016_expand <- alldata_expand[alldata_expand$visityear==2016,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2017_expand <- alldata_expand[alldata_expand$visityear==2017,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2018_expand <- alldata_expand[alldata_expand$visityear==2018,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
dat2019_expand <- alldata_expand[alldata_expand$visityear==2019,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
#dat2020_expand <- alldata_expand[alldata_expand$visityear==2020,] %>% group_by(EPICMRN) %>% filter(row_number()==1)
#demo_by_year_expand <- rbind(dat2013_expand,dat2014_expand,dat2015_expand,dat2016_expand,dat2017_expand,dat2018_expand,
#                             dat2019_expand,dat2020_expand)
demo_by_year_expand <- rbind(dat2013_expand,dat2014_expand,dat2015_expand,dat2016_expand,dat2017_expand,dat2018_expand,
                             dat2019_expand)
demo_by_year_expand <- demo_by_year_expand %>% arrange(EPICMRN,visityear)

# box plot of mean A1c by age
p <- ggplot(demo_by_year, aes(x=a1c_agecat, y=A1c_Mean)) +  geom_boxplot()

# create table of demographics and clinical characteristics by visit year
demovars <- c("Gender","Race_Ethnicity_combined","Language_combined","InsuranceCategory_combined",
              "DiabetesDuration_A1cResultDate","Age_A1cResultDate","a1c_agecat","DKAAtOnset","CGM","pump","A1c_Value","A1c_Mean","A1c_Mean_cat",
              "visitcount")
t1 <- CreateTableOne(data=demo_by_year,vars=demovars,strata="visityear")
t1 <- print(t1)

# table of whether patients had a visit in each year, following their first visit
t2 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$visityear!=2013 & 
                                                demo_by_year_expand$visityear != 2020,],vars="visit_in_year",strata="visityear")
t2 <- print(t2)

# table of patients with mean A1c <6% in a given year, and whether they had a visit the prior year
t3 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat=="<6%" 
                                              & demo_by_year_expand$visityear!=2013,],vars="visit_last_year",strata="visityear")
t3 <- print(t3)

# table of patients with mean A1c 6-12% in a given year, and whether they had a visit the prior year
t4 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat=="6-12%"
                                              & demo_by_year_expand$visityear!=2013,],vars="visit_last_year",strata="visityear")
t4 <- print(t4)

# table of patients with mean A1c >12% in a given year, and whether they had a visit the prior year
t5 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat==">12%"
                                              & demo_by_year_expand$visityear!=2013,],vars="visit_last_year",strata="visityear")
t5 <- print(t5)

# table of patients with mean A1c <7% in a given year, and whether they had a visit the prior year
t6 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat2=="<7%"
                                              & demo_by_year_expand$visityear!=2013,],vars="visit_last_year",strata="visityear")
t6 <- print(t6)

# table of patients with mean A1c 7-9% in a given year, and whether they had a visit the prior year
t7 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat2=="7 - <9%"
                                              & demo_by_year_expand$visityear!=2013,],vars="visit_last_year",strata="visityear")
t7 <- print(t7)

# table of patients with mean A1c 9-12% in a given year, and whether they had a visit the prior year
t8 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat2=="9 - <12%"
                                              & demo_by_year_expand$visityear!=2013,],vars="visit_last_year",strata="visityear")
t8 <- print(t8)

# table of patients with mean A1c >12% in a given year, and whether they had a visit the prior year
t9 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat2==">=12%"
                                              & demo_by_year_expand$visityear!=2013,],vars="visit_last_year",strata="visityear")
t9 <- print(t9)

# table of number of visits per year in those with mean A1c >12%
t10 <- CreateTableOne(data=demo_by_year_expand[demo_by_year_expand$A1c_Mean_cat2==">=12%"
                                              & demo_by_year_expand$visityear!=2013,],vars="visitcount",strata="visityear")
t10 <- print(t10)


# mixed model of mean A1c by year
mod <- lme(A1c_Mean ~ visityear + Race_Ethnicity_combined + Language_combined + InsuranceCategory_combined + DiabetesDuration_A1cResultDate +
             a1c_agecat + DKAAtOnset + CGM + pump + visitcount + visitcount*visityear,
           random=~1|EPICMRN,data = demo_by_year_expand[demo_by_year_expand$visityear!=2013 & 
                                                demo_by_year_expand$visityear != 2020,],na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

# use emmeans 
#amtime_mod_means <- emmeans::emmeans(amtime_mod,"gyl_timepoint.factor")
#amtime_mod_pairs <-  pairs(amtime_mod_means,adjust="tukey")

# what predicts A1c>12 in year 2014
log2014 <- glm(A1c_Mean_cat12 ~ Race_Ethnicity_combined + Language_combined + InsuranceCategory_combined + DiabetesDuration_A1cResultDate +
             a1c_agecat + DKAAtOnset + CGM + pump + visitcount,data = demo_by_year_expand[demo_by_year_expand$visityear==2014,],
             family = "binomial")
log2014_anova <- Anova(log2014,type="III")

log2015 <- glm(A1c_Mean_cat12 ~ Race_Ethnicity_combined + Language_combined + InsuranceCategory_combined + DiabetesDuration_A1cResultDate +
             a1c_agecat + DKAAtOnset + CGM + pump + visitcount,data = demo_by_year_expand[demo_by_year_expand$visityear==2015,],
             family = "binomial")
log2015_anova <- Anova(log2015,type="III")

log2016 <- glm(A1c_Mean_cat12 ~ Race_Ethnicity_combined + Language_combined + InsuranceCategory_combined + DiabetesDuration_A1cResultDate +
             a1c_agecat + DKAAtOnset + CGM + pump + visitcount,data = demo_by_year_expand[demo_by_year_expand$visityear==2016,],
             family = "binomial")
log2016_anova <- Anova(log2016,type="III")

log2017 <- glm(A1c_Mean_cat12 ~ Race_Ethnicity_combined + Language_combined + InsuranceCategory_combined + DiabetesDuration_A1cResultDate +
             a1c_agecat + DKAAtOnset + CGM + pump + visitcount,data = demo_by_year_expand[demo_by_year_expand$visityear==2017,],
             family = "binomial")
log2017_anova <- Anova(log2017,type="III")

log2018 <- glm(A1c_Mean_cat12 ~ Race_Ethnicity_combined + Language_combined + InsuranceCategory_combined + DiabetesDuration_A1cResultDate +
             a1c_agecat + DKAAtOnset + CGM + pump + visitcount,data = demo_by_year_expand[demo_by_year_expand$visityear==2018,],
             family = "binomial")
log2018_anova <- Anova(log2018,type="III")

log2019 <- glm(A1c_Mean_cat12 ~ Race_Ethnicity_combined + Language_combined + InsuranceCategory_combined + DiabetesDuration_A1cResultDate +
             a1c_agecat + DKAAtOnset + CGM + pump + visitcount,data = demo_by_year_expand[demo_by_year_expand$visityear==2019,],
             family = "binomial")
log2019_anova <- Anova(log2019,type="III")

# could do a GLMM with interaction between each term and year
#glmermod <- glmer(A1c_Mean_cat12 ~ Race_Ethnicity_combined|visityear + Language_combined|visityear + InsuranceCategory_combined|visityear 
#                  + DiabetesDuration_A1cResultDate|visityear + a1c_agecat|visityear + DKAAtOnset|visityear + CGM|visityear + pump|visityear 
#                  + visitcount|visityear  + (1 | EPICMRN),data = demo_by_year_expand,
#             family = binomial,verbose = T)
#glmermod_anova <- Anova(glmermod,type="III")

```

# Background

The purpose of this analysis is to examine factors that may be related to changes in clinic HbA1c values between 2013-2020 at the BDC.

# Methods

The following data manipulation steps were performed:

- Visit years 2011 and 2012 were excluded.  Should 2020 be excluded since it's a partial year?

- Language was categorized as English, Spanish, and other.

- Race/ethnicity was categorized as Hispanic, Non-Hispanic White, and other.

- Insurance was categorized as private (private insurance + military), public, or unknown.

- The mean A1c per year was calculated for each patient, and was categorized as <6%, 6-12%, and >12%.

- A variable was created to indicate whether each patient had a visit during the calendar year, beginning with the year of their first visit.

Descriptive statistics reported are mean/SD and frequency/percentage.  Years and patient groups were compared using linear models for continuous variables and chi-square tests for categorical variables.  A mixed-effects model was used to model factors associated with HbA1c over time.  Logistic regression was used to examine variables associated with having a mean HbA1c >12%.

Changes since original report:

- Updated dataset

- Excluded 2020

- Coded insulin regmimen as pump/no-pump

- Categorized age as: <6, 6-<12, 12-<18, 18-<22, >=22 and used that variable in the mixed model.

- Added second variable categorizing A1c as <7%, 7 - <9%, 9 - <12%, >=12%

- Added interaction between year and visit count to mixed model

- Added logistic models examining factors associated with having A1c>12%

# Results

## Demographic and clinical characteristics of patients having a visit in each year.  The first visit in each year was included, except for the variable A1c_Mean, which represents the mean A1c value for all visits during that year, and the variable visitcount, which represents the number of visits per  year.

```{r echo=FALSE}
kable(t1)
```
<br>

## Percentage of patients with at least one visit in each year, starting with the year of their first visit.

Note: the N reported in the first row is the total number of patients across all years, not the number of patients who had their first visit by that year.  Patients who had not yet had their first visit by that year have missing values for the variable indicating whether they had a visit in that year, so the denominator for the percentages is the number of patients who had their first visit by that year.

```{r echo=FALSE}
kable(t2)
```
<br>

## Percentage of patients with mean A1c <6% in a given year who had a visit in the prior year.

Note: the N reported in the first row is the total number of patients with mean A1c <6% in each year. Patients who had not yet had their first visit by the year prior have a missing value for the variable indicating whether they had a visit in the prior year, so the denominator for the percentages is the number of patients with mean A1c <6% and who had their first visit by the prior year.

```{r echo=FALSE}
kable(t3)
```
<br>

## Percentage of patients with mean A1c 6-12% in a given year who had a visit in the prior year.

Note: the N reported in the first row is the total number of patients with mean A1c 6-12% in each year. Patients who had not yet had their first visit by the year prior have a missing value for the variable indicating whether they had a visit in the prior year, so the denominator for the percentages is the number of patients with mean A1c 6-12% and who had their first visit by the prior year.

```{r echo=FALSE}
kable(t4)
```
<br>

## Percentage of patients with mean A1c >12% in a given year who had a visit in the prior year.

Note: the N reported in the first row is the total number of patients with mean A1c >12% in each year. Patients who had not yet had their first visit by the year prior have a missing value for the variable indicating whether they had a visit in the prior year, so the denominator for the percentages is the number of patients with mean A1c >12% and who had their first visit by the prior year.

```{r echo=FALSE}
kable(t5)
```
<br>

## Percentage of patients with mean A1c <7% in a given year who had a visit in the prior year.

Note: the N reported in the first row is the total number of patients with mean A1c <7% in each year. Patients who had not yet had their first visit by the year prior have a missing value for the variable indicating whether they had a visit in the prior year, so the denominator for the percentages is the number of patients with mean A1c <7% and who had their first visit by the prior year.

```{r echo=FALSE}
kable(t6)
```
<br>

## Percentage of patients with mean A1c 7 - <9% in a given year who had a visit in the prior year.

Note: the N reported in the first row is the total number of patients with mean A1c 7 - <9% in each year. Patients who had not yet had their first visit by the year prior have a missing value for the variable indicating whether they had a visit in the prior year, so the denominator for the percentages is the number of patients with mean A1c 7 - <9% and who had their first visit by the prior year.

```{r echo=FALSE}
kable(t7)
```
<br>

## Percentage of patients with mean A1c 9 - <12% in a given year who had a visit in the prior year.

Note: the N reported in the first row is the total number of patients with mean A1c 9 - <12% in each year. Patients who had not yet had their first visit by the year prior have a missing value for the variable indicating whether they had a visit in the prior year, so the denominator for the percentages is the number of patients with mean A1c 9 - <12% and who had their first visit by the prior year.

```{r echo=FALSE}
kable(t8)
```
<br>

## Percentage of patients with mean A1c >12% in a given year who had a visit in the prior year.

Note: the N reported in the first row is the total number of patients with mean A1c >12% in each year. Patients who had not yet had their first visit by the year prior have a missing value for the variable indicating whether they had a visit in the prior year, so the denominator for the percentages is the number of patients with mean A1c >12% and who had their first visit by the prior year.

```{r echo=FALSE}
kable(t9)
```
<br>

## Mean number of visits per year for patients with mean A1c >12% in a given year.

```{r echo=FALSE}
kable(t10)
```
<br>

## Mixed-effects model for mean A1c by year.

```{r echo=FALSE,comment=''}
kable(mod_anova)

summary(mod)
```
<br>

## Models examining predictors of having mean A1c >12% by year


### 2014

```{r echo=FALSE,comment=''}
kable(log2014_anova)

summary(log2014)
```
<br>

### 2015

```{r echo=FALSE,comment=''}
kable(log2015_anova)

summary(log2015)
```
<br>

### 2016

```{r echo=FALSE,comment=''}
kable(log2016_anova)

summary(log2016)
```
<br>

### 2017

```{r echo=FALSE,comment=''}
kable(log2017_anova)

summary(log2017)
```
<br>

### 2018

```{r echo=FALSE,comment=''}
kable(log2018_anova)

summary(log2018)
```
<br>

### 2019

```{r echo=FALSE,comment=''}
kable(log2019_anova)

summary(log2019)
```
<br>