---
title: "Tandem Patient Reported Outcomes (PROs)"
author: "Kristen Campbell"
date: "August 26, 2019"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
setwd("C:/Users/campbkri/Documents/GitHub/BDC-Code/Laurel Messer/Tandem/Code")
source("S:/Shared Material/Shared Code/R/temp_table1.R")

##functions specifically for this project
source("00_data.R")
source("01_survey_factors.R")
source("03_models.R")

table(dat$BaselineMethod)
#demographics table:
demo<-final_table(dat.all,c('B_RESPONDENT','Age','age_cat', 'duration_of_diabetes_at_baseline_years',
                        'Baseline_A1C','cgm_yn','method_cat'),
                     dat.all$cohort_complete,margin=2,single=F,2,col.names=T, summary.stat='mean')

demo.type<-final_table(dat,c('B_RESPONDENT','Age','age_cat', 'duration_of_diabetes_at_baseline_years',
                        'Baseline_A1C','cgm_yn'),
                     dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='mean')
demo.type
demo.type<-demo.type[,-2]

##survey data:
missing<-final_table(dat.all,c('baseline_complete_yn','post2m_complete_yn', 'post4m_complete_yn','post6m_complete_yn',
                               'allcomplete','cohort_complete'),
                     rep(1:nrow(dat.all)),margin=2,single=T,2,col.names=T, summary.stat='median')
missing





# #opt in data:
# 
# opt_in<-final_table(dat,c('Baseline_SurveyStudy_OptIn','post2m_BIQuse','post4m_BIQuse','post6m_BIQuse',
#                           'post2m_BIQ_YES','post4m_BIQ_YES','post6m_BIQ_YES'),
#                      rep(1:nrow(dat)),margin=2,single=F,2,col.names=T, summary.stat='median')
# opt_in
#missing data:
# missing<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="baseline_num_complete"):
#                                         which(colnames(dat)=="allcomplete"))]),
#                      rep(1:nrow(dat)),margin=2,single=F,2,col.names=T, summary.stat='median')
# missing

#baseline survey summary:
# baseline_surv<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="Baseline_1"):
#                                         which(colnames(dat)=="Baseline_12"),
#                                         which(colnames(dat)=="baseline_factor1"),
#                                         which(colnames(dat)=="baseline_factor2"))]),
#                      dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')
# baseline_surv
# mid_surv<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="mid_1"):
#                                         which(colnames(dat)=="mid_12"),
#                                         which(colnames(dat)=="mid_factor1"),
#                                         which(colnames(dat)=="mid_factor2"))]),
#                      dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')
# mid_surv
# 
# post6mo_surv<-final_table(dat,colnames(dat[,c(which(colnames(dat)=="post6m_1"):
#                                         which(colnames(dat)=="post6m_12"),
#                                         which(colnames(dat)=="post6m_factor1"),
#                                         which(colnames(dat)=="post6m_factor2"))]),
#                      dat$method_cat,margin=2,single=F,2,col.names=T, summary.stat='median')
# post6mo_surv

####Models:

```


##Summary Stats
**Table 1: Patient Demographics - By Method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(demo.type, align = c("l", "l"))
```

##Missing Data
**Table 2a: Survey Missing Data**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(missing, align = c("l", "l"))
```

There are `r nrow(dat)` patients that met the final inclusion criteria for the analysis. Of those `r nrow(dat)` patients, `r table(dat$mid_point)[1]` (`r round(table(dat$mid_point)[1]/nrow(dat)*100,2)`%) finished a 2 month survey, so that time point was used as the mid-point.  `r table(dat$mid_point)[2]` (`r round(table(dat$mid_point)[2]/nrow(dat)*100,2)`%) did not have a 2 month survey, but did have a 4 month survey, so this was used as the mid-point.

##Survey Data - plots
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
source('02_plots.R')
```
##Survey Data - results

**Table 4: Satisfaction changes, by previous method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(fac1.changes, align = c("l", "l"))
```

**Table 5: Diabetes Burden changes, by previous method**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(fac2.changes, align = c("l", "l"))
```

##Supplemental Table: 
**Table S1: Patient Demographics - Inclusion/Exclusion**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(demo, align = c("l", "l"))
```
