---
title: "SEEDS Pilot"
author: "Casey Sakamoto & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(table1)

# data upload from 3 redcap DB
pilot_hcp <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/SEEDS/SEEDSPilotHCP_DATA_2024-07-17_0910.csv")

pilot_pat <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/SEEDS/SEEDSPilotPatient_DATA_2024-07-17_0910.csv")

pilot_res <- read_csv("S:/Laura/BDC/Projects/Sarit Polsky/SEEDS/SEEDSPilotResearchSt_DATA_2024-07-17_0911.csv")
```


```{r dc, include = FALSE}
############### SEEDS Pilot HCP: HCP_Enrollment Survey*, HCP_confidence, HCP_Time&Encounter, HCP_confidence_EOS, HCP Pilot Final Questionnaire ####################
pilot_hcp = pilot_hcp %>% select(record_id:redcap_repeat_instance,hcpenroll_ptname:hcp_enrollment_survey_complete,
                                 hcpconf_cgmoverall:hcp_confidence_complete,
                                 hcpte_datetime:hcp_timeencounter_complete,
                                 hcpconf_cgmoverall_eos:hcp_confidence_eos_complete,
                                 hcpfinal_provdegree___1:hcp_pilot_final_questionnaire_complete)

pilot_hcp = pilot_hcp %>% mutate(Practice_Type = case_when(hcpenroll_practicetype == 0 ~ "Primary Care",
                                                           hcpenroll_practicetype == 1 ~ "Adult Endocrinology",
                                                           hcpenroll_practicetype == 2 ~ "Pediatric Endocrinology"),
                                 Use_EHR = case_when(hcpenroll_ehr == 0 ~ "CHCO EPIC",
                                                     hcpenroll_ehr == 1 ~ "UCH EPIC",
                                                     hcpenroll_ehr == 2 ~ "Does Not Use",
                                                     hcpenroll_ehr == 3 ~ "Other"),
                                 Apprx_Pt_Medicare = hcpenroll_ptmedicare,
                                 Apprx_Pt_Medicaid = hcpenroll_ptmedicaid,
                                 Apprx_Pt_Dual = hcpenroll_ptdualmed,
                                 Apprx_Pt_Selfpay = hcpenroll_ptselfpay,
                                 Apprx_Pt_Otherpay = hcpenroll_ptotherpay,
                                 Apprx_Pt_CommercialPrivate = hcpenroll_ptcomm,
                                 Weekly_Pts = hcpenroll_totalpts,
                                 Apprx_T1D_Pts = hcpenroll_t1dpts,
                                 Apprx_T2D_Pts_noins = hcpenroll_t2dptsnoinsulin,
                                 Apprx_T2D_Pts_ins = hcpenroll_t2dptsinsulin,
                                 Practice_Apprx_T1D_Pts = hcpenroll_t1dptsprx,
                                 Practice_Apprx_T2D_Pts_noins = hcpenroll_t2dptsprxni,
                                 Practice_Apprx_T2D_Pts_ins = hcpenroll_t2dptsprxwi)

# hcp demog table
pilot_hcp_demog = pilot_hcp %>% select(record_id, Practice_Type:Practice_Apprx_T2D_Pts_ins) %>% filter(!is.na(Practice_Type))                                 
HCP_Demographics = table1(~.|Practice_Type, data = pilot_hcp_demog %>% select(-record_id))                                 

# descriptive statistics
# 	*HCP_Enrollment_survey: analyze demographic info by practice type and overall, report of likert scale question, report percentage of patients seen in payer categories

hcp_enroll  = pilot_hcp %>% filter(is.na(redcap_repeat_instrument)) %>% select(record_id, Practice_Type ,hcpenroll_tech1:hcpenroll_tech3)
hcp_enroll = hcp_enroll %>% mutate(across(hcpenroll_tech1:hcpenroll_tech3, as.factor))
HCP_Likert = table1(~.|Practice_Type, data =hcp_enroll %>% select(-record_id)) 

#	*HCP Pilot Final Questionnaire: analyze by practice type and overall of all quantitative data (free-text not included in analysis)

hcp_finalq = pilot_hcp %>% select(record_id, Practice_Type, hcpfinal_provdegree___1:hcp_pilot_final_questionnaire_complete) %>% filter(!is.na(Practice_Type)) %>%
  mutate(prov_degree = case_when(hcpfinal_provdegree___1 == 1 ~ "MD/DO or equivalent",
                                 hcpfinal_provdegree___2 == 1 ~ "PhD",
                                 hcpfinal_provdegree___3 == 1 ~ "PharmD",
                                 hcpfinal_provdegree___4 == 1 ~ "PA/NP",
                                 hcpfinal_provdegree___5 == 1 ~ "MSN",
                                 hcpfinal_provdegree___6 == 1 ~ "RN",
                                 hcpfinal_provdegree___7 == 1 ~ "RD",
                                 hcpfinal_provdegree___8 == 1 ~ "Other"))

hcp_finalq = hcp_finalq %>% select(-c(hcpfinal_provdegree___1:hcpfinal_provdegree_other,
                                     hcpfinal_traininghelpful:hcpfinal_ptsasked,
                                     hcpfinal_partdeclineyes, hcpfinal_endosuggestionyes,
                                     hcpfinal_websiteufno, hcpfinal_websitenavno,
                                     hcpfinal_websitefree,hcpfinal_websiteunclear,hcpfinal_websitemissingyes,
                                     hcpfinal_websiteclearno,hcpfinal_workflowclearno_2,hcpfinal_workflowmissingyes,
                                     hcpfinal_insurancewffeedback,hcpfinal_nodevicestartyes,hcpfinal_prevdevyes,
                                     hcpfinal_prevdevno,hcpfinal_devicedisncontyes,hcpfinal_timesaveyes,hcpfinal_endopumpsupportyes_2,
                                     hcp_pilot_final_questionnaire_complete)) %>% mutate(across(hcpfinal_trainingprocess:hcpfinal_endoinjection, as.factor))

HCP_FinalQ = table1(~.|Practice_Type, data =hcp_finalq %>% select(-c(record_id, hcpfinal_practicetype))) 


# iii.	HCP_Time&Encounter: analyze of time spent on tasks and billing codes, by practice type and overall, and visit type and overall
hcp_te = pilot_hcp %>% select(record_id, Practice_Type, hcpte_type:hcp_timeencounter_complete) %>% fill(Practice_Type) %>% filter(!is.na(hcpte_type))

hcp_te = hcp_te %>% select(record_id, Practice_Type, hcpte_time,hcpte_bill,hcpte_bill_yes___0:hcpte_bill_yes___6) %>% 
  mutate(billing_codes = case_when(hcpte_bill_yes___0 == 1 ~ "95249",
                                   hcpte_bill_yes___1 == 1 ~ "95250",
                                   hcpte_bill_yes___2 == 1 ~ "95251",
                                   hcpte_bill_yes___3 == 1 ~ "99213 (w/modifier 25)",
                                   hcpte_bill_yes___4 == 1 ~ "99214 (w/modifier 25)",
                                   hcpte_bill_yes___5 == 1 ~ "99215 (w/modifier 25)",
                                   hcpte_bill_yes___6 == 1 ~ "Other",
                                   hcpte_bill == 2 ~ "Not Billed")) %>% select(-(hcpte_bill_yes___0:hcpte_bill_yes___6), -hcpte_bill)
HCP_TE = table1(~.|Practice_Type, data =hcp_te %>% select(-c(record_id))) 

hcp_te_device = pilot_hcp %>% select(record_id, Practice_Type, hcpte_type:hcp_timeencounter_complete) %>% fill(Practice_Type) %>% filter(!is.na(hcpte_type))
hcp_te_device = hcp_te_device %>% select(record_id, Practice_Type, hcpte_type,hcpte_devicerx:new_device_type)
hcp_te_device = hcp_te_device %>% mutate(device_prescribed = case_when(hcpte_devicerx == 1 ~"Yes",
                                                                       hcpte_devicerx == 0 ~"No",
                                                                       hcpte_type == 1 ~"Non SEEDS Pt Encounter"))

hcp_te_device = hcp_te_device %>% select(record_id, Practice_Type, device_prescribed)
HCP_Device = table1(~.|Practice_Type, data =hcp_te_device %>% select(-c(record_id))) 

#d.	CGM Data from patients who started on CGM during the pilot i.	Compare from baseline to 6-week post device start
hcp_cgm = pilot_hcp %>% select(record_id, Practice_Type, hcpconf_cgmoverall:hcp_confidence_complete, hcpconf_cgmoverall_eos:hcp_confidence_eos_complete) %>% 
  fill(Practice_Type) %>% filter(!is.na(hcpconf_cgmoverall))

hcp_cgm_long = gather(hcp_cgm, key = "eos", value = "Likert", -record_id, -Practice_Type)
hcp_cgm_long = hcp_cgm_long %>% filter(!(eos %in% c("hcp_confidence_complete", "hcp_confidence_eos_complete")))
hcp_cgm_long$eos = str_remove(hcp_cgm_long$eos, "hcpconf_")
hcp_cgm_long = hcp_cgm_long %>% mutate(eos_visit = ifelse(grepl("_eos", eos), "EOS","Baseline"))
hcp_cgm_long = hcp_cgm_long %>% mutate(likert_q = str_remove(eos, "_eos")) %>% select(-eos)

test = hcp_cgm_long %>% spread(likert_q, Likert)

# p value table
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}

pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g, paired = T)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}


#lapply(split(hcp_cgm_long, hcp_cgm_long$likert_q), function(df) t.test(Likert ~ eos_visit , data = df))

################## SEEDS Pilot Patient:PT_Enroll*Participant Withdrawal Or Removal ######################
#Should any data points be excluded?	Pilot Patients Adult-1-Unassigned and PCP-2-Unassigned  

pilot_pat = pilot_pat %>% select(record_id, pt_enroll_dob:pt_enroll_complete, worremoval_practicetype:participant_withdrawal_or_removal_complete)
pilot_pat = pilot_pat %>% filter(! record_id %in% c("Adult-1-Unassigned", "PCP-2-Unassigned"))

#patient demographics table
pilot_pat = pilot_pat %>% mutate(Race = case_when(pt_enroll_race___1 == 1 ~ "White",
                                                  pt_enroll_race___2 == 1 ~ "Black or African American",
                                                  pt_enroll_race___3 == 1 ~ "American Indian or Alaska Native",
                                                  pt_enroll_race___4 == 1 ~ "Asian",
                                                  pt_enroll_race___5 == 1 ~ "Native Hawaiian or Pacific Islander",
                                                  pt_enroll_race___6 == 1 ~"Other",
                                                  TRUE ~ "Missing"),
                                 Ethnicity = case_when(pt_enroll_ethnicity == 0 ~ "Non-Hispanic",
                                                       pt_enroll_ethnicity == 1 ~ "Hispanic or Latino"),
                                 Diabetes_Type = case_when(pt_enroll_diabetestype == 1 ~ "Type 1 Diabetes",
                                                           pt_enroll_diabetestype == 2 ~ "Type 2 Diabetes",
                                                           pt_enroll_diabetestype == 3 ~ "Other"),
                                 Diabetes_Duration = pt_enroll_diabetesdur,
                                 Insurance_Type = case_when(pt_enroll_insurancetype___0 == 1 ~ "Commercial",
                                                            pt_enroll_insurancetype___1 == 1 ~ "Medicare",
                                                            pt_enroll_insurancetype___2 == 1 ~ "Medicaid",
                                                            pt_enroll_insurancetype___3 == 1 ~ "Government or Military",
                                                            pt_enroll_insurancetype___4 == 1 ~ "None",
                                                            pt_enroll_insurancetype___5 == 1 ~ "Other"),
                                 Device_Use = factor(pt_enroll_usedevices, levels= c(0,1), labels= c("Yes", "No")),
                                 Practice_Type = case_when(pt_enroll_practicetype == 0 ~ "Primary Care",
                                                           pt_enroll_practicetype == 1 ~ "Adult Endocrinology",
                                                           pt_enroll_practicetype == 2 ~ "Pediatric Endocrinology"
                                                           ))
# c.	SEEDS Pilot Patient (PID 27895)i.	*PT_Enroll: analysis of demographics by practice type and overall
Patient_Demographics = table1(~ Race + Ethnicity + Diabetes_Type + Diabetes_Duration + Insurance_Type + Device_Use| Practice_Type, data = pilot_pat)

# ii.	Participant Withdrawal Or Removal: analyze the reason for withdrawal or removal by practice type and overall
pat_wr = pilot_pat %>% mutate(withdrawal_reason = case_when(withdrawal_reason___0 == 1 ~ "Patient Participant Moved Away",
                                                            withdrawal_reason___1 == 1 ~ "No Longer Wanted to be in Study",
                                                            withdrawal_reason___2 == 1 ~ "HCP Participant Moved Away",
                                                            withdrawal_reason___3 == 1 ~ "Other",
                                                            withdrawal_reason___4 == 1 ~ "No Reason Given"),
                              removal_reason = case_when(removal_reason___0 == 1 ~ "Patient Participant Moved Away",
                                                            removal_reason___1 == 1 ~ "No Longer Wanted to be in Study",
                                                            removal_reason___2 == 1 ~ "HCP Participant Moved Away",
                                                            removal_reason___3 == 1 ~ "Other",
                                                            removal_reason___4 == 1 ~ "No Reason Given"),
                              include = ifelse(is.na(withdrawal_or_removal), 0, 1)) %>% 
  select(record_id, Practice_Type, withdrawal_or_removal, withdrawal_reason, removal_reason, include)
PAT_WR = table1(~withdrawal_reason + removal_reason| Practice_Type, data = pat_wr %>% select(-record_id) %>% filter(include == 1))

############################ SEEDS Pilot Research Staff Research Staff Time&Encounter*###################################
pilot_res = pilot_res %>% select(record_id:redcap_repeat_instance, hcpte_datetime:research_staff_timeencounter_complete)

## i.	*Research Staff Time&Encounter: analyze of time spent by practice type and overall, visit type and overall,  
pilot_res_te = pilot_res %>% select(record_id, rste_practice_type, rste_hcp_whichvisit___0:rste_hcp_whichvisit___4, rste_time) %>% 
  mutate(visit_type = case_when(rste_hcp_whichvisit___0 == 1 ~ "Enrollment",
                                rste_hcp_whichvisit___1 == 1 ~ "Training A (Protocol/Device Expert Consultation)",
                                rste_hcp_whichvisit___2 == 1 ~ "Training B (endo.digital)",
                                rste_hcp_whichvisit___3 == 1 ~ "Patient Recruitment",
                                rste_hcp_whichvisit___4 == 1 ~ "Other/Not Tied to a Visit")) %>% select(-c(rste_hcp_whichvisit___0:rste_hcp_whichvisit___4))

pilot_res_te[57,3] = 0 

pilot_res_te = pilot_res_te %>% filter(!is.na(rste_time == 0))
pilot_res_te$rste_practice_type = factor(pilot_res_te$rste_practice_type, levels = c(0,1,2), labels = c("Patient", "HCP", "Both"))
RES_TE_vis = table1(~ rste_time| visit_type, data = pilot_res_te %>% filter(!is.na(visit_type)))
RES_TE_prac = table1(~ rste_time| rste_practice_type, data = pilot_res_te %>% filter(!is.na(rste_practice_type)))
RES_TE_both = t(table1(~ rste_time| rste_practice_type*visit_type, data = pilot_res_te %>% filter(!is.na(rste_practice_type)) %>% filter(!is.na(visit_type))))

# questionnaire_scoring: hcp conf: 1-10 likert; hcp_confidence_eos: 1-10 likert; hcp_final questionnaire: both 1-5 and 1-10 likert
```



# Data Remarks/ Questions

- did we receive CGM info for analysis plan 6d from Dreamed? I'm not sure what variables to look for.

- Tamara Oser time spent set to 0, any other input for classifications etc?


# Methods

Demographics for the overall groups and subgroups produced.

Descriptive statistics were produced for survey responses and likerts.

T-tests were run on the confidence outcomes between baseline and EOS for the overall and Patient Care groups. Insufficient data to run the tests for the other two subgroups.

# Analysis

## HCP

note:  record id Adult-2 missing a bunch of questions

### Demographics

```{r}
HCP_Demographics
```

### Enrollment Likert

1: Strong Disagree - 5: Strong Agree
tech savvy, ask for help when troubleshooting, first to try new tech

```{r}
HCP_Likert
```

### Final Questionnaire
Likerts:
1: Not at all Satisfied - 10 :Extremely Satisfied, 11 = N/A
1: Strong Disagree - 5: Strong Agree 6: N/A

1 = Yes, 0 or 2 = No , 3 : N/A
tech savvy, ask for help when troubleshooting, first to try new tech

```{r}
HCP_FinalQ
```

### Time & Encounter

data from any task or work related to an enrolled SEEDS participant; thus each subj has multiple rows per task

the devices type column was completely blank, so instead presenting whether one was prescribed

```{r}
HCP_TE
HCP_Device
```

### CGM Confidence Overall

1 = "not at all confident" and 10 = "very confident"

```{r}
table1(~.|eos_visit , data = test %>% filter(record_id != "Adult-2" )%>% select(-record_id, - Practice_Type) , overall=F, extra.col=list(`P-value`=pvalue),render.continuous=my.render.cont )
```

### CGM Confidence Primary Care

1 = "not at all confident" and 10 = "very confident"

not enough 'y' observations to run tests for this subgroup; not large enough N

```{r}
table1(~.|eos_visit , data = test %>% filter(Practice_Type == "Primary Care" ) %>% select(-record_id, - Practice_Type) , overall=F ,render.continuous=my.render.cont )
```

### CGM Confidence Adult Endo

1 = "not at all confident" and 10 = "very confident"

not enough 'y' observations to run tests for this subgroup; not large enough N

```{r}
table1(~.|eos_visit , data = test %>% filter(Practice_Type == "Adult Endocrinology" ) %>% select(-record_id, - Practice_Type) , overall=F ,render.continuous=my.render.cont )

```

### CGM Confidence Ped Endo

1 = "not at all confident" and 10 = "very confident"

t-tests able to run for some columns; low n should be taken into account

```{r}
table1(~.|eos_visit , data = test %>% filter(Practice_Type == "Pediatric Endocrinology" ) %>% select(-record_id, - Practice_Type) , overall=F,extra.col=list(`P-value`=pvalue) ,render.continuous=my.render.cont )

```

## Patient

### Demographics

```{r}
Patient_Demographics
```

### Participant Withdrawal or Removal

```{r}
PAT_WR
```



## Researcher

### Time Encounter

```{r}
RES_TE_vis
RES_TE_prac
RES_TE_both
```