---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(table1)

# Data Upload and Labeling
setwd("S:/Laura/BDC/Projects/Erin Cobry/ADA Skin and Devices/")
source("S:/Laura/BDC/Projects/Erin Cobry/ADA Skin and Devices/InsulinPumpsAndSenso_R_2024-10-18.r")
```

```{r dc, include=FALSE}
analysis_raw = data %>% select(record_id, calculated_age, gender.factor, hispanic_latino.factor, race___1.factor:race___6.factor, # demographics ##
                               diabetesduration, hba1c, cgm_use.factor, cgm_type.factor, insulin_pump.factor, type_insulin_pumo.factor,
                               device_location___1.factor:device_location___7.factor, ############
                               
                               skin_reactions.factor, problem_location___1.factor:problem_location___7.factor, # Skin Related Questions due to cgm
                               pump_skin_reaction.factor, pump_location_problem, pump_skin_appearance___1.factor:pump_skin_appearance___5.factor # Qs due to pump
                               )


```


```{r demogs, echo = FALSE}
demog_df = analysis_raw %>% select()


```

# Outstanding Data Remarks/ Questions

# Background

Skin related questions (selected items) – Skin and Diabetes Devices Survey (after Demographics)
a.	Skin reactions due to the use of glucose sensor (please describe location and skin reaction): Question #1 (skin reaction due to the use of CGM (yes/no)), Question #2 (location of skin reaction), Question #3 (description of skin reaction)
b.	Skin reaction due to the use of insulin pump (please describe location and skin reaction):  Question #4 (skin reaction due to the use of insulin pump (yes/no)), Question #5 (location of skin reaction), Question #6 (description of skin reaction)

Analysis Ideas
1.	Categorize based on race/ethnicity and gender (Race/ethnicity and gender can both be found on Demographics section of the survey)
a.	Analyze if there was a difference in skin reactions incidence with diabetes devices (combine those with CGM only, pump only, and CGM and pump reactions) _and the race/ethnicity (include how many of each category). 
b.	Analyze if there was a difference in skin reactions incidence with diabetes devices (combine those with CGM only, pump only, and CGM and pump reactions) _and the gender (include how many of each category).


# Methods

# Analysis