---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(table1)


```

```{r dc, include=FALSE}
# Data Upload and Labeling
setwd("S:/Laura/BDC/Projects/Erin Cobry/ADA Skin and Devices/")
source("S:/Laura/BDC/Projects/Erin Cobry/ADA Skin and Devices/InsulinPumpsAndSenso_R_2024-10-18.r")
analysis_raw = data %>% select(record_id, calculated_age, gender.factor, hispanic_latino.factor, race___1.factor:race___6.factor, # demographics ##
                               diabetesduration, hba1c, cgm_use.factor, cgm_type.factor, insulin_pump.factor, type_insulin_pumo.factor,
                               device_location___1.factor:device_location___7.factor, ############
                               
                               skin_reactions.factor, problem_location___1.factor:problem_location___7.factor, skin_reaction_description___1.factor:skin_reaction_description___5.factor,# Skin Related Questions due to cgm
                               pump_skin_reaction.factor, pump_location_problem, pump_skin_appearance___1.factor:pump_skin_appearance___5.factor, # Qs due to pump
                               
                               prepare_skin.factor,removal_technique.factor,skin_care.factor, adhesion_problem.factor,sensor_pump_activities.factor,overlay_tapes.factor,
                               feelings_skin_reactions___1.factor:feelings_skin_reactions___7.factor,school.factor, hobbies.factor, sleep.factor, change_sensor_pumps.factor,
                               stop_use.factor, # additional comparisons for yes/no reactions
                               score_pedsql_parent,
                               score_pedsql_young,
                               score_peds_ql_child,
                               score_pedsql_teen,
                               score_pedsql_young_adult,
                               paid_peds_score,
                               paid_pr_auto_score,
                               paid_adult_score #scores
                               )

# exclude 14 17 23 24
analysis_raw = analysis_raw %>% filter(!(record_id %in% c(14,17,23,24)))

# race classification
analysis_raw = analysis_raw %>% rowwise() %>% mutate(race_count = sum(!grepl("Unchecked", c_across(race___1.factor:race___6.factor))),
                                                     Race = case_when(race_count == 1 & race___1.factor == "Checked" ~ "American Indian/Alaska Native",
                                                                      race_count == 1 & race___2.factor == "Checked" ~ "Asian",
                                                                      race_count == 1 & race___3.factor == "Checked" ~ "African American/Black",
                                                                      race_count == 1 & race___4.factor == "Checked" ~ "Native Hawaiian/Pacific Islander",
                                                                      race_count == 1 & race___5.factor == "Checked" ~ "White",
                                                                      race_count == 1 & race___6.factor == "Checked" ~ "Do Not Wish to Provide",
                                                                      race_count > 1 ~ "Multiracial"
                                                                      ))

analysis_raw = analysis_raw %>% mutate(race_w = ifelse(Race == "White", "White", "Nonwhite"))

# pump/cgm / both/neither reactions
analysis_raw = analysis_raw %>% mutate(Reactions_Group = case_when(skin_reactions.factor == "Yes" & pump_skin_reaction.factor == "Yes" ~ "CGM and Pump Reactions",
                                                                   skin_reactions.factor == "No" & pump_skin_reaction.factor == "No" ~ "Neither",
                                                                   skin_reactions.factor == "Yes" & pump_skin_reaction.factor == "No" ~ "CGM Reactions",
                                                                   skin_reactions.factor == "No" & pump_skin_reaction.factor == "Yes" ~ "Pump Reactions",
                                                                   skin_reactions.factor == "Yes" & is.na(pump_skin_reaction.factor) ~ "CGM Reactions",
                                                                   TRUE~"Neither"))

analysis_raw = analysis_raw %>% mutate(Any_Reactions = ifelse(skin_reactions.factor == "Yes" | pump_skin_reaction.factor == "Yes", "Yes", "No"))

analysis = analysis_raw %>% filter(Reactions_Group != "Neither")

analysis_raw = analysis_raw %>% mutate(pedsql_score = coalesce(as.numeric(score_pedsql_parent),
                               as.numeric(score_pedsql_young),
                               as.numeric(score_peds_ql_child),
                               as.numeric(score_pedsql_teen),
                               as.numeric(score_pedsql_young_adult)),
                                       paid_score = coalesce( as.numeric(paid_peds_score),
                               as.numeric(paid_pr_auto_score),
                               as.numeric(paid_adult_score)
))
# appended to remove the na any skin reactions
analysis_raw = analysis_raw %>% filter(!is.na(Any_Reactions))                                                         
```


```{r tables, echo = FALSE}

# table 1
demog_df = analysis_raw %>% select( calculated_age, gender.factor, Race, race_w,hispanic_latino.factor, diabetesduration:device_location___7.factor)

dt1 = table1(~., data = demog_df)
#dt1

# skin related questions cgm

# location
cgm_skin_location_df = analysis_raw %>% filter(skin_reactions.factor == "Yes") %>% select(problem_location___1.factor:problem_location___7.factor)
csl_colnames = c("Upper Buttocks", "Abdomen", "Upper Hip", "Upper Thigh", "Upper Arm", "Forearm", "Other")

colnames(cgm_skin_location_df) = csl_colnames
csl_location = table1(~., data = cgm_skin_location_df)

# reaction
cgm_skin_reaction_df = analysis_raw %>% filter(skin_reactions.factor == "Yes") %>% select(skin_reaction_description___1.factor:skin_reaction_description___5.factor)

csr_colnames = c("Redness", "Itching", "Scarring", "Infections", "Other")
colnames(cgm_skin_reaction_df) = csr_colnames

csr_reactions = table1(~., data = cgm_skin_reaction_df)

# skin related questions pump

# location
#analysis_raw$pump_skin_reaction.factor
pump_skin_location_df = analysis_raw %>% filter(pump_skin_reaction.factor == "Yes") %>% select(pump_location_problem)

# new table
pumpskinreactiontable = c(rep("Abdomen", 21), rep("Upper Buttocks", 10), rep("Arm", 9), "Upper Hip", rep("Legs", 12), rep("All/Anywhere", 4))


# reaction
pump_skin_reaction_df = analysis_raw %>% filter(pump_skin_reaction.factor == "Yes") %>% select(pump_skin_appearance___1.factor:pump_skin_appearance___5.factor)

colnames(pump_skin_reaction_df) = csr_colnames

pump_reactions = table1(~., data = pump_skin_reaction_df)

# appended table
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g, na.action = na.omit)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- fisher.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

t2_df = analysis_raw %>% select(Any_Reactions, prepare_skin.factor:stop_use.factor, pedsql_score,paid_score, hba1c)
t2 = table1(~.|Any_Reactions, data = t2_df %>% filter(!is.na(Any_Reactions)), overall = F, extra.col = list(`P-Value`=pvalue))
```

# Outstanding Data Remarks/ Questions

- Some subjects race had multiple checked. Currently classifying in own category as multirace for demographic description
  - Distribution of race overwhelmingly white with only a few nonwhite (N = 11 nonwhite across 4 categories, 86 White) maybe Dichotomize W/NW?
   ---- using white/nw (var: race_w)
  - Hisp/Latino ("Yes" N= 11) all selected White as race (1 Prefer not to answer)
  - Gender contains n = 1 for other, creating a group split of n = 56 F, 40M, 1 Oth : recid 32
  
- pump skin location differently formatted than cgm; how to classify some of these?

# Background


# Methods
Demographics and skin related questions were produced. Due to low cell counts within race and gender groups, Fisher's Exact tests were performed to assess difference in skin reactions by gender and race.


# Analysis

## Demographics
```{r}
dt1
```

## Skin reactions due to the use of glucose sensor 

N = 53 yes, 43 No

```{r}
# Total n = Yes
table(analysis_raw$skin_reactions.factor)

# Location 
csl_location

# description
csr_reactions
```

## Skin reactions due to the use of insulin pump

N = 45 yes, 41 No

```{r}
# Total n = Yes
table(analysis_raw$pump_skin_reaction.factor)

# Location 
#table(pump_skin_location_df$pump_location_problem)

# Updated
kable(table(pumpskinreactiontable))

# description
pump_reactions
```

## Skin Reaction Groups by Gender/Race

```{r}
table(analysis_raw$Reactions_Group)

table(analysis_raw$Reactions_Group, analysis_raw$gender.factor)

table(analysis_raw$Reactions_Group, analysis_raw$race_w)
```

### Difference in Skin Reactions Groups by Gender and Race

No significant differences were found within Skin Reaction groups by Gender (p = 0.81) or Race (p = 0.82).

```{r}
fisher.test(analysis$Reactions_Group, analysis$gender.factor)
fisher.test(analysis$Reactions_Group, analysis$race_w)

# Chisq sanity check -- similar results
# chisq.test(analysis$Reactions_Group, analysis$gender.factor) 0.8
# chisq.test(analysis$Reactions_Group, analysis$race_w) 0.5
```

### APPENDED 11/13 Difference in Any Skin Reactions by Gender and Race

There were no significant association between device use and gender (p > 0.99) or race (p = 0.08)

```{r}
table(analysis_raw$Any_Reactions, analysis_raw$gender.factor)

fisher.test(analysis_raw$Any_Reactions, analysis_raw$gender.factor)
fisher.test(analysis_raw$Any_Reactions, analysis_raw$race_w)

```


# APPENDED

Participants were dichotomized into Any Skin reactions Y/N (pump and cgm combined). Fishers exact tests were run to assess difference in reactions by skin and device questions.T-tests were performed on hba1c and survey outcomes between groups.

There was a significant difference in all of the feelings due to skin reactions except embarrassed and other (p < 0.05). See below analysis.

There was a significant difference in PAID scores between skin reactions groups (p = 0.01). The any skin reactions group had a 10.6 point higher Paid score on average


```{r}
t2
```

```{r, include = F}
# prepare skin
fisher.test(analysis_raw$Any_Reactions, analysis_raw$prepare_skin.factor)
# removal technique
fisher.test(analysis_raw$Any_Reactions, analysis_raw$removal_technique.factor)
#skin care
fisher.test(analysis_raw$Any_Reactions, analysis_raw$skin_care.factor)
#adhesionproblem
fisher.test(analysis_raw$Any_Reactions, analysis_raw$adhesion_problem.factor)
# sensor pump activities
fisher.test(analysis_raw$Any_Reactions, analysis_raw$sensor_pump_activities.factor)
# overlay tapes
fisher.test(analysis_raw$Any_Reactions, analysis_raw$overlay_tapes.factor)
# self conscious ***
fisher.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___1.factor)
# upset ***
fisher.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___2.factor)
# embarrassed
fisher.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___3.factor)
# uncomfortable ***
fisher.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___4.factor)
# anger***
fisher.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___5.factor)
# none of the above***
fisher.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___6.factor)
# other
fisher.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___7.factor)
# miss school
fisher.test(analysis_raw$Any_Reactions, analysis_raw$school.factor)
# stop hobbies 
fisher.test(analysis_raw$Any_Reactions, analysis_raw$hobbies.factor)
# sleep disturbances
fisher.test(analysis_raw$Any_Reactions, analysis_raw$sleep.factor)
# change method/pump
fisher.test(analysis_raw$Any_Reactions, analysis_raw$change_sensor_pumps.factor)
# stop device use
fisher.test(analysis_raw$Any_Reactions, analysis_raw$stop_use.factor)
# chisq sanity -- same results
# # prepare skin
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$prepare_skin.factor)
# # removal technique
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$removal_technique.factor)
# #skin care
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$skin_care.factor)
# #adhesionproblem
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$adhesion_problem.factor)
# # sensor pump activities
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$sensor_pump_activities.factor)
# # overlay tapes
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$overlay_tapes.factor)
# # self conscious ***
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___1.factor)
# # upset ***
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___2.factor)
# # embarrassed
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___3.factor)
# # uncomfortable ***
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___4.factor)
# # anger***
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___5.factor)
# # none of the above***
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___6.factor)
# # other
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$feelings_skin_reactions___7.factor)
# # miss school
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$school.factor)
# # stop hobbies 
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$hobbies.factor)
# # sleep disturbances
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$sleep.factor)
# # change method/pump
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$change_sensor_pumps.factor)
# # stop device use
# chisq.test(analysis_raw$Any_Reactions, analysis_raw$stop_use.factor)
```
