---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readr)
library(table1)

# Data Upload and Labeling
setwd("S:/Laura/BDC/Projects/Erin Cobry/ADA Skin and Devices/")
source("S:/Laura/BDC/Projects/Erin Cobry/ADA Skin and Devices/InsulinPumpsAndSenso_R_2024-10-18.r")
```

```{r dc, include=FALSE}
data$skinreactio
analysis_raw = data %>% select(record_id, calculated_age, gender.factor, hispanic_latino.factor, race___1.factor:race___6.factor, # demographics ##
                               diabetesduration, hba1c, cgm_use.factor, cgm_type.factor, insulin_pump.factor, type_insulin_pumo.factor,
                               device_location___1.factor:device_location___7.factor, ############
                               
                               skin_reactions.factor, problem_location___1.factor:problem_location___7.factor, skin_reaction_description___1.factor:skin_reaction_description___5.factor,# Skin Related Questions due to cgm
                               pump_skin_reaction.factor, pump_location_problem, pump_skin_appearance___1.factor:pump_skin_appearance___5.factor # Qs due to pump
                               )
# race classification
analysis_raw = analysis_raw %>% rowwise() %>% mutate(race_count = sum(!grepl("Unchecked", c_across(race___1.factor:race___6.factor))),
                                                     Race = case_when(race_count == 1 & race___1.factor == "Checked" ~ "American Indian/Alaska Native",
                                                                      race_count == 1 & race___2.factor == "Checked" ~ "Asian",
                                                                      race_count == 1 & race___3.factor == "Checked" ~ "African American/Black",
                                                                      race_count == 1 & race___4.factor == "Checked" ~ "Native Hawaiian/Pacific Islander",
                                                                      race_count == 1 & race___5.factor == "Checked" ~ "White",
                                                                      race_count == 1 & race___6.factor == "Checked" ~ "Do Not Wish to Provide",
                                                                      race_count > 1 ~ "Multiracial"
                                                                      ))

analysis_raw = analysis_raw %>% mutate(race_w = ifelse(Race == "White", "White", "Nonwhite"))

# pump/cgm / both/neither reactions
analysis_raw = analysis_raw %>% mutate(Reactions_Group = case_when(skin_reactions.factor == "Yes" & pump_skin_reaction.factor == "Yes" ~ "CGM and Pump Reactions",
                                                                   skin_reactions.factor == "No" & pump_skin_reaction.factor == "No" ~ "Neither",
                                                                   skin_reactions.factor == "Yes" & pump_skin_reaction.factor == "No" ~ "CGM Reactions",
                                                                   skin_reactions.factor == "No" & pump_skin_reaction.factor == "Yes" ~ "Pump Reactions"))

analysis_raw = analysis_raw %>% mutate(Any_Reactions = ifelse(skin_reactions.factor == "Yes" | pump_skin_reaction.factor == "Yes", "Yes", "No"))

analysis = analysis_raw %>% filter(Reactions_Group != "Neither")


                                                                   
```


```{r tables, echo = FALSE}

# table 1
demog_df = analysis_raw %>% select( calculated_age, gender.factor, Race, race_w,hispanic_latino.factor, diabetesduration:device_location___7.factor)

dt1 = table1(~., data = demog_df)
#dt1

# skin related questions cgm

# location
cgm_skin_location_df = analysis_raw %>% filter(skin_reactions.factor == "Yes") %>% select(problem_location___1.factor:problem_location___7.factor)
csl_colnames = c("Upper Buttocks", "Abdomen", "Upper Hip", "Upper Thigh", "Upper Arm", "Forearm", "Other")

colnames(cgm_skin_location_df) = csl_colnames
csl_location = table1(~., data = cgm_skin_location_df)

# reaction
cgm_skin_reaction_df = analysis_raw %>% filter(skin_reactions.factor == "Yes") %>% select(skin_reaction_description___1.factor:skin_reaction_description___5.factor)

csr_colnames = c("Redness", "Itching", "Scarring", "Infections", "Other")
colnames(cgm_skin_reaction_df) = csr_colnames

csr_reactions = table1(~., data = cgm_skin_reaction_df)

# skin related questions pump

# location
analysis_raw$pump_skin_reaction.factor
pump_skin_location_df = analysis_raw %>% filter(pump_skin_reaction.factor == "Yes") %>% select(pump_location_problem)

# reaction
pump_skin_reaction_df = analysis_raw %>% filter(pump_skin_reaction.factor == "Yes") %>% select(pump_skin_appearance___1.factor:pump_skin_appearance___5.factor)

colnames(pump_skin_reaction_df) = csr_colnames

pump_reactions = table1(~., data = pump_skin_reaction_df)


```

```{r analysis ds, echo = F}
# 1.	Categorize based on race/ethnicity and gender (Race/ethnicity and gender can both be found on Demographics section of the survey)
  # a.	Analyze if there was a difference in skin reactions incidence with diabetes devices (combine those with CGM only, pump only, and CGM and pump reactions) _and the race/ethnicity (include how many of each category). 
  # b.	Analyze if there was a difference in skin reactions incidence with diabetes devices (combine those with CGM only, pump only, and CGM and pump reactions) _and the gender (include how many of each category).

```

# Outstanding Data Remarks/ Questions

- Some subjects race had multiple checked. Currently classifying in own category as multirace for demographic description
  - Distribution of race overwhelmingly white with only a few nonwhite (N = 11 nonwhite across 4 categories, 86 White) maybe Dichotomize W/NW?
  - Hisp/Latino ("Yes" N= 11) all selected White as race (1 Prefer not to answer)
  - Gender contains n = 1 for other, creating a group split of n = 56 F, 40M, 1 Oth
  
- pump skin location differently formatted than cgm; how to classify some of these?

# Background


# Methods
Demographics and skin related questions were produced. Due to low cell counts within race and gender groups, Fisher's Exact tests were performed to assess difference in skin reactions by gender and race.


# Analysis

## Demographics
```{r}
dt1
```

## Skin reactions due to the use of glucose sensor 

N = 53 yes, 43 No

```{r}
# Total n = Yes
table(analysis_raw$skin_reactions.factor)

# Location 
csl_location

# description
csr_reactions
```

## Skin reactions due to the use of insulin pump

N = 45 yes, 41 No

```{r}
# Total n = Yes
table(analysis_raw$pump_skin_reaction.factor)

# Location 
table(pump_skin_location_df$pump_location_problem)

# description
pump_reactions
```

## Skin Reaction Groups by Gender/Race

```{r}
table(analysis_raw$Reactions_Group)

table(analysis_raw$Reactions_Group, analysis_raw$gender.factor)

table(analysis_raw$Reactions_Group, analysis_raw$race_w)
```

### Difference in Skin Reactions Groups by Gender and Race

No significant differences were found within Skin Reaction groups by Gender (p = 0.81) or Race (p = 0.82).

```{r}
fisher.test(analysis$Reactions_Group, analysis$gender.factor)
fisher.test(analysis$Reactions_Group, analysis$race_w)

# Chisq sanity check -- similar results
# chisq.test(analysis$Reactions_Group, analysis$gender.factor) 0.8
# chisq.test(analysis$Reactions_Group, analysis$race_w) 0.5
```

### APPENDED 11/13 Difference in Any Skin Reactions by Gender and Race

There were no significant association between device use and gender (p > 0.99) or race (p = 0.08)

```{r}
table(analysis_raw$Any_Reactions, analysis_raw$gender.factor)

fisher.test(analysis_raw$Any_Reactions, analysis_raw$gender.factor)
fisher.test(analysis_raw$Any_Reactions, analysis_raw$race_w)

```


