---
title: "ISPAD Individual Group Changes"
author: "Casey Sakamoto"
date: "4/26/2022"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(Hmisc)
library(redcapAPI)
library(knitr) # kable
library(nlme) # mixed models
library(emmeans) # contrast statement
library(longCatEDA)
library(table1)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Erin Cobry/ISPAD 2022"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ISPAD 2022"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ISPAD 2022"
}
knitr::opts_knit$set(root.dir = home_dir)

#setwd("S:/Laura/BDC/Projects/Erin Cobry/Prospective HCL and sleep study")

```

```{r data import,include=FALSE}
token = read.table("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/api_token.txt")[1,1]
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",token = token)
data = exportRecords(rcon,dates = F)
ids = data$record_id
times = data$redcap_event_name
exclude = c(101,201,102,202,103,203,110,210,119,219,121,221,131,231,141,241)

# exclude 8 dyads
data = data  %>% filter(!(record_id %in% exclude)) 

######## Basic demographics & glycemic #################
child_demographics = exportRecords(rcon,forms = "child_demographics_chart_review")
child_demographics$record_id = ids
# Get glycemic measures
glyc = exportRecords(rcon,forms = "glycemic_data")
glyc$record_id = ids
glyc= glyc %>% filter(record_id >= 200,gyl_timepoint %in% c( "Baseline", "3 Months", "6 Months")) %>% 
  mutate(timepoint = case_when(gyl_timepoint == "Baseline" ~ "baseline_day_1_arm_2",
                               gyl_timepoint == "3 Months" ~ "hcl_month_3_day_1_arm_2",
                               gyl_timepoint == "6 Months" ~ "hcl_month_6_day_1_arm_2"
                               )) %>%
  select(record_id, timepoint, hba1c, gmi,sensor_g251, sensor_181_250, sensor_70_180, sensor_55_69, sensor_u54)
# Format
child_demographics$t1d_duration = 
  as.numeric(difftime(child_demographics$consent_date,
                      child_demographics$t1d_diagnosis,units = "days"))/365.25
child_demographics = child_demographics %>% 
  select(record_id,cons_age,childgender,t1d_duration) %>%
  filter(!record_id %in% exclude)
child_demographics = 
  child_demographics[rowSums(is.na(child_demographics)) < (ncol(child_demographics)-1),]
# Add Hba1c
child_demographics = left_join(child_demographics,glyc,by = "record_id")
###############################################

############### Actigraphy #####################
# Get total sleep time, sleep efficiency, and WASO
activity = exportRecords(rcon,forms = "wrist_actigraphy")
activity$record_id = ids
activity$timepoint = times
child_activity = activity %>% 
  select(record_id,timepoint,totalsleep_avg,sleepefficency_avg,waso_avg) %>% 
  filter(!record_id %in% exclude,grepl("baseline|month_3|month_6",timepoint),
         record_id >= 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(avg_sleep = as.character(totalsleep_avg),
         totalsleep_avg = as.numeric(lubridate::hms(totalsleep_avg))/60) %>%
  filter(totalsleep_avg < 21*60)
child_activity = left_join(child_activity,child_demographics) %>% distinct()
# child_activity$sufficient = apply(child_activity,1,function(r){
#   age = as.numeric(r["cons_age"])
#   time = as.numeric(r["totalsleep_avg"])/60
#   if(age <= 2){
#     if(time >= 11 & time <= 14){suff = "Yes"} else {suff = "No"}
#   } else if (age <= 5){
#     if(time >= 10 & time <= 13){suff = "Yes"} else {suff = "No"}
#   } else if (age <= 12){
#     if(time >= 9 & time <= 12){suff = "Yes"} else {suff = "No"}
#   } else if (age <= 17){
#     if(time >= 8 & time <= 10){suff = "Yes"} else {suff = "No"}
#   } else if (age > 17) {
#     if(time >= 7){suff = "Yes"} else {suff = "No"}
#   }
# })
# child_activity$sufficient = as.factor(child_activity$sufficient)

# Same for parents
parent_activity = activity %>% 
  select(record_id,timepoint,totalsleep_avg,sleepefficency_avg,waso_avg) %>% 
  filter(!record_id %in% exclude,grepl("baseline|month_3|month_6",timepoint),
         record_id < 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(avg_sleep = as.character(totalsleep_avg),
         totalsleep_avg = as.numeric(lubridate::hms(totalsleep_avg))/60,
         sufficient = ifelse(totalsleep_avg >= 7*60,"Yes","No"))
parent_activity$sufficient = as.factor(parent_activity$sufficient)


# Import select, filter, etc.
promis_peds = exportRecords(rcon,forms = "promis_pediatric_817")
promis_peds$record_id = ids
promis_peds$timepoint = times
promis_peds = promis_peds %>% select(record_id,timepoint,falling_asleep:bad_mood_sleepy) %>%
  filter(!record_id %in% exclude,grepl("enrollment|month_3|month_6",timepoint),
         record_id >= 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(across(falling_asleep:bad_mood_sleepy,.fns = as.numeric),
         through_night = 6 - through_night) %>% 
  mutate(disturbance = rowSums(select(.,falling_asleep:tossed_and_turned)),
         impairment = rowSums(select(.,daytime:bad_mood_sleepy)))
promis_peds$timepoint = sub("enrollment_","baseline_",promis_peds$timepoint)
# Convert to T scores \Erin Cobry\ADA 2022\HCL and Sleep
peds_disturbance = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/peds_sleep_disturbance.csv")
peds_impairment = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/peds_sleep_impairment.csv")
promis_peds$disturbance = 
  peds_disturbance$T.Score[match(promis_peds$disturbance,
                                 peds_disturbance$誰..Raw.Summed.Score)]
promis_peds$impairment = 
  peds_impairment$T.Score[match(promis_peds$impairment,
                                peds_impairment$誰..Raw.Summed.Score)]
# PROMIS PARENTS
# Import select, filter, etc.
promis_parent = exportRecords(rcon,forms = "promis_parent_proxy_817")
promis_parent$record_id = ids
promis_parent$timepoint = times
promis_parent = promis_parent %>% select(record_id,timepoint,asleep:bad_mood) %>%
  filter(!record_id %in% exclude,grepl("enrollment|month_3|month_6",timepoint),
         record_id < 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(across(asleep:bad_mood,.fns = as.numeric),
         night = 6 - night) %>% 
  mutate(disturbance = rowSums(select(.,asleep:tossed)),
         impairment = rowSums(select(.,sleepy:bad_mood)))
promis_parent$timepoint = sub("enrollment_","baseline_",promis_parent$timepoint)
# Convert to T scores
parent_disturbance = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/parent_proxy_sleep_disturbance.csv")
parent_impairment = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/parent_proxy_sleep_impairment.csv")
promis_parent$disturbance = 
  parent_disturbance$T.Score[match(promis_parent$disturbance,
                                   parent_disturbance$誰..Raw.Summed.Score)]
promis_parent$impairment = 
  parent_impairment$T.Score[match(promis_parent$impairment,
                                  parent_impairment$誰..Raw.Summed.Score)]

## HFS Scoring
# parents
hfs_p = data %>% select(record_id,redcap_event_name,timepoint, hypoglycemia_fear_survey_parent_timestamp:hypoglycemia_fear_survey_parent_complete) %>% 
  filter(grepl("enrollment|month_3|month_6",redcap_event_name), record_id < 200,rowSums(is.na(.)) < ncol(.)-2)  %>% 
  mutate(across(c(snacks:having_low), as.integer)) %>% mutate(across(c(snacks:having_low),function(x){x-1})) %>%
  select(-c(hfs_12_months_severe_lows___0:hypoglycemia_fear_survey_parent_complete))
 
hfs_p = hfs_p %>% rowwise() %>% 
  mutate(behave_s = sum(c_across(snacks:nighttime)),
         worry_s = sum(c_across(recognize:having_low)),
         maintain_high_s_p = sum(c_across(c(safe_side, higher_alone, away))),
         worry_lowbg_s_p = sum(c_across(c(recognize, food, dizzy, asleep_low, low_alone, no_one, seizure, complications_low, faint, having_low))),
         worry_negsc_s_p = sum(c_across(c(embarrassing, clumsy, behavior, accident, bad_evaluation))),
         total_s = sum(c_across(behave_s:worry_s))) %>% ungroup()

hfs_p$total_z_p = (hfs_p$total_s - mean(hfs_p$total_s, na.rm = T))/sd(hfs_p$total_s, na.rm = T)

hfs_p = hfs_p %>% select(record_id, redcap_event_name, maintain_high_s_p, worry_lowbg_s_p, worry_negsc_s_p, total_z_p)
# change event and id
substring(hfs_p$redcap_event_name,nchar(hfs_p$redcap_event_name),nchar(hfs_p$redcap_event_name)) ="2"
hfs_p$record_id = sub("1", "2", hfs_p$record_id)

# children
hfs_peds = data %>% select(record_id,redcap_event_name,
                           hypoglycemia_fear_survey_childteen_timestamp:hypoglycemia_fear_survey_childteen_complete) %>% 
  filter(grepl("enrollment|month_3|month_6",redcap_event_name), record_id > 199,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(across(c(bedtime_snacks:having_low_sugar), as.integer)) %>% mutate(across(c(bedtime_snacks:having_low_sugar),function(x){x-1})) 
hfs_peds = hfs_peds %>% rowwise() %>%
  mutate(behave_s = sum(c_across(bedtime_snacks:check_often)),
         worry_s = sum(c_across(recognizing_low:having_low_sugar)),
         maintain_high_s = sum(c_across(c(high_safe, alone_higher, higher_away))),
         worry_lowbg_s= sum(c_across(c(recognizing_low, food_low, passing_out, low_asleep, low_while_alone, no_help, mistake, long_term, woozy))),
         worry_negsc_s = sum(c_across(c(embarrassing_myself, looking_clumsy, trouble_at_school, seizure_low))),
         total_s = sum(c_across(behave_s:worry_s))) %>% ungroup()

hfs_peds$total_z = (hfs_peds$total_s - mean(hfs_peds$total_s, na.rm = T))/sd(hfs_peds$total_s, na.rm = T)

hfs_peds = hfs_peds %>% select(record_id, redcap_event_name, maintain_high_s, worry_lowbg_s, worry_negsc_s, total_z)

hfs_all = full_join(hfs_peds, hfs_p)
```

```{r group derive, include = FALSE}
# put parents into psqi score >5
# score parent outcomes; 1 = > 5, 0 = <=5, na else
bl_psqi_groups = data %>% mutate(psqi_group = case_when(psqi_total > 5 ~ 1,
                                              psqi_total <= 5 ~ 0)) %>%
  filter(redcap_event_name == "enrollment_arm_1") %>%
  select(record_id, psqi_total, psqi_group)
bl_psqi_groups$psqi_group = factor(bl_psqi_groups$psqi_group, levels = c(0,1), labels = c("Parent Good Sleeper", "Parent Poor Sleeper") )
# note 127 missing baseline psqi, but has a subsequent value of 4

# add a child id for merging later
bl_psqi_groups$child_id = sub("1", "2", bl_psqi_groups$record_id)
bl_psqi_groups = bl_psqi_groups %>% select(child_id, psqi_group) 
bl_psqi_groups = rename(bl_psqi_groups, c( "record_id" = "child_id"))
```

```{r data merge, include=FALSE}
# child actigraphy, demog, & glycemic already merged
# merge w parent actigraphy (rename vars to add _p)
parent_activity$record_id = sub("1", "2", parent_activity$record_id)
parent_activity = rename(parent_activity, c( "totalsleep_avg_p" = "totalsleep_avg"))
parent_activity = rename(parent_activity, c( "sleepefficiency_avg_p" = "sleepefficency_avg"))
parent_activity = rename(parent_activity, c( "waso_avg_p" = "waso_avg"))
parent_activity = parent_activity %>% select(record_id, timepoint, totalsleep_avg_p, sleepefficiency_avg_p, waso_avg_p)
# change arm so timepoint match up
substring(parent_activity$timepoint,nchar(parent_activity$timepoint),nchar(parent_activity$timepoint)) ="2"

# merge promis, child and parent (rename vars to add _p)
promis_peds = promis_peds %>% select(record_id, timepoint, impairment, disturbance)
promis_peds$timepoint = str_replace(promis_peds$timepoint, "baseline_arm_2", "baseline_day_1_arm_2")

promis_parent = promis_parent %>% select(record_id, timepoint, impairment, disturbance)
promis_parent$record_id = sub("1", "2", promis_parent$record_id)
promis_parent = rename(promis_parent, c("impairment_p" = "impairment"))
promis_parent = rename(promis_parent, c("disturbance_p" = "disturbance"))
substring(promis_parent$timepoint,nchar(promis_parent$timepoint),nchar(promis_parent$timepoint)) ="2"
promis_parent$timepoint = str_replace(promis_parent$timepoint, "baseline_arm_2", "baseline_day_1_arm_2")
promis = full_join(promis_peds, promis_parent)

# rename hfs event to timepoint
hfs_all = rename(hfs_all, c( "timepoint" = "redcap_event_name"))
hfs_all$timepoint = str_replace(hfs_all$timepoint, "enrollment_arm_2", "baseline_day_1_arm_2")


analysis = full_join(child_activity, parent_activity) %>% distinct()
analysis = full_join(analysis, promis) %>% distinct()
analysis = full_join(bl_psqi_groups, analysis) %>% distinct()
analysis = full_join(analysis, hfs_all)

analysis = analysis %>% group_by(record_id) %>% fill(psqi_group)

# change in group over time analysis
psqi_scores = data %>% select(record_id, psqi_total, redcap_event_name) %>%
  mutate(timepoint = case_when(redcap_event_name == "enrollment_arm_1" ~ "baseline_day_1_arm_2",
                               redcap_event_name == "hcl_month_3_day_1_arm_1" ~ "hcl_month_3_day_1_arm_2",
                               redcap_event_name == "hcl_month_6_day_1_arm_1" ~ "hcl_month_6_day_1_arm_2"),
         psqi_group_d = case_when(psqi_total > 5 ~ 1,
                                              psqi_total <= 5 ~ 0)) %>% 
  filter(timepoint %in% c("baseline_day_1_arm_2","hcl_month_3_day_1_arm_2","hcl_month_6_day_1_arm_2" ))

psqi_scores$psqi_group_d = factor(psqi_scores$psqi_group_d, levels = c(0,1), labels = c("Parent Good Sleeper", "Parent Poor Sleeper") )

psqi_scores$record_id = sub("1", "2", psqi_scores$record_id)
psqi_scores = psqi_scores %>% select(record_id, timepoint, psqi_group_d, psqi_total)

analysis = full_join(analysis, psqi_scores)
```

```{r model functions, include=FALSE}
fit_ind_mod = function(outcome,df,plot = T,diagnostics = F){
  
  print(outcome)
  
  # Format timepoint
  df$timepoint = as.factor(df$timepoint)
  levels(df$timepoint) = gsub("hcl_month_6_.*","Month 6",levels(df$timepoint))
  levels(df$timepoint) = gsub("hcl_month_3_.*","Month 3",levels(df$timepoint))
  levels(df$timepoint) = gsub("baseline_.*","Baseline",levels(df$timepoint))
  df$timepoint = factor(df$timepoint, levels = c("Month 6", "Month 3", "Baseline"))
  
  

  # Fit random intercept model
  f = as.formula(paste0(outcome,"~timepoint*psqi_group"))
  mod = lme(f,
            random = ~1|record_id,
            data = df,
            na.action = na.omit)
  # Anova
  mod_anova = anova.lme(mod, type="marginal")
  print(kable(mod_anova,digits = 3,caption = "Test of Overall Effect"))

 
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ timepoint:psqi_group, adjust="tukey")
  print(kable(mod_means$emmeans,digits = 3,caption = "Timepoint Means"))
  print(kable(mod_means$contrasts[c(1,2,6,13,14,15),],digits = 3,caption = "Contrasts"))
  

   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
  # Plot
  if(plot){
    p = 
      ggplot(data = df,aes_string(x = "timepoint",y = outcome,
                                  color = "psqi_group",group = "record_id")) + 
      geom_point() + geom_line() + xlab("Timepoint") +
      theme_bw() 
    print(p)
  }
  

 
   # Check model
  if(diagnostics){
    print(check_model(mod))
  }
}
```
# Methods
Parents were stratified into good/poor sleepers by their baseline psqi score (using a cutoff of >5).

Actigraphy, Survey, and Glycemic outcomes were compared at each time point between groups using contrasts from linear mixed models with a random intercept for subject and an interaction term between timepoints and sleeper group. 

For individual group analysis, contrasts were used to estimate the change over time for each outcome from Baseline to 3 Months, Baseline to 6 Months, and 3 Months to 6 Months (P-values reported will be in this order).

# Analysis
## Compare Groups at each timepoint for glycemic, actigraphy, survey data
### Child Glycemic Measures
#### Good Sleeper Group
* TBR(<54): No significant change over time (p = .99, .99, .99)
* TBR(55-69): No significant change over time (p = .96, .99, .59)
* TIR(70-180): No significant change over time (p = .52, .99, .88)
* TAR(180-250): No significant change over time (p = .17, .18, .99)
* TAR(>251): No significant change over time (p = .99, .95, .70)
* Hba1c: No significant change over time (p = .36, .62, .99)
* GMI: No significant change over time (p = .95, .99, .90)

#### Poor Sleeper Group 
* TBR(<54): No Significant change over time (p = .12, .42, .98)
* TBR(55-69): No significant change over time (p = .87, .98, .99)
* TIR(70-180): No significant change over time (p = .97, .98, .99)
* TAR(180-250): No significant change over time (p = .99, .83, .72)
* TAR(>251): No significant change over time (p = .99, .58, .88)
* Hba1c: No significant change over time (p = .95, .98, .99)
* GMI: No significant change over time (p = .99, .93, .99)

```{r glycemic measures}
fit_ind_mod("sensor_u54", analysis,plot = F)
fit_ind_mod("sensor_55_69", analysis,plot = F)
fit_ind_mod("sensor_70_180", analysis,plot = F)
fit_ind_mod("sensor_181_250", analysis,plot = F)
fit_ind_mod("sensor_g251", analysis,plot = F)
fit_ind_mod("hba1c", analysis,plot = F)
fit_ind_mod("gmi", analysis,plot = F)
```

### Actigraphy Measures
#### Good Sleeper Group
#### Parent
* Total Sleep: No significant change over time (p = .67, .19, .97)
* Sleep Efficacy: No significant change over time (p = .99, .99, .99)
* WASO: No significant change over time (p = .55, .88, .99)

#### Child
* Total Sleep: No significant change over time (p = .56, .76, .99)
* Sleep Efficacy: No significant change over time (p = .99, .99, .99)
* WASO: No significant change over time (p = .89, .51, .99)

#### Poor Sleeper Group
#### Parent
* Total Sleep:No significant change over time (p = .99, .99, .97)
* Sleep Efficacy: No significant change over time (p = .53, .60, .99)
* WASO: No significant change over time (p = .07, .07, .99)

#### Child
* Total Sleep: No significant change over time (p = .99, .99, .99)
* Sleep Efficacy: The average sleep efficiency (child) for those in the Poor Parent Sleeper group increased by 3.06 (95%CI: 0.86, 5.26) percent from Baseline to 6 Months (p = 0.049). There was no significant change over time from Baseline to 3 Months and 3 Months to 6 Months (p = .88, .55)

* WASO: No significant change over time (p = .99, .99, .99)

```{r results Actigraphy}
# Total sleep avg (minutes), wasp, efficiency
# parent
fit_ind_mod("totalsleep_avg_p", analysis,plot = F)
fit_ind_mod("sleepefficiency_avg_p",analysis,plot = F)
fit_ind_mod("waso_avg_p", analysis,plot = F)
# children
fit_ind_mod("totalsleep_avg", analysis,plot = F)
fit_ind_mod("sleepefficency_avg",analysis,plot = F)
fit_ind_mod("waso_avg", analysis,plot = F)
```

### HFS
#### Good Sleeper Groups
#### Parent
* Total Z Score: No significant change over time (p = .61, .12, .94)
* Maintain High BG: No significant change over time (p = .96, .16, .63)
* Worry Low BG: No significant change over time (p = .52, .27, .99)
* Worry Negative Social Consequences:  No significant change over time (p = .99, .92, .54)

#### Child
* Total Z Score: No significant change over time (p = .98, .87, .99)
* Maintain High BG: No significant change over time (p = .99, .99, .98)
* Worry Low BG: No significant change over time (p = .73, .98, .99)
* Worry Negative Social Consequences:  No significant change over time (p = .84, .59, .98)

#### Poor Sleeper Groups
#### Parent
* Total Z Score: The average HFS Z Score (Parent) for those in the Poor Parent Sleeper group decreased by 0.57 (95%CI: 0.170, 0.974) from Baseline to 3 Months (p = 0.040), and by 0.628 (95%CI: 0.226, 1.03) from Baseline to 6 Months. There was no significant change over time from 3 Months to 6 Months (p = .99)
* Maintain High BG: No significant change over time (p = .99, .99, .98)
* Worry Low BG: The average "Worry Low BG" Component (Parent) for those in the Poor Parent Sleeper group decreased by 5.4 points (95%CI: 2.20, 8.55) from Baseline to 3 Months (p = 0.009), and by 4.9 points (95%CI: 1.73 8.08) from Baseline to 6 Months. There was no significant change over time from 3 Months to 6 Months (p = .99)
* Worry Negative Social Consequences: No significant change over time (p = .36, .47, .99)

#### Child
* Total Z Score: No significant change over time (p = .98, .18, .63)
* Maintain High BG: No significant change over time (p = .99, .99, .97)
* Worry Low BG: The average "Worry Low BG" Component (child) for those in the Poor Parent Sleeper group decreased by 3.27 (95%CI: 0.92, 5.62) percent from Baseline to 6 Months (p = 0.047). There was no significant change over time from Baseline to 3 Months and 3 Months to 6 Months (p = .39, .92)
* Worry Negative Social Consequences: No significant change over time (p = .10, .14, .99)
```{r results hfs}
# parent
fit_ind_mod("total_z_p", analysis,plot = F)
fit_ind_mod("maintain_high_s_p", analysis,plot = F)
fit_ind_mod("worry_lowbg_s_p", analysis,plot = F)
fit_ind_mod("worry_negsc_s_p", analysis,plot = F)

# child
fit_ind_mod("total_z", analysis,plot = F)
fit_ind_mod("maintain_high_s", analysis,plot = F)
fit_ind_mod("worry_lowbg_s", analysis,plot = F)
fit_ind_mod("worry_negsc_s", analysis,plot = F)
```

### PROMIS
#### Good Sleeper Group
#### Parent Proxy
* Disturbance: No significant change over time (p = .99, .99, .99)
* Sleep Related Impairment: No significant change over time (p = .87, .99, .99)

#### Child
* Disturbance: No significant change over time (p = .99, .99, .99)
* Sleep Related Impairment: No significant change over time (p = .99, .99, .99)

#### Poor Sleeper Group
#### Parent Proxy
* Disturbance: No significant change over time (p = .99, .94, .99)
* Sleep Related Impairment: No significant change over time (p = .41, .84, .99)

#### Child
* Disturbance: No significant change over time (p = .13, .31, .99)
* Sleep Related Impairment: No significant change over time (p = .90, .99, .98)
```{r results PROMIS}
# disturbance, impairment
# parent
fit_ind_mod("disturbance_p", analysis,plot = F)
fit_ind_mod("impairment_p",analysis,plot = F)
# children
fit_ind_mod("disturbance", analysis,plot = F)
fit_ind_mod("impairment",analysis,plot = F)
```