---
title: "ISPAD Analysis"
author: "Casey Sakamoto"
date: '2022-03-21'
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(Hmisc)
library(redcapAPI)
library(knitr) # kable
library(nlme) # mixed models
library(emmeans) # contrast statement

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/BDC/Projects/Erin Cobry/ISPAD 2022"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ISPAD 2022"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Erin Cobry/ISPAD 2022"
}
knitr::opts_knit$set(root.dir = home_dir)

setwd("S:/Laura/BDC/Projects/Erin Cobry/Prospective HCL and sleep study")

```

```{r data import,include=FALSE}
token = read.table("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/api_token.txt")[1,1]
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",token = token)
data = exportRecords(rcon,dates = F)
ids = data$record_id
times = data$redcap_event_name
exclude = c(101,201,102,202,103,203,110,210,119,219,121,221,131,231,141,241)

# exclude 8 dyads
data = data  %>% filter(!(record_id %in% exclude)) 

######## Basic demographics & glycemic #################
child_demographics = exportRecords(rcon,forms = "child_demographics_chart_review")
child_demographics$record_id = ids
# Get glycemic measures
glyc = exportRecords(rcon,forms = "glycemic_data")
glyc$record_id = ids
glyc= glyc %>% filter(record_id >= 200,gyl_timepoint %in% c( "Baseline", "3 Months", "6 Months")) %>% 
  mutate(timepoint = case_when(gyl_timepoint == "Baseline" ~ "baseline_day_1_arm_2",
                               gyl_timepoint == "3 Months" ~ "hcl_month_3_day_1_arm_2",
                               gyl_timepoint == "6 Months" ~ "hcl_month_6_day_1_arm_2"
                               )) %>%
  select(record_id, timepoint, hba1c, gmi,sensor_g251, sensor_181_250, sensor_70_180, sensor_55_69, sensor_u54)
# Format
child_demographics$t1d_duration = 
  as.numeric(difftime(child_demographics$consent_date,
                      child_demographics$t1d_diagnosis,units = "days"))/365.25
child_demographics = child_demographics %>% 
  select(record_id,cons_age,childgender,t1d_duration) %>%
  filter(!record_id %in% exclude)
child_demographics = 
  child_demographics[rowSums(is.na(child_demographics)) < (ncol(child_demographics)-1),]
# Add Hba1c
child_demographics = left_join(child_demographics,glyc,by = "record_id")
###############################################

############### Actigraphy #####################
# Get total sleep time, sleep efficiency, and WASO
activity = exportRecords(rcon,forms = "wrist_actigraphy")
activity$record_id = ids
activity$timepoint = times
child_activity = activity %>% 
  select(record_id,timepoint,totalsleep_avg,sleepefficency_avg,waso_avg) %>% 
  filter(!record_id %in% exclude,grepl("baseline|month_3|month_6",timepoint),
         record_id >= 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(avg_sleep = as.character(totalsleep_avg),
         totalsleep_avg = as.numeric(lubridate::hms(totalsleep_avg))/60) %>%
  filter(totalsleep_avg < 21*60)
child_activity = left_join(child_activity,child_demographics) %>% distinct()
# child_activity$sufficient = apply(child_activity,1,function(r){
#   age = as.numeric(r["cons_age"])
#   time = as.numeric(r["totalsleep_avg"])/60
#   if(age <= 2){
#     if(time >= 11 & time <= 14){suff = "Yes"} else {suff = "No"}
#   } else if (age <= 5){
#     if(time >= 10 & time <= 13){suff = "Yes"} else {suff = "No"}
#   } else if (age <= 12){
#     if(time >= 9 & time <= 12){suff = "Yes"} else {suff = "No"}
#   } else if (age <= 17){
#     if(time >= 8 & time <= 10){suff = "Yes"} else {suff = "No"}
#   } else if (age > 17) {
#     if(time >= 7){suff = "Yes"} else {suff = "No"}
#   }
# })
# child_activity$sufficient = as.factor(child_activity$sufficient)

# Same for parents
parent_activity = activity %>% 
  select(record_id,timepoint,totalsleep_avg,sleepefficency_avg,waso_avg) %>% 
  filter(!record_id %in% exclude,grepl("baseline|month_3|month_6",timepoint),
         record_id < 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(avg_sleep = as.character(totalsleep_avg),
         totalsleep_avg = as.numeric(lubridate::hms(totalsleep_avg))/60,
         sufficient = ifelse(totalsleep_avg >= 7*60,"Yes","No"))
parent_activity$sufficient = as.factor(parent_activity$sufficient)


# Import select, filter, etc.
promis_peds = exportRecords(rcon,forms = "promis_pediatric_817")
promis_peds$record_id = ids
promis_peds$timepoint = times
promis_peds = promis_peds %>% select(record_id,timepoint,falling_asleep:bad_mood_sleepy) %>%
  filter(!record_id %in% exclude,grepl("enrollment|month_3|month_6",timepoint),
         record_id >= 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(across(falling_asleep:bad_mood_sleepy,.fns = as.numeric),
         through_night = 6 - through_night) %>% 
  mutate(disturbance = rowSums(select(.,falling_asleep:tossed_and_turned)),
         impairment = rowSums(select(.,daytime:bad_mood_sleepy)))
promis_peds$timepoint = sub("enrollment_","baseline_",promis_peds$timepoint)
# Convert to T scores \Erin Cobry\ADA 2022\HCL and Sleep
peds_disturbance = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/peds_sleep_disturbance.csv")
peds_impairment = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/peds_sleep_impairment.csv")
promis_peds$disturbance = 
  peds_disturbance$T.Score[match(promis_peds$disturbance,
                                 peds_disturbance$誰..Raw.Summed.Score)]
promis_peds$impairment = 
  peds_impairment$T.Score[match(promis_peds$impairment,
                                peds_impairment$誰..Raw.Summed.Score)]
# PROMIS PARENTS
# Import select, filter, etc.
promis_parent = exportRecords(rcon,forms = "promis_parent_proxy_817")
promis_parent$record_id = ids
promis_parent$timepoint = times
promis_parent = promis_parent %>% select(record_id,timepoint,asleep:bad_mood) %>%
  filter(!record_id %in% exclude,grepl("enrollment|month_3|month_6",timepoint),
         record_id < 200,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(across(asleep:bad_mood,.fns = as.numeric),
         night = 6 - night) %>% 
  mutate(disturbance = rowSums(select(.,asleep:tossed)),
         impairment = rowSums(select(.,sleepy:bad_mood)))
promis_parent$timepoint = sub("enrollment_","baseline_",promis_parent$timepoint)
# Convert to T scores
parent_disturbance = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/parent_proxy_sleep_disturbance.csv")
parent_impairment = read.csv("S:/Laura/BDC/Projects/Erin Cobry/ADA 2022/HCL and Sleep/Data_Clean/parent_proxy_sleep_impairment.csv")
promis_parent$disturbance = 
  parent_disturbance$T.Score[match(promis_parent$disturbance,
                                   parent_disturbance$誰..Raw.Summed.Score)]
promis_parent$impairment = 
  parent_impairment$T.Score[match(promis_parent$impairment,
                                  parent_impairment$誰..Raw.Summed.Score)]

## HFS Scoring
# parents
hfs_p = data %>% select(record_id,redcap_event_name,timepoint, hypoglycemia_fear_survey_parent_timestamp:hypoglycemia_fear_survey_parent_complete) %>% 
  filter(grepl("enrollment|month_3|month_6",redcap_event_name), record_id < 200,rowSums(is.na(.)) < ncol(.)-2)  %>% 
  mutate(across(c(snacks:having_low), as.integer)) %>% mutate(across(c(snacks:having_low),function(x){x-1})) %>%
  select(-c(hfs_12_months_severe_lows___0:hypoglycemia_fear_survey_parent_complete))
 
hfs_p = hfs_p %>% rowwise() %>% 
  mutate(behave_s = sum(c_across(snacks:nighttime)),
         worry_s = sum(c_across(recognize:having_low)),
         maintain_high_s_p = sum(c_across(c(safe_side, higher_alone, away))),
         worry_lowbg_s_p = sum(c_across(c(recognize, food, dizzy, asleep_low, low_alone, no_one, seizure, complications_low, faint, having_low))),
         worry_negsc_s_p = sum(c_across(c(embarrassing, clumsy, behavior, accident, bad_evaluation))),
         total_s = sum(c_across(behave_s:worry_s))) %>% ungroup()

hfs_p$total_z_p = (hfs_p$total_s - mean(hfs_p$total_s, na.rm = T))/sd(hfs_p$total_s, na.rm = T)

hfs_p = hfs_p %>% select(record_id, redcap_event_name, maintain_high_s_p, worry_lowbg_s_p, worry_negsc_s_p, total_z_p)
# change event and id
substring(hfs_p$redcap_event_name,nchar(hfs_p$redcap_event_name),nchar(hfs_p$redcap_event_name)) ="2"
hfs_p$record_id = sub("1", "2", hfs_p$record_id)

# children
hfs_peds = data %>% select(record_id,redcap_event_name,
                           hypoglycemia_fear_survey_childteen_timestamp:hypoglycemia_fear_survey_childteen_complete) %>% 
  filter(grepl("enrollment|month_3|month_6",redcap_event_name), record_id > 199,rowSums(is.na(.)) < ncol(.)-2) %>%
  mutate(across(c(bedtime_snacks:having_low_sugar), as.integer)) %>% mutate(across(c(bedtime_snacks:having_low_sugar),function(x){x-1})) 
hfs_peds = hfs_peds %>% rowwise() %>%
  mutate(behave_s = sum(c_across(bedtime_snacks:check_often)),
         worry_s = sum(c_across(recognizing_low:having_low_sugar)),
         maintain_high_s = sum(c_across(c(high_safe, alone_higher, higher_away))),
         worry_lowbg_s= sum(c_across(c(recognizing_low, food_low, passing_out, low_asleep, low_while_alone, no_help, mistake, long_term, woozy))),
         worry_negsc_s = sum(c_across(c(embarrassing_myself, looking_clumsy, trouble_at_school, seizure_low))),
         total_s = sum(c_across(behave_s:worry_s))) %>% ungroup()

hfs_peds$total_z = (hfs_peds$total_s - mean(hfs_peds$total_s, na.rm = T))/sd(hfs_peds$total_s, na.rm = T)

hfs_peds = hfs_peds %>% select(record_id, redcap_event_name, maintain_high_s, worry_lowbg_s, worry_negsc_s, total_z)

hfs_all = full_join(hfs_peds, hfs_p)
```

```{r group derive, include = FALSE}
# put parents into psqi score >5
# score parent outcomes; 1 = > 5, 0 = <=5, na else
bl_psqi_groups = data %>% mutate(psqi_group = case_when(psqi_total > 5 ~ 1,
                                              psqi_total <= 5 ~ 0)) %>%
  filter(redcap_event_name == "enrollment_arm_1") %>%
  select(record_id, psqi_total, psqi_group)
bl_psqi_groups$psqi_group = factor(bl_psqi_groups$psqi_group, levels = c(0,1), labels = c("Parent Good Sleeper", "Parent Poor Sleeper") )
# note 127 missing baseline psqi, but has a subsequent value of 4

# add a child id for merging later
bl_psqi_groups$child_id = sub("1", "2", bl_psqi_groups$record_id)
bl_psqi_groups = bl_psqi_groups %>% select(child_id, psqi_group) 
bl_psqi_groups = rename(bl_psqi_groups, c( "record_id" = "child_id"))
```

```{r data merge, include=FALSE}
# child actigraphy, demog, & glycemic already merged
# merge w parent actigraphy (rename vars to add _p)
parent_activity$record_id = sub("1", "2", parent_activity$record_id)
parent_activity = rename(parent_activity, c( "totalsleep_avg_p" = "totalsleep_avg"))
parent_activity = rename(parent_activity, c( "sleepefficiency_avg_p" = "sleepefficency_avg"))
parent_activity = rename(parent_activity, c( "waso_avg_p" = "waso_avg"))
parent_activity = parent_activity %>% select(record_id, timepoint, totalsleep_avg_p, sleepefficiency_avg_p, waso_avg_p)
# change arm so timepoint match up
substring(parent_activity$timepoint,nchar(parent_activity$timepoint),nchar(parent_activity$timepoint)) ="2"

# merge promis, child and parent (rename vars to add _p)
promis_peds = promis_peds %>% select(record_id, timepoint, impairment, disturbance)
promis_peds$timepoint = str_replace(promis_peds$timepoint, "baseline_arm_2", "baseline_day_1_arm_2")

promis_parent = promis_parent %>% select(record_id, timepoint, impairment, disturbance)
promis_parent$record_id = sub("1", "2", promis_parent$record_id)
promis_parent = rename(promis_parent, c("impairment_p" = "impairment"))
promis_parent = rename(promis_parent, c("disturbance_p" = "disturbance"))
substring(promis_parent$timepoint,nchar(promis_parent$timepoint),nchar(promis_parent$timepoint)) ="2"
promis_parent$timepoint = str_replace(promis_parent$timepoint, "baseline_arm_2", "baseline_day_1_arm_2")
promis = full_join(promis_peds, promis_parent)

# rename hfs event to timepoint
hfs_all = rename(hfs_all, c( "timepoint" = "redcap_event_name"))
hfs_all$timepoint = str_replace(hfs_all$timepoint, "enrollment_arm_2", "baseline_day_1_arm_2")


analysis = full_join(child_activity, parent_activity) %>% distinct()
analysis = full_join(analysis, promis) %>% distinct()
analysis = full_join(bl_psqi_groups, analysis) %>% distinct()
analysis = full_join(analysis, hfs_all)
# test = exportRecords(rcon,forms = "glycemic_data")
```

```{r model functions, include=FALSE}
fit_mod = function(outcome,df,plot = T,diagnostics = F){
  
  # Format timepoint
  df$timepoint = as.factor(df$timepoint)
  levels(df$timepoint) = gsub("baseline_.*","Baseline",levels(df$timepoint))
  levels(df$timepoint) = gsub("hcl_month_3_.*","Month 3",levels(df$timepoint))
  levels(df$timepoint) = gsub("hcl_month_6_.*","Month 6",levels(df$timepoint))
 
  # Fit random intercept model
  f = as.formula(paste0(outcome,"~timepoint*psqi_group"))
  mod = lme(f,
            random = ~1|record_id,
            data = df,
            na.action = na.omit)
  # Plot
  if(plot){
    p = 
      ggplot(data = df,aes_string(x = "timepoint",y = outcome,
                                  color = "psqi_group",group = "record_id")) + 
      geom_point() + geom_line() + xlab("Timepoint") +
      theme_bw() 
    print(p)
  }
  
  # Anova
  mod_anova = anova.lme(mod, type="marginal")
  print(kable(mod_anova,digits = 3,caption = "Test of Overall Effect"))

 
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ factor(timepoint):psqi_group, adjust="tukey")
  print(kable(mod_means$contrasts[c(3,8,12),],digits = 3,caption = "Timepoint Means"))
 # print(kable(mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE)))

   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
 
   # Check model
  if(diagnostics){
    print(check_model(mod))
  }
}


# fit change mod
fit_change_mod = function(outcome,df){
  
  # Format timepoint
  df$timepoint = as.factor(df$timepoint)
  levels(df$timepoint) = gsub("baseline_.*","Baseline",levels(df$timepoint))
  levels(df$timepoint) = gsub("hcl_month_3_.*","Month 3",levels(df$timepoint))
  levels(df$timepoint) = gsub("hcl_month_6_.*","Month 6",levels(df$timepoint))
 
  # Fit random intercept model
  f = as.formula(paste0(outcome,"~timepoint*psqi_group"))
  mod = lme(f,
            random = ~1|record_id,
            data = df,
            na.action = na.omit)

  # Anova
  mod_anova = anova.lme(mod, type="marginal")
  print(kable(mod_anova,digits = 3,caption = "Test of Overall Effect"))

 
   # Means
  mod_means = emmeans(mod,specs=pairwise ~ factor(timepoint):psqi_group, adjust="tukey")
  mod_contrasts = pairs(mod_means[[2]], simple = "contrast")
  print(mod_contrasts[c(12,26),])
   # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
 
  }

```
# Outstanding Data Questions/Remarks
* Subj 127 is missing baseline psqi score, and also has very large value for question 4.
* Subj 138 is missing Month 3 psqi score
* Subj 127, 130 missing Month 6 psqi score

# Methods
Parents were stratified into good/poor sleepers by their baseline psqi score (using a cutoff of >5).

Actigraphy, Survey, and Glycemic outcomes were compared at each time point between groups using contrasts from linear mixed models with a random intercept for subject and an interaction term between timepoints and sleeper group. 

# Results
## Stratify Parents into Poor/Good Sleepers Before Intervention
```{r results 1}
kable(analysis %>% group_by(psqi_group) %>% summarise(n = n_distinct(record_id)))
```
Parent Good Sleeper n = 15 (0.385)
Parent Poor Sleeper n = 23 (0.590)


## Compare Groups at each timepoint for glycemic, actigraphy, survey data
### Child Glycemic Measures 
* TBR(<54): No difference in groups at each time point (p = 0.65, 0.78, 0.99)
* TBR(55-69): No difference in groups at each time point (p = 0.89, 0.93, 0.86)
* TIR(70-180): No difference in groups at each time point (p = 0.97, 0.60, 0.92)
* TAR(180-250): No difference in groups at each time point (p = 0.91, 0.68, 0.18)
* TAR(>251): No difference in groups at each time point (p = 0.78, 0.72, 0.90)
* Hba1c: No difference in groups at each time point (p = 0.99, 0.45, 0.77)
* GMI: No difference in groups at each time point (p = 0.87, 0.66, 0.99)

```{r results glycemic}
# TIR, TAB, TBR, a1c, gmi
fit_mod("sensor_u54", analysis)
fit_mod("sensor_55_69",analysis)
fit_mod("sensor_70_180", analysis)
fit_mod("sensor_181_250", analysis)
fit_mod("sensor_g251", analysis)
fit_mod("hba1c", analysis)
fit_mod("gmi", analysis)
```

### Actigraphy Measures
#### Parent
* Total Sleep: No difference in groups at each time point (p = 0.99, 0.74, 0.81)
* Sleep Efficacy: No difference in groups at each time point (p = 0.62, 0.92, 0.98)
* WASO: No difference in groups at each time point (p = 0.60, 0.82, 0.91)
#### Child
* Total Sleep: No difference in groups at each time point (p = 0.87, 0.99, 0.99)
* Sleep Efficacy: No difference in groups at each time point (p = 0.68, 0.99, 0.74)
* WASO: No difference in groups at each time point (p = 0.97, 0.95, 0.79)
```{r results Actigraphy}
# Total sleep avg (minutes), wasp, efficiency
# parent
fit_mod("totalsleep_avg_p", analysis)
fit_mod("sleepefficiency_avg_p",analysis)
fit_mod("waso_avg_p", analysis)
# children
fit_mod("totalsleep_avg", analysis)
fit_mod("sleepefficency_avg",analysis)
fit_mod("waso_avg", analysis)
```

### HFS
#### Parent
* Total Z Score: No difference in groups at each time point (p = 0.99, 0.71, 0.98)
* Maintain High BG: No difference in groups at each time point (p = 0.56, 0.99, 0.99)
* Worry Low BG: No difference in groups at each time point (p = 0.65, 0.94, 0.99)
* Worry Negative Social Consequences:  No difference in groups at each time point (p = 0.94, 0.16, 0.99)
#### Child
* Total Z Score: No difference in groups at each time point (p = 0.99, 0.99, 0.74)
* Maintain High BG: No difference in groups at each time point (p = 0.62, 0.25, 0.30)
* Worry Low BG: No difference in groups at each time point (p = 0.85, 0.94, 0.93)
* Worry Negative Social Consequences:  No difference in groups at each time point (p = 0.86, 0.99, 0.98)
```{r results hfs}
# parent
fit_mod("total_z_p", analysis)
fit_mod("maintain_high_s_p", analysis)
fit_mod("worry_lowbg_s_p", analysis)
fit_mod("worry_negsc_s_p", analysis)

# child
fit_mod("total_z", analysis)
fit_mod("maintain_high_s", analysis)
fit_mod("worry_lowbg_s", analysis)
fit_mod("worry_negsc_s", analysis)
```

### PROMIS
#### Parent Proxy
* Disturbance: No difference in groups at each time point (p = 0.86, 0.99, 0.99)
* Sleep Related Impairment: No difference in groups at each time point (p = 0.45, 0.57, 0.77)
#### Child
* Disturbance: No difference in groups at each time point (p = 0.25, 0.94, 0.84)
* Sleep Related Impairment: No difference in groups at each time point (p = 0.73, 0.98, 0.90)
```{r results PROMIS}
# disturbance, impairment
# parent
fit_mod("disturbance_p", analysis)
fit_mod("impairment_p",analysis)
# children
fit_mod("disturbance", analysis)
fit_mod("impairment",analysis)
```


## Compare Groups at each timepoint for changes in  glycemic, actigraphy, survey data at BL-3mo and BL-6mo
### Child Glycemic Measures 
* TBR(<54): No difference in groups at each time difference (p = 0.32, 0.58 )
* TBR(55-69): No difference in groups at each time difference (p = 0.99, 0.46)
* TIR(70-180): No difference in groups at each time difference (p = 0.70, 0.99 )
* TAR(180-250): No difference in groups at BL-3mo difference (p = 0.24 ) There is a significant difference in groups at the BL-6mo difference (p = 0.04)
* TAR(>251): No difference in groups at each time difference (p = 0.99, 0.30 )
* Hba1c: No difference in groups at each time difference (p = 0.19, 0.40 )
* GMI: No difference in groups at each time difference (p = 0.92, 0.86)
```{r diffdiff glyc}
# TIR, TAB, TBR, a1c, gmi
fit_change_mod("sensor_u54", analysis)
fit_change_mod("sensor_55_69",analysis)
fit_change_mod("sensor_70_180", analysis)
fit_change_mod("sensor_181_250", analysis)
fit_change_mod("sensor_g251", analysis)
fit_change_mod("hba1c", analysis)
fit_change_mod("gmi", analysis)
```

### Actigraphy Measures
#### Parent
* Total Sleep: No difference in groups at each time difference (p = 0.53, 0.65)
* Sleep Efficacy: No difference in groups at each time difference (p = 0.70, 0.50)
* WASO: No difference in groups at each time difference (p = 0.85, 0.68)
#### Child
* Total Sleep: No difference in groups at each time difference (p = 0.52, 0.62)
* Sleep Efficacy: No difference in groups at each time difference (p = 0.61, 0.06)
* WASO: No difference in groups at each time difference (p = 0.57, 0.34)
```{r diffdiff Actigraphy}
# Total sleep avg (minutes), wasp, efficiency
# parent
fit_change_mod("totalsleep_avg_p", analysis)
fit_change_mod("sleepefficiency_avg_p",analysis)
fit_change_mod("waso_avg_p", analysis)
# children
fit_change_mod("totalsleep_avg", analysis)
fit_change_mod("sleepefficency_avg",analysis)
fit_change_mod("waso_avg", analysis)
```

### HFS
#### Parent
* Total Z Score: No difference in groups at each time difference (p = 0.52, 0.91)
* Maintain High BG: No difference in groups at each time difference (p = 0.41, 0.19)
* Worry Low BG: No difference in groups at each time difference (p = 0.27, 0.72)
* Worry Negative Social Consequences:  No difference in groups at each time difference (p = 0.06, 0.74)
#### Child
* Total Z Score: No difference in groups at each time difference (p = 0.99, 0.76)
* Maintain High BG: No difference in groups at each time difference (p = 0.85, 0.90)
* Worry Low BG: No difference in groups at each time difference (p = 0.97, 0.39)
* Worry Negative Social Consequences:  No difference in groups at each time difference (p = 0.71, 0.92)
```{r diffdiff hfs}
# parent
fit_change_mod("total_z_p", analysis)
fit_change_mod("maintain_high_s_p", analysis)
fit_change_mod("worry_lowbg_s_p", analysis)
fit_change_mod("worry_negsc_s_p", analysis)

# child
fit_change_mod("total_z", analysis)
fit_change_mod("maintain_high_s", analysis)
fit_change_mod("worry_lowbg_s", analysis)
fit_change_mod("worry_negsc_s", analysis)
```

### PROMIS
#### Parent Proxy
* Disturbance: No difference in groups at each time difference(p = 0.73, 0.65)
* Sleep Related Impairment: No difference in groups at each time difference(p = 0.98, 0.89)
#### Child
* Disturbance: No difference in groups at each time difference (p = 0.24, 0.40)
* Sleep Related Impairment: No difference in groups at each time difference (p = 0.79, 0.93)
```{r diffdiff PROMIS}
# disturbance, impairment
# parent
fit_change_mod("disturbance_p", analysis)
fit_change_mod("impairment_p",analysis)
# children
fit_change_mod("disturbance", analysis)
fit_change_mod("impairment",analysis)
```

## Calculate the number of parents in the poor and good sleeper groups who meet criteria for poor and good sleep over time. 

```{r aim 4}
bl_psqi_groups_long = data %>% mutate(psqi_group = case_when(psqi_total > 5 ~ 1,
                                              psqi_total <= 5 ~ 0)) %>%
  select(record_id, psqi_total, psqi_group, redcap_event_name) %>% filter(grepl("enrollment|month_3|month_6",redcap_event_name), record_id < 200)

  # Format timepoint
  bl_psqi_groups_long$redcap_event_name = as.factor(bl_psqi_groups_long$redcap_event_name)
  levels(bl_psqi_groups_long$redcap_event_name) = gsub("enrollment_.*","Baseline",levels(bl_psqi_groups_long$redcap_event_name))
  levels(bl_psqi_groups_long$redcap_event_name) = gsub("hcl_month_3_.*","Month 3",levels(bl_psqi_groups_long$redcap_event_name))
  levels(bl_psqi_groups_long$redcap_event_name) = gsub("hcl_month_6_.*","Month 6",levels(bl_psqi_groups_long$redcap_event_name))

bl_psqi_groups_long$psqi_group = factor(bl_psqi_groups_long$psqi_group, levels = c(0,1), labels = c("Parent Good Sleeper", "Parent Poor Sleeper") )
# note 127 missing baseline psqi, but has a subsequent value of 4

kable(bl_psqi_groups_long %>% group_by(redcap_event_name) %>%count(psqi_group))
# # add a child id for merging later
# bl_psqi_groups$child_id = sub("1", "2", bl_psqi_groups$record_id)
# bl_psqi_groups = bl_psqi_groups %>% select(child_id, psqi_group) 
# bl_psqi_groups = rename(bl_psqi_groups, c( "record_id" = "child_id"))
```
 * 8 Parents transitioned from poor to good sleepers within 3 months, however 3 transitioned back to poor sleepers at month 6.
 * 2 Parents transitioned from poor to good sleepers within 6 months.