---
title: "Metabolon insulin sensitivity pilot - lipidomics"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(echo = FALSE,cache = FALSE, warning=FALSE)

library(tableone)
library(limma)
library(mixOmics)

# read in data
lipids <- read.csv("T:\\CACTI\\Metabolon insulin resistance pilot\\Lipidomics\\Data_raw\\Lipidomics_Cleaning_Final_combined.csv")
# fix typo
lipids[lipids$StudyID==2008,]$StudyID <- 1008
# convert all values to numeric
for (i in 3:ncol(lipids)){
  lipids[,i] <- as.numeric(as.character(lipids[,i]))
}

# read in clamp data
# tertiles of GIR are 3.8 and 5.4 for T1D, for control are 10.3 and 13.67
clamp <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\Metabolon insulin resistance pilot\\Lipidomics\\Data_raw\\QClampHormones_All Dataset.csv")
clamp$diabetes <- ifelse(!is.na(clamp$duration),1,0)
sampledata <- clamp[,c("StudyID","S3_GIR","diabetes")]
sampledata$GIR_cat <- ifelse((sampledata$diabetes==1 & sampledata$S3_GIR<3.8) |
                               (sampledata$diabetes==0 & sampledata$S3_GIR<10.3),1,
                             ifelse((sampledata$diabetes==1 & sampledata$S3_GIR<5.4) |
                                      (sampledata$diabetes==0 & sampledata$S3_GIR<13.67),2,3))

# merge lipidomics and clamp data
alldata <- merge(lipids,sampledata,by="StudyID",all.x = T,all.y = F)

# create separate dataframes for T1D and GIR comparisons, since we want to get rid of GIRcat=2
girdata <- alldata[alldata$GIR_cat==1 | alldata$GIR_cat==3,]

# create correct matrix for limma
# first the comparison of T1D vs. control
dia_contrast <- ifelse(alldata$diabetes==1,1,0)
dia_contrast <- cbind(rep(1,nrow(alldata)),dia_contrast)
# diabetes specific tertiles of GFR
GIR_contrast <- ifelse(girdata$GIR_cat==3,1,0)
GIR_contrast <- cbind(rep(1,nrow(girdata)),GIR_contrast)

# pull out the lipids and transpose for limma
ymat <- t(alldata[,3:147])
# model for T1D vs. control
fit_dia <- lmFit(ymat,dia_contrast)
fit_dia <- eBayes(fit_dia)
# format results
results_dia <- topTable(fit_dia,coef = 2,number = nrow(ymat))
write.csv(results_dia,"T:\\CACTI\\Metabolon insulin resistance pilot\\Reports\\lipidomics_T1D.csv")

# model for GIR
ymat <- t(girdata[,3:147])
fit_gir <- lmFit(ymat,GIR_contrast)
fit_gir <- eBayes(fit_gir)
# format results
results_gir <- topTable(fit_gir,coef = 2,number = nrow(ymat))
write.csv(results_gir,"T:\\CACTI\\Metabolon insulin resistance pilot\\Reports\\lipidomics_gir.csv")

# sPLS-DA analysis for T1D vs. control
plsda_t1d <- splsda(alldata[,c(3:147)],Y=as.factor(alldata$diabetes))
```

# Individual plot from sPLS-DA by diabetes status

```{r, echo=FALSE}
plotIndiv(plsda_t1d, comp = c(1,2),
          ind.names = FALSE, 
          ellipse = TRUE, legend = TRUE, title="")
```


```{r, echo=FALSE}
t1d_top_1 <- selectVar(plsda_t1d,comp=1)
t1d_top_2 <- selectVar(plsda_t1d,comp=2)
```

# Ranking of lipids in terms of discrimination of diabetes status, for component 1

```{r, echo=FALSE}
t1d_top_1$value
```

# Ranking of lipids in terms of discrimination of diabetes status, for component 2

```{r, echo=FALSE}
t1d_top_2$value
```


```{r, echo=FALSE}
plsda.perf.t1d = perf(plsda_t1d, validation = 'Mfold', folds = 5, 
                           progressBar = FALSE, nrepeat = 10, dist = 'max.dist',auc=TRUE)
auc_save <- plsda.perf.t1d$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf.t1d$auc$comp1["AUC.mean"])
```

# Evaluation of sPLS-DA model by diabetes status

Leave-one-out cross-validation was used to examine the performance of the sPLS-DA model: 

 Overall Error rate : `r plsda.perf.t1d$error.rate$overall[1,1]`
 AUC: `r auc_save[1]`


```{r plsda,cache=TRUE, warning=FALSE}
# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
aucs <- vector(mode="numeric", length=n_perm)
for (i in 1:n_perm) {
  Y <- sample(as.factor(alldata$diabetes),replace = F)
  plsda_res <- plsda(alldata[,c(3:147)],Y,ncomp = 1)
  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
                   validation = "loo")
  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
}
ggplot(as.data.frame(aucs),aes(x=aucs)) + 
  geom_histogram(binwidth = 0.01) + 
  geom_vline(aes(xintercept=auc_true),color="red") + 
  theme_bw() + xlab("AUC")
```

Group labels were permuted `r n_perm` times. One component PLS-DA was run for each permutation and AUC was calculated using leave one out cross-validation. Red line indicates the AUC calculated for non-permuted data.


```{r, echo=FALSE}
# sPLS-DA analysis for GIR tertile
plsda_gir <- splsda(girdata[,c(3:147)],Y=as.factor(girdata$GIR_cat))
```

# Individual plot from sPLS-DA by diabetes-specific GIR tertiles

```{r, echo=FALSE}
plotIndiv(plsda_gir, comp = c(1,2),
          ind.names = FALSE, 
          ellipse = TRUE, legend = TRUE, title="")
```


```{r, echo=FALSE}
gir_top_1 <- selectVar(plsda_gir,comp=1)
gir_top_2 <- selectVar(plsda_gir,comp=2)
```

# Ranking of lipids in terms of discrimination of diabetes-specific GIR tertiles, for component 1

```{r, echo=FALSE}
gir_top_1$value
```

# Ranking of lipids in terms of discrimination of diabetes-specific GIR tertiles, for component 2

```{r, echo=FALSE}
gir_top_2$value
```

```{r, echo=FALSE}
plsda.perf.gir = perf(plsda_gir, validation = 'Mfold', folds = 5, 
                           progressBar = FALSE, nrepeat = 10, dist = 'max.dist',auc=TRUE)
auc_save_gir <- plsda.perf.gir$auc$`comp1`[1]    
auc_true_gir <- as.numeric(plsda.perf.gir$auc$comp1["AUC.mean"])
```

# Evaluation of sPLS-DA model by diabetes-specific GIR tertiles

Leave-one-out cross-validation was used to examine the performance of the sPLS-DA model: 

 Overall Error rate : `r plsda.perf.gir$error.rate$overall[1,1]`
 AUC: `r auc_save_gir[1]`

```{r plsda2,cache=TRUE, warning=FALSE}
# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
aucs <- vector(mode="numeric", length=n_perm)
for (i in 1:n_perm) {
  Y <- sample(as.factor(girdata$GIR_cat),replace = F)
  plsda_res <- plsda(girdata[,c(3:147)],Y,ncomp = 1)
  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
                   validation = "loo")
  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
}
ggplot(as.data.frame(aucs),aes(x=aucs)) + 
  geom_histogram(binwidth = 0.01) + 
  geom_vline(aes(xintercept=auc_true),color="red") + 
  theme_bw() + xlab("AUC")
```

Group labels were permuted `r n_perm` times. One component PLS-DA was run for each permutation and AUC was calculated using leave one out cross-validation. Red line indicates the AUC calculated for non-permuted data.


