---
title: "670G Combined Analysis  "
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,dpi = 600)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Dropbox/Work/Cari Berget/670G Combined")
library(Hmisc)
library(skimr)
library(arsenal)
library(skimr)
library(nlme)
library(emmeans)
library(knitr)
library(lme4)
library(Epi)
library(car)
library(broom.mixed)
library(tidyverse)
```

```{r data import and clean,include=FALSE}
na_strings <- c("","999","-999","NULL"," ")
# Kaan's data
# Read in
adult_demographics <- 
  read.csv("./Data_Cleaned/adult_demographics.csv",
           na.strings = na_strings,stringsAsFactors = F)
adult <- 
  read.csv("./Data_Cleaned/670Excel.csv",na.strings = na_strings,
           stringsAsFactors = F)
adult_2 <- read.csv("./Data_Cleaned/Data set 670 Master.csv",
                    na.strings = na_strings,stringsAsFactors = F)
adult_new <- read.csv("./Data_Cleaned/670Excel.new55patients.csv",
                      na.strings = na_strings,stringsAsFactors = F)
adult_pump <- read.csv("./Data_Cleaned/pdf_summary.csv",
                       na.strings = na_strings,stringsAsFactors = F)
adult_pump1 <- read.csv("./Data_Cleaned/pdf_summary_additional_1.csv",
                        na.strings = na_strings,stringsAsFactors = F)
adult_pump2 <- read.csv("./Data_Cleaned/pdf_summary_additional_2.csv",
                        na.strings = na_strings,stringsAsFactors = F)
# Format and combine
adult <- adult %>% select(Patient.ID,start.date,contains("HbA1C"))
adult_2 <- adult_2 %>% 
  select(Patient.ID,start.date,Ethnicity,Insurance,contains("HbA1C"))
adult <- full_join(adult,adult_2)
adult_new <- adult_new %>% 
  select(Patient.ID,start.date,Ethnicity,Insurance,contains("HbA1C"))
adult <- full_join(adult,adult_new) 
adult <- adult %>% select(Patient.ID,start.date,Ethnicity,Insurance,
                          HbA1C.Baseline,everything())
# Get demographic data
adult <- adult %>% filter(!is.na(HbA1C.Baseline))
adult <- left_join(adult,adult_demographics,by = "Patient.ID")
adult$age <- 
  as.numeric(difftime(as.Date(adult$start.date,format = "%m/%d/%Y"),
                      as.Date(adult$DOB,format = "%m/%d/%Y"),
                      units = "days")/365.25)
# Get pdf data
adult$last_name <- tolower(sapply(strsplit(adult$LastName," "),"[[",1))
adult$first_name <- tolower(sapply(strsplit(adult$FirstName," "),"[[",1))
adult <- adult %>% 
  pivot_longer(cols = contains("HbA1C"),
               names_to = c(".value","timepoint"),
               names_sep = "\\.")
adult = adult %>% 
  add_row(.,Patient.ID = unique(adult$Patient.ID),timepoint="M18") %>%
  group_by(Patient.ID) %>% fill(start.date:first_name)
# Get timepoints
adult$start.date = lubridate::mdy(adult$start.date)
b = 30 * c(0,1,3,6,9,12,18)
t = c("Baseline","M1","M3","M6","M9","M12","M18")
adult_pump1$timepoint = apply(adult_pump1,1,function(r){
  c = as.character(r["cgm_start"])
  m = which(adult$first_name == as.character(r["first_name"]) &
              adult$last_name == as.character(r["last_name"]))[1]
  s = adult$start.date[m]
  d = as.numeric(difftime(c,s,units = "days"))
  tp = t[which.min(abs(d-b))]
  ifelse(length(tp)>0,return(tp),return(NA))
})
adult_pump2$timepoint = apply(adult_pump2,1,function(r){
  c = as.character(r["cgm_start"])
  m = which(adult$first_name == as.character(r["first_name"]) &
              adult$last_name == as.character(r["last_name"]))[1]
  s = adult$start.date[m]
  d = as.numeric(difftime(c,s,units = "days"))
  tp = t[which.min(abs(d-b))]
  ifelse(length(tp)>0,return(tp),return(NA))
})
adult_pump = rbind(adult_pump,adult_pump1)
adult_pump = rbind(adult_pump,adult_pump2)
# Merge 18 month data from Cari
adult <- left_join(adult,adult_pump,
                   by = c("last_name", "first_name", "timepoint"))
# Mutate
adult <- adult %>% group_by(Patient.ID) %>% 
  mutate(hba1c_baseline = HbA1C[timepoint == "Baseline"],
         days_from_start = 
           as.numeric(difftime(as.Date(cgm_start,format = "%Y-%m-%d"),
                               as.Date(start.date,format = "%m/%d/%Y"),
                               units = "days"))) %>% ungroup()
# Delete rows without PDF information
pdf_start <- which(colnames(adult)=="days_worn")
remove_rows <- 
  which(rowSums(is.na(adult[,pdf_start:(pdf_start+28)]))==29 &
          adult$timepoint != "Baseline")
adult <- adult[-c(remove_rows),]
# Combine am exit columns to match Cari's
adult$amexits = 
  rowSums(adult[,c(which(colnames(adult) == "no_calib_exit"):
                     which(colnames(adult) == "other_exit"))])
adult <- adult %>% 
  mutate(time_am = am_use,time_mm = manual,sensor_mean = avg_sensor,
         sensor_sd = sd_sensor,gmi = 3.31 + (sensor_mean * 0.02392),
         amexit_hyper = high_sg_exit/amexits,
         amexit_manual = user_disabled_exit/amexits,
         amexits_other = other_exit/amexits,
         amexits_day = amexits / days_worn,
         cohort = "Adult",hba1c = HbA1C,
         record_id = Patient.ID,race = Ethnicity,
         insurance = Insurance,sex = Gender,
         sensor_70 = tir_under_70,
         sensor_70_180 = tir_70_180,sensor_g250 = tir_over_250,
         sensor_g180 = tir_over_180,
         tdd = daily_dose,tdd_basal = basal_perc,	
         tdd_bolus = bolus_perc,	dailymealbolus = meals_per_day)
# Re-level factors
adult$race <- as.factor(adult$race)
levels(adult$race) <- 
  c("Non-white","Non-white","White","White","Non-white","Non-white",
    "Non-white","More Than One Race","More Than One Race",
    "Unknown / Not Reported")
adult$race <- as.character(adult$race)
adult$insurance <- as.factor(adult$insurance)
levels(adult$insurance) <- c("Public","Private")
adult$insurance <- as.character(adult$insurance)
# Remove duplicate rows
adult <- adult %>% arrange(first_name,last_name,timepoint)
dups = which(duplicated(adult[,c("record_id","timepoint")]))
adult = adult[-dups,]
# Select columns
vars <- c("record_id","cohort","timepoint","days_from_start","age",
          "sex","race","insurance","hba1c_baseline","hba1c","gmi",
          "bg_checks","time_am","time_mm","sensor_wear","sensor_mean",
          "sensor_sd","sensor_70","sensor_70_180","sensor_g180",
          "sensor_g250","tdd","tdd_basal","tdd_bolus","dailymealbolus",
          "amexits","amexit_hyper","amexit_manual",
          "amexits_day","amexits_other")
adult <- adult %>%
  select(all_of(vars))
# Cari's data
source("/Users/timvigers/GitHub/BDC-Code/Cari Berget/670G Combined/data_clean.r")
# Exclude non-670G and those who never started
data <- data[as.numeric(as.character(data$record_id)) < 116,]
never_started <- data %>% group_by(record_id) %>% 
  filter(gyl_timepoint.factor == "Baseline",is.na(automode_start)) %>% 
  .$record_id
data <- data[!(data$record_id %in% never_started),]
# Format
data$cohort <- "Peds"
data$timepoint = 
  as.character(factor(data$gyl_timepoint,levels = 0:5,
                      labels = c("Baseline","M1","M3","M6","M9","M12")))
levels(data$demographics_sex.factor) = c("F","M")
# Age
data$age <- 
  as.numeric(difftime(as.Date(data$automode_start,"%Y-%m-%d"),
                      as.Date(data$demographics_dob,"%Y-%m-%d"),
                      units = "days"))/365.25
# Baseline A1c
data <- data %>% filter(!is.na(timepoint)) %>% group_by(record_id) %>% 
  mutate(hba1c_baseline = hba1c[timepoint == "Baseline"],
         age = age[timepoint == "Baseline"]) %>% ungroup()
# Make, fill down and select columns
data <- data %>%
  fill(demographics_dob,demographics_insurance.factor,
       demographics_sex.factor,demographics_race.factor,
       automode_start) %>%
  mutate(race = demographics_race.factor,
         sex = as.character(demographics_sex.factor),
         insurance = demographics_insurance.factor,
         sensor_70 = sensor_u54 + sensor_54_69,
         sensor_g180 = sensor_181_250 + sensor_g250,
         amexit_hyper = amexit_hyper / amexits,
         amexit_manual = amexit_manual / amexits,
         amexits_other = amexits_other / amexits,
         gmi = 3.31 + (sensor_mean * 0.02392),
         days_from_start = 
           as.numeric(difftime(as.Date(date_visit,"%Y-%m-%d"),
                               as.Date(automode_start,"%Y-%m-%d"),
                               units = "days"))) %>%
  select(all_of(vars))
# Remove rows with no pump data
pdf_start <- which(colnames(data)=="time_am")
remove_rows <- which(rowSums(is.na(data[,pdf_start:(pdf_start+9)]))==10)
# One data frame
df = rbind(adult,data)
factor_vars = c("record_id","sex","race","insurance","cohort")
df[,factor_vars] <- lapply(df[,factor_vars], as.factor)
# Order timepoints
df$timepoint <- factor(df$timepoint,
                       levels = c("Baseline","M1","M3","M6","M9","M12"))
# Create age and a1c groups
df$age_group <- cut(df$age,breaks = c(0,18,26,50,Inf),right = F)
levels(df$age_group) <- 
  c("Pediatric","Young Adult","Adult","Older Adult")
df$hba1c_group <- cut(df$hba1c_baseline,breaks = c(0,7,8,Inf))
levels(df$hba1c_group) <- c("<= 7%","7% - 8.9%",">= 9%")
# Remove unidentified timepoints
df = df[!is.na(df$timepoint),]
```

Pediatric participants with no automode start date and adult participants with no PDF data were excluded from this analysis.

# Table 1: Descriptive Characteristics

```{r table 1,results='asis',message=FALSE}
demographics <- df %>% group_by(record_id) %>% 
  mutate(days_to_m1 = ifelse(length(days_from_start[timepoint=="M1"])>0,
                             days_from_start[timepoint=="M1"],NA),
         days_to_m3 = ifelse(length(days_from_start[timepoint=="M3"])>0,
                             days_from_start[timepoint=="M3"],NA),
         days_to_m6 = ifelse(length(days_from_start[timepoint=="M6"])>0,
                             days_from_start[timepoint=="M6"],NA),
         days_to_m9 = ifelse(length(days_from_start[timepoint=="M9"])>0,
                             days_from_start[timepoint=="M9"],NA),
         days_to_m12 = 
           ifelse(length(days_from_start[timepoint=="M12"])>0,
                  days_from_start[timepoint=="M12"],NA),
         days_to_m18 = ifelse(length(days_from_start[timepoint=="M8"])>0,
                              days_from_start[timepoint=="M18"],NA)) %>%
  filter(row_number()==1)
# Table 1
t1 <- tableby(cohort ~ hba1c_baseline + 
                age + fe(age_group) + sex + fe(race) + fe(insurance),
              data = demographics)
# Print
summary(t1,pfootnote = T,labelTranslations = 
          list(age = "Age (years) at HCL Start",
               age_group = "Age Group at HCL Start",
               sex = "Sex",race = "Race",
               insurance = "Insurance",
               hba1c_baseline = "HbA1c (%) at Baseline",
               hba1c_group = "HbA1c Group",
               days_to_m1 = "Days to M1 Visit",
               days_to_m3 = "Days to M3 Visit",
               days_to_m6 = "Days to M6 Visit",
               days_to_m9 = "Days to M9 Visit",
               days_to_m12 = "Days to M12 Visit"))
```

# Figure 1: HbA1c Over Time by Cohort

```{r warning=FALSE,message=FALSE,cache=TRUE}
ggplot(df,aes(x = timepoint,y = hba1c)) +
  geom_line(alpha = 0.2,aes(group = record_id,color = cohort)) +
  geom_smooth(aes(group = cohort,color = cohort),se = F) +
  theme_bw() + xlab("Timepoint") + ylab("HbA1c (%)") +
  scale_color_discrete(name = "Cohort")
```

# Figure 2: HbA1c Over Time by Age Group

```{r warning=FALSE,message=FALSE,cache=TRUE}
ggplot(df[!is.na(df$age_group),],aes(x = timepoint,y = hba1c)) +
  geom_line(alpha = 0.2,aes(group = record_id,color = age_group)) +
  geom_smooth(aes(group = age_group,color = age_group),se = F) +
  theme_bw() + xlab("Timepoint") + ylab("HbA1c (%)") +
  scale_color_discrete(name = "Age Group")
```

# Figure 3: GMI Over Time by Cohort

```{r warning=FALSE,message=FALSE,cache=TRUE}
ggplot(df,aes(x = timepoint,y = gmi)) +
  geom_line(alpha = 0.2,aes(group = record_id,color = cohort)) +
  geom_smooth(aes(group = cohort,color = cohort),se = F) +
  theme_bw() + xlab("Timepoint") + ylab("GMI") +
  scale_color_discrete(name = "Cohort")
```

# Figure 4: GMI Over Time by Age Group

```{r warning=FALSE,message=FALSE,cache=TRUE}
ggplot(df[!is.na(df$age_group),],aes(x = timepoint,y = gmi)) +
  geom_line(alpha = 0.2,aes(group = record_id,color = age_group)) +
  geom_smooth(aes(group = age_group,color = age_group),se = F) +
  theme_bw() + xlab("Timepoint") + ylab("GMI") +
  scale_color_discrete(name = "Age Group")
```

# Figure 5: Violin Plot of HCL and CGM Use by Age Group (All Timepoints)

```{r warning=FALSE,message=FALSE,}
violin_df = df %>% filter(!is.na(age_group)) %>% 
  select(record_id,age_group,time_am,sensor_wear) %>%
  reshape2::melt(.,id.vars = c("record_id","age_group"))

ggplot(violin_df,aes(x = age_group,y = value,fill = variable)) +
  geom_violin() + 
  xlab("Age Group") + ylab("% Use") +
  scale_fill_discrete(name = "",labels = c("HCL","CGM")) +
  theme_bw()
```

# Model Results

Data were analyzed using a linear mixed effects model with random intercept for subject and an AR1 correlation structure to account for repeated measures. For each outcome there is a histogram of the data followed by an ANOVA table and the mixed model results. The ANOVA table shows whether a covariate in the model was significant overall, with "Timepoint:Age Group" referring to the interaction term. The mixed model table includes the results of linear contrasts to test for differences between timepoints within a group (e.g. comparing month 3 to month 1 in the pediatric cohort). P values are adjusted for multiple comparisons using the FDR method. 

```{r contrasts}
# Peds
m1_peds <- c(1,rep(0,19))
m3_peds <- c(rep(0,1),1,rep(0,18))
m6_peds <- c(rep(0,2),1,rep(0,17))
m9_peds <- c(rep(0,3),1,rep(0,16))
m12_peds <- c(rep(0,4),1,rep(0,15))
# YA
m1_ya <- c(rep(0,5),1,rep(0,14))
m3_ya <- c(rep(0,6),1,rep(0,13))
m6_ya <- c(rep(0,7),1,rep(0,12))
m9_ya <- c(rep(0,8),1,rep(0,11))
m12_ya <- c(rep(0,9),1,rep(0,10))
# Adult
m1_adult <- c(rep(0,10),1,rep(0,9))
m3_adult <- c(rep(0,11),1,rep(0,8))
m6_adult <- c(rep(0,12),1,rep(0,7))
m9_adult <- c(rep(0,13),1,rep(0,6))
m12_adult <- c(rep(0,14),1,rep(0,5))
# Older adult
m1_older <- c(rep(0,15),1,rep(0,4))
m3_older <- c(rep(0,16),1,rep(0,3))
m6_older <- c(rep(0,17),1,rep(0,2))
m9_older <- c(rep(0,18),1,rep(0,1))
m12_older <- c(rep(0,19),1)
# List
contr <- 
  list("M3 Peds vs. M1 Peds" = m3_peds - m1_peds,
       "M6 Peds vs. M1 Peds" = m6_peds - m1_peds,
       "M9 Peds vs. M1 Peds" = m9_peds - m1_peds,
       "M12 Peds vs. M1 Peds" = m12_peds - m1_peds,
       "M3 YA vs. M1 YA" = m3_ya - m1_ya,
       "M6 YA vs. M1 YA" = m6_ya - m1_ya,
       "M9 YA vs. M1 YA" = m9_ya - m1_ya,
       "M12 YA vs. M1 YA" = m12_ya - m1_ya,
       "M3 Adult vs. M1 Adult" = m3_adult - m1_adult,
       "M6 Adult vs. M1 Adult" = m6_adult - m1_adult,
       "M9 Adult vs. M1 Adult" = m9_adult - m1_adult,
       "M12 Adult vs. M1 Adult" = m12_adult - m1_adult,
       "M3 Older vs. M1 Older" = m3_older - m1_older,
       "M6 Older vs. M1 Older" = m6_older - m1_older,
       "M9 Older vs. M1 Older" = m9_older - m1_older,
       "M12 Older vs. M1 Older" = m12_older - m1_older)
m1 <- c(1,0,0,0,0)
contr_overall <- 
  list("M3 vs. M1 " = c(0,1,0,0,0)-m1,
       "M6 vs. M1 " = c(0,0,1,0,0)-m1,
       "M9 vs. M1 " = c(0,0,0,1,0)-m1,
       "M12 vs. M1 " = c(0,0,0,0,1)-m1)
# Results column names
outnames <- c("LS Mean","SE","p value vs. M1")
```

```{r model function}
# Based on preliminary analyses, an AR1 model with random intercept is best for most variables.
ar1_mod_func <- function(outcome,data = df,diagnostics=F,p.adj = "fdr"){
  data <- data[data$timepoint != "Baseline",]
  form <- as.formula(paste0(outcome,"~","timepoint*age_group"))
  # Model with interaction
  mod_int <- lme(form,random = ~1|record_id,data=data,
                 correlation=corAR1(),na.action = na.omit)
  ano <- anova(mod_int)
  rownames(ano) <- gsub("timepoint","Timepoint",rownames(ano))
  rownames(ano) <- gsub("age_group","Age Group",rownames(ano))
  # Without age group
  form <- update.formula(form,~timepoint)
  mod <- lme(form,random = ~1|record_id,data=data,
             correlation=corAR1(),na.action = na.omit)
  emm <- suppressWarnings(emmeans(mod,~timepoint))
  combined_means <- as.data.frame(emm)[,2:3]
  c <- as.data.frame(contrast(emm,method = contr_overall))
  combined_means$p <- c(NA,c$p.value)
  rownames(combined_means) <- paste0("Overall M",c(1,3,6,9,12))
  colnames(combined_means) <- outnames
  # Means by age group
  emm <- suppressWarnings(emmeans(mod_int,~timepoint*age_group))
  # Contrasts
  c <- as.data.frame(contrast(emm,method = contr))
  # Get values
  emm <- as.data.frame(emm)
  peds_means <- emm[1:5,3:4]
  ps <- c(NA,c$p.value[1:4])
  peds <- cbind(peds_means,ps)
  rownames(peds) <- paste0("Peds M",c(1,3,6,9,12))
  ya_means <- emm[6:10,3:4]
  ps <- c(NA,c$p.value[5:8])
  ya <- cbind(ya_means,ps)
  rownames(ya) <- paste0("YAs M",c(1,3,6,9,12))
  adult_means <- emm[11:15,3:4]
  ps <- c(NA,c$p.value[9:12])
  adult <- cbind(adult_means,ps)
  rownames(adult) <- paste0("Adults M",c(1,3,6,9,12))
  older_means <- emm[16:20,3:4]
  ps <- c(NA,c$p.value[13:16])
  older <- cbind(older_means,ps)
  rownames(older) <- paste0("Older Adults M",c(1,3,6,9,12))
  # Final results table
  age_groups <- bind_rows(peds,ya,adult,older)
  colnames(age_groups) <- outnames
  final <- bind_rows(combined_means,age_groups)
  final[,3] <- p.adjust(final[,3],method = p.adj)
  # Print
  print(kable(ano,digits = 3))
  print(kable(final,digits = 3))
  if(diagnostics==T){
    invisible(qqPlot(resid(mod_int)))
  }
}
model_df <- df %>% filter(time_am > 10)
```

## HbA1c

```{r results='asis'}
hist(model_df$hba1c,main = "HbA1c")
ar1_mod_func(outcome = "hba1c",data = model_df)
```

## % Time in AM

It might be worth turning this into a categorical variable, because the distribution is fairly skewed.

```{r results='asis'}
hist(model_df$time_am,main = "AM Use")
ar1_mod_func(outcome = "time_am",data = model_df)
```

## % CGM Wear

Do we want to dichotomize (or categorize) this variable? The skewed distribution is making the model residuals look pretty bad.

```{r results='asis'}
hist(model_df$sensor_wear,main = "Sensor Wear")
ar1_mod_func(outcome = "sensor_wear",data = model_df)
```

## Figure 6: HCL and CGM Use by Age Group Over Time

```{r}
# Import means and SE (manually created this table)
fig_6_df = read.csv("./Data_Cleaned/device_use_for_plot.csv",na.strings = "")
fig_6_df = fig_6_df %>% pivot_longer(M1_MEAN:M12_SE,
                                     names_to = c("Time",".value"),
                                     names_pattern = "(.*)_(.*)")
fig_6_df$Time = factor(fig_6_df$Time,levels = c("M1","M3","M6","M9","M12"))
# Plot
ggplot(fig_6_df[fig_6_df$var != "TIR",],aes(x = Time,y = MEAN, group = group)) + 
  geom_point(aes(shape = group)) + geom_line(aes(linetype = group)) + 
  geom_errorbar(aes(ymin=MEAN-SE, ymax=MEAN+SE),width = 0.1) +
  theme_bw() + ylab("%") + xlab("Time") + 
  facet_wrap("var") +
  scale_linetype_discrete(name = "") + 
  scale_shape_discrete(name = "")
```

## Figure 6: HCL and CGM Use by Age Group Over Time

```{r}
ggplot(fig_6_df[fig_6_df$var == "TIR",],aes(x = Time,y = MEAN, group = group)) + 
  geom_point(aes(shape = group)) + geom_line(aes(linetype = group)) + 
  geom_errorbar(aes(ymin=MEAN-SE, ymax=MEAN+SE),width = 0.1) +
  theme_bw() + ylab("%") + xlab("Time") + 
  facet_wrap("var") +
  scale_linetype_discrete(name = "") + 
  scale_shape_discrete(name = "")
```

## AM Exits per Day

```{r results='asis'}
hist(model_df$amexits_day,main = "AM Exits per Day")
ar1_mod_func(outcome = "amexits_day",data = model_df)
```

## BG Checks per Day

After fixing the data this looks much better.

```{r results='asis'}
hist(model_df$bg_checks,main = "BG Checks per Day")
ar1_mod_func(outcome = "bg_checks",data = model_df)
```

## Sensor Mean

```{r results='asis'}
hist(model_df$sensor_mean,main = "Mean Sensor")
ar1_mod_func(outcome = "sensor_mean",data = model_df)
```

## % TIR < 70 mg/dL

Despite the skewed distribution, this model doesn't look too bad. Peds participant 67 has 45% time < 70 at M1, is that correct?

```{r results='asis'}
hist(model_df$sensor_70,main = "% Time < 70")
# Exclude 1408816
ar1_mod_func(outcome = "sensor_70",data = model_df)
```

## % TIR 70 - 180 mg/dL

```{r results='asis'}
hist(model_df$sensor_70_180,main = "% TIR")
ar1_mod_func(outcome = "sensor_70_180",data = model_df)
```

## % TIR 180+ mg/dL

```{r results='asis'}
hist(model_df$sensor_g180[model_df$cohort=="Peds"],
     main = "% Time 180 - 250")
ar1_mod_func(outcome = "sensor_g180",data = model_df)
```

## % TIR 250+ mg/dL

This may require some more complicated models if the data are correct, because the distribution is so skewed (the means reported here are LS means from the models, so the negative values are sign that the model isn't great). The highest percentage among the adults is 34%, but many of the peds values are in the 40s and above, with the highest being 65%. Is this correct? 

This model also looks much better when the outcome is log-transformed, but like with BG checks this makes the comparisons more complicated.

```{r results='asis'}
hist(model_df$sensor_g250,main = "% Time > 250")
ar1_mod_func(outcome = "sensor_g250",data = model_df)
```

## Total Daily Dose (TDD)

```{r results='asis'}
hist(model_df$tdd,main = "TDD")
ar1_mod_func(outcome = "tdd",data = model_df)
```

## % TDD Delivered as Bolus

```{r results='asis'}
hist(model_df$tdd_bolus,main = "% TDD as Bolus")
ar1_mod_func(outcome = "tdd_bolus",data = model_df)
```

## % TDD Delivered as Basal

```{r results='asis'}
hist(model_df$tdd_basal,main = "% TDD as Basal")
ar1_mod_func(outcome = "tdd_basal",data = model_df)
```

## Meal Boluses per Day

```{r results='asis'}
hist(model_df$dailymealbolus,main = "Meal Boluses")
ar1_mod_func(outcome = "dailymealbolus",data = model_df)
```

# Discontinuers

Discontinuers were defined as any participants with <= 10% AM use at any time point. Intermittent participants are those who had <= 10% at one timepoint and > 10% AM use at a later timepoint. Those with only baseline data or baseline plus one visit are not included in these numbers because it is unclear whether they continued.

```{r discontinuers}
discontinuers <- df %>% filter(timepoint != "Baseline") %>% 
  arrange(record_id,timepoint) %>%
  group_by(record_id) %>% 
  mutate(time_am_0 = ifelse(time_am > 10,1,0)) %>%
  summarise(discontinued = 
              ifelse(1 %in% diff(time_am_0),"Intermittent",
                     ifelse(0 %in% time_am,"Discontinued","Continued")),
            cohort = cohort[1],.groups = "drop")
# Update statuses based on Cari's manual check
cari_discont_check = read.csv("./Data_Cleaned/adult_am_time_review.csv",
                              na.strings = na_strings)
discontinuers$discontinued[match(cari_discont_check$record_id,discontinuers$record_id)] = 
  cari_discont_check$Discontinued
discontinuers$discontinued = factor(tolower(discontinuers$discontinued),
                                    levels = c("continued", "discontinued","intermittent"))
kable(table(discontinuers$discontinued,discontinuers$cohort))
```

# ROC Analysis

The outcome for this ROC analysis was defined as >= 70% TIR at month 12. 

```{r cutoffs,echo=FALSE,message=FALSE}
# ROC data
roc_df = model_df %>% group_by(record_id) %>%
  summarise(outcome = ifelse(sensor_70_180[timepoint == "M12"] >= 70,1,0),
            am = mean(time_am,na.rm = T))
# ROC 
roc <- ROC(roc_df$am,roc_df$outcome,plot = "roc",MI = F,MX=F)
```

A threshold of 71.6% automode use predicted TIR >= 70% at month 12 with sensitivity of 85.1% and specificity of 58.3%. Area under the ROC curve was 0.767.

## Clinical factors associated with meeting the AM use cutoff above

A generalized linear mixed model (GzLMM) with random intercept for subject was used to determine which demographic characteristics were associated with achieving the automode use threshold (at any timepoint).

```{r}
# Make a cutoff yes/no variable
model_df$am_cutoff = cut(model_df$time_am,c(-Inf,71.6,Inf),right = F,
                         labels = c("Under","Over"))
# Also look at >/= 3 boluses/day, 3 BG checks per day and CGM wear 
model_df$bg_checks_bin = cut(model_df$bg_checks,c(-Inf,4,Inf),right = F,
                             labels = c("< 4",">= 4"))
model_df$bolus_bin = cut(model_df$dailymealbolus,c(-Inf,4,Inf),right = F,
                             labels = c("< 4",">= 4"))
# Logistic regression
thresh_mod = glmer(am_cutoff ~ age + hba1c_baseline + bg_checks_bin +
                     bolus_bin + scale(sensor_wear,scale = F) + (1|record_id),
                   data = model_df,family = "binomial")
res = tidy(thresh_mod,conf.int = T,exponentiate=T)
res = res[-which(res$group == "record_id"),]
res$term = gsub("age","Age",res$term)
res$term = gsub("hba1c_baseline","Baseline HbA1c",res$term)
res$term = gsub("bg_checks_bin","BG Checks per Day: ",res$term)
res$term = gsub("bolus_bin","Meal Boluses per Day: ",res$term)
res$term[which(res$term == "scale(sensor_wear, scale = F)")] = "Sensor Wear (Above Mean)"

res = res %>% select(term,estimate,conf.low,conf.high,p.value)
kable(res)
```

Age at HCL start, BG checks per day, and sensor wear were significantly associated with reaching the AM use cutoff. For each 1-year increase in age at HCL start, odds of reaching the AM use threshold changed by `r round(res$estimate[res$term == "Age"]-1,3)*100`%. Each 1 unit increase above average sensor wear (`r round(mean(model_df$sensor_wear),3)`%) was associated with a `r round(res$estimate[res$term == "Sensor Wear (Above Mean)"]-1,3)*100`% change in the odds of achieving the AM use threshold. Checking BG at least 4 times per day with a `r round(res$estimate[res$term == "BG Checks per Day: >= 4"]-1,3)*100`% change in the odds of success. Baseline HbA1c is no longer significant after adjusting for sensor wear. 

# Data Overview

```{r}
skim(model_df)
```

# Numbers at each timepoint

```{r}
ns = df %>% group_by(record_id) %>% filter(timepoint != "M1") %>%
  summarise(n=length(!is.na(sensor_wear))) %>% ungroup()
table(ns$n)
```

