---
title: "Erin Cobry - telemedicine and technology"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)
library(readxl)
library(dplyr)
library(ggmap)
library(tableone)
library(geosphere)
library(sjlabelled)
library(lubridate)
library(naniar)
library(lessR)
library(tidyverse)

code_dir = ifelse(.Platform$OS.type == "unix",
                  "",
                  "T:\\Erin Cobry\\Telemedicine\\Data raw\\")
setwd(code_dir)

knitr::opts_chunk$set(echo = TRUE)

# read in general clinic data
allclinic <- read.csv("General Clinic Master List_08-26-20 Rural.csv",na.strings = c(" ",""))
allclinic$DiabetesDuration_12132019<- as.numeric(as.character(allclinic$DiabetesDuration_12132019))
allclinic <- allclinic[allclinic$DiabetesDuration_12132019>0.5,]

# read in clinic data, excluding patients in telemedicine
#pump <- read_excel("S:\\Shared Projects\\Laura\\BDC\\Projects\\Erin Cobry\\Telemedicine\\Data raw\\PedsT1DPtsSeenInLast12Month_CGM_Pump_ExcludeTelemedPts_12202019.xlsx", sheet="Pts_OnPump")
#cgm <- read_excel("S:\\Shared Projects\\Laura\\BDC\\Projects\\Erin Cobry\\Telemedicine\\Data raw\\PedsT1DPtsSeenInLast12Month_CGM_Pump_ExcludeTelemedPts_12202019.xlsx", sheet="Pts_OnCGM")
#pump_cgm <- read_excel("S:\\Shared Projects\\Laura\\BDC\\Projects\\Erin Cobry\\Telemedicine\\Data raw\\PedsT1DPtsSeenInLast12Month_CGM_Pump_ExcludeTelemedPts_12202019.xlsx", sheet="Pts_OnPump_CGM")

allclinic$Gender <- as.factor(allclinic$Gender)

levels(allclinic$Pump.Only) <- c(levels(allclinic$Pump.Only),"N")
levels(allclinic$CGM.Only) <- c(levels(allclinic$CGM.Only),"N")
levels(allclinic$Pump.and.CGM) <- c(levels(allclinic$Pump.and.CGM),"N")

allclinic[is.na(allclinic$Pump.Only),]$Pump.Only <- "N"
allclinic[is.na(allclinic$CGM.Only),]$CGM.Only <- "N"
allclinic[is.na(allclinic$Pump.and.CGM),]$Pump.and.CGM <- "N"

allclinic$pump <- ifelse(allclinic$Pump.Only %in% c("Y","y"),1,0)
allclinic$cgm <- ifelse(allclinic$CGM.Only %in% c("Y","y"),1,0)
allclinic$pump_cgm <- ifelse(allclinic$Pump.and.CGM %in% c("Y","y"),1,0)
allclinic[allclinic$pump==1 & allclinic$pump_cgm==1,]$pump <- 0 
#allclinic[allclinic$cgm==1 & allclinic$pump_cgm==1,]$cgm <- 0 

# fix insurance var in clinic dataset
levels(allclinic$InsuranceType) <- c(levels(allclinic$InsuranceType),"Private/Military")
#allclinic[allclinic$InsuranceType=="dddd",]$InsuranceType <- "Unknown"
#allclinic[allclinic$InsuranceType=="Unknown",]$InsuranceType <- "Unknown"
#allclinic[allclinic$InsuranceType=="Unkown",]$InsuranceType <- "Unknown"
#allclinic[allclinic$InsuranceType=="unknown",]$InsuranceType <- "Unknown"
#allclinic[allclinic$InsuranceType=="NULL",]$InsuranceType <- "Unknown"
allclinic[allclinic$InsuranceType %in% c("Military Plans","Private"),]$InsuranceType <- "Private/Military"
allclinic$InsuranceType <- as.factor(allclinic$InsuranceType)
allclinic$InsuranceType <- droplevels(allclinic$InsuranceType)

# make one variable for tech group
allclinic$techgroup <- NA
allclinic[allclinic$pump==1,]$techgroup <- "Pump"
allclinic[allclinic$cgm==1,]$techgroup <- "CGM"
allclinic[allclinic$pump_cgm==1,]$techgroup <- "Pump and CGM"
allclinic[is.na(allclinic$techgroup),]$techgroup <- "None"

# make numeric
allclinic$DiabetesDuration_12132019 <- as.numeric(allclinic$DiabetesDuration_12132019)
allclinic$DiabetesDuration <- allclinic$DiabetesDuration_12132019/365.25
allclinic$LastA1C_OrderValue <- as.numeric(as.character(allclinic$LastA1C_OrderValue))
allclinic$Age_12132019 <- as.numeric(as.character(allclinic$Age_12132019))
allclinic$Distance.to.Clinic.Direct..Km. <- as.numeric(as.character(allclinic$Distance.to.Clinic.Direct..Km.))
allclinic$Distance.to.Clinic.Driving..Km. <- as.numeric(as.character(allclinic$Distance.to.Clinic.Driving..Km.))

# fix rural variable
allclinic$Rural <- ifelse(is.na(allclinic$Rural),"n","r")

# one row per ppt 
demo <- unique(allclinic)

# read in telemedicine data
data <- read.csv("DiabetesTelemedicine-Cobry2020ADAAbstract_DATA_rural_distance_11-5-2020.csv")
#source("T:\\Erin Cobry\\Telemedicine\\Data raw\\DiabetesTelemedicine-Cobry2020ADAAbstract_R_2020-01-05_1613.r")

# unique IDs before removing unconsented pts
n_tele_all <- length(unique(data$study_id))

# only include patients in the file Erin sent
keep <- read.csv("Telemedicine Study Patient MRNs.csv")
data <- merge(data,keep,by="study_id",all.x=F,all.y=T)
#write.csv(data,"C:\\Temp\\Erin Cobry\\Telemedicine\\export_for_erin_telemed.csv")
n_tele_consent <- length(unique(data$study_id))

# reshape redcap dataset
enroll <- data[,c("study_id","redcap_event_name", "age","dod","zip_code_of_residence","race_chart___1","race_chart___2","race_chart___3",
  "race_chart___4","race_chart___5","race_chart___8","race_chart___9","ethnicity_chart","insurance_prior_chart","pump_prior_chart","cgmuse_prior_chart","gender",
  "Rural.","Distance.to.clinic.direct..KM...as.the.crow.flies.","Distance.to.clinic.driving..km.")]
enroll <- enroll[enroll$redcap_event_name=="enrollmentpatient_arm_1",]
first <- data[,c("study_id","redcap_event_name","date_of_first_visit","age_1st_telemed","diabetes_duration","city_of_first_visit","a1c_1st_telemedicine",
                 "insurance_first_telemed","pumptype_first_telemed","cgmuse_first_telemed","zipcode_first_visit","q149")]
first <- first[first$redcap_event_name=="first_telemedicine_arm_1",]
fup <- data[, -which(names(data) %in% c("age","dod","zip_code_of_residence","race_chart___1","race_chart___2","race_chart___3",
  "race_chart___4","race_chart___5","race_chart___8","race_chart___9","ethnicity_chart","insurance_prior_chart","pump_prior_chart","cgmuse_prior_chart",
  "date_of_first_visit","age_1st_telemed","diabetes_duration","city_of_first_visit","a1c_1st_telemedicine",
                 "insurance_first_telemed","pumptype_first_telemed","cgmuse_first_telemed","zipcode_first_visit","q149","gender",
  "Rural.","Distance.to.clinic.direct..KM...as.the.crow.flies.","Distance.to.clinic.driving..km."))]
fup$date_of_fup_visit <- dplyr::na_if(fup$date_of_fup_visit,"")
fup <- fup[!is.na(fup$date_of_fup_visit),]
#fup$date_of_fup_visit <- as.Date(as.character(fup$date_of_fup_visit), format="%m-%d-%Y")
fup$date_of_fup_visit <- mdy(as.character(fup$date_of_fup_visit))
fup <- arrange(fup,study_id,date_of_fup_visit)
# for now, we just want the most recent fup visit
fup <- fup %>% 
  group_by(study_id) %>%
  slice(which.max(date_of_fup_visit))
# merge in date of first visit so we can calculate length of tele
firstvis <- first[,c("study_id","date_of_first_visit")]
fup <- merge(fup,firstvis,by="study_id")
fup$date_of_first_visit <- as.Date(as.character(fup$date_of_first_visit),format = "%m/%d/%Y")
fup$dur_tele <- as.numeric((fup$date_of_fup_visit - fup$date_of_first_visit)/365.25)
# there are 149 patients with enrollment and 1st visit data, but only 127 with fup data
# asked Erin if we want to use 1st visit data as last visit, or exclude them
# for now, I will just merge all together and exclude no one
tele <- merge(enroll, first, by="study_id",all.x=T, all.y=T)
tele <- merge(tele, fup, by="study_id",all.x=T, all.y=T)

# create variables for pump, cgm, and pump/cgm at most recent visit
tele$pump <- ifelse(!is.na(tele$date_of_fup_visit),tele$pumptype_fup_chart,tele$pumptype_first_telemed)
tele$pump <- ifelse(!is.na(tele$pump),1,0)
# recode variable for cgm use at first visit...was coded 1=yes, 2=no, 9=don't know
tele$cgmuse_first_telemed  <- ifelse(tele$cgmuse_first_telemed==1,1,
                                     ifelse(tele$cgmuse_first_telemed==2,0,NA)) 
tele$cgm <- ifelse(!is.na(tele$date_of_fup_visit),tele$cgmuse_fup_chart,tele$cgmuse_first_telemed)
tele$pump_cgm <- ifelse(is.na(tele$pump) | is.na(tele$cgm), NA,
                        ifelse(tele$pump==1 & tele$cgm==1,1,0))
tele$pump <- as.factor(tele$pump)
tele$cgm <- as.factor(tele$cgm)
tele$pump_cgm <- as.factor(tele$pump_cgm)

# check tech variables
check <- tele[,c("study_id","date_of_fup_visit","pumptype_fup_chart","pumptype_first_telemed","pump",
                 "cgmuse_fup_chart","cgmuse_first_telemed","cgm","pump_cgm")]
write.csv(check,"T:\\Erin Cobry\\Telemedicine\\Data_clean\\checking telemedicine.csv")

# create variable for insurance at most recent visit
tele$ins <- ifelse(!is.na(tele$insurance_fup_chart),tele$insurance_fup_chart,tele$insurance_first_telemed)
# now make variable to match insurance coding in clinic dataset
tele$InsuranceType <- NA
tele[is.na(tele$ins),]$InsuranceType <- NA
tele[!is.na(tele$ins) & tele$ins %in% c(1,3),]$InsuranceType <- "Private/Military"
tele[!is.na(tele$ins) & tele$ins==2,]$InsuranceType <- "Medicaid"
tele[!is.na(tele$ins) & tele$ins==4,]$InsuranceType <- NA
tele[!is.na(tele$ins) & tele$ins==9,]$InsuranceType <- NA


# find distance to BDC
tele$zip_code_of_residence<- as.character(tele$zip_code_of_residence)
allzip <- read.csv("T:\\Erin Cobry\\Telemedicine\\Data raw\\gaz1990zipcodecentroid.csv")
allzip$fipsst <- NULL
names(allzip) <- c("zip_code_of_residence","lon_residence","lat_residence")
tele <- merge(tele,allzip,by="zip_code_of_residence",all.x = T, all.y = F)
tele$bdc_long <- 104.838
tele$bdc_lat <- 39.74801
tele$dist_to_bdc_km <- distHaversine(cbind(tele$lon_residence,tele$lat_residence),cbind(tele$bdc_long,tele$bdc_lat))/1000

# code race/ethnicity
# ethnicity is 1=Hispanic, 0=NH, 9=unk
# race is 1=White, 2=Black, 3=AI/AN, 4=Asian, 5=NH/PI, 8=Other, 9=unk
tele$numrace <- rowSums(tele[,c("race_chart___1","race_chart___2","race_chart___3","race_chart___4","race_chart___5",
                           "race_chart___8","race_chart___9")],na.rm = T)
# nobody has more than 1 race
tele$race <- NA
tele[!is.na(tele$race_chart___1) & tele$race_chart___1==1,]$race <- "White"
tele[tele$race_chart___2==1,]$race <- "Black"
tele[tele$race_chart___3==1,]$race <- "AI/AN"
tele[tele$race_chart___8==1,]$race <- "Other"
tele[tele$race_chart___9==1,]$race <- "Unknown"
tele$race_eth <- NA
tele$race_eth <- ifelse(tele$ethnicity_chart==1,"Hispanic/Latino",tele$race)
tele$diabetes_duration <- as.numeric(as.character(tele$diabetes_duration))
tele$age_1st_telemed <- as.numeric(as.character(tele$age_1st_telemed))

# code new race/eth variable
tele$race2 <- NA
tele$race2[tele$race_eth=="White"] <- "NHW"
tele$race2[tele$race_eth=="Black"] <- "NHB"
tele$race2[tele$race_eth=="Hispanic/Latino"] <- "Hispanic"
tele$race2[tele$race_eth=="Unknown"] <- "Unknown"
tele$race2[tele$race_eth=="AI/AN" | tele$race_eth=="Other"] <- "Other"

# labels
set_label(tele$race_eth) <- "Race/ethnicity"
set_label(tele$race2) <- "Race/ethnicity"
set_label(tele$age_1st_telemed) <- "Age 1st telemedicine visit"
set_label(tele$cgmuse_prior_chart) <- "CGM use prior to telemedicine"
set_label(tele$pump_prior_chart) <- "Pump use prior to telemedicine"
set_label(tele$diabetes_duration) <- "Diabetes duration"
set_label(tele$pump) <- "Pump use at most recent visit"
set_label(tele$cgm) <- "CGM use at most recent visit"
set_label(tele$pump_cgm) <- "Pump/CGM use at most recent visit"
set_label(tele$dur_tele) <- "Duration of telemedicine (yrs)"
set_label(tele$Distance.to.clinic.direct..KM...as.the.crow.flies.) <- "Distance to clinic (km) direct"
set_label(tele$Distance.to.clinic.driving..km.) <- "Driving distance (km)"

# fix A1c
tele$a1c_1st_telemedicine <- as.numeric(tele$a1c_1st_telemedicine)
tele$a1c_fup_telemed <- as.numeric(tele$a1c_fup_telemed)

# make one variable for tech group
tele$techgroup <- NA
tele[tele$pump==1,]$techgroup <- "Pump"
tele[!is.na(tele$cgm) & tele$cgm==1,]$techgroup <- "CGM"
tele[!is.na(tele$pump_cgm) & tele$pump_cgm==1,]$techgroup <- "Pump and CGM"
tele[is.na(tele$techgroup),]$techgroup <- "None"

# fix factors
tele$gender <- as.factor(tele$gender)
tele$cgmuse_first_telemed <- as.factor(tele$cgmuse_first_telemed)
tele$cgmuse_prior_chart <- ifelse(tele$cgmuse_prior_chart=='9',NA,tele$cgmuse_prior_chart)
tele$cgmuse_prior_chart <- as.factor(tele$cgmuse_prior_chart)
tele$pumptype_first_telemed <- as.factor(tele$pumptype_first_telemed)
tele$pump_prior_chart <- as.factor(tele$pump_prior_chart)
tele$Rural. <- as.factor(tele$Rural.)
tele$Rural. <- droplevels(tele$Rural.)

# exclude those w/ duration <= 6 mo
tele <- tele[tele$diabetes_duration>0.5,]
n_tele_consent_6mo <- length(unique(tele$study_id))

# make combined df with techgroup and A1c
combclinic <- allclinic[,c("techgroup","LastA1C_OrderValue","Distance.to.Clinic.Driving..Km.","Distance.to.Clinic.Direct..Km.",
                           "Rural","InsuranceType")]
combclinic$hba1c <- combclinic$LastA1C_OrderValue
combclinic$LastA1C_OrderValue <- NULL
combclinic$Distance.to.clinic.driving..km. <- combclinic$Distance.to.Clinic.Driving..Km.
combclinic$Distance.to.Clinic.Driving..Km. <- NULL
combclinic$Distance.to.clinic.direct..KM...as.the.crow.flies. <- combclinic$Distance.to.Clinic.Direct..Km.
combclinic$Distance.to.Clinic.Direct..Km. <- NULL
combclinic$Rural. <- combclinic$Rural
combclinic$Rural <- NULL
combclinic$group <- "Clinic"
combtele <- tele[,c("techgroup","a1c_fup_telemed","Distance.to.clinic.direct..KM...as.the.crow.flies.","Distance.to.clinic.driving..km.",
                    "Rural.","InsuranceType")]
combtele$hba1c <- combtele$a1c_fup_telemed
combtele$a1c_fup_telemed <- NULL
combtele$group <- "Telemedicine"
comb <- rbind(combclinic,combtele)
comb$hba1c <- as.numeric(as.character(comb$hba1c))
comb$anypump <- as.factor(ifelse(comb$techgroup %in% c("Pump","Pump and CGM"),1,0))
comb$anycgm <- as.factor(ifelse(comb$techgroup %in% c("CGM","Pump and CGM"),1,0))
comb$pumponly <- as.factor(ifelse(comb$techgroup=="Pump",1,0))
comb$cgmonly <- as.factor(ifelse(comb$techgroup=="CGM",1,0))
# new variable for tech or no tech
comb$tech <- NA
comb$tech <- ifelse(comb$anypump==1 | comb$anycgm==1,"Uses tech","Doesn't use tech")

# New variable for pump and CGM yes/no
comb$pump_cgm <- as.factor(ifelse(comb$techgroup=="Pump and CGM",1,0))

# create df with pump, cgm, pump_cgm in both groups
a <- allclinic[,c("pump","cgm","pump_cgm","InsuranceType")]
a$group <- "Clinic"
b <- tele[,c("pump","cgm","pump_cgm","InsuranceType")]
b$group <- "Telemedicine"
twogroup <- rbind(a,b)

# table comparing tech use in the two groups
techtable <- CreateTableOne(vars=c("techgroup","anypump","anycgm","pumponly","cgmonly","pump_cgm","tech",
                                   "InsuranceType","Distance.to.clinic.driving..km.",
                                   "Distance.to.clinic.direct..KM...as.the.crow.flies.","Rural." ), data=comb, strata="group", test=TRUE)
techtable <- print(techtable,varLabels=TRUE,showAllLevels=TRUE)

# demographics in telemedicine patients
demovars <- c("age_1st_telemed","race2","gender","cgmuse_prior_chart","pump_prior_chart","diabetes_duration","InsuranceType",
              "a1c_1st_telemedicine","dur_tele","Distance.to.clinic.driving..km.","Distance.to.clinic.direct..KM...as.the.crow.flies.","Rural.")
nonnorm <- c("diabetes_duration","dur_tele")
t1 <- CreateTableOne(vars=demovars, data=tele)
t1 <- print(t1,varLabels=TRUE,showAllLevels=TRUE,nonnorm=nonnorm)

# demographics in BDC clinic
clinicdemovars <- c("Age_12132019","Gender","DiabetesDuration_12132019","LastA1C_OrderValue","Race_Eithnicity",
                    "InsuranceType","Distance.to.Clinic.Direct..Km.","Distance.to.Clinic.Driving..Km.","Rural" )
nonnorm <- c("DiabetesDuration_12132019")
t1a <- CreateTableOne(vars=clinicdemovars,data=allclinic)
t1a <- print(t1a,varLabels=T,showAllLevels=T,nonnorm=nonnorm)

# compare characteristics in tele pts who use tech vs. not
tele$tech <- NA
tele$tech <- ifelse(tele$cgm==1 | tele$pump==1,"Uses tech","Doesn't use tech")
tele[is.na(tele$tech),]$tech <- "Doesn't use tech"

compvars <- c("age_1st_telemed","race2","gender","cgmuse_prior_chart","pump_prior_chart","diabetes_duration","InsuranceType",
              "a1c_1st_telemedicine","a1c_fup_telemed","Distance.to.clinic.direct..KM...as.the.crow.flies.","Distance.to.clinic.driving..km.",
              "Rural.")
nonnorm <- "diabetes_duration"
t3 <- CreateTableOne(vars=compvars, data=tele, strata = "tech", test=TRUE)
t3 <- print(t3,varLabels=TRUE,showAllLevels=TRUE,nonnorm=nonnorm)

nonnorm <- "a1c_fup_telemed"
t4 <- CreateTableOne(vars="a1c_fup_telemed", data=tele, strata = "techgroup", test=TRUE)
t4 <- print(t4,varLabels=TRUE,showAllLevels=TRUE,nonnorm=nonnorm)

allclinic$LastA1C_OrderValue <- as.numeric(as.character(allclinic$LastA1C_OrderValue))
nonnorm <- "LastA1C_OrderValue"
t5 <- CreateTableOne(vars="LastA1C_OrderValue", data=allclinic, strata = "techgroup", test=TRUE)
t5 <- print(t5,varLabels=TRUE,showAllLevels=TRUE,nonnorm=nonnorm)

nonnorm <- "hba1c"
t6 <- CreateTableOne(vars=c("hba1c"), data=comb, strata = "techgroup", test=TRUE)
t6 <- print(t6,varLabels=TRUE,showAllLevels=TRUE,nonnorm=nonnorm)

plotN <- comb %>% group_by(techgroup,group) %>% summarize(N=n())
  
p <- ggplot(comb, aes(x=techgroup, y=hba1c, fill=group)) + geom_boxplot() + xlab("") + ylab("HbA1c (%)") + scale_fill_grey() + theme_classic() + guides(fill=guide_legend(title=""))

```

##  Background

The goal of this analysis is to describe technology uptake in patients who use telemedicine as a routine part of diabetes care.

##  Methods

Only consented patients in the telemedicine REDCap database were included in this analysis.  There were a total of `r n_tele_all` patients in the REDCap database, of whom `r n_tele_consent` consented.  In this version of the report, patients with <=6 months diabetes duration were excluded, leaving a total of `r n_tele_consent_6mo` patients in the analysis.

Pump, CGM, and pump/cgm use were defined as use at the most recent visit (either first telemedicine visit or follow-up visits).  Insurance type and technology use in the telemedicine cohort was compared to the BDC clinic population using chi-square tests.  Insurance types were not significantly different in the clinic and telemedicine cohorts, so all models are unadjusted.

Distance to the BDC was estimated using the great circle distance between zip codes.  Characteristics of telemedicine patients who use technology were compared to those who don't using t-tests or Mann-Whitney tests, and chi-square tests.

##  Results

```{r, echo=FALSE}
kable(t1,caption = "Table 1.  Demographics in the telemedicine cohort.  Statistics presented are mean (SD), median [25th %ile, 75th %ile], or N(%).")
```
\newpage

```{r, echo=FALSE}
kable(t1a,caption = "Table 1a.  Demographics in the BDC clinic patients.  Statistics presented are mean (SD), median [25th %ile, 75th %ile], or N(%).")
```
\newpage

```{r, echo=FALSE}
kable(techtable,caption = "Table 2.  Comparison of the telemedicine and clinic cohorts.  Statistics presented are mean (SD), median [25th %ile, 75th %ile], or N(%).")
```
\newpage

```{r, echo=FALSE}
kable(t3,caption = "Table 3.  Comparison of telemedicine patients who use technology vs. those who do not use technology. Statistics presented are mean (SD), median [25th %ile, 75th %ile], or N(%).")
```
\newpage

```{r, echo=FALSE}
kable(t4,caption = "Table 4.  Comparison of HbA1c in telemedicine patients by technology group. Statistics presented are mean (SD), median [25th %ile, 75th %ile], or N(%).")
```
\newpage

```{r, echo=FALSE}
kable(t5,caption = "Table 5.  Comparison of HbA1c in clinic patients by technology group. Statistics presented are mean (SD), median [25th %ile, 75th %ile], or N(%).")
```
\newpage

```{r, echo=FALSE}
kable(t6,caption = "Table 6.  Comparison of HbA1c and distance to clinic in all patients by technology group. Statistics presented are mean (SD), median [25th %ile, 75th %ile], or N(%).")
```

\newpage

```{r, echo=FALSE}
p
```
