---
title: "PIU! Baseline Paper"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(tidyverse)
library(arsenal)
library(skimr)
library(knitr)
library(coefficientalpha)
set.seed(1017)
```

```{r data}
# Import questionnaire and demographic data
child <- read.csv("./Kim Driscoll/PIU Baseline Paper/Data_Cleaned/child.csv",na.strings = "")
parent <- read.csv("./Kim Driscoll/PIU Baseline Paper/Data_Cleaned/parent.csv",
                   na.strings = "")
# Fill missing IDs
child <- child %>% fill(ID,.direction = "down")
# FL vs. CO
child$Site <- NA
child$Site[child$ID < 300] <- "FL"
child$Site[child$ID >= 300] <- "CO"
parent$Site <- NA
parent$Site[parent$ID < 300] <- "FL"
parent$Site[parent$ID >= 300] <- "CO"
# Pump variables
pump <- read.csv("./Kim Driscoll/Pump Variables/Data_Cleaned/summary.csv",
                 na.strings = "")
colnames(pump)[which(colnames(pump)=="study_visit")] <- "Study.Visit"
colnames(pump)[which(colnames(pump)=="subject_id")] <- "ID"
pump$Study.Visit <- as.numeric(sub("T","",pump$Study.Visit))
# Merge
child <- full_join(child,pump,by = c("ID", "Study.Visit"))
# Pump variables per day
child <- child %>%
  mutate(readings_per_day = total_readings / days_worn,
         carbs_per_day = total_carbs / days_worn,
         boluses_per_day = total_bolus / days_worn,
         bg_150_249_perc_bolus = bg_150_249_with_bolus_only / total_150_249,
         bg_150_249_perc_bolus_carb = bg_150_249_with_bolus_carb / total_150_249,
         bg_above_250_perc_bolus = bg_above_250_with_bolus_only / total_above_250,
         bg_above_250_perc_bolus_carb = bg_above_250_with_bolus_carb / total_above_250)
# Remove participant 216
child <- child[child$ID != 216,]
parent <- parent[parent$ID != 216,]
# Reverse scored psych items
# CDI
rev_cdi <- paste0("CDI",c(2,6,7,9,10,12,14,15,17,20,23,24,26,27))
child[,rev_cdi] <- lapply(child[,rev_cdi], function(x){2-x})
# CES-D
rev_ces <- paste0("CESD",c(4,8,12,16))
parent[,rev_ces] <- lapply(parent[,rev_ces], function(x){3-x})
# WECARE
rev_wecare <- paste0("WECARE",c(24,25,26,27,28,29,30,31,32,35,37))
parent[,rev_wecare] <- lapply(parent[,rev_wecare], function(x){4-x})
# Meter
meter_cols <- grep("METER",colnames(child))
child[,meter_cols] <- lapply(child[,meter_cols], function(x){
  x <- as.numeric(x)
  x[which(x==2)]<-0
  x
})
# PEDS-QL
# Reverse score and convert to 0-100 scale
peds_ql_cols <- colnames(child)[grep("PEDSQL",colnames(child))]
child[,peds_ql_cols] <- lapply(child[,peds_ql_cols], function(x){
  x <- as.numeric(x)
  x[which(x==0)]<-100
  x[which(x==1)]<-75
  x[which(x==2)]<-50
  x[which(x==3)]<-25
  x[which(x==4)]<-0
  x
})
# Means
child$pedsql_d <- apply(child[,grep("PEDSQL_D",colnames(child))],1,mean)
child$pedsql_tx1 <- apply(child[,grep("PEDSQL_TX1",colnames(child))],1,mean)
child$pedsql_tx2 <- apply(child[,grep("PEDSQL_TX2",colnames(child))],1,mean)
child$pedsql_w <- apply(child[,grep("PEDSQL_W",colnames(child))],1,mean)
child$pedsql_c <- apply(child[,grep("PEDSQL_C",colnames(child))],1,mean)
child$pds_boys <- apply(child[,paste0("PDS",1:5)],1,mean)
child$pds_girls <- apply(child[,paste0("PDS",c(1:3,6,7))],1,mean)
# Pump knowledge
pump_delete <- c("tdk_1d","tdk_1e","tdk_2d","tdk_3c","tdk_3d",
                 "tdk_4a",
                 "tdk_4b","tdk_5b","tdk_5e","tdk_6e","tdk_7d","tdk_8d",
                 "tdk_9e","tdk_10d","tdk_11d","tdk_12d","tdk_13e",
                 "tdk_14d","tdk_15d","tdk_16e","tdk_17a","tdk_17b",
                 "tdk_17e","tdk_18e","tdk_19e","tdk_20d","tdk_21d")
child[,pump_delete] <- NULL
child$tdk <- apply(child[,grep("tdk_",colnames(child))],1,sum)
# Get treatment group for parents
parent <- left_join(parent,child[,c("ID","Study.Visit","Treatment")],
                    by = c("ID", "Study.Visit"))
# Parent pump knowledge
parent[,pump_delete] <- NULL
parent$tdk <- apply(parent[,grep("tdk_",colnames(parent))],1,sum)
```

# Table 1: Self-Management Behaviors at Baseline

There was only one missing value (for participant 220, who had $\frac{0}{0}$ for BG $\geq$ 250), so simple tests were used for this table. Holly, I have one additional intervention participant, any idea why that might be?

## Child

```{r baseline self management,results='asis'}
# Baseline data
child_baseline <- child[child$Study.Visit==1,]
parent_baseline <- parent[parent$Study.Visit==1,]
# Treatment variable - make as usual the reference group
child_baseline$Treatment <- factor(child_baseline$Treatment,levels = c("Treatment as Usual","Tailored Feedback"))
# Variables
smvars <- c("days_worn","carbs_per_day","boluses_per_day",
            "bg_150_249_perc_bolus","bg_150_249_perc_bolus_carb",
            "bg_above_250_perc_bolus","bg_above_250_perc_bolus_carb")
# Table
t1 <- tableby(Treatment~kwt(days_worn)+carbs_per_day+boluses_per_day+
                bg_150_249_perc_bolus+bg_150_249_perc_bolus_carb+
                bg_above_250_perc_bolus+bg_above_250_perc_bolus_carb,
              data = child_baseline)
summary(t1,pfootnote=T)
```

# Table 2: Psychosocial Variables at Baseline

```{r fiml function,include=FALSE}
a <- list("alpha"=NA)
# lavaan models functions
fiml_latent <- function(outcome_name,prefix,data,
                        numbers = NULL,by = "Treatment"){
  vars <- colnames(data)[grep(prefix,colnames(data))]
  if(!is.null(numbers)){
    vars <- paste0(prefix,numbers)
  }
  
  vars_form <- paste(vars,collapse = " + ")
  form <- paste(outcome_name," =~ ",vars_form,"\n",
                outcome_name," ~ ",by)
  # Fit SEM
  mod <- sem(form,data = data,missing = "fiml")
  # Get mean sum of items
  control <- rep(1,length(vars))
  treat <- rep(1,length(vars)+1)
  means  <- fitted(mod)$mean 
  mean_control <- control %*% head(means,-1)
  mean_treat <- treat %*% means
  # Get se of item sum
  covar_mat_names <- c(paste0(vars,"~1"),
                       paste0(outcome_name,"~",by))
  vcov <- vcov(mod)
  vcov <- vcov[covar_mat_names,covar_mat_names]
  se_control <- sqrt(control %*% 
                       vcov[1:length(control),1:length(control)] %*% 
                       control)
  se_treat <- sqrt(treat %*% vcov[1:length(treat),1:length(treat)] %*% 
                     treat)
  # Regression estimates
  s <- as.data.frame(summary(mod)$PE)
  p <- s[which(s$lhs == outcome_name & s$rhs == by),"pvalue"]
  # Overall
  form <- paste(outcome_name," =~ ",vars_form)
  mod <- sem(form,data,missing = "fiml")
  overall <- rep(1,length(vars))
  mean_overall <- overall %*% fitted(mod)$mean
  covar_mat_names <- c(paste0(vars,"~1"))
  vcov <- vcov(mod)
  vcov <- vcov[covar_mat_names,covar_mat_names]
  se_overall <- sqrt(overall %*% vcov %*% overall)
  # Use EM robust to get alpha value
  drop <- which(rowSums(is.na(data[,vars])) == length(vars))
  if(length(drop)>0){
    try(a <- alpha(data[,vars],test = F,drop = drop,varphi = 0,
                   silent = T))
  } else {
    try(a <- alpha(data[,vars],test = F,varphi = 0,silent = T))
  }
  # Return
  res <- c(mean_overall,se_overall,mean_control,se_control,
           mean_treat,se_treat,p,a$alpha)
  res <- as.data.frame(t(res))
  colnames(res) <- res_names
  return(res)
}
fiml <- function(outcome,prefix,data,
                 by = "Treatment",numbers=NULL){
  vars <- colnames(data)[grep(prefix,colnames(data))]
  if(!is.null(numbers)){
    vars <- paste0(prefix,numbers)
  }
  form <- paste0(outcome,"~",by)
  mod <- sem(form,data,missing = "fiml")
  # Means by groups
  control <- c(1,0)
  treat <- c(1,1)
  means  <- fitted(mod)$mean 
  mean_control <- control %*% means
  mean_treat <- treat %*% means
  # SE
  covar_mat_names <- c(paste0(outcome,"~1"),paste0(outcome,"~",by))
  vcov <- vcov(mod)
  vcov <- vcov[covar_mat_names,covar_mat_names]
  se_control <- sqrt(control %*% vcov %*% control)
  se_treat <- sqrt(treat %*% vcov %*% treat)
  # Regression estimates
  s <- as.data.frame(summary(mod)$PE)
  p <- s[which(s$lhs == outcome & s$rhs == by),"pvalue"]
  # Overall
  form <- paste0(outcome,"~1")
  mod <- sem(form,data,missing = "fiml")
  mean_overall <- fitted(mod)$mean
  covar_mat_names <- paste0(outcome,"~1")
  vcov <- vcov(mod)
  vcov <- vcov[covar_mat_names,covar_mat_names]
  se_overall <- sqrt(1 %*% vcov %*% 1)
  # Use EM robust to get alpha value
  drop <- as.numeric(which(rowSums(is.na(data[,vars])) == length(vars)))
  if(length(drop)>0){
    try(a <- alpha(data[,vars],test = F,drop = drop,varphi = 0,
                   silent = T))
  } else {
    try(a <- alpha(data[,vars],test = F,varphi = 0,silent = T))
  }
  # Return
  res <- c(mean_overall,se_overall,mean_control,se_control,
           mean_treat,se_treat,p,a$alpha)
  res <- as.data.frame(t(res))
  colnames(res) <- res_names
  return(res)
}
```

## Results

Questions 1D, 3C, 5B, and 17A were excluded from the child TDK analysis due to 0 variance. 

### By Treatment Group

#### Child

```{r child models,include=FALSE}
res_names <- c("Mean Overall","SE Overall",
               "Mean Treatment as Usual","SE Treatment as Usual",
               "Mean Tailored Feedback","SE Tailored Feedback",
               "P Value","Alpha")
# CDI
cdi <- fiml_latent("depression","CDI",child_baseline)
# FOH
# Maintain
foh_maintain <- 
  fiml_latent("maintain","FOH_C",child_baseline,
              numbers = c(3,4,7))
# Helplessness
foh_helpless <- 
  fiml_latent("helpless","FOH_C",child_baseline,
              numbers = c(11,12,13,14,16,18,19,22,23))
# Social
foh_social <- 
  fiml_latent("social","FOH_C",child_baseline,
              numbers = c(15,17,20,21,25))
# PEDS QL - score is a mean rather than a sum, so slightly different approach is needed
peds_ql_d <- fiml("pedsql_d","PEDSQL_D",child_baseline)
peds_ql_tx1 <- fiml("pedsql_tx1","PEDSQL_TX1",
                    child_baseline)
peds_ql_tx2 <- fiml("pedsql_tx2","PEDSQL_TX2",
                    child_baseline)
peds_ql_w <- fiml("pedsql_w","PEDSQL_W",child_baseline)
peds_ql_c <- fiml("pedsql_c","PEDSQL_C",child_baseline)
# PDS
pds_boys <- fiml("pds_boys","PDS",numbers = c(1:5),
                 child_baseline[child_baseline$Sex==1,])
pds_girls <- fiml("pds_girls","PDS",numbers = c(1:3,6,7),
                  child_baseline[child_baseline$Sex==0,])
# Meter skills
meter <- fiml_latent("meter","METER",
                     child_baseline,numbers = c(1:6,9:18))
# Pump knowledge - Delete 4a, 4b, and those without variance
pump <- fiml("tdk","tdk_",child_baseline)
# BGMIPR
bgmipr <- fiml_latent("bgmipr","BGMIPR",
                      child_baseline,numbers = c(1:16))
# DFRQ
dfrq <- fiml_latent("dfrq","DFRQ",
                    child_baseline,numbers = c(1:17))
```

```{r child results}
child_results <- 
  bind_rows(cdi,foh_maintain,foh_helpless,foh_social, 
            peds_ql_d,peds_ql_tx1,peds_ql_tx2,peds_ql_w,peds_ql_c,
            pds_boys,pds_girls,meter,pump,bgmipr,dfrq)
rownames(child_results) <- 
  c("Depression (CDI2)","FOH (CHFS) – Maintain High BG",
    "FOH (CHFS) – Worry/Helplessness",
    "FOH (CHFS) – Worry/Social Consequences",
    "PedsQL Diabetes","PedsQL Treatment 1","PedsQL Treatment 2",
    "PedsQL Worry","PedsQL Communication","PDS - Boys",
    "PDS - Girls","Meter","tdk","BGMIPR","DFRQ")
kable(child_results,row.names = T,digits = 3)
```

#### Parent

```{r parent models,include=FALSE}
# CESD
cesd <- fiml_latent("depression","CESD",parent_baseline)
# FOH
# Maintain
foh_maintain <- fiml_latent("maintain","FOH_P",parent_baseline,
                            numbers = c(3,4,7))
# Helplessness
foh_helpless <- fiml_latent("helpless","FOH_P",parent_baseline,
                            numbers = c(13,14,15 ,17,20,23,24,25,26))
# Social
foh_social <- fiml_latent("social","FOH_P",parent_baseline,
                          numbers = c(16,18,19,21,22))
# WECARE
wecare_well_being <- 
  fiml_latent("wellbeing","WECARE",
              parent_baseline,
              numbers = c(1,2,4,14,15,16,17,18,19,20,21,22,23))
wecare_satisfaction <- 
  fiml_latent("satisfaction","WECARE",
              parent_baseline,numbers = c(3,28,29,30,31,32,35,36,37))
wecare_acceptance <- fiml_latent("ease","WECARE",parent_baseline,
                                 numbers = c(5,6,7,8,9,10))
wecare_ease <- fiml_latent("ease","WECARE",parent_baseline,
                           numbers = c(11,12,13,24,25,26,27,33,34))
# Pump knowledge - Delete 4a, 4b, and those without variance
pump <- fiml("tdk","tdk_",parent_baseline)
# BGMIPR
bgmipr <- fiml_latent("bgmipr","BGMIPR",parent_baseline,
                      numbers = c(1:16))
# DFRQ
dfrq <- fiml_latent("dfrq","DFRQ",parent_baseline,numbers = c(1:17))
```

```{r parent results}
parent_results <- 
  bind_rows(cesd,foh_maintain,foh_helpless,foh_social,
            wecare_well_being,wecare_satisfaction,wecare_acceptance,
            wecare_ease,pump,bgmipr,dfrq)
rownames(parent_results) <- 
  c("Depression (CESD)","FOH (PHFS) – Maintain High BG",
    "FOH (PHFS) – Worry/Helplessness","FOH (PHFS) – Worry/Social Consequences",
    "Psychosocial well-being","Treatment satisfaction","Acceptance of insulin administration",
    "Ease of insulin use","Pump Knowledge","BGMIPR","DFRQ")
kable(parent_results,row.names = T,digits = 3)
```

### By Site

Meter 5 was excluded for children due to lack of variability.

#### Child

```{r child models by site,include=FALSE}
res_names <- c("Mean Overall","SE Overall",
               "Mean Colorado","SE Colorado",
               "Mean Florida","SE Florida",
               "P Value","Alpha")
# CDI
cdi <- fiml_latent("depression","CDI",child_baseline,by = "Site")
# FOH
# Maintain
foh_maintain <- 
  fiml_latent("maintain","FOH_C",child_baseline,by = "Site",
              numbers = c(3,4,7))
# Helplessness
foh_helpless <- 
  fiml_latent("helpless","FOH_C",child_baseline,by = "Site",
              numbers = c(11,12,13,14,16,18,19,22,23))
# Social
foh_social <- 
  fiml_latent("social","FOH_C",child_baseline,by = "Site",
              numbers = c(15,17,20,21,25))
# PEDS QL - score is a mean rather than a sum, so slightly different approach is needed
peds_ql_d <- fiml("pedsql_d","PEDSQL_D",child_baseline,by = "Site")
peds_ql_tx1 <- fiml("pedsql_tx1","PEDSQL_TX1",child_baseline,by = "Site")
peds_ql_tx2 <- fiml("pedsql_tx2","PEDSQL_TX2",child_baseline,by = "Site")
peds_ql_w <- fiml("pedsql_w","PEDSQL_W",child_baseline,by = "Site")
peds_ql_c <- fiml("pedsql_c","PEDSQL_C",child_baseline,by = "Site")
# PDS
pds_boys <- fiml("pds_boys","PDS",child_baseline[child_baseline$Sex==1,],
                 numbers = c(1:5),by = "Site")
pds_girls <- fiml("pds_girls","PDS",
                  child_baseline[child_baseline$Sex==0,],
                  numbers = c(1:3,6,7),by = "Site")
# Meter skills
meter <- fiml_latent("meter","METER",child_baseline,
                     numbers = c(1:4,6,9:18),by = "Site")
# Pump knowledge - Delete 4a, 4b, and those without variance
pump <- fiml("tdk","tdk_",child_baseline,by = "Site")
# BGMIPR
bgmipr <- fiml_latent("bgmipr","BGMIPR",child_baseline,
                      by = "Site",numbers = c(1:16))
# DFRQ
dfrq <- fiml_latent("dfrq","DFRQ",child_baseline,
                    numbers = c(1:17),by = "Site")
```

```{r child results by site}
child_results <- 
  bind_rows(cdi,foh_maintain,foh_helpless,foh_social, 
            peds_ql_d,peds_ql_tx1,peds_ql_tx2,peds_ql_w,peds_ql_c,
            pds_boys,pds_girls,meter,pump,bgmipr,dfrq)
rownames(child_results) <- 
  c("Depression (CDI2)","FOH (CHFS) – Maintain High BG",
    "FOH (CHFS) – Worry/Helplessness",
    "FOH (CHFS) – Worry/Social Consequences",
    "PedsQL Diabetes","PedsQL Treatment 1","PedsQL Treatment 2",
    "PedsQL Worry","PedsQL Communication","PDS - Boys",
    "PDS - Girls","Meter","tdk","BGMIPR","DFRQ")
kable(child_results,row.names = T,digits = 3)
```

#### Parent

```{r parent models by site,include=FALSE}
# CESD
cesd <- fiml_latent("depression","CESD",parent_baseline,by = "Site")
# FOH
# Maintain
foh_maintain <- fiml_latent("maintain","FOH_P",parent_baseline,
                            numbers = c(3,4,7),by = "Site")
# Helplessness
foh_helpless <- fiml_latent("helpless","FOH_P",
                            parent_baseline,by = "Site",
                            numbers = c(13,14,15,17,20,23,24,25,26))
# Social
foh_social <- fiml_latent("social","FOH_P",parent_baseline,
                          numbers = c(16,18,19,21,22),by = "Site")
# WECARE
wecare_well_being <- 
  fiml_latent("wellbeing","WECARE",parent_baseline,by = "Site",
              numbers = c(1,2,4,14,15,16,17,18,19,20,21,22,23))
wecare_satisfaction <- 
  fiml_latent("satisfaction","WECARE",parent_baseline,by = "Site",
              numbers = c(3,28,29,30,31,32,35,36,37))
wecare_acceptance <- fiml_latent("ease","WECARE",
                                 parent_baseline,by = "Site",
                                 numbers = c(5,6,7,8,9,10))
wecare_ease <- fiml_latent("ease","WECARE",parent_baseline,by = "Site",
                           numbers = c(11,12,13,24,25,26,27,33,34))
# Pump knowledge - Delete 4a, 4b, and those without variance
pump <- fiml("tdk","tdk_",parent_baseline,by = "Site")
# BGMIPR
bgmipr <- fiml_latent("bgmipr","BGMIPR",parent_baseline,by = "Site",
                      numbers = c(1:16))
# DFRQ
dfrq <- fiml_latent("dfrq","DFRQ",parent_baseline,by = "Site",
                    numbers = c(1:17))
```

```{r parent results by site}
parent_results <- bind_rows(cesd,foh_maintain,foh_helpless,foh_social,
                            wecare_well_being,wecare_satisfaction,wecare_acceptance,
                            wecare_ease,pump,bgmipr,dfrq)
rownames(parent_results) <- 
  c("Depression (CESD)","FOH (PHFS) – Maintain High BG",
    "FOH (PHFS) – Worry/Helplessness","FOH (PHFS) – Worry/Social Consequences",
    "Psychosocial well-being","Treatment satisfaction","Acceptance of insulin administration",
    "Ease of insulin use","Pump Knowledge","BGMIPR","DFRQ")
kable(parent_results,row.names = T,digits = 3)
```