---
title: "PIU! Baseline Paper"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
library(tidyverse)
library(arsenal)
library(skimr)
library(knitr)
library(coefficientalpha)
set.seed(1017)
```

```{r data}
# Import questionnair and demographic data
child <- read.csv("./Kim Driscoll/PIU Baseline Paper/Data_Cleaned/child.csv",na.strings = "")
parent <- read.csv("./Kim Driscoll/PIU Baseline Paper/Data_Cleaned/parent.csv",na.strings = "")
# Fill missing IDs
child <- child %>% fill(ID,.direction = "down")
# Pump variables
pump <- read.csv("./Kim Driscoll/Pump Variables/Data_Cleaned/summary.csv",na.strings = "")
colnames(pump)[which(colnames(pump)=="study_visit")] <- "Study.Visit"
colnames(pump)[which(colnames(pump)=="subject_id")] <- "ID"
pump$Study.Visit <- as.numeric(sub("T","",pump$Study.Visit))
# Merge
child <- full_join(child,pump,by = c("ID", "Study.Visit"))
# Pump variables per day
child <- child %>%
  mutate(readings_per_day = total_readings / days_worn,
         carbs_per_day = total_carbs / days_worn,
         boluses_per_day = total_bolus / days_worn,
         bg_150_249_perc_bolus = bg_150_249_with_bolus_only / total_150_249,
         bg_150_249_perc_bolus_carb = bg_150_249_with_bolus_carb / total_150_249,
         bg_above_250_perc_bolus = bg_above_250_with_bolus_only / total_above_250,
         bg_above_250_perc_bolus_carb = bg_above_250_with_bolus_carb / total_above_250)
# Reverse scored psych items
# CDI
rev_cdi <- paste0("CDI",c(2,6,7,9,10,12,14,15,17,20,23,24,26,27))
child[,rev_cdi] <- lapply(child[,rev_cdi], function(x){2-x})
# CES-D
rev_ces <- paste0("CESD",c(4,8,12,16))
parent[,rev_ces] <- lapply(parent[,rev_ces], function(x){3-x})
# WECARE
rev_wecare <- paste0("WECARE",c(24,25,26,27,28,29,30,31,32,35,37))
parent[,rev_wecare] <- lapply(parent[,rev_wecare], function(x){4-x})
# Meter
meter_cols <- grep("METER",colnames(child))
child[,meter_cols] <- lapply(child[,meter_cols], function(x){
  x <- as.numeric(x)
  x[which(x==2)]<-0
  x
  })
```

# Table 1: Self-Management Behaviors at Baseline

There was only one missing value (for participant 220, who had $\frac{0}{0}$ for BG $\geq$ 250), so simple tests were used for this table. Holly, I have one additional intervention participant, any idea why that might be?

```{r baseline self management,results='asis'}
# Baseline data
child_baseline <- child[child$Study.Visit==1,]
# Treatment variable - make as usual the reference group
child_baseline$Treatment <- factor(child_baseline$Treatment,levels = c("Treatment as Usual","Tailored Feedback"))
# Variables
smvars <- c("days_worn","carbs_per_day","boluses_per_day",
            "bg_150_249_perc_bolus","bg_150_249_perc_bolus_carb",
            "bg_above_250_perc_bolus","bg_above_250_perc_bolus_carb")
# Table
t1 <- tableby(Treatment~kwt(days_worn)+carbs_per_day+boluses_per_day+
                bg_150_249_perc_bolus+bg_150_249_perc_bolus_carb+
                bg_above_250_perc_bolus+bg_above_250_perc_bolus_carb,
              data = child_baseline)
summary(t1,pfootnote=T)
```

# Table 2: Psychosocial Variables at Baseline

```{r fiml function}
# lavaan models function
fiml_by_treatment <- function(outcome_name,prefix,data,numbers = NULL){
  vars <- colnames(data)[grep(prefix,colnames(data))]
  if(!is.null(numbers)){
    vars <- paste0(prefix,numbers)
  }
  vars_form <- paste(vars,collapse = " + ")
  form <- paste(outcome_name," =~ ",vars_form,"\n",
                outcome_name," ~ Treatment")
  # Fit SEM
  mod <- sem(form,data = data,missing = "fiml")
  # Get mean sum of items
  n <- as.numeric(mod@Data@nobs)
  control <- rep(1,length(vars))
  treat <- rep(1,length(vars)+1)
  means  <- fitted(mod)$mean 
  mean_control <- control %*% head(means,-1)
  mean_treat <- treat %*% means
  # Get se of item sum
  covar_mat_names <- c(paste0(vars,"~1"),paste0(outcome_name,"~Treatment"))
  vcov <- vcov(mod)
  vcov <- vcov[covar_mat_names,covar_mat_names]
  se_control <- sqrt(control %*% vcov[1:length(control),1:length(control)] %*% control)
  se_treat <- sqrt(treat %*% vcov[1:length(treat),1:length(treat)] %*% treat)
  # Regression estimates
  s <- as.data.frame(summary(mod)$PE)
  p <- s[which(s$lhs == outcome_name & s$rhs == "Treatment"),"pvalue"]
  # Use EM robust to get alpha value
  drop <- which(rowSums(is.na(data[,vars])) == length(vars))
  a <- alpha(data[,vars],drop = drop,test = F,varphi = 0,silent = T)
  # Return
  return(t(round(c(mean_control,se_control,mean_treat,se_treat,p,a$alpha),3)))
}
res_names <- c("Mean Treatment as Usual","SE Treatment as Usual",
               "Mean Tailored Feedback","SE Tailored Feedback","P Value",
               "Alpha")
```

## Results

Notes:
- There may just be too much missing data to use FIML or MI on the BDI questionnaires.

```{r models,include=FALSE}
# CDI
cdi <- fiml_by_treatment("depression","CDI",child_baseline)
# Meter skills
meter <- fiml_by_treatment("meter","METER",child_baseline,numbers = c(1:6,9:18))
```

```{r results}
# CDI
kable(cdi,col.names = res_names,caption = "CDI")
# Meter
kable(meter,col.names = res_names,caption = "Meter")
```