---
title: "Latino Program - Gerard Gonzalez"
author: "Kristen Campbell"
date: "May 17, 2019"
output: word_document
editor_options: 
  chunk_output_type: console
---


```{r,echo=FALSE,message=FALSE,results='hide',warning=FALSE}
library(knitr)
setwd("C:/Users/campbkri/Documents/GitHub/BDC-Code/Gerard Gonzalez")
source('01_functions.R')
source('00_data.R')
setwd("C:/Users/campbkri/Documents/GitHub/BDC-Code/Gerard Gonzalez")
source('02_tables.R')
source('03_mixed.R')

```

#Statistical Methods


#Summary Statistics
**Table 1a: . Descriptive statistics for Latino Program group by year (of last value per time period)**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.1a, align = c("l", "l"))
```

**Table 1b:. Descriptive statistics for Control group by year (of last value per time period)**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.1b, align = c("l", "l"))
```

Table 1 presents descriptive statistics of A1c and Checks per Day by time period, regardless of age group category.

**Table 2. A1c and BMI Descriptive Statistics by year, age category and group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.2, align = c("l", "l"))
```

Table 2 presents descriptive statistics of A1c and Checks per Day by treatment group, time period and age category. 

#A1C Results

**Table 3A. A1c Over Time: WHOLE COHORT**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(year3_sum, align = c("l", "l"))
```

**Table 3b: Difference between Latino Program and Control A1c at each timepoint: WHOLE COHORT**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3, align = c("l", "l"))
```


Table 3A-3B show results of mixed effects modeling looking at whether there was a significant difference in A1c in Treatment and Control groups at baseline, year 1 and year 2 (table 3A), as well as results of specific contrasts of interest. There is not a significant difference in trend of A1c over time between groups (table 3A, p=xxx). There was no significant difference in A1c between groups at baseline, year 1 or year 2 (table 3B, p=xxx, p=xxx, p=xxx).

**Table 3c. A1c Over Time: <12 YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(year3_sum_young, align = c("l", "l"))
```

**Table 3d: Difference between Latino Program and Control A1c at each timepoint: <12 YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3_young, align = c("l", "l"))
```

**Table 3e. A1c Over Time: 12+ YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(year3_sum_old, align = c("l", "l"))
```

**Table 3f: Difference between Latino Program and Control A1c at each timepoint: 12+ YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3_old, align = c("l", "l"))
```

#Checks Per Day Results

**Table 4a.  Checks Per Day Over Time: WHOLE COHORT**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(checks_year3_sum, align = c("l", "l"))
```

**Table 4b: Difference between Latino Program and Control Checks Per Day at each timepoint: WHOLE COHORT.**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3_checks, align = c("l", "l"))
```

**Table 4c. Checks Per Day Over Time: <12 YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(checks_year3_sum_young, align = c("l", "l"))
```

**Table 4d: Difference between Latino Program and Control Checks Per Day at each timepoint: <12 YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3_checks_young, align = c("l", "l"))
```

**Table 4e. A1c Over Time: 12+ YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(checks_year3_sum_old, align = c("l", "l"))
```

**Table 4f: Difference between Latino Program and Control A1c at each timepoint: 12+ YEARS**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3_checks_old, align = c("l", "l"))
```


Table 4A-4B show results of mixed effects modeling looking at whether there was a significant difference in Checks per Day in Treatment and Control groups at baseline, year 1, year 2, and year 3 (table 4A), as well as results of specific comparisons of interest. There is not a significant difference in trend of A1c over time between groups (table 4A, p=xxx). There was no significant difference in Checks per Day between groups at baseline, year 1 or year 2 (table 4B, p=xxx, p=xxx, p=xxx).

#Pump Use Results
**Table 5a: Frequency of pump usage in greater than or equal to 12 years cohort by time period in the Latino Program group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.5a, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.5a,3)`

**Table 5B. Frequency of pump usage in greater than or equal to 12 years cohort by time period in the Control group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.5b, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.5b,3)`

**Table 6A. Frequency of pump usage in the less than 12 years cohort by time period in the Latino Program group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.6a, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.6a,3)`

**Table 6B. Frequency of pump usage in the less than 12 years cohort by time period in the Control group **
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.6b, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.6b,3)`

**Table 7A. Frequency of CGM usage in greater than or equal to 12 years cohort by time period in the Latino Program group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.7a, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.7a,3)`

**Table 7B. Frequency of CGM usage in greater than or equal to 12 years cohort by time period in the Control group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.7b, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.7b,3)`

**Table 8A. Frequency of CGM usage in the less than 12 years cohort by time period in the Latino Program group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.8a, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.8a,3)`

**Table 8B. Frequency of CGM usage in the less than 12 years cohort by time period in the Control group**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.8b, align = c("l", "l"))
```

Year 1 versus Year 3 pump use: `r round(test.8b,3)`
