---
title: "Tracking DNA-based antigen-specific TCRs"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:   
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
# required libraries
library(purrr)
library(tableone)
library(dplyr)
library(plyr)
library(zoo)
library(nlme)
library(lme4)
library(emmeans)
library(knitr)
library(car)
library(tidyr)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r, include=F}
setwd(home_dir)
# data for fig 2a
dat2a <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig2a.csv",na.strings = c(""," ","X"))

# data for fig 2b
dat2b <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig2b.csv",na.strings = c(""," ","X"))

# data for fig 2c
dat2c <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig2c.csv",na.strings = c(""," ","X"))

# data for fig 6a
dat6a <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig6a.csv",na.strings = c(""," ","X"))

# data for fig 6b
dat6b <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig6b.csv",na.strings = c(""," ","X"))

# data for fig 3a
dat3a <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig3a.csv",na.strings = c(""," ","X"))
# take maximum value per patient ID and time
dat3a <- dat3a %>% group_by(Pt.ID, Time) %>% filter(All.Islet.CD4 == max(All.Islet.CD4, na.rm=TRUE))
dat3a <- unique(dat3a)
dat3a$Time <- factor(dat3a$Time, levels=c("0-3","3-6","6-9","9-12","12+"))

# data for fig 3b
dat3b <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig3b.csv",na.strings = c(""," ","X"))
# take maximum value per patient ID and time
dat3b <- dat3b %>% group_by(Pt.ID, Time) %>% filter(PPI.CD4 == max(PPI.CD4, na.rm=TRUE))
dat3b <- unique(dat3b)
dat3b$Time <- factor(dat3b$Time, levels=c("0-3","3-6","6-9","9-12","12+"))

# data for fig 3c
dat3c <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig3c.csv",na.strings = c(""," ","X"))
# take maximum value per patient ID and time
dat3c <- dat3c %>% group_by(Pt.ID, Time) %>% filter(GAD.CD4 == max(GAD.CD4, na.rm=TRUE))
dat3c <- unique(dat3c)
dat3c$Time <- factor(dat3c$Time, levels=c("0-3","3-6","6-9","9-12","12+"))

# data for fig 3d
dat3d <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig3d.csv",na.strings = c(""," ","X"))
# take maximum value per patient ID and time
dat3d <- dat3d %>% group_by(Pt.ID, Time) %>% filter(All.Islet.CD8 == max(All.Islet.CD8, na.rm=TRUE))
dat3d <- unique(dat3d)
dat3d$Time <- factor(dat3d$Time, levels=c("0-3","3-6","6-9","9-12","12+"))

# data for fig 3e
dat3e <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig3e.csv",na.strings = c(""," ","X"))
# take maximum value per patient ID and time
dat3e <- dat3e %>% group_by(Pt.ID, Time) %>% filter(PPI.CD8 == max(PPI.CD8, na.rm=TRUE))
dat3e <- unique(dat3e)
dat3e$Time <- factor(dat3e$Time, levels=c("0-3","3-6","6-9","9-12","12+"))

# data for fig 3f
dat3f <- read.csv("/Volumes/BDC/Projects/Aaron Michels/Follow up temporal development T-cell paper/Data raw/fig3f.csv",na.strings = c(""," ","X"))
# take maximum value per patient ID and time
dat3f <- dat3f %>% group_by(Pt.ID, Time) %>% filter(ZnT8.CD8 == max(ZnT8.CD8, na.rm=TRUE))
dat3f <- unique(dat3f)
dat3f$Time <- factor(dat3f$Time, levels=c("0-3","3-6","6-9","9-12","12+"))
```

# Methods

All analyses were performed using R version 4.1 (R Core Team, Vienna).  Specific information about the models corresponding to each of the figures is below.

Figure 2a: mixed-effects models were used to test the association of templates and age.

Figure 2b: mixed-effects models were used to test the association of templates and age.

Figure 2c: linear models were used to test the association of templates and age.

Figures 3a-3f: mixed-effects models with an interaction between group and time were used to test whether the trajectories differed by group and to compare groups at each timepoint.

# Results

## Figure 2a

### All viral

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(Viral.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2a,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### Influenza

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(Flu.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2a,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### EBV

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(EBV.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2a,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### CMV

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(CMV.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2a,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### ADV

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(ADV.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2a,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

## Figure 2b

### All viral

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(Viral.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2b,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### Influenza

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(Flu.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2b,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### EBV

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(EBV.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2b,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### CMV

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(CMV.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2b,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### ADV

The p-value for "Age..yrs." in the table of type 3 tests of fixed effects tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lme(ADV.TCRβ.templates ~ Age..yrs.,random=~1|Pt.ID,data = dat2b,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

## Figure 2c

### All viral

The p-value (in column labeled "Pr(>|t|)") for "Age..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(Viral.TCRβ.templates ~ Age..yrs.,data = dat2c,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### Influenza

The p-value (in column labeled "Pr(>|t|)") for "Age..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(Flu.TCRβ.templates ~ Age..yrs.,data = dat2c,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### EBV

The p-value (in column labeled "Pr(>|t|)") for "Age..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(EBV.TCRβ.templates ~ Age..yrs.,data = dat2c,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### CMV

The p-value (in column labeled "Pr(>|t|)") for "Age..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(CMV.TCRβ.templates ~ Age..yrs.,data = dat2c,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### ADV

The p-value (in column labeled "Pr(>|t|)") for "Age..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(ADV.TCRβ.templates ~ Age..yrs.,data = dat2c,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

## Figure 3a

```{r, echo=FALSE, message=FALSE}
mod <- lme(All.Islet.CD4 ~ Time*Group,random=~1|Pt.ID,data = dat3a,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,~Time*Group)
mod_pairs <-  pairs(mod_means,adjust="tukey",simple="Group")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
kable(mod_means,caption="Time point means.",format.args = list(scientific = FALSE))
kable(mod_pairs,caption="Pair-wise comparisons between groups.",format.args = list(scientific = FALSE))
```

## Figure 3b

```{r, echo=FALSE, message=FALSE}
mod <- lme(PPI.CD4 ~ Time*Group,random=~1|Pt.ID,data = dat3b,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,~Time*Group)
mod_pairs <-  pairs(mod_means,adjust="tukey",simple="Group")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
kable(mod_means,caption="Time point means.",format.args = list(scientific = FALSE))
kable(mod_pairs,caption="Pair-wise comparisons between groups.",format.args = list(scientific = FALSE))
```

## Figure 3c

```{r, echo=FALSE, message=FALSE}
mod <- lme(GAD.CD4 ~ Time*Group,random=~1|Pt.ID,data = dat3c,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,~Time*Group)
mod_pairs <-  pairs(mod_means,adjust="tukey",simple="Group")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
kable(mod_means,caption="Time point means.",format.args = list(scientific = FALSE))
kable(mod_pairs,caption="Pair-wise comparisons between groups.",format.args = list(scientific = FALSE))
```

## Figure 3d

```{r, echo=FALSE, message=FALSE}
mod <- lme(All.Islet.CD8 ~ Time*Group,random=~1|Pt.ID,data = dat3d,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,~Time*Group)
mod_pairs <-  pairs(mod_means,adjust="tukey",simple="Group")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
kable(mod_means,caption="Time point means.",format.args = list(scientific = FALSE))
kable(mod_pairs,caption="Pair-wise comparisons between groups.",format.args = list(scientific = FALSE))
```

## Figure 3e

```{r, echo=FALSE, message=FALSE}
mod <- lme(PPI.CD8 ~ Time*Group,random=~1|Pt.ID,data = dat3e,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,~Time*Group)
mod_pairs <-  pairs(mod_means,adjust="tukey",simple="Group")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
kable(mod_means,caption="Time point means.",format.args = list(scientific = FALSE))
kable(mod_pairs,caption="Pair-wise comparisons between groups.",format.args = list(scientific = FALSE))
```

## Figure 3f

```{r, echo=FALSE, message=FALSE}
mod <- lme(ZnT8.CD8 ~ Time*Group,random=~1|Pt.ID,data = dat3f,na.action = na.omit)
mod_anova <- anova.lme(mod, type="marginal")
mod_means <- emmeans(mod,~Time*Group)
mod_pairs <-  pairs(mod_means,adjust="tukey",simple="Group")

kable(mod_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
kable(mod_means,caption="Time point means.",format.args = list(scientific = FALSE))
kable(mod_pairs,caption="Pair-wise comparisons between groups.",format.args = list(scientific = FALSE))
```

## Figure 6a

### Unique TCRs

The p-value (in column labeled "Pr(>|t|)") for "Age.Dx..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(X..Unique.TCRs ~ Age.Dx..yrs.,data = dat6a,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### Disease-associated TCR templates

The p-value (in column labeled "Pr(>|t|)") for "Age.Dx..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(X..TCRβ.templates ~ Age.Dx..yrs.,data = dat6a,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

## Figure 6b

### Unique TCRs

The p-value (in column labeled "Pr(>|t|)") for "Age.Dx..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(X..Unique.TCRs ~ Age.Dx..yrs.,data = dat6b,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

### Disease-associated TCR templates

The p-value (in column labeled "Pr(>|t|)") for "Age.Dx..yrs." tests whether there was a significant association with age.

```{r, echo=FALSE, message=FALSE}
mod <- lm(X..TCRβ.templates ~ Age.Dx..yrs.,data = dat6b,na.action = na.omit)

kable(summary(mod)$coefficients,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```