---
title: "Hispanic T1D MEGA Array"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kimber Simmons/GWAS/Data_Cleaned")
```

# Combine "_redo" and "_removeKnownFails"

```{bash combine files, eval=FALSE}
# Combine files
plink --noweb --bfile Simmons_redo --merge-list files_to_combine.txt --make-bed --out Simmons_combined
```

## Missingness pre-deletion

```{bash plink missing, eval=FALSE}
# Check missing
plink --noweb --bfile Simmons_combined --missing
```

```{r r missing plots pre-deletion, echo=FALSE}
# Read in plink missingness files
indmiss <-read.table(file="./plink.imiss",
                    header=TRUE)
snpmiss <-read.table(file="./plink.lmiss",
                    header=TRUE)
# Plots 
hist(indmiss[,6],main="Histogram proportion missing SNPs per individual",
     xlab = "Proportion Missing")
hist(snpmiss[,5],main="Histogram proportion missing individuals per SNP",
     xlab = "Proportion Missing ") 
```

## Missingness post-deletion

```{bash plink filter missing, eval=FALSE}
# Delete SNPs and individuals with high levels of missingness
# Delete SNPs
plink --noweb --bfile Simmons_combined --geno 0.02 --make-bed --out Simmons_combined
# Delete individuals
plink --noweb --bfile Simmons_combined --mind 0.02 --make-bed --out Simmons_combined
# Check missing post-deletion
plink --noweb --bfile Simmons_combined --missing --out miss_post_del
```

```{r r missing plots post-deletion, echo=FALSE}
# Post-deletion
indmiss_check <-
  read.table(file="./miss_post_del.imiss",
                    header=TRUE)
snpmiss_check <-
  read.table(file="./miss_post_del.lmiss",
                    header=TRUE)
# Plots
hist(indmiss_check[,6],main="Histogram proportion missing SNPs per individual",
     xlab = "Proportion Missing")
hist(snpmiss_check[,5],main="Histogram proportion missing individuals per SNP",
     xlab = "Proportion Missing ") 
```

## Sex check pre-deletion

```{bash plink sex check pre, eval=FALSE}
# Check sex discrepancy
plink --noweb --bfile Simmons_combined --check-sex
```

```{r r sex check pre, echo=FALSE}
# Gender file
gender <- read.table("./plink.sexcheck",
                     header=T,as.is=T)
# Histograms
hist(gender[,6],main="Gender", xlab="F")
# Males
male=subset(gender, gender$PEDSEX==1)
hist(male[,6],main="Men",xlab="F")
# Females
female=subset(gender, gender$PEDSEX==2)
hist(female[,6],main="Women",xlab="F")
```

## Sex check post-deletion

```{bash plink sex check post, eval=FALSE}
# Remove subjects with problematic X chromosome homozygosity
grep "PROBLEM" plink.sexcheck| awk '{print$1,$2}'> sex_discrepancy.txt
plink --noweb --bfile Simmons_combined --remove sex_discrepancy.txt --make-bed --out Simmons_combined
# Check sex discrepancy again
plink --noweb --bfile Simmons_combined --check-sex --out sex_post_del
```

```{r r sex check post, echo=FALSE}
# Gender file
gender <- read.table("./sex_post_del.sexcheck",
                     header=T,as.is=T)
# Histograms
hist(gender[,6],main="Gender", xlab="F")
# Males
male=subset(gender, gender$PEDSEX==1)
hist(male[,6],main="Men",xlab="F")
# Females
female=subset(gender, gender$PEDSEX==2)
hist(female[,6],main="Women",xlab="F")
```

## MAF check pre-deletion

```{bash plink maf check pre, eval=FALSE} 
# Select autosomal SNPs only (i.e., from chromosomes 1 to 22).
awk '{ if ($1 >= 1 && $1 <= 22) print $2 }' Simmons_combined.bim > snp_1_22.txt
plink --noweb --bfile Simmons_combined --extract snp_1_22.txt --make-bed --out Simmons_combined
# Check minor allele frequency
plink --noweb --bfile Simmons_combined --freq --out MAF_check
```

```{r r maf check pre, echo=FALSE}
# MAF file
maf_freq <- read.table("./MAF_check.frq", header =TRUE, as.is=T)
hist(maf_freq[,5],main = "MAF distribution", xlab = "MAF")
```

## MAF check post-deletion

```{bash plink maf check post, eval=FALSE}
# Remove those with MAF < 0.05
plink --noweb --bfile Simmons_combined --maf 0.05 --make-bed --out Simmons_combined
# Check minor allele frequency again
plink --noweb --bfile Simmons_combined --freq --out MAF_check_post
```

```{r r maf check post, echo=FALSE}
# MAF file
maf_freq <- read.table("./MAF_check_post.frq", header =TRUE, as.is=T)
hist(maf_freq[,5],main = "MAF distribution", xlab = "MAF")
```

## Hardy-Weinberg equilibrium pre-deletion

```{bash plink hw check pre, eval=FALSE}
plink --noweb --bfile Simmons_combined --hardy
# Select SNPs with HWE p-value below 0.00001
awk '{ if ($9 <0.00001) print $0 }' plink.hwe>plinkzoomhwe.hwe
```

```{r r hw check pre, echo=FALSE}
hwe<-read.table (file="./plink.hwe", header=TRUE)
hist(hwe[,9],main="Histogram HWE")
# Zoomed in
hwe_zoom<-read.table (file="./plinkzoomhwe.hwe", header=TRUE)
hist(hwe_zoom[,9],main="Histogram HWE: strongly deviating SNPs only")
```

## Hardy-Weinberg equilibrium post-deletion

```{bash plink hw check post, eval=FALSE}
# Filter for controls
plink --noweb --bfile Simmons_combined --hwe 1e-6 --make-bed --out Simmons_combined
# Filter for cases
plink --noweb --bfile Simmons_combined --hwe 1e-12 --hwe-all --make-bed --out Simmons_combined
# Check again
plink --noweb --bfile Simmons_combined --hardy --out post_del
# Select SNPs with HWE p-value below 0.00001
awk '{ if ($9 <0.00001) print $0 }' post_del.hwe>post_del_zoom.hwe
```

```{r r hw check post, echo=FALSE}
hwe<-read.table (file="./post_del.hwe", header=TRUE)
hist(hwe[,9],main="Histogram HWE")
# Zoomed in
hwe_zoom<-read.table (file="./post_del_zoom.hwe", header=TRUE)
hist(hwe_zoom[,9],main="Histogram HWE: strongly deviating SNPs only")
```

## Heterozygosity pre-deletion

```{bash plink heterozyg check pre, eval=FALSE}
plink --noweb --bfile Simmons_combined --exclude inversion.txt --range --indep-pairwise 50 5 0.2 --out indepSNP
plink --noweb --bfile Simmons_combined --extract indepSNP.prune.in --het --out R_check
```

```{r r heterozyg check pre, echo=FALSE}
het <- read.table("./R_check.het", head=TRUE)
het$HET_RATE = (het$"N.NM." - het$"O.HOM.")/het$"N.NM."
hist(het$HET_RATE, xlab="Heterozygosity Rate", ylab="Frequency", main= "Heterozygosity Rate")
# List individuals more than 3 SD from the rate mean
het$HET_RATE = (het$"N.NM." - het$"O.HOM.")/het$"N.NM."
het_fail = subset(het, (het$HET_RATE < mean(het$HET_RATE)-3*sd(het$HET_RATE)) | (het$HET_RATE > mean(het$HET_RATE)+3*sd(het$HET_RATE)));
het_fail$HET_DST = (het_fail$HET_RATE-mean(het$HET_RATE))/sd(het$HET_RATE);
write.table(het_fail, "fail-het-qc.txt", row.names=FALSE)
```

## Heterozygosity post-deletion

```{bash plink heterozyg check post, eval=FALSE}
sed 's/"// g' fail-het-qc.txt | awk '{print$1, $2}'> het_fail_ind.txt
# Remove heterozygosity rate outliers.
plink --noweb --bfile Simmons_combined --remove het_fail_ind.txt --make-bed --out Simmons_combined
plink --noweb --bfile Simmons_combined --extract indepSNP.prune.in --het --out R_check_post
```

```{r r heterozyg check post, echo=FALSE}
het <- read.table("./R_check_post.het", head=TRUE)
het$HET_RATE = (het$"N.NM." - het$"O.HOM.")/het$"N.NM."
hist(het$HET_RATE, xlab="Heterozygosity Rate", ylab="Frequency", main= "Heterozygosity Rate")
```

## Relatedness pre-deletion

```{bash plink related check pre, eval=FALSE}
plink --noweb --bfile Simmons_combined --extract indepSNP.prune.in --genome --min 0.2 --out pihat_min0.2
awk '{ if ($8 >0.9) print $0 }' pihat_min0.2.genome>zoom_pihat.genome
```

```{r r related check pre, echo=FALSE}
relatedness = read.table("pihat_min0.2.genome", header=T)
par(pch=16, cex=1)
with(relatedness,plot(Z0,Z1, xlim=c(0,1), ylim=c(0,1), type="n"))
with(subset(relatedness,RT=="PO") , points(Z0,Z1,col=4))
with(subset(relatedness,RT=="UN") , points(Z0,Z1,col=3))
legend(1,1, xjust=1, yjust=1, legend=levels(relatedness$RT), pch=16, col=c(4,3))

relatedness_zoom = read.table("zoom_pihat.genome", header=T)
par(pch=16, cex=1)
with(relatedness_zoom,plot(Z0,Z1, xlim=c(0,0.02), ylim=c(0.98,1), type="n"))
with(subset(relatedness_zoom,RT=="PO") , points(Z0,Z1,col=4))
with(subset(relatedness_zoom,RT=="UN") , points(Z0,Z1,col=3))
legend(0.02,1, xjust=1, yjust=1, legend=levels(relatedness$RT), pch=16, col=c(4,3))

relatedness = read.table("pihat_min0.2.genome", header=T)
hist(relatedness[,10],main="Histogram relatedness", xlab= "Pihat")  
```

## Relatedness post-deletion

```{bash plink related check post, eval=FALSE}
plink --noweb --bfile Simmons_combined --filter-founders --make-bed --out Simmons_combined
plink --noweb --bfile Simmons_combined --extract indepSNP.prune.in --genome --min 0.2 --out pihat_min0.2_post
awk '{ if ($8 >0.9) print $0 }' pihat_min0.2.genome>zoom_pihat_post.genome
```

```{r r related check post, echo=FALSE}
relatedness = read.table("pihat_min0.2_post.genome", header=T)
par(pch=16, cex=1)
with(relatedness,plot(Z0,Z1, xlim=c(0,1), ylim=c(0,1), type="n"))
with(subset(relatedness,RT=="PO") , points(Z0,Z1,col=4))
with(subset(relatedness,RT=="UN") , points(Z0,Z1,col=3))
legend(1,1, xjust=1, yjust=1, legend=levels(relatedness$RT), pch=16, col=c(4,3))

relatedness_zoom = read.table("zoom_pihat_post.genome", header=T)
par(pch=16, cex=1)
with(relatedness_zoom,plot(Z0,Z1, xlim=c(0,0.02), ylim=c(0.98,1), type="n"))
with(subset(relatedness_zoom,RT=="PO") , points(Z0,Z1,col=4))
with(subset(relatedness_zoom,RT=="UN") , points(Z0,Z1,col=3))
legend(0.02,1, xjust=1, yjust=1, legend=levels(relatedness$RT), pch=16, col=c(4,3))

relatedness = read.table("pihat_min0.2_post.genome", header=T)
hist(relatedness[,10],main="Histogram relatedness", xlab= "Pihat") 
```