---
title: "Mental Health Follow-Up"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(arsenal)
library(VGAM)
library(nlme)
library(emmeans)
library(splines)
library(skimr)
library(knitr)
library(reshape2)
library(tidyverse)
library(broom)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects/Shideh Majidi/Depression"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Shideh Majidi/Depression"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Shideh Majidi/Depression"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r a1c data import,include=FALSE}
# Import data
source("~/GitHub/BDC-Code/Shideh Majidi/Depression/sra.R")
a1cs <- read.csv("./Data_Cleaned/a1cs.csv")
# Clean
a1cs$A1cResultDate <- lubridate::mdy(a1cs$A1cResultDate)
temp <- data
temp$record_id <- as.numeric(as.character(temp$record_id))
# Add screen date to A1c data, find days between
a1cs <- left_join(a1cs,temp[,c("record_id","screen_date",
                               "charttreatafter.factor",
                               "charttreatbefore.factor")],
                  by = "record_id")
a1cs$days_from_screen <- 
  round(as.numeric(difftime(a1cs$A1cResultDate,a1cs$screen_date,
                            units = "days")))
# 1 year pre-/post-screening
a1cs <- a1cs[abs(a1cs$days_from_screen) < 365.25,]
# Clean up
rm(temp)
# Insurance - private vs. other
levels(data$insurance.factor) <- c("Other","Private","Other","Other","Other","Other","Other")
# Combined race
data$race.factor_comb = data$race.factor
levels(data$race.factor_comb) = c("White","Non-white","Non-white","Non-white","Non-white","Unknown/Not Reported")
```

# Table 1a: Descriptive Characteristics by Follow-Up Status

```{r table 1a,echo=FALSE,results='asis'}
t1_form = as.formula("charttreatafter.factor ~ age + diabetes_duration + a1c + 
                total_score_epic + followupscore + sex.factor + 
                fe(race.factor) + race.factor_comb + ethnicity.factor + insurance.factor + 
                pump_use.factor + cgm.factor + si_pos.factor + 
                followupphq9.factor + followupsi")
newlabels <- list(sex.factor = "Sex",race.factor = "Race",
                  race.factor_comb = "Race (Combined)",
                  ethnicity.factor = "Ethnicity",
                  insurance.factor = "Insurance",
                  insulin_type.factor = "Insulin Type",
                  pump_use.factor = "Pump Use", 
                  cgm.factor = "CGM Use",
                  followupphq9.factor = "Follow-Up PHQ 9?",
                  si_pos.factor = "Positive SI Score?",
                  followupsi = "Follow-Up SI?")
t1 <- tableby(t1_form,data)
summary(t1,pfootnote = T,labelTranslations = newlabels)
```

\pagebreak

# Table 1b: Descriptive Characteristics by SI Status ("si_pos")

```{r table 1b,echo=FALSE,results='asis'}
t1 <- tableby(update(t1_form,si_pos.factor~.-si_pos.factor),data)
summary(t1,pfootnote = T,labelTranslations = newlabels)
```

\pagebreak

# Table 1c: Descriptive Characteristics by Depressive Symptoms ("phq9_pos")

```{r table 1c,echo=FALSE,results='asis'}
t1 <- tableby(update(t1_form,phq9_pos.factor~.-total_score_epic),data)
summary(t1,pfootnote = T,labelTranslations = newlabels)
```

\pagebreak

# Table 1d: Descriptive Characteristics by Age Group at Screening

```{r table 1d,echo=FALSE,results='asis'}
# Age group
data$age_group = cut(data$age,c(-Inf,18,Inf),labels = c("< 18 Yrs.",">= 18 Yrs."),right = F)
# Table 1
t1 <- tableby(update(t1_form,age_group~.-age),data)
summary(t1,pfootnote = T,labelTranslations = newlabels)
```

\pagebreak

# HbA1c

## Pre-/post-Screening

HbA1c was not significantly different 1 year before and 1 year after screening (p = `r round(t.test(data$a1c_1yearbefore,data$a1c_1yearafter,paired = T)$p.value,3)`).

## HbA1c by Days From Screening

```{r echo=FALSE,warning=FALSE,message=FALSE,dpi=1200,fig.width=7}
ggplot(a1cs[a1cs$days_from_screen>=0,],
       aes(x = days_from_screen,y = A1cValue)) +
  geom_point(shape = ".") +
  geom_smooth(aes(linetype = charttreatafter.factor),
              se = F,color = "black") + 
  xlab("Days from Screening") + ylab("HbA1c (%)") +
  scale_y_continuous(limits = c(6,15),breaks = seq(6,15,1)) + 
  scale_linetype_discrete(name = "Mental Health Follow-up") +
  theme_bw() 
```

About `r round(length(which(a1cs$A1cValue == 15))/length(a1cs$A1cValue),3)*100`% of the values are at 15% and another `r round(length(which(a1cs$A1cValue == 14.1))/length(a1cs$A1cValue),3)*100`% of the values are at 14.1%.

\pagebreak

## Piecewise Regression

```{r echo=FALSE,include=FALSE}
# AR(1) is much better, so refit with REML and check polynomials. 
a1c_mod_poly <- lme(A1cValue ~ poly(days_from_screen,4),random = ~1|record_id,
                    data = a1cs,correlation = corAR1())
# Can use linear time. Use splines for knot at 0. Random intercept and slope for subject.
a1cs$days_star <- ifelse(a1cs$days_from_screen < 0,0,a1cs$days_from_screen)
a1c_mod_ri <- lme(A1cValue ~ days_from_screen*charttreatafter.factor + 
                    days_star*charttreatafter.factor,
                  random = ~1|record_id,data = a1cs,correlation = corAR1(),method = "ML")
a1c_mod_ris <- lme(A1cValue ~ days_from_screen*charttreatafter.factor + 
                     days_star*charttreatafter.factor,
                   random = ~days_from_screen|record_id,data = a1cs,correlation = corAR1())
AIC(a1c_mod_ri,a1c_mod_ris) # RI only better by AIC
# Correlation structure
mod <- lme(A1cValue ~ days_from_screen*charttreatafter.factor + 
             days_star*charttreatafter.factor,
           random = ~1|record_id,data = a1cs,method = "ML")
mod_ar1 <- lme(A1cValue ~ days_from_screen*charttreatafter.factor + 
                 days_star*charttreatafter.factor,random = ~1|record_id,data = a1cs,
               correlation = corAR1(),method = "ML")
AIC(mod,mod_ar1)
# Censored
a1cs$A1cCensored <- a1cs$A1cValue
a1cs$A1cCensored[a1cs$A1cCensored > 14.1] <- 14.1
cens <- vglm(A1cValue ~ days_from_screen*charttreatafter.factor + 
               days_star*charttreatafter.factor, tobit(Upper = 14.1), 
             data = a1cs)
uncens <- glm(A1cValue ~ days_from_screen*charttreatafter.factor + 
                days_star*charttreatafter.factor,data = a1cs)
```

```{r plot model,echo=FALSE,include=FALSE}
# Plot splines
ggplot(data = a1cs,aes_string(x = "days_from_screen",y = "A1cValue",
                              group = "record_id")) + 
  geom_line(aes_string(x = "days_from_screen",y = fitted(a1c_mod_ri)),
            alpha = 0.05) +
  theme_bw() +
  xlab("Days from Start") + ylab("HbA1c (%)")
```

```{r model results, echo=FALSE}
# Adjust for previous treatment 
a1c_mod_ri <- lme(A1cValue ~ days_from_screen*charttreatafter.factor + 
                    days_star*charttreatafter.factor + charttreatbefore.factor,
                  random = ~1|record_id,data = a1cs,correlation = corAR1(),
                  method = "ML")
# Spline results
# results <- format_nlme_out(a1c_mod_ri,kable = F)
vars <- c("Days","No Evidence of Psych. Treat.","Change in Slope",
          "No Previous Psych. Treat.")
# format_nlme_out(a1c_mod_ri,varnames = vars)
# emmeans
em <- emmeans(a1c_mod_ri,~days_star*charttreatafter.factor)
```

```{r echo=FALSE,warning=FALSE,message=FALSE,dpi=1200,fig.width=7}
a1cs$pred = predict(a1c_mod_ri,type = "response")
fig2 = ggplot(a1cs[a1cs$days_from_screen>=0,],
       aes(x = days_from_screen,y = pred)) +
  geom_point(aes(y = A1cValue),size = 0.5) +
  geom_smooth(aes(y = pred,linetype = charttreatafter.factor),method = "lm",
              se = F,color = "black") + 
  xlab("Days From Screening") + ylab("A1C, %") +
  scale_y_continuous(limits = c(6,15),breaks = seq(6,15,1)) + 
  scale_linetype_discrete(name = "Mental Health Follow-Up") +
  theme_bw(base_size = 20)
fig2
```

# Predicting use of psych treatment

```{r logistic model,echo=FALSE}
data$charttreatafter.factor <- relevel(data$charttreatafter.factor,ref = "No")
# Check models
mod <- glm(charttreatafter.factor ~ sex.factor + race.factor_comb + ethnicity.factor + 
             insurance.factor + a1c + total_score_epic + pump_use.factor + cgm.factor +
             charttreatbefore.factor,family = binomial(link='logit'),
           data = data[data$race.factor_comb!="Unknown/Not Reported" & 
                         data$ethnicity.factor!="Unknown/Not Reported",])
# Nice formatting
results <- tidy(mod,exponentiate = T) %>%
  mutate_at(vars(-term),round,3)
results$term = c("(Intercept)","Male","Non-white",
                 "Hispanic or Latino",
                 "Private insurance","HbA1c at Screening",
                 "PHQ9 score (EPIC)","No pump","No CGM",
                 "No previous treat.")
kable(results,row.names = F)
```

In order to get this model to work, I had to combine levels of race (white vs. non-white) and insurance (private vs. other). Also, because there were not enough "events" among those with unknown race and/or ethnicity (when keeping race and ethnicity separate), those with unknown were excluded (final n = 182). 

Males were `r (1 - results[results$term == "Male","estimate"])*100`% less likely to use psych resources after screening (p = `r results[results$term == "Male","p.value"]`). Hispanics were `r (1 - results[results$term == "Hispanic or Latino","estimate"])*100`% less likely (p = `r results[results$term == "Hispanic or Latino","p.value"]`) and those with no previous treatment were `r (1 - results[results$term == "No previous treat.","estimate"])*100`% less likely (p = `r results[results$term == "No previous treat.","p.value"]`) to get treatment after screening. For each one unit increase in PHQ9 score, participants were on average `r (results[results$term == "PHQ9 score (EPIC)","estimate"] - 1)*100`% more likely to get treatment (p = `r results[results$term == "PHQ9 score (EPIC)","p.value"]`).
