---
title: "Medicaid CGM"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nlme)
library(tableone)
library(nortest)
library(knitr)
library(tidyverse)
library(emmeans)
source('~/Documents/GitHub/Tim-and-Laura/tim_R_functions.R')
```

```{r echo=FALSE,cache=TRUE,include=FALSE}
# Read in demographic data
demographics <- read.csv("/Volumes/som-1/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sonalee Ravi/Medicaid CGM/Data_Cleaned/Demographics.csv")
# Read in CGM data
dat <- read.csv("/Volumes/som-1/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sonalee Ravi/Medicaid CGM/Data_Cleaned/30 day cgm metrics.csv",na.strings = c("","NULL"))
# Format CGM data
dat$EncounterDate <- lubridate::mdy(as.character(dat$EncounterDate))
dat$EPICMRN <- as.factor(dat$EPICMRN)
# Remove rows with no A1c
dat <- dat[which(!is.na(dat$HbA1c)),]
# Replace CGMUse = "No" with NA, fill down "Yes" values (by subject), set NAs back to "No"
# This ensures everything prior to a "Yes" is a "No" and every visit after the first "Yes" is also "Yes"
# Remove CGMUse column (not a good indicator because it's just a field in Epic)
# and Dexcom/Medtronic-specific columns.
dat$CGMUse <- NULL
dat$Dexcom.CGM.Days <- NULL
dat$Medtronic.CGM.Days <- NULL
# Fill down PumpUse "Yes", then change all NAs to no. So after first PumpUse "Yes"
# assume they were always on a pump, but not on pump before then.
dat$PumpUse[dat$PumpUse == "No"] <- NA
dat <- dat %>%
  arrange(EPICMRN,EncounterDate) %>%
  group_by(EPICMRN) %>%
  fill(PumpUse)
dat$PumpUse[is.na(dat$PumpUse)] <- "No"
# Get the first CGM use > 0 for each participant. Set this CGMUse == "Yes"
temp <- dat %>% arrange(EPICMRN,EncounterDate) %>% 
  group_by(EPICMRN) %>%
  filter(X30.day.CGM.use > 0) %>%
  filter(row_number() == 1)
temp$CGMUse <- "Yes"
# Merge the yes dataframe back into the full data set. 
dat <- left_join(dat,temp,by = colnames(dat))
# Fill down CGM use == Yes by ID, replace remaining NAs with No
dat <- dat %>% arrange(EPICMRN,EncounterDate) %>%
  group_by(EPICMRN) %>%
  fill(CGMUse)
dat$CGMUse[is.na(dat$CGMUse)] <- "No"
# Make CMG usage a percentage
dat$CGM_Perc <- (dat$X30.day.CGM.use / 30)*100
# Last off CGM to first on
# Get the last "No" and all "Yes" for each subject
no <- dat[dat$CGMUse == "No",] %>%
  arrange(EPICMRN,EncounterDate) %>%
  group_by(EPICMRN) %>%
  filter(row_number()==n())
yes <- dat[dat$CGMUse == "Yes",] %>%
  arrange(EPICMRN,EncounterDate)
# Filter out those without on/off data (all yes or all no)
no <- no[no$EPICMRN %in% yes$EPICMRN,]
yes <- yes[yes$EPICMRN %in% no$EPICMRN,]
# Bind together, calculate time difference, filter out all but last no and first yes > 75 days.
delta <- rbind(yes,no) %>%
  arrange(EPICMRN,EncounterDate) %>%
  mutate(diff = difftime(EncounterDate,EncounterDate[1],units = "days"))
yes <- delta %>% filter(CGMUse == "Yes",diff > 75) %>%
  group_by(EPICMRN) %>%
  filter(row_number()==1)
no <- no[no$EPICMRN %in% yes$EPICMRN,]
# Bind yes and no
delta <- rbind(yes, no) %>% arrange(EPICMRN,EncounterDate)

# Format demographic data
demographics$EPICMRN <- as.factor(demographics$EPICMRN)
demo_datecols <- c("DOB","DateOfDiabetesDiagnosis","FirstVisitDate") 
demographics[,demo_datecols] <- lapply(demographics[,demo_datecols], lubridate::mdy)
# New variables
demographics$AgeFirstVisit <- as.numeric(round(difftime(demographics$FirstVisitDate,
                                             demographics$DOB,units = "days")/365.25,2))
demographics$DiabetesDuration <- as.numeric(round(difftime(demographics$DateOfDiabetesDiagnosis,
                                             demographics$DOB,units = "days")/365.25,2))
# Remove demographics participants who don't have with and without CGM data
demographicsyn <- demographics[which(demographics$EPICMRN %in% unique(delta$EPICMRN)),]
```

### Table 1: Descriptive statistics for all individuals with HbA1c readings both on and off CGM

```{r echo=FALSE,include=FALSE}
# For each person, average HbA1c on CGM and off, remove those without both
avg_on_off <- dat %>%
  group_by(EPICMRN,CGMUse) %>%
  summarise(mean(HbA1c)) %>%
  group_by(EPICMRN) %>% 
  filter(n() > 1) %>% 
  arrange(EPICMRN,CGMUse)
# Compare
on_off_avg_results <- CreateTableOne("mean(HbA1c)",strata = "CGMUse",data = avg_on_off,
                                     testNonNormal = wilcox.test,argsNonNormal = list(paired = T))
on_off_avg_results <- as.data.frame(print(on_off_avg_results, nonnormal = "mean(HbA1c)"))
p <- wilcox.test(formula = `mean(HbA1c)` ~ CGMUse, data = avg_on_off,paired = T)$p.value
rownames(on_off_avg_results)[2] <- "Mean HbA1c"
on_off_avg_results$p <- as.character(on_off_avg_results$p)
on_off_avg_results$p[2] <- round(p,5)
# Demographics
demographicsoo <- demographics[which(demographics$EPICMRN %in% unique(avg_on_off$EPICMRN)),]
vars <- c("AgeFirstVisit","DiabetesDuration","Gender","Race.Ethnicity")
nonnormal <- norm.check(demographicsoo[,c("AgeFirstVisit","DiabetesDuration")])
t1_on_off <- CreateTableOne(vars,data = demographicsoo)
t1_on_off <- as.data.frame(print(t1_on_off,nonnormal = nonnormal))
```

```{r echo=FALSE}
kable(t1_on_off)
```

### Table 2: Mean HbA1c on CGM and off CGM (CGM Use "Yes" or "No")

```{r echo=FALSE}
kable(on_off_avg_results[,1:3])
```

For this comparison we excluded participants who were never on CGM, and those who only had data while on CGM. HbA1c values while on and off CGM were averaged for each remaining participant. These average HbA1c values were compared using a Wilcoxon signed rank test with continuity correction. 

### Table 3: Descriptive statistics for all individuals with an HbA1c reading on CGM, at least 75 days after their last HbA1c reading off CGM 

```{r echo=FALSE,include=FALSE}
# Table 1 for cohort of 134
vars <- c("AgeFirstVisit","DiabetesDuration","Gender","Race.Ethnicity")
nonnormal <- norm.check(demographicsyn[,c("AgeFirstVisit","DiabetesDuration")])
t1_last_first <- CreateTableOne(vars,data = demographicsyn)
t1_last_first <- as.data.frame(print(t1_last_first,nonnormal = nonnormal))
```

```{r echo=FALSE}
kable(t1_last_first)
```

This cohort is similar to the comparison above, but instead of using the average HbA1c on and off CGM, we compared each participant's last HbA1c off CGM to their first HbA1c on CGM (requiring at least 75 days between the two measurements). A total of 6 participants were excluded because they did not have an HbA1c measure more than 75 days after their last off CGM reading.  

### Table 4: Mixed model comparing last HbA1c not on CGM to first HbA1c on CGM, adjusting for pump status

```{r echo=FALSE,include=FALSE}
delta <- delta %>% 
  unite(cgm_pump,CGMUse,PumpUse,remove = F) %>%
  mutate(cgm_pump = factor(cgm_pump,
                           levels = c("No_No","No_Yes","Yes_No","Yes_Yes")))
# Mixed model testing last no/first yes
# Random intercept
mod_int <- lme(HbA1c ~ cgm_pump,random=~1|EPICMRN,data = delta,na.action = na.omit)
on_off_results <- as.data.frame(summary(mod_int)$tTable)
rownames(on_off_results) <- c("cgmNo:pumpNo","cgmNo:pumpYes","cgmYes:pumpNo","cgmYes:pumpYes")
# Means model
mod_int_mean <- lme(HbA1c ~ cgm_pump-1,random=~1|EPICMRN,data = delta,na.action = na.omit)
on_off_means <- as.data.frame(summary(mod_int_mean)$tTable)
rownames(on_off_means) <- c("cgmNo:pumpNo","cgmNo:pumpYes","cgmYes:pumpNo","cgmYes:pumpYes")
```

```{r echo=FALSE}
kable(on_off_results, caption = "Fixed Effects")
kable(anova.lme(mod_int), caption = "Type 3 Tests of Fixed Effects")
kable(on_off_means, caption = "Group Means")
```

The above model effectively uses CGM and pump status as a treatment group, and examines the effect on HbA1c. The fixed effects table indicates the differences between the no CGM and no pump group and each other group respectively (so only the cgm and pump group was significantly different). The type 3 test of fixed effects indicates that combined CGM and pump status was significant overall. The group means table shows the average HbA1c for each group. The p values in the group means table indicate whether or not the average HbA1c is close to 0. 

The interpretation is similar for the following model.

Pump use for last A1c off CGM:

```{r echo=FALSE}
kable(no %>% ungroup(no) %>% count(PumpUse))
```

Pump use for first A1c on CGM:

```{r echo=FALSE}
kable(yes %>% ungroup(yes) %>% count(PumpUse))
```

### Table 5a: Mean HbA1c and descriptive statistics for individuals with some CGM time vs. those never on CGM

```{r echo=FALSE,include=FALSE}
# Add DOB to 
# Compare A1c in those who never went on CGM to those on CGM
some_cgm <- dat[which(dat$CGMUse == "Yes"),] %>%
  group_by(EPICMRN) %>%
  summarise(mean(HbA1c)) %>%
  mutate(CGMUse = "Some")
never_cgm_ids <- unique(dat$EPICMRN[which(!(dat$EPICMRN %in% some_cgm$EPICMRN))])
never_cgm <- dat[which(dat$EPICMRN %in% never_cgm_ids),] %>%
  group_by(EPICMRN) %>%
  summarise(mean(HbA1c)) %>%
  mutate(CGMUse = "Never")
never_some <- rbind(never_cgm,some_cgm)
never_some <- left_join(never_some,demographics,by = "EPICMRN")
# Race ethnicity indicator variables
never_some$AI_AN <- as.factor(ifelse(never_some$Race.Ethnicity == "American Indian/Alaska Native","Yes","No"))
never_some$Asian <- as.factor(ifelse(never_some$Race.Ethnicity == "Asian","Yes","No"))
never_some$Hispanic <- as.factor(ifelse(never_some$Race.Ethnicity == "Hispanic","Yes","No"))
never_some$more_than_one <- as.factor(ifelse(never_some$Race.Ethnicity == "More than one Race","Yes","No"))
never_some$NativeH_OPI <- as.factor(ifelse(never_some$Race.Ethnicity == "Native Hawaiian/Other Pacific Islander","Yes","No"))
never_some$NH_Black <- as.factor(ifelse(never_some$Race.Ethnicity == "Non-Hispanic Black","Yes","No"))
never_some$NH_White <- as.factor(ifelse(never_some$Race.Ethnicity == "Non-Hispanic White","Yes","No"))
never_some$Other <- as.factor(ifelse(never_some$Race.Ethnicity == "Other","Yes","No"))
never_some$Unknown <- as.factor(ifelse(never_some$Race.Ethnicity == "Unknown","Yes","No"))
never_some$Other_Unknown <- pmax(as.numeric(never_some$Other),as.numeric(never_some$Unknown))
never_some$Other_Unknown <- ifelse(never_some$Other_Unknown == 1,"No","Yes")
# Pump use
pumpuse <- dat 
pumpuse$PumpUse <- ifelse(pumpuse$PumpUse == "Yes",1,0)
pumpuse <- pumpuse %>%
  group_by(EPICMRN) %>%
  summarise(Pump = max(PumpUse,na.rm = T))
pumpuse$Pump <- ifelse(pumpuse$Pump == 1,"Yes","No")
never_some <- left_join(never_some,pumpuse,by = "EPICMRN")
vars[length(vars)-1] <- "Pump"
# Tableone
never_some_comp <- CreateTableOne(c("mean(HbA1c)",vars,"AI_AN","Asian","Hispanic","more_than_one","NativeH_OPI","NH_Black","NH_White","Other_Unknown"),strata = "CGMUse",never_some)
never_some_comp <- print(never_some_comp,nonnormal = c("mean(HbA1c)",nonnormal),exact = c("AI_AN","Asian","more_than_one","NativeH_OPI"))
```

```{r echo=FALSE}
kable(never_some_comp)
```

### Table 5b: Mixed model of HbA1c by technology status, adjusted for age and T1D Duration

```{r echo=FALSE,include=FALSE}
# Compare A1c in those who never went on CGM to those on CGM
some_cgm_ids <- unique(dat$EPICMRN[which(dat$CGMUse == "Yes")])
dat$CGMUseSomeNever <- ifelse(dat$EPICMRN %in% some_cgm_ids,"Some","Never")
# Group into never CGM yes pump, etc.
dat <- dat %>% 
  unite(cgm_ever_pump,CGMUseSomeNever,PumpUse,remove = F)
# Add T1D duration
dat <- left_join(dat,demographics[,c("EPICMRN","DateOfDiabetesDiagnosis")],by = "EPICMRN")
dat$T1D_Duration <- round(difftime(dat$EncounterDate,dat$DateOfDiabetesDiagnosis,units = "days")/365.25,3)
dat$cgm_ever_pump[dat$cgm_ever_pump == "Never_No"] <- "CGM-|Pump-"
dat$cgm_ever_pump[dat$cgm_ever_pump == "Never_Yes"] <- "CGM-|Pump+"
dat$cgm_ever_pump[dat$cgm_ever_pump == "Some_No"] <- "CGM+|Pump-"
dat$cgm_ever_pump[dat$cgm_ever_pump == "Some_Yes"] <- "CGM+|Pump+"
# Mixed model with random intercept
mod_cgm_ever <- lme(HbA1c ~ cgm_ever_pump + Age_AtEncounterDate + T1D_Duration,
                    random=~1|EPICMRN,data = dat,na.action = na.omit)
means <- emmeans(mod_cgm_ever,"cgm_ever_pump")
```

```{r echo=FALSE}
kable(anova(mod_cgm_ever),caption = "Type 3 Tests of Overall Effect")
kable(means)
kable(pairs(means))
```

<!-- CGM uptake = $\frac{177}{892}=19.84\text{ %}$ -->

<!-- ### Figure 1: HbA1c by % CGM Use -->

<!-- ```{r echo=FALSE,include=FALSE} -->
<!-- # Add demographic info to main data -->
<!-- dat <- left_join(dat,demographics,by = "EPICMRN") -->
<!-- dat$DiabetesDuration <- difftime(dat$EncounterDate,dat$DateOfDiabetesDiagnosis,units = "weeks") -->
<!-- # Model A1c continuously vs. CGM percentage, adjusting for pump, age, diabetes duration -->
<!-- mod_adj_int <- lme(HbA1c ~ CGM_Perc + Age_AtEncounterDate + DiabetesDuration + PumpUse + Race.Ethnicity-1, -->
<!--                random=~1|EPICMRN,data = dat, na.action = na.omit) -->
<!-- # Model selection -->
<!-- summary(mod_adj_int) -->
<!-- ``` -->

<!-- ```{r echo=FALSE,warning=FALSE,dpi=600} -->
<!-- # Plot -->
<!-- mm_plot <- ggplot(data = dat, aes(x = CGM_Perc, y = HbA1c,group = EPICMRN)) + -->
<!--   geom_smooth(aes(x = CGM_Perc, y = HbA1c),colour="black",method='lm',inherit.aes = F,size=0.5) + -->
<!--   xlab("% CGM Use") + -->
<!--   ylab("HbA1c (%)") + -->
<!--   theme_bw() -->
<!-- mm_plot -->
<!-- ``` -->

<!-- ### Table 6: Mixed model HbA1c by % CGM Use, adjusted for age, race, diabetes duration, and pump use -->

<!-- ```{r echo=FALSE} -->
<!-- kable(summary(mod_adj_int)$tTable,caption = "Fixed Effects") -->
<!-- ``` -->

<!-- The interpretation of this mixed effects model is similar to a simple linear model. For example, on average each 1 point increase in % CGM use resulted in a decrease in HbA1c of 0.005 after adjusting for age, race, diabetes duration, and pump use. -->

<!-- Median [IQR] CGM use in days across all subjects and all visits (excluding missing values): -->
<!-- ```{r echo=FALSE} -->
<!-- print(paste0(median(dat$X30.day.CGM.use,na.rm = T)," [", -->
<!--              quantile(dat$X30.day.CGM.use,na.rm = T)[2],",", -->
<!--              quantile(dat$X30.day.CGM.use,na.rm = T)[4],"]")) -->
<!-- ``` -->

<!-- Percent of CGM users with at least 85% wear: -->

<!-- ```{r echo=FALSE} -->
<!-- wear <- dat %>% -->
<!--   filter(CGMUseSomeNever == "Some") %>% -->
<!--   group_by(EPICMRN) %>% -->
<!--   summarise(max_wear = max(X30.day.CGM.use,na.rm = T), -->
<!--             min_wear = min(X30.day.CGM.use,na.rm = T)) -->
<!-- round(length(which(wear$max_wear >= 25)) / nrow(wear) * 100,3) -->
<!-- ``` -->

<!-- Percent of CGM users with 0% wear: -->

<!-- ```{r echo=FALSE} -->
<!-- round(length(which(wear$min_wear == 0)) / nrow(wear) * 100,3) -->
<!-- ``` -->

<!-- ### Figure 2: % CGM Use Over Time -->

<!-- ```{r echo=FALSE} -->
<!-- # Average CGM use change over time -->
<!-- dat <- dat %>% -->
<!--   arrange(EPICMRN,EncounterDate) %>% -->
<!--   group_by(EPICMRN) %>% -->
<!--   mutate(Days = as.numeric(difftime(EncounterDate,EncounterDate[1],units = "days"))) -->
<!-- # Plot -->
<!-- use_plot <- ggplot(data = dat, aes(x = Days, y = CGM_Perc,group = as.factor(EPICMRN))) + -->
<!--   geom_line(size = 0.2,aes(color = EPICMRN)) + -->
<!--   ylab("% CGM Use") + -->
<!--   xlab("Days from First Visit") + -->
<!--   theme(legend.position = "none") -->
<!-- ``` -->

<!-- ```{r echo=FALSE,warning=FALSE,dpi=600} -->
<!-- use_plot -->
<!-- ``` -->

<!-- ```{r echo=FALSE,dpi=600} -->
<!-- hist(dat$CGM_Perc,xlab = "CGM_Perc") -->
<!-- hist(dat$Days,xlab = "Days") -->
<!-- ``` -->

<!-- ### Table 7: Mixed model % CGM Use by Days from First Visit -->

<!-- ```{r echo=FALSE} -->
<!-- # Model -->
<!-- use_mod <- lme(CGM_Perc ~ Days,random = ~1|EPICMRN,data = dat,na.action = na.omit) -->
<!-- ``` -->

<!-- ```{r echo=FALSE} -->
<!-- # kable(summary(use_mod)$tTable,caption = "Fixed Effects") -->
<!-- # plot(use_mod) -->
<!-- ``` -->

### Table 8: Descriptive statistics for all individuals with TIR data

```{r echo=FALSE,include=FALSE}
# Average TIR
tir <- dat %>%
  filter(CGM_Perc > 0)
tir_dem <- tir %>%
  group_by(EPICMRN) %>%
  summarise(Mean_CGM_Low = mean(CGMLowPercent,na.rm = T),
            Mean_CGM_OK = mean(CGMOKPercent,na.rm = T),
            Mean_CGM_High = mean(CGMHighPercent,na.rm = T),
            Mean_HbA1c = mean(HbA1c,na.rm = T))
tir_dem[tir_dem == "NaN"] <- NA
tir_dem <- left_join(tir_dem,demographics,by = "EPICMRN")
vars <- c("AgeFirstVisit","DiabetesDuration","Mean_HbA1c","Mean_CGM_Low","Mean_CGM_OK","Mean_CGM_High","Gender","Race.Ethnicity")
nonnormal <- norm.check(tir_dem,c("AgeFirstVisit","DiabetesDuration","Mean_HbA1c","Mean_CGM_Low","Mean_CGM_OK","Mean_CGM_High"))
tir_t1 <- CreateTableOne(vars,data = tir_dem)
tir_t1 <- print(tir_t1,nonnormal = c(nonnormal,"Mean_CGM_High"))
```

```{r echo=FALSE}
kable(tir_t1)
```

<!-- ### Figure 3: HbA1c by % Time in Range -->

<!-- ```{r echo=FALSE,warning=FALSE,dpi=600} -->
<!-- tir_plot <- ggplot(data = tir, aes(x = CGMOKPercent, y = HbA1c,group = EPICMRN)) + -->
<!--   geom_smooth(aes(x = CGMOKPercent, y = HbA1c),colour="black",method='lm',inherit.aes = F,size=0.5) + -->
<!--   xlab("% Time in Range") + -->
<!--   ylab("HbA1c (%)") + -->
<!--   theme_bw() -->
<!-- tir_plot -->
<!-- ``` -->

<!-- ### Table 9: Mixed model HbA1c by % Time in Range -->

<!-- ```{r echo=FALSE} -->
<!-- # Model -->
<!-- tir_mod <- lme(HbA1c ~ CGMOKPercent,random = ~1|EPICMRN,data = tir,na.action = na.omit) -->
<!-- ``` -->

<!-- ```{r echo=FALSE} -->
<!-- kable(summary(tir_mod)$tTable,caption = "Fixed Effects") -->
<!-- ``` -->

<!-- Here the intercept refers to the average HbA1c for participants with 0 % TIR. For each 1 percentage point increase in TIR, HbA1c decreases by 0.04 on average. -->