---
title: "Holly's Dissertation"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(car)
library(skimr)
library(knitr)
library(MASS)
library(lme4)
library(nlme)
library(tidyverse)
```

```{r echo=FALSE,include=FALSE}
# Import data
dat <- 
  read.csv("/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Holly O'Donnell/Dissertation/Data_Raw/diss_data_7.22.19.csv",
           na.strings = "")
# format and fill down dates and income level
dat$DateofBirth <- lubridate::mdy(dat$DateofBirth)
dat$DiagnosisDate <- lubridate::mdy(dat$DiagnosisDate)
dat$PumpStartDate <- lubridate::mdy(dat$PumpStartDate)
dat$VisitDate <- lubridate::mdy(dat$VisitDate)
dat <- dat %>% group_by(ID) %>% fill(DateofBirth,IncomeLevel,DiagnosisDate,
                                     PumpStartDate,Insurance,Sex) 
# Calculate age, T1D duration, and pump duration at each visit
dat$age_days <- difftime(dat$VisitDate,dat$DateofBirth,units = "days")
dat$age_years <- as.numeric(dat$age_days / 365.25)
dat$T1D_dur_days <- difftime(dat$VisitDate,dat$DiagnosisDate,units = "days")
dat$T1D_dur_years <- as.numeric(dat$T1D_dur_days / 365.25)
dat$pump_dur_days <- difftime(dat$VisitDate,dat$PumpStartDate,units = "days")
dat$pump_dur_yrs <- as.numeric(dat$pump_dur_days / 365.25)
# Add site field
dat$Site <- ifelse(as.numeric(dat$ID) < 300,"Florida","Colorado")
# Make study visit and treatment categorical
dat$StudyVisit <- as.factor(dat$StudyVisit)
dat$Treatment <- as.factor(dat$Treatment)
dat$IncomeLevel <- as.factor(dat$IncomeLevel)
dat$Sex <- as.factor(dat$Sex)
dat$Insurance <- as.factor(dat$Insurance)
# Days from baseline, baseline levels of outcome
dat <- dat %>% group_by(ID) %>% 
  mutate(Days_since_baseline = 
           as.numeric(difftime(VisitDate,VisitDate[1],units = "days")),
         readings_per_day_b = readings_per_day[1],
         highBG_without_carb_with_bolus_b = highBG_without_carb_with_bolus_piu[1],
         carb_inputs_per_day_b = carb_inputs_per_day[1],
         boluses_per_day_b = boluses_per_day[1],
         extremeBG_without_carb_with_bolus_b = 
           extremeBG_without_carb_with_bolus_piu[1])
# Change scores
dat <- dat %>% group_by(ID) %>%
  mutate(readings_per_day_change = readings_per_day - readings_per_day[1],
         carb_inputs_per_day_change = carb_inputs_per_day - carb_inputs_per_day[1],
         boluses_per_day_change = boluses_per_day - boluses_per_day[1],
         highBG_change = highBG_without_carb_with_bolus_piu - 
           highBG_without_carb_with_bolus_piu[1],
         extremeBG_change = extremeBG_without_carb_with_bolus_piu - 
           extremeBG_without_carb_with_bolus_piu[1]) %>% ungroup()
# Center age and HbA1c
dat <- dat %>% group_by(StudyVisit) %>% 
  mutate(HbA1c = HbA1c - mean(HbA1c,na.rm=T),
         age_years = age_years - mean(age_years,na.rm=T)) %>% ungroup()
# Collapse income categories
dat$income_three <- as.character(dat$IncomeLevel)
dat$income_three[dat$income_three == "1" | dat$income_three == "2"] <- 1
dat$income_three[dat$income_three == "3" | dat$income_three == "4"] <- 2
dat$income_three[dat$income_three == "5" | dat$income_three == "6"] <- 3
dat$income_three <- factor(dat$income_three,labels = c("Low","Medium","High"))
# Check insurance pediction by income
mylogit <- glm(Insurance ~ income_three, data = dat[dat$StudyVisit==1,], family = "binomial")
summary(mylogit) # SE looks good
# Only include those with >= 40 days of wear
dat <- dat %>% filter(days_worn >= 40)
```

# BG Checks per Day

## Final model

```{r}
bg_mod <- lmer(readings_per_day ~ Treatment*StudyVisit + HbA1c + Sex + 
                age_years + income_three + readings_per_day_b+(1|Site/ID),
               data = dat)
```

# Percent high boluses without carb

## Final model

```{r}
highBG_mod <- lme(highBG_without_carb_with_bolus_piu ~ Treatment*StudyVisit + 
                   HbA1c + Sex + age_years + income_three + 
                   highBG_without_carb_with_bolus_b, random = ~1|Site/ID, 
                  data = dat,na.action = na.omit)
```

```{r include=FALSE}
plot(highBG_mod,main="% bolused when high (150-250) and no carb")
```

### Results

```{r echo=FALSE}
kable(anova(highBG_mod),caption = "Overall Effects")
kable(summary(highBG_mod)$tTable,caption = "Fixed Effects")
```

# Carb inputs per day

## Final model

```{r}
carb_mod <- lme(carb_inputs_per_day ~ Treatment*StudyVisit + HbA1c + Sex + 
                  age_years + income_three + carb_inputs_per_day_b, random = ~1|Site/ID,
                data = dat,na.action = na.omit)
```

```{r include=FALSE}
plot(carb_mod,main="Carb Inputs per Day",idLabels=~ID,id=0.0001)
```

### Results

```{r echo=FALSE}
kable(anova(carb_mod),caption = "Overall Effects")
kable(summary(carb_mod)$tTable,caption = "Fixed Effects")
```

# Boluses per day

## Final model

```{r}
bolus_day_mod <- lmer(boluses_per_day ~ Treatment*StudyVisit + HbA1c + Sex + 
                  age_years + income_three + boluses_per_day_b + (1|Site/ID),
                  data = dat)
```

# Percent extreme boluses without carb

## Final model

```{r}
extreme_bolus_mod <- lmer(extremeBG_without_carb_with_bolus_piu ~ 
                            Treatment*StudyVisit + HbA1c + Sex + age_years + 
                            income_three + extremeBG_without_carb_with_bolus_b + (1|Site/ID),data = dat)
```

# Psych Variables

## Check convergence

```{r}
# BG readings
reading_mod <- lmer(readings_per_day ~ Treatment*zDEPRESSION + StudyVisit + 
                   Treatment:StudyVisit + Sex + age_years + income_three + 
                   readings_per_day_b + (1|Site/ID), data = dat)
# High BG without carb
high_mod <- lmer(highBG_without_carb_with_bolus_piu ~ Treatment*zDEPRESSION + 
                   StudyVisit + Treatment:StudyVisit + Sex + age_years + 
                   income_three + highBG_without_carb_with_bolus_b + (1|Site/ID),
                 data = dat)
```

The models still aren't working with the combined income categories, so I think it's best to still use change scores.

```{r echo=FALSE,eval=FALSE}
# Model selection
# Make non-missing dataframe
step_data <- dat %>% select(ID,readings_per_day_change,Treatment,zDEPRESSION,
                            HbA1c,StudyVisit,Sex,age_years,T1D_dur_years,
                            IncomeLevel,Insurance,income_three)
step_data <- step_data[complete.cases(step_data),]
# Stepwise selection by AIC
full_mod_l <- lme(readings_per_day_change ~ Treatment*zDEPRESSION*StudyVisit +
                     Sex + age_years + T1D_dur_years + Treatment*income_three +
                    HbA1c,random = ~1|ID,data = step_data,
                  na.action = na.omit,method = "ML")
step_mod <- stepAIC(full_mod_l)
# The model readings_per_day_change ~ Treatment + zDEPRESSION + StudyVisit +
# Insurance + Treatment:StudyVisit + zDEPRESSION:StudyVisit
# is the best, and also makes the most sense in terms of the question.

# Check another outcome
step_data <- dat %>% select(ID,highBG_change,Treatment,zDEPRESSION,
                            HbA1c,StudyVisit,Sex,age_years,T1D_dur_years,
                            IncomeLevel,Insurance,income_three)
step_data <- step_data[complete.cases(step_data),]
full_mod_l <- lme(highBG_change ~ Treatment*zDEPRESSION*StudyVisit +
                     Sex + age_years + T1D_dur_years + Treatment*income_three +
                     HbA1c,random = ~1|ID,data = step_data,
                  na.action = na.omit,method = "ML")
step_mod <- stepAIC(full_mod_l)

# highBG_change ~ Treatment + zDEPRESSION + StudyVisit + Sex + age_years +
# T1D_dur_years + IncomeLevel + Insurance + Treatment:zDEPRESSION +
# Treatment:IncomeLevel + Treatment:Insurance
# is the best model that includes Treatment:zDEPRESSION.

# One more outcome.
step_data <- dat %>% select(ID,carb_inputs_per_day_change,Treatment,zDEPRESSION,
                            HbA1c,StudyVisit,Sex,age_years,T1D_dur_years,
                            IncomeLevel,Insurance)
step_data <- step_data[complete.cases(step_data),]
full_mod_l <- lme(carb_inputs_per_day_change ~ Treatment*zDEPRESSION*StudyVisit +
                     Sex + age_years + T1D_dur_years + Treatment*IncomeLevel +
                     Treatment*Insurance,random = ~1|ID,data = step_data,
                  na.action = na.omit,method = "ML")
step_mod <- stepAIC(full_mod_l)
# carb_inputs_per_day_change ~ Treatment + zDEPRESSION + StudyVisit +
# Treatment:zDEPRESSION
# is the best. Because high BG change has a lot of covariates but the other ones
# don't, I think it makes sense to stick with a model similar to the outcomes
# above, but without HbA1c:
# outcome ~ Treatment + zDEPRESSION + StudyVisit + Treatment:StudyVisit +
# zDEPRESSION:StudyVisit + Sex + age_years + IncomeLevel + Insurance
```

## Change Scores Baseline to T3

```{r echo=FALSE}
# Split the change score model into two linear models without random effects, e.g. change at T2 ~ depression, change at T3 ~ depression etc. Keep same covariates as before, since we don't have to worry about singular fits, etc. Change at T3 only for now.
t3 <- dat %>% filter(StudyVisit == 3)
vars <- c("HbA1c","Sex","age_years","T1D_dur_years","Insurance","income_three")
t1 <- CreateTableOne(vars,strata = "Treatment",data = t3)
kableone(t1)
```

Because there is a significant difference in sex at T3, we'll adjust for it in the linear models. Also, even though insurance technically isn't significantly different, it was at T1 and we adjusted for it earlier, so we'll include it in these models as well. Everything else is the same between the two groups, and we're not interested in the effects, so we can leave them out of the models.

All p values FDR-adjusted for multiple comparisons.

```{r eval=FALSE,echo=FALSE}
# More model selection
# Singular fit
bg_checks_mod <- lmer(readings_per_day_change ~ Treatment*zDEPRESSION +
                       StudyVisit + Treatment:StudyVisit +
                       Sex + age_years + IncomeLevel + Insurance + (1|ID),
                     data = dat)
# Should I start with this model and remove least important covariates until the fit isn't singular?
# I think the issue that we basically have 2 measures per person, so can't support a complicated model?

# Okay fit
bg_checks_mod <- lmer(readings_per_day_change ~ Treatment*zDEPRESSION +
                       StudyVisit + Treatment:StudyVisit + (1|ID),
                     data = dat)
plot(bg_checks_mod)
bg_checks_mod <- lme(readings_per_day_change ~ Treatment*zDEPRESSION +
                       StudyVisit + Treatment:StudyVisit,random = ~1|ID,
                     data = dat,na.action = na.omit)
kable(summary(bg_checks_mod)$tTable)
```

```{r echo=FALSE}
# Depression
bg_depress_mod <- lm(readings_per_day_change ~ Treatment*zDEPRESSION + Sex + income_three,data = t3)
results <- as.data.frame(summary(bg_depress_mod)$coefficients)
results$model <- "bg_depress_mod"
adjust_dep <- results

carb_depress_mod <- lm(carb_inputs_per_day_change ~ Treatment*zDEPRESSION + Sex + income_three,data = t3)
results <- as.data.frame(summary(carb_depress_mod)$coefficients)
results$model <- "carb_depress_mod"
adjust_dep <- rbind(adjust_dep,results)

bolus_depress_mod <- lm(boluses_per_day_change ~ Treatment*zDEPRESSION + Sex + income_three,data = t3)
results <- as.data.frame(summary(bolus_depress_mod)$coefficients)
results$model <- "bolus_depress_mod"
adjust_dep <- rbind(adjust_dep,results)

highbg_depress_mod <- lm(highBG_change ~ Treatment*zDEPRESSION + Sex + income_three,data = t3)
results <- as.data.frame(summary(highbg_depress_mod)$coefficients)
results$model <- "highbg_depress_mod"
adjust_dep <- rbind(adjust_dep,results)

extremebg_depress_mod <- lm(extremeBG_change ~ Treatment*zDEPRESSION + Sex + income_three,data = t3)
results <- as.data.frame(summary(extremebg_depress_mod)$coefficients)
results$model <- "extremebg_depress_mod"
adjust_dep <- rbind(adjust_dep,results)

#FOH Help
bg_help_mod <- lm(readings_per_day_change ~ Treatment*FOH_C_Help + Sex + income_three,data = t3)
results <- as.data.frame(summary(bg_help_mod)$coefficients)
results$model <- "bg_help_mod"
adjust_help <- results

carb_help_mod <- lm(carb_inputs_per_day_change ~ Treatment*FOH_C_Help + Sex + income_three,data = t3)
results <- as.data.frame(summary(carb_help_mod)$coefficients)
results$model <- "carb_help_mod"
adjust_help <- rbind(adjust_help,results)

bolus_help_mod <- lm(boluses_per_day_change ~ Treatment*FOH_C_Help + Sex + income_three,data = t3)
results <- as.data.frame(summary(bolus_help_mod)$coefficients)
results$model <- "bolus_help_mod"
adjust_help <- rbind(adjust_help,results)

highbg_help_mod <- lm(highBG_change ~ Treatment*FOH_C_Help + Sex + income_three,data = t3)
results <- as.data.frame(summary(highbg_help_mod)$coefficients)
results$model <- "highbg_help_mod"
adjust_help <- rbind(adjust_help,results)

extremebg_help_mod <- lm(extremeBG_change ~ Treatment*FOH_C_Help + Sex + income_three,data = t3)
results <- as.data.frame(summary(extremebg_help_mod)$coefficients)
results$model <- "extremebg_help_mod"
adjust_help <- rbind(adjust_help,results)

# FOH Maintain
bg_maint_mod <- lm(readings_per_day_change ~ Treatment*FOH_C_Maint + Sex + income_three,data = t3)
results <- as.data.frame(summary(bg_maint_mod)$coefficients)
results$model <- "bg_maint_mod"
adjust_maint <- results

carb_maint_mod <- lm(carb_inputs_per_day_change ~ Treatment*FOH_C_Maint + Sex + income_three,data = t3)
results <- as.data.frame(summary(carb_maint_mod)$coefficients)
results$model <- "carb_maint_mod"
adjust_maint <- rbind(adjust_maint,results)

bolus_maint_mod <- lm(boluses_per_day_change ~ Treatment*FOH_C_Maint + Sex + income_three,data = t3)
results <- as.data.frame(summary(bolus_maint_mod)$coefficients)
results$model <- "bolus_maint_mod"
adjust_maint <- rbind(adjust_maint,results)

highbg_maint_mod <- lm(highBG_change ~ Treatment*FOH_C_Maint + Sex + income_three,data = t3)
results <- as.data.frame(summary(highbg_maint_mod)$coefficients)
results$model <- "highbg_maint_mod"
adjust_maint <- rbind(adjust_maint,results)

extremebg_maint_mod <- lm(extremeBG_change ~ Treatment*FOH_C_Maint + Sex + income_three,data = t3)
results <- as.data.frame(summary(extremebg_maint_mod)$coefficients)
results$model <- "extremebg_maint_mod"
adjust_maint <- rbind(adjust_maint,results)
# Adjust p values
adjust_dep$`Pr(>|t|)` <- p.adjust(adjust_dep$`Pr(>|t|)`,"fdr")
adjust_help$`Pr(>|t|)` <- p.adjust(adjust_help$`Pr(>|t|)`,"fdr")
adjust_maint$`Pr(>|t|)` <- p.adjust(adjust_maint$`Pr(>|t|)`,"fdr")
```

## Depression

### BG Checks per Day

```{r echo=FALSE}
results <- adjust_dep[adjust_dep$model == "bg_depress_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","zDEPRESSION","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:zDEPRESSION")
kable(results)
```

### Carb Inputs per Day

```{r echo=FALSE}
results <- adjust_dep[adjust_dep$model == "carb_depress_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","zDEPRESSION","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:zDEPRESSION")
kable(results)
```

### Boluses per Day

```{r echo=FALSE}
results <- adjust_dep[adjust_dep$model == "bolus_depress_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","zDEPRESSION","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:zDEPRESSION")
kable(results)
```

### High BG Without Carb With Bolus

```{r echo=FALSE}
results <- adjust_dep[adjust_dep$model == "highbg_depress_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","zDEPRESSION","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:zDEPRESSION")
kable(results)
```

### Extreme BG Without Carb With Bolus

```{r echo=FALSE}
results <- adjust_dep[adjust_dep$model == "extremebg_depress_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","zDEPRESSION","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:zDEPRESSION")
kable(results)
```

## FOH Help

### BG Checks per Day

```{r echo=FALSE}
results <- adjust_help[adjust_help$model == "bg_help_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Help","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Help")
kable(results)
```

### Carb Inputs per Day

```{r echo=FALSE}
results <- adjust_help[adjust_help$model == "carb_help_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Help","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Help")
kable(results)
```

### Boluses per Day

```{r echo=FALSE}
results <- adjust_help[adjust_help$model == "bolus_help_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Help","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Help")
kable(results)
```

### High BG Without Carb With Bolus

```{r echo=FALSE}
results <- adjust_help[adjust_help$model == "highbg_help_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Help","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Help")
kable(results)
```

### Extreme BG Without Carb With Bolus

```{r echo=FALSE}
results <- adjust_help[adjust_help$model == "extremebg_help_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Help","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Help")
kable(results)
```

## FOH Maintain

### BG Checks per Day

```{r echo=FALSE}
results <- adjust_maint[adjust_maint$model == "bg_maint_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Maint","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Maint")
kable(results)
```

### Carb Inputs per Day

```{r echo=FALSE}
results <- adjust_maint[adjust_maint$model == "carb_maint_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Maint","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Maint")
kable(results)
```

### Boluses per Day

```{r echo=FALSE}
results <- adjust_maint[adjust_maint$model == "bolus_maint_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Maint","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Maint")
kable(results)
```

### High BG Without Carb With Bolus

```{r echo=FALSE}
results <- adjust_maint[adjust_maint$model == "highbg_maint_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Maint","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Maint")
kable(results)
```

### Extreme BG Without Carb With Bolus

```{r echo=FALSE}
results <- adjust_maint[adjust_maint$model == "extremebg_maint_mod",1:4]
rownames(results) <- c("(Intercept)","Treatment1","FOH_C_Maint","Sex1",
                       "income_threeMedium","income_threeHigh",
                       "Treatment1:FOH_C_Maint")
kable(results)
```
