---
title: "TCL Analysis - Maki"
author: "Kristen Campbell"
date: "1/17/2020"
output: word_document
---

```{r,echo=FALSE,message=FALSE,results='hide',warning=FALSE}
library(knitr)
#library(labelled)
options(scipen=999)
#code for data prep: 
source("C:/Users/campbkri/Documents/GitHub/BDC-Code/Maki Nakayama/TCL/00_analysis.R")


```

#Methods

Figure 1: One way ANOVA was used to test for differences in the average frequency of clonotypes by the number of mice sharing that clonotype.  The overall F-test was reported to test for any differences in average frequency by mice number, and Tukey’s Honestly Significant Difference (Tukey’s HSD) was used to perform pairwise comparisons while maintaining an alpha of 0.05.

Figure 3: A linear mixed model was fit to %shared unique clonotypes by rank.  An interaction between categorical rank by group (PLN vs. blood) was included.  Constrasts using least square (LS) means were used to test for differences between PLN and blood at each rank. 

Figure 4: A linear mixed model was fit to %public NY8.3 over time, with a random intercept for mouse to account for correlation within a mouse over time.  An interaction between categorical time by group was included.  Constrasts using least square (LS) means were used to test for differences between groups at each time point.  Fisher's exact tests were conducted to test for differences in `%`>0.5 and `%`>0.6 between wild-type NOD and insulin-KO NOD.

# Summary Statistics:

## Figure 1B:
No statistics necessary, this is considered a histogram (a description of clonotypes by mice)

## Figure 1c (alpha):
Unit of observation is the clonotype.  For each clonotype, how many mice shared that clonotype (group 1-7), and what was the average frequency of the clonotype?

*Table 1c: ALPHA, comparisons between number of mice sharing unique clonotypes*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tuk_a, align = c("l", "l"))
```

Overall, there was a significant difference between number of mice sharing a clonotype and average frequency of clonotypes (p=`r all.p_a`).

## Figure 1c (beta):

*Table 1c: BETA, comparisons between number of mice sharing unique clonotypes*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tuk_b, align = c("l", "l"))
```

Overall, there was a significant difference between number of mice sharing a clonotype and average frequency of clonotypes (p=`r all.p_b`).

## Figure 3:

*Table 3a: ALPHA, comparisons beetween PLN and Blood*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3a, align = c("l", "l"))
```

*Table 3b: BETA, comparisons beetween PLN and Blood*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_3b, align = c("l", "l"))
```

## Figure 4:

*Table 4a: %PUBLIC - comparisons at each timepoint*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_4a, align = c("l", "l"))
```

*Table 4b: %EXTENDED - comparisons at each timepoint*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(contrasts_4b, align = c("l", "l"))
```

*Table 4a: %PUBLIC - >0.5% VS. <0.5%*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.4c, align = c("l", "l"))
```

*Table 4B: %EXTENDED - >0.5% VS. <0.5%*
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab.4d, align = c("l", "l"))
```