---
title: "Health Related Quality of Life During COVID (ADA 2022)"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(redcapAPI)
library(skimr)
library(knitr)
library(tidyverse)
library(Hmisc)
library(arsenal)
library(emmeans)
library(lme4)
library(lmerTest)
library(DT)
library(broom.mixed)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# Import
source("./Erin Cobry/ADA 2022/COVID/Data raw/HealthRelatedQuality_R_2022-02-01_1723.r")
# Exclude participants per Erin (missing consent)
exclude = c("1012","2012","1018","2018","1028","2028","1037","2037",
            "1045","2045","1060","2060","1064","2064","1079","2079",
            "1088","2088","1103","2103","1108","2108","1115","2115")
exclude = c(sapply(exclude, grep,data$record_id))
data = data[-exclude,]
```

```{r}
# Score surveys
# Exposure
exposure = data %>% select(record_id,redcap_event_name,stay_at_home.factor:died.factor)
data$exposure_sum = apply(exposure[,3:ncol(exposure)],1,function(r){
  sum(r == "Yes")
})
# Impact
impact = data %>% select(record_id,redcap_event_name,parenting:mood)
# Convert 5 to NA
impact[,3:ncol(impact)] = lapply(impact[,3:ncol(impact)],function(c){
  c[c==5]=NA
  c
})
# Mean of all items - NA if more than 3 are missing
data$impact_mean = apply(impact[,3:ncol(impact)],1,function(r){
  if(sum(is.na(r))>3){NA}else{
    mean(as.numeric(r),na.rm = T)
  }
})
# Distress
distress = data %>% select(record_id,redcap_event_name,distress:child_distress)
data$family_distress = apply(distress[,3:ncol(distress)],1,mean,na.rm = T)
data$family_distress[is.nan(data$family_distress)] = NA

# Clean up
rm(distress,exposure,impact,exclude)
```

```{r results='asis'}
# Score PEDS-QL
source("~/GitHub/BDC-Code/Erin Cobry/COVID Quality of Life/score_pedsql.r")
```

# Table 1

```{r results='asis'}
# Get demographic information - per Erin "The child demographic information is under the child IDs (so all of the 2000 IDs, since the 1000 IDs are all of the parents)"
# Convert age to years
demographics = data %>% 
  filter(grepl("baseline",redcap_event_name),as.numeric(substr(record_id,1,4))>=2000) %>%
  mutate(age = as.numeric(child_age) / 365.25,
         t1d_duration = as.numeric(t1d_duration) / 365.25)
# Table 1 
t1_form = as.formula(" ~ age + t1d_duration + gender.factor + child_insurance.factor + race.factor")
newlabels = list(age = "Age (Years)",t1d_duration = "T1D Duration",race.factor = "Race",
                 gender.factor = "Gender", child_insurance.factor = "Insurance")
t1 = tableby(t1_form,demographics)
summary(t1,labelTranslations = newlabels)
```

\newpage

# CEFIS Scores

```{r}
# Filter out unnecessary rows and rename time points ("arm" refers to child's age)
scores = data %>% 
  filter(as.numeric(substr(record_id,1,4))<2000) %>%
  mutate(Timepoint = sub("_arm_.*","",redcap_event_name))
scores$num_id = as.numeric(substr(scores$record_id,1,4))
scores$Timepoint = factor(scores$Timepoint,levels = c("baseline","3_month",
                                                      "6_month","9_month","12_month"),
                          labels = c("Baseline","3 Month","6 Month","9_month","12 Month"))
scores$num_time = recode(scores$Timepoint,"Baseline" = 0,"3 Month" = 3,
                         "6 Month" = 6,"9 Month"=9,"12 Month" = 12)

scores_child = data %>% 
  filter(as.numeric(substr(record_id,1,4))>=2000) %>%
  mutate(Timepoint = sub("_arm_.*","",redcap_event_name))
scores_child$num_id = as.numeric(substr(scores_child$record_id,1,4))
scores_child$Timepoint = factor(scores_child$Timepoint,levels = c("baseline","3_month",
                                                      "6_month","9_month","12_month"),
                          labels = c("Baseline","3 Month","6 Month","9 Month","12 Month"))
scores_child$num_time = recode(scores_child$Timepoint,"Baseline" = 0,"3 Month" = 3,
                         "6 Month" = 6,"9 Month"=9,"12 Month" = 12)

```

```{r plot function}
model_and_plot = function(outcome_name,ylab,group_var,df,pois = F){
  # Convert to numeric
  df[,outcome_name] = as.numeric(df[,outcome_name])
  # Fit a mixed model. Some of these are essentially count data and may require a Poisson model
  mod = lmer(as.formula(paste0(outcome_name,"~Timepoint+(1|",group_var,")")),df)
  # Means by timepoint
  means = as.data.frame(emmeans(mod,~Timepoint))
  means$num_time = c(0,3,6,9,12)
  # Plot
  p = ggplot(df,aes_string(x = "num_time",y = outcome_name, group = group_var)) +
    geom_line(alpha=0.1) +
    geom_line(data=means,aes(x=num_time,y=as.numeric(emmean),group=1),
            size = 1,inherit.aes = F) +
    theme_bw() +xlab("Timepoint") + ylab(ylab) +
    scale_x_continuous(breaks=c(0,3,6,9,12),labels=c("Baseline","3 Month","6 Month","9 Month","12 Month"))
  # Print plot and model results
  print(p)
  print(kable(anova(mod),digits = 3,caption = "Type III ANOVA"))
  print(kable(summary(mod)$coefficients,digits = 3,caption = "Fixed Effects"))
}
```

```{r plot_no12 function}
model_and_plot_no12 = function(outcome_name,ylab,group_var,df,pois = F){
  # Convert to numeric
  df[,outcome_name] = as.numeric(df[,outcome_name])
  # Fit a mixed model. Some of these are essentially count data and may require a Poisson model
  mod = lmer(as.formula(paste0(outcome_name,"~Timepoint+(1|",group_var,")")),df)
  # Means by timepoint
  means = as.data.frame(emmeans(mod,~Timepoint))
  means$num_time = c(0,3,6,9)
  # Plot
  p = ggplot(df,aes_string(x = "num_time",y = outcome_name, group = group_var)) +
    geom_line(alpha=0.1) +
    geom_line(data=means,aes(x=num_time,y=as.numeric(emmean),group=1),
            size = 1,inherit.aes = F) +
    theme_bw() +xlab("Timepoint") + ylab(ylab) +
    scale_x_continuous(breaks=c(0,3,6,9),labels=c("Baseline","3 Month","6 Month","9 Month"))
  # Print plot and model results
  print(p)
  print(kable(anova(mod),digits = 3,caption = "Type III ANOVA"))
  print(kable(summary(mod)$coefficients,digits = 3,caption = "Fixed Effects"))
}
```

```{r tvc function}
# Trying to make a function for a mixed model with time varying covariates
tvc = function(outcome_name,group_var,df,pois = F){
  # Convert to numeric
  df[,outcome_name] = as.numeric(df[,outcome_name])
  # Fit a mixed model. Some of these are essentially count data and may require a Poisson model
  mod = lmer(as.formula(paste0(outcome_name,"~num_time +",group_var,"+ (num_time|record_id)")),df)
  # Means by timepoint
  #means = as.data.frame(emmeans(mod,~Timepoint))
  #means$num_time = c(0,3,6)
  # Plot
  #p = ggplot(df,aes_string(x = "num_time",y = outcome_name, group = group_var)) +
  #  geom_line(alpha=0.1) +
  #  geom_line(data=means,aes(x=num_time,y=as.numeric(emmean),group=1),
  #          size = 1,inherit.aes = F) +
  #  theme_bw() +xlab("Timepoint") + ylab(ylab) +
  #  scale_x_continuous(breaks=c(0,3,6),labels=c("Baseline","3 Month","6 Month"))
  # Print plot and model results
  #print(p)
  print(kable(anova(mod),digits = 3,caption = "Type III ANOVA"))
  print(kable(summary(mod)$coefficients,digits = 3,caption = "Fixed Effects"))
}
```

## Exposure

```{r results='asis'}
model_and_plot("exposure_sum","Exposure","record_id",scores[scores$exposure_sum>0,])
```

## Impact

```{r results='asis'}
model_and_plot("impact_mean","Impact","record_id",scores[scores$exposure_sum>0,])
```

## Distress

```{r results='asis'}
model_and_plot("distress","Distress","record_id",scores[scores$exposure_sum>0,])
```

## Child Distress

```{r results='asis'}
model_and_plot("child_distress","Child Distress","record_id",scores[scores$exposure_sum>0,])
```

## Family Distress

```{r results='asis'}
model_and_plot("family_distress","Family Distress","record_id",scores[scores$exposure_sum>0,])
```

\newpage

# Peds-QL Diabetes Module Scores

```{r}
# Filter out unnecessary rows and rename time points ("arm" refers to child's age)
child_pedsql = final_pedsql %>% 
  filter(as.numeric(substr(record_id,1,4))>=2000) %>%
  mutate(Timepoint = sub("_arm_.*","",redcap_event_name))
child_pedsql$num_id = as.numeric(substr(child_pedsql$record_id,1,4))
child_pedsql$Timepoint = factor(child_pedsql$Timepoint,levels = c("baseline","3_month",
                                                      "6_month","9_month","12_month"),
                          labels = c("Baseline","3 Month","6 Month","9 Month","12 Month"))
child_pedsql$num_time = recode(child_pedsql$Timepoint,"Baseline" = 0,"3 Month" = 3,
                         "6 Month" = 6,"9 Month" = 9,"12 Month" = 12)

parent_pedsql = final_pedsql %>% 
  filter(as.numeric(substr(record_id,1,4))<2000) %>%
  mutate(Timepoint = sub("_arm_.*","",redcap_event_name))
parent_pedsql$num_id = as.numeric(substr(parent_pedsql$record_id,1,4))
parent_pedsql$Timepoint = factor(parent_pedsql$Timepoint,levels = c("baseline","3_month",
                                                      "6_month","9_month","12_month"),
                          labels = c("Baseline","3 Month","6 Month","9 Month","12 Month"))
parent_pedsql$num_time = recode(parent_pedsql$Timepoint,"Baseline" = 0,"3 Month" = 3,
                         "6 Month" = 6,"9 Month" = 9,"12 Month" = 12)
```

## Child Diabetes

```{r results='asis'}
model_and_plot_no12("diabetes_score","Diabetes score","record_id",child_pedsql[child_pedsql$diabetes_score>0,])
```

## Child Treatment I

```{r results='asis'}
model_and_plot_no12("tx1_score","Treatment I score","record_id",child_pedsql[child_pedsql$tx1_score>0,])
```

## Child Treatment II

```{r results='asis'}
model_and_plot_no12("tx2_score","Treatment II score","record_id",child_pedsql[child_pedsql$tx2_score>0,])
```

## Child Worry

```{r results='asis'}
model_and_plot_no12("worry_score","Worry score","record_id",child_pedsql[child_pedsql$worry_score>0,])
```

## Child Communication

```{r results='asis'}
model_and_plot_no12("comm_score","Communication score","record_id",child_pedsql[child_pedsql$comm_score>0,])
```

## Child Total Score

```{r results='asis'}
model_and_plot_no12("pedsql_total_score","Total score","record_id",child_pedsql[child_pedsql$pedsql_total_score>0,])
```

## Parent Diabetes

```{r results='asis'}
model_and_plot_no12("diabetes_score","Diabetes score","record_id",parent_pedsql[parent_pedsql$diabetes_score>0,])
```

## Parent Treatment I

```{r results='asis'}
model_and_plot_no12("tx1_score","Treatment I score","record_id",parent_pedsql[parent_pedsql$tx1_score>0,])
```

## Parent Treatment II

```{r results='asis'}
model_and_plot_no12("tx2_score","Treatment II score","record_id",parent_pedsql[parent_pedsql$tx2_score>0,])
```

## Parent Worry

```{r results='asis'}
model_and_plot_no12("worry_score","Worry score","record_id",parent_pedsql[parent_pedsql$worry_score>0,])
```

## Parent Communication

```{r results='asis'}
model_and_plot_no12("comm_score","Communication score","record_id",parent_pedsql[parent_pedsql$comm_score>0,])
```

## Parent Total Score

```{r results='asis'}
model_and_plot_no12("pedsql_total_score","Total score","record_id",parent_pedsql[parent_pedsql$pedsql_total_score>0,])
```

# Association between Peds-QL Diabetes Module Scores and CEFIS scores

How to interpret these models: in the table labeled "Fixed Effects," the estimate for the CEFIS measure is the change in the Peds-QL measure for a one unit change in the CEFIS measure.

```{r}
# Looks like children did not complete CEFIS
#child_pedsql_cefis <- merge(child_pedsql,scores,by=c("record_id","Timepoint"),all.x = T, all.y = F)
parent_pedsql_cefis <- merge(parent_pedsql,scores,by=c("record_id","Timepoint"),all.x = T, all.y = F)
parent_pedsql_cefis$num_time <- parent_pedsql_cefis$num_time.x
parent_pedsql_cefis$num_id <- parent_pedsql_cefis$num_id.x

# Link child PedsQL total score and parent CEFIS
childtot <- child_pedsql[,c("num_id","Timepoint","pedsql_total_score")]
colnames(childtot) <- c("num_id","Timepoint","pedsql_total_score_child")
childtot$num_id <- childtot$num_id-1000
parent_pedsql_cefis_child <- merge(parent_pedsql_cefis,childtot,by=c("num_id","Timepoint"),all.x = T,all.y = F)
```

## Exposure and Parent Total Score

```{r results='asis'}
tvc("pedsql_total_score","exposure_sum",parent_pedsql_cefis[parent_pedsql_cefis$exposure_sum>0,])
```

## Impact and Parent Total Score

```{r results='asis'}
tvc("pedsql_total_score","impact_mean",parent_pedsql_cefis[parent_pedsql_cefis$exposure_sum>0,])
```

## Distress and Parent Total Score

```{r results='asis'}
tvc("pedsql_total_score","distress",parent_pedsql_cefis[parent_pedsql_cefis$exposure_sum>0,])
```

## Exposure and Child Total Score

```{r results='asis'}
tvc("pedsql_total_score_child","exposure_sum",parent_pedsql_cefis_child[parent_pedsql_cefis_child$exposure_sum>0,])
```

## Impact and Child Total Score

```{r results='asis'}
tvc("pedsql_total_score_child","impact_mean",parent_pedsql_cefis_child[parent_pedsql_cefis_child$exposure_sum>0,])
```

## Distress and Child Total Score

```{r results='asis'}
tvc("pedsql_total_score_child","distress",parent_pedsql_cefis_child[parent_pedsql_cefis_child$exposure_sum>0,])
```

```{r api import,include=FALSE}
token = read.table("./Erin Cobry/ADA 2022/HCL and Sleep/api_token.txt")[1,1]
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",token = token)
data = exportRecords(rcon,dates = F)
ids = data$record_id
times = data$redcap_event_name
exclude = c(101,201,102,202,103,203,110,210,119,219,121,221,131,231,141,241)
```

```{r model functions}
fit_mod = function(outcome,df,plot = T,diagnostics = F){
  # Format timepoint
  df$timepoint = as.factor(df$timepoint)
  levels(df$timepoint) = gsub("baseline_.*","Baseline",levels(df$timepoint))
  levels(df$timepoint) = gsub("hcl_month_3_.*","Month 3",levels(df$timepoint))
  levels(df$timepoint) = gsub("hcl_month_6_.*","Month 6",levels(df$timepoint))
  # Fit random intercept model
  f = as.formula(paste0(outcome,"~timepoint+(1|record_id)"))
  mod = lmer(f,data = df)
  # Plot
  if(plot){
    p = 
      ggplot(data = df,aes_string(x = "timepoint",y = outcome,
                                  color = "record_id",group = "record_id")) + 
      geom_point() + geom_line() + xlab("Timepoint") +
      theme_bw() + theme(legend.position = "none")
    print(p)
  }
  # Type 3
  type3 = data.frame(car::Anova(mod,type = 3))
  colnames(type3)[3] = "p.value"
  print(kable(type3,digits = 3,caption = "Test of Overall Effect"))
  # Model results
  res = tidy(mod,"fixed")
  res$term = sub("timepoint","",res$term)
  res$effect = NULL
  print(kable(res,digits = 3,caption = "Comparison to Baseline"))
  # Means
  means = data.frame(emmeans(mod,~timepoint))
  print(kable(means,digits = 3,caption = "Model Means"))
  # Number of obs
  cat("\n")
  cat(paste("This model was fit using",nobs(mod),"observations."))
  cat("\n")
  # Check model
  if(diagnostics){
    print(check_model(mod))
  }
}
```

# PROMIS

## PROMIS: Sleep Disturbance

### Children

```{r}
#source("./Erin Cobry/ADA 2022/COVID/Data raw/HealthRelatedQuality_R_2022-01-01_2151.r")
#promis_peds$timepoint = promis_peds$redcap_event_name

# had been using Tim's code from the HCL and sleep study but this is a different form

```

```{r results='asis'}
#fit_mod(outcome = "disturbance",df = promis_peds)
```

### Parents

```{r}
# Import select, filter, etc.
```

```{r results='asis'}
#fit_mod(outcome = "disturbance",df = promis_parent)
```

## PROMIS: Impairment

### Children

```{r results='asis'}
#fit_mod(outcome = "impairment",df = promis_peds)
```

### Parents

```{r results='asis'}
#fit_mod(outcome = "impairment",df = promis_parent)
```