---
title: "Retinopathy Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Sarit Polsky/Retinopathy/Data_Raw/PregnancyCGM_R_2019-08-12_1528.r")
library(tableone)
library(skimr)
library(knitr)
library(tidyverse)
```

```{r echo=FALSE}
# Get retinopathy data, wide to extra long
retino <- data %>% filter(redcap_event_name == "retinopathy_arm_1") %>% 
  select(subject_id,retinop_precon_right.factor:retinop_pp_left.factor) %>%
  gather(.,visit,value,retinop_precon_right.factor:retinop_pp_left.factor) %>%
  mutate(eye = str_replace_all(visit,".*_|.factor",""))
# Get visit number, retinopathy and visit as factors
retino$visit <- sapply(strsplit(retino$visit,"_"), `[`,2)
retino$visit <- factor(retino$visit,levels = c("precon","1st","2nd","3rd","pp"))
retino$value <- factor(retino$value,
                       levels = c("None","Mild nonproliferative diabetic retinopathy",
                          "Mild-moderate nonproliferative diabetic retinopathy",
                          "Moderate nonproliferative diabetic retinopathy",
                          "Moderate-severe nonproliferative diabetic retinopathy",
                          "Severe nonproliferative diabetic retinopathy",
                          "Proliferative retinopathy"))
# Extra long to long
retino <- retino %>% spread(eye,value)
# Study group
group <- data %>% filter(redcap_event_name == "baseline_informati_arm_1") %>%
  select(subject_id,study_group.factor)
group$Group <- NA
group$Group[grep("alone",group$study_group.factor)] <- "Alone"
group$Group[grep("Share",group$study_group.factor)] <- "Share"
retino <- left_join(retino,group,by="subject_id") %>% 
  select(subject_id,visit,left,right,Group)
# Colnames and formatting
colnames(retino) <- c("subject_id","Visit","Left Eye","Right Eye","Group")
retino$Visit <- as.factor(retino$Visit)
```

# Retinopathy Status, CGM Alone vs. CGM Share

The number and percentage of women in each group with each grade of retinopathy, split by time point. Missing data is given as a percentage throughout the report.

## Table 1: Pre-Conception Visit

```{r echo=FALSE}
table1 <- CreateTableOne(c("Left Eye","Right Eye"),strata = "Group",data = retino[retino$Visit == "precon",])
table1 <- print(table1,exact=c("Left Eye","Right Eye"),printToggle = F,missing = T)
kable(table1)
```

## Table 2: 1st Trimester Visit

```{r echo=FALSE}
table2 <- CreateTableOne(c("Left Eye","Right Eye"),strata = "Group",data = retino[retino$Visit == "1st",])
table2 <- print(table2,exact=c("Left Eye","Right Eye"),printToggle = F,missing = T)
kable(table2)
```

## Table 3: 2nd Trimester Visit

```{r echo=FALSE}
table3 <- CreateTableOne(c("Left Eye","Right Eye"),strata = "Group",data = retino[retino$Visit == "2nd",])
table3 <- print(table3,exact=c("Left Eye","Right Eye"),printToggle = F,missing = T)
kable(table3)
```

## Table 4: 3rd Trimester Visit

```{r echo=FALSE}
table4 <- CreateTableOne(c("Left Eye","Right Eye"),strata = "Group",data = retino[retino$Visit == "3rd",])
table4 <- print(table4,exact=c("Left Eye","Right Eye"),printToggle = F,missing = T)
kable(table4)
```

## Table 5: Post-Partum Visit

```{r echo=FALSE}
table5 <- CreateTableOne(c("Left Eye","Right Eye"),strata = "Group",data = retino[retino$Visit == "pp",])
table5 <- print(table5,exact=c("Left Eye","Right Eye"),printToggle = F,missing = T)
kable(table5)
```

There were no differences in retinopathy grades between the two groups at any time point.

# Retinopathy Progression, CGM Alone vs. CGM Share

A participant was said to have progressed if their most severe retinopathy grade during pregnancy was worse than their grade at the pre-conception visit. Regression was similarly defined as a post-delivery retinopathy grade less severe than the worst grade during pregnancy. 

Participants without pre-conception retinopathy data were counted as missing for the progression numbers, and those without post-delivery data were treated as missing for regression.

```{r echo=FALSE,warning=FALSE}
progression <- retino
progression$`Left Eye` <- as.numeric(progression$`Left Eye`)
progression$`Right Eye` <- as.numeric(progression$`Right Eye`)
# Find worst diagnoses during pregnancy
progression <- progression %>% group_by(subject_id) %>%
  summarise(precon_left = `Left Eye`[1],
            preg_max_left = max(`Left Eye`[2:4],na.rm=T),
            postdel_left = `Left Eye`[5],
            precon_right = `Right Eye`[1],
            preg_max_right = max(`Right Eye`[2:4],na.rm=T),
            postdel_right = `Right Eye`[5],
            Group = Group[1])
# Remove -Inf (from those who had no measures during pregnancy)
progression <- as.data.frame(progression)
progression[!is.na(progression) & progression == -Inf] <- NA
# Determine if they progressed during pregnancy/resolved post-delivery
# Left
progression$`Left Eye Progressed` <- ifelse(progression$preg_max_left > progression$precon_left,T,F)
progression$`Left Eye Regressed` <- ifelse(progression$preg_max_left > progression$postdel_left,T,F)
# Right
progression$`Right Eye Progressed` <- ifelse(progression$preg_max_right > progression$precon_right,T,F)
progression$`Right Eye Regressed` <- ifelse(progression$preg_max_right > progression$postdel_right,T,F)
```

## Table 6: Eyes Affected

```{r echo=FALSE}
vars = c("Left Eye Progressed","Left Eye Regressed","Right Eye Progressed","Right Eye Regressed")
t6 <- CreateTableOne(vars,strata = "Group",data = progression)
t6 <- print(t6,exact = vars,missing = T,printToggle = F,showAllLevels = T)
kable(t6)
```

There were no differences in progression or regression in either eye between the two groups.

## Table 7: Individuals Affected

```{r echo=FALSE,message=FALSE}
# Progression in both eyes, one eye, or neither eye.
# Unite eye columns
progression <- progression %>% 
  unite(Progressed,`Left Eye Progressed`,`Right Eye Progressed`,remove = F) %>%
  unite(Regressed,`Left Eye Regressed`,`Right Eye Regressed`,remove = F)
# Change values
progression$Progressed <- 
  as.factor(plyr::mapvalues(progression$Progressed,
                            from = c("FALSE_FALSE","FALSE_NA","FALSE_TRUE","NA_NA","TRUE_FALSE","TRUE_TRUE"),
                            to = c("Neither Eye","Neither Eye","One Eye",NA,"One Eye","Both Eyes")))
progression$Regressed <- 
  as.factor(plyr::mapvalues(progression$Regressed,
                            from = c("FALSE_FALSE","FALSE_NA","FALSE_TRUE","NA_NA","TRUE_FALSE","TRUE_TRUE"),
                            to = c("Neither Eye","Neither Eye","One Eye",NA,"One Eye","Both Eyes")))
# Table
t7 <- CreateTableOne(vars = c("Progressed","Regressed"),strata = "Group",data = progression)
t7 <- print(t7,missing = T,printToggle = F,showAllLevels = T,exact = c("Progressed","Regressed"))
kable(t7)
```

There were no differences in the number of individuals who progressed or regressed between the two groups.

Participant 13 did not have right eye data during pregnancy, but her left eye did not change, so she was counted as "Neither Eye" for both progression and regression.

None of the participants regressed in both eyes.