---
title: "Analysis of AHA omics data"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(arsenal)
library(skimr)
library(knitr)
library(mixOmics)
library(tidyverse)
library(limma)
library(webchem)
library(omu)
library(PCAtools)
library(openxlsx)
library(dplyr)
library(multtest)

home_dir = ifelse(.Platform$OS.type != "unix","T:\\Janet Snell-Bergeon\\AHA collaborative grant",
                  "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant")
knitr::opts_knit$set(root.dir = home_dir)

# Basic workflow:
# log transform prior to analyses
# Moderated t-tests
# Linear model in limma containing interaction with sex
# sPLS-DA
# No multiple correction
```

```{r read AHA data,echo=FALSE}
load(file="T:/Janet Snell-Bergeon/AHA collaborative grant/aha_master_data.Rdata")

# log-transform
# NOTE: per Qibin, lipids are already log transformed
df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins)] = 
  apply(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins)], 2, log)

# delete records with missing CACp
df <- df[!is.na(df$CACp),]

# remove low variance items
no_var = caret::nearZeroVar(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)])
df = df[,-c(no_var)]
# removed 23 columns, leaves 1816 columns

# remove the compounds with missing data and low variance from the strings of names
untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
global_proteins <- global_proteins[global_proteins %in% colnames(df)]
glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
lipids <- lipids[lipids %in% colnames(df)]

# remove markers with >=80% missing - 61 columns
miss <- df[,which(colMeans(is.na(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]))>=0.8)]
miss <- miss[,colnames(miss) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]
df <- df %>% select(-one_of(colnames(miss)))
# left with 1755 columns

# remove the compounds with missing data and low variance from the strings of names
untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
global_proteins <- global_proteins[global_proteins %in% colnames(df)]
glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
lipids <- lipids[lipids %in% colnames(df)]
```

```{r moderated ttests,echo=FALSE}
cac_contrast <- ifelse(df$CACp=="Yes",1,0)
cac_contrast <- cbind(rep(1,nrow(df)),cac_contrast)

# untargeted metabolites
ymat <- t(df[,names(df) %in% untargeted_metabs])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_untargeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_untargeted_metabs <- results_untargeted_metabs[order(results_untargeted_metabs$adj.P.Val),] 
# trying permutation test based adjustment
resT_untargeted_metabs<-mt.maxT(ymat,cac_contrast[,2])
resP_untargeted_metabs<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# targeted metabolites
ymat <- t(df[,names(df) %in% targeted_metabs])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_targeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_targeted_metabs <- results_targeted_metabs[order(results_targeted_metabs$adj.P.Val),] 
# trying permutation test based adjustment
#resT_targeted_metabs<-mt.maxT(ymat,cac_contrast[,2])
#resP_targeted_metabs<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# global proteomics
ymat <- t(df[,names(df) %in% global_proteins])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_global_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_global_proteins <- results_global_proteins[order(results_global_proteins$adj.P.Val),] 
# trying permutation test based adjustment
resT_global_proteins<-mt.maxT(ymat,cac_contrast[,2])
resP_global_proteins<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# glycated proteomics
ymat <- t(df[,names(df) %in% glycated_proteins])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_glycated_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_glycated_proteins <- results_glycated_proteins[order(results_glycated_proteins$adj.P.Val),] 
# trying permutation test based adjustment
resT_glycated_proteins<-mt.maxT(ymat,cac_contrast[,2])
resP_glycated_proteins<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# lipidomics
ymat <- t(df[,names(df) %in% lipids])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_lipids <- topTable(fit,coef = 2,number = nrow(ymat))
results_lipids <- results_lipids[order(results_lipids$adj.P.Val),] 
# trying permutation test based adjustment
resT_lipids<-mt.maxT(ymat,cac_contrast[,2])
resP_lipids<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# write to file
wb <- createWorkbook()
addWorksheet(wb,"untargeted_metab_moderated_FDR")
writeData(wb,"untargeted_metab_moderated_FDR",results_untargeted_metabs,rowNames = T)
addWorksheet(wb,"untargeted_metab_maxT")
writeData(wb,"untargeted_metab_maxT",resT_untargeted_metabs,rowNames = T)
addWorksheet(wb,"untargeted_metab_minP")
writeData(wb,"untargeted_metab_minP",resP_untargeted_metabs,rowNames = T)
addWorksheet(wb,"targeted_metab_moderated_FDR")
writeData(wb,"targeted_metab_moderated_FDR",results_targeted_metabs,rowNames = T)
#addWorksheet(wb,"targeted_metabolomics_maxT")
#writeData(wb,"resT_targeted_metabolomics",results_untargeted_metabs,rowNames = T)
#addWorksheet(wb,"targeted_metabolomics_minP")
#writeData(wb,"resP_targeted_metabolomics",results_untargeted_metabs,rowNames = T)
addWorksheet(wb,"global_prot_moderated_FDR")
writeData(wb,"global_prot_moderated_FDR",results_global_proteins,rowNames = T)
addWorksheet(wb,"global_prot_maxT")
writeData(wb,"global_prot_maxT",resT_global_proteins,rowNames = T)
addWorksheet(wb,"global_prot_minP")
writeData(wb,"global_prot_minP",resP_global_proteins,rowNames = T)
addWorksheet(wb,"glycated_prot_moderated_FDR")
writeData(wb,"glycated_prot_moderated_FDR",results_glycated_proteins,rowNames = T)
addWorksheet(wb,"glycated_prot_maxT")
writeData(wb,"glycated_prot_maxT",resT_glycated_proteins,rowNames = T)
addWorksheet(wb,"glycated_prot_minP")
writeData(wb,"glycated_prot_minP",resP_glycated_proteins,rowNames = T)
addWorksheet(wb,"lipidomics_moderated_FDR")
writeData(wb,"lipidomics_moderated_FDR",results_lipids,rowNames = T)
addWorksheet(wb,"lipidomics_maxT")
writeData(wb,"lipidomics_maxT",resT_lipids,rowNames = T)
addWorksheet(wb,"lipidomics_minP")
writeData(wb,"lipidomics_minP",resP_lipids,rowNames = T)
saveWorkbook(wb,"AHA_omics_ttests.xlsx",overwrite = TRUE)

```

```{r sex interaction,echo=FALSE}
df$sexnum <- ifelse(df$gender=="M",0,1)
df$CACp_num <- ifelse(df$CACp=="Yes",1,0)
int_contrast <- model.matrix(~CACp_num*sexnum,df)

# untargeted metabolites
ymat <- t(df[,names(df) %in% untargeted_metabs])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_untargeted_metabs_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_untargeted_metabs_int <- results_untargeted_metabs_int[order(results_untargeted_metabs_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# targeted metabolites
ymat <- t(df[,names(df) %in% targeted_metabs])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_targeted_metabs_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_targeted_metabs_int <- results_targeted_metabs_int[order(results_targeted_metabs_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# global proteomics
ymat <- t(df[,names(df) %in% global_proteins])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_global_proteins_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_global_proteins_int <- results_global_proteins_int[order(results_global_proteins_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# glycated proteomics
ymat <- t(df[,names(df) %in% glycated_proteins])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_glycated_proteins_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_glycated_proteins_int <- results_glycated_proteins_int[order(results_glycated_proteins_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# lipidomics
ymat <- t(df[,names(df) %in% lipids])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_lipids_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_lipids_int <- results_lipids_int[order(results_lipids_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# write to file
wb_int <- createWorkbook()
addWorksheet(wb_int,"untargeted_metabolomics")
writeData(wb_int,"untargeted_metabolomics",results_untargeted_metabs_int,rowNames = T)
addWorksheet(wb_int,"targeted_metabolomics")
writeData(wb_int,"targeted_metabolomics",results_targeted_metabs_int,rowNames = T)
addWorksheet(wb_int,"global_proteomics")
writeData(wb_int,"global_proteomics",results_global_proteins_int,rowNames = T)
addWorksheet(wb_int,"glycated_proteins")
writeData(wb_int,"glycated_proteins",results_glycated_proteins_int,rowNames = T)
addWorksheet(wb_int,"lipidomics")
writeData(wb_int,"lipidomics",results_lipids_int,rowNames = T)
saveWorkbook(wb_int,"AHA_omics_sex_interaction.xlsx",overwrite = TRUE)
```

# Background

The purpose of this analysis is to examine differences in targeted and untargeted metabolomics, global and glycated proteomics, and lipidomics in CACTI participants with and without CAC progression.

# Methods

what needs to be done:
selectvar
permutation tests on lambda

Data were log transformed prior to analysis with the exception of the lipid data, which had previously been log transformed.  Markers with >=80% missing data or near-zero variance were removed prior to analysis.

Moderated t-tests were used to compare CAC progression groups.  Linear models were used to test whether there was a significant interaction between CAC progression groups and sex.

Sparse PLS-DA was used to identify markers that best discriminated between the CAC progression groups.  Each sPLS-DA model was tuned to identify the number of components and the number of proteins that resulted in the lowest error rate.  Permutation testing with 1000 permutations was used to evaluate the extent of overfitting.

# Results

For moderated t-tests and linear models with interaction between CACp and sex, see separate files

```{r splsda utm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# untargeted metabolites
Y <- df$CACp
X <- df[,names(df) %in% untargeted_metabs]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA of untargeted metabolites by CACp",
         pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
# plsda.perf = perf(splsda.utm, validation = "loo",auc=TRUE)
# auc_save <- plsda.perf$auc$`comp1`[1]    
# auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% untargeted_metabs],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda tm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# targeted metabolites
Y <- df$CACp
X <- df[,names(df) %in% targeted_metabs]
#tune.splsda <- tune.splsda(X, Y, ncomp = 2) 
#choice.keepX <- tune.splsda$choice.keepX[1:2]
choice.keepX <- 10
## sPLS-DA function
splsda.tm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.tm,title = "sPLS-DA of targeted metabolites by CACp",
         pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.tm,comp=1)
top_2 <- selectVar(splsda.tm,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
# plsda.perf = perf(splsda.tm, validation = "loo",auc=TRUE)
# auc_save <- plsda.perf$auc$`comp1`[1]    
# auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% targeted_metabs],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda gp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# global proteins
Y <- df$CACp
X <- df[,names(df) %in% global_proteins]
#tune.splsda <- tune.splsda(X, Y, ncomp =  2) 
#choice.keepX <- tune.splsda$choice.keepX[1:2]
choice.keepX <- 10
## sPLS-DA function
splsda.globalp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.globalp,title = "sPLS-DA of global proteins by CACp",
         pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.globalp,comp=1)
top_2 <- selectVar(splsda.globalp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
# plsda.perf = perf(splsda.globalp, validation = "loo",auc=TRUE)
# auc_save <- plsda.perf$auc$`comp1`[1]    
# auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% global_proteins],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda glyp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# glycated proteins
Y <- df$CACp
X <- df[,names(df) %in% glycated_proteins]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.glycp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.glycp,title = "sPLS-DA of glycated proteins by CACp",
         pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.glycp,comp=1)
top_2 <- selectVar(splsda.glycp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
# plsda.perf = perf(splsda.glycp, validation = "loo",auc=TRUE)
# auc_save <- plsda.perf$auc$`comp1`[1]    
# auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% glycated_proteins],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda lipids,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# lipids
Y <- df$CACp
X <- df[,names(df) %in% lipids]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.lipids <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.lipids,title = "sPLS-DA of lipids by CACp",
         pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.lipids,comp=1)
top_2 <- selectVar(splsda.lipids,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
# plsda.perf = perf(splsda.lipids, validation = "loo",auc=TRUE)
# auc_save <- plsda.perf$auc$`comp1`[1]    
# auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% lipids],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda all,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# everything
Y <- df$CACp
X <- df[,names(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]
#tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
#choice.keepX <- tune.splsda$choice.keepX[1:2]
choice.keepX <- 10
## sPLS-DA function
splsda.all <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.all,title = "sPLS-DA of all markers by CACp",
           pch = 20,ind.names = F,ellipse = T,legend = T)
 
top_1 <- selectVar(splsda.all,comp=1)
top_2 <- selectVar(splsda.all,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
# plsda.perf = perf(splsda.all, validation = "loo",auc=TRUE)
# auc_save <- plsda.perf$auc$`comp1`[1]    
# auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```