---
title: "Analysis of AHA omics data"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(arsenal)
library(skimr)
library(knitr)
library(mixOmics)
library(tidyverse)
library(limma)
library(webchem)
library(omu)
library(PCAtools)
library(openxlsx)

home_dir = ifelse(.Platform$OS.type != "unix","T:/Janet Snell-Bergeon/AHA collaborative grant",
                  "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant")
knitr::opts_knit$set(root.dir = home_dir)

# Basic workflow:
# log transform prior to analyses
# Moderated t-tests
# Linear model in limma containing interaction with sex
# sPLS-DA
# No multiple correction
```

```{r read AHA data,echo=FALSE}
load(file="T:/Janet Snell-Bergeon/AHA collaborative grant/aha_master_data.Rdata")

# log-transform
df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)] = 
  apply(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)], 2, log)

# delete records with missing CACp
df <- df[!is.na(df$CACp),]

# remove low variance items
no_var = caret::nearZeroVar(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)])
df = df[,-c(no_var)]

# remove the compounds with missing data and low variance from the strings of names
untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
global_proteins <- global_proteins[global_proteins %in% colnames(df)]
glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
lipids <- lipids[lipids %in% colnames(df)]

# remove markers with >=80% missing
df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)] = 
  df[,-c(which(colMeans(is.na(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]))>=0.8))]

# remove the compounds with missing data and low variance from the strings of names
untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
global_proteins <- global_proteins[global_proteins %in% colnames(df)]
glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
lipids <- lipids[lipids %in% colnames(df)]
```

```{r moderated ttests,echo=FALSE}
cac_contrast <- ifelse(df$CACp=="Yes",1,0)
cac_contrast <- cbind(rep(1,nrow(df)),cac_contrast)

# untargeted metabolites
a <- t(df)
b <- a[,names(a) %in% untargeted_metabs]
ymat <- t(df[,names(df) %in% untargeted_metabs])
temp <- as.data.frame(ymat)
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_untargeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_untargeted_metabs <- results_untargeted_metabs[order(results_untargeted_metabs$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# targeted metabolites
ymat <- t(df[,names(df) %in% targeted_metabs])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_targeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_targeted_metabs <- results_targeted_metabs[order(results_targeted_metabs$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# global proteomics
ymat <- t(df[,names(df) %in% global_proteins])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_global_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_global_proteins <- results_global_proteins[order(results_global_proteins$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# glycated proteomics
ymat <- t(df[,names(df) %in% glycated_proteins])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_glycated_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_glycated_proteins <- results_glycated_proteins[order(results_glycated_proteins$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# lipidomics
ymat <- t(df[,names(df) %in% lipids])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_lipids <- topTable(fit,coef = 2,number = nrow(ymat))
results_lipids <- results_lipids[order(results_lipids$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# write to file
wb <- createWorkbook()
addWorksheet(wb,"untargeted_metabolomics")
writeData(wb,"untargeted_metabolomics",results_untargeted_metabs,rowNames = T)
addWorksheet(wb,"targeted_metabolomics")
writeData(wb,"targeted_metabolomics",results_targeted_metabs,rowNames = T)
addWorksheet(wb,"global_proteomics")
writeData(wb,"global_proteomics",results_global_proteins,rowNames = T)
addWorksheet(wb,"glycated_proteins")
writeData(wb,"glycated_proteins",results_glycated_proteins,rowNames = T)
addWorksheet(wb,"lipidomics")
writeData(wb,"lipidomics",results_lipids,rowNames = T)
saveWorkbook(wb,"AHA_omics_ttests.xlsx",overwrite = TRUE)

```

```{r sex interaction,echo=FALSE}
df$sexnum <- ifelse(df$gender=="M",0,1)
df$CACp_num <- ifelse(df$CACp=="Yes",1,0)
int_contrast <- model.matrix(~CACp_num*sexnum,df)

# untargeted metabolites
ymat <- t(df[,names(df) %in% untargeted_metabs])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_untargeted_metabs_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_untargeted_metabs_int <- results_untargeted_metabs_int[order(results_untargeted_metabs_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# targeted metabolites
ymat <- t(df[,names(df) %in% targeted_metabs])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_targeted_metabs_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_targeted_metabs_int <- results_targeted_metabs_int[order(results_targeted_metabs_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# global proteomics
ymat <- t(df[,names(df) %in% global_proteins])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_global_proteins_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_global_proteins_int <- results_global_proteins_int[order(results_global_proteins_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# glycated proteomics
ymat <- t(df[,names(df) %in% glycated_proteins])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_glycated_proteins_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_glycated_proteins_int <- results_glycated_proteins_int[order(results_glycated_proteins_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# lipidomics
ymat <- t(df[,names(df) %in% lipids])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_lipids_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_lipids_int <- results_lipids_int[order(results_lipids_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# write to file
wb_int <- createWorkbook()
addWorksheet(wb_int,"untargeted_metabolomics")
writeData(wb_int,"untargeted_metabolomics",results_untargeted_metabs_int,rowNames = T)
addWorksheet(wb_int,"targeted_metabolomics")
writeData(wb_int,"targeted_metabolomics",results_targeted_metabs_int,rowNames = T)
addWorksheet(wb_int,"global_proteomics")
writeData(wb_int,"global_proteomics",results_global_proteins_int,rowNames = T)
addWorksheet(wb_int,"glycated_proteins")
writeData(wb_int,"glycated_proteins",results_glycated_proteins_int,rowNames = T)
addWorksheet(wb_int,"lipidomics")
writeData(wb_int,"lipidomics",results_lipids_int,rowNames = T)
saveWorkbook(wb_int,"AHA_omics_sex_interaction.xlsx",overwrite = TRUE)
```

# Background

The purpose of this analysis is to examine differences in targeted and untargeted metabolomics, global and glycated proteomics, and lipidomics in CACTI participants with and without CAC progression.

# Methods

Data were log transformed prior to analysis.  Markers with >=80% missing data or near-zero variance were removed prior to analysis.

Moderated t-tests were used to compare CAC progression groups.  Linear models were used to test whether there was a significant interaction between CAC progression groups and sex.

Sparse PLS-DA was used to identify markers that best discriminated between the CAC progression groups.  Each sPLS-DA model was tuned to identify the number of components and the number of proteins that resulted in the lowest error rate.  Permutation testing with 1000 permutations was used to evaluate the extent of overfitting.

# Results

For moderated t-tests and linear models with interaction between CACp and sex, see separate files

```{r splsda utm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE}
# untargeted metabolites
Y <- df$CACp
X <- df[,names(df) %in% untargeted_metabs]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA of untargeted metabolites by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

# something really strange is going on with the loadings - what prints in the console does not match the html
top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
kable(top_1d)
kable(top_2d)

set.seed(3654)
plsda.perf = perf(splsda.utm, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
aucs <- vector(mode="numeric", length=n_perm)
for (i in 1:n_perm) {
 Y <- sample(as.factor(df$CACp),replace = F)
 plsda_res <- plsda(df[,names(df) %in% untargeted_metabs],Y,ncomp = 2)
 perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
                  validation = "loo")
 aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
}
ggplot(as.data.frame(aucs),aes(x=aucs)) +
geom_histogram(binwidth = 0.01) +
geom_vline(aes(xintercept=auc_true),color="red") +
theme_bw() + xlab("AUC")

```

```{r splsda tm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE}
# targeted metabolites
Y <- df$CACp
X <- df[,names(df) %in% targeted_metabs]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.tm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.tm,title = "sPLS-DA of targeted metabolites by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

# something really strange is going on with the loadings - what prints in the console does not match the html
top_1 <- selectVar(splsda.tm,comp=1)
top_2 <- selectVar(splsda.tm,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
kable(top_1d)
kable(top_2d)

set.seed(3654)
plsda.perf = perf(splsda.tm, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% targeted_metabs],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda gp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE}
# global proteins
Y <- df$CACp
X <- df[,names(df) %in% global_proteins]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.globalp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.globalp,title = "sPLS-DA of global proteins by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

# something really strange is going on with the loadings - what prints in the console does not match the html
top_1 <- selectVar(splsda.globalp,comp=1)
top_2 <- selectVar(splsda.globalp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
kable(top_1d)
kable(top_2d)

set.seed(3654)
plsda.perf = perf(splsda.globalp, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% global_proteins],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda glyp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE}
# glycated proteins
Y <- df$CACp
X <- df[,names(df) %in% glycated_proteins]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.glycp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.glycp,title = "sPLS-DA of glycated proteins by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

# something really strange is going on with the loadings - what prints in the console does not match the html
top_1 <- selectVar(splsda.glycp,comp=1)
top_2 <- selectVar(splsda.glycp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
kable(top_1d)
kable(top_2d)

set.seed(3654)
plsda.perf = perf(splsda.glycp, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% glycated_proteins],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda lipids,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE}
# lipids
Y <- df$CACp
X <- df[,names(df) %in% lipids]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.lipids <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.lipids,title = "sPLS-DA of lipids by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

# something really strange is going on with the loadings - what prints in the console does not match the html
top_1 <- selectVar(splsda.lipids,comp=1)
top_2 <- selectVar(splsda.lipids,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
kable(top_1d)
kable(top_2d)

set.seed(3654)
plsda.perf = perf(splsda.lipids, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% lipids],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```

```{r splsda all,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE}
# everything
Y <- df$CACp
X <- df[,names(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.all <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.all,title = "sPLS-DA of all markers by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

# something really strange is going on with the loadings - what prints in the console does not match the html
top_1 <- selectVar(splsda.all,comp=1)
top_2 <- selectVar(splsda.all,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
kable(top_1d)
kable(top_2d)

set.seed(3654)
plsda.perf = perf(splsda.all, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
# n_perm <- 1000
# aucs <- vector(mode="numeric", length=n_perm)
# for (i in 1:n_perm) {
#  Y <- sample(as.factor(df$CACp),replace = F)
#  plsda_res <- plsda(df[,names(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)],Y,ncomp = 2)
#  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
#                   validation = "loo")
#  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
# }
# ggplot(as.data.frame(aucs),aes(x=aucs)) +
# geom_histogram(binwidth = 0.01) +
# geom_vline(aes(xintercept=auc_true),color="red") +
# theme_bw() + xlab("AUC")

```