---
title: "Analysis of AHA omics data"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(arsenal)
library(skimr)
library(knitr)
library(mixOmics)
library(tidyverse)
library(limma)
library(webchem)
library(omu)
library(PCAtools)
library(openxlsx)
library(dplyr)
library(multtest)
library(pls)
library(parallel)
home_dir = ifelse(.Platform$OS.type != "unix","/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant",
                  "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/")
#home_dir = ("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant")
#knitr::opts_knit$set(root.dir = home_dir)

#setwd("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant")

# Basic workflow:
# log transform prior to analyses
# Moderated t-tests
# Linear model in limma containing interaction with sex
# sPLS-DA
# No multiple correction
```

```{r read AHA data,echo=FALSE}
#load(file="T:/Janet Snell-Bergeon/AHA collaborative grant/aha_master_data.Rdata")
load(file="~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant//aha_master_data.Rdata")

# before log transform, convert 0 lineolic acid to a small number
df$linoleic.acid <- ifelse(df$linoleic.acid==0,0.001,df$linoleic.acid)

# log-transform
# NOTE: per Qibin, lipids are already log transformed
df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins)] = 
  apply(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins)], 2, log)

# delete records with missing eIS
df <- df[!is.na(df$eISV1),]

# remove low variance items
# no_var = caret::nearZeroVar(df[,colnames(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)])
# df = df[,-c(no_var)]

# remove the compounds with missing data and low variance from the strings of names
# untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
# targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
# global_proteins <- global_proteins[global_proteins %in% colnames(df)]
# glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
# lipids <- lipids[lipids %in% colnames(df)]

# remove markers with >=80% missing 
# miss <- df[,colnames(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)] %>% dplyr::select(where(~mean(is.na(.)) > 0.8))
# don't want to select clinical variables with >80% missing
# miss <- miss[,colnames(miss) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]
# df <- df %>% dplyr::select(-one_of(colnames(miss)))

# remove the compounds with missing data and low variance from the strings of names
# untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
# targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
# global_proteins <- global_proteins[global_proteins %in% colnames(df)]
# glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
# lipids <- lipids[lipids %in% colnames(df)]

# calculate change in square-root transformed CAC volume
# df$delta_cac_v2 <- (df$C2 - df$C1)
# df$delta_cac_v3 <- (df$C3 - df$C1)
# df$delta_cac_v4 <- (df$C4 - df$C1)

# create scaled version of df for pls-da of all measures
df_z <- df[,colnames(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]
df_z <- scale(df_z)

# create tertiles of insulin sensitivity
df$eISV1 <- as.numeric(df$eISV1)
df <- df %>% mutate(ins_median=ntile(df$eISV1,2))
```

```{r moderated ttests,echo=FALSE, message=FALSE, results="hide"}
# CAC progression
#ins_contrast <- ifelse(df$CACp=="Yes",1,0)
#ins_contrast <- cbind(rep(1,nrow(df)),ins_contrast)

# insulin tertiles
# code the contrast so that high insulin sensitivity=1 and low=0
ins_contrast <- ifelse(df$ins_median==2,1,0)
ins_contrast <- cbind(rep(1,nrow(df)),ins_contrast)

# untargeted metabolites
ymat <- df[,names(df) %in% untargeted_metabs]
ymat <- t(ymat)
fit <- lmFit(ymat,ins_contrast)
fit <- eBayes(fit)
results_untargeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_untargeted_metabs <- results_untargeted_metabs[order(results_untargeted_metabs$P.Value),] 
# trying permutation test based adjustment
resT_untargeted_metabs<-mt.maxT(ymat,ins_contrast[,2])
resP_untargeted_metabs<-mt.minP(ymat,ins_contrast[,2])
ymat <- NULL
fit <- NULL

# targeted metabolites
ymat <- t(df[,names(df) %in% targeted_metabs])
fit <- lmFit(ymat,ins_contrast)
fit <- eBayes(fit)
results_targeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_targeted_metabs <- results_targeted_metabs[order(results_targeted_metabs$P.Value),] 
# trying permutation test based adjustment
#resT_targeted_metabs<-mt.maxT(ymat,ins_contrast[,2])
#resP_targeted_metabs<-mt.minP(ymat,ins_contrast[,2])
ymat <- NULL
fit <- NULL

# global proteomics
ymat <- t(df[,names(df) %in% global_proteins])
fit <- lmFit(ymat,ins_contrast)
fit <- eBayes(fit)
results_global_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_global_proteins <- results_global_proteins[order(results_global_proteins$P.Value),] 
# trying permutation test based adjustment
resT_global_proteins<-mt.maxT(ymat,ins_contrast[,2])
resP_global_proteins<-mt.minP(ymat,ins_contrast[,2])
ymat <- NULL
fit <- NULL

# glycated proteomics
ymat <- t(df[,names(df) %in% glycated_proteins])
fit <- lmFit(ymat,ins_contrast)
fit <- eBayes(fit)
results_glycated_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_glycated_proteins <- results_glycated_proteins[order(results_glycated_proteins$P.Value),] 
# trying permutation test based adjustment
resT_glycated_proteins<-mt.maxT(ymat,ins_contrast[,2])
resP_glycated_proteins<-mt.minP(ymat,ins_contrast[,2])
ymat <- NULL
fit <- NULL

# lipidomics
ymat <- t(df[,names(df) %in% lipids])
fit <- lmFit(ymat,ins_contrast)
fit <- eBayes(fit)
results_lipids <- topTable(fit,coef = 2,number = nrow(ymat))
results_lipids <- results_lipids[order(results_lipids$P.Value),] 
# trying permutation test based adjustment
resT_lipids<-mt.maxT(ymat,ins_contrast[,2])
resP_lipids<-mt.minP(ymat,ins_contrast[,2])
ymat <- NULL
fit <- NULL

# write to file
wb <- createWorkbook()
addWorksheet(wb,"untargeted_metab_moderated_FDR")
writeData(wb,"untargeted_metab_moderated_FDR",results_untargeted_metabs,rowNames = T)
#addWorksheet(wb,"untargeted_metab_maxT")
#writeData(wb,"untargeted_metab_maxT",resT_untargeted_metabs,rowNames = T)
#addWorksheet(wb,"untargeted_metab_minP")
#writeData(wb,"untargeted_metab_minP",resP_untargeted_metabs,rowNames = T)
addWorksheet(wb,"targeted_metab_moderated_FDR")
writeData(wb,"targeted_metab_moderated_FDR",results_targeted_metabs,rowNames = T)
#addWorksheet(wb,"targeted_metabolomics_maxT")
#writeData(wb,"resT_targeted_metabolomics",results_untargeted_metabs,rowNames = T)
#addWorksheet(wb,"targeted_metabolomics_minP")
#writeData(wb,"resP_targeted_metabolomics",results_untargeted_metabs,rowNames = T)
addWorksheet(wb,"global_prot_moderated_FDR")
writeData(wb,"global_prot_moderated_FDR",results_global_proteins,rowNames = T)
#addWorksheet(wb,"global_prot_maxT")
#writeData(wb,"global_prot_maxT",resT_global_proteins,rowNames = T)
#addWorksheet(wb,"global_prot_minP")
#writeData(wb,"global_prot_minP",resP_global_proteins,rowNames = T)
addWorksheet(wb,"glycated_prot_moderated_FDR")
writeData(wb,"glycated_prot_moderated_FDR",results_glycated_proteins,rowNames = T)
#addWorksheet(wb,"glycated_prot_maxT")
#writeData(wb,"glycated_prot_maxT",resT_glycated_proteins,rowNames = T)
#addWorksheet(wb,"glycated_prot_minP")
#writeData(wb,"glycated_prot_minP",resP_glycated_proteins,rowNames = T)
addWorksheet(wb,"lipidomics_moderated_FDR")
writeData(wb,"lipidomics_moderated_FDR",results_lipids,rowNames = T)
#addWorksheet(wb,"lipidomics_maxT")
#writeData(wb,"lipidomics_maxT",resT_lipids,rowNames = T)
#addWorksheet(wb,"lipidomics_minP")
#writeData(wb,"lipidomics_minP",resP_lipids,rowNames = T)
#saveWorkbook(wb,"AHA_omics_ttests_baseline_eIS.xlsx",overwrite = TRUE)

saveWorkbook(wb,"~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/AHA_omics_ttests_baseline_eIS.xlsx",overwrite = TRUE)

```


# Background

The purpose of this analysis is to examine differences in targeted and untargeted metabolomics, global and glycated proteomics, and lipidomics in CACTI participants above and below the median of eIS at baseline.

# Methods

Data were log transformed prior to analysis with the exception of the lipid data, which had previously been log transformed.  Markers with >=80% missing data or near-zero variance were removed prior to analysis.

Moderated t-tests were used to compare CAC progression groups with FDR correction for multiple testing.  In addition, the maxT and minP permutation methods were used to correct for multiple testing.  

Linear models were used to test whether there was a significant interaction between CAC progression groups and sex.  The interpretation of the FC value for the interaction is the average additional FC for females with CAC progression.

Sparse PLS-DA was used to identify markers that best discriminated between the CAC progression groups.  Each sPLS-DA model was tuned to identify the number of components and the number of proteins that resulted in the lowest error rate.  Permutation testing with 1000 permutations was used to evaluate the extent of overfitting.

# Results

For moderated t-tests, see separate files.

## sPLS-DA of untargeted metabolites

```{r splsda utm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# untargeted metabolites
Y <- df$ins_median
X <- df[,names(df) %in% untargeted_metabs]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA of untargeted metabolites by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

```{r splsda tm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# targeted metabolites
Y <- df$ins_median
X <- df[,colnames(df) %in% targeted_metabs]
tune.splsda <- tune.splsda(X, Y, ncomp=5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
#choice.keepX <- 10
## sPLS-DA function
splsda.tm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.tm,title = "sPLS-DA of targeted metabolites by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.tm,comp=1)
top_2 <- selectVar(splsda.tm,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.tm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## sPLS-DA of global proteins

```{r splsda gp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# global proteins
Y <- df$ins_median
X <- df[,names(df) %in% global_proteins]
tune.splsda <- tune.splsda(X, Y, ncomp =  5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
#choice.keepX <- 10
## sPLS-DA function
splsda.globalp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.globalp,title = "sPLS-DA of global proteins by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.globalp,comp=1)
top_2 <- selectVar(splsda.globalp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.globalp, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## sPLS-DA of glycated proteins

```{r splsda glyp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# glycated proteins
Y <- df$ins_median
X <- df[,names(df) %in% glycated_proteins]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.glycp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.glycp,title = "sPLS-DA of glycated proteins by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.glycp,comp=1)
top_2 <- selectVar(splsda.glycp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.glycp, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```


## sPLS-DA of lipids

```{r splsda lipids,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# lipids
Y <- df$ins_median
X <- df[,names(df) %in% lipids]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.lipids <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.lipids,title = "sPLS-DA of lipids by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.lipids,comp=1)
top_2 <- selectVar(splsda.lipids,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.lipids, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```


## sPLS-DA of all data combined

```{r splsda all,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# everything
Y <- df$ins_median
X <- df_z
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
#choice.keepX <- 10
## sPLS-DA function
splsda.all <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.all,title = "sPLS-DA of all markers by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.all,comp=1)
top_2 <- selectVar(splsda.all,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
plsda.perf = perf(splsda.all, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

