---
title: "Analysis of AHA omics data"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(arsenal)
library(skimr)
library(knitr)
library(mixOmics)
library(tidyverse)
library(limma)
library(webchem)
library(omu)
library(PCAtools)
library(openxlsx)
library(dplyr)
library(multtest)
library(pls)
library(parallel)
home_dir = ifelse(.Platform$OS.type != "unix","Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant",
                  "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/")
knitr::opts_knit$set(root.dir = home_dir)

# Basic workflow:
# log transform prior to analyses
# Moderated t-tests
# Linear model in limma containing interaction with sex
# sPLS-DA
# No multiple correction
```

```{r read AHA data,echo=FALSE}
#load(file="T:/Janet Snell-Bergeon/AHA collaborative grant/aha_master_data.Rdata")
load(file="./aha_master_data.Rdata")

# before log transform, convert 0 lineolic acid to a small number
df$linoleic.acid <- ifelse(df$linoleic.acid==0,0.001,df$linoleic.acid)

# log-transform
# NOTE: per Qibin, lipids are already log transformed
df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins)] = 
  apply(df[,c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins)], 2, log)

# delete records with missing CACp
df <- df[!is.na(df$CACp),]

# remove low variance items
no_var = caret::nearZeroVar(df[,colnames(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)])
df = df[,-c(no_var)]

# remove the compounds with missing data and low variance from the strings of names
untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
global_proteins <- global_proteins[global_proteins %in% colnames(df)]
glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
lipids <- lipids[lipids %in% colnames(df)]

# remove markers with >=80% missing 
miss <- df[,colnames(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)] %>% select(where(~mean(is.na(.)) > 0.8))
# don't want to select clinical variables with >80% missing
miss <- miss[,colnames(miss) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]
df <- df %>% select(-one_of(colnames(miss)))

# remove the compounds with missing data and low variance from the strings of names
untargeted_metabs <- untargeted_metabs[untargeted_metabs %in% colnames(df)]
targeted_metabs <- targeted_metabs[targeted_metabs %in% colnames(df)]
global_proteins <- global_proteins[global_proteins %in% colnames(df)]
glycated_proteins <- glycated_proteins[glycated_proteins %in% colnames(df)]
lipids <- lipids[lipids %in% colnames(df)]

# calculate change in square-root transformed CAC volume
df$delta_cac_v2 <- (df$c2 - df$c1)
df$delta_cac_v3 <- (df$c3 - df$c1)
df$delta_cac_v4 <- (df$c4 - df$c1)

# create scaled version of df for pls-da of all measures
df_z <- df[,colnames(df) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)]
df_z <- scale(df_z)
```

```{r moderated ttests,echo=FALSE, message=FALSE, results="hide"}
cac_contrast <- ifelse(df$CACp=="Yes",1,0)
cac_contrast <- cbind(rep(1,nrow(df)),cac_contrast)

# untargeted metabolites
ymat <- t(df[,names(df) %in% untargeted_metabs])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_untargeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_untargeted_metabs <- results_untargeted_metabs[order(results_untargeted_metabs$P.Value),] 
# trying permutation test based adjustment
resT_untargeted_metabs<-mt.maxT(ymat,cac_contrast[,2])
resP_untargeted_metabs<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# targeted metabolites
ymat <- t(df[,names(df) %in% targeted_metabs])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_targeted_metabs <- topTable(fit,coef = 2,number = nrow(ymat))
results_targeted_metabs <- results_targeted_metabs[order(results_targeted_metabs$P.Value),] 
# trying permutation test based adjustment
#resT_targeted_metabs<-mt.maxT(ymat,cac_contrast[,2])
#resP_targeted_metabs<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# global proteomics
ymat <- t(df[,names(df) %in% global_proteins])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_global_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_global_proteins <- results_global_proteins[order(results_global_proteins$P.Value),] 
# trying permutation test based adjustment
resT_global_proteins<-mt.maxT(ymat,cac_contrast[,2])
resP_global_proteins<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# glycated proteomics
ymat <- t(df[,names(df) %in% glycated_proteins])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_glycated_proteins <- topTable(fit,coef = 2,number = nrow(ymat))
results_glycated_proteins <- results_glycated_proteins[order(results_glycated_proteins$P.Value),] 
# trying permutation test based adjustment
resT_glycated_proteins<-mt.maxT(ymat,cac_contrast[,2])
resP_glycated_proteins<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# lipidomics
ymat <- t(df[,names(df) %in% lipids])
fit <- lmFit(ymat,cac_contrast)
fit <- eBayes(fit)
results_lipids <- topTable(fit,coef = 2,number = nrow(ymat))
results_lipids <- results_lipids[order(results_lipids$P.Value),] 
# trying permutation test based adjustment
resT_lipids<-mt.maxT(ymat,cac_contrast[,2])
resP_lipids<-mt.minP(ymat,cac_contrast[,2])
ymat <- NULL
fit <- NULL

# write to file
wb <- createWorkbook()
addWorksheet(wb,"untargeted_metab_moderated_FDR")
writeData(wb,"untargeted_metab_moderated_FDR",results_untargeted_metabs,rowNames = T)
addWorksheet(wb,"untargeted_metab_maxT")
writeData(wb,"untargeted_metab_maxT",resT_untargeted_metabs,rowNames = T)
addWorksheet(wb,"untargeted_metab_minP")
writeData(wb,"untargeted_metab_minP",resP_untargeted_metabs,rowNames = T)
addWorksheet(wb,"targeted_metab_moderated_FDR")
writeData(wb,"targeted_metab_moderated_FDR",results_targeted_metabs,rowNames = T)
#addWorksheet(wb,"targeted_metabolomics_maxT")
#writeData(wb,"resT_targeted_metabolomics",results_untargeted_metabs,rowNames = T)
#addWorksheet(wb,"targeted_metabolomics_minP")
#writeData(wb,"resP_targeted_metabolomics",results_untargeted_metabs,rowNames = T)
addWorksheet(wb,"global_prot_moderated_FDR")
writeData(wb,"global_prot_moderated_FDR",results_global_proteins,rowNames = T)
addWorksheet(wb,"global_prot_maxT")
writeData(wb,"global_prot_maxT",resT_global_proteins,rowNames = T)
addWorksheet(wb,"global_prot_minP")
writeData(wb,"global_prot_minP",resP_global_proteins,rowNames = T)
addWorksheet(wb,"glycated_prot_moderated_FDR")
writeData(wb,"glycated_prot_moderated_FDR",results_glycated_proteins,rowNames = T)
addWorksheet(wb,"glycated_prot_maxT")
writeData(wb,"glycated_prot_maxT",resT_glycated_proteins,rowNames = T)
addWorksheet(wb,"glycated_prot_minP")
writeData(wb,"glycated_prot_minP",resP_glycated_proteins,rowNames = T)
addWorksheet(wb,"lipidomics_moderated_FDR")
writeData(wb,"lipidomics_moderated_FDR",results_lipids,rowNames = T)
addWorksheet(wb,"lipidomics_maxT")
writeData(wb,"lipidomics_maxT",resT_lipids,rowNames = T)
addWorksheet(wb,"lipidomics_minP")
writeData(wb,"lipidomics_minP",resP_lipids,rowNames = T)
saveWorkbook(wb,"AHA_omics_ttests.xlsx",overwrite = TRUE)
```

```{r sex interaction,echo=FALSE}
df$sexnum <- ifelse(df$gender=="M",0,1)
df$CACp_num <- ifelse(df$CACp=="Yes",1,0)
int_contrast <- model.matrix(~CACp_num*sexnum,df)

# untargeted metabolites
ymat <- t(df[,names(df) %in% untargeted_metabs])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_untargeted_metabs_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_untargeted_metabs_int <- results_untargeted_metabs_int[order(results_untargeted_metabs_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# targeted metabolites
ymat <- t(df[,names(df) %in% targeted_metabs])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_targeted_metabs_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_targeted_metabs_int <- results_targeted_metabs_int[order(results_targeted_metabs_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# global proteomics
ymat <- t(df[,names(df) %in% global_proteins])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_global_proteins_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_global_proteins_int <- results_global_proteins_int[order(results_global_proteins_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# glycated proteomics
ymat <- t(df[,names(df) %in% glycated_proteins])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_glycated_proteins_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_glycated_proteins_int <- results_glycated_proteins_int[order(results_glycated_proteins_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# lipidomics
ymat <- t(df[,names(df) %in% lipids])
fit <- lmFit(ymat,int_contrast)
fit <- eBayes(fit)
results_lipids_int <- topTable(fit,coef = 4,number = nrow(ymat))
results_lipids_int <- results_lipids_int[order(results_lipids_int$adj.P.Val),] 
ymat <- NULL
fit <- NULL

# write to file
wb_int <- createWorkbook()
addWorksheet(wb_int,"untargeted_metabolomics")
writeData(wb_int,"untargeted_metabolomics",results_untargeted_metabs_int,rowNames = T)
addWorksheet(wb_int,"targeted_metabolomics")
writeData(wb_int,"targeted_metabolomics",results_targeted_metabs_int,rowNames = T)
addWorksheet(wb_int,"global_proteomics")
writeData(wb_int,"global_proteomics",results_global_proteins_int,rowNames = T)
addWorksheet(wb_int,"glycated_proteins")
writeData(wb_int,"glycated_proteins",results_glycated_proteins_int,rowNames = T)
addWorksheet(wb_int,"lipidomics")
writeData(wb_int,"lipidomics",results_lipids_int,rowNames = T)
saveWorkbook(wb_int,"AHA_omics_sex_interaction.xlsx",overwrite = TRUE)
```

# Background

The purpose of this analysis is to examine differences in targeted and untargeted metabolomics, global and glycated proteomics, and lipidomics in CACTI participants with and without CAC progression.

# Methods

Data were log transformed prior to analysis with the exception of the lipid data, which had previously been log transformed.  Markers with >=80% missing data or near-zero variance were removed prior to analysis.

Moderated t-tests were used to compare CAC progression groups with FDR correction for multiple testing.  In addition, the maxT and minP permutation methods were used to correct for multiple testing.  

Linear models were used to test whether there was a significant interaction between CAC progression groups and sex.  The interpretation of the FC value for the interaction is the average additional FC for females with CAC progression.

Sparse PLS-DA was used to identify markers that best discriminated between the CAC progression groups.  Each sPLS-DA model was tuned to identify the number of components and the number of proteins that resulted in the lowest error rate.  Permutation testing with 1000 permutations was used to evaluate the extent of overfitting.

# Results

For moderated t-tests and linear models with interaction between CACp and sex, see separate files

## sPLS-DA of untargeted metabolites

```{r splsda utm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# untargeted metabolites
Y <- df$CACp
X <- df[,names(df) %in% untargeted_metabs]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA of untargeted metabolites by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## PLS regression of untargeted metabolites, change in CAC volume from v1-v4

```{r PLSR utm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
## PLS regression of untargeted metabolites
nomiss <- df[!is.na(df$delta_cac_v4),]
plsrx <- as.matrix(nomiss[,names(nomiss) %in% untargeted_metabs])
# get rid of rows with all missing
missrow <- apply(plsrx, 1, function(x) all(is.na(x)))
plsrx <- plsrx[!missrow,]
plsry <- as.matrix(nomiss$delta_cac_v4)
# remove rows of y corresponding to missrow
plsry <- plsry[!missrow,]

# nipals algorithm to impute missings
# once we get rid of rows with all missing, there are no missing values
#nipals.tune = nipals(plsrx, ncomp = 10)$eig
#barplot(nipals.tune, xlab = 'number of components', ylab = 'explained variance')
#nipals.X <- impute.nipals(plsrx, ncomp = 5)$rec

#SPLS
spls_utm <- spls(plsrx, plsry, ncomp = 2, mode = "regression",keepX = c(20,20))  # where ncomp is the number of dimensions/components to choose
# can't get tuning function to work
#tune.spls <- tune.spls(plsrx, plsry, ncomp=5, folds=5, measure="MSE", test.keepX = c(10,10,10,10,10), nrepeat = 3)
#plot(tune.pls$measures$Q2.total$values)
#plotIndiv(spls_utm)
#plotVar(spls_utm)
#par(mar=c(0.1,0.1,0.1,0.1))
#network(spls_utm, comp = 1:2, shape.node = c("rectangle", "rectangle"),
#        color.node = c("white", "pink"), color.edge =  color.GreenRed(50), )
c1 <- spls_utm$loadings$X[,1]
c1 <- as.data.frame(c1[c1 != 0])
c2 <- spls_utm$loadings$X[,2]
c2 <- as.data.frame(c2[c2 != 0])
colnames(c1) <- "Loading"
c1 <- c1 %>% arrange(Loading)
colnames(c2) <- "Loading"
c2 <- c2 %>% arrange(Loading) 

kable(c1)
kable(c2)
```

## sPLS-DA of targeted metabolites

```{r splsda tm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# targeted metabolites
Y <- df$CACp
X <- df[,colnames(df) %in% targeted_metabs]
tune.splsda <- tune.splsda(X, Y, ncomp=5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
#choice.keepX <- 10
## sPLS-DA function
splsda.tm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.tm,title = "sPLS-DA of targeted metabolites by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.tm,comp=1)
top_2 <- selectVar(splsda.tm,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.tm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## PLS regression of targeted metabolites, change in CAC volume from v1-v4

```{r PLSR tm,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
nomiss <- df[!is.na(df$delta_cac_v4),]
plsrx <- as.matrix(nomiss[,names(nomiss) %in% targeted_metabs])
# get rid of rows with all missing
missrow <- apply(plsrx, 1, function(x) all(is.na(x)))
plsrx <- plsrx[!missrow,]
plsry <- as.matrix(nomiss$delta_cac_v4)
# remove rows of y corresponding to missrow
plsry <- plsry[!missrow,]

#SPLS
spls_tm <- spls(plsrx, plsry, ncomp = 2, mode = "regression",keepX = c(20,20))  # where ncomp is the number of dimensions/components to choose
# can't get tuning function to work
#tune.spls <- tune.spls(plsrx, plsry, ncomp=5, folds=5, measure="MSE", test.keepX = c(10,10,10,10,10), nrepeat = 3)
#plot(tune.pls$measures$Q2.total$values)
#plotIndiv(spls_utm)
#plotVar(spls_utm)
#par(mar=c(0.1,0.1,0.1,0.1))
#network(spls_utm, comp = 1:2, shape.node = c("rectangle", "rectangle"),
#        color.node = c("white", "pink"), color.edge =  color.GreenRed(50), )
c1 <- spls_tm$loadings$X[,1]
c1 <- as.data.frame(c1[c1 != 0])
c2 <- spls_tm$loadings$X[,2]
c2 <- as.data.frame(c2[c2 != 0])
colnames(c1) <- "Loading"
c1 <- c1 %>% arrange(Loading)
colnames(c2) <- "Loading"
c2 <- c2 %>% arrange(Loading) 

kable(c1)
kable(c2)

```

## sPLS-DA of global proteins

```{r splsda gp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# global proteins
Y <- df$CACp
X <- df[,names(df) %in% global_proteins]
tune.splsda <- tune.splsda(X, Y, ncomp =  5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
#choice.keepX <- 10
## sPLS-DA function
splsda.globalp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.globalp,title = "sPLS-DA of global proteins by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.globalp,comp=1)
top_2 <- selectVar(splsda.globalp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.globalp, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## PLS regression of global proteins, change in CAC volume from v1-v4

```{r PLSR globp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
nomiss <- df[!is.na(df$delta_cac_v4),]
plsrx <- as.matrix(nomiss[,names(nomiss) %in% global_proteins])
# get rid of rows with all missing
missrow <- apply(plsrx, 1, function(x) all(is.na(x)))
plsrx <- plsrx[!missrow,]
plsry <- as.matrix(nomiss$delta_cac_v4)
# remove rows of y corresponding to missrow
plsry <- plsry[!missrow,]

#SPLS
spls_globalp <- spls(plsrx, plsry, ncomp = 2, mode = "regression",keepX = c(20,20))  # where ncomp is the number of dimensions/components to choose
# can't get tuning function to work
#tune.spls <- tune.spls(plsrx, plsry, ncomp=5, folds=5, measure="MSE", test.keepX = c(10,10,10,10,10), nrepeat = 3)
#plot(tune.pls$measures$Q2.total$values)
#plotIndiv(spls_utm)
#plotVar(spls_utm)
#par(mar=c(0.1,0.1,0.1,0.1))
#network(spls_utm, comp = 1:2, shape.node = c("rectangle", "rectangle"),
#        color.node = c("white", "pink"), color.edge =  color.GreenRed(50), )
c1 <- spls_globalp$loadings$X[,1]
c1 <- as.data.frame(c1[c1 != 0])
c2 <- spls_globalp$loadings$X[,2]
c2 <- as.data.frame(c2[c2 != 0])
colnames(c1) <- "Loading"
c1 <- c1 %>% arrange(Loading)
colnames(c2) <- "Loading"
c2 <- c2 %>% arrange(Loading) 

kable(c1)
kable(c2)

```

## sPLS-DA of glycated proteins

```{r splsda glyp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# glycated proteins
Y <- df$CACp
X <- df[,names(df) %in% glycated_proteins]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.glycp <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.glycp,title = "sPLS-DA of glycated proteins by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.glycp,comp=1)
top_2 <- selectVar(splsda.glycp,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.glycp, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## PLS regression of glycated proteins, change in CAC volume from v1-v4

```{r PLSR glyp,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
nomiss <- df[!is.na(df$delta_cac_v4),]
plsrx <- as.matrix(nomiss[,names(nomiss) %in% glycated_proteins])
# get rid of rows with all missing
missrow <- apply(plsrx, 1, function(x) all(is.na(x)))
plsrx <- plsrx[!missrow,]
plsry <- as.matrix(nomiss$delta_cac_v4)
# remove rows of y corresponding to missrow
plsry <- plsry[!missrow,]

#SPLS
spls_glyp <- spls(plsrx, plsry, ncomp = 2, mode = "regression",keepX = c(20,20))  # where ncomp is the number of dimensions/components to choose
# can't get tuning function to work
#tune.spls <- tune.spls(plsrx, plsry, ncomp=5, folds=5, measure="MSE", test.keepX = c(10,10,10,10,10), nrepeat = 3)
#plot(tune.pls$measures$Q2.total$values)
#plotIndiv(spls_utm)
#plotVar(spls_utm)
#par(mar=c(0.1,0.1,0.1,0.1))
#network(spls_utm, comp = 1:2, shape.node = c("rectangle", "rectangle"),
#        color.node = c("white", "pink"), color.edge =  color.GreenRed(50), )
c1 <- spls_glyp$loadings$X[,1]
c1 <- as.data.frame(c1[c1 != 0])
c2 <- spls_glyp$loadings$X[,2]
c2 <- as.data.frame(c2[c2 != 0])
colnames(c1) <- "Loading"
c1 <- c1 %>% arrange(Loading)
colnames(c2) <- "Loading"
c2 <- c2 %>% arrange(Loading) 

kable(c1)
kable(c2)

```

## sPLS-DA of lipids

```{r splsda lipids,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# lipids
Y <- df$CACp
X <- df[,names(df) %in% lipids]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
## sPLS-DA function
splsda.lipids <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.lipids,title = "sPLS-DA of lipids by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.lipids,comp=1)
top_2 <- selectVar(splsda.lipids,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

set.seed(3654)
plsda.perf = perf(splsda.lipids, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## PLS regression of lipids, change in CAC volume from v1-v4

```{r PLSR lipids,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
nomiss <- df[!is.na(df$delta_cac_v4),]
plsrx <- as.matrix(nomiss[,names(nomiss) %in% lipids])
# get rid of rows with all missing
missrow <- apply(plsrx, 1, function(x) all(is.na(x)))
plsrx <- plsrx[!missrow,]
plsry <- as.matrix(nomiss$delta_cac_v4)
# remove rows of y corresponding to missrow
plsry <- plsry[!missrow,]

#SPLS
spls_lipids <- spls(plsrx, plsry, ncomp = 2, mode = "regression",keepX = c(20,20))  # where ncomp is the number of dimensions/components to choose
# can't get tuning function to work
#tune.spls <- tune.spls(plsrx, plsry, ncomp=5, folds=5, measure="MSE", test.keepX = c(10,10,10,10,10), nrepeat = 3)
#plot(tune.pls$measures$Q2.total$values)
#plotIndiv(spls_utm)
#plotVar(spls_utm)
#par(mar=c(0.1,0.1,0.1,0.1))
#network(spls_utm, comp = 1:2, shape.node = c("rectangle", "rectangle"),
#        color.node = c("white", "pink"), color.edge =  color.GreenRed(50), )
c1 <- spls_lipids$loadings$X[,1]
c1 <- as.data.frame(c1[c1 != 0])
c2 <- spls_lipids$loadings$X[,2]
c2 <- as.data.frame(c2[c2 != 0])
colnames(c1) <- "Loading"
c1 <- c1 %>% arrange(Loading)
colnames(c2) <- "Loading"
c2 <- c2 %>% arrange(Loading) 

kable(c1)
kable(c2)

```

## sPLS-DA of all data combined

```{r splsda all,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
# everything
Y <- df$CACp
X <- df_z
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2]
#choice.keepX <- 10
## sPLS-DA function
splsda.all <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.all,title = "sPLS-DA of all markers by CACp",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.all,comp=1)
top_2 <- selectVar(splsda.all,comp=2)
top_1d <- top_1$value
top_2d <- top_2$value
top_1d
top_2d

# set.seed(3654)
plsda.perf = perf(splsda.all, validation = "loo",auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])

# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(df$CACp),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## PLS regression of all data, change in CAC volume from v1-v4

```{r PLSR all,echo=FALSE,cache=TRUE,warning=FALSE,message=FALSE, comment=""}
nomiss <- df[!is.na(df$delta_cac_v4),]
plsrx <- as.matrix(nomiss[,names(nomiss) %in% c(untargeted_metabs,targeted_metabs,global_proteins,glycated_proteins,lipids)])
# get rid of rows with all missing
missrow <- apply(plsrx, 1, function(x) all(is.na(x)))
plsrx <- plsrx[!missrow,]
plsry <- as.matrix(nomiss$delta_cac_v4)
# remove rows of y corresponding to missrow
plsry <- plsry[!missrow,]

#SPLS
spls_all <- spls(plsrx, plsry, ncomp = 2, mode = "regression",keepX = c(20,20))  # where ncomp is the number of dimensions/components to choose
# can't get tuning function to work
#tune.spls <- tune.spls(plsrx, plsry, ncomp=5, folds=5, measure="MSE", test.keepX = c(10,10,10,10,10), nrepeat = 3)
#plot(tune.pls$measures$Q2.total$values)
#plotIndiv(spls_utm)
#plotVar(spls_utm)
#par(mar=c(0.1,0.1,0.1,0.1))
#network(spls_utm, comp = 1:2, shape.node = c("rectangle", "rectangle"),
#        color.node = c("white", "pink"), color.edge =  color.GreenRed(50), )
c1 <- spls_all$loadings$X[,1]
c1 <- as.data.frame(c1[c1 != 0])
c2 <- spls_all$loadings$X[,2]
c2 <- as.data.frame(c2[c2 != 0])
colnames(c1) <- "Loading"
c1 <- c1 %>% arrange(Loading)
colnames(c2) <- "Loading"
c2 <- c2 %>% arrange(Loading) 

kable(c1)
kable(c2)

```