---
title: "AHA clinical model - all variables - addtional analyses for ADA"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(broom)
library(survival)
library(performance)
library(knitr)
library(pROC)
library(naniar)
library(readxl)
library(psych)
library(pheatmap)
library(tableone)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggplot2)
library(reshape2)
knitr::opts_chunk$set(echo = FALSE,fig.height = 10,fig.width = 10)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}

knitr::opts_knit$set(root.dir = home_dir)
#rm(home_dir)
```

```{r}
# Function and master data
source("~/Documents/Github/shared-resources/Machine Learning/Tim - ElasticNet CV/easy_elasticnet.R")
load("~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/aha_master_data_no_snps.Rdata")
# omics selected by moderated t-test and PLSDA
omics = unique(c("gly_P02671.7","gly_P02671.5","gly_P02671.4","gly_P02671.3",
          "AAA01201000179.0807263.4","gly_P02679","HMDB00510161.0696402.9",
          "AAA01201000179.0809235.7","gly_P02675.4","gly_P02675","gly_P02647",
          "P02655","gly_P02647.1","HMDB00510161.0694436","gly_P02652",
          "gly_P01009","gly_P01834.11","P01817","AcCa 10:3",
          "HMDB0028822260.1378628.6","AAA01201000179.0807263.4",
          "gly_P02671.3","gly_P02671.7"))
limited_omics = c("AAA01201000179.0807263.4","gly.P02671.7","gly.P02671.3",
          "gly.P02671.5","gly.P02671.4")
# Take out people without diabetes, remove diabetes status from predictors
df = df[df$diabetic == 1,]
clinical_predictors = clinical_predictors[-which(clinical_predictors == "dia")]
# Log transform 
df[,omics] = log(df[,omics])
# Fix names
colnames(df)[which(colnames(df) %in% omics)] = make.names(omics,unique = T,allow_ = F)
omics = make.names(omics,unique = T,allow_ = F)
# Variables for null model
null_vars <- c("age","sex","durcatV1")

# add in post-baseline A1c values
sample_info <- read_csv("~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Janet Snell-Bergeon/AHA collaborative grant/Combined predictive model/QAllVisits_Long_V1_V4_Glycemic_Vars.csv")
# calculate within person variability
sample_info <- sample_info[,c("StudyID","hba1c","glu","egdr","eis_nf")]
sample_info$StudyID <- as.factor(as.character(sample_info$StudyID))
sample_info_var <- sample_info %>% group_by(StudyID) %>% summarise(across(.cols = where(is.numeric),.fns = list(SD=sd, MEAN=mean), na.rm=T, .names="{col}_{fn}"))
# merge in omics data
corr_df <- merge(sample_info_var,df_labelled[,c("StudyID",limited_omics)],by="StudyID",all.x = F, all.y = T)

# exploratory correlations
explor <- corr_df
explor <- merge(explor, df[,c("StudyID",clinical_predictors)], by="StudyID", all.x=T, all.y=F)
```

# Descriptive statistics - participants included in the clinical model

```{r include=FALSE}
t1data <- df[complete.cases(df[,clinical_predictors]),]
t1data[,c("CACprogV3","CAD","CVD")] <- apply(t1data[,c("CACprogV3","CAD","CVD")],2,as.factor)
t1 <- CreateTableOne(data=t1data,vars=c("CACprogV3","CAD","CVD","age","sex","durationV1"))
t1 <- print(t1)
```

```{r}
kable(t1)
```

# Descriptive statistics - participants included in the omics model

```{r include=FALSE}
t2data <- df[complete.cases(df[,c(clinical_predictors,limited_omics)]),]
t2data[,c("CACprogV3","CAD","CVD")] <- apply(t2data[,c("CACprogV3","CAD","CVD")],2,as.factor)
t2 <- CreateTableOne(data=t2data,vars=c("CACprogV3","CAD","CVD","age","sex","durationV1"))
t2 <- print(t2)
```

```{r}
kable(t2)
```


# ElasticNet on clinical variables only

## CAC progression at visit 3

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
cac_prog_select = easy_elasticnet(data = df,outcome = "CACprogV3",
                           predictors = clinical_predictors,
                           model_type = "binomial")
```

### Null model (age, sex, T1D duration only)

```{r}
n_cac_prog = as.formula(paste0("CACprogV3~",paste0(null_vars,collapse = "+")))
null_mod = glm(formula = n_cac_prog,data = df,family = "binomial")
kable(tidy(null_mod, exponentiate=TRUE, conf.int=TRUE),digits = 3)
```

### Clinical model results

```{r}
f_cac_prog = as.formula(paste0("CACprogV3~",paste0(cac_prog_select,collapse = "+")))
clin_mod = glm(formula = f_cac_prog,data = df,family = "binomial")
kable(tidy(clin_mod, exponentiate=TRUE, conf.int=TRUE),digits = 3)
```

```{r}
df_labelled <- df
df_labelled$acV1 <- set_label(df_labelled$acV1,"Albumin:creatinine*")
df_labelled$age <- set_label(df_labelled$age,"Age (yrs)")
df_labelled$avesystbpV1 <- set_label(df_labelled$avesystbpV1,"SBP (mmHg)*")
df_labelled$bmiV1 <- set_label(df_labelled$bmiV1,"BMI (kg/m2)")
df_labelled$cholV1 <- set_label(df_labelled$cholV1,"Total cholesterol (mg/dL)")
df_labelled$durcatV1 <- set_labels(df_labelled$durcatV1,labels=c("T1D duration <20 yrs", "T1D duration 20-39 yrs","T1D duration 40+ yrs"))
df_labelled$hba1cV1 <- set_label(df_labelled$hba1cV1,"HbA1c (%)*")
df_labelled$hdlcV1 <- set_label(df_labelled$hdlcV1,"HDL (mg/dL)*")
df_labelled$homoV1 <- set_label(df_labelled$homoV1,"Homocysteine (umol/L)*")
df_labelled$insdoseperkgV1 <- set_label(df_labelled$insdoseperkgV1,"Insulin dose (units/kg)*")
df_labelled$l45sqfV1 <- set_label(df_labelled$l45sqfV1,"Subq fat L4-L5 (mm3)*")
df_labelled$onhypermedsV1 <- set_labels(df_labelled$onhypermedsV1,labels=c("No anti-hypertensives","Anti-hypertensives"))
df_labelled$onlipidmedsV1 <- set_labels(df_labelled$onlipidmedsV1,labels=c("No lipid medications","Lipid medications"))
df_labelled$PAT_V1 <- set_label(df_labelled$PAT_V1,"Pericardial fat (cm3)*")
df_labelled$sex <- set_labels(df_labelled$sex,labels=c("Male","Female"))
df_labelled$smknum <- set_labels(df_labelled$smknum,labels=c("Non-smoker","Current smoker"))
df_labelled$triV1 <- set_label(df_labelled$triV1,"Triglycerides (mg/dL)*")
df_labelled$UA_V1 <- set_label(df_labelled$UA_V1,"Serum uric acid (mg/dL)*")
df_labelled$AAA01201000179.0807263.4 <- set_label(df_labelled$AAA01201000179.0807263.4,"D-Glucosamine/D-Galactosamine*")
df_labelled$gly.P02671.3 <- set_label(df_labelled$gly.P02671.3,"Glycated fibrinogen 3*")
df_labelled$gly.P02671.7 <- set_label(df_labelled$gly.P02671.7,"Glycated fibrinogen 7*")
df_labelled$gly.P02671.5 <- set_label(df_labelled$gly.P02671.5,"Glycated fibrinogen 5*")
df_labelled$gly.P02671.4 <- set_label(df_labelled$gly.P02671.4,"Glycated fibrinogen 4*")
clin_mod_labelled = glm(formula = f_cac_prog,data = df_labelled,family = "binomial")
plot_model(clin_mod_labelled, title="Clinical model of CAC progression", auto.label = TRUE)
```

### Omics model results (no imputation)

```{r}
omics_mod = glm(formula = update(f_cac_prog,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
          data = df,family = "binomial")
kable(tidy(omics_mod, exponentiate=TRUE, conf.int=TRUE),digits = 3)
```

```{r}
omics_mod_labelled = glm(formula = update(f_cac_prog,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
          data = df_labelled,family = "binomial")
plot_model(omics_mod_labelled, title="Omics model of CAC progression", auto.label = TRUE) 

```

### Model comparison

Green indicates the omics model, blue is the clinical model, and red is the null model (age, sex, T1D duration only).

```{r message=FALSE}
# Predicted values
clin_pred = predict(clin_mod,type="response")
omics_pred = predict(omics_mod,type="response")
null_pred = predict(null_mod,type="response")
# Plot
par(pty="s")
plot(roc(clin_mod$model$CACprogV3, clin_pred), 
                 levels = c("No Progression","Progression"),direction = "<",
                 print.auc = TRUE, col = "blue",quiet = T,print.auc.y = .4)
plot(roc(omics_mod$model$CACprogV3, omics_pred), 
                 levels = c("No Progression","Progression"),direction = "<",
                 print.auc = TRUE, col = "black",add = T,quiet = T,print.auc.y = .5)
plot(roc(null_mod$model$CACprogV3, null_pred), 
                 levels = c("No Progression","Progression"),direction = "<",
                 print.auc = TRUE, col = "red",add = T,quiet = T,print.auc.y = .3)
```

## CAC progression at last visit

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
cac_prog_any_select = easy_elasticnet(data = df,outcome = "cac_prog_last_vis",
                           predictors = clinical_predictors,
                           model_type = "binomial")
```

### Null model (age, sex, T1D duration only)

```{r}
n_cac_prog_any = as.formula(paste0("cac_prog_last_vis~",paste0(null_vars,collapse = "+")))
null_mod = glm(formula = n_cac_prog_any,data = df,family = "binomial")
kable(tidy(null_mod, exponentiate=TRUE, conf.int=TRUE),digits = 3)
```

### Clinical model results

```{r}
f_cac_prog_any = as.formula(paste0("cac_prog_last_vis~",paste0(cac_prog_any_select,
                                                               collapse = "+")))
clin_mod = glm(formula = f_cac_prog_any,data = df,family = "binomial")
kable(tidy(clin_mod, exponentiate=TRUE, conf.int=TRUE),digits = 3)
```

### Omics model results (no imputation)

```{r}
omics_mod = glm(formula = 
                  update(f_cac_prog_any,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
          data = df,family = "binomial")
kable(tidy(omics_mod, exponentiate=TRUE, conf.int=TRUE),digits = 3)
```

### Model comparison

Green indicates the omics model and blue is the clinical model.

```{r message=FALSE}
# Predicted values
clin_pred = predict(clin_mod,type="response")
omics_pred = predict(omics_mod,type="response")
null_pred = predict(null_mod,type="response")
# Plot
plot(roc(clin_mod$model$cac_prog_last_vis, clin_pred), 
                 levels = c("No Progression","Progression"),direction = "<",
                 print.auc = TRUE, col = "blue",quiet = T)
plot(roc(omics_mod$model$cac_prog_last_vis, omics_pred), 
                 levels = c("No Progression","Progression"),direction = "<",
                 print.auc = TRUE, col = "green",add = T,quiet = T,print.auc.y = .4)
plot(roc(null_mod$model$cac_prog_last_vis, null_pred), 
                 levels = c("No Progression","Progression"),direction = "<",
                 print.auc = TRUE, col = "red",add = T,quiet = T,print.auc.y = .2)
```

## Change in CAC per year

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
cac_change_select = easy_elasticnet(data = df,outcome = "cac_change_per_yr",
                           predictors = clinical_predictors,
                           model_type = "gaussian")
```

### Null model (age, sex, T1D duration only)

```{r}
n_cac_change  = as.formula(paste0("cac_change_per_yr~",paste0(null_vars,collapse = "+")))
null_mod = lm(formula = n_cac_change,data = df)
kable(tidy(null_mod),digits = 3)
```

### Clinical model results

```{r}
f_cac_change = as.formula(paste0("cac_change_per_yr~",
                                 paste0(cac_change_select,collapse = "+")))
clin_mod = lm(formula = f_cac_change,data = df)
kable(tidy(clin_mod),digits = 3)
```

### Omics model results (no imputation)

```{r}
omics_mod = lm(formula = update(f_cac_change,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
         data = df)
kable(tidy(omics_mod),digits = 3)
```

### Model comparison

```{r}
mc = as.data.frame(compare_performance(clin_mod,omics_mod,null_mod,
                                       metrics = c("R2","R2_adj"),verbose = F))
mc$Name = c("Clinical Model","Omics Model","Null model")
kable(mc,digits = 3)
```

## Death

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
death_select = easy_elasticnet(data = df,outcome = "Deceased",
                           predictors = clinical_predictors,
                           model_type = "cox",time = "PersonYrsDeath")
```

### Null model (age, sex, T1D duration only)

```{r}
n_death  = as.formula(paste0("Surv(time = df$PersonYrsDeath,event = df$Deceased)~",paste0(null_vars,collapse = "+")))
null_mod = coxph(formula = n_death,data = df,id = StudyID)
kable(tidy(null_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Clinical model results

```{r}
f_death = as.formula(paste0("Surv(time = df$PersonYrsDeath,event = df$Deceased)~",
                      paste0(death_select,collapse = "+")))
clin_mod = coxph(formula = f_death,data = df,id = StudyID)
kable(tidy(clin_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Omics model results (no imputation)

The coefficient for `durcat` refuses to converge.

```{r}
omics_mod = coxph(formula = update(f_death,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
            data = df,id = StudyID,control = coxph.control(iter.max = 1000))
kable(tidy(omics_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Model comparison

Concordance in the clinical model was `r round(as.numeric(summary(clin_mod)$concordance[1]),3)` compared to `r round(as.numeric(summary(omics_mod)$concordance[1]),3)` for the omics model and `r round(as.numeric(summary(null_mod)$concordance[1]),3)` for the null model.

## CAD

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
cad_select = easy_elasticnet(data = df,outcome = "CAD",
                           predictors = clinical_predictors,
                           model_type = "cox",time = "PersonYrsCAD")
```

### Null model (age, sex, T1D duration only)

```{r}
n_cad  = as.formula(paste0("Surv(time = df$PersonYrsCAD,event = df$CAD)~",paste0(null_vars,collapse = "+")))
null_mod = coxph(formula = n_cad,data = df,id = StudyID)
kable(tidy(null_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Clinical model results

```{r}
f_cad = as.formula(paste0("Surv(time = df$PersonYrsCAD,event = df$CAD)~",
                      paste0(cad_select,collapse = "+")))
clin_mod = coxph(formula = f_cad,data = df,id = StudyID)
kable(tidy(clin_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Omics model results (no imputation)

```{r}
omics_mod = coxph(formula = update(f_cad,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
            data = df,id = StudyID)
kable(tidy(omics_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Model comparison

Concordance in the clinical model was `r round(as.numeric(summary(clin_mod)$concordance[1]),3)` compared to `r round(as.numeric(summary(omics_mod)$concordance[1]),3)` for the omics model and `r round(as.numeric(summary(null_mod)$concordance[1]),3)` for the null model.

## Hard CAD

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
hard_cad_select = easy_elasticnet(data = df,outcome = "HardCAD",
                           predictors = clinical_predictors,
                           model_type = "cox",time = "PersonYrsHardCAD")
```

### Null model (age, sex, T1D duration only)

```{r}
n_hard_cad  = as.formula(paste0("Surv(time = df$PersonYrsHardCAD,event = df$HardCAD)~",paste0(null_vars,collapse = "+")))
null_mod = coxph(formula = n_hard_cad,data = df,id = StudyID)
kable(tidy(null_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Clinical model results

```{r}
f_hard_cad = as.formula(paste0("Surv(time = df$PersonYrsHardCAD,event = df$HardCAD)~",
                      paste0(hard_cad_select,collapse = "+")))
clin_mod = coxph(formula = f_hard_cad,data = df,id = StudyID)
kable(tidy(clin_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Omics model results (no imputation)

`NHW` and `smknum` refuse to converge.

```{r}
omics_mod = coxph(formula = update(f_hard_cad,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
            data = df,id = StudyID,control = coxph.control(iter.max = 1000))
kable(tidy(omics_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Model comparison

Concordance in the clinical model was `r round(as.numeric(summary(clin_mod)$concordance[1]),3)` compared to `r round(as.numeric(summary(omics_mod)$concordance[1]),3)` for the omics model and `r round(as.numeric(summary(null_mod)$concordance[1]),3)` for the null model.

## CVD

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
cvd_select = easy_elasticnet(data = df,outcome = "CVD",
                           predictors = clinical_predictors,
                           model_type = "cox",time = "PersonYrsCVD")
```

### Null model (age, sex, T1D duration only)

```{r}
n_cvd  = as.formula(paste0("Surv(time = df$PersonYrsCVD,event = df$CVD)~",paste0(null_vars,collapse = "+")))
null_mod = coxph(formula = n_cvd,data = df,id = StudyID)
kable(tidy(null_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Clinical model results

```{r}
f_cvd = as.formula(paste0("Surv(time = df$PersonYrsCVD,event = df$CVD)~",
                      paste0(cvd_select,collapse = "+")))
clin_mod = coxph(formula = f_cvd,data = df,id = StudyID)
kable(tidy(clin_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Omics model results (no imputation)

```{r}
omics_mod = coxph(formula = update(f_cvd,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
            data = df,id = StudyID)
kable(tidy(omics_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r}
f_cvd_labelled = as.formula(paste0("Surv(time = PersonYrsCVD,event = CVD)~",
                      paste0(cvd_select,collapse = "+")))
omics_mod_labelled = coxph(formula = update(f_cvd,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
            data = df_labelled,id = StudyID)
plot_model(omics_mod_labelled, title="Omics model of CVD events", auto.label = TRUE)
```


### Model comparison

Concordance in the clinical model was `r round(as.numeric(summary(clin_mod)$concordance[1]),3)` compared to `r round(as.numeric(summary(omics_mod)$concordance[1]),3)` for the omics model and `r round(as.numeric(summary(null_mod)$concordance[1]),3)` for the null model.

## Hard CVD

```{r include=FALSE}
# Sometimes using too many alpha/lambda values results in duplicate values for the 
# error metric. This doesn't seem to be a problem but throws a lot of warnings
hard_cvd_select = easy_elasticnet(data = df,outcome = "HardCVD",
                           predictors = clinical_predictors,
                           model_type = "cox",time = "PersonYrsHardCVD")
```

### Null model (age, sex, T1D duration only)

```{r}
n_hard_cvd  = as.formula(paste0("Surv(time = df$PersonYrsHardCVD,event = df$HardCVD)~",paste0(null_vars,collapse = "+")))
null_mod = coxph(formula = n_hard_cvd,data = df,id = StudyID)
kable(tidy(null_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Clinical model results

```{r}
f_hard_cvd = as.formula(paste0("Surv(time = df$PersonYrsHardCVD,event = df$HardCVD)~",
                      paste0(hard_cvd_select,collapse = "+")))
clin_mod = coxph(formula = f_hard_cvd,data = df,id = StudyID)
kable(tidy(clin_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Omics model results (no imputation)

Basically this whole model refused to converge.

```{r}
omics_mod = coxph(formula = update(f_hard_cvd,paste0("~ . +",paste0(limited_omics,collapse = "+"))),
            data = df,id = StudyID,control = coxph.control(iter.max = 1000))
kable(tidy(omics_mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

### Model comparison

Concordance in the clinical model was `r round(as.numeric(summary(clin_mod)$concordance[1]),3)` compared to `r round(as.numeric(summary(omics_mod)$concordance[1]),3)` for the omics model and `r round(as.numeric(summary(null_mod)$concordance[1]),3)` for the null model.

# Correlation between omics markers and glycemic variability

```{r}
c = corr.test(corr_df[,-1],use = "pairwise.complete.obs")
pheatmap(c$r,display_numbers = T)
```

# Exploratory correlations

## Heat map of all clinical variables

```{r}
explor$smknum <- NULL
explor[,-1] <- apply(explor[,-1],2,as.numeric)
ce = corr.test(explor[,-1],use = "pairwise.complete.obs")
pheatmap(ce$r,display_numbers = T)
```

## Heat map of significant variables

```{r}

# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
ce_flat <- flattenCorrMatrix(ce$r, ce$p)
ce_flat <- ce_flat[ce_flat$row %in% limited_omics,]
ce_flat <- ce_flat[!ce_flat$column %in% limited_omics,]
ce_flat$cor <- round(ce_flat$cor,3)
ce_flat$p <- round(ce_flat$p,3)
ce_flat_sig <- ce_flat[ce_flat$p<0.05,]

explor_sig <- explor[,colnames(explor) %in% ce_flat_sig$row | colnames(explor) %in% ce_flat_sig$column]
colnames(explor_sig) <- c("D-Glucosamine","Glycated fibrinogen 7","Glycated fibrinogen 3","Glycated fibrinogen 5",
                          "HDL","Homocysteine","Pericardial fat","Sex","Triglycerides","Serum uric acid","Waist:hip ratio")
ce_sig = corr.test(explor_sig,use = "pairwise.complete.obs")
pheatmap(ce_sig$r,display_numbers = T)
```

## Omics markers by sex

```{r, include=FALSE}
stable <- CreateTableOne(data=df, vars = limited_omics, strata = "sex")
stable <- print(stable)

sdata <- df_labelled[,c("sex",limited_omics)]
sdata_long <- melt(sdata)
sdata_long$variable <- ifelse(sdata_long$variable=="AAA01201000179.0807263.4","D-Glucosamine",
                             ifelse(sdata_long$variable=="gly.P02671.7","Glycated fibrinogen 7",
                             ifelse(sdata_long$variable=="gly.P02671.3","Glycated fibrinogen 3",
                             ifelse(sdata_long$variable=="gly.P02671.5","Glycated fibrinogen 5",
                             ifelse(sdata_long$variable=="gly.P02671.4","Glycated fibrinogen 4",NA)))))
                              
p <- ggplot(sdata_long, aes(x=variable,y=value,fill=sex)) + geom_boxplot(lwd=1) +
  xlab("") + ylab("") + theme(axis.text.x=element_text(angle=45, vjust = 1, hjust = 1)) +
  scale_fill_manual(values=c("#ed2213","#809fd1"),name="Sex",labels=c("Male","Female")) + ylim(-2,2.5) + theme(text=element_text(size=30))
  
p
```

```{r}
kable(stable)
```

