---
title: "AHA Clinical Model Imputation"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(caret)
library(mice)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# Import
load("./Janet Snell-Bergeon/AHA collaborative grant/aha_master_data.Rdata")
# Make new variables
df$smknum = as.numeric(!(df$SmkStatusV1 == "Former" | df$SmkStatusV1 == "Never"))
df$SmkPackYrsV1[df$SmkStatusV1 == "Former" | df$SmkStatusV1 == "Never"] = 0
df$smknum = df$smknum * df$SmkPackYrsV1
df$duration_cat = cut(df$durationV1,breaks = c(-Inf,23,Inf),right = F)
# Columns from Laura's model
columns_from_laura = read.csv("./Janet Snell-Bergeon/AHA collaborative grant/Combined predictive model/selected_vars.csv")
columns_from_laura$X = NULL
columns_from_laura = colnames(columns_from_laura)
columns_from_laura = c(columns_from_laura,"gly_P02671.7","gly_P02671.5","gly_P02671.4",
                       "gly_P02671.3","AAA01201000179.0807263.4","gly_P02679",
                       "HMDB00510161.0696402.9","AAA01201000179.0809235.7",
                       "gly_P02675.4","gly_P02675","gly_P02647","P02655","gly_P02647.1",
                       "HMDB00510161.0694436","gly_P02652","gly_P01009","gly_P01834.11",
                       "P01817","AcCa 10:3","HMDB0028822260.1378628.6",
                       "AAA01201000179.0807263.4","gly_P02671.3","gly_P02671.7")
#columns_from_laura = c(snps,columns_from_laura)
outcomes = c("CACprogV3_num","cac_change_per_yr","death","CAD","HardCAD","CVD","HardCVD")
# Make new variables
df$nhw = as.factor(df$NHW)
df$sex_0m1f = factor(df$gender,levels = c("M","F"))
colnames(df) = sub("V1$","",colnames(df))
df$cholV1 = df$chol
# Factor variables
cat_vars = c("onhypermeds","onlipidmeds")
df[,cat_vars] = lapply(df[,cat_vars], as.factor)
# Log transform some
to_transform = c("bmi","avesystbp","hdlc","tri","CKDepi","ac",
                 "insdoseperkg","hba1c","l45sqf","homo","pai1","UA",
                 "l45vsf","pfatcm_v1","crp","fib")
log_transformed = paste0("l.",sub("V1","",to_transform))
df[,log_transformed] = log(df[,to_transform])
df[,to_transform] = NULL
# Remove duplicates
df = df[!df$StudyID %in% df$StudyID[duplicated(df$StudyID)],]
# Rename columns - mice has trouble with spaces and punctuation
columns_from_laura = gsub(" |[[:punct:]]","_",columns_from_laura)
colnames(df) = gsub(" |[[:punct:]]","_",colnames(df))
# Remove variables no variance for imputation model
exclude = c("StudyID",snps)
t = df %>% select(all_of(columns_from_laura)) %>%
  select(-any_of(exclude))
```

# Imputation

## Missing data

```{r fig.height=10,fig.width=10}
p = md.pattern(t, plot = T,rotate.names = T)
```

The numbers on the left side of the above plot indicate the number of observations with that pattern of missing data (blue = observed and red = missing), and the numbers on the right indicate the number of missing variables in that pattern. The bottom numbers indicate the raw number of missing observations per variable ad the total. 

## Complete cases regression



## Compared to Laura's analysis dataset

```{r}
laura_data = read.csv("./Janet Snell-Bergeon/AHA collaborative grant/Combined predictive model/ada_data_cacp.csv")
laura_data$X = NULL
colnames(laura_data) = gsub(" |[[:punct:]]","_",colnames(laura_data))
```