---
title: "Mixed Model Walkthrough"
author: "Tim Vigers"
date: "5/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nlme)
library(lme4)
library(ggplot2)
library(knitr)
```

## The Data

For this tutorial we will be using data on the average reaction time per day for subjects in a sleep deprivation study (data is included with the lme4 package). On day 0 the subjects had their normal amount of sleep. Starting that night they were restricted to 3 hours of sleep per night.

First, load the data and look at its structure:

```{r}
data("sleepstudy")
```

## Plots

Make a spaghetti plot (one line per subject) of reaction over time.

```{r spaghetti plot}

```

Now make a LOESS plot (one smooth line for the whole cohort):

```{r loess plot,message=FALSE}

```

## Models

Fit two models, random intercept:

$$
Reaction_i = \beta_0 + \beta_1Days_i + b_{0i} + \epsilon_i
$$

And random intercept and random slope:

$$
Reaction_i = \beta_0 + (\beta_1+b_{1i})Days_i + b_{0i} + \epsilon_i
$$

You can use either the nlme package or the lme4. What are some advantages and disadvantages of each?

```{r models}

```

Compare the two models. Use one numeric test and examine the residuals:

```{r compare}

```

Which model is best? Extract the results from your preferred model object and interpret them.

```{r results}

```
