---
title: "JDRF Models"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,dpi = 600)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Greg Forlenza/JDRF")
library(Hmisc)
library(arsenal)
library(skimr)
library(knitr)
library(Epi)
library(plotROC)
library(glmnet)
library(emmeans)
library(tidyverse)
```

```{r import data,echo=FALSE}
source("~/GitHub/BDC-Code/Greg Forlenza/JDRF/data_import.R")
# Exclude participants after 115 and those who never started AM
exclude <- data$record_id[which(data$gyl_timepoint==0 & is.na(data$automode_start))]
data <- data %>% filter(record_id <= 115,!(record_id %in% exclude))
```

```{r data cleaning,echo=FALSE,message=FALSE}
# Binary endpoint and predictors
# The primary endpoint of interest will be percent time in closed loop at 1 year 
# captured as a continuous percentage for the 2 weeks prior to the 1 year clinical 
# follow up visit, and will be converted to a binary variable with ≥60% representing 
# successful AP use and <60% representing unsuccessful AP use.
df <- left_join(data %>% group_by(record_id) %>% 
                  filter(gyl_timepoint.factor =="12 Months" | 
                           gyl_timepoint.factor =="9 Months") %>%
                  filter(row_number() == n()) %>%
                  summarise(am_final = time_am), 
                data %>% filter(grepl("baseline",redcap_event_name)),
                by = "record_id")
# Remove duplicates
df <- df[-c(which(duplicated(df$record_id))-1),]
# PAID
df$c_paid_total <- 
  apply(df[,which(colnames(df)=="c_paid1"):
             which(colnames(df)=="c_paid20")],1,
        function(x){mean(5-x)*25})
df$ya_paid_total <- 
  apply(df[,which(colnames(df)=="ya_paid1"):
             which(colnames(df)=="ya_paid20")],1,
        function(x){sum(x)*1.25})
# HFS behave
df$c_hfs_behave_total <- 
  apply(df[,which(colnames(df)=="c_hfs_behave1"):
             which(colnames(df)=="c_hfs_behave10")],1,
        function(x){sum(x)})
df$ya_hfs_behave_total <- 
  apply(df[,which(colnames(df)=="ya_hfs_b_behave1"):
             which(colnames(df)=="ya_hfs_b_behave15")],1,
        function(x){sum(x)})
# HFS worry
df$c_hfs_worry_total <- 
  apply(df[,which(colnames(df)=="c_hfs_worry11"):
             which(colnames(df)=="c_hfs_worry25")],1,
        function(x){sum(x)})
df$ya_hfs_worry_total <- 
  apply(df[,which(colnames(df)=="ya_hfs_b_worry16"):
             which(colnames(df)=="ya_hfs_b_worry33")],1,
        function(x){sum(x)})
# Add 1 month variables, select relevant columns
df <- left_join(df,data %>% 
                  filter(gyl_timepoint.factor == "Month 1/ Training F/U") %>%
                  mutate(m1_sensor_wear = sensor_wear,
                         m1_time_am = time_am,
                         m1_tir = sensor_70_180,
                         m1_boluses = dailybolus) %>%
                  select(record_id,m1_sensor_wear,m1_time_am,m1_tir,m1_boluses),
                by = "record_id") %>%
  select(record_id,am_final,hba1c,c_paid_total,ya_paid_total,
         c_hfs_behave_total,ya_hfs_behave_total,
         c_hfs_worry_total,ya_hfs_worry_total,
         m1_sensor_wear,m1_time_am,m1_tir,m1_boluses,
         demographics_t1d_duration,demographics_age,
         demographics_sex.factor,demographics_insurance.factor,
         demographics_race.factor,demographics_ethnicity.factor,
         demographics_cgmhx.factor,demographics_pumphx.factor)
# Binary outcome
df$success <- as.factor(ifelse(df$am_final >= 60,1,0))
levels(df$success) <- c("Failure","Success")
# Categorical variables
df$Age <- df$demographics_age
df$demographics_age <- cut(df$demographics_age,c(5,13,18,Inf),right = T,
                           labels = c("6 - 13","14 - 18","18+"))
levels(df$demographics_race.factor) = 
  c("Non-white","Non-white","Non-white","Non-white","White","Non-white","Non-white")
levels(df$demographics_cgmhx.factor) = 
  c("<= 6 months","<= 6 months","6+ months","6+ months","6+ months","6+ months",
    "<= 6 months")
levels(df$demographics_pumphx.factor) = 
  c("<= 6 months","<= 6 months","6+ months","6+ months","6+ months","6+ months",
    "<= 6 months")
# Make df for lasso
lasso <- df %>% select(-ya_paid_total,-ya_hfs_behave_total,-ya_hfs_worry_total,
                       -c_paid_total,-c_hfs_behave_total,-c_hfs_worry_total)
```

# Lasso model selection

## M1 data

For model selection, I examined baseline HbA1c, sensor wear,AM use, TIR, daily boluses, T1D duration, age, sex, insurance status, race, ethnicity, CGM history, and pump history. Because some reviewers were concerned about using baseline age as a categorical variable given the different cohorts in the model development and validation datasets, age was kept as a continuous variable in all models. 

```{r without interact,echo=FALSE,warning=FALSE}
set.seed(1000)
# Make model matrix and outcome vector
form <- as.formula("success ~ hba1c + m1_sensor_wear + m1_time_am + m1_tir + m1_boluses +
demographics_t1d_duration + Age + demographics_sex.factor + 
demographics_insurance.factor + demographics_race.factor + 
demographics_ethnicity.factor + demographics_cgmhx.factor + 
demographics_pumphx.factor")
cc <- lasso[complete.cases(lasso),]
x <- model.matrix(form,cc)[,-1]
y <- cc$success
# Cross validation for lambda
cv.fit <- cv.glmnet(x, y, alpha=1, family = "binomial")
coef(cv.fit)
```

With no constraints on the lasso, baseline HbA1c, sensor wear, TIR, daily boluses, sex, ethnicity, and pump history were all selected. 

\newpage

## Comparing AM Use to Sensor Wear

```{r echo=FALSE}
# Model formulas
model_sens_form <- 
  as.formula("success ~ scale(hba1c,scale=F) + scale(m1_sensor_wear,scale=F) +
             scale(m1_tir,scale=F) + scale(m1_boluses,scale=F) + demographics_sex.factor + 
             demographics_ethnicity.factor + demographics_pumphx.factor")
model_am_form <- 
  as.formula("success ~ scale(hba1c,scale=F) + scale(m1_time_am,scale=F) +
             scale(m1_tir,scale=F) + scale(m1_boluses,scale=F) + demographics_sex.factor + 
             demographics_ethnicity.factor + demographics_pumphx.factor")
# Fit
model_sens <- glm(model_sens_form,family = "binomial",df)
model_am <- glm(model_am_form,family = "binomial",df)
# Compare
r_sens <- ROC(form = model_sens_form,data = df,MX=F,MI=F,PV=F,plot = "ROC",
              main = "Sensor Wear Model")
r_am <- ROC(form = model_am_form,data = df,MX=F,MI=F,PV=F,plot = "ROC",
            main = "AM Use Model")
```

Models using AM use at month 1 were compared to models using sensor wear at month 1. Technically, sensor wear was slightly better in terms of AIC and AUC, but the predictive abilities of the models were comparable and both excellent (AUC = 0.928 for sensor wear and AUC = 0.909 for AM use). So, the following models used time in AM instead of sensor wear, given that the validation cohort only has data for AM time.

Some of the models would not converge when both ethnicity and pump history were included, so only ethnicity was retained in the following models and participants of unknown ethnicity were excluded.

## Comparing 1 month variables to 3 month variables

```{r echo=FALSE}
# Get three month data
three_month <- data %>% 
  filter(gyl_timepoint.factor == "3 Months") %>%
  mutate(m3_sensor_wear = sensor_wear,
         m3_time_am = time_am,
         m3_tir = sensor_70_180,
         m3_boluses = dailybolus) %>%
  select(record_id,m3_sensor_wear,m3_time_am,m3_tir,m3_boluses)
# Add to df
df <- left_join(df,three_month,by = "record_id")
df = df %>% filter(demographics_ethnicity.factor != "Unknown / Not Reported")
# Compare 1 month to 3 month
model1 <- 
  as.formula("success ~ scale(hba1c,scale=F) + scale(m1_time_am,scale=F) +
             scale(m1_tir,scale=F) + scale(m1_boluses,scale=F) + demographics_sex.factor +
             demographics_ethnicity.factor")
model3 <- 
  as.formula("success ~ scale(hba1c,scale=F) + scale(m3_time_am,scale=F) +
             scale(m3_tir,scale=F) + scale(m3_boluses,scale=F) + demographics_sex.factor +
             demographics_ethnicity.factor")
# Fit
model_m1 <- glm(model1,family = "binomial",df)
model_m3 <- glm(model3,family = "binomial",df)
# Compare
r_sens <- ROC(form = model1,data = df,MX=F,MI=F,PV=F,plot = "ROC",
              main = "1-Month Model")
r_am <- ROC(form = model3,data = df,MX=F,MI=F,PV=F,plot = "ROC",
            main = "3-Month Model")
```

The month 3 and month 1 models have essentially the same predictive ability (AUC 0.894 vs. 0.900).

\newpage

## Results

```{r stanford,echo=FALSE,message=FALSE,warning=FALSE}
# Read in
stanford <- read.csv("./Data_Cleaned/stanford.csv",
                     na.strings = c("","NA","na","N/A"),stringsAsFactors = F)
# Format variables to match BDC data
num_vars <-
  c("Baseline.A1c","Auto.Mode...1","Time.in.range.....1","Age","X..Boluses.day...Visit")
stanford[,num_vars] <- lapply(stanford[,num_vars], as.numeric)
# Insurance - combine checkboxes
stanford$insurance <- 
  ifelse(stanford$Insurance.Type..choice.Private.HMO. == "Checked" | 
           stanford$Insurance.Type..choice.Private.PPO. == "Checked" |
           stanford$Insurance.Type..choice.Other.Private. == "Checked","Private",
         ifelse(stanford$Insurance.Type..choice.CCS. == "Checked" | 
                  stanford$Insurance.Type..choice.MediCal. == "Checked" |
                  stanford$Insurance.Type..choice.MediCare. == "Checked" |
                  stanford$Insurance.Type..choice.Other.government. == "Checked",
                "Public","Other"))
# Race
stanford$race <- 
  ifelse(stanford$Race..choice.White. == "Checked","White","Non-white")
# Pump and CGM history
stanford = stanford %>% group_by(Participant.ID) %>% 
  fill(CGM.Start.Date,Pump.Start.Date,Auto.Mode.Start,.direction = "downup")

stanford[,c("CGM.Start.Date","Pump.Start.Date")] = 
  lapply(stanford[,c("CGM.Start.Date","Pump.Start.Date")],lubridate::ym)
stanford$Auto.Mode.Start = lubridate::mdy(stanford$Auto.Mode.Start)

stanford$pump.hx = difftime(stanford$Auto.Mode.Start,stanford$Pump.Start.Date,units = "days")
stanford$cgm.hx = difftime(stanford$Auto.Mode.Start,stanford$CGM.Start.Date,units = "days")
# Get variables from appropriate time points
stan_df <- stanford %>% group_by(Participant.ID) %>%
  summarise(hba1c = Baseline.A1c[Event.Name=="670G Start"],
            m3_time_am = Auto.Mode...1[Event.Name=="3 month visit"],
            m3_tir = Time.in.range.....1[Event.Name=="3 month visit"],
            m3_boluses = X..Boluses.day...Visit[Event.Name=="3 month visit"],
            Age = Age[Event.Name=="Baseline"],
            demographics_sex.factor = Gender[Event.Name=="Baseline"],
            demographics_t1d_duration = 
              as.numeric(T1D.Duration[Event.Name=="Baseline"]),
            demographics_insurance.factor = insurance[Event.Name=="Baseline"],
            demographics_race.factor = race[Event.Name=="Baseline"],
            demographics_ethnicity.factor = Ethnicity[Event.Name=="Baseline"],
            demographics_cgmhx.factor = 
              ifelse(cgm.hx[Event.Name=="Baseline"]>180,"6+ months","<= 6 months"),
            demographics_pumphx.factor = 
              ifelse(pump.hx[Event.Name=="Baseline"]>180,"6+ months","<= 6 months"),
            Age = Age[Event.Name=="Baseline"]) %>%
  filter(demographics_sex.factor != "Trans female")
# Correct factor levels
stan_df$demographics_sex.factor <- as.factor(stan_df$demographics_sex.factor)
stan_df$demographics_insurance.factor <- 
  factor(stan_df$demographics_insurance.factor,
         levels = levels(df$demographics_insurance.factor))
stan_df$demographics_race.factor <- 
  factor(stan_df$demographics_race.factor,
         levels = levels(df$demographics_race.factor))
stan_df$demographics_ethnicity.factor <- 
  factor(stan_df$demographics_ethnicity.factor)
levels(stan_df$demographics_ethnicity.factor) <- 
  levels(df$demographics_ethnicity.factor)
# Get am use at final timepoint
success <- stanford %>% group_by(Participant.ID) %>%
  filter(Event.Name == "12 month visit" | 
           Event.Name =="9 month visit") %>%
  filter(row_number() == n()) %>%
  summarise(am_final = Auto.Mode...1)
# Combine covariates and success dfs
stan_df <- left_join(stan_df,success,by = "Participant.ID")
stan_df$success <- as.factor(ifelse(stan_df$am_final >= 60,"Success","Failure"))
stan_df = stan_df %>% filter(demographics_ethnicity.factor != "Unknown / Not Reported")
# Cohort
stan_df$Cohort <- "Stanford"
colnames(stan_df)[1] <- "record_id"
df$Cohort <- "BDC"
df$record_id = as.character(df$record_id)
t1_df <- full_join(df,stan_df)
# Combine questionnaires
t1_df = t1_df %>% 
  unite(paid_total,c_paid_total,ya_paid_total,na.rm = T,remove = F) %>%
  unite(hfs_behave_total,c_hfs_behave_total,ya_hfs_behave_total,na.rm = T,remove = F) %>%
  unite(hfs_worry_total,c_hfs_worry_total,ya_hfs_worry_total,na.rm = T,remove = F)
t1_df[t1_df==""] = NA
t1_df[,c("paid_total","hfs_behave_total","hfs_worry_total")] = 
  lapply(t1_df[,c("paid_total","hfs_behave_total","hfs_worry_total")],as.numeric)
```

### Table 1a: Descriptive Characteristics by Cohort

```{r table 1,echo=FALSE,results='asis'}
# Table 1 design formula. Visually non-normal variables tested with K-W, categorical
# variables with < 5 in any cell tested with Fisher's exact
t1_form <- "Cohort~Age+kwt(am_final)+hba1c+kwt(m1_sensor_wear)+kwt(m3_sensor_wear)+kwt(m1_time_am)+kwt(m3_time_am)+m1_tir+m3_tir+kwt(demographics_t1d_duration)+success+Age+c_paid_total+ya_paid_total+c_hfs_worry_total+ya_hfs_worry_total+hfs_behave_total+demographics_sex.factor+fe(demographics_insurance.factor)+demographics_race.factor+fe(demographics_ethnicity.factor)+demographics_pumphx.factor+demographics_cgmhx.factor"

t1 <- tableby(as.formula(t1_form),data = t1_df[!is.na(t1_df$success),])
summary(t1,labelTranslations = 
          list(am_final = "AM Use at Final Visit",hba1c = "Baseline HbA1c",
               m1_sensor_wear = "M1 Sensor Wear",m3_sensor_wear = "M3 Sensor Wear",
               m1_time_am = "M1 Time in AM",m3_time_am = "M3 Time in AM",
               m1_tir = "M1 TIR",m3_tir = "M3 TIR",
               c_paid_total = "PAID Peds Total",
               ya_paid_total = "PAID YA Total",
               c_hfs_worry_total = "Peds HFS Worry Total",
               ya_hfs_worry_total = "YA HFS Worry Total",
               hfs_behave_total = "HFS Behave Total",
               success = "Success",
               demographics_t1d_duration = "T1D Duration",
               demographics_sex.factor = "Sex",
               demographics_insurance.factor = "Insurance",
               demographics_race.factor = "Race",
               demographics_ethnicity.factor = "Ethnicity",
               demographics_cgmhx.factor = "CGM History",
               demographics_pumphx.factor = "CSII Pump History"),
        pfootnote=T)
```

### Table 1b: Descriptive Characteristics by Sucess (BDC Cohort Only)

```{r table 1b,echo=FALSE,results='asis'}
# Table 1 design formula. Visually non-normal variables tested with K-W, categorical
# variables with < 5 in any cell tested with Fisher's exact
t1_form <- update(as.formula(t1_form),success ~ .)

t1 <- tableby(t1_form,data = t1_df[!is.na(t1_df$success) & t1_df$Cohort == "BDC",])
summary(t1,labelTranslations = 
          list(am_final = "AM Use at Final Visit",hba1c = "Baseline HbA1c",
               m1_sensor_wear = "M1 Sensor Wear",m3_sensor_wear = "M3 Sensor Wear",
               m1_time_am = "M1 Time in AM",m3_time_am = "M3 Time in AM",
               m1_tir = "M1 TIR",m3_tir = "M3 TIR",
               c_paid_total = "PAID Peds Total",
               ya_paid_total = "PAID YA Total",
               c_hfs_worry_total = "Peds HFS Worry Total",
               ya_hfs_worry_total = "YA HFS Worry Total",
               hfs_behave_total = "HFS Behave Total",
               success = "Success",
               demographics_t1d_duration = "T1D Duration",
               demographics_sex.factor = "Sex",
               demographics_insurance.factor = "Insurance",
               demographics_race.factor = "Race",
               demographics_ethnicity.factor = "Ethnicity",
               demographics_cgmhx.factor = "CGM History",
               demographics_pumphx.factor = "CSII Pump History"),
        pfootnote=T)
```

### Final model (1 month variables)

```{r m1 time_am,echo=FALSE}
final_model <- as.formula("success ~ scale(hba1c,scale=F) + scale(m1_time_am,scale=F) +
             scale(m1_tir,scale=F) + scale(m1_boluses,scale=F) + demographics_sex.factor +
             demographics_ethnicity.factor")
final_mod = glm(final_model,family = "binomial",df)
t = broom::tidy(final_mod,conf.int = T,exponentiate=T)
t$term[2:nrow(t)] = c("Baseline HbA1c","M1 AM %","M1 TIR","M1 Boluses per Day","Male","NOT Hispanic or Latino")
t[,2:ncol(t)] <- lapply(t[,2:ncol(t)],function(x){round(x,4)})

r = ROC(form = final_model,data = df,MX=F,MI=F,PV=F,plot = "ROC",
        main = "1-Month Model")
kable(t)
```

### Final model (3 month variables)

```{r m3 time am model,echo=FALSE}
final_model <- as.formula("success ~ scale(hba1c,scale=F) + scale(m3_time_am,scale=F) +
             scale(m3_tir,scale=F) + scale(m3_boluses,scale=F) + demographics_sex.factor +
             demographics_ethnicity.factor")
final_mod = glm(final_model,family = "binomial",df)
t = broom::tidy(final_mod,conf.int = T)
t$term[2:nrow(t)] = c("Baseline HbA1c","M3 AM %","M3 TIR","M3 Boluses per Day","Male","NOT Hispanic or Latino")
t[,2:ncol(t)] <- lapply(t[,2:ncol(t)],function(x){round(x,4)})

r = ROC(form = final_model,data = df,MX=F,MI=F,PV=F,plot = "ROC",main = "3-Month Model")
kable(t)
```

### Validation

The 3-month model above was applied to the Stanford data, and resulted in fair prediction compared to the excellent prediction in the BDC dataset.

```{r test model,echo=FALSE}
# Model formula
final_model <- as.formula("success ~ scale(hba1c,scale=F) + scale(m3_time_am,scale=F) +
             scale(m3_tir,scale=F) + scale(m3_boluses,scale=F) + demographics_sex.factor +
             demographics_ethnicity.factor")
final_mod = glm(final_model,family = "binomial",df)
# Test on Stanford's data
pred = predict(final_mod,stan_df)
# ROC
r = ROC(test = pred,stat = stan_df$success,MX=F,MI=F,PV=F,plot = "ROC",
        main = "Validation")
```

#### AUC comparison

```{r warning=FALSE}
# Get data and combine into one df
df$train_prob = predict(final_mod,df,type = "response")
stan_df$train_prob = predict(final_mod,stan_df,type = "response")
plot_df = rbind(df[,c("record_id","train_prob","success","Cohort")],
                stan_df[,c("record_id","train_prob","success","Cohort")])
# Plot
p = ggplot(plot_df, aes(d = success, m = train_prob, linetype = Cohort)) + 
  geom_roc(n.cuts = 0) + theme_bw() +
  xlab("1 - Specificity") + ylab("Sensitivity")
p
```

# References

1. 	Tibshirani R. Regression Shrinkage and Selection Via the Lasso. J R Stat Soc Ser B Methodol. 1996;58(1):267-288. doi:10.1111/j.2517-6161.1996.tb02080.x

2. 	Hastie T, Tibshirani R, Friedman JH. The Elements of Statistical Learning: Data Mining, Inference, and Prediction. 2nd ed. Springer; 2009

3. 	Friedman J, Hastie T, Tibshirani R. Regularization Paths for Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33(1). doi:10.18637/jss.v033.i01
