---
title: "JDRF Models"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Greg Forlenza/JDRF")
library(Hmisc)
library(arsenal)
library(skimr)
library(knitr)
library(glmnet)
library(tidyverse)
```

```{r import data,echo=FALSE}
source("/Users/timvigers/Documents/GitHub/BDC-Code/Greg Forlenza/JDRF/data_import.R")
# Exclude participants after 115 and those who never started AM
exclude <- data$record_id[which(data$gyl_timepoint==0 & is.na(data$automode_start))]
data <- data %>% filter(record_id <= 115,!(record_id %in% exclude))
```

```{r data cleaning,echo=FALSE,message=FALSE}
# Binary endpoint and predictors
# The primary endpoint of interest will be percent time in closed loop at 1 year 
# captured as a continuous percentage for the 2 weeks prior to the 1 year clinical 
# follow up visit, and will be converted to a binary variable with ≥60% representing 
# successful AP use and <60% representing unsuccessful AP use.
df <- left_join(data %>% group_by(record_id) %>% 
                  filter(gyl_timepoint.factor =="12 Months" | 
                           gyl_timepoint.factor =="9 Months") %>%
                  filter(row_number() == n()) %>%
                  summarise(am_final = time_am), 
                data %>% filter(grepl("baseline",redcap_event_name)),
                by = "record_id")
# Remove duplicates
df <- df[-c(which(duplicated(df$record_id))-1),]
# PAID
df$c_paid_total <- 
  apply(df[,which(colnames(df)=="c_paid1"):
                  which(colnames(df)=="c_paid20")],1,
        function(x){mean(5-x)*25})
df$ya_paid_total <- 
  apply(df[,which(colnames(df)=="ya_paid1"):
                  which(colnames(df)=="ya_paid20")],1,
        function(x){sum(x)*1.25})
# HFS behave
df$c_hfs_behave_total <- 
  apply(df[,which(colnames(df)=="c_hfs_behave1"):
                  which(colnames(df)=="c_hfs_behave10")],1,
        function(x){sum(x)})
df$ya_hfs_behave_total <- 
  apply(df[,which(colnames(df)=="ya_hfs_b_behave1"):
                  which(colnames(df)=="ya_hfs_b_behave15")],1,
        function(x){sum(x)})
# HFS worry
df$c_hfs_worry_total <- 
  apply(df[,which(colnames(df)=="c_hfs_worry11"):
                  which(colnames(df)=="c_hfs_worry25")],1,
        function(x){sum(x)})
df$ya_hfs_worry_total <- 
  apply(df[,which(colnames(df)=="ya_hfs_b_worry16"):
                  which(colnames(df)=="ya_hfs_b_worry33")],1,
        function(x){sum(x)})
# Add 1 month variables, select relevant columns
df <- left_join(df,data %>% 
                  filter(gyl_timepoint.factor == "Month 1/ Training F/U") %>%
                  mutate(m1_sensor_wear = sensor_wear,
                         m1_time_am = time_am,
                         m1_tir = sensor_70_180) %>%
                  select(record_id,m1_sensor_wear,m1_time_am,m1_tir),
                by = "record_id") %>%
  select(record_id,am_final,hba1c,c_paid_total,ya_paid_total,
         c_hfs_behave_total,ya_hfs_behave_total,
         c_hfs_worry_total,ya_hfs_worry_total,
         m1_sensor_wear,m1_time_am,m1_tir,
         demographics_t1d_duration,demographics_age,
         demographics_sex.factor,demographics_insurance.factor,
         demographics_race.factor,demographics_ethnicity.factor,
         demographics_cgmhx.factor,demographics_pumphx.factor)
# Binary outcome
df$success <- as.factor(ifelse(df$am_final >= 60,"Success","Failure"))
# Categorical variables
df$demographics_age <- cut(df$demographics_age,c(5,13,18,Inf),right = T,
                           labels = c("6 - 13","14 - 18","18+"))
levels(df$demographics_race.factor) = 
  c("Non-white","Non-white","Non-white","Non-white","White","Non-white","Non-white")
levels(df$demographics_cgmhx.factor) = 
  c("<= 6 months","<= 6 months","6+ months","6+ months","6+ months","6+ months",
    "<= 6 months")
levels(df$demographics_pumphx.factor) = 
  c("<= 6 months","<= 6 months","6+ months","6+ months","6+ months","6+ months",
    "<= 6 months")
```

# Skim

```{r echo=FALSE}
skim(df)
```

# Table 1: Descriptive Characteristics

```{r table 1,echo=FALSE,results='asis'}
t1_form <- paste(colnames(df)[2:ncol(df)],collapse = "+")
t1 <- tableby(as.formula(paste("~",t1_form)),data = df)
summary(t1)
```

# Models

Using the lasso for model selection requires complete data, and the survey variables have a lot of missing data. The sample size of YA participants is too small to use the lasso, but none of the surveys are associated with the outcome based on simple logistic models of the form:

$$
logit(success) = \beta_0+\beta_1*paid+\beta_2*worry+\beta_3*behave + \epsilon
$$

Because of all this, surveys were excluded from the lasso model selection, in order to keep the data in one cohort and to avoid missing data problems.

## Lasso

### Without interaction terms

```{r without interact,echo=FALSE,warning=FALSE}
# Remove surveys
lasso <- df %>% select(-ya_paid_total,-ya_hfs_behave_total,-ya_hfs_worry_total,
                       -c_paid_total,-c_hfs_behave_total,-c_hfs_worry_total)
# Lasso
# Make model matrix and outcome vector
form <- as.formula("success ~ hba1c + m1_sensor_wear + m1_time_am + m1_tir + 
demographics_t1d_duration + demographics_age + demographics_sex.factor + 
demographics_insurance.factor + demographics_race.factor + 
demographics_ethnicity.factor + demographics_cgmhx.factor + 
demographics_pumphx.factor")
cc <- lasso[complete.cases(lasso),]
x <- model.matrix(form,cc)[,-1]
y <- cc$success
# Fit
fit <- glmnet(x,y,family = "binomial")
plot(fit,xvar = "lambda")
# Cross validation for lambda
cv.fit <- cv.glmnet(x, y, alpha=1, family = "binomial")
coef(cv.fit)
```

### With interaction terms

Because sensor wear and time in automode at month 1 are correlated, only sensor wear was included in further model selection.

```{r with interact,echo=FALSE,warning=FALSE}
# Lasso
# Make model matrix and outcome vector
form <- as.formula("success ~ demographics_age*hba1c + 
demographics_age*m1_sensor_wear + demographics_age*m1_tir + 
demographics_age*demographics_sex.factor + 
demographics_age*demographics_ethnicity.factor + 
demographics_age*demographics_pumphx.factor")
cc <- lasso[complete.cases(lasso),]
x <- model.matrix(form,cc)[,-1]
y <- cc$success
# Fit
fit <- glmnet(x,y,family = "binomial")
plot(fit,xvar = "lambda")
# Cross validation for lambda
cv.fit <- cv.glmnet(x, y, alpha=1, family = "binomial")
coef(cv.fit)
```

## Figures

### HbA1c by age

```{r hba1c age plot,echo=FALSE}
data <- data %>% group_by(record_id) %>% fill(demographics_dob)
data$age <- as.numeric(difftime(as.Date(data$date_visit),
                                as.Date(data$demographics_dob)))/365.25
data$age_group <- cut(data$age,c(5,13,18,Inf),right = T,
                           labels = c("6 - 13","14 - 18","18+"))
ggplot(data,aes(x = age,y = hba1c,group = record_id)) + geom_line() + xlim(0,25)
```

# Notes

Used largest value of lambda such that error is within 1 standard error of the minimum.

Deviance is the loss function used for CV (10 fold).
