---
title: "Vira Kravets - Mixed model - updates for PLOS BIO"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(nlme)
library(robustlmm)
library(geoR)
library(sjstats)
library(car)
library(readxl)
library(dplyr)
library(emmeans)
library(data.table)
library(knitr)
library(stringr)

if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects/Vira Kravets/PLOS BIO resubmission"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
setwd(home_dir)
```

```{r, include=FALSE}
# read in fig 1f
fig1f <- read.csv("./Fig 1F By Cell_IslNorm.csv")
colnames(fig1f) <- c("ID",colnames(fig1f[,2:8]))
fig1f$mouse <- substring(fig1f$ID,1,5)
fig1f$islet <- sub(".*_", "", fig1f$ID)
# need to assign a cell ID within islets
fig1f$unique_islet <- NA
for (i in 1:nrow(fig1f)) {
  fig1f[i,]$unique_islet <- paste0(fig1f[i,]$mouse,fig1f[i,]$islet)
}
fig1f <- fig1f %>% group_by(unique_islet) %>% mutate(cell = sequence(n()))
# make a long dataset
fig1f_long <- reshape(fig1f, direction="long", idvar = c("mouse","islet","cell"), 
                      varying = c("Islet","X1st.resp","w.or","w.end","last.resp","hub.like..p1.","hub.like..p2."),
                      v.names = "value",timevar = "Characteristic",
                      times=c("Islet","X1st.resp","w.or","w.end","last.resp","hub.like..p1.","hub.like..p2."))
fig1f_long <- fig1f_long %>% arrange(mouse,islet,cell)
fig1f_long$l_value <- log(fig1f_long$value)
fig1f_long$Characteristic <- as.factor(fig1f_long$Characteristic)

# read in fig 1g
fig1g <- read.csv("./Fig 1G byCell_IslNorm.csv")
colnames(fig1g) <- c("ID",colnames(fig1g[,2:8]))
fig1g$mouse <- substring(fig1g$ID,1,5)
fig1g$islet <- sub(".*_", "", fig1g$ID)
# need to assign a cell ID within islets
fig1g$unique_islet <- NA
for (i in 1:nrow(fig1g)) {
  fig1g[i,]$unique_islet <- paste0(fig1g[i,]$mouse,fig1g[i,]$islet)
}
fig1g <- fig1g %>% group_by(unique_islet) %>% mutate(cell = sequence(n()))
# make a long dataset
fig1g_long <- reshape(fig1g, direction="long", idvar = c("mouse","islet","cell"), 
                      varying = c("Islet","X1st.resp","w.or","w.end","last.resp","hub.like..p1.","hub.like..p2."),
                      v.names = "value",timevar = "Characteristic",
                      times=c("Islet","X1st.resp","w.or","w.end","last.resp","hub.like..p1.","hub.like..p2."))
fig1g_long <- fig1g_long %>% arrange(mouse,islet,cell)
fig1g_long$Characteristic <- as.factor(fig1g_long$Characteristic)

# read in fig S2
files <- list.files(path = "./Fig S2 individual sheets",
                          full.names = TRUE,recursive = T)
figs2 <- NULL
for (f in 1:length(files)) {
  table <- read.csv(files[f],stringsAsFactors = FALSE,
                               header = TRUE,
                               na.strings = "")
  table$id <- as.factor(as.character(str_extract_all(files[f], "(?<=sheets/).+(?=.csv)")))
  figs2 <- rbind(figs2, table)
}
figs2$ï.. <- NULL
figs2$Distance.to.1st.resp <- NULL
figs2$um <- NULL
colnames(figs2) <- c("Response_time","Distance_1st_resp","ID")
figs2$mouse <- str_sub(figs2$ID, 1, 6)
figs2$islet <-str_sub(figs2$ID, 8, 12)
figs2$unique_islet <- figs2$ID

# read in fig 2K
fig2k <- read.csv("./1st resp Large bins (Fig 2K).csv")
fig2k$X <- NULL
fig2k$id <- fig2k$ï..
fig2k$ï..<- NULL
fig2k_sum <- fig2k %>% group_by(id) %>% summarise(X0h=mean(X0h, na.rm=T), X.6.12.h=mean(X.6.12.h, na.rm=T),
                                                  X.18.24.h=mean(X.18.24.h, na.rm=T), X.30.36.h=mean(X.30.36.h, na.rm=T),
                                                  X.42.48.h=mean(X.42.48.h, na.rm=T))
fig2k_sum$X0h <- ifelse(is.nan(fig2k_sum$X0h),NA,fig2k_sum$X0h)
fig2k_sum$X.6.12.h <- ifelse(is.nan(fig2k_sum$X.6.12.h),NA,fig2k_sum$X.6.12.h)
fig2k_sum$X.18.24.h <- ifelse(is.nan(fig2k_sum$X.18.24.h),NA,fig2k_sum$X.18.24.h)
fig2k_sum$X.30.36.h <- ifelse(is.nan(fig2k_sum$X.30.36.h),NA,fig2k_sum$X.30.36.h)
fig2k_sum$X.42.48.h <- ifelse(is.nan(fig2k_sum$X.42.48.h),NA,fig2k_sum$X.42.48.h)
fig2k_sum$mouse <- str_sub(fig2k_sum$id, 1, 2)
fig2k_sum$islet <- str_sub(fig2k_sum$id, 4, 5)
fig2k_sum$cell <- sub(".*_", "", fig2k_sum$id)
# now make a long dataset
fig2k_sum_long <- reshape(fig2k_sum, direction="long", idvar = c("mouse","islet","cell"), 
                      varying = c("X0h","X.6.12.h","X.18.24.h","X.30.36.h","X.42.48.h"),
                      v.names = "value",timevar = "time",
                      times=c("0","6-12","18-24","30-36","42-48"))
fig2k_sum_long <- fig2k_sum_long %>% arrange(mouse,islet,cell,time)
fig2k_sum_long$time <- as.factor(fig2k_sum_long$time)
fig2k_sum_long$time <- factor(fig2k_sum_long$time, levels=c("0","6-12","18-24","30-36","42-48"))

# read in fig 2L
fig2L <- read.csv("./last resp Large bins (Fig 2L).csv")
fig2L$X <- NULL
fig2L$id <- fig2L$ï..
fig2L$ï..<- NULL
fig2L_sum <- fig2L %>% group_by(id) %>% summarise(X0h=mean(X0h, na.rm=T), X.6.12.h=mean(X.6.12.h, na.rm=T),
                                                  X.18.24.h=mean(X.18.24.h, na.rm=T), X.30.36.h=mean(X.30.36.h, na.rm=T),
                                                  X.42.48.h=mean(X.42.48.h, na.rm=T))
fig2L_sum$X0h <- ifelse(is.nan(fig2L_sum$X0h),NA,fig2L_sum$X0h)
fig2L_sum$X.6.12.h <- ifelse(is.nan(fig2L_sum$X.6.12.h),NA,fig2L_sum$X.6.12.h)
fig2L_sum$X.18.24.h <- ifelse(is.nan(fig2L_sum$X.18.24.h),NA,fig2L_sum$X.18.24.h)
fig2L_sum$X.30.36.h <- ifelse(is.nan(fig2L_sum$X.30.36.h),NA,fig2L_sum$X.30.36.h)
fig2L_sum$X.42.48.h <- ifelse(is.nan(fig2L_sum$X.42.48.h),NA,fig2L_sum$X.42.48.h)
fig2L_sum$mouse <- str_sub(fig2L_sum$id, 1, 2)
fig2L_sum$islet <- str_sub(fig2L_sum$id, 4, 5)
fig2L_sum$cell <- sub(".*_", "", fig2L_sum$id)
# now make a long dataset
fig2L_sum_long <- reshape(fig2L_sum, direction="long", idvar = c("mouse","islet","cell"), 
                      varying = c("X0h","X.6.12.h","X.18.24.h","X.30.36.h","X.42.48.h"),
                      v.names = "value",timevar = "time",
                      times=c("0","6-12","18-24","30-36","42-48"))
fig2L_sum_long <- fig2L_sum_long %>% arrange(mouse,islet,cell,time)
fig2L_sum_long$time <- as.factor(fig2L_sum_long$time)
fig2L_sum_long$time <- factor(fig2L_sum_long$time, levels=c("0","6-12","18-24","30-36","42-48"))

# read in non-normalized data for Tresp, lags, links
tll_nonnormalized <- read.csv("./Hierarchy of Tresp,Lags,Links not normalized.csv")
tll_nonnormalized$mouse <- str_sub(tll_nonnormalized$X, 1, 6)
tll_nonnormalized$islet <- str_sub(tll_nonnormalized$X, 8, 9)

# read in normalized data for Tresp, lags, links
tll_normalized <- read.csv("./Hierarchy of Tresp,Lags,Links normalized.csv")


###########
# FIG 1f  #
###########

#mod_1f <- lme(value ~ Characteristic,random=~1|unique_islet,data = fig1f_long,na.action=na.omit)
mod_1f <- lme(value ~ Characteristic,random=~1|mouse/islet,data = fig1f_long,na.action=na.omit,
              correlation = corCompSymm())
resid_1f <- residuals(mod_1f)
shapiro.test(resid(mod_1f))
# QQplot
a_1f <- qqnorm(mod_1f,abline = c(0,1))
# ANOVA table
mod_1f_anova <- anova.lme(mod_1f, type="marginal")
# use emmeans 
mod_1f_means <- emmeans(mod_1f,"Characteristic")
mod_1f_pairs <-  pairs(mod_1f_means)

###########
# FIG 1g  #
###########

mod_1g <- lme(value ~ Characteristic,random=~1|mouse/islet,data = fig1g_long,na.action=na.omit,
              correlation = corCompSymm())
resid_1g <- residuals(mod_1g)
shapiro.test(resid(mod_1g))
# QQplot
a_1g <- qqnorm(mod_1g,abline = c(0,1))
# ANOVA table
mod_1g_anova <- anova.lme(mod_1g, type="marginal")
# use emmeans 
mod_1g_means <- emmeans(mod_1g,"Characteristic")
mod_1g_pairs <-  pairs(mod_1g_means)

###########
# FIG S2  #
###########

ctrl <- lmeControl(maxIter = 100000);
mod_s2 <- lme(Response_time ~ Distance_1st_resp,random=~1 + Distance_1st_resp|unique_islet,data = figs2,na.action=na.omit,
              correlation = corAR1(), control = ctrl)
resid_s2 <- residuals(mod_s2)
shapiro.test(resid(mod_s2))
# QQplot
a_s2 <- qqnorm(mod_s2,abline = c(0,1))
# ANOVA table
mod_s2_anova <- anova.lme(mod_s2, type="marginal")

###########
# FIG 2K  #
###########
mod_2k <- lme(value ~ time,random=~1|mouse/islet,data = fig2k_sum_long,na.action=na.omit,
              correlation = corAR1(), control = ctrl)
# ANOVA table
mod_2k_anova <- anova.lme(mod_2k, type="marginal")
# use emmeans 
mod_2k_means <- emmeans(mod_2k,"time")
mod_2k_pairs <-  pairs(mod_2k_means)
# contrast each mean with 50
mod2k_con <- contrast(mod_2k_means,
                 list(m0 = c(1, 0, 0, 0, 0), m6_12 = c(0, 1, 0, 0, 0), m18_24 = c(0, 0, 1, 0, 0),
                      m30_36 = c(0, 0, 0, 1, 0), m42_48 = c(0, 0, 0, 0, 1)), offset = c(-50, -50, -50, -50, -50))

###########
# FIG 2L  #
###########
mod_2L <- lme(value ~ time,random=~1|mouse/islet,data = fig2L_sum_long,na.action=na.omit,
              correlation = corAR1(), control = ctrl)
# ANOVA table
mod_2L_anova <- anova.lme(mod_2L, type="marginal")
# use emmeans 
mod_2L_means <- emmeans(mod_2L,"time")
mod_2L_pairs <-  pairs(mod_2L_means)
# contrast each mean with 50
mod2L_con <- contrast(mod_2L_means,
                 list(m0 = c(1, 0, 0, 0, 0), m6_12 = c(0, 1, 0, 0, 0), m18_24 = c(0, 0, 1, 0, 0),
                      m30_36 = c(0, 0, 0, 1, 0), m42_48 = c(0, 0, 0, 0, 1)), offset = c(-50, -50, -50, -50, -50))


```

# Figure 1f

This is the overall summary of the model.  However, most of the information you will need is below in the ANOVA table, estimated means, and pairwise comparisons.

The parameter labeled "rho" is the correlation between measures on the same mouse or islet. Because the sample size is relatively small, I needed to use a simple correlation structure with the same correlation between measures within a mouse as between measures within an islet. The model still accounts for both levels of correlation.

```{r echo=FALSE, comment=""}
summary(mod_1f)
```

## ANOVA table

The p-value for Characteristic is the overall test of the effect.

```{r echo=FALSE, comment=""}
kable(mod_1f_anova)
```

## Estimated means

The table below provides estimates of each mean.

```{r echo=FALSE, comment=""}
kable(mod_1f_means)
```

## Pairwise comparison of means with Tukey's HSD adjustment for multiple testing

The table below provides a comparison of each pairwise combination of means.

```{r echo=FALSE, comment=""}
kable(mod_1f_pairs)
```

# Figure 1g

This is the overall summary of the model.  However, most of the information you will need is below in the ANOVA table, estimated means, and pairwise comparisons.

The parameter labeled "rho" is the correlation between measures on the same mouse or islet. Because the sample size is relatively small, I needed to use a simple correlation structure with the same correlation between measures within a mouse as between measures within an islet. The model still accounts for both levels of correlation.

```{r echo=FALSE, comment=""}
summary(mod_1g)
```

## ANOVA table

The p-value for Characteristic is the overall test of the effect.

```{r echo=FALSE, comment=""}
kable(mod_1g_anova)
```

## Estimated means

The table below provides estimates of each mean.

```{r echo=FALSE, comment=""}
kable(mod_1g_means)
```

## Pairwise comparison of means with Tukey's HSD adjustment for multiple testing

The table below provides a comparison of each pairwise combination of means.

```{r echo=FALSE, comment=""}
kable(mod_1g_pairs)
```

# Figure S2

The model would not converge with a correlation structure of cell within islet within mouse, so a simpler correlation structure was used: cell within islet. This was compared to a model with a correlation structure of cell within mouse, but the first model had a better fit to the data. The parameter labeled "phi" is the estimate of the correlation.

This is the overall summary of the model. The slope of the regression of response time on distance can be found in the table labeled "fixed effects," under the column labeled "Value," in the row labeled "Distance_1st_resp."

```{r echo=FALSE, comment=""}
summary(mod_s2)
```

## ANOVA table

The p-value for Distance_1st_resp is the test of whether there is a significant non-zero slope.

```{r echo=FALSE, comment=""}
kable(mod_s2_anova)
```

## Individual slopes

These are the estimates of the islet-level intercepts and slopes.

```{r echo=FALSE, comment=""}
kable(coef(mod_s2))
```

# Figure 2K

This is the overall summary of the model. The parameter labeled "phi" is the correlation between times within the same cell.

```{r echo=FALSE, comment=""}
summary(mod_2k)
```

## ANOVA table

The p-value for time is the overall test of the effect of time.

```{r echo=FALSE, comment=""}
kable(mod_2k_anova)
```

## Estimated means

The table below provides estimates of each mean.

```{r echo=FALSE, comment=""}
kable(mod_2k_means)
```

## Pairwise comparison of means with Tukey's HSD adjustment for multiple testing

The table below provides a comparison of each pairwise combination of means.

```{r echo=FALSE, comment=""}
kable(mod_2k_pairs)
```

## Comparison of each mean to fixed value islet median=50

```{r echo=FALSE, comment=""}
kable(mod2k_con)
```

# Figure 2L

This is the overall summary of the model.  The parameter labeled "phi" is the correlation between times within the same cell.

```{r echo=FALSE, comment=""}
summary(mod_2L)
```

## ANOVA table

The p-value for time is the overall test of the effect of time.

```{r echo=FALSE, comment=""}
kable(mod_2L_anova)
```

## Estimated means

The table below provides estimates of each mean.

```{r echo=FALSE, comment=""}
kable(mod_2L_means)
```

## Pairwise comparison of means with Tukey's HSD adjustment for multiple testing

The table below provides a comparison of each pairwise combination of means.

```{r echo=FALSE, comment=""}
kable(mod_2L_pairs)
```

## Comparison of each mean to fixed value islet median=50

```{r echo=FALSE, comment=""}
kable(mod2L_con)
```