---
title: "Vira Kravets - Mixed model"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(nlme)
library(robustlmm)
library(geoR)
library(sjstats)
library(car)
library(readxl)
library(dplyr)
library(emmeans)

# read in data for fig 2c, log transform after adding a small number to account for zeroes
data <- read.csv("B:\\Projects\\Vira Kravets\\Data raw\\T-Tswap-Median,% - Fig2C.csv")
data$um <- NULL
colnames(data) <- c("islet","initial","repeated","distance")
data$l_initial <- log(data$initial+0.0001)

# this seems to be the same file as "data" but without distance
c1 <- read.csv("B:\\Projects\\Vira Kravets\\Data raw\\(C1) Orig.Fig. 2C.csv")
colnames(c1) <- c("islet","initial","repeated")
c1$l_initial <- log(c1$initial+0.0001)

# read in data for fig 2d
fig2d <- read.csv("B:\\Projects\\Vira Kravets\\Data raw\\c2.csv")
fig2d$um <- NULL
colnames(fig2d) <- c("islet","initial","repeated","distance")
fig2d$l_initial <- log(fig2d$initial+0.0001)

# read in data for 3b
fig3b <- read.csv("B:\\Projects\\Vira Kravets\\Data raw\\(C4) Figure 3 - Source Data fig 3b.csv")
fig3b$l_glucose <- log(fig3b$glucose+0.0001)
fig3b$l_glibencl <- log(fig3b$glibencl+0.0001)

# read in data for figure 2e
fig2e <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure2 for LP mixed model.xlsx",sheet = "Fig.2 E 1st rep")
fig2e$...2 <- NULL
fig2e$Islet <- NULL
colnames(fig2e) <- c("islet","initial","repeated")
fig2e$l_initial <- log(fig2e$initial+0.0001)

# read in data for figure 2f
fig2f <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure2 for LP mixed model.xlsx",sheet = "Fig.2 F Last Rep")
fig2f$...2 <- NULL
fig2f$Islet <- NULL
colnames(fig2f) <- c("islet","initial","repeated")
fig2f$l_initial <- log(fig2f$initial+0.0001)

# read in data for figure 2j
fig2j <- read.csv("B:\\Projects\\Vira Kravets\\Data raw\\fig2j.csv")
fig2j$l_measure <- log(fig2j$Measure+0.0001)

# read in data for figure 2k
fig2k <- read.csv("B:\\Projects\\Vira Kravets\\Data raw\\fig2k.csv")
colnames(fig2k) <- c("ID","Time","Measure")
fig2k$l_measure <- log(fig2k$Measure+0.0001)

# read in data for figure 2l
fig2l <- read.csv("B:\\Projects\\Vira Kravets\\Data raw\\fig2l.csv")
fig2l$l_measure <- log(fig2l$Measure+0.0001)

# read in data for 3d
fig3d <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure 3 mixed model for LP.xlsx",sheet = "Fig.3 D 1st")
fig3d$Islet <- NULL
colnames(fig3d) <- c("Islet","glucose","glibenclamide")
fig3d$l_glucose <- log(fig3d$glucose+0.0001)
fig3d$l_glibenclamide <- log(fig3d$glibenclamide+0.0001)

# read in data for 3e
fig3e <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure 3 mixed model for LP.xlsx",sheet = "Fig.3 E last")
fig3e$Islet <- NULL
colnames(fig3e) <- c("Islet","glucose","glibenclamide")
fig3e$l_glucose <- log(fig3e$glucose+0.0001)
fig3e$l_glibenclamide <- log(fig3e$glibenclamide+0.0001)

# read in data for figure 4
fig4 <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure4 for LP mixed model.xlsx",sheet = "Fig.4 E NAD(P)H")

# read in data for figure 5e
fig5e <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure5 for LP mixed model.xlsx",sheet = "Fig.5 E 1st")
fig5e$islet <- NULL
colnames(fig5e) <- c("islet","pre-ablation","post-ablation")

# read in data for figure 5f
fig5f <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure5 for LP mixed model.xlsx",sheet = "Fig.5 F last")
fig5f$islet <- NULL
colnames(fig5f) <- c("islet","pre-ablation","post-ablation")

# read in data for figure 5h
fig5h <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure5 for LP mixed model.xlsx",sheet = "Fig.5 H 1st")
fig5h$islet <- NULL
colnames(fig5h) <- c("islet","pre-ablation","post-ablation")

# read in data for figure 5l
fig5i <- read_xlsx("B:\\Projects\\Vira Kravets\\Data raw\\Figure5 for LP mixed model.xlsx",sheet = "Fig.5 I last")
fig5i$islet <- NULL
colnames(fig5i) <- c("islet","pre-ablation","post-ablation")


###########
# FIG 2c  #
###########

# run a mixed model with log-transformed initial values
mod_2c <- lme(repeated ~ l_initial,random=~1|islet,data = c1,na.action=na.omit)
resid_2c <- residuals(mod_2c)
# QQplot
a_2c <- qqnorm(mod_2c,abline = c(0,1))
x_2c <- a_2c$panel.args[[1]]$x
y_2c <- a_2c$panel.args[[1]]$y
points_2c <- cbind(x_2c,y_2c)
write.csv(points_2c,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2c.csv")


###########
# FIG 2d  #
###########

# run a mixed model with log-transformed initial values
mod_2d <- lme(repeated ~ l_initial,random=~1|islet,data = fig2d)
resid_2d  <- residuals(mod_2d)
# QQplot
a_2d  <- qqnorm(mod_2d,abline = c(0,1))
x_2d  <- a_2d $panel.args[[1]]$x
y_2d  <- a_2d $panel.args[[1]]$y
points_2d  <- cbind(x_2d ,y_2d )
write.csv(points_2d ,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2d.csv")

###########
# FIG 3b  #
###########

# run a mixed model with log-transformed initial values
mod_3b <- lme(glibencl ~ l_glucose,random=~1|islet,data = fig3b,na.action=na.omit)
resid_3b <- residuals(mod_3b)
# QQplot
a_3b <- qqnorm(mod_3b,abline = c(0,1))
x_3b <- a_3b$panel.args[[1]]$x
y_3b <- a_3b$panel.args[[1]]$y
points_3b <- cbind(x_3b,y_3b)
write.csv(points_3b,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig3b.csv")

###############
# FIG s3a <40 #
###############

# run a mixed model with log-transformed initial values
mod_2c_lt40 <- lme(repeated ~ l_initial,random=~1|islet,data = data[data$distance<40,])
resid_2c_lt40 <- residuals(mod_2c_lt40)
# QQplot
a_2c_lt40 <- qqnorm(mod_2c_lt40,abline = c(0,1))
x_2c_lt40 <- a_2c_lt40$panel.args[[1]]$x
y_2c_lt40 <- a_2c_lt40$panel.args[[1]]$y
points_2c_lt40 <- cbind(x_2c_lt40,y_2c_lt40)
write.csv(points_2c_lt40,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2c_lt40.csv")

###############
# FIG s3a<50  #
###############

# run a mixed model with log-transformed initial values
mod_2c_lt50 <- lme(repeated ~ l_initial,random=~1|islet,data = data[data$distance<50,])
resid_2c_lt50 <- residuals(mod_2c_lt50)
# QQplot
a_2c_lt50 <- qqnorm(mod_2c_lt50,abline = c(0,1))
x_2c_lt50 <- a_2c_lt50$panel.args[[1]]$x
y_2c_lt50 <- a_2c_lt50$panel.args[[1]]$y
points_2c_lt50 <- cbind(x_2c_lt50,y_2c_lt50)
write.csv(points_2c_lt50,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2c_lt50.csv")

###############
# FIG s3a <60 #
###############

# run a mixed model with log-transformed initial values
mod_2c_lt60 <- lme(repeated ~ l_initial,random=~1|islet,data = data[data$distance<60,])
resid_2c_lt60 <- residuals(mod_2c_lt60)
# QQplot
a_2c_lt60 <- qqnorm(mod_2c_lt60,abline = c(0,1))
x_2c_lt60 <- a_2c_lt60$panel.args[[1]]$x
y_2c_lt60 <- a_2c_lt60$panel.args[[1]]$y
points_2c_lt60 <- cbind(x_2c_lt60,y_2c_lt60)
write.csv(points_2c_lt60,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2c_lt60.csv")

###########
# FIG 2e  #
###########

# run a mixed model with log-transformed initial values
mod_2e <- lme(repeated ~ l_initial,random=~1|islet,data = fig2e)
resid_2e <- residuals(mod_2e)
# QQplot
a_2e <- qqnorm(mod_2e,abline = c(0,1))
x_2e <- a_2e$panel.args[[1]]$x
y_2e <- a_2e$panel.args[[1]]$y
points_2e <- cbind(x_2e,y_2e)
write.csv(points_2e,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2e.csv")

# calculate delta and do t-test
fig2e$delta <- fig2e$repeated - fig2e$initial
ttest_2e <- t.test(fig2e$delta,mu=0)

###########
# FIG 2f  #
###########

# run a mixed model with log-transformed initial values
mod_2f <- lme(repeated ~ l_initial,random=~1|islet,data = fig2f)
resid_2f <- residuals(mod_2f)
# QQplot
a_2f <- qqnorm(mod_2f,abline = c(0,1))
x_2f <- a_2f$panel.args[[1]]$x
y_2f <- a_2f$panel.args[[1]]$y
points_2f <- cbind(x_2f,y_2f)
write.csv(points_2f,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2f.csv")

# calculate delta and do t-test
fig2f$delta <- fig2f$repeated - fig2f$initial
ttest_2f <- t.test(fig2f$delta,mu=0)

###########
# FIG 2j  #
###########

# run a mixed model with log-transformed initial values
mod_2j <- lme(Measure ~ Time,random=~1|ID,data = fig2j,na.action=na.omit)
resid_2j <- residuals(mod_2j)
# QQplot
a_2j <- qqnorm(mod_2j,abline = c(0,1))
x_2j <- a_2j$panel.args[[1]]$x
y_2j <- a_2j$panel.args[[1]]$y
points_2j <- cbind(x_2j,y_2j)
write.csv(points_2j,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2j.csv")

means2j <- emmeans(mod_2j,"Time")
pairs2j <- pairs(means2j,adjust="tukey")

###########
# FIG 2k  #
###########

# run a mixed model with log-transformed initial values
mod_2k <- lme(Measure ~ Time,random=~1|ID,data = fig2k,na.action=na.omit)
resid_2k <- residuals(mod_2k)
# QQplot
a_2k <- qqnorm(mod_2k,abline = c(0,1))
x_2k <- a_2k$panel.args[[1]]$x
y_2k <- a_2k$panel.args[[1]]$y
points_2k <- cbind(x_2k,y_2k)
write.csv(points_2k,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2k.csv")

means2k <- emmeans(mod_2k,"Time")
pairs2k <- pairs(means2k,adjust="tukey")

###########
# FIG 2l  #
###########

# run a mixed model with log-transformed initial values
mod_2l <- lme(Measure ~ Time,random=~1|ID,data = fig2l,na.action=na.omit)
resid_2l <- residuals(mod_2l)
# QQplot
a_2l <- qqnorm(mod_2l,abline = c(0,1))
x_2l <- a_2l$panel.args[[1]]$x
y_2l <- a_2l$panel.args[[1]]$y
points_2l <- cbind(x_2l,y_2l)
write.csv(points_2l,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig2l.csv")

###########
# FIG 3d  #
###########

# run a mixed model with log-transformed initial values
mod_3d <- lme(glibenclamide ~ l_glucose,random=~1|Islet,data = fig3d,na.action=na.omit)
resid_3d <- residuals(mod_3d)
# QQplot
a_3d <- qqnorm(mod_3d,abline = c(0,1))
x_3d <- a_3d$panel.args[[1]]$x
y_3d <- a_3d$panel.args[[1]]$y
points_3d <- cbind(x_3d,y_3d)
write.csv(points_3d,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig3d.csv")

# calculate delta and do t-test
fig3d$delta <- fig3d$glucose - fig3d$glibenclamide
ttest_3d <- t.test(fig3d$delta,mu=0)

###########
# FIG 3e  #
###########

# run a mixed model with log-transformed initial values
mod_3e <- lme(glibenclamide ~ l_glucose,random=~1|Islet,data = fig3e,na.action=na.omit)
resid_3e <- residuals(mod_3e)
# QQplot
a_3e <- qqnorm(mod_3e,abline = c(0,1))
x_3e <- a_3e$panel.args[[1]]$x
y_3e <- a_3e$panel.args[[1]]$y
points_3e <- cbind(x_3e,y_3e)
write.csv(points_3e,"B:\\Projects\\Vira Kravets\\Output\\qqplot_fig3e.csv")

# calculate delta and do t-test
fig3e$delta <- fig3e$glucose - fig3e$glibenclamide
ttest_3e <- t.test(fig3e$delta,mu=0)

###########
# FIG 4  #
###########

# calculate delta and do t-test
fig4$delta <- fig4$`1st (2mM)` - fig4$`1st (11mM)`
ttest_4 <- t.test(fig4$delta,mu=0)

###########
# FIG 5E  #
###########

# calculate delta and do t-test
fig5e$delta <- fig5e$`pre-ablation` - fig5e$`post-ablation`
ttest_5e <- t.test(fig5e$delta,mu=0)

###########
# FIG 5F  #
###########

# calculate delta and do t-test
fig5f$delta <- fig5f$`pre-ablation` - fig5f$`post-ablation`
ttest_5f <- t.test(fig5f$delta,mu=0)

###########
# FIG 5H  #
###########

# calculate delta and do t-test
fig5h$delta <- fig5h$`pre-ablation` - fig5h$`post-ablation`
ttest_5h <- t.test(fig5h$delta,mu=0)

###########
# FIG 5I  #
###########

# calculate delta and do t-test
fig5i$delta <- fig5i$`pre-ablation` - fig5i$`post-ablation`
ttest_5i <- t.test(fig5i$delta,mu=0)


```

# Figure 2c 

```{r echo=FALSE, comment=""}
summary(mod_2c)
```

Interpretation of model above: for a 1% increase in initial, repeated changes by `r summary(mod_2c)$tTable[2,1]`%.

```{r echo=FALSE, comment=""}
a_2c
shapiro.test(resid(mod_2c))
r2(mod_2c)
```
<br>

# Figure 2d 

```{r echo=FALSE, comment=""}
summary(mod_2d)
```

Interpretation of model above: for a 1% increase in initial, repeated changes by `r summary(mod_2d)$tTable[2,1]`%.

```{r echo=FALSE, comment=""}
a_2d
shapiro.test(resid(mod_2d))
r2(mod_2d)
```
<br>

# Figure S3a

## Distance <40

```{r echo=FALSE, comment=""}
summary(mod_2c_lt40)
```

Interpretation of model above: for a 1% increase in initial, repeated changes by `r summary(mod_2c_lt40)$tTable[2,1]`%.

```{r echo=FALSE, comment=""}
a_2c_lt40
shapiro.test(resid(mod_2c_lt40))
r2(mod_2c_lt40)
```
<br>

## Distance <50

```{r echo=FALSE, comment=""}
summary(mod_2c_lt50)
```

Interpretation of model above: for a 1% increase in initial, repeated changes by `r summary(mod_2c_lt50)$tTable[2,1]`%.

```{r echo=FALSE, comment=""}
a_2c_lt50
shapiro.test(resid(mod_2c_lt50))
r2(mod_2c_lt50)
```
<br>

## Distance <60

```{r echo=FALSE, comment=""}
summary(mod_2c_lt60)
```

Interpretation of model above: for a 1% increase in initial, repeated changes by `r summary(mod_2c_lt60)$tTable[2,1]`%.

```{r echo=FALSE, comment=""}
a_2c_lt60
shapiro.test(resid(mod_2c_lt60))
r2(mod_2c_lt60)
```
<br>

# Figure 3b

```{r echo=FALSE, comment=""}
summary(mod_3b)
```

Interpretation of model above: for a 1% increase in glucose, glibencl changes by `r summary(mod_3b)$tTable[2,1]`%.

```{r echo=FALSE, comment=""}
a_3b
shapiro.test(resid(mod_3b))
r2(mod_3b)
```
<br>

# Figure 2e

Because there are only 2 measures per cell/islet, we can calculate the difference between initial and repeated and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_2e
```
<br>

# Figure 2f

Because there are only 2 measures per cell/islet, we can calculate the difference between initial and repeated and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_2f
```
<br>

# Figure 2j

```{r echo=FALSE, comment=""}
summary(mod_2j)
Anova(mod_2j,type="III")
means2j
pairs2j
```

```{r echo=FALSE, comment=""}
a_2j
shapiro.test(resid(mod_2j))
r2(mod_2j)
```
<br>

# Figure 2k

```{r echo=FALSE, comment=""}
summary(mod_2k)
Anova(mod_2k,type="III")
means2k
pairs2k
```

```{r echo=FALSE, comment=""}
a_2k
shapiro.test(resid(mod_2k))
r2(mod_2k)
```
<br>

# Figure 2l

```{r echo=FALSE, comment=""}
summary(mod_2l)
Anova(mod_2l,type="III")
```

```{r echo=FALSE, comment=""}
a_2l
shapiro.test(resid(mod_2l))
r2(mod_2l)
```
<br>

# Figure 3d

Because there are only 2 measures per cell/islet, we can calculate the difference between measures and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_3d
```
<br>

# Figure 3e

Because there are only 2 measures per cell/islet, we can calculate the difference between measures and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_3e
```
<br>

# Figure 4

Because there are only 2 measures per cell/islet, we can calculate the difference between measures and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_4
```
<br>

# Figure 5e

Because there are only 2 measures per cell/islet, we can calculate the difference between measures and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_5e
```
<br>

# Figure 5f

Because there are only 2 measures per cell/islet, we can calculate the difference between measures and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_5f
```
<br>

# Figure 5h

Because there are only 2 measures per cell/islet, we can calculate the difference between measures and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_5h
```
<br>

# Figure 5i

Because there are only 2 measures per cell/islet, we can calculate the difference between measures and use a one-sample t-test to test whether the difference is equal to 0.

```{r echo=FALSE, comment=""}
ttest_5i
```
<br>