---
title: "FDA Examples"
output: slidy_presentation
---

```{r setup, include=FALSE}
library(redcapAPI)
library(tidyverse)
library(lubridate)
library(tidyfun)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,cache = FALSE)
home_dir = 
  ifelse(.Platform$OS.type != "unix",
         "Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Jenna Sopfe/",
         "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Jenna Sopfe/")
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
# REDCap import
api <- read.table("./api_token.txt",header = T,sep = "\t")
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = as.character(api[1,1]))
data_full <- exportRecords(rcon)
# Remove participant 10
data_full <- data_full %>% filter(study_id != "10")
# Fill down Tx date
data_full <- data_full %>% group_by(study_id) %>% fill(date_of_tx)
# Clean
cgm = read.csv("./CGM Accuracy in BMT/Data_Cleaned/Prospective HSCT CGM Data Updated 6.24.19 FULL DATA SET.csv",na.strings = "")
cgm$Time = mdy_hm(cgm$Time)
cgm$Tx_date = mdy(cgm$Tx_date)
cgm$dx_cat = factor(cgm$dx_cat)
cgm$tx_type = factor(cgm$tx_type)
# Get other dates
datecols = c("time_engraft","time_plt_engraft")
dates = data_full %>% select(study_id,bmi,tx_type,all_of(datecols)) %>% 
  group_by(study_id) %>% fill(all_of(datecols),.direction = "updown") %>%
  filter(row_number() == 1)
cgm$tx_type = NULL
dates$study_id = as.integer(dates$study_id)
cgm = left_join(cgm,dates,by = c("SubjectID" = "study_id"))
cgm$time_engraft_minus_3 = cgm$time_engraft - 3*24*3600
# Time since engraftment
cgm$Time = round_date(cgm$Time,unit = "15 minutes")
cgm$time_from_engraft = as.numeric(difftime(cgm$Time,cgm$time_engraft_minus_3,
                                 units = "mins",tz = "UTC"))
# Format data for presentation
cgm = cgm %>% select(SubjectID,tx_type,time_from_engraft,Historic_Glucose) %>%
  group_by(SubjectID) %>% fill(tx_type) %>% ungroup() %>%
  filter(time_from_engraft >= 0 & time_from_engraft < 10080) %>%
  rename(id = SubjectID,transplant_type = tx_type,time = time_from_engraft,glucose = Historic_Glucose)
```

## Data structure

- Most of this information comes from Jeff Goldsmith's website (https://jeffgoldsmith.com/tidyfun.html)
- Functional data don't fit into the usual "tidy data" paradigm
  - Normally we want a single value in each cell of a matrix
  - FDA replaces a single-valued outcome $Y$ with a function
    - We want a single cell to contain functional data
    
## `tidyfun`

- The `tidyfun` package is a nice bridge between the tidyverse and FDA
- Focused on data cleaning, exploration, and visualization
  - Can do some light FDA with `tidyfun`, but not much at the moment
  
## Starting with traditional tidy data

- Example data is a study of blood glucose after engraftment in kids who underwent bone marrow transplant.
- We have a "long" dataset, where each row is a sensor glucose value from a CGM.
  - Sensor reading every 15 minutes.
- The goal of our analysis is to compare two different transplant types. 

```{r}
kable(head(cgm,10))
```

## `tidyfun` objects

- The workhorse data structure in `tidyfun` is the `tf` vector.
- `tidyfun` provides functions for converting `tidyverse` data to `tf`
- `tf_nest` is the `tidyfun` version of ` tidyr::nest()`, and converts long data to a `tf`:

```{r echo=TRUE}
fda_df = cgm %>% tf_nest(glucose,.id = id,.arg = time)
fda_df
```

## Easy spaghetti plots

```{r echo=TRUE,warning=FALSE}
fda_df %>% ggplot(aes(y = glucose,color = transplant_type)) + 
  geom_spaghetti(alpha = 0.3) +
  theme_bw()
```

## Easy spaghetti plots

```{r echo=TRUE,warning=FALSE}
fda_df %>% ggplot(aes(y = glucose,color = transplant_type)) + 
  geom_spaghetti(alpha = 0.3) +
  theme_bw() +
  facet_grid(~transplant_type)
```

## Adding a little sauce

```{r echo=TRUE,warning=FALSE}
fda_df %>% gglasagna(glucose) + theme_bw() + 
  facet_wrap(~transplant_type,scales = "free")
```

## Converting to basis representation

```{r echo=TRUE,warning=FALSE,message=FALSE}
fda_df %>% group_by(transplant_type) %>% 
  summarize(mean_gluc = mean(glucose,na.rm = T)) %>% 
  mutate(smooth_mean = tfb(mean_gluc,basis = "spline")) %>% 
  ggplot(aes(y = smooth_mean,color = transplant_type)) + 
  geom_spaghetti(size = 1.25, alpha = 1) + 
  geom_meatballs(aes(y = mean_gluc), alpha = .1) + theme_bw()
```

## A little bit of FDA

```{r cache=TRUE,echo=TRUE,warning=FALSE}
# Functional PCA with refundr
library(refundr)
cgm_fpca = rfr_fpca(Y = "glucose", data = fda_df)
# Plot model predictions
fda_df %>% 
  modelr::add_predictions(cgm_fpca, var = "glucose_fits") %>% 
  ggplot(aes(y = glucose_fits)) + 
  geom_spaghetti() + 
  geom_spaghetti(aes(y = glucose), color = "blue", alpha = 0.1) + 
  facet_grid(~transplant_type) + theme_bw()
```

## More data wrangling for `refund`

- Unfortunately, the `refund` package does not yet work directly with `tf` objects.
- `tidyfun` makes conversion back and forth easy though.
- For FPCA, it needs data in a "wide" format.

```{r echo=TRUE}
wide = fda_df %>% tf_unnest(glucose) %>% 
  pivot_wider(names_from = glucose_arg,values_from = glucose_value)
fda_wide = wide[,1:2]
fda_wide$glucose = wide %>% select(`0`:tail(names(.),1))
dim(fda_wide)
head(fda_wide$glucose,10)
```

```{r include=FALSE}
fda_wide$glucose = data.matrix(fda_wide$glucose)
```

## `refund.shiny` FPCA

```{r echo=TRUE,eval=FALSE}
library(refund)
library(refund.shiny)
fit.fpca = fpca.sc(fda_wide$glucose)
plot_shiny(fit.fpca)
```

## `refund.shiny` function-on-scalar

```{r echo=TRUE,eval=FALSE}
DTI = DTI[complete.cases(DTI),]
fit.fosr = bayes_fosr(cca ~ pasat + sex, data = DTI)
plot_shiny(fit.fosr)
```
