---
title: ""
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(Seurat) # need these for Differential expression testing
library(SeuratObject)
library(data.table)
library(broom)
library(pROC)
library(caTools)
library(lubridate)
library(caret)
library(tidymodels)
library(doParallel)
library(DataExplorer)
library(knitr)
library(ranger)
# update the dfs as change (changes in first chunk of version 3 on the Lambda)
#perc_express_cluster <- read.csv("S:/Laura/BDC/Projects/Howard Davidson/perc_express_cluster_TEST.csv")

# check donor cluster level
perc_express_donorcluster  <- read.csv("S:/Laura/BDC/Projects/Howard Davidson/perc_express_donorcluster_TEST.csv")
#rownames(df2) = colnames(perc_express_donorcluster)

# data
home_dir = "S:/Laura/BDC/Projects/Howard Davidson/R01/Data_Raw/Davidson pilot data/"
knitr::opts_knit$set(root.dir = home_dir)
#setwd("S:/Laura/BDC/Projects/Howard Davidson/R01/Data_Raw/Davidson pilot data/")
df2 = readRDS(paste0(home_dir, "OneDrive_1_6-28-2023/normalized.rds"))

```

```{r fun, include = FALSE}
gene_list = function(df, cutoff){
  # df is one cluster w/ all donors at perc_expr level
  dft = as.data.frame(t(df)[-1,])
  donors  = colnames(dft)
  donorgenes = list()
  for(i in donors){
    nam = dft %>% select(i)
    donor_vec = nam %>% filter(.[[1]] > cutoff)  %>% rownames()
    donorgenes[[i]] = donor_vec
  }
  return(donorgenes)
}

# Tims elnet
#source("D:/Repositories/shared-resources/Machine Learning/Tim - ElasticNet CV/easy_elasticnet.R")
#source("C:/Repositories/shared-resources/Machine Learning/Tim - ElasticNet CV/easy_elasticnet_cs.R")
```

```{r col clusters, include = FALSE}
# 22.3 for 505 
# Cluster 0
c_0 = perc_express_donorcluster %>% filter(cluster == 0) %>% select(-cluster)
a = gene_list(c_0, 22.3) 
cluster0_genes = Reduce(intersect, a) 


# Cluster 1
c_1 = perc_express_donorcluster %>% filter(cluster == 1) %>% select(-cluster)
a = gene_list(c_1, 28) # 28 for 502
cluster1_genes = Reduce(intersect, a)

# Cluster 2
c_2 = perc_express_donorcluster %>% filter(cluster == 2) %>% select(-cluster)
a = gene_list(c_2, 26)
cluster2_genes = Reduce(intersect, a) #26% for 500 exactly hot dog

# Cluster 3
c_3 = perc_express_donorcluster %>% filter(cluster == 3) %>% select(-cluster)
a = gene_list(c_3, 28)
cluster3_genes = Reduce(intersect, a)
# Cluster 4
c_4 = perc_express_donorcluster %>% filter(cluster == 4) %>% select(-cluster)
a = gene_list(c_4, 31.6)
cluster4_genes = Reduce(intersect, a)
# Cluster 5
c_5 = perc_express_donorcluster %>% filter(cluster == 5) %>% select(-cluster)
a = gene_list(c_5, 29.4)
cluster5_genes = Reduce(intersect, a)
# Cluster 6
c_6 = perc_express_donorcluster %>% filter(cluster == 6) %>% select(-cluster)
a = gene_list(c_6, 21)
cluster6_genes = Reduce(intersect, a)
# Cluster 7
c_7 = perc_express_donorcluster %>% filter(cluster == 7) %>% select(-cluster)
a = gene_list(c_7,26)
cluster7_genes = Reduce(intersect, a)
# Cluster 8
c_8 = perc_express_donorcluster %>% filter(cluster == 8) %>% select(-cluster)
a = gene_list(c_8, 26.3)
cluster8_genes = Reduce(intersect, a)
# Cluster 9
c_9 = perc_express_donorcluster %>% filter(cluster == 9) %>% select(-cluster)
a = gene_list(c_9,19.3)
cluster9_genes = Reduce(intersect, a)
# Cluster 10
c_10 = perc_express_donorcluster %>% filter(cluster == 10) %>% select(-cluster)
a = gene_list(c_10,6.5)
cluster10_genes = Reduce(intersect, a)
# Cluster 11
c_11 = perc_express_donorcluster %>% filter(cluster == 11) %>% select(-cluster)
a = gene_list(c_11,0)
cluster11_genes = Reduce(intersect, a) # only 400 shared genes between these
# Cluster 12
c_12 = perc_express_donorcluster %>% filter(cluster == 12) %>% select(-cluster)
a = gene_list(c_12,0)
cluster12_genes = Reduce(intersect, a) # only 275 shared genes between donors
rm(a)

# create one dataframe with all clusters genes, renaming to each cluster

df_list = list(c_0=c_0,c_1=c_1,c_2=c_2,c_3=c_3,c_4=c_4,c_5=c_5,c_6=c_6,c_7=c_7,c_8=c_8,c_9=c_9,c_10=c_10,c_11=c_11,c_12=c_12)
genes_list = list(c_0=cluster0_genes,c_1=cluster1_genes,c_2=cluster2_genes,c_3=cluster3_genes,c_4=cluster4_genes,c_5=cluster5_genes,c_6=cluster6_genes,
                  c_7=cluster7_genes,c_8=cluster8_genes,c_9=cluster9_genes,c_10=cluster10_genes,c_11=cluster11_genes,c_12=cluster12_genes)

for(df_name in names(df_list)) {
  df_list[[df_name]] = df_list[[df_name]] %>% select(donor, genes_list[[df_name]])
  df_list[[df_name]] = df_list[[df_name]] %>%
    rename_with(.fn = ~paste(., df_name, sep = "_"), .cols = c(genes_list[[df_name]]))
}

rm(c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8,c_9,c_10,c_11,c_12)

expressions_df = as.data.frame(df_list %>% reduce(left_join, by = "donor"))
```

```{r outcomes derive, include = FALSE}
# outcomes data
read_excel_allsheets <- function(filename, tibble = FALSE) {
    # I prefer straight data.frames
    # but if you like tidyverse tibbles (the default with read_excel)
    # then just pass tibble = TRUE
    sheets <- readxl::excel_sheets(filename)
    x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
    if(!tibble) x <- lapply(x, as.data.frame)
    names(x) <- sheets
    x
}


a1ccpep = read_excel_allsheets("S:/Laura/BDC/Projects/Howard Davidson/R01/Data_Raw/Davidson pilot data/IDDA1C_Cpep.xlsx")
# remove all df as it looks incomplete
a1ccpep = a1ccpep[-1]

# change to numeric
outcomes_df <- Reduce(function(x, y) merge(x, y, all = TRUE), a1ccpep)
outcomes_df = outcomes_df %>% filter(!is.na(`HWD ID`)) %>% mutate(a1c = as.numeric( case_when(A1c_Value == "NULL" ~ NA,
                                                                        TRUE ~ A1c_Value)),
                                                        bmi = as.numeric( case_when(BMI == "NULL" ~ NA,
                                                                        TRUE ~ BMI))); rm(a1ccpep)
# add in visit num and baseline values
outcomes_df = outcomes_df %>% arrange(`HWD ID`, as.Date(VisitDate)) %>% group_by(`HWD ID`) %>% mutate(visit = row_number(),
                                                                                                      ecpep = case_when(`Est C-pep` >= 0 ~ `Est C-pep` ,
                                                                                                                        `Est C-pep` < 0 ~ 0 ),
                                                                                                      hwdid=`HWD ID`)

# IDDA1c
idda1c_df = outcomes_df %>% select(`HWD ID`, IDDA1C, VisitDate, visit, bmi, Gender, `Age at visit`)
idda1c_df = idda1c_df %>% group_by(`HWD ID`) %>% mutate(first_visit = case_when(visit== 1 ~ VisitDate),
                                                            idda1c_bl = case_when(visit==1~IDDA1C)) %>% fill(first_visit,idda1c_bl) %>% ungroup() 
# group into 3month bins
idda1c_df = idda1c_df %>% mutate(month_from_bl = as.duration(first_visit %--% VisitDate)/dmonths(1))
idda1c_df = idda1c_df %>% mutate(visit_binned = floor(month_from_bl/3))

analysis_idda1c = idda1c_df %>% group_by(`HWD ID`, visit_binned) %>% mutate(IDDA1C_binned = mean(IDDA1C, na.rm=T),
                                                                 bmi_binned = mean(bmi, na.rm=T),
                                                                 age_binned = mean(`Age at visit`, na.rm=T)) %>%
  select(`HWD ID`, IDDA1C_binned, bmi_binned, age_binned, Gender, idda1c_bl, visit_binned) %>% unique()

# est_cpep
ecpep_df = outcomes_df %>% select(`HWD ID`, ecpep, VisitDate, visit, bmi, Gender, `Age at visit`)
ecpep_df = ecpep_df %>% group_by(`HWD ID`) %>% mutate(first_visit = case_when(visit== 1 ~ VisitDate),
                                                            ecpep_bl = case_when(visit==1~ecpep)) %>% fill(first_visit,ecpep_bl) %>% ungroup() 
# group into 3month bins
ecpep_df = ecpep_df %>% mutate(month_from_bl = as.duration(first_visit %--% VisitDate)/dmonths(1))
ecpep_df = ecpep_df %>% mutate(visit_binned = floor(month_from_bl/3))

analysis_ecpep = ecpep_df %>% group_by(`HWD ID`, visit_binned) %>% mutate(ecpep_binned = mean(ecpep, na.rm=T),
                                                                 bmi_binned = mean(bmi, na.rm=T),
                                                                 age_binned = mean(`Age at visit`, na.rm=T)) %>%
  select(`HWD ID`, ecpep_binned, bmi_binned, age_binned, Gender, ecpep_bl, visit_binned) %>% unique()

```


```{r test mixed models pckg}
# test that mixed models work
library(nlme)
library(lme4)

analysis_idda1c$hwdid = analysis_idda1c$`HWD ID`
analysis_ecpep$hwdid = analysis_ecpep$`HWD ID`
idda1c_mod_nogenes = lme(IDDA1C_binned ~ Gender + age_binned + bmi_binned + visit_binned + idda1c_bl,
            random = ~1|hwdid,
            data = analysis_idda1c,
            na.action = na.omit)

ecpep_mod_nogenes = lme(ecpep_binned ~ Gender + age_binned + bmi_binned + visit_binned + ecpep_bl,
            random = ~1|hwdid,
            data = analysis_ecpep,
            na.action = na.omit)

summary(idda1c_mod_nogenes)
summary(ecpep_mod_nogenes)

```


# Methods

Inital dataframe was first subsetted by donor and cluster. Percent expressions were calculated by cluster such that each cluster was pared down to ~ 500 genes. This cutoff was between 20-30% by cluster (total genes n = 5815). Data were then scaled and centered. This set was further reduced via univariable linear regression with each outcome, with genes with p > 0.1 further removed (total genes n = 433). 

Leave One Out cross validation was performed for elasticnet variable selection. Models within 1 SE of the MSE of the best model were examined, with the best model being chosen by MSE. 

# A1c Slope Analysis

Elasticnets were performed on the large gene list (n = 5815 genes). The best models by Cross Validated Mean error selected 0 features from the list. The best model with > 0 features selected included 5776 features, but slightly higher CVMe. 

Elasticnets were then performed on the smaller gene list (n = 433 genes). The best model by CVMe included 61 genes listed below. Each suffix "_c_X" represents cluster.

```{r enet a1cslope}
print(a1c_slope_en_small)
```


