---
title: "Shah hip structural analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(stringr)
library(Hmisc)
library(dplyr)
library(skimr)
library(tableone)
library(knitr)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(car)
library(hablar)
library(crayon)
library(emmeans)
library(ggpubr)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
home_dir = ifelse(.Platform$OS.type == "unix",
                  "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Viral Shah/R01 bone quality in long standing T1D/Raw data",
                  "T:\\Viral Shah\\R01 bone quality in long standing T1D\\Raw data\\")
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# options(scipen=999, digits = 4)
data <- read.csv("./QBMDLifespan_Table1_updated HSA.csv")
data$T1DStatus <- as.factor(data$T1DStatus)

vars <- c("NN_BMD","NN_CSA","NN_CSMI","NN_WIDTH","NN_ED","NN_ACT","NN_PCD","NN_CMP","NN_SECT_MOD","NN_BR","IT_BMD","IT_CSA","IT_CSMI",
          "IT_WIDTH","IT_ED","IT_ACT","IT_PCD","IT_CMP","IT_SECT_MOD","IT_BR","FS_BMD","FS_CSA","FS_CSMI","FS_WIDTH","FS_ED","FS_ACT",
          "FS_PCD","FS_CMP","FS_SECT_MOD","FS_BR","NECK_BMD", "HTOT_BMD")
NN_vars <- c("NN_CSA","NN_CSMI","NN_WIDTH","NN_ED","NN_ACT","NN_PCD","NN_CMP","NN_SECT_MOD","NN_BR")
IT_vars <- c("IT_CSA","IT_CSMI","IT_WIDTH","IT_ED","IT_ACT","IT_PCD","IT_CMP","IT_SECT_MOD","IT_BR")
FS_vars <- c("FS_CSA","FS_CSMI","FS_WIDTH","FS_ED","FS_ACT","FS_PCD","FS_CMP","FS_SECT_MOD","FS_BR")

# code menopause as Y/N -1 indicates N
data$menopause <- NA
data$menopause <- ifelse(data$MENOPAUSE_YEAR != -1,1,0)
data$menopause <- as.factor(data$menopause)

# read in ethnicity from older version of file
eth <- read.csv("./HipHSA with age at dx.csv")
eth$StudyID <- eth$ï..Subject.Id
eth <- eth %>% select(StudyID, Ethnicity)
eth <- eth %>% filter(!is.na(StudyID))
eth <- unique(eth)
data$StudyID <- str_replace(data$StudyID, "_F", "")
data$StudyID <- str_replace(data$StudyID, "_E", "")
data <- merge(data, eth, by="StudyID", all.x = T, all.y = F)
data$Ethnicity <- as.factor(data$Ethnicity)

# data frame with only females
females <- data[data$THSA_Analysis_SEX=="F",]

# data frame with only T1D
t1d <- data[data$T1DStatus=="Y",]

# diagnosed before age 20?
t1d$age_dx <- t1d$AgeYr - t1d$T1D_Duration
t1d$Diagnosis_Age_lt20 <- ifelse(t1d$age_dx<20,1,0)

#t1d$yrs_since_dx <- t1d$Age - t1d$Diagnosis_Age
# use duration instead

std <- data[,vars]
standard <- function(dataframe = std){
  var.names <- names(dataframe)
  dataframe[paste0("s.", var.names)] <- lapply(dataframe[var.names], function(x) c(scale(x)))
  dataframe
}
std <- standard()
std <- cbind(std,data[,c("StudyID","T1DStatus")])



```

# Background

The purpose of this analysis was to examine differences between T1D and non-T1D participants in hip bone quality variables while controlling for potential confounders.

# Methods

Linear models were used to compare groups while adjusting for age, sex, and BMI.  An additional set of models controlled for BMD (using NN_BMD, IT_BMD, or FS_BMD, depending on the outcome of the model), in addition to age, sex, and BMI.  Finally, interaction terms were added to the fully adjusted models: (1) interaction bdtween menopause status and T1D status, in females only; (2) interaction between age and T1D status, controlling for menopause; (3) interaction between sex and T1D status, in males and females.

# Results

## Models comparing T1D and control, adjusted for age, sex, and BMI

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
#for (i in 1:length(vars)) {
#  mod <- lm(data = data,data[,vars[i]]~T1DStatus+Age+as.factor(Sex)+BMI)
#  print(noquote(c("Adjusted model for: ", noquote(vars[i]))))
#  print(summary(mod))
#  means <- emmeans(mod,"T1DStatus")
#  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
#}
```

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, and NN BMD - NN variables

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + NN_BMD"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, and FS BMD - FS variables

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in FS_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + FS_BMD"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, and IT BMD - IT variables

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in IT_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + IT_BMD"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, BMI, menopause and NN BMD - NN variables, with interaction between menopause and T1D status, females only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+BMI + NN_BMD + menopause + T1DStatus*menopause" ))
  mod <- lm(form,data = females)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,pairwise ~ T1DStatus|menopause)
  print(means)
  p <- emmip(mod, T1DStatus ~ menopause)
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, BMI, menopause and FS BMD - FS variables, with interaction between menopause and T1D status, females only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in FS_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+BMI + FS_BMD + menopause + T1DStatus*menopause" ))
  mod <- lm(form,data = females)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,pairwise ~ T1DStatus|menopause)
  print(means)
  p <- emmip(mod, T1DStatus ~ menopause)
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, BMI, menopause and IT BMD - IT variables, with interaction between menopause and T1D status, females only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in IT_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+BMI + IT_BMD + menopause + T1DStatus*menopause" ))
  mod <- lm(form,data = females)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,pairwise ~ T1DStatus|menopause)
  print(means)
  p <- emmip(mod, T1DStatus ~ menopause)
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, BMI, menopause and NN BMD - NN variables, with interaction between age and T1D status, females only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+BMI + NN_BMD + menopause + AgeYr*menopause" ))
  mod <- lm(form,data = females)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  #means <- emmeans(mod,pairwise ~ T1DStatus|menopause)
  #print(means)
  p <- emmip(mod, menopause ~ AgeYr, cov.reduce=range)
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, BMI, menopause and FS BMD - FS variables, with interaction between age and T1D status, females only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in FS_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+BMI + FS_BMD + menopause + AgeYr*menopause" ))
  mod <- lm(form,data = females)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  #means <- emmeans(mod,pairwise ~ T1DStatus|menopause)
  #print(means)
  p <- emmip(mod, menopause ~ AgeYr, cov.reduce=range)
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, BMI, menopause and IT BMD - IT variables, with interaction between age and T1D status, females only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in IT_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+BMI + IT_BMD + menopause + AgeYr*menopause" ))
  mod <- lm(form,data = females)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  #means <- emmeans(mod,pairwise ~ T1DStatus|menopause)
  #print(means)
  p <- emmip(mod, menopause ~ AgeYr, cov.reduce=range)
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, and NN BMD - NN variables, with interaction between sex and T1D status

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + NN_BMD + as.factor(THSA_Analysis_SEX)*T1DStatus" ))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,pairwise ~ as.factor(THSA_Analysis_SEX)|T1DStatus)
  print(means)
  p <- emmip(mod, T1DStatus ~ as.factor(THSA_Analysis_SEX))
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, and FS BMD - FS variables, with interaction between sex and T1D status

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in FS_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + FS_BMD + as.factor(THSA_Analysis_SEX)*T1DStatus" ))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,pairwise ~ as.factor(THSA_Analysis_SEX)|T1DStatus)
  print(means)
  p <- emmip(mod, T1DStatus ~ as.factor(THSA_Analysis_SEX))
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, and IT BMD - IT variables, with interaction between sex and T1D status

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in IT_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + IT_BMD + as.factor(THSA_Analysis_SEX)*T1DStatus" ))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,pairwise ~ as.factor(THSA_Analysis_SEX)|T1DStatus)
  print(means)
  p <- emmip(mod, T1DStatus ~ as.factor(THSA_Analysis_SEX))
  print(p)
   #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing those diagnosed age <20 vs. >=20, adjusted for sex and BMI, T1D only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "Diagnosis_Age_lt20," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~Diagnosis_Age_lt20+as.factor(THSA_Analysis_SEX)+BMI "))
  mod <- lm(form,data = t1d)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"Diagnosis_Age_lt20")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing those diagnosed age <20 vs. >=20, adjusted for sex, BMI, and NN BMD - NN variables, T1D only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "Diagnosis_Age_lt20," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~Diagnosis_Age_lt20+as.factor(THSA_Analysis_SEX)+BMI + NN_BMD"))
  mod <- lm(form,data = t1d)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"Diagnosis_Age_lt20")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing those diagnosed age <20 vs. >=20, adjusted for sex, BMI, and FS BMD - FS variables, T1D only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "Diagnosis_Age_lt20," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in FS_vars) {
  form = as.formula(paste0(v,"~Diagnosis_Age_lt20+as.factor(THSA_Analysis_SEX)+BMI + FS_BMD"))
  mod <- lm(form,data = t1d)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"Diagnosis_Age_lt20")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing those diagnosed age <20 vs. >=20, adjusted for sex, BMI, and IT BMD - IT variables, T1D only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "Diagnosis_Age_lt20," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in IT_vars) {
  form = as.formula(paste0(v,"~Diagnosis_Age_lt20+as.factor(THSA_Analysis_SEX)+BMI + IT_BMD"))
  mod <- lm(form,data = t1d)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"Diagnosis_Age_lt20")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models testing years since diagnosis, adjusted for sex, BMI, and NN BMD - NN variables, T1D only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "Diagnosis_Age_lt20," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~T1D_Duration+as.factor(THSA_Analysis_SEX)+BMI + NN_BMD"))
  mod <- lm(form,data = t1d)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  #means <- emmeans(mod,"Diagnosis_Age_lt20")
  #print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models testing years since diagnosis, adjusted for sex, BMI, and FS BMD - FS variables, T1D only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "Diagnosis_Age_lt20," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in FS_vars) {
  form = as.formula(paste0(v,"~T1D_Duration+as.factor(THSA_Analysis_SEX)+BMI + FS_BMD"))
  mod <- lm(form,data = t1d)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  #means <- emmeans(mod,"Diagnosis_Age_lt20")
  #print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models testing years since diagnosis, adjusted for sex, BMI, and IT BMD - IT variables, T1D only

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "Diagnosis_Age_lt20," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in IT_vars) {
  form = as.formula(paste0(v,"~T1D_Duration+as.factor(THSA_Analysis_SEX)+BMI + IT_BMD"))
  mod <- lm(form,data = t1d)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
 # means <- emmeans(mod,"Diagnosis_Age_lt20")
#  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```


## Models comparing T1D and control, adjusted for age, sex, BMI, race, menopause, and NN BMD - NN variables

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in NN_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + NN_BMD + Ethnicity + menopause"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, race, menopause, and FS BMD - FS variables

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in FS_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + FS_BMD + Ethnicity + menopause"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, race, menopause, and IT BMD - IT variables

The output from each model can be interpreted as follows: refer to the table labeled "Coefficients."  In the row labeled "T1DStatusY," the p-value is under the column labeled "Pr(>|t|)."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in IT_vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI + IT_BMD + Ethnicity + menopause"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

## Models comparing T1D and control, adjusted for age, sex, BMI, race, and menopause - all variables

```{r echo=FALSE,comment=''}
# group comparisons adjusted for age, sex, BMI
for (v in vars) {
  form = as.formula(paste0(v,"~T1DStatus+AgeYr+as.factor(THSA_Analysis_SEX)+BMI+ Ethnicity + menopause"))
  mod <- lm(form,data = data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"T1DStatus")
  print(means)
  #a <- Anova(mod,type="III")
  #print(a)
}
```

# Figures

```{r echo=FALSE,comment=''}
std %>% ggplot(aes(x=T1DStatus,y=s.NN_CSA,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.IT_CSA,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.FS_CSA,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))

std %>% ggplot(aes(x=T1DStatus,y=s.NN_CSMI,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.IT_CSMI,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.FS_CSMI,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))

std %>% ggplot(aes(x=T1DStatus,y=s.NN_ED,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.IT_ED,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.FS_ED,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))

std %>% ggplot(aes(x=T1DStatus,y=s.NN_ACT,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.IT_ACT,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.FS_ACT,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))

std %>% ggplot(aes(x=T1DStatus,y=s.NN_PCD,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.IT_PCD,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.FS_PCD,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))

std %>% ggplot(aes(x=T1DStatus,y=s.NN_BR,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.IT_BR,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
std %>% ggplot(aes(x=T1DStatus,y=s.FS_BR,fill=T1DStatus)) + geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent",draw_quantiles = .5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(-3,5) + stat_compare_means(method = "wilcox.test",aes(label = paste0("p=", ..p.format..)))
```