---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---



```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readxl)
library(table1)
library(nlme)
library(emmeans)
library(readr)

# data upload
# first sheet demog
lyumjev_demog <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/Lyumjuv/Clean excel_lyumjev.xlsx", 
    sheet = "Demographics")

# seondsheet the pre post data
lyumjev_data <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/Lyumjuv/Clean excel_lyumjev.xlsx", 
    sheet = "Data")

# new data "S:\Laura\BDC\Projects\Erin Cobry\op5_ur_ins\mm_r2_OP5_URinsulin.xlsx"

# add group into demogs
groups = lyumjev_data %>% select(1,`Insulin Type` )
names(groups) = c("MRN", "Group")
label(lyumjev_demog$Age_09212023) = "Age"
label(lyumjev_demog$DiabetesDuration_09212023) = "Diabetes Duration"
lyumjev_demog = left_join(groups, lyumjev_demog)

# auc 2 4 hr
hr2_cgm_vars <- read_csv("S:/Laura/BDC/Projects/Erin Cobry/op5_ur_ins/2hr_cgm_vars.csv")
hr4_cgm_vars <- read_csv("S:/Laura/BDC/Projects/Erin Cobry/op5_ur_ins/4hr_cgm_vars.csv")


groups_refs <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/Lyumjuv/mm_r2_OP5_URinsulin.xlsx",sheet = "Complete List")
groups_refs$subject_id = as.character(groups_refs$...1)
```

```{r dc, include = F}
# hr 2 pt id
hr2_cgm_vars = hr2_cgm_vars %>% mutate(patient_id = (sapply(strsplit(subject_id,"_"), `[`, 1)),
                                       visit = sapply(strsplit(subject_id,"_"), `[`, 2) )

hr2_cgm_vars = hr2_cgm_vars %>% mutate(subject_id = substr(patient_id, 3,nchar(patient_id)))

analysis_hr2 = left_join(hr2_cgm_vars,groups_refs, by = "subject_id") %>% filter(!is.na(subject_id))
analysis_hr2$visit = factor(analysis_hr2$visit, levels = c("Pre", "Post"))
analysis_hr2$insulin_type = tolower(analysis_hr2$`insulin type`)

analysis_hr2 = analysis_hr2 %>% filter(!is.na(`insulin type`))%>% filter(!is.na(visit))

# hr4 pt id
hr4_cgm_vars = hr4_cgm_vars %>% mutate(patient_id = (sapply(strsplit(subject_id,"_"), `[`, 1)),
                                       visit = sapply(strsplit(subject_id,"_"), `[`, 2) )

hr4_cgm_vars = hr4_cgm_vars %>% mutate(subject_id = substr(patient_id, 3,nchar(patient_id)))

analysis_hr4 = left_join(hr4_cgm_vars,groups_refs, by = "subject_id") %>% filter(!is.na(subject_id))
analysis_hr4$visit = factor(analysis_hr4$visit, levels = c("Pre", "Post"))

analysis_hr4$insulin_type = tolower(analysis_hr4$`insulin type`)
analysis_hr4 = analysis_hr4 %>% filter(!is.na(`insulin type`)) %>% filter(!is.na(visit))
```

```{r data prep, include = FALSE}
demog_df = groups_refs %>% select(Age, `Diabetes Duration`, `insulin type`,Gender) %>% mutate(insulin_type = tolower(`insulin type`)) %>% select(-`insulin type`)
```

```{r mods, include = FALSE}
fit_mod = function(outcome, df){

    # Fit model
  f = as.formula(paste0(outcome,"~","visit"))
  mod = lme(f,
            random = ~1|subject_id,
            data = df,
            na.action = na.omit)
  
  cat("\n")
  print(outcome)
  cat("\n")
  print(summary(mod))

}

fit_mod2 = function(outcome, df){

    # Fit model
  f = as.formula(paste0(outcome,"~","visit + insulin_type"))
  mod = lme(f,
            random = ~1|subject_id,
            data = df,
            na.action = na.omit)
  
  cat("\n")
  print(outcome)
  cat("\n")
  print(summary(mod))

}
```


Analysis plan: 
                Background: Lyumjev and Fiasp are new ultra-rapid acting insulins now on the market, currently labelled for use in Medtronic and Tandem. These insulins are designed to work-faster in the hopes of reducing post-prandial excursions. OP5 is an AID that is commonly used, but there is no published real-world data on the use of these insulins with this specific pump model. We are comparing glycemic and behavioral measures before and after switching to either lyumjev or aspart. We anticipate we will likely see no difference in glycemic measures but potentially in behavioral measure (ie overrides, and bolus/day). Would like 3 subgroup analysis: pre vs post as a whole, pre vs post lyumjev only, pre vs post fiasp only. We are not doing head-to-head of lyumjev to fiasp. We are aware n will be small for the single-type insulin analysis. 
                
Demographics we would like to include: 
Age
Sex
Diabetes duration
% lyumjev
% Fiasp

outcomes

2hr auc
4hr auc

# Methods

Demographics by group and overall were produced in a table.

Linear mixed models with random intercept for MRN and time period (Pre/Post) as a covariate were fit for the entire cohort, lyumjev only, and fiasp only. 


# Analysis

## Updated Demographics Table

```{r t1, echo=FALSE}
t1_full = table1(~.|insulin_type, data = demog_df)
t1_full
```

## Original Demogs table

```{r}
t1 = table1(~ Age_09212023 + Gender + DiabetesDuration_09212023| Group, data = lyumjev_demog)
t1
```

## Auc tables


```{r}
hr2 = table1(~ total_auc |insulin_type*visit, data = analysis_hr2)
hr2


hr4 = table1(~ total_auc |insulin_type*visit, data = analysis_hr4)
hr4
```

## Models

### 2hr AUC

There were no significant Pre/Post differences in 2hr auc for the whole cohort (p = 0.51), Fiasp (p = 0.55), or Lyumjev groups (p = 0.79)

```{r}
fit_mod("total_auc", analysis_hr2)
fit_mod("total_auc", analysis_hr2%>% filter(insulin_type == "fiasp"))
fit_mod("total_auc", analysis_hr2%>% filter(insulin_type == "lyumjev"))
```

### 4hr AUC

There were no significant Pre/Post differences in 4hr auc for the whole cohort (p = 0.57), Fiasp (p = 0.48), or Lyumjev groups (p = 0.91)

```{r}
fit_mod("total_auc", analysis_hr4)
fit_mod("total_auc", analysis_hr4%>% filter(insulin_type == "fiasp"))
fit_mod("total_auc", analysis_hr4%>% filter(insulin_type == "lyumjev"))
```
