---
title: ""
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---



```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(readxl)
library(table1)
library(nlme)
library(emmeans)
library(readr)

# data upload
# first sheet demog
lyumjev_demog <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/Lyumjuv/Clean excel_lyumjev.xlsx", 
    sheet = "Demographics")

# seondsheet the pre post data
lyumjev_data <- read_excel("S:/Laura/BDC/Projects/Erin Cobry/Lyumjuv/Clean excel_lyumjev.xlsx", 
    sheet = "Data")

# new data "S:\Laura\BDC\Projects\Erin Cobry\op5_ur_ins\mm_r2_OP5_URinsulin.xlsx"

# add group into demogs
groups = lyumjev_data %>% select(1,`Insulin Type` )
names(groups) = c("MRN", "Group")
label(lyumjev_demog$Age_09212023) = "Age"
label(lyumjev_demog$DiabetesDuration_09212023) = "Diabetes Duration"
lyumjev_demog = left_join(groups, lyumjev_demog)

# auc 2 4 hr
hr2_cgm_vars <- read_csv("S:/Laura/BDC/Projects/Erin Cobry/op5_ur_ins/2hr_cgm_vars.csv")
hr4_cgm_vars <- read_csv("S:/Laura/BDC/Projects/Erin Cobry/op5_ur_ins/4hr_cgm_vars.csv")
```

```{r data prep, include = FALSE}
colnames(lyumjev_data)

df_tir = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_TIR, Post_TIR) %>% rename(MRN = `...1`)
df_tir = df_tir %>% pivot_longer(cols = c("Pre_TIR", "Post_TIR"), 
                      names_to = "visit",
                      values_to = "TIR")


df_tbr = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_TBR, Post_TBR) %>% rename(MRN = `...1`)
df_tbr = df_tbr %>% pivot_longer(cols = c("Pre_TBR", "Post_TBR"), 
                      names_to = "visit",
                      values_to = "TBR")

df_tar = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_TAR, Post_TAR) %>% rename(MRN = `...1`)
df_tar = df_tar %>% pivot_longer(cols = c("Pre_TAR", "Post_TAR"), 
                      names_to = "visit",
                      values_to = "TAR")

df_GMI = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_GMI, Post_GMI) %>% rename(MRN = `...1`)
df_GMI = df_GMI %>% pivot_longer(cols = c("Pre_GMI", "Post_GMI"), 
                      names_to = "visit",
                      values_to = "GMI")

df_A1c = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_A1c, Post_A1c) %>% rename(MRN = `...1`)
df_A1c = df_A1c %>% pivot_longer(cols = c("Pre_A1c", "Post_A1c"), 
                      names_to = "visit",
                      values_to = "A1c")

df_SD = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_SD, Post_SD) %>% rename(MRN = `...1`)
df_SD = df_SD %>% pivot_longer(cols = c("Pre_SD", "Post_SD"), 
                      names_to = "visit",
                      values_to = "SD")

df_CoV = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_CoV, Post_CoV) %>% rename(MRN = `...1`)
df_CoV$Post_CoV = as.numeric(df_CoV$Post_CoV)
df_CoV = df_CoV %>% pivot_longer(cols = c("Pre_CoV", "Post_CoV"), 
                      names_to = "visit",
                      values_to = "CoV")

df_override = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_override, Post_overide) %>% rename(MRN = `...1`)
df_override$Post_overide = as.numeric(df_override$Post_overide)

df_override = df_override %>% pivot_longer(cols = c("Pre_override", "Post_overide"), 
                      names_to = "visit",
                      values_to = "override")

df_BpD = lyumjev_data %>% select(`...1`, `Insulin Type`,Pre_BpD, Post_BpD) %>% rename(MRN = `...1`)
df_BpD = df_BpD %>% pivot_longer(cols = c("Pre_BpD", "Post_BpD"), 
                      names_to = "visit",
                      values_to = "BpD")

```

```{r functions, include = FALSE}
fit_mod = function(outcome, df){

    # Fit model
  f = as.formula(paste0(outcome,"~","visit"))
  mod = lme(f,
            random = ~1|MRN,
            data = df,
            na.action = na.omit)
  
  cat("\n")
  print(outcome)
  cat("\n")
  print(summary(mod))

}
```


Analysis plan: 
                Background: Lyumjev and Fiasp are new ultra-rapid acting insulins now on the market, currently labelled for use in Medtronic and Tandem. These insulins are designed to work-faster in the hopes of reducing post-prandial excursions. OP5 is an AID that is commonly used, but there is no published real-world data on the use of these insulins with this specific pump model. We are comparing glycemic and behavioral measures before and after switching to either lyumjev or aspart. We anticipate we will likely see no difference in glycemic measures but potentially in behavioral measure (ie overrides, and bolus/day). Would like 3 subgroup analysis: pre vs post as a whole, pre vs post lyumjev only, pre vs post fiasp only. We are not doing head-to-head of lyumjev to fiasp. We are aware n will be small for the single-type insulin analysis. 
                
Demographics we would like to include: 
Age
Sex
Diabetes duration
% lyumjev
% Fiasp

outcomes

"Pre_TIR"  
"Pre_TBR"    
"Pre_TAR"             
"Pre_GMI"
"Pre_A1c" 
"Pre_SD"    
"Pre_CoV"  
"Pre_override"
"Pre_BpD"    

# Methods

Demographics by group and overall were produced in a table.

Linear mixed models with random intercept for MRN and time period (Pre/Post) as a covariate were fit for the entire cohort, lyumjev only, and fiasp only. 


# Analysis

## Demographics Table

```{r t1, echo=FALSE}
t1 = table1(~ Age_09212023 + Gender + DiabetesDuration_09212023| Group, data = lyumjev_demog)
t1
```

## Models

### TIR

There were no significant Pre/Post differences in TIR for the whole cohort (p = 0.46), Fiasp (p = 0.39), or Lyumjev groups (p = 0.65)

```{r}
fit_mod("TIR", df_tir)
fit_mod("TIR", df_tir %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("TIR", df_tir %>% filter(`Insulin Type` == "Lyumjev"))
```

### TBR

There were no significant Pre/Post differences in TBR for the whole cohort (p = 0.68), Fiasp (p = 0.99), or Lyumjev groups (p = 0.68)

```{r}
fit_mod("TBR", df_tbr)
fit_mod("TBR", df_tbr %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("TBR", df_tbr %>% filter(`Insulin Type` == "Lyumjev"))
```

### TAR

There were no significant Pre/Post differences in TAR for the whole cohort (p = 0.26), Fiasp (p = 0.39), or Lyumjev groups (p = 0.47)

```{r}
fit_mod("TAR", df_tar)
fit_mod("TAR", df_tar %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("TAR", df_tar %>% filter(`Insulin Type` == "Lyumjev"))
```

### GMI

There were no significant Pre/Post differences in GMI for the  Fiasp subgroups (p = 0.09) 

A significant difference in Pre/Post GMI was found in the Whole and Lyumjev groups (p = 0.01, 0.04)

For the whole groups, Pre GMI was on average 0.15 (SE 0.05) less than post 
For the lyumjev group, Pre GMI was on average 0.12(SE 0.06) less than post 

```{r}
fit_mod("GMI", df_GMI)
fit_mod("GMI", df_GMI %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("GMI", df_GMI %>% filter(`Insulin Type` == "Lyumjev"))
```


### A1c

There were no significant Pre/Post differences in a1c for the whole cohort (p = 0.34), Fiasp (p = 0.49), or Lyumjev groups (p = 0.71)


```{r}
fit_mod("A1c", df_A1c)
fit_mod("A1c", df_A1c %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("A1c", df_A1c %>% filter(`Insulin Type` == "Lyumjev"))
```

### SD

There were no significant Pre/Post differences in sd for the whole cohort (p = 0.25), Fiasp (p = 0.30), or Lyumjev groups (p = 0.06)


```{r}
fit_mod("SD", df_SD)
fit_mod("SD", df_SD %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("SD", df_SD %>% filter(`Insulin Type` == "Lyumjev"))
```

### CoV

There were no significant Pre/Post differences in CoV for the whole cohort (p = 0.06) or , Fiasp (p = 0.88) groups

There was a significant difference in pre/post cov for the lyumjev cohort (0.03). On average, PreCoV was 2.9 (se = 1.3) higher than Post.


```{r}
fit_mod("CoV", df_CoV)
fit_mod("CoV", df_CoV %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("CoV", df_CoV %>% filter(`Insulin Type` == "Lyumjev"))
```

### override

There were no significant Pre/Post differences in override for the  Fiasp (p = 0.25) group

There was a significant difference in pre/post override for the whole cohort (0.02). On average, Preoverride was 5.4 (se = 2.1) higher than Post.

There was a significant difference in pre/post override for the lyumjev cohort (0.02). On average, Preoverride was 5.1 (se = 2.1) higher than Post.


```{r}
fit_mod("override", df_override)
fit_mod("override", df_override %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("override", df_override %>% filter(`Insulin Type` == "Lyumjev"))
```

### BpD

There were no significant Pre/Post differences in BpD for the Whole (p = 0.82) Fiasp (p = 0.50) and lyumjev (p = 0.71) groups


```{r}
fit_mod("BpD", df_BpD)
fit_mod("BpD", df_BpD %>% filter(`Insulin Type` == "Fiasp"))
fit_mod("BpD", df_BpD %>% filter(`Insulin Type` == "Lyumjev"))
```