---
title: ""
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(glinternet)
library(boot)
library(knitr)
set.seed(1000)
```

# TDD Model

## Methods

All analyses were performed using R version 4.1.0. Model selection was based on 10-fold cross validation with hierarchical group-lasso regularization as implemented in the glinternet R package, version 1.0.11. Model selection was performed on variables: Age, Sex, Race/Ethnicity, Initial_A1c, DKA, and Wt_onset and all interactions with Average_TDD as the outcome. Bicarb was categorized as a  into severe (< 5), moderate (5 - 10), and mild (> 10). Race and Ethnicity were combined per Kimber's 6/22 email.

Next, a linear model was re-fit. A linear mixed model with random effect for provider produced convergence warnings and was not much better than a simple linear model. 

## 10-Fold Cross Validation

```{r}
# Format data for glinternet
df = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/Kimber Simmons/New Onset Dosing/Data_Cleaned/cleaned_data.csv",stringsAsFactors = T,na.strings = "")
lasso_df = df %>% select(Age,Sex,Race,Initial_A1c,DKA,Wt_onset,Average_TDD)
lasso_df = lasso_df[complete.cases(lasso_df),]
lasso_df$Race = droplevels(lasso_df$Race)
Y = lasso_df$Average_TDD
X = lasso_df %>% select(Age,Sex,Race,Initial_A1c,DKA,Wt_onset)
numLevels = unlist(lapply(X,function(c){length(levels(c))}))
numLevels[numLevels==0] = 1
X = as.data.frame(lapply(X,function(x){as.numeric(x)-1}))
# CV
cv_fit = glinternet.cv(X, Y, numLevels)
# Plot
plot(cv_fit)
# Best lambda
best_lambda <- which(cv_fit$lambda == cv_fit$lambdaHat)
```

The lambda index with the lowest error was number `r which(cv_fit$lambdaHat == cv_fit$lambda)`, so this value was used to find the correct model. The value which produced the simplest model within 1 SD of the minimum error resulted in only age being selected, which is not particularly useful. 

```{r}
cat = names(numLevels)[numLevels > 1]
cont = names(numLevels)[numLevels == 1]
# Get coefficients
coefs <- coef(cv_fit$glinternetFit)[[best_lambda]]
```

## Continuous Main Effects

```{r}
kable(cont[coefs$mainEffects$cont],col.names = "")
```

## Categorical Main Effects

```{r}
kable(cat[coefs$mainEffects$cat],col.names = "")
```

## Continuous-Continuous Interaction Effects

```{r}
kable(cont[coefs$interactions$contcont],col.names = "")
```

## Categorical-Categorical Interaction Effects

```{r}
kable(cat[coefs$interactions$catcat],col.names = "")
```

## Continuous-categorical Interaction Effects.

None

## Final Model Coefficients

Race was not included in the final model because there are too many categories for an interpretable model.

```{r}
mod = lm(Average_TDD~Age+Initial_A1c+Wt_onset+Sex+DKA,df)
kable(summary(mod)$coefficients)
```

# Correction Equations

## Methods

Week 1 bolus values were assumed to be independent of TDD. For each participant, we calculated $$y_i = \text{Correction}_i*\text{TDD}_i$$ Next, we regressed $y_i$ (the numerator in the correction equation) on age, to see if there was any association. Finally, we used a 10,000 sample bootstrap to calculate confidence intervals for $y_i$ using the percentile method.

## BG Correction

### Association with age

```{r warning=FALSE,message=FALSE}
# Create "X" variable
df$week1_correction1 = as.numeric(as.character(df$week1_correction1))
df$short_acting = df$week1_correction1 * df$Average_TDD
# Categorical age
df$age_cat = cut(df$Age,c(12,36,156,Inf),labels = c("Toddler","Elementary","Adolescent"))
# Association with age
mod = lm(short_acting~Age,df)
ggplot(df,aes(x = Age,y = short_acting)) + 
  geom_point() + geom_smooth(method = "lm",color="blue",se=F) +
  theme_bw() + ylab("Numerator") + xlab("Age (months)")
kable(summary(mod)$coefficients)
```

There does not appear to be a significant association with age. This was also true when age was converted to a categorical variable. 

### Numerator

```{r}
y = df$short_acting[!is.na(df$short_acting)]
m = function(data,i){
  mean(data[i],na.rm = T)
}
b = boot(data = y,statistic = m,R = 10000,parallel = "multicore",ncpus = 8)
ci = boot.ci(b,type = "perc")
```

The average numerator in this group was `r round((mean(df$short_acting,na.rm=T)))`, with a 95% confidence interval of (`r round(ci$percent[4])` - `r round(ci$percent[5])`).

## I:C Correction

### Association with age

```{r warning=FALSE,message=FALSE}
# Create "X" variable
df$short_acting = rowMeans(df[,c("week1_carbbreakfast","week1_carblunch","week1_carbdinner")]) * df$Average_TDD
# Association with age
mod = lm(short_acting~Age,df)
ggplot(df,aes(x = Age,y = short_acting)) + 
  geom_point() + geom_smooth(method = "lm",color="blue",se=F) +
  theme_bw() + ylab("Numerator") + xlab("Age (months)")
kable(summary(mod)$coefficients)
```

There does not appear to be a significant association with age. This was also true when age was converted to a categorical variable. 

### Numerator

```{r}
y = df$short_acting
b = boot(data = y,statistic = m,R = 10000,parallel = "multicore",ncpus = 8)
ci = boot.ci(b,type = "perc")
```

The average numerator in this group was `r round((mean(df$short_acting,na.rm=T)))`, with a 95% confidence interval of (`r round(ci$percent[4])` - `r round(ci$percent[5])`).
